### 2021-09-10 documents/config/own_abbreviations.tex
initial commit

\newGlossingAbbrev{inter}{interrogative}
\DeclareAcronym{tam}{short=TAM,long=Tense-Aspect-Mood,class=other}

### 2021-09-10 documents/sections/extensions.tex
initial commit

\section{The innovative \gl{1}\gl{s_a_} markers}
\label{sec:extensions}
As discussed in \cref{sec:extensions_intro}, the person marker extensions which did not spread through the whole lexicon all have in common that they are innovative \gl{1}\gl{s_a_} markers.
There are six attested such innovations, three of which can be reconstructed to intermediate proto-languages.
The new \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
\cref{sec:pekodian} investigates the innovation of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
\cref{sec:waiwaian} takes a closer look at the extension of \rc{k-} in \PWai, which was already shown in \cref{sec:extensions_intro}.
\cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir or \PTar, reflected in modern \trio and \akuriyo.
\crefrange{sec:akuriyo}{sec:yukpa} look at innovative first person markers which are only attested in single modern languages:
\obj{k-} in \akuriyo, \obj{j(i)-} in \carijo, and \obj{j(ɨ)-} \yukpa.

\subsection{\PPek \rc{k-}}
\label{sec:pekodian}
The Pekodian branch was suggested by \textcite{meira2005southern}, as the result of fieldwork on \bakairi by Meira and the availability of more material on \ikpeng. % \parencites{pacheco1997relativizacao}{ikpengpacheco1997}{campetela1997analise}{pacheco1998nominalizacao}{ikpengpacheco2001}{campetela2002aspectos}{pacheco2002aspectos}{pacheco2003intransitivos}.
It consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
\textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work has been done on \PPek morphosyntax.
All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \exref{pekreg}.
Its presence can thus be reconstructed to \PPek.

\ex<pekreg>
\begin{tabular}[t]{@{}llll@{}}
& \bakairi \qu{to go up}  & \arara \qu{to dance}  & \ikpeng \qu{to run} \\
& \parencite[4]{meira2003bakairi} & \parencite[150]{alves2017arara} &  \parencite[52]{ikpengpacheco2001}\\
\gl{1}\gl{s_a_} & \obj{\emp{k-}əku-} & \obj{\emp{k-}origu-} & \obj{\emp{k-}aranme-} \\
\gl{2}\gl{s_a_} & \obj{m-əku-} & \obj{m-origu-} & \obj{m-aranme-} \\
\gl{1+2}\gl{s_a_} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
\gl{3}\gl{s_a_} & \obj{n-əku-} & \obj{Ø-origu-} & \obj{Ø-aranme-} \\
\end{tabular}
\xe

For \arara, \textcite{alves2017arara} reports six\footnote{Seven under her analysis, which treats the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.} \gl{s_a_} verbs which have a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
We took these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.

\ex<arairr> \arara \parencite[153]{alves2017arara} \\
\begin{tabular}[t]{@{}ll@{}}
\obj{wɨ-genɨ} & \qu{I said}\\
\obj{w-it͡ʃinɨ} & \qu{I was, lied down}\\
\obj{w-ebɨnɨ} & \qu{I came}\\
\obj{w-ibɨnɨ} & \qu{I bathed}\\
\obj{w-iptoŋrɨ} & \qu{I went down}\\
\obj{w-ɨdolɨ} & \qu{I went}\\
\end{tabular}
\xe

In his brief discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.
The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}, contrasting with regular \obj{əku} \qu{to go up} in \exref{pekreg} above.
\qu{to bathe} is also found in the \obj{w-}list for \arara.

\ex<bak-33>\bakairi \parencite[][4]{meira2003bakairi}\\
\begingl
\gla w-i-də//
\glb \gl{1}\gl{s_a_}-bathe-\gl{imm}//
\glft \qu{I bathed}//
\endgl
\xe
%
While \textcite[4]{meira2003bakairi} does list \obj{ge} \qu{to say}, \obj{tə} \qu{to go}, and \obj{əe(wɨ)} \qu{to come} as examples of \gl{s_a_} verbs, he does not indicate whether they belong to the class of \gl{s_a_}-1 verbs, with first person \obj{k-}, or the \gl{s_a_}-2 verbs, with \obj{w-}.\footnote{It should be noted at this point that \textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}. However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}. Given the lack of data for other verbs, I will not further discuss this potential additional pattern. If the characterization by \citeauthor{meira2003bakairi} is accurate, then the pattern is fully parallel to the distribution of the first person prefixes.}
Luckily, while \textcite{von1892bakairi} did not accurately record all phonemic distinctions in \bakairi \parencite{meira2005bakairi}, he does provide inflected forms of cognates to the \arara verbs in \exref{arairr}.
I present them according to the phonemic analysis of \bakairi by \textcites{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.

\pex[everyglpreamble=]<bakverbs> \bakairi \parencite[][131, 397, 76, 137, 374]{von1892bakairi}
\begin{multicols}{2}
\a<bak-3>
\begingl
\glpreamble \ort{u-ɣépa} //
\gla u-ge-pa//
\glb \gl{1}\gl{s_a_}-say-\gl{neg}//
\glft \qu{I don't say.}//
\endgl
\a<bak-5>
\begingl
\glpreamble \ort{wi-táki} / \ort{wi-tági} //
\gla w-i-taki//
\glb \gl{1}\gl{s_a_}-be-\gl{int}//
\glft \qu{I was.}//
\endgl
\a<bak-4>
\begingl
\glpreamble \ort{kχaewí-le} //
\gla k-əewɨ-lɨ//
\glb \gl{1}\gl{s_a_}-come-\gl{imm}//
\glft \qu{I came.}//
\endgl
\a<bak-6>
\begingl
\glpreamble \ort{kχ-itaké-he} //
\gla k-ɨtəgɨ-se//
\glb \gl{1}\gl{s_a_}-go.down-\gl{npst}?//
\glft \qu{I go down.}//
\endgl
\a<bak-2>
\begingl
\glpreamble \ort{ū́ta} / \ort{uúta}//
\gla u-tə//
\glb \gl{1}\gl{s_a_}-go//
\glft \qu{I go.}//
\endgl
\end{multicols}
\xe

Turning to the third Pekodian language, all available descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
However, most \ikpeng cognates of the \arara verbs in \exref{arairr} actually do not take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down} \exref{ikp-182}, but no first person forms are attested \perscommpar{Angela Chagas}.

\pex<ikpw>\ikpeng
\a<ikp-168>
\begingl
\gla ɨ-ge-lɨ//
\glb \gl{1}-say-\gl{rec}//
\glft \qu{I said.} \parencite[][209]{ikpengpacheco2001}//
\endgl
\a<ikp-169>
\begingl
\gla Ø-et͡ʃi-lɨ//
\glb \gl{1}-be-\gl{rec}//
\glft \qu{I was.} \parencite[][139]{ikpengpacheco2001}//
\endgl
\a<ikp-167>
\begingl
\gla at͡ʃagotpop Ø-ip-t͡ʃi ik-gwa-kt͡ʃi//
\glb always \gl{1}-bathe-\gl{npst} river-\gl{loc}.aquatic-\gl{all}//
\glft \qu{I always bathe in this river.} \parencite[][68]{ikpengpacheco1997}//
\endgl
\xe

\ex<ikp-170>\ikpeng \parencite[][80]{ikpengpacheco2001}\\
\begingl
\gla k-aran-t͡ʃi//
\glb \gl{1}-go-\gl{npst}//
\glft \qu{I'm going.}//
\endgl
\xe

\ex<ikp-182>\ikpeng \parencite[][320]{chagas2019ikpeng}\\
\begingl
\gla Ø-iptoŋ-lan//
\glb \gl{3}\gl{s_a_}-sit.down-\gl{pst}//
\glft \qu{She sat down.}//
\endgl
\xe
%
While clearly not a regular sound change, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \exref{ikpwloss}.
Thus, we suggest that the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}, and was subject to phonological erosion of \rc{w}, a phenomenon which will also be encountered in languages discussed in following sections.

\ex<ikpwloss> %\arara-\ikpeng \obj{w}:Ø correspondences\\
\begin{tabular}[t]{@{}lll@{}}
\PXin & \arara & \ikpeng\\
\rc{wɨna} \qu{\gl{dat}} & \obj{wɨna} & \obj{ɨna}\\
\rc{wokori} \qu{dog, jaguar} & \obj{wokori} & \obj{akari}\\
\rc{tawe} \qu{capuchin monkey} & \obj{tawe} & \obj{tae}\\
\rc{watke} \qu{to defecate} & \obj{watke} & \obj{atke}\\
\rc{wɨnkɨ} \qu{to sleep} & \obj{wɨnkɨ} & \obj{ɨnkɨ}\\
& (\obj{o-wɨnkɨ} \qu{sleep!}) & (\obj{j-ɨnkɨ-lɨ} \qu{s/he slept})\\
\end{tabular}\\
\parencites[80]{alves2017arara}[164, 67, 150]{ikpengpacheco2001}[9]{desouza2010arara}[68]{souza1993arara}[91]{alves2013verbo}[40]{campetela1997analise} 
\xe

\begin{table}
	\centering
	\caption{Verbs preserving \gl{1}\gl{s_a_} in \PPek}
	\label{tab:pekverbs}
	\begin{tabular}{@{}lllll@{}}
	\mytoprule
& \PPek & \arara & \ikpeng & \bakairi \\
\mytoprule
\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} & \obj{u-ge-} \\
\qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{Ø-ip-} & \obj{w-i-} \\
\qu{be} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
\qu{come} & \rc{w-epɨ}, \rc{k-əd-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
\qu{go down} & \rc{w-[ɨ/i]ptə-} & \obj{w-iptoŋ-} & ?\obj{-iptoŋ-} & \obj{k-ɨtəgɨ-} \\
\qu{go} & \rc{w-ɨtən-} & \obj{w-ɨdo-} & \obj{k-aran-} & \obj{u-tə-} \\
	\mybottomrule
	\end{tabular}
\end{table}

\Cref{tab:pekverbs} gives an overview of the first person forms of the six verbs under discussion, along with our \PPek reconstruction.
The form for \rc{ge} \qu{to say} is straightforward to reconstruct; the \bakairi change of \rc{wɨ-} to \obj{u-} is not implausible given correspondences like \bakairi \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
\qu{to bathe} is straightforwardly reconstructible as \rc{ipɨ}.
For \qu{to be}, \ikpeng \obj{e} is very likely the original vowel, given the \PC form \rc{eti} \pcref{sec:be}.

The forms for \qu{to come} are not fully cognate; \ikpeng and \bakairi both show a reflex of the detransitivizer in combination with a root reconstructible as \rc{epɨ}.
In contrast, the \arara first person form is directly based on this root \rc{epɨ}.
However, other person values show a reflex of \detrz in \arara, too \exref{ara-123}.

\ex<ara-123>\arara \parencite[][150]{alves2017arara}\\
\begingl
\gla m-odebɨ-nɨ//
\glb \gl{2}\gl{s_a_}-come-\gl{rec}//
\glft \qu{You came.}//
\endgl
\xe
%
Thus, it seems that there was some variation in \PPar as to whether or not the first person form of \qu{to go} was based on \rc{ətepɨ} or \rc{epɨ}.
Given that \bakairi and \ikpeng agree about both the first person prefix and the reflex of \rc{ət-}, I reconstruct two distinct first person forms, \rc{w-epɨ-} and \rc{k-ət-epɨ-}.\footnote{It may also be the case that \bakairi and \ikpeng both independently regularized the paradigm to show \rc{ədepɨ} with all persons. If that were the case, the connection between \rc{k-} and \rc{ədepɨ} would still exist.}

The forms for \qu{to go down} are again not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
Rather, it seems that some roots gained a final \obj{ŋ} in \PXin, for so far unclear reasons: \begin{inlinelist}
\item \PC \rc{əne} \qu{to see}, \arara and \ikpeng \obj{eneŋ} \parencites[8]{gildea2007greenberg}[56]{alves2017arara}[25]{ikpengpacheco2001}
\item \PC \rc{əta} \qu{to hear}, \arara \obj{taŋ}, \ikpeng \obj{iraŋ} \parencites[8]{gildea2007greenberg}[144]{alves2017arara}[270]{ikpengpacheco2001}
\item \PC \rc{ənə} \qu{to eat meat}, \arara \obj{oŋoŋ} \qu{to bite} \parencites[8]{gildea2007greenberg}[57]{alves2017arara}
\end{inlinelist}.
Also, based on the fact \bakairi has generally lost much segmental material, I suggest that \bakairi \obj{gɨ} is yet another addition to a root \rc{ɨptə} or \rc{iptə}, rather than a conservative form.

The forms for \qu{to go} are all cognate, but while \arara and \bakairi point to \rc{ɨtə}, the \ikpeng form is rather divergent.
However, it is in fact compatible with a reconstruction \rc{ɨtən}, when one considers that \ikpeng \obj{a} is an attested outcome of \rc{ə} (\PXin \rc{o}/\obj{e}), for example \obj{akari} \qu{dog} in \exref{ikpwloss} above, \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg}, or \obj{jaj} \qu{tree} from \PC \rc{jəje}.
This attested change of \rc{ə} to \obj{a} need only be preceded by a assimilatory lowering of initial \rc{ɨ} to \rc{ə}, to yield the form \obj{aran} from \rc{ɨtən}.
Other \ikpeng reflexes of \qu{to go} offer evidence for the suggested intermediate stage \rc{ətən}: \obj{ero-lɨ} \qu{s/he went} \parencite[25]{ikpengpacheco2001}.

Summing up, an innovative \gl{1}\gl{s_a_} marker \rc{k-} is reconstructible to \PPek.
Six verbs can be reconstructed as having resisted this innovation, and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
For \qu{to come}, two distinct forms \rc{w-epɨ-} and \rc{k-əd-epɨ-} must be reconstructed.
In later, individual developments, \bakairi introduced \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.

\subsection{\PWai \rc{k-}}
\label{sec:waiwaian}
This extension was briefly discussed in \cref{sec:extensions_intro}; the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was later extended to \gl{1}\gl{s_a_} in \PWai.
This created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \exref{wais}.

\ex<wais> Waiwaian \rc{eɸurka} \qu{to fall} (\gl{s_a_}) and \rc{wɨnɨkɨ} \qu{to sleep} (\gl{s_p_})\\
\begin{tabular}[t]{@{}lllllll@{}}
& \multicolumn{2}{l}{\PWai} & \multicolumn{2}{l}{\hixka}  & \multicolumn{2}{l}{\waiwai} \\
 & \gl{s_a_} & \gl{s_p_} & \gl{s_a_} & \gl{s_p_} & \multicolumn{2}{l}{\gl{s}} \\
\gl{1} & \rc{\emp{k-}eɸurka-} & \rc{kɨ-wɨnɨkɨ-} & \obj{\emp{k-}ehurka-} & \obj{kɨ-nɨkɨ-} & \obj{\emp{k-}eɸɨrka-} & \obj{kɨ-wɨnk-} \\
\gl{2} & \rc{m-eɸurka-} & \rc{o-wɨnɨkɨ-} & \obj{m-ehurka-} & \obj{o-wnɨkɨ-} & \obj{m-eɸɨrka-} & \obj{mɨ-wɨnk-} \\
\gl{1+2} & \rc{t-eɸurka-} & \rc{tɨt-wɨnɨkɨ-} & \obj{t-ehurka-} & \obj{tɨ-nɨkɨ-} & \obj{t͡ʃ-eɸɨrka-} & \obj{tɨt-wɨnk-} \\
\gl{3} & \rc{ɲ-eɸurka-} & \rc{nɨ-wɨnɨkɨ-} & \obj{ɲ-ehurka-} & \obj{nɨ-nɨkɨ-} & \obj{ɲ-eɸɨrka-} & \obj{nɨ-wɨnk-} \\
\end{tabular}\\
\parencites[510]{howard2001wrought}[189--191]{hixkaryanaderby1985}[209--211]{hawkins1953waiwai}[50]{waiwaihawkins1998}
\xe
%
The \waiwai verb \obj{ka} \qu{to say} was already shown have resisted this extension in \cref{sec:extensions_intro}, showing \obj{wɨ-}, a reflex of the \PC \gl{1}\gl{s_a_} marker \rc{w-}.
Its \hixka counterpart also takes a reflex of \gl{1}\gl{s_a_} \rc{w-}, which has the form \obj{ɨ-} \exref{hixɨ.hix-118}.
A formally identical prefix is found on transitive verbs in \gl{1}>\gl{3} scenarios \exref{hixɨ.hix-17}, where it also corresponds to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.

\pex<hixɨ>
\a<hix-118> \hixka \parencite[][124]{hixkaryanaderby1985}\\
\begingl
\glpreamble roxehra nay hamɨ Kaywerye ɨkekonɨ//
\gla ro-ʃe-hɨra n-a-je hamɨ kajwerʲe ɨ-ka-jakonɨ//
\glb \gl{1}-\gl{des}-\gl{neg} \gl{3}-be-\gl{npst}.\gl{uncert} \gl{evid} K. \gl{1}\gl{s_a_}-say-\gl{rem}.\gl{cont}//
\glft \qu{I said (to myself), “Kaywerye evidently doesn't like me”.}//
\endgl
\a<hix-17> \hixka \parencite[][191]{hixkaryanaderby1985}\\
\begingl
\gla ɨ-koroka-no//
\glb \gl{1}>\gl{3}-wash-\gl{imm}//
\glft \qu{I washed him.}//
\endgl
\a<wai-159> \waiwai \parencite[][192]{waiwaihawkins1998}\\
\begingl
\glpreamble wîyesî//
\gla wɨ-jo-jasɨ//
\glb \gl{1}>\gl{3}-boil-\gl{npst}//
\glft \qu{I will boil it.}//
\endgl
\xe
%
Thus, similar to the the case of \ikpeng, \hixka has lost the \rc{w} of the \envr{}{C} allomorph \rc{wɨ-}.
Notably, \textcite{hixkaryanaderby1985} does not see this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_}, but as the regular \gl{1}>\gl{3} marker, because he considers \hixka \obj{ka} \qu{to say} to be transitive, an issue which will be discussed in \cref{sec:say}.

There are two more verbs which did not take innovative \rc{k-} in \PWai.
One of them is the copula \rc{eʃi}/\obj{a(h)}, reconstructed in \exref{parpar.parcop}.
The other is \qu{to go} \exref{parpar.pargo}, somewhat of a special case.

\pex<parpar>
\a<parcop> Parukotoan \qu{to be}\\
\begin{tabular}[t]{@{}llll@{}}
& \PPar & \hixka & \waiwai \\
\gl{1}\gl{s_a_} & \rc{\emp{w-}eʃi-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eʃe-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eeʃi-}/\obj{\emp{w-}a-}\\
\gl{2}\gl{s_a_} & \rc{m-eʃi-}/\obj{m-ah-} & \obj{m-eʃe-}/\obj{m-ah-} & \obj{m-eeʃi-}/\obj{m-a-}\\
\gl{1+2}\gl{s_a_} & \rc{t-eʃi-}/\obj{t-ah-} & \obj{t-eʃe-}/\obj{t-ah-} & \obj{t͡ʃ-eeʃi-}/\obj{t-a-}\\
\gl{3}\gl{s_a_} & \rc{n-eʃi-}/\obj{n-ah-} & \obj{n-eʃe-}/\obj{n-ah-} & \obj{n-eeʃi-}/\obj{n-a-}\\
\end{tabular}\\
(\cite[197--198]{hixkaryanaderby1985} and Spike Gildea, p.c.)
\a<pargo> Parukotoan \qu{to go}\\
\begin{tabular}[t]{@{}llll@{}}
& \PPar & \hixka & \waiwai\\
\gl{1}\gl{s_a_} & \rc{\emp{wɨ-}to-} & \obj{\emp{ɨ-}to-} & \obj{kɨ\emp{w-}to-}\\
\gl{2}\gl{s_a_} & \rc{mɨ-to-} & \obj{mɨ-to-} & \obj{mɨɨ-to-}\\
\gl{1+2}\gl{s_a_} & \rc{tɨt-to-} & \obj{tɨ-to-} & \obj{tɨ-to-}/\obj{tɨh-t͡ʃe-}\\
\gl{3}\gl{s_a_} & \rc{nɨ-to-} & \obj{n-to-} & \obj{nɨɨ-to-}\\
\end{tabular}\\
(\cites[69--70, 211]{hixkaryanaderby1985}[179]{waiwaihawkins1998} and Spike Gildea, p.c.)
\xe
%
While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \gl{1}\gl{s_a_} \obj{k-} with the old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
Alternatively, this form was influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \exref{waiwto}.

\pex<waiwto> \waiwai reflexes of the \gl{s_a_} class marker \rc{w-}
\a \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}
\a \obj{o-\emp{w}to-t͡ʃhe} \qu{after I went} \parencite[165]{waiwaihawkins1998}
\a \obj{kɨ-\emp{w}to-me} \qu{for us to go} \parencite[204]{waiwaihawkins1998}
\xe
%
In any case, \hixka \qu{to go} was clearly not affected by the extension of \rc{k-}, allowing us to reconstruct a \PWai first person form \rc{wɨ-to-}.

Summing up, I reconstruct the three verbs \rc{et͡ʃi}/\obj{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{to} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.

\subsection{\PTir \rc{t-}}
\label{sec:taranoan}
The Taranoan languages were already grouped together by \textcite{girard1971proto}, consisting of closely related \trio and \akuriyo (subsumed here under the moniker Tiriyoan), and the more distant member \carijo.
\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar.
The \setone paradigms of \trio and \akuriyo contain an interesting puzzle: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
This is illustrated by comparing reconstructible \PTir forms with more conservative \wayana \exref{sat} and \kalina \exref{13w}, which can be taken as representative for \PC.

\ex<sat>
\begin{tabular}[t]{@{}llllll@{}}
\gl{1}\gl{s_a_} & \PC & \wayana & \PTir & \trio & \akuriyo\\
\qu{yawn} & \rc{w-e-mɨta-po-} & \obj{w-emtapɨ-} & \rc{t͡ʃ-entapo-} & \obj{s-entapo-} & \obj{t͡ʃ-entapo-}\\
\qu{bathe} & \rc{w-e-pɨ-} & \obj{w-epɨ-} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-}\\
\qu{talk} & \rc{w-ət-uru} & \obj{w-əturu-} & \rc{t-əturu-} & \obj{t-əturu-} & \obj{k-əturu-}\footnotemark\\
\qu{learn} & \rc{w-ət-eme-pa} & \obj{w-əhepa-} & \rc{t-əempa-}\footnotemark & \obj{t-əenpa-} & \obj{k-əempa-}\\
\end{tabular}\\
\parencites[28]{camargo2010wayana}[292, 664]{triomeira1999}[77, 86, 82]{gildea1994akuriyo}[50, 208, 213]{wayanatavares2005}
\xe
\addtocounter{footnote}{-1}
\footnotetext{\PTir \rc{t-} was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
\addtocounter{footnote}{1}
\footnotetext{
The \trio form is not due to dissimilation from \rc{mp} to \obj{np}, but rather a phonemic transcription rendering the syllable-final archiphoneme \pnm{N} as \obj{n} \parencite[31]{triomeira1999}.
Further, \PTar intervocalic \rc{t͡ʃ} was lost entirely in \trio (except after \rc{i}) and is reflected as \obj{ʔ} in \akuriyo \parencite[31--32]{meira1998proto}.
Given the original shape of the detransitivizers \rc{ət(e)-} and \rc{e-}, \trio \obj{ə-enpa} \qu{to learn} contains a reflex of the former with loss of \rc{t}, so \akuriyo \obj{əempa} would be expected to have a glottal stop: ?\obj{əʔempa}.
Its absence could either be due to \trio influence or not being transcribed by \textcite{gildea1994akuriyo}.
Since there is no evidence for a glottal stop in \PTir (\rc{əʔempa}), I do not reconstruct it.}

\ex<13w> 
\begin{tabular}[t]{@{}llllll@{}}
\gl{1}>\gl{3}& \PC & \kalina & \PTir & \trio & \akuriyo \\
\qu{see} & \rc{t-ene-} & \obj{s-ene-} & \rc{w-ene-} & \obj{w-ene-} & \obj{(w-)ene-} \\
\qu{bring} & \rc{t-enepɨ-} & \obj{s-enepɨ-} & \rc{w-enepɨ-} & \obj{w-enepɨ-} & Ø\obj{-enepɨ-} \\
\end{tabular}\\
\parencites[425--426]{hoff1968carib}[28, 304]{triomeira1999}[86--87]{gildea1994akuriyo}
\xe
%
Thus, \wayana preserves the \PC \gl{1}\gl{s_a_} marker \rc{w-}, while it was replaced by \rc{t-} (\phonr{}{\rc{t͡ʃ-}}{\rc{e}}) in \PTir \exref{sat}.
%Both \qu{to yawn} and \qu{to bathe} are transparently derived \gl{s_a_} verbs, from \rc{mɨta-po} \qu{to open the mouth of} and \rc{pɨ} \qu{to wash}, respectively.
On the other hand, \gl{1}>\gl{3} scenarios show a reflex of \rc{t-} in \kalina, while they took on \rc{w-} in \PTir \exref{13w}.

The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}\footnote{In fact, the issue of \emph{when} this happened is also open. It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation. Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.}
The relevant aspect of this extension is of course that not all verbs were affected by it.
\textcite{meira1998proto} lists the four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}.

\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-}\\
\begin{tabular}[t]{@{}llll@{}}
& \PTir & \trio & \akuriyo\\
\qu{go} & \rc{wɨ-tə(mɨ)-} & \obj{wɨ-tə(mɨ)-} & \obj{ə-təmɨ-}/\obj{wɨ-təemɨ-}  \\
\qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
\qu{come} & \rc{w-əʔepɨ-/}\obj{w-eepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
\qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
\end{tabular}\\
\parencites[112--115]{meira1998proto}[85]{gildea1994akuriyo}
%\a<aku12> \akuriyo \obj{kəʔ-eepɨ} \qu{we came} \parencite[114]{meira1998proto}
%\a<aku13> \akuriyo \obj{(w)i-toka} \qu{I hit him/her} \parencite[86]{gildea1994akuriyo}
%\a<tricome> \trio \obj{n-epɨ} \qu{s/he came} \parencite[114]{meira1998proto}
\xe

For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} gives a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
The two forms can be reconciled by a specific idiosyncratic combination of sound changes:
I suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
Also, \obj{w-} appears to have been subject to ongoing erosion in \akuriyo, also evidenced by its absence in \qu{to come} and \qu{to be}, but its presence in \qu{to say} \exref{ptirw}.
This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix (\getfullref{13w},  \getfullref{exp.aku-164}), as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.

\pex<exp>
\a<aku-159> \akuriyo \parencite[][114]{meira1998proto}\\
\begingl
\gla kəʔ-eepɨ//
\glb \gl{1+2}-come//
\glft \qu{We came.}//
\endgl
\a<aku-164> \akuriyo \parencite[][86]{gildea1994akuriyo}\\
\begingl
\gla (w)i-toka//
\glb \gl{1}>\gl{3}-hit//
\glft \qu{I hit him/her.}//
\endgl
\xe

For \qu{to come}, a similar situation as in Pekodian is encountered, in that \trio shows a reflex of \rc{ət-epɨ}, while the \akuriyo form points to bare \rc{epɨ}.
Given that \trio, like \arara, also has forms without a reflex of \rc{ət-} \exref{tri-130}, I reconstruct both \rc{w-epɨ-} and \rc{w-əʔepɨ-} for \PTir.

\ex<tri-130> \trio \parencite[][114]{triomeira1999}\\
\begingl
\gla n-epɨ//
\glb \gl{3}-come//
\glft \qu{S/he came.}//
\endgl
\xe

\textcite{gildea1994akuriyo} also recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-}.
They are listed in \exref{akumov.aku}.
I have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in that it takes \obj{s-} \exref{akumov.tri}.
Thus, it is uncertain whether these verbs can be reconstructed as taking \rc{w-} in \PTir, although its presence in \akuriyo does suggest so.

\pex<akumov>
\a<aku> \akuriyo \gl{1}\gl{s_a_} \rc{w-} \parencite[84--86]{gildea1994akuriyo}\\
\begin{tabular}[t]{@{}ll@{}}
\qu{return} & Ø\obj{-erama-}\\
\qu{get up} & Ø\obj{-eokahtə-}\\
\qu{jump} & \obj{w-ejahka-}\\
\qu{go out} & \obj{w-ekɨrɨka-}\\
\end{tabular}
\a<tri> \trio \obj{s-erama-} \parencite[301]{triomeira1999}
\xe

Finally, there are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
They have \gl{1}\gl{s_a_} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
At least for \qu{to go down}, the \akuriyo cognate suggests that the irregular first person form \rc{p-ɨhtə-} can be reconstructed to \PTir.
Whatever their origins, they were not affected by the extension of \rc{t-}.

\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes\\
\begin{tabular}[t]{@{}lll@{}}
& \trio & \akuriyo\\
\qu{go down} & \obj{p-ɨhtə-} & \obj{p-ɨtə-}\\
\qu{defecate} & \obj{k-oeka-} & ?\\
\end{tabular}\\
\parencites[294]{triomeira1999}[84]{gildea1994akuriyo}
\xe

In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} was extended to \gl{1} verbs.
Four verbs are solidly reconstructible as preserving \rc{w-} in \PTir, and \rc{ɨhtə} \qu{to go down} with irregular \rc{p-} was also not affected.
There are four movement verbs in \akuriyo which appear to not have been affected by the extension of \rc{t-}, as well as another irregular verb \obj{weka} \qu{to defecate} in \trio, with first person \obj{k-oeka-}.

\subsection{\akuriyo \obj{k-}}
\label{sec:akuriyo}
After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
It seems to have replaced \rc{t-} only in specific places, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to me \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
\textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
He also suggests that \obj{k-} could be more recent, with which I agree -- since phonologically distributed \rc{t-}/\rc{t͡ʃ-} is reconstructible to \PTir, a scenario whereby \obj{k-} replaces \rc{t-}, but not \rc{t͡ʃ-}, is the most straightforward.
The few \obj{t-} recorded by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.

\begin{table}
	\centering
	\caption{\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes}
	\label{tab:aku1sa}
	\begin{tabular}{@{}lll@{}}
	\mytoprule
first person \obj{k-} & first person \obj{t͡ʃ-} \\
\mymidrule
\obj{ənɨkɨ} \qu{to sleep} & \obj{eepɨ} \qu{to bathe} \\
\obj{əməmɨ} \qu{to enter} & \obj{ewai} \qu{to sit down} \\
\obj{əturu} \qu{to talk} & \obj{etonema} \qu{to lie down} \\
\obj{əət͡ʃena} \qu{to cry} & \obj{ekɨɨrɨka} \qu{to stay back} \\
\obj{ətajiŋka} \qu{to run} & \obj{entapo} \qu{to yawn} \\
\obj{əiwa} \qu{to tremble} &  \\
\obj{əempa} \qu{to learn} \\
	\mybottomrule
	\end{tabular}
\end{table}

%Let us first consider a scenario where \obj{k-} was the older \gl{1}\gl{s_a_} marker, occurring on both \obj{e}- and \obj{ə}-initial verbs.
%Next, \obj{t-}/\obj{t͡ʃ-} would have come in and replaced \obj{k-} before \obj{e}, but (mostly) not before \obj{ə}, resulting in the observable distribution.
%This scenario is not very satisfactory, since \obj{k-} is replaced in one completely unmotivated phonological environment, but not in the other.
%
%The other scenario is that \rc{t-} was the old first person marker occurring on the vast majority of \gl{s_a_} verbs.
%Subsequently, it developed into \rc{t͡ʃ-} / \_\obj{e}, but remained \rc{t-} / \_\obj{ə}; \obj{t͡ʃ}\goodtilde\obj{ʃ}\goodtilde\obj{s} and \obj{t} are phonemically contrastive in \akuriyo \parencite[16]{meira1998proto}.
%Thus, \obj{k-} can be seen as replacing \obj{t-}, but not \obj{t͡ʃ-}.
%The occurrences of \obj{t-} / \_\obj{ə} mentioned by \citeauthor{meira1998proto} are then remnants of the earlier prevalence of \obj{t-}.
%While this scenario also lacks a clear motivation, it at least explains the largely phonologically complementary distribution of \obj{k-} and \obj{t-} by the fact that the marker which was replaced had a phonemically different form than the one which was preserved.
%The other scenario which sees \rc{k-} as older would require \rc{t-} replacing \obj{k-} only in a specific environment.\footnote{Note that this process would be plausible even if the instances of \obj{t-} / \_\obj{ə} were due to \trio influence, as suggested by \textcite[108]{meira1998proto} -- these would then be a later reintroduction, replacing \obj{k-}, instead of remnants of the introduction of it.}

It has been speculated that the \trio \gl{s_a_} verb with a first person \obj{k-} marker, \obj{weka}/\obj{oeka} \qu{to defecate} might have something to do with \akuriyo \obj{k-}, which would potentially make the innovation of \rc{k-} a \PTir matter \parencite[116]{meira1998proto}.
However, this hypothesis faces a problem in that the first person form for \akuriyo \qu{to defecate} is \obj{j-ereina-} \parencite[88]{gildea1994akuriyo}.
Another hypothesis is that it originated in the corresponding \kaxui form \obj{ku-weka-} (form provided by Spike Gildea, p.c.), although the occurrence of \obj{o} in \trio would still need explanation.
%Rather, it seems likely that this irregular first person form originated in a class switch from \gl{s_p_} to \gl{s_a_} (\Cref{sec:resistant_verbs}), although it is unclear how exactly that switch resulted in the irregular form.

\subsection{\carijo \obj{j(i)-}}
\label{sec:carijo}
As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.
This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \exref{carpar}.

\ex<carpar> \carijo \parencites[106]{meira1998proto}[173]{robayo2000avance}\\
\begin{tabular}[t]{@{}lll@{}}
& \obj{tuda} \qu{to arrive} & \obj{eharaga} \qu{to dance}\\
\gl{1} & \obj{ji-tuda-} & \obj{j-eharaga-}\\
\gl{2} & \obj{mɨ-tuda-} & \obj{m-egaraga-}\\
\gl{1+2} & \obj{kɨsi-tuda-} & \obj{kɨs-eharaga-}\\
\gl{3} & \obj{ni-tuda-} & \obj{n-eharaga-}\\
\end{tabular}
\xe
%
Thus, also detransitivized verbs like \obj{ehɨnəhɨ} \qu{to fight}, which are \gl{s_a_} verbs in other languages, take \obj{j(i)-} \exref{car-30}.
The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.

\ex<car-30>\carijo \parencite[][79]{koch1908hiana}\\
\begingl
\glpreamble hẹnẹ(x)tónoko-mā́ɽẹ y-e̱-hẹnẹ́(x)yai̯//
\gla hɨnəhtono-ko=marə j-e-hɨnəh-jai//
\glb enemy-\gl{pl}=with \gl{1}-\gl{detrz}-kill-\gl{npst}.\gl{cert}//
\glft \qu{I fight with the enemies.}//
\endgl
\xe
%

\pex<car1p>\carijo \parencite[][59, 53]{guerrero2016carijo}
\begin{multicols}{2}
\a<car-1>
\begingl
\gla jɨ-hɨnəh-ɨ//
\glb \gl{3}>\gl{1}-kill-\gl{pfv}//
\glft \qu{S/he killed me.}//
\endgl
\a<car-31>
\begingl
\gla j-owo-rɨ//
\glb \gl{1}\gl{poss}-uncle-\gl{pert}//
\glft \qu{my uncle}//
\endgl
\end{multicols}
\xe
%

As noted in \cref{sec:taranoan}, this extension also erased any traces of a putative \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.

\pex<car1w>\carijo \parencite[][5, 42]{guerrero2016karihona}
\a<car-24>
\begingl
\gla wɨ-tə-e=rehe//
\glb \gl{1}-go-\gl{npst}=\gl{frust}//
\glft \qu{I almost go (but I a, not going to go).}//
\endgl
\a<car-25>
\begingl
\gla əji-marə-ne w-a-e//
\glb \gl{2}-with-\gl{pl} \gl{1}-be-\gl{npst}//
\glft \qu{I am with you all.}//
\endgl
\xe
%
%Interestingly enough, the verb \qu{to be} only shows \obj{w-} in its \obj{a} root allomorph, but not in the \obj{et͡ʃi} root allomorph, where it has \obj{j-} \exref{car1y.mivida-12}.
%In a somewhat foreshadowing move, I will note that \PTar \rc{(ət)epɨ} \qu{to come} was also affected by the \carijo extension, as seen in \exref{car1y.car-18}.
%As discussed in \cref{sec:taranoan}, this verb preserves \rc{w-} in \akuriyo and \trio \exref{tri-105}.
%
%\pex<car1y>\carijo
%\a<mivida-12>
%\begingl
%\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
%\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
%\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
%\glft \qu{Then I was already grown up.} \parencite[][177]{robayo1989rame}//
%\endgl
%\a<car-18>
%\begingl
%\gla əji-wa-e j-eh-ɨ//
%\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
%\glft \qu{I came looking for you.} \parencite[][102]{guerrero2019carijo}//
%\endgl
%\xe
%
%\ex<tri-105>\trio \parencite[][337]{triocarlin2004}\\
%\begingl
%\gla ene w-əe//
%\glb see.\gl{sup} \gl{1}\gl{s_a_}-come//
%\glft \qu{I came to see him/her.}//
%\endgl
%\xe

\subsection{\yukpa \obj{j-}}
\label{sec:yukpa}
\yukpa has lost most of its \setone constructions, mostly employing innovative ones \parencite{meira2006syntactic}.
It preserves the \setone prefixes in the immediate past, where the \gl{1}\gl{s_p_} marker appears to have been extended to \gl{s_a_} verbs, as evidenced by the paradigms in \exref{yukpaintr}.
The loss of split-\gl{s} in favor of the \gl{1}\gl{s_p_} marker is accompanied by the \gl{2}\gl{s_a_} marker \obj{m(ɨ)-} displacing \gl{2}\gl{s_p_} \rc{ə(j)-}, and \gl{1+2} being lost as an inflectional value.
%It has lost the split-\gl{s} system in general

\ex<yukpaintr> \yukpa \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}\\
\begin{tabular}[t]{@{}llll@{}}
& \obj{otum} \qu{to wash self} & \obj{nɨ} \qu{to sleep} & \obj{ata} \qu{to fall}\\
\gl{1} & \obj{j-otum-} & \obj{jɨ-nɨ-} & \obj{j-ata-}\\
\gl{2} & \obj{m-otum-} & \obj{mɨ-nɨ-} & \obj{m-ata-}\\
\gl{3} & \obj{n-otum-} & \obj{nɨ-nɨ-} & \obj{n-ata-}\\
\end{tabular}
\xe
%
The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of \detrz, like \obj{otum} \qu{to wash self} in \exref{yukpaintr}.
The analysis that this is a reflex of the \gl{1}\gl{s_p_} marker \rc{u(j)-} is supported by the same form occurring in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.

\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw j-esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw Ø-esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian,sec:akuriyo}, I argue that this zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
The same zero marking is attested for a single intransitive verb: \obj{to} \qu{to go} \exref{yuk-7}.
Since regular C-initial verbs take \obj{jɨ-}, like \qu{to sleep} in \exref{yukpaintr}, I argue that \obj{to} \qu{to go} preserved the old \gl{1}\gl{s_a_} marker \rc{w(ɨ)-}, which was lost due to phonological erosion.

\ex<yuk-7>\yukpa \parencite[][139]{meira2006syntactic}\\
\begingl
\gla aw Ø-to//
\glb \gl{1}\gl{pro} \gl{1}\gl{s_a_}-go//
\glft \qu{I went.}//
\endgl
\xe

### 2021-09-10 documents/sections/introduction.tex
initial commit

\section{Introduction}
\label{sec:intro}

\input{stuff/cariban_tree}

### 2021-09-10 documents/sections/proto_carib.tex
initial commit

\section{Cariban verbal person marking and marker extensions}
The archaic person markers discussed in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system.
This section provides the necessary background for said ancestral system \pcref{sec:pc_person}, the split-\gl{s} system which will be relevant later \pcref{sec:split}, and the concept of person marker extensions \pcref{sec:extensions_intro}.

\subsection{Verbal person marking in \PC}
\label{sec:pc_person}
\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.\footnote{I use standard IPA symbols in my transcription of Cariban languages, with the exception of coronal rhotics, which I simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc. In languages with strong morphophonological processes and/or subphonemic orthography I show the original transcription in an additional surface line when presented in an interlinearized glossed example. I follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
The locuphoric markers, those referring to \gl{sap} participants, had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
This was not the case for third person referents, which only had one marker, since third person only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, and third-person prefixes did not distinguish different different third persons.
Local scenarios were expressed in a non-transparent manner, by using the \gl{1+2} prefix \rc{k-} in both cases.\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value. This is expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented in this table.}

\begin{table}
	\centering
	\caption{\PC \setone (main clause) person markers \parencites[495]{meira2010origin}[497]{gildea2016referential}}
	\label{tab:pcpers}
\begin{subtable}[b]{.49\linewidth}
\caption{Transitive}
\label{tab:pctrans}
\centering
	\begin{tabular}{@{}lllll@{}}
	\mytoprule
\gl{a}/\gl{p}&		\gl{1}	&	\gl{2}		&	\gl{1+2}	&	\gl{3}	\\
\mymidrule
\gl{1}	&		&	\rc{k-}	&				&	\rc{t(i)-}		\\	
\gl{2}	&	\rc{k-}			&&				&	\rc{m(i)-}		\\
\gl{1+2}&		&				&				&	\rc{kɨt(i)-}		\\
\gl{3}	&	\rc{u(j)-}	&	\rc{ə(j)-}	&	\rc{k-}			&	\rc{n(i)-}		\\
	\mybottomrule
	\end{tabular}
\end{subtable}%
\begin{subtable}[b]{.49\linewidth}
\caption{Intransitive}
\label{tab:pcintrans}
\centering
\begin{tabular}{@{}lll@{}}
\mytoprule
& \gl{s_a_} & \gl{s_p_}  \\
\mymidrule
\gl{1} & \rc{w-} & \rc{u(j)-} \\
\gl{2} & \rc{m-} & \rc{ə(j)-}\\
\gl{1+2} & \rc{kɨt-} & \rc{k-}\\
\gl{3} & \rc{n-} & \rc{n(i)-}\\
\mybottomrule
\end{tabular}	
\end{subtable}
\end{table}

Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}), as well as the absence of \rc{i} after all \gl{s_a_} prefixes.
On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones.
The third person marker in \gl{s_p_} verbs was identical to the one in \gl{3}>\gl{3} scenarios (\rc{n(i)-)}, while the \gl{3}\gl{s_a_} marker did not have an \rc{i}, like the other \gl{s_a_} prefixes.

As in many other split-\gl{s} systems found in languages of the world, the intransitive verbal lexicon was divided into the two classes, with verbs inherently being \gl{s_a_} or \gl{s_p_} verbs.
That is, while in transitive verbs, the choice of the person marker had a crucial semantic contribution, it is predictable in intransitive verbs.
This is illustrated with modern \kalina data in \exref[kar1]{kar2}.

\pex<kar1> \kalina
\a<kar-60>
\begingl
\gla mi-kupi-ja//
\glb \gl{2}>\gl{3}-bathe-\gl{prs}//
\glft \qu{You bathe him/her.} \parencite[][160]{hoff1968carib}//
\endgl
\a<kar-62>
\begingl
\gla a-kupi-ja//
\glb \gl{3}>\gl{2}-bathe-\gl{prs}//
\glft \qu{S/he bathes you.} \parencite[][63]{yamada2011evidentiality}//
\endgl
\xe
%
In \exref{kar1}, the choice between the second person \gl{a}- and \gl{p}-oriented markers \obj{mi-} and \obj{a-} depends on the scenario:
The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
While intransitive verbs show the same (or very similar) person markers, they contribute no semantic difference here \exref{kar2}.

\pex<kar2> \kalina
\a<kar-59>
\begingl
\gla sipi tɨnka-rɨ m-ekema-non hen//
\glb net pull-\gl{nmlz} \gl{2}-be.afraid-\gl{prs}.\gl{uncert} eh?//
\glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
\endgl
\a<kar-61>
\begingl
\gla aj-awoi-ja//
\glb \gl{2}-get.up-\gl{prs}//
\glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
\endgl
\xe
%
Rather, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.\footnote{The root \obj{awomɨ} \qu{to get up} is subject to syllable reduction and assimilation to the prefix-initial \obj{j}.}

In fact, the split-\gl{s} system is not only fully lexically conditioned, but there are clear semantic mismatches between class membership and semantics:
\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, in fact  \qu{you're afraid of pulling up the net} clearly has a second person patient (experiencer).
Similarly, \qu{to get up} with a \dbqu{patientive} marker is a clear semantic mismatch as well.
These mismatches are not isolated cases, as will be discussed in \cref{sec:split}.

\subsection{Defining features and origins of the split-\gl{s} system}
\label{sec:split}
As seen in the previous section, the split-\gl{s} defined two inflectional classes for intransitive verbs within the \setone system.
%\kalina paradigms illustrating the split for each person are given in \cref{tab:kalinaparadigms}.
%\gl{3}\gl{s_a_} differs from \gl{3}\gl{s_p_} by \begin{inlinelist}
%\item not showing the allomorph \obj{ni-} before consonants
%\item not triggering the umlaut of verb-initial \phon{\obj{o}}{\obj{e}}
%\end{inlinelist}, which is shown in \cref{tab:kalinaparadigms}.
%
%\begin{table}[h]
%	\centering
%	\caption{\kalina \gl{s_a_} and \gl{s_p_} person paradigms \parencites[178]{hoff1968carib}[430, 73]{courtz2008carib}[481]{meira2010origin}}
%	\label{tab:kalinaparadigms}
%	\begin{tabular}{@{}lll@{}}
%	\mytoprule
%Person & \gl{s_a_} \obj{opɨ} \qu{to come} & \gl{s_p_} \obj{[o/e]kanumɨ} \qu{to run}\\
%	\mymidrule
%\gl{1} & \obj{w-opɨ-} & \obj{j-ekanumɨ-}\\
%\gl{2} & \obj{m-opɨ-} & \obj{aj-ekanumɨ-}\\
%\gl{1+2} & \obj{kɨt-opɨ-} & \obj{k-okanumɨ-}\\
%\gl{3} & \obj{n-opɨ-} & \obj{n-ekajumɨ-}\\
%	\mybottomrule
%	\end{tabular}
%\end{table}
%
However, there were some other morphological criteria distinguishing \gl{s_a_} from \gl{s_p_} verbs in \PC:
\begin{inlinelist}
\item presence vs absence of the \gl{s_a_} marker \rc{w-}
\item absence vs presence of the second person prefix \rc{ə(j)-} in imperatives
\item presence vs absence of a derivational detransitivizing prefix
\end{inlinelist}.

Many languages show an \gl{s_a_} class marker in deverbalized forms, which can be reconstructed to \PC as \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles} for participles, and in \cref{tab:nmlz} for nominalizations.
%\kaxui does not preserve \rc{w-} in its participles (\obj{tehurkat͡ʃe} \qu{fallen}, not \ungr{tɨwehurkat͡ʃe}), while \apalai only preserves it there, but not in nominalizations (\obj{jepɨtopo} \qu{my bathing place}, not \ungr{joepɨtopo}).

\begin{table}
	\centering
	\caption{Participles of \gl{s_a_} and \gl{s_p_} verbs}
	\label{tab:participles}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
Language & \gl{s_a_} verb & \gl{s_p_} verb & Source(s)\\
	\mymidrule
\kaxui & \obj{t-ehurka-t͡ʃe} \qu{fallen} & \obj{tɨ-jaʔ-so} \qu{burnt} & \textcite[39]{schuring2018kaxuyana} \\
\arara & \obj{t-\emp{o-}ep-te} \qu{come} & \obj{t-oregrum-te} \qu{sad} & \textcite[207, 118]{alves2017arara} \\
\trio & \obj{tɨ-\emp{w-}əturu-e} \qu{talked} & \obj{t-əpəə-se} \qu{tired} & \textcite[333--334]{triomeira1999} \\
\wayana & \obj{tə-\emp{w-}epɨ-he} \qu{bathed} & \obj{t-onopɨ-he} \qu{painted} & \textcite[400]{wayanatavares2005} \\
\kalina & \obj{tu-\emp{w-}oʔka-se} \qu{come out} & \obj{t-okari-se} \qu{told} & \textcite[433, 430]{hoff1968carib} \\
\apalai & \obj{t-\emp{o-}erama-se} \qu{returned} & \obj{t-anorɨ-se} \qu{dried} & \textcite[42, 30]{camargo2002lexico}\\
\panare & \obj{t-\emp{o-}tatɨhpə-se} \qu{wailed} & \obj{tɨ-sɨrɨke-t͡ʃe} \qu{tired} & \textcite[232, 244]{panarepayne2013} \\
	\mybottomrule
	\end{tabular}
\end{table}

\begin{table}
	\centering
	\caption{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs}
	\label{tab:nmlz}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
Language & \gl{s_a_} verb & \gl{s_p_} verb & Source(s)\\
	\mymidrule
\kaxui & \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} & \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} & \textcite[49, 74]{schuring2018kaxuyana}\\
\arara & \obj{\emp{w-}orik-tubo} \qu{dancing place} & \obj{ereŋmɨ-tpo} \qu{killing instrument} & \textcite[97]{alves2017arara} \\
\trio & \obj{ji-\emp{w-}əturu-to} \qu{(for) my talking} & \obj{j-emamina-to} \qu{(for) my playing} & \textcite[246]{triomeira1999} \\
\wayana & \obj{ɨ-\emp{w-}əturu-topo} \qu{my story} & \obj{j-ɨnɨkɨ-topo} \qu{my object for sleeping} & \textcite[409, 130]{wayanatavares2005}\\
\kalina & \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} & \obj{aj-ereʔna-Ø} \qu{your fainting} & \textcite[392, 135]{hoff1968carib}\\
\apalai & \obj{j-epɨ-topo} \qu{my bath place} & \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} & \textcites[90]{koehn1986apalai}[ner2/1]{koehns1994textos}\\
\panare & \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} & \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} & \textcites[390]{panarepayne2013}[23]{mattei1994diccionario}\\
	\mybottomrule
	\end{tabular}
\end{table}

The distinction between \gl{s_a_} and \gl{s_p_} is also borne out in imperatives, the suffix for which can be reconstructed as \PC \rc{-kə}.
Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed.
This is illustrated with reflexes in various modern languages in \cref{tab:imp}.
As in the case of the \gl{s_a_} marker \rc{w-} participles and nominalizations, some languages have lost the distinction between \gl{s_a_} and \gl{s_p_} verbs in imperatives, for example \panare, or \kaxui (not shown in \cref{tab:imp}).

\begin{table}
	\centering
	\caption{Imperatives of \gl{s_a_} and \gl{s_p_} verbs}
	\label{tab:imp}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
Language & \gl{s_a_} & \gl{s_p_} & Source(s)\\
	\mymidrule
\hixka & \obj{omoh-ko} \qu{come!} & \obj{\emp{oj-}okajɨm-ko} \qu{go up!} & \textcite[89, 44]{derbyshire1965textos}\\
\arara & \obj{odotpot-ko} \qu{come back!} & \obj{\emp{o-}alum-ko} \qu{jump!} & \textcite[161]{alves2017arara}\\
\trio & \obj{epɨ-kə} \qu{bathe!} & \obj{\emp{ə-}eremina-kə} \qu{sing!} & \textcite[323]{triomeira1999}\\
\wayana & \obj{əməm-kə} \qu{enter!} & \obj{\emp{əw-}eremi-kə} \qu{sing!} &  \textcite[227]{wayanatavares2005} \\
\kalina & \obj{oʔmaʔ-ko} \qu{stop!} & \obj{\emp{aj-}awon-ko} \qu{get up!} & \textcite[190]{hoff1968carib}\\
\apalai & \obj{otuʔ-ko} \qu{eat!} & \obj{\emp{o-}nɨʔ-ko} \qu{sleep!} & \textcites[62]{koehn1986apalai}[Mopo/20]{koehns1994textos} \\
\panare & \obj{ape-ʔ} \qu{flee!} & \obj{ahpən-kə} \qu{jump!} & \textcite[17, 5]{mattei1994diccionario}\\
	\mybottomrule
	\end{tabular}
\end{table}

%\subsubsection{The detransitivizing prefix}
There is one further property uniting \gl{s_a_} verbs, which is not based on inflectional morphology.
As mentioned in \cref{sec:pc_person}, mismatches between the semantics of intransitive verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common.
However, the Cariban split-\gl{s} system goes further than all other known such systems, in that its division of the verbal lexicon does not follow any discernible semantic criteria whatsoever.
\textcite{meira2000split} takes a sizable corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
He shows that neither (non\-)activities, % \parencite[210]{meira2000split}, 
(non-)agency, % \citeyearpar[212]{meira2000split}, 
(in-)animacy, % \citeyearpar[213]{meira2000split},
nor Aktionsart % \citeyearpar[215]{meira2000split} 
satisfactorily predict the class membership of intransitive verbs.

Rather, the reason for a verb to take the \gl{a}- or \gl{p}-oriented prefix is (at least diachronically) a morphological one.
\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
\begin{quotation}
Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still exisiting transitive sources) or diachronically (with reconstructible but no longer existing transitive sources) \parencite[201]{meira2000split}
\end{quotation}
\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
\begin{inlinelist}
	\item there are a few underived \gl{s_a_} verbs, with no detransitivizing prefix
	\item \gl{s_a_} verbs can develop irregular semantics compared to their transitive counterparts
	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{s_a_} verb
	\item some originally derived \gl{s_a_} verbs have lost their transitive counterparts
	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{s_a_})} from \qu{to dance with (\gl{tr})}
\end{inlinelist}.
He also notes that this leads to an inflectional split not based in meaning, but rather morphology:

\begin{quotation}
Apparently, the morphological behavior of the \gl{s_a_} verb class is an accidental consequence of the fact that detransitivization, as far back as we can reconstruct, entails all the morphology described […] as typical of \gl{s_a_} verbs. The alignment of person-marking prefixes appears not to be driven by any semantic forces in the language; it is as though they were being dragged by the evolution of the reflexive marker. \parencite[226]{meira2000split}
\end{quotation}

As for the form of this reflexive marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although they have since merged into a single morpheme in modern Cariban languages.
Reflexes of \detrz show diverse meanings, which are best subsumed under \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.

\ex<triodetrz> \trio \parencites[218--219]{meira2000split}[128, 256]{triomeira1999}\\
\begin{tabular}[t]{@{}llll@{}}
\\
\obj{nonta}  & → & \obj{e-nonta}, & \qu{abandon each other}\\
\qu{abandon} & & \obj{əi-nonta} &  (reciprocal) \\
\\
\obj{suka} & → & \obj{e-suka}, & \qu{wash self}\\
\qu{wash} & & \obj{əi-suka} & (reflexive)\\
\\
\obj{pahka} & → & \obj{e-pahka} & \qu{break (\gl{intr})}\\
\qu{break (\gl{tr})} & & & (anticausative)\\
\\
\obj{puunəpɨ} & → & \obj{əh-puunəpɨ}, & \qu{think, meditate}\\
\qu{think about} & & \obj{əi-puunəpɨ} & (antipassive)\\
\end{tabular}
\xe

%\pex<triodetrz> Semantics of the \trio detransitivizer \parencite[218--219]{meira2000split}
%\a \obj{e-suka}, \obj{əi-suka} \qu{to wash oneself} (reflexive)
%\a \obj{e-nonta}, \obj{əi-nonta} \qu{to abandon each other} (reciprocal)
%\a \obj{e-pahka} \qu{to break} (anticausative)
%\a \obj{əi-puunəəpɨ} \qu{to think} (antipassive)
%\xe
%
The morphological variation featured in \qu{to wash oneself} and \qu{to abandon each other} is due to the mentioned collapse between the two \PC prefixes:
\obj{e-} is a reflex of the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
However, both can occur with either meaning -- at least for these two verbs.

%\subsubsection{Underived \gl{s_a_} verbs}
As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitives, and do not have a reflex of \detrz, either.
\textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
\textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
These verbs will be discussed in more detail in \cref{sec:verbs}.

\begin{table}
	\centering
	\caption{Underived \PC \gl{s_a_} verbs \parencite[30]{gildea2007greenberg}}
	\label{tab:underived2007}
	\begin{tabular}{@{}lll@{}}
	\mytoprule
Form & Meaning\\
\mymidrule
\rc{tə(mə)} & \qu{to go}\\
\rc{ətepɨ} & \qu{to come\subs{1}}\\
\rc{ka(ti)} & \qu{to say}\\
\rc{əmə(mɨ)} & \qu{to enter}\\
\rc{eti} & \qu{to dwell, be\subs{2}}\\
\rc{a(p)} & \qu{to be\subs{1}, say}\\
\rc{əməkɨ} & \qu{to come\subs{2}}\\
	\mybottomrule
	\end{tabular}
\end{table}

\subsection{Person marker extensions in intransitive verbs}
\label{sec:extensions_intro}
The original \PC split-\gl{s} system has been subject to change in many languages, in some even to the point of total loss.
Besides such modifications as for example losing the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} or losing certain person values altogether, these changes have largely been due to person prefixes being extended to new verbs.\footnote{
One could also imagine a scenario whereby verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
However, the verbs he sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial and explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system includes markers being extended to new verbs.}
There have been many such person marker extensions in Cariban languages, and some are still ongoing.
This is shown by \textcite{gildea1998}, using the Parukotoan languages as an example.
I have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription -- and in the case of \kaxui, the addition of Ø/\obj{j-} as an alternative \gl{1}\gl{s_p_} marker \perscommpar{Spike Gildea}.
Apart from segmental changes to individual morphemes, the following restructuring innovations happened in the \setone paradigm in Parukotoan:

\begin{enumerate}
\item \PPar \begin{enumerate}
	\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
	\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
	\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
\end{enumerate}
\item \PWai \begin{enumerate}
\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
\end{enumerate}
\item \waiwai \begin{enumerate}
\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
\end{enumerate}

\begin{figure}[hbt]
\centering
\setlength{\tabcolsep}{2pt}
\begin{tikzpicture}[
	every node/.append style={align=center},
	level distance=80pt,
]
\Tree[.{\PC{}\pcone}
[.{\PPar{}\ppone}
{\kaxui{}\kaxone}
[.{\PWai{}\pwone}
{\hixka{}\hixone}
{\waiwai{}\waione}
]
]
]
\end{tikzpicture}
  \caption{Person marking extensions in Parukotoan, after \textcite[94]{gildea1998}}
  \label{fig:par_ext}
\end{figure}

\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows variation in the first person and \gl{1+2} prefixes.
\waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the system.
That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
Similarly, the \gl{2}\gl{s_p_} prefix \obj{a-} was extended to imperatives of (former) \gl{s_a_} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.

\pex<wailoss>
\a<wai-158> \waiwai \parencite[][98]{waiwaihawkins1998}\\
\begingl
\gla k-eɸɨrka-t͡ʃhe//
\glb \gl{1+2}-fall-\gl{advz}.after//
\glft \qu{after our fall}//
\endgl
\a<kax-138> \kaxui \parencite[][49]{schuring2018kaxuyana}\\
\begingl
\gla ku-w-ehurka-tpɨrɨ//
\glb \gl{1+2}-\gl{s_a_}-fall-\gl{nmlz}.\gl{pst}.\gl{pert}//
\glft \qu{our fall}//
\endgl
\a<wai-112> \waiwai \parencite[][177]{waiwaihawkins1998}\\
\begingl
\gla a-mo-ko//
\glb \gl{2}-come-\gl{imp}//
\glft \qu{come!}//
\endgl
\a<hix-43> \hixka \parencite[][191]{hixkaryanaderby1985}\\
\begingl
\gla m-omokɨ-no//
\glb \gl{2}\gl{s_a_}-come-\gl{imm}//
\glft \qu{You have come.}//
\endgl
\xe

%Such extensions, often only affecting one person value at a time, and the eventual loss of split-\gl{s} are not entirely unexpected, given that the system lacks any semantic basis.

While different cases of loss of split-\gl{s} are discussed by \textcite[91--96]{gildea1998}, this paper focuses on a so far neglected aspect of these person marking extensions.
I argue that they are executed via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.

First of all, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which is what one would expect from a change in progress.
%todo put in GOOD kaxui data
Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
Finally, the process of lexical diffusion is seen most clearly where it was incomplete.
Not all person marker extensions spread through the entire lexicon, but rather stopped short of some verbs.
This is illustrated for \trio first person forms of \gl{s_a_} verbs in \exref{trio1}; regular \gl{s_a_} verbs take phonologically determined allomorphs \obj{s-} (\envr{}{\obj{e}}) and \obj{t-} (\envr{}{\obj{ə}}), but \obj{tən} \qu{to go} takes an \dbqu{irregular} or \dbqu{archaic} prefix \obj{wɨ-}; it was not affected by the spread of innovative \obj{t-}/\obj{s-} and therefore preserves the old prefix.

\ex<trio1> \trio \gl{1}\gl{s_a_} verbs \parencite[292, 294]{triomeira1999}\\
\begin{tabular}[t]{@{}ll@{}}
\obj{s-epɨ} & \qu{I bathed} \\
\obj{s-entapo} & \qu{I yawned}\\
\obj{t-əturu} & \qu{I talked}\\
\obj{t-əənɨkɨ} & \qu{I slept}\\
\obj{wɨ-tən} & \qu{I went}\\
\end{tabular}
\xe
%
This paper is primarily about these incomplete extensions, and the verbs that were not affected by them.

%
%
%For instance, while most \gl{s_a_} verbs in \hixka and \waiwai have a first person marker \obj{k-}, not all do:
%
%\ex<TAG> \waiwai \qu{to say} \perscommpar{Spike Gildea}\\
%\begin{tabular}[t]{@{}ll@{}}
%\gl{1} & \obj{wɨɨ-ka-}\\
%\gl{2} & \obj{mɨɨ-ka-}\\
%\gl{1+2} & \obj{tɨt-ka-}\\
%\gl{3} & \obj{nɨɨ-ka-}\\
%\end{tabular}
%\xe
%%
%From a synchronic perspective, \waiwai \qu{to say} has an irregular first person marker \obj{wɨ-} instead of regular \obj{kɨ-}.
%From a diachronic perspective, this verb is archaic, preserving the old \gl{1}\gl{s_a_} marker \rc{w-}.

That is, apparently ongoing changes like the situation discussed above for \kaxui will not be investigated in detail.
The same is true for extensions that affected the entire verbal lexicon, although we will briefly comment on them.
We have investigated the 19 person marker extensions we are aware of, but only 6 of them left a group of irregularly or archaically inflected verbs.
Interestingly, these verbs are always \gl{s_a_} verbs, and the irregular marker is always a first person one.
While there have been extensions for other person values as well, they never affect \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affect the entire lexicon, at least based on the attested data.
Besides the completed extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}, examples include the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set in \waimiri \pcref{tab:waimiri}.

\begin{table}[h]
\caption{Some examples for completed extensions \parencite[90--92]{gildea1998}}
\label{tab:completed}
\begin{subtable}[h]{0.24\textwidth}
\centering
\caption{\apalai}
\label{tab:apalai}
\begin{tabular}{@{}lll@{}}
\mytoprule
& \gl{s_a_} & \gl{s_p_}\\
\mymidrule
\gl{1} & Ø/\obj{ɨ-} & \obj{ɨ-}/\obj{j-}\\
\gl{2} & \obj{m(ɨ)-} & \obj{o-}\\
\gl{1+2} & \multicolumn{2}{c}{\emp{\obj{s(ɨ)-}}}\\
\gl{3} & \multicolumn{2}{c}{\obj{n(ɨ)-}}\\
\mybottomrule
\end{tabular}
\end{subtable}
\hfill
\begin{subtable}[h]{0.24\textwidth}
\centering
\caption{\panare}
\label{tab:panare}
\begin{tabular}{@{}lll@{}}
\mytoprule
& \gl{s_a_} & \gl{s_p_}\\
\mymidrule
\gl{1} & \obj{w(ɨ)-} & Ø/\obj{j-}\\
\gl{2} & \multicolumn{2}{c}{\emp{\obj{m(ɨ)-}}}\\
\gl{1+2} & \multicolumn{2}{c}{\obj{n(ɨ)-}}\\
\gl{3} & \multicolumn{2}{c}{\obj{n(ɨ)-}}\\
\mybottomrule
\end{tabular}
\end{subtable}
\hfill
%\begin{subtable}[h]{0.24\textwidth}
%\centering
%\caption{\pemon}
%\label{tab:pemon}
%\begin{tabular}{@{}lll@{}}
%\mytoprule
%& \gl{s_a_} & \gl{s_p_}\\
%\mymidrule
%\gl{1} & \multicolumn{2}{c}{\emp{Ø}}\\
%\gl{2} & \multicolumn{2}{c}{\emp{\obj{m-}}}\\
%\gl{1+2} & \multicolumn{2}{c}{\obj{n-}}\\
%\gl{3} & \multicolumn{2}{c}{\obj{n-}}\\
%\mybottomrule
%\end{tabular}
%\end{subtable}
%\hfill
\begin{subtable}[h]{0.24\textwidth}
\centering
\caption{\waimiri}
\label{tab:waimiri}
\begin{tabular}{@{}ll@{}}
\mytoprule
& \gl{s}\\
\mymidrule
\gl{1} & \emp{\obj{w(ɨ)-}/\obj{i-}}\\
\gl{2} & \emp{\obj{m(ɨ)-}}\\
\gl{1+2} & \emp{\obj{h(ɨ)-}}\\
\gl{3} & \obj{n-}/Ø\\
\mybottomrule
\end{tabular}
\end{subtable}
\end{table}

The remainder of this paper is structured as follows:
\cref{sec:extensions} investigates each of these six incomplete extensions individually, listing unaffected verbs and reconstructing proto-paradigms, where necessary.
\cref{sec:verbs} takes a comparative look at the verbs that were not affected by these extensions and searches for possible motivations.
Finally, \cref{sec:discussion} discusses the findings and puts them in a general context of language change and morphology.

%For instance, while the extension of \PWai \gl{1}\gl{s_p_} \rc{k-} affected most verbs, like \rc{eɸurka} \qu{to fall}, verbs like \rc{ka(ti)} \qu{to say} were not affected \exref{pwaiext}.

%\ex<pwaiext> Incomplete extension of \PWai \rc{k-} to \gl{1}\gl{s_a_}\\
%Sources: \textcites[510]{howard2001wrought}[227, 39, 4, 198, 249]{hixkaryanaderby1985}[175, 174, 43, 55, 60]{waiwaihawkins1998} \\
%\begin{tabular}[t]{@{}llll@{}}
%& \PWai & \hixka & \waiwai \\
%\qu{fall} & \rc{k-eɸurka-} & \obj{k-ehurka-} & \obj{k-eɸɨrka}\\
%\qu{choke self} & \rc{k-oseʃewnuk-} & \obj{k-oseʃemnuk-} & \obj{k-eʃeʃewnuk-}\\
%\qu{treat self medically} & \rc{k-osoht͡ʃema-} & \obj{k-osoht͡ʃema-} & \obj{k-eseht͡ʃem-}\\
%\qu{come} & \rc{k-omokɨ-} & \obj{k-omokɨ-} & \obj{k-mok-}\\
%\qu{say} & \rc{wɨ-ka-} & \obj{ɨ-ka-} & \obj{wɨɨ-ka-}\\
%\qu{be} & \rc{w-eʃi-} & \obj{w-eʃe-} & \obj{w-eeʃi-}\\
%\end{tabular}
%\xe
%

### 2021-09-10 documents/sections/verbs.tex
initial commit

\section{Resistant verbs in a comparative perspective}
\label{sec:verbs}
In \cref{sec:extensions}, six distinct extensions of personal prefixes into \gl{1}\gl{s_a_} territory were introduced, with certain verbs resisting the innovation.
The set of unaffected verbs was rather small in most cases, and there is a considerable etymological overlap between the (proto-)language-specific verb groups.
In this section, I present these verbs from a comparative perspective and discuss their reconstructibility.
Attentive readers will have noticed that certain verbs reoccurred in these lists, not only matching in meaning but also largely in form.
In this section, I will discuss these verbs from a comparative perspective.
\cref{sec:be} treats the copular verb \rc{eti}/\obj{a(p)} \qu{to be}, \cref{sec:say} \rc{ka(ti)} \qu{to say}, \cref{sec:go} \qu{ɨtə(mə)} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstructed as \gl{s_a_} verbs that were not derived from transitive verbs.
\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek; \cref{sec:bathe} investigates \PPek \rc{ɨpɨ} \qu{to bathe}, \cref{sec:shit} \trio \qu{to defecate}, and \cref{sec:movement} the \akuriyo movement verbs.
The \obj{e}-initial verbs not affected by the extension of \rc{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.

%\begin{table}
%	\centering
%	\caption{XXX}
%	\label{tab:XXX}
%	\begin{tabular}{@{}llllllllllllll@{}}
%	\mytoprule
%& \PPeks \rc{k-} & \acs{ara_lg} & \acs{ikp_lg} & \acs{bak_lg} & \PWais \rc{k-} & \acs{wai_lg} & \acs{hix_lg} & \PTirs \rc{t-} & \acs{tri_lg} & \acs{aku_lg} & \acs{aku_lg} \obj{k-} & \acs{car_lg} \obj{j(i)-} & \acs{yuk_lg} \obj{j-} \\
%\rc{ɨtə(mə)} \qu{go} & × & × & \checkmark & × & × & (\checkmark) & × & × & × & × & × & × & × \\
%\rc{ka(ti)} \qu{say} & × & × & × & × & × & × & × & × & × & × & × & ? & ? \\
%\rc{ət-epɨ} \qu{come} & \checkmark & -- & \checkmark & \checkmark & -- & -- & -- & × & × & -- & -- & -- & ? \\
%\rc{jəpɨ} \qu{come} & × & × & -- & -- & -- & -- & -- & × & × & × & × & \checkmark & ? \\
%\rc{ɨptə} \qu{go down} & × & × & ? & \checkmark & -- & -- & -- & × & × & × & × & ? & -- \\
%\PPeks\rc{ipɨ} \qu{bathe} & × & × & × & × & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
%\rc{eti} \qu{be} & × & × & × & × & × & × & × & × & × & × & × & \checkmark & \checkmark \\
%\rc{a(p)} \qu{be} & × & -- & -- & × & × & × & × & × & × & × & × & × & \checkmark \\
%	\mybottomrule
%	\end{tabular}
%\end{table}

\subsection{\rc{eti}/\obj{a(p)} \qu{to be}}
\label{sec:be}
For a comprehensive comparative overview for this verb, I refer the reader to \textcite[375--382]{gildea2018reconstructing}, who reconstructs two distinct roots serving as verbs \qu{to be} in modern Cariban languages.
One is the older copula \rc{a(p)}, which can be reconstructed as already having various irregularities in \PC.
The other is a root \rc{eti} reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam} value.
Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
\akuriyo \obj{eʔi} was not a potential target of innovative \obj{k-}, due to the verb being \obj{e}-initial.
\carijo innovated \obj{j-}, but only in the \obj{et͡ʃi} root allomorph \exref{carcop.mivida-12}; the \obj{a} root preserves \obj{w-} \exref{carcop.car-25}.
\yukpa innovated \obj{j-} for reflexes of both \rc{a(p)} and \rc{eti}, which are preserved as auxiliaries in certain constructions \exref{yukcop}.

\pex<carcop>\carijo
\a<mivida-12>
\begingl
\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
\glft \qu{Then I was already grown up.} \parencite[][177]{robayo1989rame}//
\endgl
\a<car-25>
\begingl
\gla əji-marə-ne w-a-e//
\glb \gl{2}-\gl{com}-\gl{pl} \gl{1}-be-\gl{npst}//
\glft \qu{I am with you all.} \parencite[][42]{guerrero2016karihona}//
\endgl
\xe

\ex<yukcop> cliticized copula forms in \yukpa \parencite[143--144]{meira2006syntactic}\\
\begin{tabular}[t]{@{}lll@{}}
& \gl{npst} & \gl{pst}\\
\gl{1} & \obj{=j-a(-s)}&\obj{=j-e}\\
\gl{2} & \obj{=mak(o)}&\obj{=m-e}\\
\gl{3} & \obj{=mak(o)}&\obj{=n-e}\\
\end{tabular}
\xe

\subsection{\rc{ka(ti)} \qu{to say}}
\label{sec:say}
Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed based on some imperative forms \exref{kati}.

\pex<kati> \PC \rc{kati-kə} \qu{say!}
\a \apalai \obj{kaʃi-ko} \parencite[35]{koehn1986apalai}
\a \wayana \obj{kai-kə} \parencite[181]{wayanatavares2005}
\a \hixka \obj{kas-ko} \parencite[128]{hixkaryanaderby1985}
\a \panare \obj{kah-kə} \parencite[102]{mattei1994diccionario}
\xe

This verb was not affected by the extensions found in \PPek, \PWai, and \PTir.
It was not a potential target of \akuriyo \obj{k-}; I do not have data for its \carijo and \yukpa reflexes.

As briefly mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes this verb as transitive in \hixka.
This analytical choice is not only motivated by avoiding an idiosyncratic intransitive first person prefix \obj{ɨ-}, with the usual prefix being \obj{k-}.
\hixka \obj{ka} also behaves like a transitive verb in other ways, for instance by showing the complementary distribution of the third person marker \obj{n-} and preceding objects -- in this case direct speech or ideophones \exref{hixsay}.

\pex<hixsay>\hixka
\a<hix-119>
\begingl
\glpreamble onɨ wyaro nkekonɨ bɨryekomo, tɨyonɨ wya//
\gla onɨ wjaro n-ka-jakonɨ bɨrʲekomo tɨ-jonɨ wja//
\glb this like \gl{3}-say-\gl{rem}.\gl{cont} boy \gl{cor}-mother \gl{obl}//
\glft \qu{This is what the boy said to his mother.} \parencite[][36]{hixkaryanaderby1985}//
\endgl
\a<ordpl-35>
\begingl
\glpreamble moro ha, ketxkoná hatá.//
\gla moro ha ka-jat͡ʃkonɨ hatɨ//
\glb \gl{med}.\gl{dem}.\gl{inan} \gl{ints} say-\gl{rem}.\gl{cont}.\gl{pl} \gl{hsy}//
\glft \qu{“That one there” they said.} \parencite[][14]{derbyshire1965textos}//
\endgl
\xe
In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}).
In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
This complementary distribution is otherwise only found with transitive verbs.
The verb shows the same pattern, albeit inconsistently, in \trio \parencite[267]{triocarlin2004}.

Further comparative evidence also points to \rc{ka(ti)} \qu{to say} showing transitive traits:
\trio \obj{ka} is characterized as the only intransitive verb being able to take the causative suffix \obj{-po} and the agentive nominalizer \obj{-ne} \parencite[263, 169]{triomeira1999}.
The exceptionality of \obj{ka} \qu{to say} taking \obj{-po} \qu{\gl{caus}.\gl{tr}} has also been noted for \kalina \parencite[82]{courtz2008carib} and \wayana \parencite[258]{wayanatavares2005}.
Reflexes of another transitive causativizer \rc{-metipo} \parencite{gildea2015valency} are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
\panare has innovated a gnomic or nonspecific verbal suffix \obj{-ɲe} from the agent nominalizer \rc{-ne} \parencite[184]{gildea1998}.
Its occurrence on \obj{ka} leads \parencite[214]{panarepayne2013} to analyze the verb as transitive, in contrast to \textcite[102]{mattei1994diccionario}, who categorizes it as intransitive.

The view of \qu{to say} as intransitive is supported primarily by prefix patterns:
\kalina offers a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.\footnote{Interestingly, the \kalina causativized form \obj{kapo} \qu{to make say} does not have the regular \gl{1}>\gl{3} prefix \obj{s(i)-}, but irregular \obj{w(ɨ)-} \parencite[430]{courtz2008carib}.}
Similarly, \qu{to say} in Pekodian languages has a reflex of \gl{1}\gl{s} \rc{w-} \pcref{sec:pekodian}, and not \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
Additionally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{kar-84}, and it is probably reflected in vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.

\ex<kar-84>\kalina \parencite[][202]{courtz2008carib}\\
\begingl
\glpreamble Òmakon \`wa oti ywykàpo kaiko.//
\gla o-ʔma-kon ʔwa oti ɨ-wɨ-ka-ʔpo kai-ko//
\glb \gl{2}-child-\gl{pl} \gl{obl} greeting \gl{1}-\gl{s_a_}-say-\gl{pst}.\gl{nmlz} say-\gl{imp}//
\glft \qu{Pass my greetings to your children.}//
\endgl
\xe
%
Summing up, this verb could be reconstructed as being intransitive based on its prefixes, but transitive based on some suffixes.
\hixka has lost the main intransitive criteria, making its reflex look more like a transitive verb.

\subsection{\rc{ɨtə(mə)} \qu{to go}}
\label{sec:go}
This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə(mə)}, with the second syllable only occurring in some reflexes, in certain contexts.
It is true that many reflexes are clearly \obj{t}-initial, for example \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}.
However, a comparison of selected forms with unambiguously C-initial \rc{ka(ti)} \qu{to say} suggests that an initial vowel \rc{ɨ} should be reconstructed \exref{gosaycomp}.\footnote{Many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.}

\ex<gosaycomp>
\begin{tabular}[t]{@{}lllll@{}}
& go-\gl{imp} & go-\gl{neg} & say-\gl{imp} & say-\gl{neg} \\
\wayana & \obj{ɨtə-kə} & \obj{ɨtə-ra} & \obj{kai-kə} & \obj{ka-ra} \\
\hixka & \obj{ɨto-ko} & \obj{ɨto-hra} & \obj{kas-ko} & \obj{ka-hra} \\
\apalai & \obj{ɨto-ko} & \obj{ɨto-pɨra} & \obj{kaʃi-ko} & \obj{ka-ra} \\
\end{tabular}\\
\parencites[66, 98]{camargo2010wayana}[235, 258]{wayanatavares2005}[47, 54 194]{hixkaryanaderby1985}[65]{derbyshire1965textos}[kuruaz 033, 055]{koehns1994textos}[100]{camargo2002lexico}[107]{koehn1986apalai}
\xe
%

This verb was not affected by the extensions found in \PPek, \PWai, \PTir, \carijo, and \yukpa.
It was not a potential target of \akuriyo \obj{k-}.

\subsection{\rc{(ət-)jəpɨ} \qu{to come}}
\label{sec:come}
\definecolor{come1}{RGB}{161.0,201.0,244.0}
\definecolor{come2}{RGB}{255.0,180.0,130.0}
\definecolor{come3}{RGB}{141.0,229.0,161.0}
\definecolor{come4}{RGB}{255.0,159.0,155.0}
\begin{table}
	\centering
	\caption{Reflexes of \qu{to come}}
    \label{tab:come}
	\begin{tabular}{@{}lll@{}}
	\mytoprule
	Language & Form & Source\\
	\mymidrule
\kaxui & \colorbox{come3}{\obj{ehɨ}} & \perscomm{Spike Gildea}\\
\kaxui & \colorbox{come1}{\obj{johɨ}} & \perscomm{Spike Gildea}\\
\kaxui & \colorbox{come2}{\obj{oohɨ}} & \perscomm{Spike Gildea}\\
\arara & \colorbox{come3}{\obj{ebɨ}} & \textcite[150]{alves2017arara}\\
\arara & \colorbox{come4}{\obj{odebɨ}} & \textcite[150]{alves2017arara}\\
\ikpeng & \colorbox{come4}{\obj{arep}} & \textcite[265]{ikpengpacheco2001}\\
\bakairi & \colorbox{come4}{\obj{əewɨ}} & \textcite[4]{meira2003bakairi}\\
\trio & \colorbox{come4}{\obj{əepɨ}} & \textcite[294]{triomeira1999}\\
\trio & \colorbox{come3}{\obj{epɨ}} & \textcite[294]{triomeira1999}\\
\akuriyo & \colorbox{come3}{\obj{epɨ}} & \textcite[168]{meira1998proto}\\
\carijo & \colorbox{come3}{\obj{ehɨ}} & \textcite[178]{robayo2000avance}\\
\apalai & \colorbox{come4}{\obj{oepɨ}} & \textcite[37]{koehn1986apalai}\\
\kalina & \colorbox{come2}{\obj{opɨ}} & \textcite[429]{courtz2008carib}\\
\maqui & \colorbox{come3}{\obj{ehə}} & \textcite[438]{maquiritaricaceres2011}\\
\akawaio & \colorbox{come4}{\obj{əsipɨ}} & \textcite[160]{stegeman2014akawaio}\\
\akawaio & \colorbox{come1}{\obj{jepɨ}} & \textcite[125]{akawaiocaesar2003}\\
\ingariko & \colorbox{come1}{\obj{jepə}} & \textcite[415]{cruz2005fonologia}\\
\ingariko & \colorbox{come1}{\obj{jə}} & \textcite[299]{cruz2005fonologia}\\
\patamona & \colorbox{come1}{\obj{jepɨ}} & \perscomm{Spike Gildea}\\
\patamona & \colorbox{come1}{\obj{jəpɨ}} & \perscomm{Spike Gildea}\\
\pemon & \colorbox{come1}{\obj{jepɨ}} & \textcite[102]{alvarez2000construcciones}\\
\macushi & \colorbox{come1}{\obj{ipɨ}} & \textcite[32]{macushiabbott1991}\\
\panare & \colorbox{come2}{\obj{əpɨ}} & \textcite[65]{panarepayne2013}\\
\yawarana & \colorbox{come2}{\obj{əpɨ}} & \textcite[68]{mendez1959yawarana}\\
\mapoyo & \colorbox{come3}{\obj{epɨ}} & \textcite[74]{muller1975mapoyo}\\
\uxc & \colorbox{come2}{\obj{ee}} & \textcite[182]{meira2005southern}\\
\mybottomrule
	\end{tabular}
\end{table}
This verb is reconstructed as monomorphemic \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a more complex story.
In particular, the Pemongan languages and one of its allomorphs in \kaxui suggest a root \colorbox{come1}{\rc{jəpɨ}}.
However, the one form in \kaxui only occurs in the Progressive \exref{kaxprog.kax-82}, and \obj{j(-)} may be a reflex of the \settwo third person marker \rc{i-} and the \gl{s_a_} class marker \rc{w-}, as suggested by the occurrence of the latter with other person values \exref{kaxprog.kax-54}.
On the other hand, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
Thus, it seems more likely that the \obj{j} is indeed part of the verb, rather than an outcome of \rc{i-w-}.

\pex<kaxprog>\kaxui \perscomm{Spike Gildea}
\a<kax-82>
\begingl
\gla johɨ-rɨ//
\glb \gl{3}.come-\gl{prog}//
\glft \qu{S/he is coming.}//
\endgl
\a<kax-54>
\begingl
\gla o-w-ohɨ-rɨ//
\glb \gl{2}-\gl{s_a_}-come-\gl{prog}//
\glft \qu{You are coming.}//
\endgl
\a<kax-77>
\begingl
\gla i-nkɨ-rɨ//
\glb \gl{3}-sleep-\gl{prog}//
\glft \qu{S/he is sleeping.}//
\endgl
\a<kax-83>
\begingl
\gla Ø-osone-rɨ//
\glb \gl{3}-dream-\gl{prog}//
\glft \qu{S/he is dreaming.}//
\endgl
\xe

There is further evidence supporting the reconstruction of both \rc{j} and \rc{ə}.
First, there are many forms reflecting \colorbox{come2}{\rc{əpɨ}}, occurring in different branches of the family.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
Second, there are also forms reflecting \colorbox{come3}{\rc{epɨ}}, and umlaut of \rc{ə} to \rc{e} is conditioned by preceding \rc{j} \parencite{meira2010origin}.
That is, a unifying account of these forms starts from a root \rc{jəpɨ}, which is the subject to two major sound changes: \begin{inlinelist}
 \item \rc{j}-loss
 \item \rc{ə}-umlaut	
 \end{inlinelist}.
However, these sound changes appear to have applied irregularly, and not always in the same order.
For example, the \kalina form \obj{opɨ} can only be explained if \rc{j} was lost before the umlaut of \rc{ə} to \rc{e}, which happened elsewhere after \rc{j}.
On the other hand, forms like \maqui \obj{ehə} must be the result of \phonl{\rc{ə}}{\rc{e}}{\rc{j}}, with subsequent loss of \rc{j}.

I have so far ignored the forms which form the basis for \posscite{gildea2007greenberg} \rc{ətepɨ}; these are compatible with my reconstruction \rc{jəpɨ}, with the addition of the detransitivizer \rc{ət(e)-}: \colorbox{come4}{\rc{ətjəpɨ}}.
In fact, the \obj{i} in the \akawaio form \obj{əsipɨ} is better explained by a sequence \rc{jə}, which has the same outcome in the \macushi reflex of \rc{jəpɨ}, \obj{ipɨ}.
While the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
For example, \gl{s_p_} \rc{wənɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
\waiwai \qu{go to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced by its status in \kaxui, \kalina, \panare \exref{pan-128}, \arara, \trio, and \akuriyo.

\ex<pan-128>\panare \parencite[][65]{panarepayne2013}\\
\begingl
\gla ju-w-əəpɨ-n ka=m kanoʔ//
\glb \gl{3}-\gl{s_a_}-come-\gl{nspec} \gl{q}=\gl{2}.\gl{aux} rain//
\glft \qu{Do you think it is gonna rain?}//
\endgl
\xe

While these sound changes and the addition of \rc{ət(e)-} do account for the majority of the forms in \cref{tab:come},\footnote{The only exception is the \apalai form \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}. While \obj{oepɨ} would be a regular outcome of \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-}. The form may be due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}. Alternatively, \apalai \obj{oepɨ} could be from \wayana, which has lost its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
the distribution within the family is rather chaotic.
Not only do very closely related languages show different forms, like \yawarana and \mapoyo, but distinct forms can even be found within the same language, usually conditioned by different prefixes.
This was discussed in \cref{sec:pekodian} for \arara, which has reflexes of \rc{ətjəpɨ} and \rc{jəpɨ} within the same paradigm.
A similar situation is found in \trio, where the \setone paradigm shows a reflex of \rc{ətjəpɨ} for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}

\ex<tricome> \trio \parencite[294]{triomeira1999}\\
\begin{tabular}[t]{@{}ll@{}}
\gl{1} & \obj{w-\colorbox{come4}{əepɨ}} \\
\gl{2} &  \obj{mən-\colorbox{come3}{epɨ}} \\ 
\gl{1+2} &  \obj{ke-\colorbox{come3}{epɨ}} \\
\gl{3} &  \obj{n-\colorbox{come3}{epɨ}} \\
\end{tabular}
\xe
%
Summing up, this verb is highly irregular, both from a synchronic and diachronic perspective.
It seems that reflexes of the detransitivizer \rc{ət(e)-} were optionally added to an \gl{s_a_} verb root \rc{jəpɨ}, which further underwent umlaut and loss of \rc{j}, but in no systematic manner, resulting in the chaotic picture in \cref{tab:come}.

As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \arara reflex of \rc{jəpɨ}, but not on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}.
Both reflexes of \rc{ətjəpɨ} (\trio) and of \rc{jəpɨ} (\akuriyo) resisted the introduction of \rc{t-} in \PTir.
\carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
It is unknown whether there is a \yukpa reflex of this verb, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come}.

\ex<car-18>\carijo \parencite[][102]{guerrero2019carijo}\\
\begingl
\gla əji-wa-e j-eh-ɨ//
\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
\glft \qu{I came looking for you.}//
\endgl
\xe

\subsection{\rc{ɨpɨtə} \qu{to go down}}
\label{sec:godown}
Reflexes of this verb were not affected by the extensions of \rc{t-} in \PTir \pcref{sec:taranoan} and \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
Rather, an irregular first person form \rc{p-ɨptə} can be reconstructed for \PTir.
While it seems to have resisted the extension of \rc{k-} in \PPek, \bakairi subsequently introduced it \pcref{sec:pekodian}.
It was also affected by the extension of \obj{j(i)-} in \carijo \exref{car-32}.

\ex<car-32>\carijo \perscommpar{David Felipe Guerrero}\\
\begingl
\gla irə wat͡ʃinakano tae j-ehɨtə-e//
\glb \gl{inan}.\gl{ana} body.of.water along.bounded \gl{1}-go.down-\gl{npst}//
\glft \qu{…I go down through that guachinacán.}//
\endgl
\xe
%
The situation in \yukpa is unclear, as it is an open question whether \obj{ew(uh)tu} \qu{to go down} \parencite{meira2003primeras} is cognate.

A comparative view \pcref{tab:descend_cog} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
In one language, \wayana, the verb shows traits of both classes, leading me to consider it a \dbqu{mixed} verb.
It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.

\pex<waygodown>\wayana \parencite[][200]{wayanatavares2005}
\a<way-73>
\begingl
\glpreamble ïwïptëë//
\gla ɨ-w-ɨptə-rɨ//
\glb \gl{1}-\gl{s_a_}-go.down-\gl{nmlz}//
\glft \qu{my going down}//
\endgl
\a<way-71>
\begingl
\gla əw-ɨptə-k//
\glb \gl{2}-go.down-\gl{imp}//
\glft \qu{Go down!}//
\endgl
\xe
%
Its causativized form is \obj{ɨptəka} \parencite[255]{wayanatavares2005}, containing a reflex of the transitivizer \rc{-ka}, which was restricted to \gl{s_p_} verbs \parencite{gildea2019overview}.

Comparative evidence suggests that \rc{ɨpɨtə} \qu{to go down} was originally an \gl{s_p_} verb which switched its class to \gl{s_a_} in some languages, incompletely so in \wayana.
The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:descend_cog} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
Both causativized forms contain a reflex of the (rare) transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms \parencite[263]{triomeira1999}.
Thus, it seems that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.

\begin{table}
	\centering
	\caption{Reflexes of \rc{ɨpɨtə} \qu{to go down}}
	\label{tab:descend_cog}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
Language & Verb & Class & Source \\
\mymidrule
\kaxui & \obj{ɨhɨto} & \gl{s_p_} & \perscomm{Spike Gildea} \\
\hixka & \obj{hto} & ? & \textcite[196]{hixkaryanaderby1979}\\
\waiwai & \obj{hto} & -- & \textcite[55]{waiwaihawkins1998}\\
\PPek & \rc{ɨptə} & \gl{s_a_} & \cref{sec:pekodian}\\
\PTir & \rc{ɨhtə}  (\gl{1} \rc{p-})& \gl{s_a_} & \cref{sec:taranoan} \\
\carijo & \obj{ehɨtə} & -- & \textcite[118]{guerrero2019carijo}\\
\wayana & \obj{ɨptə} & \gl{s_a_}/\gl{s_p_} & \textcite[44]{camargo2010wayana}\\
\apalai & \obj{ɨhto} & \gl{s_p_} & \textcite[99]{camargo2002lexico} \\
\kalina & (\obj{onɨʔto}) & \gl{s_a_} & \textcite[263]{courtz2008carib}\\
\maqui & \obj{əʔtə} & \gl{s_p_} & \textcite[450]{maquiritaricaceres2011}\\
\kapon & (\obj{uʔtə}) & -- & \textcite[139]{stegeman2014akawaio}\\
\pemon & (\obj{uʔtə}) & -- & \textcite[139]{alvarez2008clausulas}\\
\macushi & (\obj{autɨ}) & -- & \textcite[34]{macushiabbott1991}\\
\panare & \obj{əhtə} & \gl{s_a_} & \textcite[88]{mattei1994diccionario}\\
\yawarana & \obj{əhtə} & -- & \textcite[68]{mendez1959yawarana}\\
\waimiri & \obj{ɨtɨ} & -- & \textcite[58]{bruno1996dictionary}\\
	\mybottomrule
	\end{tabular}
\end{table}

\subsection{\PPek \rc{ipɨ} \qu{to bathe}}
\label{sec:bathe}
This verb only emerged as resisting an extension in the case of \PPek \pcref{sec:pekodian}.
Verbs for intransitive \qu{to bathe} are usually based on transitive verbs in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in Pemongan, \panare, \kalina, and \maqui\footnote{For \maqui, note that while intransitive \obj{eʔhi} points to \rc{e-kupi}, transitive \obj{ɨhɨ} looks like a reflex of \rc{pɨ}.} \pcref{tab:bathe}.
As I have shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
Thus, while the transitive verbs in Pekodian have perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch.
The reasons for this are unknown; I have not encountered \obj{i-} as a reflex of the detransitivizer in Pekodian, see also \textcite[506]{meira2010origin}.
However, it should be noted that other languages also show unexpected developments in this verb, consider the apparent addition of \obj{ew-} in \hixka or the chaotic mixture of \rc{pɨ} and \rc{kupi} in languages spoken in Venezuela.

\begin{table}
	\centering
	\caption{Intransitive and transitive \qu{to bathe}}
    \label{tab:bathe}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
	Language & \gl{intr} & \gl{tr} & Source\\
	\mymidrule
\kaxui & \obj{eehɨ} & \obj{ɨhɨ} & \perscomm{Spike Gildea}\\
\hixka & \obj{ewehɨ} & \obj{ɨhɨ} & \textcites[198]{hixkaryanaderby1979}\\
\waiwai & \obj{ejepu} & \obj{pɨ} & \textcites[166, 192]{waiwaihawkins1998}\\
\arara & \obj{ibɨ} & \obj{ɨp} & \textcites[150, 162]{alves2017arara}\\
\ikpeng & \obj{ip} & \obj{ɨp} & \textcites[103]{ikpengpacheco1997}[123]{campetela1997analise}\\
\bakairi & \obj{i} & \obj{ɨ} & \textcites[4]{meira2003bakairi}[285]{meira2005bakairi}\\
\trio & \obj{epɨ} & \obj{pɨ} & \textcites[697]{triomeira1999}\\
\akuriyo & \obj{epɨ} & \obj{pɨ} & \textcites[87]{gildea1994akuriyo}\\
\wayana & \obj{epɨ} & \obj{upɨ} & \textcites[24, 52]{camargo2010wayana}\\
\apalai & \obj{epɨ} & \obj{pɨ} & \textcites[218]{meira2000split}\\
\kalina & \obj{ekupi} & \obj{kupi} & \textcites[304]{courtz2008carib}\\
\maqui & \obj{eʔhi} & \obj{ɨhɨ} & \textcites[439, 454]{maquiritaricaceres2011}\\
\kapon & \obj{ekuʔpi} & \obj{kuʔpi} & \textcites[37]{stegeman2014akawaio}\\
\pemon & \obj{ekupɨ} & \obj{pɨ} & \textcites[34, 129]{pemondearmellada1944dic}\\
\panare & \obj{akupɨ} & \obj{ɨpɨ}/\obj{kupɨ} & \textcites[8, 294]{mattei1994diccionario}\\
\mybottomrule
	\end{tabular}
\end{table}

\subsection{\trio \obj{weka}/\obj{oeka} \qu{to defecate}}
\label{sec:shit}
As in the case of \qu{to go down} \pcref{sec:godown}, \rc{weka} \qu{to defecate} was likely originally \gl{s_p_}, with most languages showing an \gl{s_p_} reflex \pcref{tab:shit_cog}.
Another parallel to \qu{to go down} is the mixed \gl{s_a_}/\gl{s_p_} status in \wayana.
The \bakairi reflex \obj{əeke} seems to contain a reflex of \rc{ət-}, which would explain its status as an \gl{s_a_} verb.
While the class memberships of \panare \obj{aiʔka} and \obj{iʔka} \qu{to defecate} are both unknown, the first form seems to contain a reflex of the detransitivizer, but not the second form.

\begin{table}
	\centering
	\caption{\rc{weka} \qu{to defecate} as a class-switching \gl{s_p_} verb}
	\label{tab:shit_cog}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
Language & Verb & Class & Source \\
\mymidrule
\kalina & \obj{uweka} & \gl{s_p_} & \textcites[418]{courtz2008carib} \\
\maqui & \obj{weka} & \gl{s_p_} & \textcites[455]{maquiritaricaceres2011} \\
\kaxui & \obj{weka} & \gl{s_p_} & \perscomm{Spike Gildea} \\
\arara & \obj{watke} & \gl{s_p_} & \textcite[44]{souza1993arara}\\
\ikpeng & \obj{atke} & \gl{s_p_} & \textcite[118]{alves2013verbo}\\
\wayana & \obj{uika} & \gl{s_a_}/\gl{s_p_} & \textcite[86, 206]{wayanatavares2005}\\
\trio & \obj{weka} (\gl{1} \obj{koeka}) & \gl{s_a_}  & \textcites[294]{triomeira1999} \\
\bakairi & \obj{əeke} & \gl{s_a_} & \parencite{meira2005bakairi} \\
\apalai & \obj{weka} & ? & \textcites[96]{camargo2002lexico} \\
\panare & \obj{(a)iʔka} & ? & \textcite[319]{mattei1994diccionario}\\
	\mybottomrule
	\end{tabular}
\end{table}

\subsection{The \akuriyo movement verbs}
\label{sec:movement}
As discussed in \cref{sec:akuriyo}, four \akuriyo \gl{s_a_} verbs are attested as not having the first person marker \obj{t͡ʃ-} in \textcite{gildea1994akuriyo}; instead, they have \obj{w-} or its phonologically reduced form Ø.
These verbs share two properties: they are all \obj{e-}initial, and they are all movement verbs.
However, there are also \obj{e-}initial movement verbs with a first person marker \obj{t͡ʃ-} \exref{aku-106}.
As for their etymology, I can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
For the other three verbs, I have not encountered potential cognates, but given their \obj{e-}initial nature and their semantics, it seems plausible that they, too, are derived from transitive verbs with meanings like \qu{to put up} etc.
In any case, \akuriyo is only sparsely documented, and there are many open questions about it.

\ex<aku-106>\akuriyo \parencite[][84]{gildea1994akuriyo}\\
\begingl
\gla t͡ʃ-ewai//
\glb \gl{1}-sit.down//
\glft \qu{I sat down.}//
\endgl
\xe

### 2021-09-10 documents/talks_abstracts/carib_irreg_SLE.tex
initial commit

\documentclass[9pt]{beamer}
\input{stuff/sle_beamerconfig.tex}
\input{glossingConfigurationBeamer}
\usepackage{datetime2}
\usepackage{phonrule}
\usepackage[edges]{forest}
\input{iswbibConfiguration}
\addbibresource{cariban_references.bib}
\input{cariban_language_names}

\newcommand{\pcone}{\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \rc{t(i)-} & \rc{w-} & \rc{u(j)-} & \rc{u(j)-}\\
\gl{2} & \rc{m(i)-} & \rc{m-} & \rc{ə(j)-} & \rc{ə(j)-}\\
\gl{1+2} & \rc{kɨt(i)-} & \rc{kɨt-} & \rc{k-} & \rc{k-}\\
\end{tabular}}

\newcommand{\ppone}{\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \emp{\rc{w-}} & \rc{w-} & \emp{\rc{k-}} / *Ø/\obj{j-} & *Ø/\obj{j-}\\
\gl{2} & \rc{m-} & \rc{m-} & \rc{o(j)-} & \rc{o(j)-}\\
\gl{1+2} & \rc{kɨt-} & \rc{kɨt-} & \rc{k-}, \emp{\rc{kɨt-}} & \rc{k-}\\
\end{tabular}}

\newcommand{\kaxone}{\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \obj{w-} & \obj{w-} & \obj{k-}, Ø/\obj{j-} & Ø/\obj{j-}\\
\gl{2} & \obj{m-} & \obj{m-} & \obj{o-} & \obj{o-}\\
\gl{1+2} & \obj{kɨt-} & \obj{kɨt-} & \obj{k-} / \obj{kɨt-} & \obj{k-}\\
\end{tabular}}

\newcommand{\pwone}{\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \rc{w-} & \emp{\rc{k-}} & \rc{k-} & \emp{\rc{owɨ(ro) j-}}\\
\gl{2} & \rc{m-} & \rc{m-} & \rc{o(j)-} & \rc{o(j)-}\\
\gl{1+2} & \rc{tɨt-} & \rc{tɨt-} & \rc{tɨt-} & \rc{k-}\\
\end{tabular}}

\newcommand{\hixone}{\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \obj{w-}/\obj{ɨ-} & \obj{k-} & \obj{k-} & \obj{r(o)-}\\
\gl{2} & \obj{m-} & \obj{m-} & \obj{o(j)-} & \obj{o(j)-}\\
\gl{1+2} & \obj{t-} & \obj{t-} & \obj{t-} & \obj{k-}\\
\end{tabular}}

\newcommand{\waione}{\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \obj{w-} & \obj{k-} & \obj{k-} & \obj{o(j)-}\\
\gl{2} & \obj{m-} & \obj{m-} & \emp{\obj{m-}} & \obj{a(w)-}\\
\gl{1+2} & \obj{t(ɨt)-} & \obj{t(ɨt)-} & \obj{t(ɨt)-} & \obj{k-}\\
\end{tabular}}

\newcommand{\setone}{Set I\xspace}
\addbibresource{non_cariban_references.bib}
\title{Irregular first person inflections in Cariban: converging factors for morphological (dis-)similarity}
\institute{SLE 2021 (Online)}
\date{\DTMdate{2021-08-31}}

\begin{document}

\frame{\maketitle}

\section{Irregular first person inflections}

\subsection{Cariban verbal person marking}

\begin{frame}{\PC \setone (main clause) person markers}
	\begin{tabular}{@{}lllll@{}}
	\mytoprule
\gl{a}/\gl{p}&		\gl{1}	&	\gl{2}		&	\gl{1+2}	&	\gl{3}	\\
\mymidrule
\gl{1}	&		&	\rc{k-}	&				&	\rc{t(i)-}		\\	
\gl{2}	&	\rc{k-}			&&				&	\rc{m(i)-}		\\
\gl{1+2}&		&				&				&	\rc{kɨt(i)-}		\\
\gl{3}	&	\rc{u(j)-}	&	\rc{ə(j)-}	&	\rc{k-}			&	\rc{n(i)-}		\\
	\mybottomrule
	\end{tabular}
	\hfill
\begin{tabular}{@{}lll@{}}
\mytoprule
& \gl{s_a_} & \gl{s_p_}  \\
\mymidrule
\gl{1} & \rc{w-} & \rc{u(j)-} \\
\gl{2} & \rc{m-} & \rc{ə(j)-}\\
\gl{1+2} & \rc{kɨt-} & \rc{k-}\\
\gl{3} & \rc{n-} & \rc{n(i)-}\\
\mybottomrule
\end{tabular}	
\end{frame}

\begin{frame}{The split-\gl{s} system}
\begin{itemize}
	\item conserved in many modern languages
	\item \gl{p}-markers identical to \gl{s_p_}
	\item \gl{a}-markers = \gl{s_a_} + \rc{i-} (except first person)
	\item \emp{no semantic basis} (\hyperlink{kalina}{illustration})
\end{itemize}
\end{frame}

\begin{frame}{What conditions the split? \parencite{meira2000split}}
\begin{itemize}
	\item not explicable by (non\-)activities, (non-)agency, (in-)animacy, Aktionsart
	\item purely lexical, semantically arbitrary
	\item but: vast majority of \gl{s_a_} verbs have reflex of detransitivizer \rc{ət(e)-}/\obj{e-}
	\item derivational prefix attaching mostly to transitive verbs, with various semantics (\hyperlink{triodetrz}{illustration})
\end{itemize}
\end{frame}

\subsection{Person marker extensions}
\begin{frame}{Example for person marker extensions: Parukotoan \parencite[94]{gildea1998}}
\centering
\begin{tikzpicture}[
	every node/.append style={align=center},
	level distance=100pt
]
\tikzset{edge from parent/.append style={thick,->}}
\Tree[.\node[draw]{\PC{}\pcone};
\node[draw]{\PPar{}\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \subnode{ppar13}{\emp{\rc{w-}}} & \subnode{ppar1w}{\rc{w-}} & \emp{\rc{k-}} / *Ø/\obj{j-} & *Ø/\obj{j-}\\
\gl{2} & \rc{m-} & \rc{m-} & \rc{o(j)-} & \rc{o(j)-}\\
\gl{1+2} & \rc{kɨt-} & \rc{kɨt-} & \rc{k-}, \emp{\rc{kɨt-}} & \rc{k-}\\
\end{tabular}};
%{\kaxui{}\kaxone}
%[.{\PWai{}\pwone}
%{\hixka{}\hixone}
%{\waiwai{}\waione}
%]
%]
]
\end{tikzpicture}
% \begin{tikzpicture}[overlay,>=latex,shorten >=1pt,->,remember picture]
% \draw[black, ->] (ppar1w) -- (ppar13);
%\end{tikzpicture}
\end{frame}

\begin{frame}{Example for person marker extensions: Parukotoan}
\centering
\begin{tikzpicture}[
	every node/.append style={align=center,scale=0.9},
	level distance=100pt,
]
\tikzset{edge from parent/.append style={thick,->}}
\Tree[.\node[draw]{\PPar\\\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \rc{w-} & \rc{w-} & \rc{k-} / *Ø/\obj{j-} & *Ø/\obj{j-}\\
\gl{2} & \rc{m-} & \rc{m-} & \rc{o(j)-} & \rc{o(j)-}\\
\gl{1+2} & \rc{kɨt-} & \rc{kɨt-} & \rc{k-}, \rc{kɨt-} & \rc{k-}\\
\end{tabular}};%{\PPar{}}
\node[draw]{\kaxui{}\kaxone};
\node[draw]{\PWai{}\pwone};
%{\hixka{}\hixone}
%{\waiwai{}\waione}
%]
%]
]
\end{tikzpicture}
\end{frame}

\begin{frame}{Example for person marker extensions: Parukotoan}
\centering
\begin{tikzpicture}[
	every node/.append style={align=center},
	level distance=100pt,
]
\tikzset{edge from parent/.append style={thick,->}}
\Tree[.\node[draw,label=above:\PWai{}]{\small\begin{tabular}{@{}rllll@{}}
& \gl{a} & \gl{s_a_} & \gl{s_p_} & \gl{p}\\
\gl{1} & \rc{w-} & \rc{k-} & \rc{k-} & \rc{owɨ(ro) j-}\\
\gl{2} & \rc{m-} & \rc{m-} & \rc{o(j)-} & \rc{o(j)-}\\
\gl{1+2} & \rc{tɨt-} & \rc{tɨt-} & \rc{tɨt-} & \rc{k-}\\
\end{tabular}};
\node[draw]{\hixka{}\hixone};
\node[draw]{\waiwai{}\waione};
%]
%]
]
\end{tikzpicture}
\end{frame}

\begin{frame}{Person marker extensions: lexical diffusion}
\begin{itemize}
	\item gradual spread through the verbal lexicon, not all at once: \begin{enumerate}
	\item visible in ongoing changes in \kaxui
	\item place switch of \PC \rc{t-} (\gl{1}>\gl{3}) and \rc{w-} (\gl{1}\gl{s_a_}) in Tiriyoan likely by lexical diffusion \parencite[112]{meira1998proto}
	\item incomplete extensions -- diffusion stops short of some verbs
\end{enumerate}
\end{itemize}
\end{frame}

\subsection{Incomplete extensions}
%\begin{frame}{Incomplete extensions}
%\begin{itemize}
%	\item t \exref{arasa}
%\end{itemize}
%\ex<arasa> \arara \gl{s_a_} verbs\\ \parencite[61, 99, 150, 164, 177]{alves2017arara} \\
%\begin{tabular}[t]{@{}llll@{}}
%&  \qu{to dance} & \qu{to enter} & \qu{to bathe}\\
%\gl{1} & \obj{k-origu-} & \obj{k-omomɨ-} & \obj{\emp{w-}ibɨ-}\\
%\gl{2} & \obj{m-origu-} & \obj{m-omomɨ-} & \obj{m-ibɨ-}\\
%\gl{1+2} & \obj{kud-origu-} & \obj{kut-omomɨ} & \obj{kut͡ʃ-ibɨ-}\\
%\gl{3} & Ø-\obj{origu} & Ø-\obj{omomɨ-} & Ø\obj{-ibɨ-}\\
%\end{tabular}
%\xe
%\end{frame}

\begin{frame}{Incomplete extension: \PPek \rc{k-}}
\label{ppek}
%\begin{itemize}
%	\item likely from \gl{1+2}\gl{s_p_} prefix \rc{k-}
%%	\item so far no reconstructions of \PPek
%	\item newly identified shared innovation of the Pekodian branch
%\end{itemize}
\ex<pekreg>
\begin{tabular}[t]{@{}llll@{}}
 & \arara \qu{to dance}  & \ikpeng \qu{to run} & \bakairi \qu{to go up} \\
& \parencite[150]{alves2017arara} &  \parencite[52]{ikpengpacheco2001}& \parencite[4]{meira2003bakairi} \\
\gl{1}\gl{s_a_}  & \obj{\emp{k-}origu-} & \obj{\emp{k-}aranme-} & \obj{\emp{k-}əku-}\\
\gl{2}\gl{s_a_}  & \obj{m-origu-} & \obj{m-aranme-} & \obj{m-əku-}\\
\gl{1+2}\gl{s_a_}  & \obj{kud-origu-} & \obj{kw-aranme-} & \obj{kɨd-əku-}\\
\gl{3}\gl{s_a_} & \obj{Ø-origu-} & \obj{Ø-aranme-} & \obj{n-əku-}\\
\end{tabular}
\xe
\end{frame}

\begin{frame}{\PPek \rc{k-}: unaffected verbs}
\begin{itemize}
	\item those verbs not affected by specific extension retain old markers -- \dbqu{irregular} or \dbqu{archaic}
\end{itemize}
\begin{center}
\begin{tabular}{@{}lllll@{}}
	\mytoprule
& \PPek & \arara & \ikpeng & \bakairi \\
\mytoprule
\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} & \obj{u-ge-} \\
\qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{Ø-ip-} & \obj{w-i-} \\
\qu{be} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
& \rc{w-ap-} & \obj{w-ap-} & -- & \obj{w-a-}\\
\qu{come} & \rc{w-epɨ} / \rc{k-əd-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
\qu{go down} & \rc{w-[ɨ/i]ptə-} & \obj{w-iptoŋ-} & ?\obj{-iptoŋ-} & \obj{k-ɨtəgɨ-} \\
\qu{go} & \rc{w-ɨtən-} & \obj{w-ɨdo-} & \obj{k-aran-} & \obj{u-tə-} \\
	\mybottomrule
	\end{tabular}\\
\footnotesize \parencites[153, 200]{alves2017arara}[130]{von1892bakairi}[4]{meira2003bakairi}[80, 139, 209]{ikpengpacheco2001}[68; p.c., Angela Fabíola Alves Chagas]{ikpengpacheco1997}
\end{center}
\end{frame}

\begin{frame}{\qu{to come} in Pekodian}
\small
\begin{itemize}
	\item \ikpeng \obj{k-arep-} and \bakairi \obj{k-əewɨ-} point to \rc{k-əd-ebɨ} (root \rc{ebɨ} with detransitivizing prefix)
	\item \arara \obj{w-ebɨ-} points to \rc{w-ebɨ-} (bare root)
	\item \arara also has reflexes of complex \rc{əd-ebɨ} elsewhere in the paradigm \exref{ara-123}
	\item extension of \rc{k-} to morphologically complex form either in \PPek (coexisting variants), or subsequently in daughters (paradigm leveling)
\end{itemize}
\ex<ara-123> \arara \parencite[][150]{alves2017arara}
\begingl
\gla m-odebɨ-nɨ//
\glb \gl{2}\gl{s_a_}-come-\gl{rec}//
\glft \qu{You came.}//
\endgl
\xe

\end{frame}

\begin{frame}{Incomplete person marker extensions in Cariban}
\begin{itemize}
	\item six incomplete extensions (of 19 investigated extensions)
	\item \emp{all innovative \gl{1}\gl{s_a_} markers}
	\item i.e., what remains is a small class of verbs with reflexes of \PC \gl{1}\gl{s_a_} \rc{w-}
	\item other extensions: \begin{itemize}
	\item non-first persons only innovate \gl{s_p_} markers, never \gl{s_a_}
	\item always affect all verbs (unless still ongoing)\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Incomplete extensions of new \gl{1}\gl{s_a_} markers}
\begin{enumerate}
	\item \PPek \rc{k-} (\hyperlink{sec:ppek}{details})
	\item \PWai \rc{k-} (\hyperlink{sec:pwai}{details})
	\item \PTir \rc{t-} (\hyperlink{sec:ptir}{details})
	\item \akuriyo \obj{k-} (\hyperlink{sec:aku}{details})
	\item \carijo \obj{j-} (\hyperlink{sec:car}{details})
	\item \yukpa \obj{j-} (\hyperlink{sec:yuk}{details})
\end{enumerate}
\end{frame}

\begin{frame}{Distribution in the family}
\begin{figure}
\centering
\tiny
\begin{forest}
for tree={
s sep = -1ex,
grow=east,
forked edges,
node options={align=center },
anchor=west,edge path={\noexpand\path[draw, gray, line width=1pt, \forestoption{edge}](!u.parent anchor) |- (.child anchor)\forestoption{edge label};},
before typesetting nodes={if content={}{shape=coordinate}{}},}
    [Cariban, tier=4
        [
            [{\uxc}, tier=0]
        ]
        [
            [{\waimiri}, tier=0]
        ]
        [
            [{\apalai}, tier=0]
        ]
        [
            [{\wayana}, tier=0]
        ]
        [
            [{\kalina}, tier=0]
        ]
        [Yukpan,tier=1
            [{\emp{\yukpa}}, tier=0]
            [{\japreria}, tier=0]
        ]
        [Taranoan, tier=2
            [\emp{Tiriyoan}
                [{\trio}, tier=0]
                [{\emp{\akuriyo}}, tier=0]
            ]
            [{\emp{\carijo}}, tier=0]
        ]
        [Venezuelan, tier=3
            [{\maqui/\dekwana}, tier=0]
            [Core Venezuelan, tier=2
            	[%\pemon-\panare
            	[{\panare}, tier=0]
                [Pemongan, tier=1
                    [{\ingariko}, tier=0]
                    [{\macushi}, tier=0]
                    [{\pemon}, tier=0]
                ]
            	]
            	[
            	  [{\kumana}, tier=0]
                [{\tamanaku}, tier=0]
                [
                [{\mapoyo}, tier=0]
                [{\yawarana}, tier=0]
                ]
            	]
            ]
        ]
        [\emp{Pekodian}, tier=2
            [Xinguan, tier=1
                [{\arara}, tier=0]
                [{\ikpeng}, tier=0]
            ]
            [{\bakairi}, tier=0]
        ]
        [Parukotoan, tier=2
            [\emp{Waiwaian}, tier=1
                [{\hixka}, tier=0]
                [{\waiwai}, tier=0]
            ]
            [{\kaxui}, tier=0]
        ]
    ]
\end{forest}
\caption{The Cariban language family}
\label{fig:caribantree}
\end{figure}
\end{frame}

\begin{frame}{Geographical distribution}
\begin{center}
\includegraphics[height=0.9\textheight]{stuff/cariban_phylo_map}
\end{center}
\end{frame}

%\begin{frame}{Example: regular \gl{s_a_} in Pekodian}
%\footnotesize
%\begin{tabular}[t]{@{}llll@{}}
%\mytoprule
%& \bakairi \qu{to go up}  & \arara \qu{to dance}  & \ikpeng \qu{to run} \\
%& \parencite[4]{meira2003bakairi} & \parencite[150]{alves2017arara} &  \parencite[52]{ikpengpacheco2001}\\
%\mymidrule
%\gl{1}\gl{s_a_} & \obj{\emp{k-}əku-} & \obj{\emp{k-}origu-} & \obj{\emp{k-}aranme-} \\
%\gl{2}\gl{s_a_} & \obj{m-əku-} & \obj{m-origu-} & \obj{m-aranme-} \\
%\gl{1+2}\gl{s_a_} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
%\gl{3}\gl{s_a_} & \obj{n-əku-} & \obj{Ø-origu-} & \obj{Ø-aranme-} \\
%\mybottomrule
%\end{tabular}
%\end{frame}
%
%\begin{frame}{Example: irregular \gl{1}\gl{s_a_} in Pekodian}
%\footnotesize
%	\begin{tabular}{@{}lllll@{}}
%	\mytoprule
%& \PPek & \arara & \ikpeng & \bakairi \\
%\mytoprule
%\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} & \obj{u-ge-} \\
%\qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{Ø-ip-} & \obj{w-i-} \\
%\qu{be} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
%\qu{come} & \rc{w-epɨ}, \rc{k-əd-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
%\qu{go} & \rc{w-ɨtən-} & \obj{w-ɨdo-} & \obj{k-aran-} & \obj{u-tə-} \\
%\qu{go down} & \rc{w-[ɨ/i]ptə-} & \obj{w-iptoŋ-} & ?\obj{-iptoŋ-} & \obj{k-ɨtəgɨ-} \\
%	\mybottomrule
%	\end{tabular}
%\end{frame}

\begin{frame}{What verbs are unaffected by extensions?}
\small
\begin{center}
	\begin{tabular}{@{}lllllll@{}}
	\mytoprule
PC & \PPeks \rc{k-} & \PWais \rc{k-} & \PTirs \rc{t-} & \acs{aku_lg} \obj{k-} & \acs{car_lg} \obj{j-} & \acs{yuk_lg} \obj{j-} \\
\mymidrule
\rc{ɨtə(mɨ)} \qu{go} & × & × & × & × & × & × \\
\rc{ka(ti)} \qu{say} & × & × & × & × & ? & ? \\
\rc{(ət-)jəpɨ} \qu{come} & \checkmark/× & -- & × & × & \checkmark & -- \\
%\rc{ət-epɨ} \qu{come} & \checkmark & -- & × & -- & -- & -- \\
%\rc{jəpɨ} \qu{come} & × & -- & × & × & \checkmark & -- \\
\rc{ɨpɨtə} \qu{go down} & × & -- & × & × & \checkmark & -- \\
\rc{eti} \qu{be} & × & × & × & × & \checkmark & \checkmark \\
\rc{a(p)} \qu{be} & × & × & × & × & × & \checkmark \\
\rc{e-pɨ} \qu{bathe} & × & \checkmark & \checkmark & × & ? & -- \\
	\mybottomrule
	\end{tabular}\\
\emp{\checkmark} innovative marker \quad \emp{×} archaic marker\quad \emp{--} unattested \quad\emp{?} unknown
\end{center}
\begin{itemize}
	\item strong overlap among languages in the verbs retaining \gl{1}\gl{s_a_} \rc{w-}
	\item not shown: 1) many \obj{e}-initial \akuriyo verbs, 2) 4 \akuriyo movement verbs, 3) \trio \qu{to defecate}
\end{itemize}
\end{frame}

\subsection{Potential causes}
\begin{frame}{What causes verbs to adapt or resist innovative markers?}
%\small
\begin{itemize}
	\item possible explanations based on the network model of morphology by \textcites{bybee1985morphology}{bybee1995regular}:
\begin{enumerate}
	\item morphology-based lexical connections → same inflection class
	\item phonology-based lexical connections → same inflection class
	\item frequency: high-frequency lexemes are resistant to innovation
	\item (semantics play no role in this case)
\end{enumerate}
	\item strong overlap of all three factors in the Cariban case:\begin{enumerate}
	\item vast majority of \gl{s_a_} verbs have reflex of detransitivizer \rc{ət(e)-}/\obj{e-}
	\item therefore, they are \obj{e}- or \obj{ə}/\obj{o}-initial
	\item other, underived \gl{s_a_} verbs often high-frequency
	\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}{Morphology: \PPek}
\begin{itemize}
	\item none of the unaffected verbs have detransitivizer
	\item normally detransitivized \rc{e-pɨ} \qu{to bathe} has become \rc{ibɨ} in \PPek (\hyperlink{tobathe}{details})\begin{itemize}
	\item[→] loss of transparent morphology, not affected by innovative marker \end{itemize}
	\item \bakairi and \ikpeng show \rc{k-} with \rc{əd-ebɨ} \qu{to come}, but \arara has \rc{w-} with root \rc{ebɨ} (\hyperlink{tocome}{details})\begin{itemize}
	\item[→] presence of detransitivizer clearly triggers innovative prefix\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[allowframebreaks]{Phonology: \akuriyo, \carijo, \yukpa}
\begin{itemize}
	\item \akuriyo \obj{k-} only on \obj{ə}-initial verbs, not \obj{e}-initial (\hyperlink{aku}{details})
	\item \carijo: all \obj{e}- and \obj{ə}-initial verbs affected, even those without detransitivizer \exref{care}
	\item \yukpa: all V-initial verbs, only \obj{to} \qu{to go} unaffected (\hyperlink{yukpaj}{details}) -- but very patchy data!
\end{itemize}
\footnotesize
\pex<care>\carijo
\a<car-18>
\begingl
\gla əji-wa-e \emp{j-e}h-ɨ//
\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
\glft \qu{I came looking for you.} \parencite[][102]{guerrero2019carijo}//
\endgl
\a<car-32>
\begingl
\gla irə wat͡ʃinakano tae \emp{j-e}hɨtə-e//
\glb \gl{inan}.\gl{ana} body.of.water along.bounded \gl{1}-go.down-\gl{npst}//
\glft \qu{…I go down through that guachinacán.} (p.c., David Felipe Guerrero)//
\endgl
\a<mivida-12>
\begingl
\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
\gla ireti-bə et͡ʃi-nə=me gərə \emp{j-e}t͡ʃi-ɨ//
\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
\glft \qu{Then I was already grown up.} \parencite[][177]{robayo1989rame}//
\endgl
\xe
\end{frame}

\begin{frame}{Converging factors}
\begin{itemize}
	\item most resistant verbs: \rc{ka(ti)} \qu{to say} and \rc{ɨtə(mɨ)} \qu{to go}\begin{enumerate}
	\item no detransitivizer
	\item C-initial (in these languages)
	\item very frequent (although no text counts)
	\end{enumerate}
\item often no definite answer for resistance to innovation
\item \PC: underived \gl{s_a_} verbs also most frequent, other irregularities
\item ultimate origins of split-\gl{s} system still mysterious
\end{itemize}
\end{frame}

\renewcommand*{\bibfont}{\tiny}
\begin{frame}[allowframebreaks]{References}
\printbibliography
\end{frame}

\section{appx: Individual extensions}

\subsection{\PPek \rc{k-}}
\label{sec:ppek}
\begin{frame}{Irregular \arara verbs}
\ex<arairr> \arara \parencite[153]{alves2017arara} \\
\begin{tabular}[t]{@{}ll@{}}
\obj{wɨ-ge-nɨ} & \qu{I said}\\
\obj{w-it͡ʃi-nɨ} & \qu{I was, lied down}\\
\obj{w-ebɨ-nɨ} & \qu{I came}\\
\obj{w-ibɨ-nɨ} & \qu{I bathed}\\
\obj{w-iptoŋ-rɨ} & \qu{I went down}\\
\obj{w-ɨdo-lɨ} & \qu{I went}\\
\end{tabular}
\xe
\end{frame}

\begin{frame}{Cognate verbs in \bakairi: \obj{u-}/\obj{w-}, two \obj{k-}}
\footnotesize
\pex[everyglpreamble=,interpartskip=-1ex]<bakverbs> \bakairi \parencites[][131, 397, 76, 137, 374]{von1892bakairi}[4]{meira2003bakairi}
\begin{multicols}{2}
\a<bak-3>
\begingl
\glpreamble \ort{u-ɣépa} //
\gla \emp{u-}ge-pa//
\glb \gl{1}\gl{s_a_}-say-\gl{neg}//
\glft \qu{I don't say.}//
\endgl
\a<bak-5>
\begingl
\glpreamble \ort{wi-táki} / \ort{wi-tági} //
\gla \emp{w-}i-taki//
\glb \gl{1}\gl{s_a_}-be-\gl{int}//
\glft \qu{I was.}//
\endgl
\a<bak-4>
\begingl
\glpreamble \ort{kχaewí-le} //
\gla \emp{k-}əewɨ-lɨ//
\glb \gl{1}\gl{s_a_}-come-\gl{imm}//
\glft \qu{I came.}//
\endgl
\a<bak-6>
\begingl
\glpreamble \ort{kχ-itaké-he} //
\gla \emp{k-}ɨtəgɨ-se//
\glb \gl{1}\gl{s_a_}-go.down-\gl{npst}?//
\glft \qu{I go down.}//
\endgl
\a<bak-2>
\begingl
\glpreamble \ort{ū́ta} / \ort{uúta}//
\gla \emp{u-}tə//
\glb \gl{1}\gl{s_a_}-go//
\glft \qu{I go.}//
\endgl
\a
\begingl
\gla \emp{w-}i-də//
\glb \gl{1}\gl{s_a_}-bathe-\gl{imm}//
\glft \qu{I bathed}//
\endgl
\end{multicols}
\xe
\end{frame}

\begin{frame}[allowframebreaks]{Cognate verbs in \ikpeng: \obj{ɨ-}/Ø-, two \obj{k-}}
\small
\pex<ikpw>
\a<ikp-168>
\begingl
\gla \emp{ɨ-}ge-lɨ//
\glb \gl{1}-say-\gl{rec}//
\glft \qu{I said.} \parencite[][209]{ikpengpacheco2001}//
\endgl
\a<ikp-169>
\begingl
\gla \emp{Ø-}et͡ʃi-lɨ//
\glb \gl{1}-be-\gl{rec}//
\glft \qu{I was.} \parencite[][139]{ikpengpacheco2001}//
\endgl
\a<ikp-167>
\begingl
\gla at͡ʃagotpop \emp{Ø-}ip-t͡ʃi ik-gwa-kt͡ʃi//
\glb always \gl{1}-bathe-\gl{npst} river-\gl{loc}.aquatic-\gl{all}//
\glft \qu{I always bathe in this river.} \parencite[][68]{ikpengpacheco1997}//
\endgl
\xe

\pex<ikpk> Subsequent innovations:
\a<ikp-170>
\begingl
\gla \emp{k-}aran-t͡ʃi//
\glb \gl{1}-go-\gl{npst}//
\glft \qu{I'm going.} \parencite[][80]{ikpengpacheco2001}//
\endgl
\a<ikp-171>
\begingl
\gla \emp{k-}arep-lɨ//
\glb \gl{1}-come-\gl{rec}//
\glft \qu{I came.} \parencite[][80]{ikpengpacheco2001}//
\endgl
\xe
Unattested first person form: \obj{iptoŋ} \qu{to sit down} (p.c., Angela Fabíola Alves Chagas)
\end{frame}

\begin{frame}{\PPek \rc{k-}: unaffected verbs}
\begin{center}
\hyperlink{ppek}{Overview table}	
\end{center}
\end{frame}

\subsection{\PWai \rc{k-}}
\label{sec:pwai}
\begin{frame}[allowframebreaks]{\PWai \rc{k-}}
\begin{itemize}
	\item first extension of \gl{1+2}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_p_} in \PPar
	\item then further extension from \gl{1}\gl{s_p_} to \gl{1}\gl{s_a_}, creating unified first person category:
\end{itemize}
\small
\ex<wais> \PWai \\
\begin{tabular}[t]{@{}lll@{}}
\gl{1} & \rc{\emp{k-}eɸurka-} & \rc{kɨ-wɨnɨkɨ-}\\
\gl{2} & \rc{m-eɸurka-} & \rc{o-wɨnɨkɨ-}\\
\gl{1+2} & \rc{t-eɸurka-} & \rc{tɨt-wɨnɨkɨ-}\\
\gl{3} & \rc{ɲ-eɸurka-} & \rc{nɨ-wɨnɨkɨ-}\\
& \qu{to fall} (\gl{s_a_}) & \qu{to sleep} (\gl{s_p_})\\
\end{tabular}
\xe
\pex Modern reflexes\\
\a \hixka \parencites[510]{howard2001wrought}[189--191]{hixkaryanaderby1985} \\ \begin{tabular}[t]{@{}lll@{}}
	\gl{1} & \obj{k-ehurka-} & \obj{kɨ-nɨkɨ-}\\
\gl{2} & \obj{m-ehurka-} & \obj{o-wnɨkɨ-}\\
\gl{1+2} & \obj{t-ehurka-} & \obj{tɨ-nɨkɨ-}\\
\gl{3} & \obj{ɲ-ehurka-} & \obj{nɨ-nɨkɨ-}\\
\end{tabular}
\a \waiwai \parencites[510]{howard2001wrought}[209--211]{hawkins1953waiwai}[50]{waiwaihawkins1998} \\ \begin{tabular}[t]{@{}lll@{}}
\gl{1} & \obj{k-eɸɨrka-} & \obj{kɨ-wɨnk-}\\
\gl{2} & \obj{m-eɸɨrka-} & \obj{mɨ-wɨnk-}\\
\gl{1+2} & \obj{t͡ʃ-eɸɨrka-} & \obj{tɨt-wɨnk-}\\
\gl{3} & \obj{ɲ-eɸɨrka-} & \obj{nɨ-wɨnk-}\\
\end{tabular}
\xe
\end{frame}

\begin{frame}{\PWai \rc{k-}: unaffected verbs}
\ex<parpar> Archaic \gl{1}\gl{s_a_} forms \\ (\cite[124, 197--198, 209]{hixkaryanaderby1985} and Spike Gildea, p.c.)\\
\begin{tabular}[t]{@{}llll@{}}
 & \PPar & \hixka & \waiwai \\
\qu{to say} & \rc{\emp{wɨ-}ka-} & \obj{\emp{ɨ-}to-} & \obj{\emp{wɨɨ-}ka}\\
\qu{to be} & \rc{\emp{w-}eʃi-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eʃe-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eeʃi-}/\obj{\emp{w-}a-}\\
\qu{to go} & \rc{\emp{wɨ-}to-} & \obj{\emp{ɨ-}to-} & \obj{kɨ\emp{w-}to-}\\
\end{tabular}\\
\xe
\end{frame}

\subsection{\PTir \rc{t-}}
\label{sec:ptir}
\begin{frame}[allowframebreaks]{\PTir \rc{t-}}
\begin{itemize}
	\item in \trio and \akuriyo, \PC \rc{t-} (\gl{1}>\gl{3}) and \rc{w-} (\gl{1}\gl{s_a_}) switched places \parencite[107--112]{meira1998proto}
	\item \envr{\rc{t-}}{\rc{ə}}, \envr{\rc{t͡ʃ-}}{\rc{e}}
	\item maybe in \PTar, maybe in \PTir{} -- \carijo removed any evidence by the \hyperlink{sec:car}{extension of \obj{j(i)-}}
\end{itemize}
\end{frame}
\begin{frame}[allowframebreaks]{\PTir \rc{t-}: unaffected verbs}
\small
\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-}\\
\begin{tabular}[t]{@{}llll@{}}
& \PTir & \trio & \akuriyo\\
\qu{go} & \rc{wɨ-tə(mɨ)-} & \obj{wɨ-tə(mɨ)-} & \obj{ə-təmɨ-}/\obj{wɨ-təemɨ-}  \\
\qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
\qu{come} & \rc{w-əʔepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
\qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
& \rc{w-eʔi-} & \obj{w-ei-} & --\\
\end{tabular}\\
\parencites[112--115]{meira1998proto}[85]{gildea1994akuriyo}[339]{triomeira1999}
\xe
\pex<akumov>
\a<aku> \akuriyo \gl{1}\gl{s_a_} \rc{w-} in 4 movement verbs\\ \parencite[84--86]{gildea1994akuriyo}\\
\begin{tabular}[t]{@{}ll@{}}
\qu{return} & Ø\obj{-erama-}\\
\qu{get up} & Ø\obj{-eokahtə-}\\
\qu{jump} & \obj{w-ejahka-}\\
\qu{go out} & \obj{w-ekɨrɨka-}\\
\end{tabular}
\a<tri> \trio \obj{s-erama-} \parencite[301]{triomeira1999}
\xe
\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes\\
\begin{tabular}[t]{@{}lll@{}}
& \trio & \akuriyo\\
\qu{go down} & \obj{p-ɨhtə-} & \obj{p-ɨtə-}\\
\qu{defecate} & \obj{k-oeka-} & ?\\
\end{tabular}\\
\parencites[294]{triomeira1999}[84]{gildea1994akuriyo}
\xe
\end{frame}

\subsection{\akuriyo \obj{k-}}
\label{sec:aku}
\begin{frame}[allowframebreaks]{\akuriyo \obj{k-}}
\label{aku}
\begin{itemize}
	\item after split of Tiriyoan, \akuriyo innovated \gl{1}\gl{s_a_} \obj{k-}
	\item apparently only replaced \rc{t-} (\envr{}{\rc{ə}}):
\end{itemize}
\begin{center}
\begin{tabular}{@{}lll@{}}
	\mytoprule
first person \obj{k-} & first person \obj{t͡ʃ-} \\
\mymidrule
\obj{ənɨkɨ} \qu{to sleep} & \obj{eepɨ} \qu{to bathe} \\
\obj{əməmɨ} \qu{to enter} & \obj{ewai} \qu{to sit down} \\
\obj{əturu} \qu{to talk} & \obj{etonema} \qu{to lie down} \\
\obj{əət͡ʃena} \qu{to cry} & \obj{ekɨɨrɨka} \qu{to stay back} \\
\obj{ətajiŋka} \qu{to run} & \obj{entapo} \qu{to yawn} \\
\obj{əiwa} \qu{to tremble} &  \\
\obj{əempa} \qu{to learn} & \parencite{gildea1994akuriyo}	\\
	\mybottomrule
	\end{tabular}
\end{center}
\begin{itemize}
	\item \textcite[107]{meira1998proto} largely agrees, mentioning \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs)
	\item no data provided; either leftovers not affected by extension of \obj{k-}, or later influence by \trio
\end{itemize}
\end{frame}

\subsection{\carijo \obj{j-}}
\label{sec:aku}
\begin{frame}{\carijo \obj{j-}}
\begin{itemize}
	\item \carijo extended \gl{1}\gl{s_p_} \obj{j(i)-} to \gl{s_a_} verbs
	\item (alongside \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨs-} to \gl{s_p_})
\end{itemize}
\ex<carpar> \carijo \parencites[106]{meira1998proto}[173]{robayo2000avance}\\
\begin{tabular}[t]{@{}lll@{}}
& \obj{tuda} \qu{to arrive} & \obj{eharaga} \qu{to dance}\\
\gl{1} & \obj{ji-tuda-} & \obj{\emp{j-}eharaga-}\\
\gl{2} & \obj{mɨ-tuda-} & \obj{m-eharaga-}\\
\gl{1+2} & \obj{kɨsi-tuda-} & \obj{kɨs-eharaga-}\\
\gl{3} & \obj{ni-tuda-} & \obj{n-eharaga-}\\
\end{tabular}
\xe
\end{frame}
\begin{frame}{\carijo \obj{j-}: unaffected verbs}
\begin{itemize}
	\item attested are only \obj{tə} \qu{to go} and \obj{a} \qu{to be}
\end{itemize}
\pex<car1w>\carijo \parencite[][5, 42]{guerrero2016karihona}
\a<car-24>
\begingl
\gla \emp{wɨ-}tə-e=rehe//
\glb \gl{1}-go-\gl{npst}=\gl{frust}//
\glft \qu{I almost go (but I am not going to go).}//
\endgl
\a<car-25>
\begingl
\gla əji-marə-ne \emp{w-}a-e//
\glb \gl{2}-with-\gl{pl} \gl{1}-be-\gl{npst}//
\glft \qu{I am with you all.}//
\endgl
\xe
\end{frame}

\subsection{\yukpa \obj{j-}}
\label{sec:yuk}
\begin{frame}[allowframebreaks]{\yukpa \obj{j-}}
\label{yukpaj}
\begin{itemize}
	\item same innovation as \carijo{}: \gl{1}\gl{s_p_} \obj{j-} to \gl{1}\gl{s_a_} \exref{yukpaintr}
	\item not affected: only \obj{to} \qu{to go} attested \exref{yuk-7}
	\item \PC \rc{w-} is reflected as Ø \exref{yukpatr}
\end{itemize}
\ex<yukpaintr> \yukpa \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}\\
\begin{tabular}[t]{@{}llll@{}}
& \obj{otum} \qu{to wash self} & \obj{nɨ} \qu{to sleep} & \obj{ata} \qu{to fall}\\
\gl{1} & \obj{\emp{j-}otum-} & \obj{jɨ-nɨ-} & \obj{j-ata-}\\
\gl{2} & \obj{m-otum-} & \obj{mɨ-nɨ-} & \obj{m-ata-}\\
\gl{3} & \obj{n-otum-} & \obj{nɨ-nɨ-} & \obj{n-ata-}\\
\end{tabular}
\xe
\ex<yuk-7>\yukpa \parencite[][139]{meira2006syntactic}
\begingl
\gla aw \emp{Ø-}to//
\glb \gl{1}\gl{pro} \gl{1}\gl{s_a_}-go//
\glft \qu{I went.}//
\endgl
\xe
\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw \emp{j-}esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw \emp{Ø-}esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
\end{frame}

\section{appx: Resistant verbs}

\subsection{\rc{ɨtə(mɨ)}}
\begin{frame}{\rc{ɨtə(mɨ)} \qu{to go}}
\label{togo}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{tabular}[t]{@{}ll@{}}
\toprule
 Language &          First person form \\
\midrule
  \waiwai &              \obj{kɨw-to-} \\
   \hixka &                \obj{ɨ-to-} \\
   \arara &               \obj{w-ɨdo-} \\
  \ikpeng &              \obj{k-aran-} \\
 \bakairi &                 \obj{u-tə} \\
 \akuriyo &   \obj{wɨ-təmɨ-/wɨ-təemɨ-} \\
    \trio &              \obj{wɨ-tən-} \\
  \carijo &             \obj{wɨ-təmə-} \\
   \yukpa &  \obj{{\normalfont Ø}-to-} \\
\bottomrule
\end{tabular}
\tiny
\parencites[85]{waiwaihawkins1998}[70]{hixkaryanaderby1985}[153]{alves2017arara}[80]{ikpengpacheco2001}[374]{von1892bakairi}[112]{meira1998proto}[292]{triomeira1999}[139]{meira2006syntactic}\end{column}
\begin{column}{0.5\textwidth}
	\begin{itemize}
		\item has become C-initial in many languages
		\item combination of innovative \rc{k-} and archaic \rc{w-} in \waiwai
		\item innovative \obj{k-} in \ikpeng due to phonological form / reanalysis as detransitivizer?
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\subsection{\rc{ka(ti)}}
\begin{frame}{\rc{ka(ti)} \qu{to say}}
\label{tosay}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{tabular}[t]{@{}ll@{}}
\toprule
 Language & First person form \\
\midrule
  \waiwai &      \obj{wɨ-ka-} \\
   \hixka &       \obj{ɨ-ka-} \\
   \arara &      \obj{wɨ-ge-} \\
  \ikpeng &       \obj{ɨ-ge-} \\
 \bakairi &       \obj{u-ge-} \\
 \akuriyo &      \obj{wɨ-ka-} \\
    \trio &      \obj{wɨ-ka-} \\
  \carijo &                 ? \\
   \yukpa &                 ? \\
\bottomrule
\end{tabular}\\
\tiny
\parencites[71]{waiwaihawkins1998}[4]{hixkaryanaderby1979}[150]{alves2017arara}[209]{ikpengpacheco2001}[131]{von1892bakairi}[113]{meira1998proto}[294]{triomeira1999}
\end{column}
\begin{column}{0.5\textwidth}
	\begin{itemize}
		\item only C-initial \gl{s_a_} verb in \PC
		\item some properties of transitive verbs (causativizer \rc{-po}, nominalizer \rc{-ne})
		\item has become (more) transitive in \hixka (interaction with \gl{3}\gl{p} \obj{n-})
		\item no innovative person markers attested
%		\item \carijo and \yukpa both have \obj{ka}, but first person forms unknown
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\subsection{\rc{ət-epɨ} / \rc{jəpɨ}}
\begin{frame}{\rc{ət-epɨ} / \rc{jəpɨ} \qu{to come}}
\label{tocome}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{tabular}[t]{@{}ll@{}}
\toprule
 Language &            First person form \\
\midrule
  \waiwai &                            – \\
   \hixka &                            – \\
   \arara &                 \obj{w-ebɨ-} \\
  \ikpeng &                \obj{k-arep-} \\
 \bakairi &                \obj{k-əewɨ-} \\
 \akuriyo &  \obj{{\normalfont Ø}-eepɨ-} \\
    \trio &                \obj{w-əepɨ-} \\
  \carijo &                 \obj{j-ehɨ-} \\
   \yukpa &                            – \\
\bottomrule
\end{tabular}\\
\tiny
\parencites[153]{alves2017arara}[42]{ikpengpacheco2001}[76]{von1892bakairi}[114]{meira1998proto}[294]{triomeira1999}[102]{guerrero2019carijo}
\end{column}
\begin{column}{0.5\textwidth}
\small
	\begin{itemize}
		\item reflexes of \rc{jəpɨ}, \rc{epɨ}, \rc{ət-epɨ}, likely with \rc{jəpɨ} > \rc{epɨ}
		\item unexpected detransitivizer, since root is already intransitive
		\item reflexes of detransitivizer are inconsistent even paradigm-internally (\trio or \arara)
		\item Pekodian shows \rc{k-} with complex \rc{əd-ebɨ}, but \rc{w-} with root \rc{ebɨ}
		\item innovative marker in \carijo, due to initial \obj{e}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\subsection{\rc{ɨpɨtə}}
\begin{frame}{\rc{ɨpɨtə} \qu{to go down}}
\label{togodown}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{tabular}[t]{@{}ll@{}}
\toprule
 Language & First person form \\
\midrule
  \waiwai &                 – \\
   \hixka &                 – \\
   \arara &    \obj{w-iptoŋ-} \\
  \ikpeng &                 ? \\
 \bakairi &    \obj{k-ɨtəgɨ-} \\
 \akuriyo &      \obj{p-ɨtə-} \\
    \trio &     \obj{p-ɨhtə-} \\
  \carijo &    \obj{j-ehɨtə-} \\
   \yukpa &                 – \\
\bottomrule
\end{tabular}\\
\tiny
\parencites[153]{alves2017arara}[137]{von1892bakairi}[84]{gildea1994akuriyo}[116; p.c., David Felipe Guerrero]{meira1998proto}
\end{column}
\begin{column}{0.5\textwidth}
	\begin{itemize}
		\item originally an \gl{s_p_} verb (\kaxui, \apalai, + morphological evidence)
		\item switched class (\PPek, \PTir, \panare, partially in \wayana)
		\item irregular \PTir marker \rc{p-}
		\item innovative in \carijo (phonology) and \bakairi (why?)
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\subsection{\rc{a(p)}}
\begin{frame}{\rc{a(p)} \qu{to be}}
\label{tobe1}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{tabular}[t]{@{}ll@{}}
\toprule
 Language &         First person form \\
\midrule
  \waiwai &                \obj{w-a-} \\
   \hixka &               \obj{w-ah-} \\
   \arara &               \obj{w-ap-} \\
  \ikpeng &                         – \\
 \bakairi &                \obj{w-a-} \\
 \akuriyo &  \obj{{\normalfont Ø}-a-} \\
    \trio &                \obj{w-a-} \\
  \carijo &                \obj{w-a-} \\
   \yukpa &                \obj{j-a-} \\
\bottomrule
\end{tabular}\\
\tiny
\parencites[197]{hixkaryanaderby1985}[200]{alves2017arara}[130]{von1892bakairi}[115]{meira1998proto}[142; p.c., Spike Gildea]{meira2006syntactic}
\end{column}
\begin{column}{0.5\textwidth}
	\begin{itemize}
		\item older copula \parencite[375--382]{gildea2018reconstructing}
		\item root co-occurs with \hyperlink{tobe2}{\rc{eti}}
		\item \rc{p} only present in some reflexes
		\item innovative person prefix only in \yukpa
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\subsection{\rc{eti}}
\begin{frame}{\rc{eti} \qu{to be}}
\label{tobe2}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{tabular}[t]{@{}ll@{}}
\toprule
 Language &             First person form \\
\midrule
  \waiwai &                 \obj{w-eeʃi-} \\
   \hixka &                  \obj{w-eʃe-} \\
   \arara &                \obj{w-it͡ʃi-} \\
  \ikpeng &  \obj{{\normalfont Ø}-et͡ʃi-} \\
 \bakairi &                    \obj{w-i-} \\
 \akuriyo &                             – \\
    \trio &                    \obj{w-ei} \\
  \carijo &                \obj{j-et͡ʃi-} \\
   \yukpa &                    \obj{j-e-} \\
\bottomrule
\end{tabular}\\
\tiny
\parencites[198]{hixkaryanaderby1985}[153]{alves2017arara}[139]{ikpengpacheco2001}[397]{von1892bakairi}[339]{triomeira1999}[92]{guerrero2019carijo}[143; p.c., Spike Gildea]{meira2006syntactic}
\end{column}
\begin{column}{0.5\textwidth}
	\begin{itemize}
		\item newer copula root \parencite[375--382]{gildea2018reconstructing}
		\item root co-occurs with \hyperlink{tobe1}{\rc{a(p)}}
		\item original meaning maybe \qu{to dwell}
		\item innovative person prefixes in \carijo and \yukpa (both phonological)
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

\subsection{\rc{e-pɨ}}
\begin{frame}{\rc{e-pɨ} \qu{to bathe}}
\label{tobathe}
\begin{itemize}
	\item fully regular \gl{s_a_} verb in most languages, with reflex of detransitivizer \rc{e-}
	\item \rc{e-pɨ} has become \rc{ibɨ} in \PPek, for unknown reasons
\end{itemize}
\begin{center}
\scriptsize
	\begin{tabular}{@{}llll@{}}
	\mytoprule
	Language & \gl{intr} & \gl{tr} & Source\\
	\mymidrule
\kaxui & \obj{eehɨ} & \obj{ɨhɨ} & p.c., Spike Gildea\\
\hixka & \obj{ewehɨ} & \obj{ɨhɨ} & \textcites[198]{hixkaryanaderby1979}\\
\waiwai & \obj{ejepu} & \obj{pɨ} & \textcites[166, 192]{waiwaihawkins1998}\\
\arara & \obj{ibɨ} & \obj{ɨp} & \textcites[150, 162]{alves2017arara}\\
\ikpeng & \obj{ip} & \obj{ɨp} & \textcites[103]{ikpengpacheco1997}[123]{campetela1997analise}\\
\bakairi & \obj{i} & \obj{ɨ} & \textcites[4]{meira2003bakairi}[285]{meira2005bakairi}\\
\trio & \obj{epɨ} & \obj{pɨ} & \textcites[697]{triomeira1999}\\
\akuriyo & \obj{epɨ} & \obj{pɨ} & \textcites[87]{gildea1994akuriyo}\\
\wayana & \obj{epɨ} & \obj{upɨ} & \textcites[24, 52]{camargo2010wayana}\\
\apalai & \obj{epɨ} & \obj{pɨ} & \textcites[218]{meira2000split}\\
\kalina & \obj{ekupi} & \obj{kupi} & \textcites[304]{courtz2008carib}\\
\maqui & \obj{eʔhi} & \obj{ɨhɨ} & \textcites[439, 454]{maquiritaricaceres2011}\\
\kapon & \obj{ekuʔpi} & \obj{kuʔpi} & \textcites[37]{stegeman2014akawaio}\\
\pemon & \obj{ekupɨ} & \obj{pɨ} & \textcites[34, 129]{pemondearmellada1944dic}\\
\panare & \obj{akupɨ} & \obj{ɨpɨ}/\obj{kupɨ} & \textcites[8, 294]{mattei1994diccionario}\\
\mybottomrule
	\end{tabular}
\end{center}
\end{frame}

\section{appx: split-\gl{s}}

\begin{frame}[allowframebreaks]{Semantic mismatches: \kalina}
\label{kalina}
\pex<kar1>
\a<kar-60>
\begingl
\gla mi-kupi-ja//
\glb \gl{2}>\gl{3}-bathe-\gl{prs}//
\glft \qu{You bathe him/her.} \parencite[][160]{hoff1968carib}//
\endgl
\a<kar-62>
\begingl
\gla a-kupi-ja//
\glb \gl{3}>\gl{2}-bathe-\gl{prs}//
\glft \qu{S/he bathes you.} \parencite[][63]{yamada2011evidentiality}//
\endgl
\xe
\pex<kar2>
\a<kar-59>
\begingl
\gla sipi tɨnka-rɨ m-ekema-non hen//
\glb net pull-\gl{nmlz} \gl{2}-be.afraid-\gl{prs}.\gl{uncert} eh?//
\glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
\endgl
\a<kar-61>
\begingl
\gla aj-awoi-ja//
\glb \gl{2}-get.up-\gl{prs}//
\glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
\endgl
\xe
\begin{itemize}
	\item \gl{s_a_} \obj{ekema} \qu{to be afraid}
	\item \gl{s_p_} \obj{awomɨ} \qu{to get up}
\end{itemize}
\end{frame}

\begin{frame}{Generic meaning of the detransitivizer}
\label{triodetrz}
\small
\ex<TAG> \trio \parencites[218--219]{meira2000split}[128, 256]{triomeira1999}
\begin{tabular}[t]{@{}llll@{}}
\\
\obj{nonta}  & → & \obj{e-nonta}, & \qu{abandon each other}\\
\qu{abandon} & & \obj{əi-nonta} &  (reciprocal) \\
\\
\obj{suka} & → & \obj{e-suka}, & \qu{wash self}\\
\qu{wash} & & \obj{əi-suka} & (reflexive)\\
\\
\obj{pahka} & → & \obj{e-pahka} & \qu{break (\gl{intr})}\\
\qu{break (\gl{tr})} & & & (anticausative)\\
\\
\obj{puunəpɨ} & → & \obj{əh-puunəpɨ}, & \qu{think, meditate}\\
\qu{think about} & & \obj{əi-puunəpɨ} & (antipassive)\\
\end{tabular}
\xe
\end{frame}

\end{document}

### 2021-09-10 documents/talks_abstracts/cariban_underived_abstract_submission.tex
initial commit

\input{stuff/config.tex}
\newcommand{\detrz}{\rc{əte}\slash{}\obj{e-}}
\usepackage{phonrule}
\usepackage{amsthm,etoolbox,xcolor}

\title{Irregular first person markers in Cariban\\}

\author{Florian Matter \\ University of Bern}
%\date{}

\begin{document}
\pagestyle{empty}
\maketitle

\begin{abstract}
\PC is reconstructed as having an verbal inflectional paradigm called \dbqu{\setone} \parencite{gildea1998}.
Most modern languages have reflexes of this system, which shows hierarchically conditioned access to index slots in transitive verbs, and two lexically conditioned inflection classes, \gl{s_a_} and \gl{s_p_}, in intransitive verbs.
This split-\gl{s} division is not based on any known semantic or pragmatic criteria, but rather, the overwhelming majority of \gl{s_a_} verbs are synchronically or diachronically derived from transitive verbs by means of a detransitivizing prefix \parencite{meira2000split}, which had the forms \detrz{} \parencite{meira2010origin}.
In many languages, person markers in this system have been subject to innovations, often by the extension of person markers from somewhere in paradigm to a new location.
In intransitive verbs, this often lead to the erosion or even loss of the split-\gl{s} system.
These person marker extensions happened by way of lexical diffusion, spreading within the verbal lexicon.

Six of these extensions did not fully run their course, but left a small group of verbs with old markers (i.e., irregularly inflected verbs).
This is illustrated in \cref{tab:hixka} for \hixka (Parukotoan), where the regular first person prefix is \obj{k(ɨ)-}, but the verb \obj{to} \qu{to go} shows an irregular first person marker \obj{ɨ-}, since it was not affected by the extension of innovative \obj{k(ɨ)-}.
%
\begin{table}[b]
\caption{\hixka \gl{s_a_} verbs \parencites{hixkaryanaderby1985}{howard2001wrought} }
\label{tab:hixka}
\centering
\begin{tabular}{@{}lllll@{}}
\mytoprule
{} &  \qu{to be standing} &     \qu{to fall} &  \qu{to be afraid} &    \qu{to go} \\
\mymidrule
\gl{1}   &  \obj{k-osowaʃehto-} &  \obj{k-ehurka-} &  \obj{k-oserʲehɨ-} &   \obj{ɨ-to-} \\
\gl{2}   &  \obj{m-osowaʃehto-} &  \obj{m-ehurka-} &  \obj{m-oserʲehɨ-} &  \obj{mɨ-to-} \\
\gl{1+2} &  \obj{t-osowaʃehto-} &  \obj{t-ehurka-} &  \obj{t-oserʲehɨ-} &  \obj{tɨ-to-} \\
\gl{3}   &  \obj{n-osowaʃehto-} &  \obj{ɲ-ehurka-} &  \obj{n-oserʲehɨ-} &   \obj{n-to-} \\
\mybottomrule
\end{tabular}
\end{table}
%
Interestingly, these incomplete extensions, three of which can be reconstructed to proto-languages, all introduced new first person \gl{s_a_} markers.
In addition, the \gl{s_a_} verbs which resisted the extensions show a considerable etymological overlap across the different cases of extension.

I reconstruct these verbs, which often show other irregularities, and investigate potential reasons for why these particular verbs were not affected by different extensions.
Using \posscite{bybee1985morphology} network model of morphology, I show that some cases of incomplete extension can be argued to be due to lexical connections based on morphology (presence vs absence of the detransitivizing prefix), while others appear to be due to phonological connections (\obj{e}- and/or \obj{ə}-initial vs others).
While semantic connections appear to play no role, many of the resistant verbs are high frequency verbs, a factor predicting conservativism in Bybee’s model.
In fact, most cases of resistant verbs are predicted by all three factors: morphological connections, phonological connections, and frequency.
Thus, while the network model offers attractive explanations for the (non-)spread of innovative \gl{1}\gl{s_a_} markers, these predicting factors strongly overlap in many of the investigated cases of extension.
This in turn is largely due to the fact that the most high-frequency \gl{s_a_} verbs were exceptional already at the level of \PC, since they did not contain the detransitivizing prefix \detrz.

\printbibliography

\end{abstract}

\end{document}

### 2021-09-10 documents/talks_abstracts/extensions_SLE_abstract.tex
initial commit

\documentclass[a4paper]{article}
\usepackage[top=2cm]{geometry}
\usepackage{fontspec}
\setmainfont{Linux Libertine}
\input{iswbibConfiguration.tex}
\addbibresource{cariban_references.bib}
\addbibresource{non_cariban_references.bib}
\usepackage{titlesec}
\input{booktabs_config}
\titleformat*{\section}{\normalsize\bfseries}
\usepackage{glossingtool}
\usepackage{graphicx}
\usepackage{subcaption}
\input{cariban_language_names.tex}
\usepackage[inline]{enumitem}
\newlist{inlinelist}{enumerate*}{1}
\setlist[inlinelist]{label={\alph*)}, itemjoin={{; }}, itemjoin*={{; and }}}
\newcommand{\emp}[1]{\textbf{#1}}
\usepackage{hyperref}
\usepackage[capitalise]{cleveref}
\providecommand{\goodtilde}{\char`~\xspace}

\begin{document}

\begin{center}
\bfseries
\large{Irregular first person inflections in Cariban}\\
\vspace{.05cm}
\normalsize{converging factors for morphological (dis-)similarity}\\
\vspace{.1cm}
\normalfont
Florian Matter\\(University of Bern	)
\end{center}
Keywords: Cariban, Morphology, Irregular inflection, Morphological change, Network model
\vspace{.4cm}

\noindent \PC is reconstructed as having an untypical split-\gl{s} system, where the division between \gl{s_a_} and \gl{s_p_} is not based on any known semantic criteria \parencite{meira2000split}.
Rather, verbs derived via the detransitivizer \rc{ət(e)-}/\obj{e-} \parencite{meira2010origin} end up in the \gl{s_a_} category, while almost all other intransitive verbs are \gl{s_p_} \parencite{meira2000split}.
The category of verbs is primarily identified via the person marking prefixes they take (\cref{tab:intr}).

\begin{table}[h]
	\centering
	\caption{\PC split-\gl{s} person marking}
	\label{tab:intr}
\begin{tabular}{@{}lll@{}}
\mytoprule
& \gl{s_a_} & \gl{s_p_}  \\
\mymidrule
\gl{1} & \rc{w-} & \rc{uj-} \\
\gl{2} & \rc{m-} & \rc{əj-}\\
\gl{1+2} & \rc{kɨt-} & \rc{k-}\\
\gl{3} & \rc{n-} & \rc{ni-}\\
\mybottomrule
\end{tabular}
\end{table}

Many Cariban languages have modified this system, introducing new person markers \parencite[80--84, 90--92]{gildea1998}.
These innovations most likely happened via lexical diffusion, as suggested by \textcite{meira2000proto} for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} in \trio and \akuriyo, and as evidenced by the \obj{kɨt͡ʃ-} \goodtilde \obj{k-} \qu{\gl{1+2}\gl{s_p_}} variation attested in \kaxui (Spike Gildea, p.c.).
Of interest are new prefixes which did not spread completely, leaving over a small group of verbs with synchronically irregular inflection patterns, remnants of old \gl{s_a_} markers.
This is illustrated in (\getref{hix}--\getref{yuk}), where (a--b) show verbs with regular marking, and (c) shows verbs with irregular marking -- reflexes of \PC \rc{w-} \qu{\gl{1}\gl{s_a_}}.

%\ex<arairr> \arara first person forms \parencite[153]{alves2017arara} \\
%\begin{tabular}[t]{@{}llll@{}}
%\obj{k-omomɨlɨ} & \qu{enter} & \obj{wɨ-genɨ} & \qu{say}\\
%\obj{k-onkulɨ} & \qu{ascend} & \obj{w-it͡ʃinɨ} & \qu{be}\\
%\obj{k-origulɨ} & \qu{dance} & \obj{w-ebɨnɨ} & \qu{come}\\
%\obj{k-ot͡ʃimtabriŋelɨ} & \qu{eat} & \obj{w-ibɨnɨ} & \qu{bathe}\\
%\obj{k-odakpilɨ} & \qu{be drunk} & \obj{w-iptoŋrɨ} & \qu{descend}\\
%\obj{k-odubulɨ} & \qu{remain} & \obj{w-ɨdolɨ} & \qu{go}\\
%\end{tabular}
%\xe

\begin{multicols}{2}
\pex<hix> \hixka \parencite[188, 209]{hixkaryanaderby1985}
\a \obj{k-ratano} \qu{I wept}
\a \obj{kɨ-kɨtano} \qu{I rushed}
\a \obj{\emp{ɨ-}tono} \qu{I went}
\xe

\pex<ara> \arara \parencite[153]{alves2017arara}
\a \obj{k-omomɨlɨ} \qu{I entered}
\a \obj{k-onkulɨ} \qu{I ascended}
\a \obj{\emp{w-}ebɨnɨ} \qu{I came}
\xe

\pex \trio \parencite[293--294]{triomeira1999}
\a \obj{t-əturu} \qu{I talked}
\a \obj{t-əənɨkɨ} \qu{I slept}
\a \obj{\emp{w-}əepɨ} \qu{I came}
\xe

\pex<yuk> \yukpa \parencites{meira2003primeras}
\a \obj{j-otɨrɨ} \qu{I stayed}
\a \obj{jɨ-nke} \qu{I slept}
\a \emp{Ø-}\obj{to} \qu{I went}
\xe

\end{multicols}

\pex \carijo \parencites[79]{koch1908hiana}[70]{guerrero2016karihona}[5]{guerrero2016carijo}
\a \obj{j-ehɨhəhjai} \qu{I fight}
\a \obj{j-ejae} \qu{I come}
\a \obj{\emp{wɨ-}təe} \qu{I go}
\xe

Interestingly, of 18 investigated person marker extensions affecting intransitive verbs, the 6 extensions leaving such irregular verbs all introduced \gl{1}\gl{s}\textsubscript{(A)} markers.
Three of these innovative markers are reconstructible to intermediate proto-languages: \PPar \rc{k-}, \PWai \rc{k-}, and \PTir \rc{t-}.
%In the first two cases, some daughter languages further extended the new markers to a previously conservative verb.
The other three are found in single extant languages: \akuriyo \obj{k-}, \carijo \obj{j-}, and \yukpa \obj{j(ɨ)-}.
Besides some language-specific irregular verbs, the six innovations show considerable overlap in what verbs they did not affect \exref{irrverb}.

\pex<irrverb> \PC verbs with multiple irregularly inflected reflexes
\a \rc{a(p)}/\obj{eti} \qu{to be}
\a \rc{ka(ti)} \qu{to say}
\a \rc{ɨtə(mɨ)} \qu{to go}
\a \rc{(ət-)jəpɨ} \qu{to come}
\a \rc{ɨpɨtə} \qu{to go down}
\xe

Using \posscite{bybee1985morphology} network model of morphology, I show that some cases of incomplete extension can be argued to be due to lexical connections based on morphology (presence vs absence of the detransitivizing prefix), while the distribution of others reveals phonological connections between verbs with innovative markers (\obj{e}- and/or \obj{ə}-initial vs others).
While no semantic connections have emerged as relevant, many of the resistant verbs are high frequency verbs, a factor predicting conservativism in Bybee’s model.
In fact, most cases of resistant verbs are predicted simultaneously by three factors: morphological connections, phonological connections, and frequency.
Thus, while the network model offers attractive explanations for the (non-)spread of innovative \gl{1}\gl{s_a_} markers, these predicting factors strongly overlap in many of the investigated cases of extension.
This in turn is largely due to the fact that the most high-frequency \gl{s_a_} verbs were different from normal \gl{s_a_} verbs already at the level of \PC, since they did not contain the detransitivizing prefix \rc{ət(e)-}/\obj{e-}, and were therefore morphologically and phonologically distinct from normal \gl{s_a_} verbs.
%\renewcommand{\bibfont}{\footnotesize}
\printbibliography
\end{document}

### 2021-09-10 workflow/create_intr_tables.py

@@ -1,0 +1,160 @@
+import pandas as pd
+from cariban_helpers import *
+import seaborn as sns
+
+def print_source(string):
+    entries = string.split("; ")
+    output = ""
+    for entry in entries:
+        bibkey = entry.split("[")[0]
+        pages = entry.split("[")[1].replace("]","")
+        if bibkey == "pc":
+            return "\\perscomm{%s}" % pages
+        else:
+            output += "[%s]{%s}" % (pages, bibkey)
+    return "\\textcites%s" % output
+
+def shorthandify(lg_id):
+    return "\\" + get_shorthand(lg_id)
+
+def objectify(string, obj_string="obj"):
+    if string in ["", "?", " ", "  "]:
+        return string
+    if string[0] == " ":
+        return string
+    out_list = []
+    morph_list = string.split("; ")
+    for i, morph in enumerate(morph_list):
+        out_list.append(r"\%s{%s}" % (obj_string, morph))
+        if i > 0: obj_string = "obj"
+    return "/".join(out_list)
+    
+a = pd.read_csv("come_roots.csv")
+# print(a)
+ 
+a["Language_ID"] = pd.Categorical(a['Language_ID'], lg_order().keys())
+a.sort_values("Language_ID", inplace=True)
+
+a["Source"] = a["Source"].apply(print_source)
+a["Form"] = a["Form"].apply(lambda x: "\obj{%s}" % x)
+a["Language"] = a["Language_ID"].apply(shorthandify)
+
+cogsets = {
+    1: "jəpɨ",
+    2: "əpɨ",
+    3: "epɨ",
+    4: "ətjəpɨ",
+}
+
+def get_palette(n):
+    palette = sns.color_palette("pastel", n)
+    for i, color in enumerate(palette):
+        palette[i] = [str(x*255) for x in color]
+    return dict(zip(range(1,len(palette)+1), palette))
+
+colors = get_palette(len(cogsets))
+
+cmd_out = ""
+for i, j in colors.items():
+    cmd_out += "\\definecolor{come%s}{RGB}{%s}\n" % (i, ",".join(j))
+# print(cmd_out)
+
+output = ""
+df = a
+for index, row in df.iterrows():
+    row_colors = "\\colorbox{come%s}" % row["Cognateset_ID"]
[…]

### 2021-09-10 workflow/irregular_paradigms.py

@@ -1,0 +1,79 @@
+import pyradigms as pyd
+import pandas as pd
+import cldf_helpers
+import cariban_helpers
+
+normalfonties = ["Ø", "?", "V", "G", "C"]
+def objectify(string, obj_string="obj"):
+    if string in ["", "?", " ", "  ", "–"]:
+        return string
+    if string[0] == " ":
+        return string
+    output = []
+    entry_list = string.split(" OR ")
+    for entry in entry_list:
+        out_list = []
+        morph_list = entry.split("; ")
+        for i, morph in enumerate(morph_list):
+            out_list.append(r"\%s{%s}" % (obj_string, morph))
+            if i > 0: obj_string = "obj"
+        # morph_list = [r"\%s{%s}" % (obj_string, morph) for morph in morph_list]
+        output.append("/".join(out_list))
+    for i, cell in enumerate(output):
+        if r"\env" in cell:
+            output[i] = cell.replace(r"\obj", "")
+    output = ", ".join(output)
+    for normalfonty in normalfonties:
+        output = output.replace(normalfonty, r"{\normalfont %s}" % normalfonty)
+    return output
+
+def get_sources(df, parens=True):
+    pyd.content_string = "Source"
+    table = pyd.compose_paradigm(df)
+    ref_list = [j for i in table.values.tolist() for j in i]
+    source_string = cldf_helpers.cite_a_bunch(ref_list, parens=parens)
+    pyd.content_string = "Form"
+    return source_string
+
+def print_latex(df, ex=False):
+    lines = df.to_latex(escape=False, index=False).split("\n")
+    lines[0] = lines[0].replace("tabular}{l", "tabular}[t]{@{}l").replace("l}", "l@{}}")
+    if ex:
+        del lines[1:4]
+        del lines[-1]
+        del lines[-2]
+        return "\n".join(lines)
+    else:
+        return "\n".join(lines).replace(u"\toprule", u"\mytoprule")
+        
+df = pd.read_csv("irregular_verb_data.csv")
+
+lg_list = ["PWai", "wai", "hix", "PPek", "PXin", "ara", "ikp", "bak", "PTir", "aku", "tri", "car", "yuk"]
+for i in ["PWai", "PPek", "PXin", "PTir"]:
+    lg_list.remove(i)
+    
+sh_dict = {x : "\\" + cariban_helpers.get_shorthand(x) for x in lg_list}
[…]

### 2021-09-10 documents/floats/nominalizations.tex
added split-s illustration data, more examples

\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &                                         \gl{s_a_} &                                             \gl{s_p_} \\
\midrule
\kaxui  &       \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} &            \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} \\
\arara  &        \obj{\emp{w-}orik-tubo} \qu{dancing place} &              \obj{ereŋmɨ-tpo} \qu{killing instrument} \\
\trio   &   \obj{ji-\emp{w-}əturu-to} \qu{(for) my talking} &              \obj{j-emamina-to} \qu{(for) my playing} \\
\wayana &          \obj{ɨ-\emp{w-}əturu-topo} \qu{my story} &        \obj{j-ɨnɨkɨ-topo} \qu{my object for sleeping} \\
\apalai &            \obj{j-epɨ-topo} \qu{my bathing place} &   \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} \\
\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &                  \obj{aj-ereʔna-∅} \qu{your fainting} \\
\panare &     \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} &  \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} \\
\bottomrule
\end{tabular}

### 2021-09-10 documents/floats/participles.tex
added split-s illustration data, more examples

\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &                               \gl{s_a_} &                        \gl{s_p_} \\
\midrule
\kaxui  &         \obj{t-ehurka-t͡ʃe} \qu{fallen} &       \obj{tɨ-jaʔ-so} \qu{burnt} \\
\arara  &         \obj{t-\emp{o-}ep-te} \qu{come} &      \obj{t-oregrum-te} \qu{sad} \\
\trio   &    \obj{tɨ-\emp{w-}əturu-e} \qu{talked} &       \obj{t-əpəə-se} \qu{tired} \\
\wayana &     \obj{tə-\emp{w-}epɨ-he} \qu{bathed} &    \obj{t-onopɨ-he} \qu{painted} \\
\apalai &  \obj{t-\emp{o-}erama-se} \qu{returned} &      \obj{t-anorɨ-se} \qu{dried} \\
\kalina &  \obj{tu-\emp{w-}oʔka-se} \qu{come out} &       \obj{t-okari-se} \qu{told} \\
\panare &  \obj{t-\emp{o-}tatɨhpə-se} \qu{wailed} &  \obj{tɨ-sɨrɨke-t͡ʃe} \qu{tired} \\
\bottomrule
\end{tabular}

### 2021-09-10 documents/sections/proto_carib.tex
added split-s illustration data, more examples

@@
 	\centering
-	\caption{Participles of \gl{s_a_} and \gl{s_p_} verbs}
+	\caption{Participles of \gl{s_a_} and \gl{s_p_} verbs \parencites[118, 207]{alves2017arara}[30, 42]{camargo2002lexico}[430, 433]{hoff1968carib}[333, 334]{triomeira1999}[400]{wayanatavares2005}[39]{schuring2018kaxuyana}[232, 244]{panarepayne2013}}
 	\label{tab:participles}
-	\begin{tabular}{@{}llll@{}}
-	\mytoprule
-Language & \gl{s_a_} verb & \gl{s_p_} verb & Source(s)\\
-	\mymidrule
-\kaxui & \obj{t-ehurka-t͡ʃe} \qu{fallen} & \obj{tɨ-jaʔ-so} \qu{burnt} & \textcite[39]{schuring2018kaxuyana} \\
-\arara & \obj{t-\emp{o-}ep-te} \qu{come} & \obj{t-oregrum-te} \qu{sad} & \textcite[207, 118]{alves2017arara} \\
-\trio & \obj{tɨ-\emp{w-}əturu-e} \qu{talked} & \obj{t-əpəə-se} \qu{tired} & \textcite[333--334]{triomeira1999} \\
-\wayana & \obj{tə-\emp{w-}epɨ-he} \qu{bathed} & \obj{t-onopɨ-he} \qu{painted} & \textcite[400]{wayanatavares2005} \\
-\kalina & \obj{tu-\emp{w-}oʔka-se} \qu{come out} & \obj{t-okari-se} \qu{told} & \textcite[433, 430]{hoff1968carib} \\
-\apalai & \obj{t-\emp{o-}erama-se} \qu{returned} & \obj{t-anorɨ-se} \qu{dried} & \textcite[42, 30]{camargo2002lexico}\\
-\panare & \obj{t-\emp{o-}tatɨhpə-se} \qu{wailed} & \obj{tɨ-sɨrɨke-t͡ʃe} \qu{tired} & \textcite[232, 244]{panarepayne2013} \\
-	\mybottomrule
-	\end{tabular}
+\input{floats/participles}
+\end{table}
+
+\begin{table}
+	\centering
+	\caption{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs \parencites[49, 74]{schuring2018kaxuyana}[97]{alves2017arara}[246]{triomeira1999}[130, 409]{wayanatavares2005}[90]{koehn1986apalai}[ner2-003]{koehns1994textos}[135, 392]{hoff1968carib}[390]{panarepayne2013}[23]{mattei1994diccionario}}
+	\label{tab:nmlz}
+	\input{floats/nominalizations}
 \end{table}
@@
 This is shown by \textcite{gildea1998}, using the Parukotoan languages as an example.
-I have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription -- and in the case of \kaxui, the addition of Ø/\obj{j-} as an alternative \gl{1}\gl{s_p_} marker \perscommpar{Spike Gildea}.
+I have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription -- and in the case of \kaxui, the addition of ∅/\obj{j-} as an alternative \gl{1}\gl{s_p_} marker \perscommpar{Spike Gildea}.
 Apart from segmental changes to individual morphemes, the following restructuring innovations happened in the \setone paradigm in Parukotoan:
@@
 \mymidrule
-\gl{1} & Ø/\obj{ɨ-} & \obj{ɨ-}/\obj{j-}\\
+\gl{1} & ∅/\obj{ɨ-} & \obj{ɨ-}/\obj{j-}\\
 \gl{2} & \obj{m(ɨ)-} & \obj{o-}\\
@@
 \mymidrule
-\gl{1} & \obj{w(ɨ)-} & Ø/\obj{j-}\\
+\gl{1} & \obj{w(ɨ)-} & ∅/\obj{j-}\\
 \gl{2} & \multicolumn{2}{c}{\emp{\obj{m(ɨ)-}}}\\
@@
 %\mymidrule
-%\gl{1} & \multicolumn{2}{c}{\emp{Ø}}\\
+%\gl{1} & \multicolumn{2}{c}{\emp{∅}}\\
 %\gl{2} & \multicolumn{2}{c}{\emp{\obj{m-}}}\\
@@
 \gl{1+2} & \emp{\obj{h(ɨ)-}}\\
-\gl{3} & \obj{n-}/Ø\\
+\gl{3} & \obj{n-}/∅\\
 \mybottomrule

### 2021-09-26 documents/sections/extensions.tex
added split-s illustration data, more examples
added additional split-s data, bakairi examples, rewrote up to sec 3.4

@@
 \label{sec:extensions}
-As discussed in \cref{sec:extensions_intro}, the person marker extensions which did not spread through the whole lexicon all have in common that they are innovative \gl{1}\gl{s_a_} markers.
+As stated in \cref{sec:extensions_intro}, the person marker extensions which did not spread through the entire lexicon all have in common that they feature innovative \gl{1}\gl{s_a_} markers.
 There are six attested such innovations, three of which can be reconstructed to intermediate proto-languages.
@@
 \cref{sec:pekodian} investigates the innovation of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
-\cref{sec:waiwaian} takes a closer look at the extension of \rc{k-} in \PWai, which was already shown in \cref{sec:extensions_intro}.
+\cref{sec:waiwaian} takes a closer look at the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
 \cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir or \PTar, reflected in modern \trio and \akuriyo.
@@
 \gl{1+2}\gl{s_a_} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
-\gl{3}\gl{s_a_} & \obj{n-əku-} & \obj{Ø-origu-} & \obj{Ø-aranme-} \\
+\gl{3}\gl{s_a_} & \obj{n-əku-} & ∅\obj{-origu-} & ∅\obj{-aranme-} \\
 \end{tabular}
@@
 For \arara, \textcite{alves2017arara} reports six\footnote{Seven under her analysis, which treats the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.} \gl{s_a_} verbs which have a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
-We took these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.
+We take these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.
 
@@
 \obj{w-ɨdolɨ} & \qu{I went}\\
+\end{tabular}
+\xe
+%
+From a comparative view, this list is not quite complete, as there is also a reflex of the copula \rc{a(p)}, which is analayzed as a postposition by \textcite[199--201]{alves2017arara}; it is however inflected with \setone prefixes and also shows first person \obj{w-} \exref{araap}.
+
+\ex<araap>
+\begin{tabular}[t]{@{}ll@{}}
+\gl{1} & \obj{w-aptam}\\
+\gl{2} & \obj{m-od-aptam}\\
+\gl{1+2} & \obj{kud-aptam}\\
+\gl{3} & ∅\obj{-aptam}\\
 \end{tabular}
@@
 However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
-Given the lack of data for other verbs, I will not further discuss this potential additional pattern.
+Given the lack of data for other verbs, we will not further discuss this potential additional pattern.
 If the characterization by \citeauthor{meira2003bakairi} is accurate, then the pattern is fully parallel to the distribution of the first person prefixes.}
 Luckily, while \textcite{von1892bakairi} did not accurately record all phonemic distinctions in \bakairi \parencite{meira2005bakairi}, he does provide inflected forms of cognates to the \arara verbs in \exref{arairr}.
-I present them according to the phonemic analysis of \bakairi by \textcites{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.
+We present them according to the phonemic analysis of \bakairi by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.
 
-\pex[everyglpreamble=]<bakverbs> \bakairi \parencite[][131, 397, 76, 137, 374]{von1892bakairi}
+\pex[everyglpreamble=]<bakverbs> \bakairi \parencite[][131, 397, 76, 137, 374, 130]{von1892bakairi}
 \begin{multicols}{2}
@@
 \endgl
+\a<bak-35>
+\begingl
+\glpreamble \ort{töre-w-akine}//
+\gla tərə w-a-kɨne//
+\glb there \gl{1}\gl{s_a_}-be-\gl{pst}.\gl{cont}//
+\glft \qu{I was there.}//
+\endgl
 \end{multicols}
@@
 
-Turning to the third Pekodian language, all available descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
-However, most \ikpeng cognates of the \arara verbs in \exref{arairr} actually do not take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
+All available descriptions of the third Pekodian language, \ikpeng, list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
+However, most \ikpeng cognates of the \arara verbs with \gl{1}\gl{s_a_} \obj{w-} actually do not take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
 The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
-There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down} \exref{ikp-182}, but no first person forms are attested \perscommpar{Angela Chagas}.
+There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
+While there are reflexes of \rc{a(p)} \qu{to be} in \ikpeng, only \rc{eti} forms are attested inflected for first person \parencite[401]{gildea2018reconstructing}.
 
@@
 
-\ex<ikp-182>\ikpeng \parencite[][320]{chagas2019ikpeng}\\
-\begingl
-\gla Ø-iptoŋ-lan//
-\glb \gl{3}\gl{s_a_}-sit.down-\gl{pst}//
-\glft \qu{She sat down.}//
-\endgl
-\xe
+%\ex<ikp-182>\ikpeng \parencite[][320]{chagas2019ikpeng}\\
+%\begingl
+%\gla ∅-iptoŋ-lan//
+%\glb \gl{3}\gl{s_a_}-sit.down-\gl{pst}//
+%\glft \qu{She sat down.}//
+%\endgl
+%\xe
 %
 While clearly not a regular sound change, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \exref{ikpwloss}.
-Thus, we suggest that the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}, and was subject to phonological erosion of \rc{w}, a phenomenon which will also be encountered in languages discussed in following sections.
+Thus, we suggest that the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}, and was subject to phonological erosion of \rc{w}, a phenomenon which will also be seen in other languages discussed in following sections.
 
@@
 \qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{Ø-ip-} & \obj{w-i-} \\
-\qu{be} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
+\qu{be-1} & \rc{w-ap-} & \obj{w-ap-} & -- & \obj{w-a-} \\
+\qu{be-2} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
 \qu{come} & \rc{w-epɨ}, \rc{k-əd-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
@@
 
-\Cref{tab:pekverbs} gives an overview of the first person forms of the six verbs under discussion, along with our \PPek reconstruction.
-The form for \rc{ge} \qu{to say} is straightforward to reconstruct; the \bakairi change of \rc{wɨ-} to \obj{u-} is not implausible given correspondences like \bakairi \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+\Cref{tab:pekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with our \PPek reconstruction.
+The form for \rc{ge} \qu{to say} is straightforward to reconstruct; the \bakairi change of \rc{wɨ-} to \obj{u-} is not unexpected given correspondences like \bakairi \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
 \qu{to bathe} is straightforwardly reconstructible as \rc{ipɨ}.
@@
 %
-Thus, it seems that there was some variation in \PPar as to whether or not the first person form of \qu{to go} was based on \rc{ətepɨ} or \rc{epɨ}.
-Given that \bakairi and \ikpeng agree about both the first person prefix and the reflex of \rc{ət-}, I reconstruct two distinct first person forms, \rc{w-epɨ-} and \rc{k-ət-epɨ-}.\footnote{It may also be the case that \bakairi and \ikpeng both independently regularized the paradigm to show \rc{ədepɨ} with all persons.
-If that were the case, the connection between \rc{k-} and \rc{ədepɨ} would still exist.}
+We can imagine two scenarios leading to this situation:
+One possibility is that there was some variation in \PPar as to whether or not the first person form of \qu{to go} was based on \rc{ətepɨ} or \rc{epɨ}.
+Alternatively, the \PPek paradigm had a similar person-based distribution of \rc{əd-} as \arara, and \bakairi and \ikpeng both independently regularized the paradigm to show \rc{əd-epɨ} with all persons.
+However, in both scenarios there is a correlation of \rc{w-} and \rc{ebɨ}, and \rc{k-} and \rc{əd-ebɨ}.
 
-The forms for \qu{to go down} are again not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
+The forms for \qu{to go down} are also not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
 Rather, it seems that some roots gained a final \obj{ŋ} in \PXin, for so far unclear reasons: \begin{inlinelist}
@@
 \end{inlinelist}.
-Also, based on the fact \bakairi has generally lost much segmental material, I suggest that \bakairi \obj{gɨ} is yet another addition to a root \rc{ɨptə} or \rc{iptə}, rather than a conservative form.
+Also, based on the fact \bakairi has generally lost much segmental material, we suggest that \bakairi \obj{gɨ} is yet another addition to a root \rc{ɨptə} or \rc{iptə}, rather than a conservative form.
 
-The forms for \qu{to go} are all cognate, but while \arara and \bakairi point to \rc{ɨtə}, the \ikpeng form is rather divergent.
+The forms for \qu{to go} are all cognate, but while \arara and \bakairi point to \rc{ɨtə}, the \ikpeng form seems rather divergent.
 However, it is in fact compatible with a reconstruction \rc{ɨtən}, when one considers that \ikpeng \obj{a} is an attested outcome of \rc{ə} (\PXin \rc{o}/\obj{e}), for example \obj{akari} \qu{dog} in \exref{ikpwloss} above, \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg}, or \obj{jaj} \qu{tree} from \PC \rc{jəje}.
@@
 Summing up, an innovative \gl{1}\gl{s_a_} marker \rc{k-} is reconstructible to \PPek.
-Six verbs can be reconstructed as having resisted this innovation, and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
-For \qu{to come}, two distinct forms \rc{w-epɨ-} and \rc{k-əd-epɨ-} must be reconstructed.
+Seven verbs can be reconstructed as having resisted this innovation, and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
+For \qu{to come}, two distinct forms \rc{w-epɨ-} and \rc{k-əd-epɨ-} can be reconstructed, although the latter may be two parallel developments in \bakairi and \ikpeng.
 In later, individual developments, \bakairi introduced \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
@@
 %
-The \waiwai verb \obj{ka} \qu{to say} was already shown have resisted this extension in \cref{sec:extensions_intro}, showing \obj{wɨ-}, a reflex of the \PC \gl{1}\gl{s_a_} marker \rc{w-}.
+\waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
 Its \hixka counterpart also takes a reflex of \gl{1}\gl{s_a_} \rc{w-}, which has the form \obj{ɨ-} \exref{hixɨ.hix-118}.
@@
 \pex<hixɨ>
-\a<hix-118> \hixka \parencite[][124]{hixkaryanaderby1985}\\
+\a<wai-108> \waiwai \parencite[][71]{waiwaihawkins1998}
+\begingl
+\glpreamble wɨɨkekɲe//
+\gla wɨ-ka-jakɲe//
+\glb \gl{1}-say-\gl{pst}//
+\glft \qu{I said.}//
+\endgl
+\a<hix-118> \hixka \parencite[][124]{hixkaryanaderby1985}
 \begingl
@@
 \endgl
-\a<hix-17> \hixka \parencite[][191]{hixkaryanaderby1985}\\
+\a<hix-17> \hixka \parencite[][191]{hixkaryanaderby1985}
 \begingl
@@
 \endgl
-\a<wai-159> \waiwai \parencite[][192]{waiwaihawkins1998}\\
+\a<wai-159> \waiwai \parencite[][192]{waiwaihawkins1998}
 \begingl
@@
 
-There are two more verbs which did not take innovative \rc{k-} in \PWai.
-One of them is the copula \rc{eʃi}/\obj{a(h)}, reconstructed in \exref{parpar.parcop}.
+There are three more verbs which did not take innovative \rc{k-} in \PWai.
+Two roots are part of the copula \rc{eʃi}/\obj{a(h)}, straightforwardly reconstructed in \exref{parpar.parcop}.
 The other is \qu{to go} \exref{parpar.pargo}, somewhat of a special case.
@@
 
-Summing up, I reconstruct the three verbs \rc{et͡ʃi}/\obj{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{to} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
+Summing up, we reconstruct the three verbs \rc{et͡ʃi}/\obj{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{to} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
 
@@
 The \setone paradigms of \trio and \akuriyo contain an interesting puzzle: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
-This is illustrated by comparing reconstructible \PTir forms with more conservative \wayana \exref{sat} and \kalina \exref{13w}, which can be taken as representative for \PC.
-
-\ex<sat>
-\begin{tabular}[t]{@{}llllll@{}}
-\gl{1}\gl{s_a_} & \PC & \wayana & \PTir & \trio & \akuriyo\\
-\qu{yawn} & \rc{w-e-mɨta-po-} & \obj{w-emtapɨ-} & \rc{t͡ʃ-entapo-} & \obj{s-entapo-} & \obj{t͡ʃ-entapo-}\\
-\qu{bathe} & \rc{w-e-pɨ-} & \obj{w-epɨ-} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-}\\
-\qu{talk} & \rc{w-ət-uru} & \obj{w-əturu-} & \rc{t-əturu-} & \obj{t-əturu-} & \obj{k-əturu-}\footnotemark\\
-\qu{learn} & \rc{w-ət-eme-pa} & \obj{w-əhepa-} & \rc{t-əempa-}\footnotemark & \obj{t-əenpa-} & \obj{k-əempa-}\\
-\end{tabular}\\
-\parencites[28]{camargo2010wayana}[292, 664]{triomeira1999}[77, 86, 82]{gildea1994akuriyo}[50, 208, 213]{wayanatavares2005}
-\xe
-\addtocounter{footnote}{-1}
-\footnotetext{\PTir \rc{t-} was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
-\addtocounter{footnote}{1}
-\footnotetext{
-The \trio form is not due to dissimilation from \rc{mp} to \obj{np}, but rather a phonemic transcription rendering the syllable-final archiphoneme \pnm{N} as \obj{n} \parencite[31]{triomeira1999}.
-Further, \PTar intervocalic \rc{t͡ʃ} was lost entirely in \trio (except after \rc{i}) and is reflected as \obj{ʔ} in \akuriyo \parencite[31--32]{meira1998proto}.
-Given the original shape of the detransitivizers \rc{ət(e)-} and \rc{e-}, \trio \obj{ə-enpa} \qu{to learn} contains a reflex of the former with loss of \rc{t}, so \akuriyo \obj{əempa} would be expected to have a glottal stop: ?\obj{əʔempa}.
-Its absence could either be due to \trio influence or not being transcribed by \textcite{gildea1994akuriyo}.
-Since there is no evidence for a glottal stop in \PTir (\rc{əʔempa}), I do not reconstruct it.}
-
-\ex<13w> 
-\begin{tabular}[t]{@{}llllll@{}}
-\gl{1}>\gl{3}& \PC & \kalina & \PTir & \trio & \akuriyo \\
-\qu{see} & \rc{t-ene-} & \obj{s-ene-} & \rc{w-ene-} & \obj{w-ene-} & \obj{(w-)ene-} \\
-\qu{bring} & \rc{t-enepɨ-} & \obj{s-enepɨ-} & \rc{w-enepɨ-} & \obj{w-enepɨ-} & Ø\obj{-enepɨ-} \\
-\end{tabular}\\
-\parencites[425--426]{hoff1968carib}[28, 304]{triomeira1999}[86--87]{gildea1994akuriyo}
-\xe
+%This is illustrated by comparing reconstructible \PTir forms with more conservative \wayana \exref{sat} and \kalina \exref{13w}, which can be taken as representative for \PC.
 %
-Thus, \wayana preserves the \PC \gl{1}\gl{s_a_} marker \rc{w-}, while it was replaced by \rc{t-} (\phonr{}{\rc{t͡ʃ-}}{\rc{e}}) in \PTir \exref{sat}.
-%Both \qu{to yawn} and \qu{to bathe} are transparently derived \gl{s_a_} verbs, from \rc{mɨta-po} \qu{to open the mouth of} and \rc{pɨ} \qu{to wash}, respectively.
-On the other hand, \gl{1}>\gl{3} scenarios show a reflex of \rc{t-} in \kalina, while they took on \rc{w-} in \PTir \exref{13w}.
-
-The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}\footnote{In fact, the issue of \emph{when} this happened is also open.
+%\ex<sat>
+%\begin{tabular}[t]{@{}llllll@{}}
+%\gl{1}\gl{s_a_} & \PC & \wayana & \PTir & \trio & \akuriyo\\
+%\qu{yawn} & \rc{w-e-mɨta-po-} & \obj{w-emtapɨ-} & \rc{t͡ʃ-entapo-} & \obj{s-entapo-} & \obj{t͡ʃ-entapo-}\\
+%\qu{bathe} & \rc{w-e-pɨ-} & \obj{w-epɨ-} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-}\\
+%\qu{talk} & \rc{w-ət-uru} & \obj{w-əturu-} & \rc{t-əturu-} & \obj{t-əturu-} & \obj{k-əturu-}\footnotemark\\
+%\qu{learn} & \rc{w-ət-eme-pa} & \obj{w-əhepa-} & \rc{t-əempa-}\footnotemark & \obj{t-əenpa-} & \obj{k-əempa-}\\
+%\end{tabular}\\
+%\parencites[28]{camargo2010wayana}[292, 664]{triomeira1999}[77, 86, 82]{gildea1994akuriyo}[50, 208, 213]{wayanatavares2005}
+%\xe
+%\addtocounter{footnote}{-1}
+%\footnotetext{\PTir \rc{t-} was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
+%\addtocounter{footnote}{1}
+%\footnotetext{
+%The \trio form is not due to dissimilation from \rc{mp} to \obj{np}, but rather a phonemic transcription rendering the syllable-final archiphoneme \pnm{N} as \obj{n} \parencite[31]{triomeira1999}.
+%Further, \PTar intervocalic \rc{t͡ʃ} was lost entirely in \trio (except after \rc{i}) and is reflected as \obj{ʔ} in \akuriyo \parencite[31--32]{meira1998proto}.
+%Given the original shape of the detransitivizers \rc{ət(e)-} and \rc{e-}, \trio \obj{ə-enpa} \qu{to learn} contains a reflex of the former with loss of \rc{t}, so \akuriyo \obj{əempa} would be expected to have a glottal stop: ?\obj{əʔempa}.
+%Its absence could either be due to \trio influence or not being transcribed by \textcite{gildea1994akuriyo}.
+%Since there is no evidence for a glottal stop in \PTir (\rc{əʔempa}), I do not reconstruct it.}
+%
+%\ex<13w> 
+%\begin{tabular}[t]{@{}llllll@{}}
+%\gl{1}>\gl{3}& \PC & \kalina & \PTir & \trio & \akuriyo \\
+%\qu{see} & \rc{t-ene-} & \obj{s-ene-} & \rc{w-ene-} & \obj{w-ene-} & \obj{(w-)ene-} \\
+%\qu{bring} & \rc{t-enepɨ-} & \obj{s-enepɨ-} & \rc{w-enepɨ-} & \obj{w-enepɨ-} & Ø\obj{-enepɨ-} \\
+%\end{tabular}\\
+%\parencites[425--426]{hoff1968carib}[28, 304]{triomeira1999}[86--87]{gildea1994akuriyo}
+%\xe
+%%
+%Thus, \wayana preserves the \PC \gl{1}\gl{s_a_} marker \rc{w-}, while it was replaced by \rc{t-} (\phonr{}{\rc{t͡ʃ-}}{\rc{e}}) in \PTir \exref{sat}.
+%%Both \qu{to yawn} and \qu{to bathe} are transparently derived \gl{s_a_} verbs, from \rc{mɨta-po} \qu{to open the mouth of} and \rc{pɨ} \qu{to wash}, respectively.
+%On the other hand, \gl{1}>\gl{3} scenarios show a reflex of \rc{t-} in \kalina, while they took on \rc{w-} in \PTir \exref{13w}.
+%
+The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{In fact, the issue of \emph{when} this happened is also open.
 It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation.
 Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.}
-The relevant aspect of this extension is of course that not all verbs were affected by it.
-\textcite{meira1998proto} lists the four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}.
+As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}; we offer our reconstructed \PTir forms.
 
-\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-}\\
+\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-} \parencites[112--115]{meira1998proto}[85]{gildea1994akuriyo}\\
 \begin{tabular}[t]{@{}llll@{}}
@@
 \qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
-\qu{come} & \rc{w-əʔepɨ-/}\obj{w-eepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
+\qu{come} & \rc{w-əʔepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
 \qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
-\end{tabular}\\
-\parencites[112--115]{meira1998proto}[85]{gildea1994akuriyo}
+\end{tabular}
 %\a<aku12> \akuriyo \obj{kəʔ-eepɨ} \qu{we came} \parencite[114]{meira1998proto}
@@
 
-For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} gives a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
+For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} registered a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
 The two forms can be reconciled by a specific idiosyncratic combination of sound changes:
-I suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
+We suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
 Also, \obj{w-} appears to have been subject to ongoing erosion in \akuriyo, also evidenced by its absence in \qu{to come} and \qu{to be}, but its presence in \qu{to say} \exref{ptirw}.
-This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix (\getfullref{13w},  \getfullref{exp.aku-164}), as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
+This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix \exref{exp.aku-164}, as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
 
@@
 \xe
+%
+For the form of the verb, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for the first person, and \rc{eepɨ} for the others, based on the idiosyncratic form in \trio and the vowel length in \akuriyo.
+\akuriyo (and \carijo) would then have levelled the original distribution of \rc{əepɨ} and \rc{eepɨ}, similar to what we have suggested for Pekodian in \cref{sec:pekodian}.
+We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly cognate with \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ}.
 
-For \qu{to come}, a similar situation as in Pekodian is encountered, in that \trio shows a reflex of \rc{ət-epɨ}, while the \akuriyo form points to bare \rc{epɨ}.
-Given that \trio, like \arara, also has forms without a reflex of \rc{ət-} \exref{tri-130}, I reconstruct both \rc{w-epɨ-} and \rc{w-əʔepɨ-} for \PTir.
-
-\ex<tri-130> \trio \parencite[][114]{triomeira1999}\\
-\begingl
-\gla n-epɨ//
-\glb \gl{3}-come//
-\glft \qu{S/he came.}//
-\endgl
-\xe
+%\ex<tri-130> \trio \parencite[][114]{triomeira1999}\\
+%\begingl
+%\gla n-epɨ//
+%\glb \gl{3}-come//
+%\glft \qu{S/he came.}//
+%\endgl
+%\xe
 
@@
 They are listed in \exref{akumov.aku}.
-I have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in that it takes \obj{s-} \exref{akumov.tri}.
+We have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
 Thus, it is uncertain whether these verbs can be reconstructed as taking \rc{w-} in \PTir, although its presence in \akuriyo does suggest so.
@@
 
-\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes\\
+\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
 \begin{tabular}[t]{@{}lll@{}}
@@
 \qu{defecate} & \obj{k-oeka-} & ?\\
-\end{tabular}\\
-\parencites[294]{triomeira1999}[84]{gildea1994akuriyo}
+\end{tabular}
 \xe
 
-In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} was extended to \gl{1} verbs.
+In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} largely replaced \gl{1}\gl{s_a_} \rc{w-}.
 Four verbs are solidly reconstructible as preserving \rc{w-} in \PTir, and \rc{ɨhtə} \qu{to go down} with irregular \rc{p-} was also not affected.

### 2021-09-26 documents/sections/proto_carib.tex
added additional split-s data, bakairi examples, rewrote up to sec 3.4

@@
 
-\begin{table}
-	\centering
-	\caption{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs}
-	\label{tab:nmlz}
-	\begin{tabular}{@{}llll@{}}
-	\mytoprule
-Language & \gl{s_a_} verb & \gl{s_p_} verb & Source(s)\\
-	\mymidrule
-\kaxui & \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} & \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} & \textcite[49, 74]{schuring2018kaxuyana}\\
-\arara & \obj{\emp{w-}orik-tubo} \qu{dancing place} & \obj{ereŋmɨ-tpo} \qu{killing instrument} & \textcite[97]{alves2017arara} \\
-\trio & \obj{ji-\emp{w-}əturu-to} \qu{(for) my talking} & \obj{j-emamina-to} \qu{(for) my playing} & \textcite[246]{triomeira1999} \\
-\wayana & \obj{ɨ-\emp{w-}əturu-topo} \qu{my story} & \obj{j-ɨnɨkɨ-topo} \qu{my object for sleeping} & \textcite[409, 130]{wayanatavares2005}\\
-\kalina & \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} & \obj{aj-ereʔna-Ø} \qu{your fainting} & \textcite[392, 135]{hoff1968carib}\\
-\apalai & \obj{j-epɨ-topo} \qu{my bath place} & \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} & \textcites[90]{koehn1986apalai}[ner2/1]{koehns1994textos}\\
-\panare & \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} & \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} & \textcites[390]{panarepayne2013}[23]{mattei1994diccionario}\\
-	\mybottomrule
-	\end{tabular}
-\end{table}
+%\begin{table}
+%	\centering
+%	\caption{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs}
+%	\label{tab:nmlz}
+%	\begin{tabular}{@{}llll@{}}
+%	\mytoprule
+%Language & \gl{s_a_} verb & \gl{s_p_} verb & Source(s)\\
+%	\mymidrule
+%\kaxui & \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} & \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} & \textcite[49, 74]{schuring2018kaxuyana}\\
+%\arara & \obj{\emp{w-}orik-tubo} \qu{dancing place} & \obj{ereŋmɨ-tpo} \qu{killing instrument} & \textcite[97]{alves2017arara} \\
+%\trio & \obj{ji-\emp{w-}əturu-to} \qu{(for) my talking} & \obj{j-emamina-to} \qu{(for) my playing} & \textcite[246]{triomeira1999} \\
+%\wayana & \obj{ɨ-\emp{w-}əturu-topo} \qu{my story} & \obj{j-ɨnɨkɨ-topo} \qu{my object for sleeping} & \textcite[409, 130]{wayanatavares2005}\\
+%\kalina & \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} & \obj{aj-ereʔna-Ø} \qu{your fainting} & \textcite[392, 135]{hoff1968carib}\\
+%\apalai & \obj{j-epɨ-topo} \qu{my bath place} & \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} & \textcites[90]{koehn1986apalai}[ner2/1]{koehns1994textos}\\
+%\panare & \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} & \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} & \textcites[390]{panarepayne2013}[23]{mattei1994diccionario}\\
+%	\mybottomrule
+%	\end{tabular}
+%\end{table}
 
@@
 This is illustrated with reflexes in various modern languages in \cref{tab:imp}.
-As in the case of the \gl{s_a_} marker \rc{w-} participles and nominalizations, some languages have lost the distinction between \gl{s_a_} and \gl{s_p_} verbs in imperatives, for example \panare, or \kaxui (not shown in \cref{tab:imp}).
+As in the case of the \gl{s_a_} marker \rc{w-} participles and nominalizations, some languages have lost the distinction between \gl{s_a_} and \gl{s_p_} verbs in imperatives, for example \panare.
 
@@
 	\centering
-	\caption{Imperatives of \gl{s_a_} and \gl{s_p_} verbs}
+	\caption{Imperatives of \gl{s_a_} and \gl{s_p_} verbs \parencites[44, 89]{derbyshire1965textos}[161]{alves2017arara}[323]{triomeira1999}[227]{wayanatavares2005}[62]{koehn1986apalai}[Mopo/20]{koehns1994textos}[190]{hoff1968carib}[5, 17]{mattei1994diccionario}}
 	\label{tab:imp}
-	\begin{tabular}{@{}llll@{}}
-	\mytoprule
-Language & \gl{s_a_} & \gl{s_p_} & Source(s)\\
-	\mymidrule
-\hixka & \obj{omoh-ko} \qu{come!} & \obj{\emp{oj-}okajɨm-ko} \qu{go up!} & \textcite[89, 44]{derbyshire1965textos}\\
-\arara & \obj{odotpot-ko} \qu{come back!} & \obj{\emp{o-}alum-ko} \qu{jump!} & \textcite[161]{alves2017arara}\\
-\trio & \obj{epɨ-kə} \qu{bathe!} & \obj{\emp{ə-}eremina-kə} \qu{sing!} & \textcite[323]{triomeira1999}\\
-\wayana & \obj{əməm-kə} \qu{enter!} & \obj{\emp{əw-}eremi-kə} \qu{sing!} &  \textcite[227]{wayanatavares2005} \\
-\kalina & \obj{oʔmaʔ-ko} \qu{stop!} & \obj{\emp{aj-}awon-ko} \qu{get up!} & \textcite[190]{hoff1968carib}\\
-\apalai & \obj{otuʔ-ko} \qu{eat!} & \obj{\emp{o-}nɨʔ-ko} \qu{sleep!} & \textcites[62]{koehn1986apalai}[Mopo/20]{koehns1994textos} \\
-\panare & \obj{ape-ʔ} \qu{flee!} & \obj{ahpən-kə} \qu{jump!} & \textcite[17, 5]{mattei1994diccionario}\\
-	\mybottomrule
-	\end{tabular}
+	\input{floats/imperatives}
 \end{table}
@@
 As for the form of this reflexive marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although they have since merged into a single morpheme in modern Cariban languages.
-Reflexes of \detrz show diverse meanings, which are best subsumed under \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
+Modern reflexes of \detrz show a range of meanings, which can all be characterized as \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
 
@@
 \xe
-
-%\pex<triodetrz> Semantics of the \trio detransitivizer \parencite[218--219]{meira2000split}
-%\a \obj{e-suka}, \obj{əi-suka} \qu{to wash oneself} (reflexive)
-%\a \obj{e-nonta}, \obj{əi-nonta} \qu{to abandon each other} (reciprocal)
-%\a \obj{e-pahka} \qu{to break} (anticausative)
-%\a \obj{əi-puunəəpɨ} \qu{to think} (antipassive)
-%\xe
 %
-The morphological variation featured in \qu{to wash oneself} and \qu{to abandon each other} is due to the mentioned collapse between the two \PC prefixes:
+The morphological variation featured in \qu{to abandon each other} and \qu{to wash self} is due to the mentioned collapse between the two \PC prefixes:
 \obj{e-} is a reflex of the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
@@
 One could also imagine a scenario whereby verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
-However, the verbs he sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial and explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
+However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
 Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
-For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system includes markers being extended to new verbs.}
+For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system features markers being extended to new verbs.}
 There have been many such person marker extensions in Cariban languages, and some are still ongoing.
 This is shown by \textcite{gildea1998}, using the Parukotoan languages as an example.
-I have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription -- and in the case of \kaxui, the addition of ∅/\obj{j-} as an alternative \gl{1}\gl{s_p_} marker \perscommpar{Spike Gildea}.
+We have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription -- and in the case of \kaxui, the addition of ∅/\obj{j-} as an alternative \gl{1}\gl{s_p_} marker \perscommpar{Spike Gildea}.
 Apart from segmental changes to individual morphemes, the following restructuring innovations happened in the \setone paradigm in Parukotoan:
@@
 While different cases of loss of split-\gl{s} are discussed by \textcite[91--96]{gildea1998}, this paper focuses on a so far neglected aspect of these person marking extensions.
-I argue that they are executed via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+We argue that they are executed via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
 
@@
 The remainder of this paper is structured as follows:
-\cref{sec:extensions} investigates each of these six incomplete extensions individually, listing unaffected verbs and reconstructing proto-paradigms, where necessary.
-\cref{sec:verbs} takes a comparative look at the verbs that were not affected by these extensions and searches for possible motivations.
-Finally, \cref{sec:discussion} discusses the findings and puts them in a general context of language change and morphology.
+In \cref{sec:extensions}, we discuss each of these six incomplete extensions individually, listing all attested unaffected verbs, and reconstructing proto-paradigms where necessary.
+In \cref{sec:verbs}, we take a comparative look at the verbs that were not affected by these extensions and searches for possible motivations.
+Finally, we discuss the findings and put them in a general context of language change and morphology in \cref{sec:discussion}.
 

### 2021-09-26 workflow/irregular_paradigms.py

@@ -13,7 +13,7 @@ pyd.z = []
 verb_list = ["come-1", "go", "say", "go.down", "be-1", "be-2"]
 pyd.y_sort = lg_list
 pyd.filters = {"Inflection": ["1"], "Cognateset_ID": verb_list, "Language_ID": lg_list}
-pyd.content_string = "Prefix_Cognateset_ID"
+# pyd.content_string = "Prefix_Cognateset_ID"
 table = pyd.compose_paradigm(df)
 print(table)
 #

### 2021-09-26 documents/sections/extensions.tex
reworded 3.4

@@
 After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
-It seems to have replaced \rc{t-} only in specific places, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to me \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
+It seems to have replaced \rc{t-} only in specific places, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to us \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} could be more recent, with which I agree -- since phonologically distributed \rc{t-}/\rc{t͡ʃ-} is reconstructible to \PTir, a scenario whereby \obj{k-} replaces \rc{t-}, but not \rc{t͡ʃ-}, is the most straightforward.
-The few \obj{t-} recorded by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
+He also suggests that \obj{k-} could be more recent, with which we agree -- since phonologically distributed \rc{t-}/\rc{t͡ʃ-} is reconstructible to \PTir, a scenario whereby \obj{k-} replaces \rc{t-}, but not \rc{t͡ʃ-}, is the most straightforward.
+The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
 

### 2021-09-26 documents/sections/extensions.tex
slimmed down 3.5

@@
 %
-Thus, also detransitivized verbs like \obj{ehɨnəhɨ} \qu{to fight}, which are \gl{s_a_} verbs in other languages, take \obj{j(i)-} \exref{car-30}.
-The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
+Thus, also detransitivized verbs like \obj{ehɨnəhɨ} \qu{to fight}, which are \gl{s_a_} verbs in other languages, take \obj{j-} \exref{car-30}.
+%The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
 
@@
 
-\pex<car1p>\carijo \parencite[][59, 53]{guerrero2016carijo}
-\begin{multicols}{2}
-\a<car-1>
-\begingl
-\gla jɨ-hɨnəh-ɨ//
-\glb \gl{3}>\gl{1}-kill-\gl{pfv}//
-\glft \qu{S/he killed me.}//
-\endgl
-\a<car-31>
-\begingl
-\gla j-owo-rɨ//
-\glb \gl{1}\gl{poss}-uncle-\gl{pert}//
-\glft \qu{my uncle}//
-\endgl
-\end{multicols}
-\xe
+%\pex<car1p>\carijo \parencite[][59, 53]{guerrero2016carijo}
+%\begin{multicols}{2}
+%\a<car-1>
+%\begingl
+%\gla jɨ-hɨnəh-ɨ//
+%\glb \gl{3}>\gl{1}-kill-\gl{pfv}//
+%\glft \qu{S/he killed me.}//
+%\endgl
+%\a<car-31>
+%\begingl
+%\gla j-owo-rɨ//
+%\glb \gl{1}\gl{poss}-uncle-\gl{pert}//
+%\glft \qu{my uncle}//
+%\endgl
+%\end{multicols}
+%\xe
 %
@@
 \glb \gl{1}-go-\gl{npst}=\gl{frust}//
-\glft \qu{I almost go (but I a, not going to go).}//
+\glft \qu{I almost go (but I am not going to go).}//
 \endgl

### 2021-09-26 documents/sections/extensions.tex
proofread 3.6

@@
 
-\subsection{\carijo \obj{j(i)-}}
+\subsection{\carijo \obj{j-}}
 \label{sec:carijo}
@@
 \begingl
-\gla aw Ø-esare//
+\gla aw {\normalfont ∅}-esare//
 \glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
@@
 On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian,sec:akuriyo}, I argue that this zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian,sec:akuriyo}, we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
 The same zero marking is attested for a single intransitive verb: \obj{to} \qu{to go} \exref{yuk-7}.
-Since regular C-initial verbs take \obj{jɨ-}, like \qu{to sleep} in \exref{yukpaintr}, I argue that \obj{to} \qu{to go} preserved the old \gl{1}\gl{s_a_} marker \rc{w(ɨ)-}, which was lost due to phonological erosion.
+Based on regular C-initial verbs taking \obj{jɨ-}, like \qu{to sleep} in \exref{yukpaintr}, it is clear that also C-initial \obj{to} \qu{to go} preserved the old \gl{1}\gl{s_a_} marker \rc{w(ɨ)-}, which was lost due to phonological erosion.
 
@@
 \begingl
-\gla aw Ø-to//
+\gla aw {\normalfont ∅}-to//
 \glb \gl{1}\gl{pro} \gl{1}\gl{s_a_}-go//

### 2021-09-26 documents/sections/verbs.tex
slimmed down intro to section 4

@@
-\section{Resistant verbs in a comparative perspective}
+\section{Resistant verbs from a comparative perspective}
 \label{sec:verbs}
-In \cref{sec:extensions}, six distinct extensions of personal prefixes into \gl{1}\gl{s_a_} territory were introduced, with certain verbs resisting the innovation.
+In \cref{sec:extensions}, we introduced six distinct extensions of personal prefixes into \gl{1}\gl{s_a_} territory, and identified verbs resistant to each innovation.
 The set of unaffected verbs was rather small in most cases, and there is a considerable etymological overlap between the (proto-)language-specific verb groups.
-In this section, I present these verbs from a comparative perspective and discuss their reconstructibility.
-Attentive readers will have noticed that certain verbs reoccurred in these lists, not only matching in meaning but also largely in form.
-In this section, I will discuss these verbs from a comparative perspective.
-\cref{sec:be} treats the copular verb \rc{eti}/\obj{a(p)} \qu{to be}, \cref{sec:say} \rc{ka(ti)} \qu{to say}, \cref{sec:go} \qu{ɨtə(mə)} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstructed as \gl{s_a_} verbs that were not derived from transitive verbs.
+In this section, we present these verbs from a comparative perspective and discuss their reconstructability.
+\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a(p)} \qu{to be}, \cref{sec:say} \rc{ka(ti)} \qu{to say}, \cref{sec:go} \rc{ɨtə(mə)} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstructed as \gl{s_a_} verbs that were not derived from transitive verbs.
 \cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek; \cref{sec:bathe} investigates \PPek \rc{ɨpɨ} \qu{to bathe}, \cref{sec:shit} \trio \qu{to defecate}, and \cref{sec:movement} the \akuriyo movement verbs.

### 2021-09-26 documents/sections/verbs.tex
rewrote 4.1-4.3

@@
 \label{sec:be}
-For a comprehensive comparative overview for this verb, I refer the reader to \textcite[375--382]{gildea2018reconstructing}, who reconstructs two distinct roots serving as verbs \qu{to be} in modern Cariban languages.
+For a comprehensive comparative overview for this verb, we refer the reader to \textcite[375--382]{gildea2018reconstructing}, who reconstructs two distinct roots serving as verbs \qu{to be} in modern Cariban languages.
 One is the older copula \rc{a(p)}, which can be reconstructed as already having various irregularities in \PC.
@@
 Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
-\akuriyo \obj{eʔi} was not a potential target of innovative \obj{k-}, due to the verb being \obj{e}-initial.
+\akuriyo \obj{eʔi} was not a potential target of innovative \obj{k-} \pcref{sec:akuriyo}, due to the verb being \obj{e}-initial.
 \carijo innovated \obj{j-}, but only in the \obj{et͡ʃi} root allomorph \exref{carcop.mivida-12}; the \obj{a} root preserves \obj{w-} \exref{carcop.car-25}.
-\yukpa innovated \obj{j-} for reflexes of both \rc{a(p)} and \rc{eti}, which are preserved as auxiliaries in certain constructions \exref{yukcop}.
+\yukpa innovated \obj{j-} for reflexes of both \rc{a(p)} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
 
@@
 \label{sec:say}
-Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed based on some imperative forms \exref{kati}.
+Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, illustrated with modern imperative forms in \exref{kati}.
 
@@
 This verb was not affected by the extensions found in \PPek, \PWai, and \PTir.
-It was not a potential target of \akuriyo \obj{k-}; I do not have data for its \carijo and \yukpa reflexes.
+It was not a potential target of \akuriyo \obj{k-}; we do not know the first person forms of its \carijo and \yukpa reflexes.
 
@@
 In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
-This complementary distribution is otherwise only found with transitive verbs.
+This complementary distribution is otherwise only found with transitive verbs \parencite[59--60]{gildea1998}.
 The verb shows the same pattern, albeit inconsistently, in \trio \parencite[267]{triocarlin2004}.
@@
 
-The view of \qu{to say} as intransitive is supported primarily by prefix patterns:
+Our classification of \qu{to say} as an intransitive verb is supported primarily by prefix patterns:
 \kalina offers a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.\footnote{Interestingly, the \kalina causativized form \obj{kapo} \qu{to make say} does not have the regular \gl{1}>\gl{3} prefix \obj{s(i)-}, but irregular \obj{w(ɨ)-} \parencite[430]{courtz2008carib}.}
@@
 \label{sec:go}
-This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə(mə)}, with the second syllable only occurring in some reflexes, in certain contexts.
+This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə(mə)}, with the second syllable only rarely occurring, as in the case of \rc{ka(ti)} \qu{to say}.
 It is true that many reflexes are clearly \obj{t}-initial, for example \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}.
-However, a comparison of selected forms with unambiguously C-initial \rc{ka(ti)} \qu{to say} suggests that an initial vowel \rc{ɨ} should be reconstructed \exref{gosaycomp}.\footnote{Many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.}
+However, a comparison of selected forms with unambiguously C-initial \rc{ka(ti)} \qu{to say} suggests that an initial vowel \rc{ɨ} should be reconstructed \exref{gosaycomp},\footnote{Many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.} which was subsequently reanalyzed as part of the prefixes in many languages.
 

### 2021-09-26 documents/sections/verbs.tex
proofread 4.6

@@
 In particular, the Pemongan languages and one of its allomorphs in \kaxui suggest a root \colorbox{come1}{\rc{jəpɨ}}.
-However, the one form in \kaxui only occurs in the Progressive \exref{kaxprog.kax-82}, and \obj{j(-)} may be a reflex of the \settwo third person marker \rc{i-} and the \gl{s_a_} class marker \rc{w-}, as suggested by the occurrence of the latter with other person values \exref{kaxprog.kax-54}.
+However, the one form in \kaxui only occurs in the Progressive \exref{kaxprog.kax-82}, and \obj{j(-)} may be a reflex the \settwo third person marker \rc{i-} combined with  the \gl{s_a_} class marker \rc{w-}, as suggested by the occurrence of the latter with other person values \exref{kaxprog.kax-54}.
 On the other hand, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
@@
 
-I have so far ignored the forms which form the basis for \posscite{gildea2007greenberg} \rc{ətepɨ}; these are compatible with my reconstruction \rc{jəpɨ}, with the addition of the detransitivizer \rc{ət(e)-}: \colorbox{come4}{\rc{ətjəpɨ}}.
-In fact, the \obj{i} in the \akawaio form \obj{əsipɨ} is better explained by a sequence \rc{jə}, which has the same outcome in the \macushi reflex of \rc{jəpɨ}, \obj{ipɨ}.
+we have so far ignored the forms which form the basis for \posscite{gildea2007greenberg} \rc{ətepɨ}; these are compatible with our reconstructed root \rc{jəpɨ}, with the addition of the detransitivizer \rc{ət(e)-}: \colorbox{come4}{\rc{ətjəpɨ}}.
+In fact, the \obj{i} in the \akawaio form \obj{əsipɨ} is better explained by a sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of \rc{jəpɨ}, \obj{ipɨ}.
 While the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
@@
 The form may be due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}.
-Alternatively, \apalai \obj{oepɨ} could be from \wayana, which has lost its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
+Alternatively, \apalai \obj{oepɨ} could be a fossilized loan \wayana, which has lost its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
 the distribution within the family is rather chaotic.
@@
 Rather, an irregular first person form \rc{p-ɨptə} can be reconstructed for \PTir.
-While it seems to have resisted the extension of \rc{k-} in \PPek, \bakairi subsequently introduced it \pcref{sec:pekodian}.
-It was also affected by the extension of \obj{j(i)-} in \carijo \exref{car-32}.
+While it originally resisted the extension of \rc{k-} in \PPek, \bakairi subsequently introduced it \pcref{sec:pekodian}.
+It was also affected by the extension of \obj{j-} in \carijo \exref{car-32}.
 
@@
 The situation in \yukpa is unclear, as it is an open question whether \obj{ew(uh)tu} \qu{to go down} \parencite{meira2003primeras} is cognate.
-
-A comparative view \pcref{tab:descend_cog} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
-The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
-In one language, \wayana, the verb shows traits of both classes, leading me to consider it a \dbqu{mixed} verb.
-It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
-It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.
-
-\pex<waygodown>\wayana \parencite[][200]{wayanatavares2005}
-\a<way-73>
-\begingl
-\glpreamble ïwïptëë//
-\gla ɨ-w-ɨptə-rɨ//
-\glb \gl{1}-\gl{s_a_}-go.down-\gl{nmlz}//
-\glft \qu{my going down}//
-\endgl
-\a<way-71>
-\begingl
-\gla əw-ɨptə-k//
-\glb \gl{2}-go.down-\gl{imp}//
-\glft \qu{Go down!}//
-\endgl
-\xe
-%
-Its causativized form is \obj{ɨptəka} \parencite[255]{wayanatavares2005}, containing a reflex of the transitivizer \rc{-ka}, which was restricted to \gl{s_p_} verbs \parencite{gildea2019overview}.
-
-Comparative evidence suggests that \rc{ɨpɨtə} \qu{to go down} was originally an \gl{s_p_} verb which switched its class to \gl{s_a_} in some languages, incompletely so in \wayana.
-The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:descend_cog} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
-Both causativized forms contain a reflex of the (rare) transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
-Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms \parencite[263]{triomeira1999}.
-Thus, it seems that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.
 
@@
 \end{table}
+
+A comparative view \pcref{tab:descend_cog} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
+The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
+In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb.
+It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
+It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.
+
+\pex<waygodown>\wayana \parencite[][200]{wayanatavares2005}
+\a<way-73>
+\begingl
+\glpreamble ïwïptëë//
+\gla ɨ-w-ɨptə-rɨ//
+\glb \gl{1}-\gl{s_a_}-go.down-\gl{nmlz}//
+\glft \qu{my going down}//
+\endgl
+\a<way-71>
+\begingl
+\gla əw-ɨptə-k//
+\glb \gl{2}-go.down-\gl{imp}//
+\glft \qu{Go down!}//
+\endgl
+\xe
+%
+Its causativized form is \obj{ɨptəka} \parencite[255]{wayanatavares2005}, containing a reflex of the transitivizer \rc{-ka}, which was restricted to \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
+
+Further comparative evidence suggests that \rc{ɨpɨtə} \qu{to go down} was originally an \gl{s_p_} verb which switched its class to \gl{s_a_} in some languages, incompletely so in \wayana.
+The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:descend_cog} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
+Both causativized forms contain a reflex of the (rare) transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
+Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms \parencite[263]{triomeira1999}.
+Thus, it seems that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.
 

### 2021-09-26 workflow/irregular_paradigms.py
started adding PTir forms

@@ -4,14 +4,14 @@ import pandas as pd
 df = pd.read_csv("../data/irregular_verb_data.csv")
 
 lg_list = ["PWai", "wai", "hix", "PPek", "PXin", "ara", "ikp", "bak", "PTir", "aku", "tri", "car", "yuk"]
-for i in ["PWai", "PPek", "PXin", "PTir"]:
-    lg_list.remove(i)
+# for i in ["PWai", "PPek", "PXin", "PTir"]:
+#     lg_list.remove(i)
     
 pyd.x = ["Cognateset_ID"]
 pyd.y = ["Language_ID"]
 pyd.z = []
 verb_list = ["come-1", "go", "say", "go.down", "be-1", "be-2"]
-pyd.y_sort = lg_list
+pyd.x_sort = lg_list
 pyd.filters = {"Inflection": ["1"], "Cognateset_ID": verb_list, "Language_ID": lg_list}
 # pyd.content_string = "Prefix_Cognateset_ID"
 table = pyd.compose_paradigm(df)

### 2021-09-28 documents/floats/imperatives.tex
added imperative data, some optics

\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &                         \gl{s_a_} &                             \gl{s_p_} \\
\midrule
\hixka  &          \obj{omoh-ko} \qu{come!} &  \obj{\emp{oj-}okajɨm-ko} \qu{go up!} \\
\arara  &  \obj{odotpot-ko} \qu{come back!} &      \obj{\emp{o-}alum-ko} \qu{jump!} \\
\trio   &          \obj{epɨ-kə} \qu{bathe!} &   \obj{\emp{ə-}eremina-kə} \qu{sing!} \\
\wayana &         \obj{əməm-kə} \qu{enter!} &    \obj{\emp{əw-}eremi-kə} \qu{sing!} \\
\apalai &           \obj{otuʔ-ko} \qu{eat!} &      \obj{\emp{o-}nɨʔ-ko} \qu{sleep!} \\
\kalina &         \obj{oʔmaʔ-ko} \qu{stop!} &   \obj{\emp{aj-}awon-ko} \qu{get up!} \\
\panare &            \obj{ape-ʔ} \qu{flee!} &             \obj{ahpən-kə} \qu{jump!} \\
\bottomrule
\end{tabular}

### 2021-09-28 documents/sections/extensions.tex
started adding PTir forms
added imperative data, some optics

--- 2021-09-26 documents/sections/extensions.tex
@@
 Luckily, while \textcite{von1892bakairi} did not accurately record all phonemic distinctions in \bakairi \parencite{meira2005bakairi}, he does provide inflected forms of cognates to the \arara verbs in \exref{arairr}.
-We present them according to the phonemic analysis of \bakairi by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.
+We present them according to the phonemic analysis of \bakairi by \textcites{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.
 
-\pex[everyglpreamble=]<bakverbs> \bakairi \parencite[][131, 397, 76, 137, 374, 130]{von1892bakairi}
+\pex[everyglpreamble=]<bakverbs> \bakairi \parencite[][131, 397, 76, 137, 374]{von1892bakairi}
 \begin{multicols}{2}
@@
 \endgl
-\a<bak-35>
-\begingl
-\glpreamble \ort{töre-w-akine}//
-\gla tərə w-a-kɨne//
-\glb there \gl{1}\gl{s_a_}-be-\gl{pst}.\gl{cont}//
-\glft \qu{I was there.}//
-\endgl
 \end{multicols}
@@
 All available descriptions of the third Pekodian language, \ikpeng, list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
-However, most \ikpeng cognates of the \arara verbs with \gl{1}\gl{s_a_} \obj{w-} actually do not take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
+However, most \ikpeng cognates of the \arara verbs in \exref{arairr} actually do not take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
 The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
-There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-While there are reflexes of \rc{a(p)} \qu{to be} in \ikpeng, only \rc{eti} forms are attested inflected for first person \parencite[401]{gildea2018reconstructing}.
+There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down} \exref{ikp-182}, but no first person forms are attested \perscommpar{Angela Chagas}.
 
@@
 
-%\ex<ikp-182>\ikpeng \parencite[][320]{chagas2019ikpeng}\\
-%\begingl
-%\gla ∅-iptoŋ-lan//
-%\glb \gl{3}\gl{s_a_}-sit.down-\gl{pst}//
-%\glft \qu{She sat down.}//
-%\endgl
-%\xe
+\ex<ikp-182>\ikpeng \parencite[][320]{chagas2019ikpeng}\\
+\begingl
+\gla ∅-iptoŋ-lan//
+\glb \gl{3}\gl{s_a_}-sit.down-\gl{pst}//
+\glft \qu{She sat down.}//
+\endgl
+\xe
 %
 While clearly not a regular sound change, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \exref{ikpwloss}.
-Thus, we suggest that the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}, and was subject to phonological erosion of \rc{w}, a phenomenon which will also be seen in other languages discussed in following sections.
+Thus, we suggest that the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}, and was subject to phonological erosion of \rc{w}, a phenomenon which can also be found in other languages discussed in following sections.
 
@@
 \qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{Ø-ip-} & \obj{w-i-} \\
-\qu{be-1} & \rc{w-ap-} & \obj{w-ap-} & -- & \obj{w-a-} \\
-\qu{be-2} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
+\qu{be} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
 \qu{come} & \rc{w-epɨ}, \rc{k-əd-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
@@
 
-\Cref{tab:pekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with our \PPek reconstruction.
+\Cref{tab:pekverbs} gives an overview of the first person forms of the six verbs under discussion, along with our \PPek reconstruction.
 The form for \rc{ge} \qu{to say} is straightforward to reconstruct; the \bakairi change of \rc{wɨ-} to \obj{u-} is not unexpected given correspondences like \bakairi \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
@@
 %
-We can imagine two scenarios leading to this situation:
-One possibility is that there was some variation in \PPar as to whether or not the first person form of \qu{to go} was based on \rc{ətepɨ} or \rc{epɨ}.
-Alternatively, the \PPek paradigm had a similar person-based distribution of \rc{əd-} as \arara, and \bakairi and \ikpeng both independently regularized the paradigm to show \rc{əd-epɨ} with all persons.
-However, in both scenarios there is a correlation of \rc{w-} and \rc{ebɨ}, and \rc{k-} and \rc{əd-ebɨ}.
+Thus, it seems that there was some variation in \PPar as to whether or not the first person form of \qu{to go} was based on \rc{ətepɨ} or \rc{epɨ}.
+Given that \bakairi and \ikpeng agree about both the first person prefix and the reflex of \rc{ət-}, we reconstruct two distinct first person forms, \rc{w-epɨ-} and \rc{k-ət-epɨ-}.\footnote{It may also be the case that the \PPek paradigm had a similar person-based distribution of \rc{əd-} as \arara, and \bakairi and \ikpeng both independently regularized the paradigm to show \rc{əd-epɨ} with all persons.
+If that were the case, the connection between \rc{k-} and \rc{əd-epɨ} would still exist.}
 
-The forms for \qu{to go down} are also not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
+The forms for \qu{to go down} are again not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
 Rather, it seems that some roots gained a final \obj{ŋ} in \PXin, for so far unclear reasons: \begin{inlinelist}
@@
 
-The forms for \qu{to go} are all cognate, but while \arara and \bakairi point to \rc{ɨtə}, the \ikpeng form seems rather divergent.
+The forms for \qu{to go} are all cognate, but while \arara and \bakairi point to \rc{ɨtə}, the \ikpeng form is rather divergent.
 However, it is in fact compatible with a reconstruction \rc{ɨtən}, when one considers that \ikpeng \obj{a} is an attested outcome of \rc{ə} (\PXin \rc{o}/\obj{e}), for example \obj{akari} \qu{dog} in \exref{ikpwloss} above, \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg}, or \obj{jaj} \qu{tree} from \PC \rc{jəje}.
@@
 Seven verbs can be reconstructed as having resisted this innovation, and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
-For \qu{to come}, two distinct forms \rc{w-epɨ-} and \rc{k-əd-epɨ-} can be reconstructed, although the latter may be two parallel developments in \bakairi and \ikpeng.
+For \qu{to come}, two distinct forms \rc{w-epɨ-} and \rc{k-əd-epɨ-} must be reconstructed.
 In later, individual developments, \bakairi introduced \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
@@
 %
-\waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
+The \waiwai verb \obj{ka} \qu{to say} was already shown have resisted this extension in \cref{sec:extensions_intro}, showing \obj{wɨ-}, a reflex of the \PC \gl{1}\gl{s_a_} marker \rc{w-}.
 Its \hixka counterpart also takes a reflex of \gl{1}\gl{s_a_} \rc{w-}, which has the form \obj{ɨ-} \exref{hixɨ.hix-118}.
@@
 \pex<hixɨ>
-\a<wai-108> \waiwai \parencite[][71]{waiwaihawkins1998}
-\begingl
-\glpreamble wɨɨkekɲe//
-\gla wɨ-ka-jakɲe//
-\glb \gl{1}-say-\gl{pst}//
-\glft \qu{I said.}//
-\endgl
-\a<hix-118> \hixka \parencite[][124]{hixkaryanaderby1985}
+\a<hix-118> \hixka \parencite[][124]{hixkaryanaderby1985}\\
 \begingl
@@
 \endgl
-\a<hix-17> \hixka \parencite[][191]{hixkaryanaderby1985}
+\a<hix-17> \hixka \parencite[][191]{hixkaryanaderby1985}\\
 \begingl
@@
 \endgl
-\a<wai-159> \waiwai \parencite[][192]{waiwaihawkins1998}
+\a<wai-159> \waiwai \parencite[][192]{waiwaihawkins1998}\\
 \begingl
@@
 
-There are three more verbs which did not take innovative \rc{k-} in \PWai.
-Two roots are part of the copula \rc{eʃi}/\obj{a(h)}, straightforwardly reconstructed in \exref{parpar.parcop}.
+There are two more verbs which did not take innovative \rc{k-} in \PWai.
+One of them is the copula \rc{eʃi}/\obj{a(h)}, reconstructed in \exref{parpar.parcop}.
 The other is \qu{to go} \exref{parpar.pargo}, somewhat of a special case.
@@
 
-Summing up, we reconstruct the three verbs \rc{et͡ʃi}/\obj{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{to} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
+Summing up, I reconstruct the three verbs \rc{et͡ʃi}/\obj{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{to} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
 
@@
 The \setone paradigms of \trio and \akuriyo contain an interesting puzzle: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
-%This is illustrated by comparing reconstructible \PTir forms with more conservative \wayana \exref{sat} and \kalina \exref{13w}, which can be taken as representative for \PC.
+This is illustrated by comparing reconstructible \PTir forms with more conservative \wayana \exref{sat} and \kalina \exref{13w}, which can be taken as representative for \PC.
+
+\ex<sat>
+\begin{tabular}[t]{@{}llllll@{}}
+\gl{1}\gl{s_a_} & \PC & \wayana & \PTir & \trio & \akuriyo\\
+\qu{yawn} & \rc{w-e-mɨta-po-} & \obj{w-emtapɨ-} & \rc{t͡ʃ-entapo-} & \obj{s-entapo-} & \obj{t͡ʃ-entapo-}\\
+\qu{bathe} & \rc{w-e-pɨ-} & \obj{w-epɨ-} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-}\\
+\qu{talk} & \rc{w-ət-uru} & \obj{w-əturu-} & \rc{t-əturu-} & \obj{t-əturu-} & \obj{k-əturu-}\footnotemark\\
+\qu{learn} & \rc{w-ət-eme-pa} & \obj{w-əhepa-} & \rc{t-əempa-}\footnotemark & \obj{t-əenpa-} & \obj{k-əempa-}\\
+\end{tabular}\\
+\parencites[28]{camargo2010wayana}[292, 664]{triomeira1999}[77, 86, 82]{gildea1994akuriyo}[50, 208, 213]{wayanatavares2005}
+\xe
+\addtocounter{footnote}{-1}
+\footnotetext{\PTir \rc{t-} was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
+\addtocounter{footnote}{1}
+\footnotetext{
+The \trio form is not due to dissimilation from \rc{mp} to \obj{np}, but rather a phonemic transcription rendering the syllable-final archiphoneme \pnm{N} as \obj{n} \parencite[31]{triomeira1999}.
+Further, \PTar intervocalic \rc{t͡ʃ} was lost entirely in \trio (except after \rc{i}) and is reflected as \obj{ʔ} in \akuriyo \parencite[31--32]{meira1998proto}.
+Given the original shape of the detransitivizers \rc{ət(e)-} and \rc{e-}, \trio \obj{ə-enpa} \qu{to learn} contains a reflex of the former with loss of \rc{t}, so \akuriyo \obj{əempa} would be expected to have a glottal stop: ?\obj{əʔempa}.
+Its absence could either be due to \trio influence or not being transcribed by \textcite{gildea1994akuriyo}.
+Since there is no evidence for a glottal stop in \PTir (\rc{əʔempa}), I do not reconstruct it.}
+
+\ex<13w> 
+\begin{tabular}[t]{@{}llllll@{}}
+\gl{1}>\gl{3}& \PC & \kalina & \PTir & \trio & \akuriyo \\
+\qu{see} & \rc{t-ene-} & \obj{s-ene-} & \rc{w-ene-} & \obj{w-ene-} & \obj{(w-)ene-} \\
+\qu{bring} & \rc{t-enepɨ-} & \obj{s-enepɨ-} & \rc{w-enepɨ-} & \obj{w-enepɨ-} & Ø\obj{-enepɨ-} \\
+\end{tabular}\\
+\parencites[425--426]{hoff1968carib}[28, 304]{triomeira1999}[86--87]{gildea1994akuriyo}
+\xe
 %
-%\ex<sat>
-%\begin{tabular}[t]{@{}llllll@{}}
-%\gl{1}\gl{s_a_} & \PC & \wayana & \PTir & \trio & \akuriyo\\
-%\qu{yawn} & \rc{w-e-mɨta-po-} & \obj{w-emtapɨ-} & \rc{t͡ʃ-entapo-} & \obj{s-entapo-} & \obj{t͡ʃ-entapo-}\\
-%\qu{bathe} & \rc{w-e-pɨ-} & \obj{w-epɨ-} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-}\\
-%\qu{talk} & \rc{w-ət-uru} & \obj{w-əturu-} & \rc{t-əturu-} & \obj{t-əturu-} & \obj{k-əturu-}\footnotemark\\
-%\qu{learn} & \rc{w-ət-eme-pa} & \obj{w-əhepa-} & \rc{t-əempa-}\footnotemark & \obj{t-əenpa-} & \obj{k-əempa-}\\
-%\end{tabular}\\
-%\parencites[28]{camargo2010wayana}[292, 664]{triomeira1999}[77, 86, 82]{gildea1994akuriyo}[50, 208, 213]{wayanatavares2005}
-%\xe
-%\addtocounter{footnote}{-1}
-%\footnotetext{\PTir \rc{t-} was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
-%\addtocounter{footnote}{1}
-%\footnotetext{
-%The \trio form is not due to dissimilation from \rc{mp} to \obj{np}, but rather a phonemic transcription rendering the syllable-final archiphoneme \pnm{N} as \obj{n} \parencite[31]{triomeira1999}.
-%Further, \PTar intervocalic \rc{t͡ʃ} was lost entirely in \trio (except after \rc{i}) and is reflected as \obj{ʔ} in \akuriyo \parencite[31--32]{meira1998proto}.
-%Given the original shape of the detransitivizers \rc{ət(e)-} and \rc{e-}, \trio \obj{ə-enpa} \qu{to learn} contains a reflex of the former with loss of \rc{t}, so \akuriyo \obj{əempa} would be expected to have a glottal stop: ?\obj{əʔempa}.
-%Its absence could either be due to \trio influence or not being transcribed by \textcite{gildea1994akuriyo}.
-%Since there is no evidence for a glottal stop in \PTir (\rc{əʔempa}), I do not reconstruct it.}
-%
-%\ex<13w> 
-%\begin{tabular}[t]{@{}llllll@{}}
-%\gl{1}>\gl{3}& \PC & \kalina & \PTir & \trio & \akuriyo \\
-%\qu{see} & \rc{t-ene-} & \obj{s-ene-} & \rc{w-ene-} & \obj{w-ene-} & \obj{(w-)ene-} \\
-%\qu{bring} & \rc{t-enepɨ-} & \obj{s-enepɨ-} & \rc{w-enepɨ-} & \obj{w-enepɨ-} & Ø\obj{-enepɨ-} \\
-%\end{tabular}\\
-%\parencites[425--426]{hoff1968carib}[28, 304]{triomeira1999}[86--87]{gildea1994akuriyo}
-%\xe
-%%
-%Thus, \wayana preserves the \PC \gl{1}\gl{s_a_} marker \rc{w-}, while it was replaced by \rc{t-} (\phonr{}{\rc{t͡ʃ-}}{\rc{e}}) in \PTir \exref{sat}.
-%%Both \qu{to yawn} and \qu{to bathe} are transparently derived \gl{s_a_} verbs, from \rc{mɨta-po} \qu{to open the mouth of} and \rc{pɨ} \qu{to wash}, respectively.
-%On the other hand, \gl{1}>\gl{3} scenarios show a reflex of \rc{t-} in \kalina, while they took on \rc{w-} in \PTir \exref{13w}.
-%
-The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{In fact, the issue of \emph{when} this happened is also open.
+Thus, \wayana preserves the \PC \gl{1}\gl{s_a_} marker \rc{w-}, while it was replaced by \rc{t-} (\phonr{}{\rc{t͡ʃ-}}{\rc{e}}) in \PTir \exref{sat}.
+%Both \qu{to yawn} and \qu{to bathe} are transparently derived \gl{s_a_} verbs, from \rc{mɨta-po} \qu{to open the mouth of} and \rc{pɨ} \qu{to wash}, respectively.
+On the other hand, \gl{1}>\gl{3} scenarios show a reflex of \rc{t-} in \kalina, while they took on \rc{w-} in \PTir \exref{13w}.
+
+The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}\footnote{In fact, the issue of \emph{when} this happened is also open.
 It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation.
 Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.}
-As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}; we offer our reconstructed \PTir forms.
+The relevant aspect of this extension is of course that not all verbs were affected by it.
+\textcite{meira1998proto} lists the four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}.
 
-\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-} \parencites[112--115]{meira1998proto}[85]{gildea1994akuriyo}\\
+\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-}\\
 \begin{tabular}[t]{@{}llll@{}}
@@
 \qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
-\qu{come} & \rc{w-əʔepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
+\qu{come} & \rc{w-əʔepɨ-/}\obj{w-eepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
 \qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
-\end{tabular}
+\end{tabular}\\
+\parencites[112--115]{meira1998proto}[85]{gildea1994akuriyo}
 %\a<aku12> \akuriyo \obj{kəʔ-eepɨ} \qu{we came} \parencite[114]{meira1998proto}
@@
 
-For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} registered a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
+For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} gives a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
 The two forms can be reconciled by a specific idiosyncratic combination of sound changes:
-We suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
+I suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
 Also, \obj{w-} appears to have been subject to ongoing erosion in \akuriyo, also evidenced by its absence in \qu{to come} and \qu{to be}, but its presence in \qu{to say} \exref{ptirw}.
-This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix \exref{exp.aku-164}, as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
+This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix (\getfullref{13w},  \getfullref{exp.aku-164}), as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
 
@@
 \xe
-%
-For the form of the verb, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for the first person, and \rc{eepɨ} for the others, based on the idiosyncratic form in \trio and the vowel length in \akuriyo.
-\akuriyo (and \carijo) would then have levelled the original distribution of \rc{əepɨ} and \rc{eepɨ}, similar to what we have suggested for Pekodian in \cref{sec:pekodian}.
-We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly cognate with \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ}.
 
-%\ex<tri-130> \trio \parencite[][114]{triomeira1999}\\
-%\begingl
-%\gla n-epɨ//
-%\glb \gl{3}-come//
-%\glft \qu{S/he came.}//
-%\endgl
-%\xe
+For \qu{to come}, a similar situation as in Pekodian is encountered, in that \trio shows a reflex of \rc{ət-epɨ}, while the \akuriyo form points to bare \rc{epɨ}.
+Given that \trio, like \arara, also has forms without a reflex of \rc{ət-} \exref{tri-130}, I reconstruct both \rc{w-epɨ-} and \rc{w-əʔepɨ-} for \PTir.
+
+\ex<tri-130> \trio \parencite[][114]{triomeira1999}\\
+\begingl
+\gla n-epɨ//
+\glb \gl{3}-come//
+\glft \qu{S/he came.}//
+\endgl
+\xe
 
@@
 They are listed in \exref{akumov.aku}.
-We have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
+I have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in that it takes \obj{s-} \exref{akumov.tri}.
 Thus, it is uncertain whether these verbs can be reconstructed as taking \rc{w-} in \PTir, although its presence in \akuriyo does suggest so.
@@
 
-\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
+\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes\\
 \begin{tabular}[t]{@{}lll@{}}
@@
 \qu{defecate} & \obj{k-oeka-} & ?\\
-\end{tabular}
+\end{tabular}\\
+\parencites[294]{triomeira1999}[84]{gildea1994akuriyo}
 \xe
 
-In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} largely replaced \gl{1}\gl{s_a_} \rc{w-}.
+In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} was extended to \gl{1} verbs.
 Four verbs are solidly reconstructible as preserving \rc{w-} in \PTir, and \rc{ɨhtə} \qu{to go down} with irregular \rc{p-} was also not affected.
@@
 After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
-It seems to have replaced \rc{t-} only in specific places, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to us \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
+It seems to have replaced \rc{t-} only in specific places, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to me \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} could be more recent, with which we agree -- since phonologically distributed \rc{t-}/\rc{t͡ʃ-} is reconstructible to \PTir, a scenario whereby \obj{k-} replaces \rc{t-}, but not \rc{t͡ʃ-}, is the most straightforward.
-The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
+He also suggests that \obj{k-} could be more recent, with which I agree -- since phonologically distributed \rc{t-}/\rc{t͡ʃ-} is reconstructible to \PTir, a scenario whereby \obj{k-} replaces \rc{t-}, but not \rc{t͡ʃ-}, is the most straightforward.
+The few \obj{t-} recorded by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
 

### 2021-09-28 documents/sections/verbs.tex

@@
 Verbs for intransitive \qu{to bathe} are usually based on transitive verbs in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in Pemongan, \panare, \kalina, and \maqui\footnote{For \maqui, note that while intransitive \obj{eʔhi} points to \rc{e-kupi}, transitive \obj{ɨhɨ} looks like a reflex of \rc{pɨ}.} \pcref{tab:bathe}.
-As I have shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
-Thus, while the transitive verbs in Pekodian have perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch.
-The reasons for this are unknown; I have not encountered \obj{i-} as a reflex of the detransitivizer in Pekodian, see also \textcite[506]{meira2010origin}.
+As we have shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
+Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch, changing \rc{e-} to \rc{i}.
+The reasons for this are unknown; we are not aware of \rc{i-} as a regular development of the detransitivizer in Pekodian, see also \textcite[506]{meira2010origin}.
 However, it should be noted that other languages also show unexpected developments in this verb, consider the apparent addition of \obj{ew-} in \hixka or the chaotic mixture of \rc{pɨ} and \rc{kupi} in languages spoken in Venezuela.
@@
 \label{sec:movement}
-As discussed in \cref{sec:akuriyo}, four \akuriyo \gl{s_a_} verbs are attested as not having the first person marker \obj{t͡ʃ-} in \textcite{gildea1994akuriyo}; instead, they have \obj{w-} or its phonologically reduced form Ø.
+As discussed in \cref{sec:akuriyo}, four \akuriyo \gl{s_a_} verbs are attested as not having the first person marker \obj{t͡ʃ-} in \textcite{gildea1994akuriyo}; instead, they have \obj{w-} or its phonologically reduced form ∅.
 These verbs share two properties: they are all \obj{e-}initial, and they are all movement verbs.
 However, there are also \obj{e-}initial movement verbs with a first person marker \obj{t͡ʃ-} \exref{aku-106}.
-As for their etymology, I can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
-For the other three verbs, I have not encountered potential cognates, but given their \obj{e-}initial nature and their semantics, it seems plausible that they, too, are derived from transitive verbs with meanings like \qu{to put up} etc.
+As for their etymology, we can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
+For the other three verbs, we have not encountered potential cognates, but given their \obj{e-}initial nature and their semantics, it seems plausible that they, too, are derived from transitive verbs with meanings like \qu{to put up} etc.
 In any case, \akuriyo is only sparsely documented, and there are many open questions about it.

### 2021-09-28 workflow/irregular_paradigms.py
working on overview of etxensions and verbs

@@ -1,7 +1,9 @@
 import pyradigms as pyd
 import pandas as pd
-        
+
+cs_df = pd.read_csv("../data/cognate_sets.csv")
 df = pd.read_csv("../data/irregular_verb_data.csv")
+print(cs_df)
 
 lg_list = ["PWai", "wai", "hix", "PPek", "PXin", "ara", "ikp", "bak", "PTir", "aku", "tri", "car", "yuk"]
 # for i in ["PWai", "PPek", "PXin", "PTir"]:
@@ -11,10 +13,21 @@ pyd.x = ["Cognateset_ID"]
 pyd.y = ["Language_ID"]
 pyd.z = []
 verb_list = ["come-1", "go", "say", "go.down", "be-1", "be-2"]
-pyd.x_sort = lg_list
+pyd.x_sort = verb_list
+pyd.y_sort = lg_list
 pyd.filters = {"Inflection": ["1"], "Cognateset_ID": verb_list, "Language_ID": lg_list}
-# pyd.content_string = "Prefix_Cognateset_ID"
+pyd.content_string = "Prefix_Cognateset_ID"
 table = pyd.compose_paradigm(df)
+table.replace("", "–", inplace=True)
+
+def identify_unaffected(value):
+    if ("1w" in value or "1p" in value):
+        return "y"
+    elif value in ["?", "–"]:
+        return value
+    else:
+        return "n"
+table = table.applymap(identify_unaffected)
 print(table)
 #
 # for verb in verb_list:

### 2021-09-29 documents/sections/verbs.tex
working on overview of etxensions and verbs
work on verbs section, added verb cognates

@@
 \label{sec:come}
-\definecolor{come1}{RGB}{161.0,201.0,244.0}
-\definecolor{come2}{RGB}{255.0,180.0,130.0}
-\definecolor{come3}{RGB}{141.0,229.0,161.0}
-\definecolor{come4}{RGB}{255.0,159.0,155.0}
 \begin{table}
 	\centering
-	\caption{Reflexes of \qu{to come}}
+	\caption{Reflexes of \qu{to come} \parencites[150]{alves2017arara}[265]{ikpengpacheco2001}[4]{meira2003bakairi}[294]{triomeira1999}[168]{meira1998proto}[178]{robayo2000avance}[37]{koehn1986apalai}[429]{courtz2008carib}[438]{maquiritaricaceres2011}[160]{stegeman2014akawaio}[125]{akawaiocaesar2003}[299, 415]{cruz2005fonologia}[102]{alvarez2000construcciones}[32]{macushiabbott1991}[65]{panarepayne2013}[68]{mendez1959yawarana}[74]{muller1975mapoyo}[182; p.c., Spike Gildea]{meira2005southern}}
     \label{tab:come}
-	\begin{tabular}{@{}lll@{}}
-	\mytoprule
-	Language & Form & Source\\
-	\mymidrule
-\kaxui & \colorbox{come3}{\obj{ehɨ}} & \perscomm{Spike Gildea}\\
-\kaxui & \colorbox{come1}{\obj{johɨ}} & \perscomm{Spike Gildea}\\
-\kaxui & \colorbox{come2}{\obj{oohɨ}} & \perscomm{Spike Gildea}\\
-\arara & \colorbox{come3}{\obj{ebɨ}} & \textcite[150]{alves2017arara}\\
-\arara & \colorbox{come4}{\obj{odebɨ}} & \textcite[150]{alves2017arara}\\
-\ikpeng & \colorbox{come4}{\obj{arep}} & \textcite[265]{ikpengpacheco2001}\\
-\bakairi & \colorbox{come4}{\obj{əewɨ}} & \textcite[4]{meira2003bakairi}\\
-\trio & \colorbox{come4}{\obj{əepɨ}} & \textcite[294]{triomeira1999}\\
-\trio & \colorbox{come3}{\obj{epɨ}} & \textcite[294]{triomeira1999}\\
-\akuriyo & \colorbox{come3}{\obj{epɨ}} & \textcite[168]{meira1998proto}\\
-\carijo & \colorbox{come3}{\obj{ehɨ}} & \textcite[178]{robayo2000avance}\\
-\apalai & \colorbox{come4}{\obj{oepɨ}} & \textcite[37]{koehn1986apalai}\\
-\kalina & \colorbox{come2}{\obj{opɨ}} & \textcite[429]{courtz2008carib}\\
-\maqui & \colorbox{come3}{\obj{ehə}} & \textcite[438]{maquiritaricaceres2011}\\
-\akawaio & \colorbox{come4}{\obj{əsipɨ}} & \textcite[160]{stegeman2014akawaio}\\
-\akawaio & \colorbox{come1}{\obj{jepɨ}} & \textcite[125]{akawaiocaesar2003}\\
-\ingariko & \colorbox{come1}{\obj{jepə}} & \textcite[415]{cruz2005fonologia}\\
-\ingariko & \colorbox{come1}{\obj{jə}} & \textcite[299]{cruz2005fonologia}\\
-\patamona & \colorbox{come1}{\obj{jepɨ}} & \perscomm{Spike Gildea}\\
-\patamona & \colorbox{come1}{\obj{jəpɨ}} & \perscomm{Spike Gildea}\\
-\pemon & \colorbox{come1}{\obj{jepɨ}} & \textcite[102]{alvarez2000construcciones}\\
-\macushi & \colorbox{come1}{\obj{ipɨ}} & \textcite[32]{macushiabbott1991}\\
-\panare & \colorbox{come2}{\obj{əpɨ}} & \textcite[65]{panarepayne2013}\\
-\yawarana & \colorbox{come2}{\obj{əpɨ}} & \textcite[68]{mendez1959yawarana}\\
-\mapoyo & \colorbox{come3}{\obj{epɨ}} & \textcite[74]{muller1975mapoyo}\\
-\uxc & \colorbox{come2}{\obj{ee}} & \textcite[182]{meira2005southern}\\
-\mybottomrule
+\begin{tabular}[t]{@{}ll@{}}
+\toprule
+Language &                               Form \\
+\midrule
+\kaxui    &  \obj{johɨ}, \obj{ehɨ}, \obj{oohɨ} \\
+\arara    &             \obj{ebɨ}, \obj{odebɨ} \\
+\ikpeng   &                         \obj{arep} \\
+\bakairi  &                         \obj{əewɨ} \\
+\trio     &              \obj{epɨ}, \obj{əepɨ} \\
+\akuriyo  &                         \obj{eepɨ} \\
+\carijo   &                          \obj{ehɨ} \\
+\apalai   &                         \obj{oepɨ} \\
+\kalina   &                          \obj{opɨ} \\
+\maqui    &                          \obj{ehə} \\
+\akawaio  &            \obj{jepɨ}, \obj{əsipɨ} \\
+\ingariko &               \obj{jepə}, \obj{jə} \\
+\patamona &             \obj{jəpɨ}, \obj{jepɨ} \\
+\pemon    &                         \obj{jepɨ} \\
+\macushi  &                          \obj{ipɨ} \\
+\panare   &                          \obj{əpɨ} \\
+\yawarana &                          \obj{əpɨ} \\
+\mapoyo   &                          \obj{epɨ} \\
+\uxc      &                           \obj{ee} \\
+\bottomrule
 \end{tabular}
 \end{table}
+
 This verb is reconstructed as monomorphemic \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a more complex story.
-In particular, the Pemongan languages and one of its allomorphs in \kaxui suggest a root \colorbox{come1}{\rc{jəpɨ}}.
+Different languages present rather different forms, which we argue are all (partially) cognate, with the reconstructed \PC verb having the form \rc{(ət-)jəpɨ}.
+Reflexes of a bare \rc{j-}initial root \rc{jəpɨ} are found in the Pemongan languages and \kaxui.
 However, the one form in \kaxui only occurs in the Progressive \exref{kaxprog.kax-82}, and \obj{j(-)} may be a reflex the \settwo third person marker \rc{i-} combined with  the \gl{s_a_} class marker \rc{w-}, as suggested by the occurrence of the latter with other person values \exref{kaxprog.kax-54}.
 On the other hand, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
-Thus, it seems more likely that the \obj{j} is indeed part of the verb, rather than an outcome of \rc{i-w-}.
+Thus, it seems more likely that the \obj{j} is indeed part of the root, rather than an outcome of \rc{i-w-}.
 
@@
 
-There is further evidence supporting the reconstruction of both \rc{j} and \rc{ə}.
-First, there are many forms reflecting \colorbox{come2}{\rc{əpɨ}}, occurring in different branches of the family.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
-Second, there are also forms reflecting \colorbox{come3}{\rc{epɨ}}, and umlaut of \rc{ə} to \rc{e} is conditioned by preceding \rc{j} \parencite{meira2010origin}.
-That is, a unifying account of these forms starts from a root \rc{jəpɨ}, which is the subject to two major sound changes: \begin{inlinelist}
+Other languages -- across the family -- have forms reflecting the same root without the initial consonant, \rc{əpɨ}: \kalina, \panare, \yawarana, and \uxc.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
+There are also forms reflecting \rc{epɨ}, again across the family: \kaxui, \arara, \trio, \carijo, \maqui, and \mapoyo.
+The \akuriyo form \obj{eepɨ} looks like a reflex thereof as well, but the length is unexpected, and is analyzed by \textcite[]{meira1998proto} as reflecting an earlier diphthong \rc{əe}.
+A unifying account of these forms starts from a root \rc{jəpɨ}, which is the subject to two major sound changes: \begin{inlinelist}
  \item \rc{j}-loss
- \item \rc{ə}-umlaut	
+ \item \rc{ə}-umlaut after \rc{j}
  \end{inlinelist}.
+Both phenomena are widespread in the family \parencite{meira2010origin}.
 However, these sound changes appear to have applied irregularly, and not always in the same order.
@@
 
-we have so far ignored the forms which form the basis for \posscite{gildea2007greenberg} \rc{ətepɨ}; these are compatible with our reconstructed root \rc{jəpɨ}, with the addition of the detransitivizer \rc{ət(e)-}: \colorbox{come4}{\rc{ətjəpɨ}}.
-In fact, the \obj{i} in the \akawaio form \obj{əsipɨ} is better explained by a sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of \rc{jəpɨ}, \obj{ipɨ}.
-While the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
+We have so far ignored the forms which form the basis for \posscite{gildea2007greenberg} \rc{ətepɨ}; these are compatible with our reconstructed root \rc{jəpɨ}, with the addition of the detransitivizer \rc{ət(e)-}: \rc{ətjəpɨ}.
+While most of the modern forms are lacking evidence for \rc{j} and thus look like a reflex of \rc{ətepɨ}, the \obj{i} in the \akawaio form \obj{əsipɨ} is evidence of the sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of \rc{jəpɨ}, \obj{ipɨ}.
+Although the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
 For example, \gl{s_p_} \rc{wənɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
@@
 
-While these sound changes and the addition of \rc{ət(e)-} do account for the majority of the forms in \cref{tab:come},\footnote{The only exception is the \apalai form \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
+While these sound changes and the addition of \rc{ət(e)-} do account for the majority of the forms in \cref{tab:come},\footnote{
+Apart from aforementioned \akuriyo \obj{eepɨ}, another exception is the \apalai form \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
 While \obj{oepɨ} would be a regular outcome of \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-}.
@@
 This was discussed in \cref{sec:pekodian} for \arara, which has reflexes of \rc{ətjəpɨ} and \rc{jəpɨ} within the same paradigm.
-A similar situation is found in \trio, where the \setone paradigm shows a reflex of \rc{ətjəpɨ} for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}
+A similar situation is found in \trio, where the \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}
 
@@
 \begin{tabular}[t]{@{}ll@{}}
-\gl{1} & \obj{w-\colorbox{come4}{əepɨ}} \\
-\gl{2} &  \obj{mən-\colorbox{come3}{epɨ}} \\ 
-\gl{1+2} &  \obj{ke-\colorbox{come3}{epɨ}} \\
-\gl{3} &  \obj{n-\colorbox{come3}{epɨ}} \\
+\gl{1} & \obj{w-əepɨ} \\
+\gl{2} &  \obj{mən-epɨ} \\ 
+\gl{1+2} &  \obj{ke-epɨ} \\
+\gl{3} &  \obj{n-epɨ} \\
 \end{tabular}
@@
 As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \arara reflex of \rc{jəpɨ}, but not on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}.
-Both reflexes of \rc{ətjəpɨ} (\trio) and of \rc{jəpɨ} (\akuriyo) resisted the introduction of \rc{t-} in \PTir.
+Both reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ} and/or \rc{epɨ} (\akuriyo) resisted the introduction of \rc{t-} in \PTir.
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
@@
 	\mybottomrule
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\centering
+	\caption{Reflexes of \rc{ɨpɨtə} \qu{to go down} \parencites[294, 325, 331]{triomeira1999}[196]{hixkaryanaderby1979}[55]{waiwaihawkins1998}[118]{guerrero2019carijo}[44]{camargo2010wayana}[99]{camargo2002lexico}[263]{courtz2008carib}[450]{maquiritaricaceres2011}[139]{stegeman2014akawaio}[139]{alvarez2008clausulas}[34]{macushiabbott1991}[88]{mattei1994diccionario}[68]{mendez1959yawarana}[58]{bruno1996dictionary}[84]{gildea1994akuriyo}[153]{alves2017arara}[137]{von1892bakairi}[116; p.c., Angela Fabíola Alves Chagas, David Felipe Guerrero, Spike Gildea]{meira1998proto}}
+	\label{tab:XXX}
+\begin{tabular}[t]{@{}lll@{}}
+\toprule
+Language &            Form &      Class \\
+\midrule
+\PPar     &     \obj{ɨɸɨto} &  \gl{s_p_} \\
+\kaxui    &     \obj{ɨhɨto} &  \gl{s_p_} \\
+\hixka    &       \obj{hto} &          ? \\
+\waiwai   &       \obj{hto} &          – \\
+\PPek     &      \obj{ɨptə} &  \gl{s_a_} \\
+\arara    &     \obj{iptoŋ} &  \gl{s_a_} \\
+\ikpeng   &     \obj{iptoŋ} &          ? \\
+\bakairi  &     \obj{ɨtəgɨ} &  \gl{s_a_} \\
+\PTir     &      \obj{ɨhtə} &  \gl{s_a_} \\
+\trio     &      \obj{ɨhtə} &  \gl{s_a_} \\
+\akuriyo  &       \obj{ɨtə} &  \gl{s_a_} \\
+\carijo   &     \obj{ehɨtə} &          – \\
+\wayana   &      \obj{ɨptə} &  \gl{s_p_} \\
+\apalai   &      \obj{ɨhto} &  \gl{s_p_} \\
+\kalina   &  (\obj{onɨʔto}) &  \gl{s_a_} \\
+\maqui    &      \obj{əʔtə} &  \gl{s_p_} \\
+\kapon    &    (\obj{uʔtə}) &          – \\
+\pemon    &    (\obj{uʔtə}) &          – \\
+\macushi  &    (\obj{autɨ}) &          – \\
+\panare   &      \obj{əhtə} &  \gl{s_a_} \\
+\yawarana &      \obj{əhtə} &          – \\
+\waimiri  &       \obj{ɨtɨ} &          – \\
+\bottomrule
 \end{tabular}

### 2021-10-02 documents/floats/defecate.tex
added 'bathe' data, rendering tables

\begin{table}
\centering
\caption{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
\label{tab:defecate}
\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &           Form &      Class \\
\midrule
\kaxui   &     \obj{weka} &  \gl{s_p_} \\
\arara   &    \obj{watke} &  \gl{s_p_} \\
\ikpeng  &     \obj{atke} &  \gl{s_p_} \\
\bakairi &     \obj{əeke} &  \gl{s_a_} \\
\trio    &     \obj{weka} &  \gl{s_a_} \\
\wayana  &     \obj{uika} &  \gl{s_p_} \\
\apalai  &     \obj{weka} &          ? \\
\kalina  &    \obj{uweka} &  \gl{s_p_} \\
\maqui   &     \obj{weka} &  \gl{s_p_} \\
\panare  &  \obj{(a)iʔka} &          ? \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-02 documents/sections/verbs.tex
added 'bathe' data, rendering tables

@@
 \label{sec:come}
-\begin{table}
-	\centering
-	\caption{Reflexes of \qu{to come} \parencites[150]{alves2017arara}[265]{ikpengpacheco2001}[4]{meira2003bakairi}[294]{triomeira1999}[168]{meira1998proto}[178]{robayo2000avance}[37]{koehn1986apalai}[429]{courtz2008carib}[438]{maquiritaricaceres2011}[160]{stegeman2014akawaio}[125]{akawaiocaesar2003}[299, 415]{cruz2005fonologia}[102]{alvarez2000construcciones}[32]{macushiabbott1991}[65]{panarepayne2013}[68]{mendez1959yawarana}[74]{muller1975mapoyo}[182; p.c., Spike Gildea]{meira2005southern}}
-    \label{tab:come}
-\begin{tabular}[t]{@{}ll@{}}
-\toprule
-Language &                               Form \\
-\midrule
-\kaxui    &  \obj{johɨ}, \obj{ehɨ}, \obj{oohɨ} \\
-\arara    &             \obj{ebɨ}, \obj{odebɨ} \\
-\ikpeng   &                         \obj{arep} \\
-\bakairi  &                         \obj{əewɨ} \\
-\trio     &              \obj{epɨ}, \obj{əepɨ} \\
-\akuriyo  &                         \obj{eepɨ} \\
-\carijo   &                          \obj{ehɨ} \\
-\apalai   &                         \obj{oepɨ} \\
-\kalina   &                          \obj{opɨ} \\
-\maqui    &                          \obj{ehə} \\
-\akawaio  &            \obj{jepɨ}, \obj{əsipɨ} \\
-\ingariko &               \obj{jepə}, \obj{jə} \\
-\patamona &             \obj{jəpɨ}, \obj{jepɨ} \\
-\pemon    &                         \obj{jepɨ} \\
-\macushi  &                          \obj{ipɨ} \\
-\panare   &                          \obj{əpɨ} \\
-\yawarana &                          \obj{əpɨ} \\
-\mapoyo   &                          \obj{epɨ} \\
-\uxc      &                           \obj{ee} \\
-\bottomrule
-\end{tabular}
-\end{table}
+
+\input{floats/come}
 
@@
 
-\begin{table}
-	\centering
-	\caption{Reflexes of \rc{ɨpɨtə} \qu{to go down}}
-	\label{tab:descend_cog}
-	\begin{tabular}{@{}llll@{}}
-	\mytoprule
-Language & Verb & Class & Source \\
-\mymidrule
-\kaxui & \obj{ɨhɨto} & \gl{s_p_} & \perscomm{Spike Gildea} \\
-\hixka & \obj{hto} & ? & \textcite[196]{hixkaryanaderby1979}\\
-\waiwai & \obj{hto} & -- & \textcite[55]{waiwaihawkins1998}\\
-\PPek & \rc{ɨptə} & \gl{s_a_} & \cref{sec:pekodian}\\
-\PTir & \rc{ɨhtə}  (\gl{1} \rc{p-})& \gl{s_a_} & \cref{sec:taranoan} \\
-\carijo & \obj{ehɨtə} & -- & \textcite[118]{guerrero2019carijo}\\
-\wayana & \obj{ɨptə} & \gl{s_a_}/\gl{s_p_} & \textcite[44]{camargo2010wayana}\\
-\apalai & \obj{ɨhto} & \gl{s_p_} & \textcite[99]{camargo2002lexico} \\
-\kalina & (\obj{onɨʔto}) & \gl{s_a_} & \textcite[263]{courtz2008carib}\\
-\maqui & \obj{əʔtə} & \gl{s_p_} & \textcite[450]{maquiritaricaceres2011}\\
-\kapon & (\obj{uʔtə}) & -- & \textcite[139]{stegeman2014akawaio}\\
-\pemon & (\obj{uʔtə}) & -- & \textcite[139]{alvarez2008clausulas}\\
-\macushi & (\obj{autɨ}) & -- & \textcite[34]{macushiabbott1991}\\
-\panare & \obj{əhtə} & \gl{s_a_} & \textcite[88]{mattei1994diccionario}\\
-\yawarana & \obj{əhtə} & -- & \textcite[68]{mendez1959yawarana}\\
-\waimiri & \obj{ɨtɨ} & -- & \textcite[58]{bruno1996dictionary}\\
-	\mybottomrule
-	\end{tabular}
-\end{table}
+\input{floats/godown}
 
-\begin{table}
-	\centering
-	\caption{Reflexes of \rc{ɨpɨtə} \qu{to go down} \parencites[294, 325, 331]{triomeira1999}[196]{hixkaryanaderby1979}[55]{waiwaihawkins1998}[118]{guerrero2019carijo}[44]{camargo2010wayana}[99]{camargo2002lexico}[263]{courtz2008carib}[450]{maquiritaricaceres2011}[139]{stegeman2014akawaio}[139]{alvarez2008clausulas}[34]{macushiabbott1991}[88]{mattei1994diccionario}[68]{mendez1959yawarana}[58]{bruno1996dictionary}[84]{gildea1994akuriyo}[153]{alves2017arara}[137]{von1892bakairi}[116; p.c., Angela Fabíola Alves Chagas, David Felipe Guerrero, Spike Gildea]{meira1998proto}}
-	\label{tab:XXX}
-\begin{tabular}[t]{@{}lll@{}}
-\toprule
-Language &            Form &      Class \\
-\midrule
-\PPar     &     \obj{ɨɸɨto} &  \gl{s_p_} \\
-\kaxui    &     \obj{ɨhɨto} &  \gl{s_p_} \\
-\hixka    &       \obj{hto} &          ? \\
-\waiwai   &       \obj{hto} &          – \\
-\PPek     &      \obj{ɨptə} &  \gl{s_a_} \\
-\arara    &     \obj{iptoŋ} &  \gl{s_a_} \\
-\ikpeng   &     \obj{iptoŋ} &          ? \\
-\bakairi  &     \obj{ɨtəgɨ} &  \gl{s_a_} \\
-\PTir     &      \obj{ɨhtə} &  \gl{s_a_} \\
-\trio     &      \obj{ɨhtə} &  \gl{s_a_} \\
-\akuriyo  &       \obj{ɨtə} &  \gl{s_a_} \\
-\carijo   &     \obj{ehɨtə} &          – \\
-\wayana   &      \obj{ɨptə} &  \gl{s_p_} \\
-\apalai   &      \obj{ɨhto} &  \gl{s_p_} \\
-\kalina   &  (\obj{onɨʔto}) &  \gl{s_a_} \\
-\maqui    &      \obj{əʔtə} &  \gl{s_p_} \\
-\kapon    &    (\obj{uʔtə}) &          – \\
-\pemon    &    (\obj{uʔtə}) &          – \\
-\macushi  &    (\obj{autɨ}) &          – \\
-\panare   &      \obj{əhtə} &  \gl{s_a_} \\
-\yawarana &      \obj{əhtə} &          – \\
-\waimiri  &       \obj{ɨtɨ} &          – \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-A comparative view \pcref{tab:descend_cog} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
+A comparative view \pcref{tab:godown} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
 The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
@@
 
-\begin{table}
-	\centering
-	\caption{Intransitive and transitive \qu{to bathe}}
-    \label{tab:bathe}
-	\begin{tabular}{@{}llll@{}}
-	\mytoprule
-	Language & \gl{intr} & \gl{tr} & Source\\
-	\mymidrule
-\kaxui & \obj{eehɨ} & \obj{ɨhɨ} & \perscomm{Spike Gildea}\\
-\hixka & \obj{ewehɨ} & \obj{ɨhɨ} & \textcites[198]{hixkaryanaderby1979}\\
-\waiwai & \obj{ejepu} & \obj{pɨ} & \textcites[166, 192]{waiwaihawkins1998}\\
-\arara & \obj{ibɨ} & \obj{ɨp} & \textcites[150, 162]{alves2017arara}\\
-\ikpeng & \obj{ip} & \obj{ɨp} & \textcites[103]{ikpengpacheco1997}[123]{campetela1997analise}\\
-\bakairi & \obj{i} & \obj{ɨ} & \textcites[4]{meira2003bakairi}[285]{meira2005bakairi}\\
-\trio & \obj{epɨ} & \obj{pɨ} & \textcites[697]{triomeira1999}\\
-\akuriyo & \obj{epɨ} & \obj{pɨ} & \textcites[87]{gildea1994akuriyo}\\
-\wayana & \obj{epɨ} & \obj{upɨ} & \textcites[24, 52]{camargo2010wayana}\\
-\apalai & \obj{epɨ} & \obj{pɨ} & \textcites[218]{meira2000split}\\
-\kalina & \obj{ekupi} & \obj{kupi} & \textcites[304]{courtz2008carib}\\
-\maqui & \obj{eʔhi} & \obj{ɨhɨ} & \textcites[439, 454]{maquiritaricaceres2011}\\
-\kapon & \obj{ekuʔpi} & \obj{kuʔpi} & \textcites[37]{stegeman2014akawaio}\\
-\pemon & \obj{ekupɨ} & \obj{pɨ} & \textcites[34, 129]{pemondearmellada1944dic}\\
-\panare & \obj{akupɨ} & \obj{ɨpɨ}/\obj{kupɨ} & \textcites[8, 294]{mattei1994diccionario}\\
-\mybottomrule
-	\end{tabular}
-\end{table}
+\input{floats/bathe}
 
@@
 
+\input{floats/defecate}
+
 \subsection{The \akuriyo movement verbs}

### 2021-10-05 documents/floats/defecate.tex
wrapped up verbs section for now

@@
 \centering
-\caption{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
+\caption{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
 \label{tab:defecate}
@@
 \trio    &     \obj{weka} &              \gl{s_a_} \\
-\wayana  &     \obj{uika} &  \gl{s_p_} \\
+\wayana  &     \obj{uika} &  \gl{s_a_} / \gl{s_p_} \\
 \apalai  &     \obj{weka} &                      ? \\

### 2021-10-05 documents/sections/verbs.tex
wrapped up verbs section for now

@@
 
-%\begin{table}
-%	\centering
-%	\caption{XXX}
-%	\label{tab:XXX}
-%	\begin{tabular}{@{}llllllllllllll@{}}
-%	\mytoprule
 %& \PPeks \rc{k-} & \acs{ara_lg} & \acs{ikp_lg} & \acs{bak_lg} & \PWais \rc{k-} & \acs{wai_lg} & \acs{hix_lg} & \PTirs \rc{t-} & \acs{tri_lg} & \acs{aku_lg} & \acs{aku_lg} \obj{k-} & \acs{car_lg} \obj{j(i)-} & \acs{yuk_lg} \obj{j-} \\
@@
 Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
-\akuriyo \obj{eʔi} was not a potential target of innovative \obj{k-} \pcref{sec:akuriyo}, due to the verb being \obj{e}-initial.
+\akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while \obj{eʔi} is not attested in a first-person form. 
 \carijo innovated \obj{j-}, but only in the \obj{et͡ʃi} root allomorph \exref{carcop.mivida-12}; the \obj{a} root preserves \obj{w-} \exref{carcop.car-25}.
@@
 
-This verb was not affected by the extensions found in \PPek, \PWai, \PTir, \carijo, and \yukpa.
-It was not a potential target of \akuriyo \obj{k-}.
+This verb was not affected by the any of the extensions discussed in \cref{sec:extensions}.
 
@@
 
-This verb is reconstructed as monomorphemic \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a more complex story.
-Different languages present rather different forms, which we argue are all (partially) cognate, with the reconstructed \PC verb having the form \rc{(ət-)jəpɨ}.
+This verb is reconstructed as monomorphemic \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
+Different languages present rather different forms, which we argue are all (partially) cognate, with the reconstructed \PC verb having the form \rc{(ət-)jəpɨ}, with an optional detransitivizer.
 Reflexes of a bare \rc{j-}initial root \rc{jəpɨ} are found in the Pemongan languages and \kaxui.
@@
 
-Other languages -- across the family -- have forms reflecting the same root without the initial consonant, \rc{əpɨ}: \kalina, \panare, \yawarana, and \uxc.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
-There are also forms reflecting \rc{epɨ}, again across the family: \kaxui, \arara, \trio, \carijo, \maqui, and \mapoyo.
+Other languages -- across the family -- have forms reflecting the same root without the initial consonant, i.e.\ \rc{əpɨ}: \kalina, \panare, \yawarana, and \uxc.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
+Also found across the family are forms reflecting \rc{epɨ}, in \kaxui, \arara, \trio, \carijo, \maqui, and \mapoyo.
 The \akuriyo form \obj{eepɨ} looks like a reflex thereof as well, but the length is unexpected, and is analyzed by \textcite[]{meira1998proto} as reflecting an earlier diphthong \rc{əe}.
@@
 We have so far ignored the forms which form the basis for \posscite{gildea2007greenberg} \rc{ətepɨ}; these are compatible with our reconstructed root \rc{jəpɨ}, with the addition of the detransitivizer \rc{ət(e)-}: \rc{ətjəpɨ}.
-While most of the modern forms are lacking evidence for \rc{j} and thus look like a reflex of \rc{ətepɨ}, the \obj{i} in the \akawaio form \obj{əsipɨ} is evidence of the sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of \rc{jəpɨ}, \obj{ipɨ}.
+While most of the modern forms with a reflex of the detransitivizer are lacking evidence for \rc{j} (\rc{ətepɨ}), the \obj{i} in the \akawaio form \obj{əsipɨ} is evidence of the sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of bare \rc{jəpɨ}, \obj{ipɨ}.
 Although the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
@@
 \waiwai \qu{go to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced by its status in \kaxui, \kalina, \panare \exref{pan-128}, \arara, \trio, and \akuriyo.
+The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced by its status in \kaxui, \kalina, \panare \exref{pan-128}, \arara, and \trio.
 
@@
 Apart from aforementioned \akuriyo \obj{eepɨ}, another exception is the \apalai form \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
-While \obj{oepɨ} would be a regular outcome of \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-}.
+While \obj{oepɨ} would be a regular outcome of \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
 The form may be due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}.
-Alternatively, \apalai \obj{oepɨ} could be a fossilized loan \wayana, which has lost its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
+Alternatively, \apalai \obj{oepɨ} could be a fossilized loan from \wayana, which has replaced its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
 the distribution within the family is rather chaotic.
@@
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
-It is unknown whether there is a \yukpa reflex of this verb, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come}.
+It is unknown whether there is a \yukpa reflex of this verb, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come} and was thus not a potential target of innovative \rc{k-}.
 
@@
 %
-The situation in \yukpa is unclear, as it is an open question whether \obj{ew(uh)tu} \qu{to go down} \parencite{meira2003primeras} is cognate.
+While it seems that \yukpa \obj{ewu(h)tu} \qu{to go down} \parencite{meira2003primeras} is cognate, it was affected by the extension of \obj{j-}.
+%todo maybe insert a yukpa example here? the form is just jewuhtu
 
@@
 
-A comparative view \pcref{tab:godown} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
+A comparative view \pcref{tab:godown} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
 The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
-In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb.
+In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb in the synchronic analysis of \wayana.
 It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
@@
 Its causativized form is \obj{ɨptəka} \parencite[255]{wayanatavares2005}, containing a reflex of the transitivizer \rc{-ka}, which was restricted to \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
+This leads us to posit the hypothesis that the verb may have been a normal member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class -- fully so in the languages where it is \gl{s_a_}.
 
-Further comparative evidence suggests that \rc{ɨpɨtə} \qu{to go down} was originally an \gl{s_p_} verb which switched its class to \gl{s_a_} in some languages, incompletely so in \wayana.
+\wayana-external comparative evidence supports this hypothesis:
 The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:descend_cog} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
@@
 \label{sec:bathe}
-This verb only emerged as resisting an extension in the case of \PPek \pcref{sec:pekodian}.
+This verb only emerged as resisting an extension in the case of \PPek \pcref{sec:pekodian}, but is widely attested in the family.
 Verbs for intransitive \qu{to bathe} are usually based on transitive verbs in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in Pemongan, \panare, \kalina, and \maqui\footnote{For \maqui, note that while intransitive \obj{eʔhi} points to \rc{e-kupi}, transitive \obj{ɨhɨ} looks like a reflex of \rc{pɨ}.} \pcref{tab:bathe}.
@@
 \label{sec:shit}
-As in the case of \qu{to go down} \pcref{sec:godown}, \rc{weka} \qu{to defecate} was likely originally \gl{s_p_}, with most languages showing an \gl{s_p_} reflex \pcref{tab:shit_cog}.
-Another parallel to \qu{to go down} is the mixed \gl{s_a_}/\gl{s_p_} status in \wayana.
+The comparative picture for the resistant \trio \gl{s_a_} verb \qu{to defecate} \pcref{tab:defecate} shows that \rc{weka} was clearly originally an \gl{s_p_} verb.
+As in the case of \qu{to go down} \pcref{sec:godown}, this verb has a mixed \gl{s_a_}/\gl{s_p_} status in \wayana.
 The \bakairi reflex \obj{əeke} seems to contain a reflex of \rc{ət-}, which would explain its status as an \gl{s_a_} verb.
-While the class memberships of \panare \obj{aiʔka} and \obj{iʔka} \qu{to defecate} are both unknown, the first form seems to contain a reflex of the detransitivizer, but not the second form.
-
-\begin{table}
-	\centering
-	\caption{\rc{weka} \qu{to defecate} as a class-switching \gl{s_p_} verb}
-	\label{tab:shit_cog}
-	\begin{tabular}{@{}llll@{}}
-	\mytoprule
-Language & Verb & Class & Source \\
-\mymidrule
-\kalina & \obj{uweka} & \gl{s_p_} & \textcites[418]{courtz2008carib} \\
-\maqui & \obj{weka} & \gl{s_p_} & \textcites[455]{maquiritaricaceres2011} \\
-\kaxui & \obj{weka} & \gl{s_p_} & \perscomm{Spike Gildea} \\
-\arara & \obj{watke} & \gl{s_p_} & \textcite[44]{souza1993arara}\\
-\ikpeng & \obj{atke} & \gl{s_p_} & \textcite[118]{alves2013verbo}\\
-\wayana & \obj{uika} & \gl{s_a_}/\gl{s_p_} & \textcite[86, 206]{wayanatavares2005}\\
-\trio & \obj{weka} (\gl{1} \obj{koeka}) & \gl{s_a_}  & \textcites[294]{triomeira1999} \\
-\bakairi & \obj{əeke} & \gl{s_a_} & \parencite{meira2005bakairi} \\
-\apalai & \obj{weka} & ? & \textcites[96]{camargo2002lexico} \\
-\panare & \obj{(a)iʔka} & ? & \textcite[319]{mattei1994diccionario}\\
-	\mybottomrule
-	\end{tabular}
-\end{table}
+While the class membership of \panare \obj{aiʔka} and \obj{iʔka} \qu{to defecate} is unknown, the former seems to contain a reflex of the detransitivizer, while the latter only consists of the root, meaning that one would assume \gl{s_a_} and \gl{s_p_} membership, respectively.
 
@@
 These verbs share two properties: they are all \obj{e-}initial, and they are all movement verbs.
-However, there are also \obj{e-}initial movement verbs with a first person marker \obj{t͡ʃ-} \exref{aku-106}.
-As for their etymology, we can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
+However, the same dataset also contains \obj{e-}initial movement verbs with a first person marker \obj{t͡ʃ-} \exref{aku-106}.
+As for the etymology of these four verbs, we can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
 For the other three verbs, we have not encountered potential cognates, but given their \obj{e-}initial nature and their semantics, it seems plausible that they, too, are derived from transitive verbs with meanings like \qu{to put up} etc.
-In any case, \akuriyo is only sparsely documented, and there are many open questions about it.
+In any case, \akuriyo is only sparsely documented, and there are many open questions about it; the \akuriyo data in \textcite{gildea1994akuriyo} are not fully reliable.
 

### 2021-10-11 documents/floats/defecate.tex
cleaned up data, started work on discussion section

@@
 \centering
-\caption{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
+\caption[\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb]{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
 \label{tab:defecate}
@@
 \midrule
+\apalai  &     \obj{weka} &                      ? \\
+\arara   &    \obj{watke} &              \gl{s_p_} \\
+\bakairi &     \obj{əeke} &              \gl{s_a_} \\
+\ikpeng  &     \obj{atke} &              \gl{s_p_} \\
+\kalina  &    \obj{uweka} &              \gl{s_p_} \\
 \kaxui   &     \obj{weka} &              \gl{s_p_} \\
-\arara   &    \obj{watke} &              \gl{s_p_} \\
-\ikpeng  &     \obj{atke} &              \gl{s_p_} \\
-\bakairi &     \obj{əeke} &              \gl{s_a_} \\
+\maqui   &     \obj{weka} &              \gl{s_p_} \\
+\panare  &  \obj{(a)iʔka} &                      ? \\
 \trio    &     \obj{weka} &              \gl{s_a_} \\
 \wayana  &     \obj{uika} &  \gl{s_a_} / \gl{s_p_} \\
-\apalai  &     \obj{weka} &                      ? \\
-\kalina  &    \obj{uweka} &              \gl{s_p_} \\
-\maqui   &     \obj{weka} &              \gl{s_p_} \\
-\panare  &  \obj{(a)iʔka} &                      ? \\
 \bottomrule

### 2021-10-11 documents/floats/imperatives.tex
cleaned up data, started work on discussion section

@@
+\begin{table}
+\centering
+\caption[Imperatives of \gl{s_a_} and \gl{s_p_} verbs]{Imperatives of \gl{s_a_} and \gl{s_p_} verbs \parencites[62]{koehn1986apalai}[Mopo/20]{koehns1994textos}[44, 89]{derbyshire1965textos}[5, 17]{mattei1994diccionario}[190]{hoff1968carib}[161]{alves2017arara}[323]{triomeira1999}[227]{wayanatavares2005}}
+\label{tab:imperatives}
 \begin{tabular}[t]{@{}lll@{}}
@@
 \midrule
+\apalai &           \obj{otuʔ-ko} \qu{eat!} &      \obj{\emp{o-}nɨʔ-ko} \qu{sleep!} \\
 \hixka  &          \obj{omoh-ko} \qu{come!} &  \obj{\emp{oj-}okajɨm-ko} \qu{go up!} \\
+\panare &            \obj{ape-ʔ} \qu{flee!} &             \obj{ahpən-kə} \qu{jump!} \\
+\kalina &         \obj{oʔmaʔ-ko} \qu{stop!} &   \obj{\emp{aj-}awon-ko} \qu{get up!} \\
 \arara  &  \obj{odotpot-ko} \qu{come back!} &      \obj{\emp{o-}alum-ko} \qu{jump!} \\
@@
 \wayana &         \obj{əməm-kə} \qu{enter!} &    \obj{\emp{əw-}eremi-kə} \qu{sing!} \\
-\apalai &           \obj{otuʔ-ko} \qu{eat!} &      \obj{\emp{o-}nɨʔ-ko} \qu{sleep!} \\
-\kalina &         \obj{oʔmaʔ-ko} \qu{stop!} &   \obj{\emp{aj-}awon-ko} \qu{get up!} \\
-\panare &            \obj{ape-ʔ} \qu{flee!} &             \obj{ahpən-kə} \qu{jump!} \\
 \bottomrule
 \end{tabular}
+\end{table}

### 2021-10-11 documents/floats/nominalizations.tex
added additional split-s data, bakairi examples, rewrote up to sec 3.4
cleaned up data, started work on discussion section

@@
+\begin{table}
+\centering
+\caption[Nominalizations of \gl{s_a_} and \gl{s_p_} verbs]{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs \parencites[90]{koehn1986apalai}[ner2-003]{koehns1994textos}[390]{panarepayne2013}[23]{mattei1994diccionario}[135, 392]{hoff1968carib}[97]{alves2017arara}[49, 74]{schuring2018kaxuyana}[246]{triomeira1999}[130, 409]{wayanatavares2005}}
+\label{tab:nominalizations}
 \begin{tabular}[t]{@{}lll@{}}
@@
 \midrule
+\apalai &            \obj{j-epɨ-topo} \qu{my bathing place} &   \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} \\
+\panare &     \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} &  \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} \\
+\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &    \obj{aj-ereʔna-{\normalfont ∅}} \qu{your fainting} \\
+\arara  &        \obj{\emp{w-}orik-tubo} \qu{dancing place} &              \obj{ereŋmɨ-tpo} \qu{killing instrument} \\
 \kaxui  &       \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} &            \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} \\
-\arara  &        \obj{\emp{w-}orik-tubo} \qu{dancing place} &              \obj{ereŋmɨ-tpo} \qu{killing instrument} \\
 \trio   &   \obj{ji-\emp{w-}əturu-to} \qu{(for) my talking} &              \obj{j-emamina-to} \qu{(for) my playing} \\
 \wayana &          \obj{ɨ-\emp{w-}əturu-topo} \qu{my story} &        \obj{j-ɨnɨkɨ-topo} \qu{my object for sleeping} \\
-\apalai &            \obj{j-epɨ-topo} \qu{my bathing place} &   \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} \\
-\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &                  \obj{aj-ereʔna-∅} \qu{your fainting} \\
-\panare &     \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} &  \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} \\
 \bottomrule
 \end{tabular}
+\end{table}

### 2021-10-11 documents/sections/proto_carib.tex
cleaned up data, started work on discussion section

@@
 With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
-Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles} for participles, and in \cref{tab:nmlz} for nominalizations.
+Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles} for participles, and in \cref{tab:nominalizations} for nominalizations.
 %\kaxui does not preserve \rc{w-} in its participles (\obj{tehurkat͡ʃe} \qu{fallen}, not \ungr{tɨwehurkat͡ʃe}), while \apalai only preserves it there, but not in nominalizations (\obj{jepɨtopo} \qu{my bathing place}, not \ungr{joepɨtopo}).
 
-\begin{table}
-	\centering
-	\caption{Participles of \gl{s_a_} and \gl{s_p_} verbs \parencites[118, 207]{alves2017arara}[30, 42]{camargo2002lexico}[430, 433]{hoff1968carib}[333, 334]{triomeira1999}[400]{wayanatavares2005}[39]{schuring2018kaxuyana}[232, 244]{panarepayne2013}}
-	\label{tab:participles}
 \input{floats/participles}
-\end{table}
 
-\begin{table}
-	\centering
-	\caption{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs \parencites[49, 74]{schuring2018kaxuyana}[97]{alves2017arara}[246]{triomeira1999}[130, 409]{wayanatavares2005}[90]{koehn1986apalai}[ner2-003]{koehns1994textos}[135, 392]{hoff1968carib}[390]{panarepayne2013}[23]{mattei1994diccionario}}
-	\label{tab:nmlz}
 \input{floats/nominalizations}
-\end{table}
-
-%\begin{table}
-%	\centering
-%	\caption{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs}
-%	\label{tab:nmlz}
-%	\begin{tabular}{@{}llll@{}}
-%	\mytoprule
-%Language & \gl{s_a_} verb & \gl{s_p_} verb & Source(s)\\
-%	\mymidrule
-%\kaxui & \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} & \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} & \textcite[49, 74]{schuring2018kaxuyana}\\
-%\arara & \obj{\emp{w-}orik-tubo} \qu{dancing place} & \obj{ereŋmɨ-tpo} \qu{killing instrument} & \textcite[97]{alves2017arara} \\
-%\trio & \obj{ji-\emp{w-}əturu-to} \qu{(for) my talking} & \obj{j-emamina-to} \qu{(for) my playing} & \textcite[246]{triomeira1999} \\
-%\wayana & \obj{ɨ-\emp{w-}əturu-topo} \qu{my story} & \obj{j-ɨnɨkɨ-topo} \qu{my object for sleeping} & \textcite[409, 130]{wayanatavares2005}\\
-%\kalina & \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} & \obj{aj-ereʔna-Ø} \qu{your fainting} & \textcite[392, 135]{hoff1968carib}\\
-%\apalai & \obj{j-epɨ-topo} \qu{my bath place} & \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} & \textcites[90]{koehn1986apalai}[ner2/1]{koehns1994textos}\\
-%\panare & \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} & \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} & \textcites[390]{panarepayne2013}[23]{mattei1994diccionario}\\
-%	\mybottomrule
-%	\end{tabular}
-%\end{table}
 
@@
 Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed.
-This is illustrated with reflexes in various modern languages in \cref{tab:imp}.
+This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
 As in the case of the \gl{s_a_} marker \rc{w-} participles and nominalizations, some languages have lost the distinction between \gl{s_a_} and \gl{s_p_} verbs in imperatives, for example \panare.
 
-\begin{table}
-	\centering
-	\caption{Imperatives of \gl{s_a_} and \gl{s_p_} verbs \parencites[44, 89]{derbyshire1965textos}[161]{alves2017arara}[323]{triomeira1999}[227]{wayanatavares2005}[62]{koehn1986apalai}[Mopo/20]{koehns1994textos}[190]{hoff1968carib}[5, 17]{mattei1994diccionario}}
-	\label{tab:imp}
 \input{floats/imperatives}
-\end{table}
 

### 2021-10-11 workflow/process_data.py

@@ -1,0 +1,300 @@
+import pyradigms as pyd
+import pandas as pd
+pd.options.mode.chained_assignment = None
+import cariban_helpers as cah
+import cldf_helpers as cldfh
+import pynterlinear as pynt
+
+lg_list = cah.lg_order().keys()
+cognate_list = ["go", "say", "come", "be-1", "be-2", "go.down", "bathe"]
+cs_df = pd.read_csv("../data/cognate_sets.csv")
+v_df = pd.read_csv("../data/verb_stem_data.csv")
+
+#functions for creating latex tables
+
+def print_latex(df, ex=False, keep_index=False):
+    if keep_index:
+        df.columns.name = df.index.name
+    df.index.name = None
+    with pd.option_context("max_colwidth", 1000):
+        lines = df.to_latex(escape=False).split("\n")
+    lines[0] = lines[0].replace("tabular}{l", "tabular}[t]{@{}l").replace("l}", "l@{}}")
+    if ex:
+        del lines[1:4]
+        del lines[-1]
+        del lines[-2]
+        return "\n".join(lines)
+    else:
+        return "\n".join(lines).replace(u"\toprule", u"\mytoprule")
+        
+def save_float(tabular, label, caption, filename=None, short=None):
+    if not filename:
+        filename = label
+    if not short:
+        shorttext = ""
+    else:
+        shorttext = f"[{short}]"
+    output = f"""\\begin{{table}}
+\centering
+\caption{shorttext}{{{caption}}}
+\label{{tab:{label}}}
+{tabular}\end{{table}}"""
+    f = open(f"../documents/floats/{filename}.tex", "w")
+    f.write(output)
+    f.close()
+
+normalfonties = ["Ø", "?", "V", "G", "C", "∅"]
+def objectify(string, obj_string="obj"):
+    if string in ["", "?", " ", "  ", "–"]:
+        return string
+    if string is None:
+        return ""
+    if string[0] == " ":
+        return string
+    output = []
+    entry_list = string.split(" OR ")
+    for entry in entry_list:
+        out_list = []
+        morph_list = entry.split("; ")
+        for i, morph in enumerate(morph_list):
+            out_list.append(r"\%s{%s}" % (obj_string, morph))
+            if i > 0: obj_string = "obj"
+        output.append("/".join(out_list))
[…]

### 2021-10-11 documents/floats/imperatives.tex
discarded hypothesis about PPek *k-əd-ebɨ

--- 2021-09-28 documents/floats/imperatives.tex
@@
+\begin{table}
+\centering
+\caption[Imperatives of \gl{s_a_} and \gl{s_p_} verbs]{Imperatives of \gl{s_a_} and \gl{s_p_} verbs \parencites[44, 89]{derbyshire1965textos}[161]{alves2017arara}[323]{triomeira1999}[227]{wayanatavares2005}[62]{koehn1986apalai}[Mopo/20]{koehns1994textos}[190]{hoff1968carib}[5, 17]{mattei1994diccionario}}
+\label{tab:imperatives}
 \begin{tabular}[t]{@{}lll@{}}
@@
 \end{tabular}
+\end{table}

### 2021-10-11 documents/floats/nominalizations.tex
discarded hypothesis about PPek *k-əd-ebɨ

--- 2021-09-10 documents/floats/nominalizations.tex
@@
+\begin{table}
+\centering
+\caption[Nominalizations of \gl{s_a_} and \gl{s_p_} verbs]{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs \parencites[49, 74]{schuring2018kaxuyana}[97]{alves2017arara}[246]{triomeira1999}[130, 409]{wayanatavares2005}[90]{koehn1986apalai}[ner2-003]{koehns1994textos}[135, 392]{hoff1968carib}[390]{panarepayne2013}[23]{mattei1994diccionario}}
+\label{tab:nominalizations}
 \begin{tabular}[t]{@{}lll@{}}
@@
 \apalai &            \obj{j-epɨ-topo} \qu{my bathing place} &   \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} \\
-\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &                  \obj{aj-ereʔna-∅} \qu{your fainting} \\
+\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &    \obj{aj-ereʔna-{\normalfont ∅}} \qu{your fainting} \\
 \panare &     \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} &  \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} \\

### 2021-10-11 documents/floats/participles.tex
cleaned up data, started work on discussion section
discarded hypothesis about PPek *k-əd-ebɨ

@@
+\begin{table}
+\centering
+\caption[Participles of \gl{s_a_} and \gl{s_p_} verbs]{Participles of \gl{s_a_} and \gl{s_p_} verbs \parencites[39]{schuring2018kaxuyana}[118, 207]{alves2017arara}[333, 334]{triomeira1999}[400]{wayanatavares2005}[35]{koehn1986apalai}[kuruaz-154]{koehns1994textos}[430, 433]{hoff1968carib}[232, 244]{panarepayne2013}}
+\label{tab:participles}
 \begin{tabular}[t]{@{}lll@{}}
@@
 \wayana &     \obj{tə-\emp{w-}epɨ-he} \qu{bathed} &    \obj{t-onopɨ-he} \qu{painted} \\
-\apalai &  \obj{t-\emp{o-}erama-se} \qu{returned} &      \obj{t-anorɨ-se} \qu{dried} \\
+\apalai &        \obj{t-\emp{o-}ɨto-se} \qu{gone} &   \obj{t-ɨhto-se} \qu{gone down} \\
 \kalina &  \obj{tu-\emp{w-}oʔka-se} \qu{come out} &       \obj{t-okari-se} \qu{told} \\

### 2021-10-11 documents/sections/extensions.tex
working on overview of etxensions and verbs
discarded hypothesis about PPek *k-əd-ebɨ

--- 2021-09-26 documents/sections/extensions.tex
@@
 \crefrange{sec:akuriyo}{sec:yukpa} look at innovative first person markers which are only attested in single modern languages:
-\obj{k-} in \akuriyo, \obj{j(i)-} in \carijo, and \obj{j(ɨ)-} \yukpa.
+\obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.
 
@@
 %
-We can imagine two scenarios leading to this situation:
-One possibility is that there was some variation in \PPar as to whether or not the first person form of \qu{to go} was based on \rc{ətepɨ} or \rc{epɨ}.
-Alternatively, the \PPek paradigm had a similar person-based distribution of \rc{əd-} as \arara, and \bakairi and \ikpeng both independently regularized the paradigm to show \rc{əd-epɨ} with all persons.
-However, in both scenarios there is a correlation of \rc{w-} and \rc{ebɨ}, and \rc{k-} and \rc{əd-ebɨ}.
+On the other hand, both \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
+Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm.
 
@@
 As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}; we offer our reconstructed \PTir forms.
+In addition, \trio \obj{ei} \qu{to be}, the other copular root, also shows first person \obj{w-} \parencite[294]{triomeira1999}, allowing us to reconstruct a first-person \rc{w-eʔi-} for \PTir.
 

### 2021-10-11 workflow/create_intr_tables.py
deleted old python scripts

@@ -1,160 +1,0 @@
-import pandas as pd
-from cariban_helpers import *
-import seaborn as sns
-
-def print_source(string):
-    entries = string.split("; ")
-    output = ""
-    for entry in entries:
-        bibkey = entry.split("[")[0]
-        pages = entry.split("[")[1].replace("]","")
-        if bibkey == "pc":
-            return "\\perscomm{%s}" % pages
-        else:
-            output += "[%s]{%s}" % (pages, bibkey)
-    return "\\textcites%s" % output
-
-def shorthandify(lg_id):
-    return "\\" + get_shorthand(lg_id)
-
-def objectify(string, obj_string="obj"):
-    if string in ["", "?", " ", "  "]:
-        return string
-    if string[0] == " ":
-        return string
-    out_list = []
-    morph_list = string.split("; ")
-    for i, morph in enumerate(morph_list):
-        out_list.append(r"\%s{%s}" % (obj_string, morph))
-        if i > 0: obj_string = "obj"
-    return "/".join(out_list)
-    
-a = pd.read_csv("come_roots.csv")
-# print(a)
- 
-a["Language_ID"] = pd.Categorical(a['Language_ID'], lg_order().keys())
-a.sort_values("Language_ID", inplace=True)
-
-a["Source"] = a["Source"].apply(print_source)
-a["Form"] = a["Form"].apply(lambda x: "\obj{%s}" % x)
-a["Language"] = a["Language_ID"].apply(shorthandify)
-
-cogsets = {
-    1: "jəpɨ",
-    2: "əpɨ",
-    3: "epɨ",
-    4: "ətjəpɨ",
-}
-
-def get_palette(n):
-    palette = sns.color_palette("pastel", n)
-    for i, color in enumerate(palette):
-        palette[i] = [str(x*255) for x in color]
-    return dict(zip(range(1,len(palette)+1), palette))
-
-colors = get_palette(len(cogsets))
-
-cmd_out = ""
-for i, j in colors.items():
-    cmd_out += "\\definecolor{come%s}{RGB}{%s}\n" % (i, ",".join(j))
-# print(cmd_out)
-
-output = ""
-df = a
-for index, row in df.iterrows():
-    row_colors = "\\colorbox{come%s}" % row["Cognateset_ID"]
[…]

### 2021-10-14 documents/sections/extensions.tex
various revisions

@@
 %
-From a comparative view, this list is not quite complete, as there is also a reflex of the copula \rc{a(p)}, which is analayzed as a postposition by \textcite[199--201]{alves2017arara}; it is however inflected with \setone prefixes and also shows first person \obj{w-} \exref{araap}.
+From a comparative view, this list is not quite complete, as there is also a reflex of the copula \rc{a[p]}, which is analayzed as a postposition by \textcite[199--201]{alves2017arara}; it is however inflected with \setone prefixes and also shows first person \obj{w-} \exref{araap}.
 
@@
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-While there are reflexes of \rc{a(p)} \qu{to be} in \ikpeng, only \rc{eti} forms are attested inflected for first person \parencite[401]{gildea2018reconstructing}.
+While there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} forms are attested inflected for first person \parencite[401]{gildea2018reconstructing}.
 
@@
 %
-The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{In fact, the issue of \emph{when} this happened is also open.
+The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
+In fact, even the issue of \emph{when} this happened is open.
 It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation.
-Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.}
-As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}; we offer our reconstructed \PTir forms.
-In addition, \trio \obj{ei} \qu{to be}, the other copular root, also shows first person \obj{w-} \parencite[294]{triomeira1999}, allowing us to reconstruct a first-person \rc{w-eʔi-} for \PTir.
+Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
+We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it happening already in \PTar.}
+%
+As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the first four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}, for which we offer our reconstructed \PTir forms.
+To this list, we add the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}}, which has first person \obj{w-} at least in \trio.
 
-\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-} \parencites[112--115]{meira1998proto}[85]{gildea1994akuriyo}\\
+\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-} \parencites[112--115]{meira1998proto}[294]{triomeira1999}\\ %[85]{gildea1994akuriyo}
 \begin{tabular}[t]{@{}llll@{}}
 & \PTir & \trio & \akuriyo\\
-\qu{go} & \rc{wɨ-tə(mɨ)-} & \obj{wɨ-tə(mɨ)-} & \obj{ə-təmɨ-}/\obj{wɨ-təemɨ-}  \\
+\qu{go} & \rc{wɨ-tə(mɨ)-} & \obj{wɨ-tə(mɨ)-} & \obj{ə-təmɨ-} \\% (\obj{wɨ-təemɨ-})\\
 \qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
@@
 \qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
+& \rc{w-eʔi-} & \obj{w-ei-} & ?-\obj{eʔi-}\\
 \end{tabular}
@@
 
-For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} registered a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
-The two forms can be reconciled by a specific idiosyncratic combination of sound changes:
-We suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
-Also, \obj{w-} appears to have been subject to ongoing erosion in \akuriyo, also evidenced by its absence in \qu{to come} and \qu{to be}, but its presence in \qu{to say} \exref{ptirw}.
-This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix \exref{exp.aku-164}, as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
+We agree with \posscite[113]{meira1998proto} identification of the idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} as a reflex of \rc{wɨ-}.
+Both components of the irregular change \rc{wɨ-} > \obj{ə-} -- loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə} -- are found in other person prefixes \exref{exp}.
 
 \pex<exp>
+\a<aku-164> \akuriyo \parencite[][86]{gildea1994akuriyo}\\
+\begingl
+\gla (w)i-toka//
+\glb \gl{1}>\gl{3}-hit//
+\glft \qu{I hit him/her.}//
+\endgl
 \a<aku-159> \akuriyo \parencite[][114]{meira1998proto}\\
@@
 \endgl
-\a<aku-164> \akuriyo \parencite[][86]{gildea1994akuriyo}\\
-\begingl
-\gla (w)i-toka//
-\glb \gl{1}>\gl{3}-hit//
-\glft \qu{I hit him/her.}//
-\endgl
 \xe
 %
-For the form of the verb, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for the first person, and \rc{eepɨ} for the others, based on the idiosyncratic form in \trio and the vowel length in \akuriyo.
-\akuriyo (and \carijo) would then have levelled the original distribution of \rc{əepɨ} and \rc{eepɨ}, similar to what we have suggested for Pekodian in \cref{sec:pekodian}.
-We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly cognate with \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ}.
+%For \akuriyo \qu{to go}, \textcite{gildea1994akuriyo} registered a different prefix \obj{wɨ-}, rather than \posscite{meira1998proto} \obj{ə-}.\footnote{The \akuriyo recorded by Gildea potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998}.}
+%The two forms can be reconciled by a specific idiosyncratic combination of sound changes:
+%We suggest that \obj{ə} is the outcome of the lowering of the \rc{ɨ} in the prefix \rc{wɨ-}; the same has happened to the vowel in the \gl{1+2} prefix \rc{kɨt-} \exref{exp.aku-159}.
+%Also, \obj{w-} appears to have been subject to ongoing erosion in \akuriyo, also evidenced by its absence in \qu{to come} and \qu{to be}, but its presence in \qu{to say} \exref{ptirw}.
+%This erosion is also found in the etymologically related \gl{1}>\gl{3} prefix \exref{exp.aku-164}, as well as in other Cariban languages, like \ikpeng \pcref{sec:pekodian} or \hixka \pcref{sec:waiwaian}.
+%
+%
+%
+For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for the first person, and \rc{eepɨ} for the others, based on the idiosyncratic form in \trio and the vowel length in \akuriyo.
+\akuriyo (and \carijo) then levelled the original distribution of \rc{əepɨ} and \rc{eepɨ}, similar to what we have suggested for Pekodian \pcref{sec:pekodian}.
+We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly cognate with \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form must have been \rc{əʔepɨ}.
 
@@
 
-\textcite{gildea1994akuriyo} also recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-}.
-They are listed in \exref{akumov.aku}.
+Further, here are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
+They have \gl{1}\gl{s_a_} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
+At least for \qu{to go down}, the \akuriyo cognate suggests that the irregular first person form \rc{p-ɨhtə-} can be reconstructed to \PTir.
+Whatever their origins, they were not affected by the extension of \rc{t-}.
+
+\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
+\begin{tabular}[t]{@{}lll@{}}
+& \trio & \akuriyo\\
+\qu{go down} & \obj{p-ɨhtə-} & \obj{p-ɨtə-}\\
+\qu{defecate} & \obj{k-oeka-} & ?\\
+\end{tabular}
+\xe
+
+Finally, \textcite{gildea1994akuriyo} also recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-} \exref{akumov.aku}, all \obj{e}-initial movement verbs.
 We have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
-Thus, it is uncertain whether these verbs can be reconstructed as taking \rc{w-} in \PTir, although its presence in \akuriyo does suggest so.
+Further, these verbs are not mentioned by \textcite{meira1998proto}, who was also working with \posscite{gildea1994akuriyo} data.
+Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by Meira's data, we cannot reconstruct these verbs as not being affected by the extension of \PTir \rc{t-}.
 
@@
 
-Finally, there are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
-They have \gl{1}\gl{s_a_} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
-At least for \qu{to go down}, the \akuriyo cognate suggests that the irregular first person form \rc{p-ɨhtə-} can be reconstructed to \PTir.
-Whatever their origins, they were not affected by the extension of \rc{t-}.
-
-\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
-\begin{tabular}[t]{@{}lll@{}}
-& \trio & \akuriyo\\
-\qu{go down} & \obj{p-ɨhtə-} & \obj{p-ɨtə-}\\
-\qu{defecate} & \obj{k-oeka-} & ?\\
-\end{tabular}
-\xe
-
-In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} largely replaced \gl{1}\gl{s_a_} \rc{w-}.
-Four verbs are solidly reconstructible as preserving \rc{w-} in \PTir, and \rc{ɨhtə} \qu{to go down} with irregular \rc{p-} was also not affected.
-There are four movement verbs in \akuriyo which appear to not have been affected by the extension of \rc{t-}, as well as another irregular verb \obj{weka} \qu{to defecate} in \trio, with first person \obj{k-oeka-}.
+%In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} largely replaced \gl{1}\gl{s_a_} \rc{w-}.
+%Five verbs are solidly reconstructible as preserving \rc{w-} in \PTir, and \rc{ɨhtə} \qu{to go down} with irregular \rc{p-} was also not affected.
+%There is another irregular verb \obj{weka} \qu{to defecate} in \trio, with first person \obj{k-oeka-}.
 
@@
 After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
-It seems to have replaced \rc{t-} only in specific places, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to us \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
+It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to us \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} could be more recent, with which we agree -- since phonologically distributed \rc{t-}/\rc{t͡ʃ-} is reconstructible to \PTir, a scenario whereby \obj{k-} replaces \rc{t-}, but not \rc{t͡ʃ-}, is the most straightforward.
+He also suggests that \obj{k-} might be more recent, with which we agree -- since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
 The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
@@
 \label{sec:carijo}
-As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.
+As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since almost all \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} is attested.}
 This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \exref{carpar}.
@@
 \gl{1} & \obj{ji-tuda-} & \obj{j-eharaga-}\\
-\gl{2} & \obj{mɨ-tuda-} & \obj{m-egaraga-}\\
+\gl{2} & \obj{mɨ-tuda-} & \obj{m-eharaga-}\\
 \gl{1+2} & \obj{kɨsi-tuda-} & \obj{kɨs-eharaga-}\\
@@
 %
-Thus, also detransitivized verbs like \obj{ehɨnəhɨ} \qu{to fight}, which are \gl{s_a_} verbs in other languages, take \obj{j-} \exref{car-30}.
+Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by taking a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance} \exref{car-30}.
 %The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
@@
 \glpreamble hẹnẹ(x)tónoko-mā́ɽẹ y-e̱-hẹnẹ́(x)yai̯//
-\gla hɨnəhtono-ko=marə j-e-hɨnəh-jai//
+\gla hɨnəhtono-ko=marə j-e-hɨnəhɨ-jai//
 \glb enemy-\gl{pl}=with \gl{1}-\gl{detrz}-kill-\gl{npst}.\gl{cert}//
@@
 
-As noted in \cref{sec:taranoan}, this extension also erased any traces of a putative \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
+As noted in \cref{sec:taranoan}, this extension also erased any traces of a potential \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
 However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.

### 2021-10-14 documents/sections/verbs.tex
cleaned up data, started work on discussion section
various revisions

@@
 In this section, we present these verbs from a comparative perspective and discuss their reconstructability.
-\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a(p)} \qu{to be}, \cref{sec:say} \rc{ka(ti)} \qu{to say}, \cref{sec:go} \rc{ɨtə(mə)} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstructed as \gl{s_a_} verbs that were not derived from transitive verbs.
-\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek; \cref{sec:bathe} investigates \PPek \rc{ɨpɨ} \qu{to bathe}, \cref{sec:shit} \trio \qu{to defecate}, and \cref{sec:movement} the \akuriyo movement verbs.
+\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstructed as \gl{s_a_} verbs that were not derived from transitive verbs.
+\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek; \cref{sec:bathe} investigates \PPek \rc{ɨpɨ} \qu{to bathe}, and \cref{sec:shit} \trio \qu{to defecate}.
+%, and \cref{sec:movement} the \akuriyo movement verbs.
 The \obj{e}-initial verbs not affected by the extension of \rc{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
 
-%& \PPeks \rc{k-} & \acs{ara_lg} & \acs{ikp_lg} & \acs{bak_lg} & \PWais \rc{k-} & \acs{wai_lg} & \acs{hix_lg} & \PTirs \rc{t-} & \acs{tri_lg} & \acs{aku_lg} & \acs{aku_lg} \obj{k-} & \acs{car_lg} \obj{j(i)-} & \acs{yuk_lg} \obj{j-} \\
-%\rc{ɨtə(mə)} \qu{go} & × & × & \checkmark & × & × & (\checkmark) & × & × & × & × & × & × & × \\
-%\rc{ka(ti)} \qu{say} & × & × & × & × & × & × & × & × & × & × & × & ? & ? \\
-%\rc{ət-epɨ} \qu{come} & \checkmark & -- & \checkmark & \checkmark & -- & -- & -- & × & × & -- & -- & -- & ? \\
-%\rc{jəpɨ} \qu{come} & × & × & -- & -- & -- & -- & -- & × & × & × & × & \checkmark & ? \\
-%\rc{ɨptə} \qu{go down} & × & × & ? & \checkmark & -- & -- & -- & × & × & × & × & ? & -- \\
-%\PPeks\rc{ipɨ} \qu{bathe} & × & × & × & × & -- & -- & -- & -- & -- & -- & -- & -- & -- \\
-%\rc{eti} \qu{be} & × & × & × & × & × & × & × & × & × & × & × & \checkmark & \checkmark \\
-%\rc{a(p)} \qu{be} & × & -- & -- & × & × & × & × & × & × & × & × & × & \checkmark \\
-%	\mybottomrule
-%	\end{tabular}
-%\end{table}
-
-\subsection{\rc{eti}/\obj{a(p)} \qu{to be}}
+\subsection{\rc{eti}/\obj{a[p]} \qu{to be}}
 \label{sec:be}
 For a comprehensive comparative overview for this verb, we refer the reader to \textcite[375--382]{gildea2018reconstructing}, who reconstructs two distinct roots serving as verbs \qu{to be} in modern Cariban languages.
-One is the older copula \rc{a(p)}, which can be reconstructed as already having various irregularities in \PC.
+One is the older copula \rc{a[p]}, which can be reconstructed as already having various irregularities in \PC.
 The other is a root \rc{eti} reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
@@
 \carijo innovated \obj{j-}, but only in the \obj{et͡ʃi} root allomorph \exref{carcop.mivida-12}; the \obj{a} root preserves \obj{w-} \exref{carcop.car-25}.
-\yukpa innovated \obj{j-} for reflexes of both \rc{a(p)} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
+\yukpa innovated \obj{j-} for reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
 
@@
 
-\subsection{\rc{ka(ti)} \qu{to say}}
+\subsection{\rc{ka[ti]} \qu{to say}}
 \label{sec:say}
@@
 
-Further comparative evidence also points to \rc{ka(ti)} \qu{to say} showing transitive traits:
+Further comparative evidence also points to \rc{ka[ti]} \qu{to say} showing transitive traits:
 \trio \obj{ka} is characterized as the only intransitive verb being able to take the causative suffix \obj{-po} and the agentive nominalizer \obj{-ne} \parencite[263, 169]{triomeira1999}.
@@
 
-\subsection{\rc{ɨtə(mə)} \qu{to go}}
+\subsection{\rc{ɨtə[mə]} \qu{to go}}
 \label{sec:go}
-This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə(mə)}, with the second syllable only rarely occurring, as in the case of \rc{ka(ti)} \qu{to say}.
+This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə[mə]}, with the second syllable only rarely occurring, as in the case of \rc{ka[ti]} \qu{to say}.
 It is true that many reflexes are clearly \obj{t}-initial, for example \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}.
-However, a comparison of selected forms with unambiguously C-initial \rc{ka(ti)} \qu{to say} suggests that an initial vowel \rc{ɨ} should be reconstructed \exref{gosaycomp},\footnote{Many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.} which was subsequently reanalyzed as part of the prefixes in many languages.
+However, a comparison of selected forms with unambiguously C-initial \rc{ka[ti]} \qu{to say} suggests that an initial vowel \rc{ɨ} should be reconstructed \exref{gosaycomp},\footnote{Many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.} which was subsequently reanalyzed as part of the prefixes in many languages.
 
@@
 \input{floats/defecate}
-
-\subsection{The \akuriyo movement verbs}
-\label{sec:movement}
-As discussed in \cref{sec:akuriyo}, four \akuriyo \gl{s_a_} verbs are attested as not having the first person marker \obj{t͡ʃ-} in \textcite{gildea1994akuriyo}; instead, they have \obj{w-} or its phonologically reduced form ∅.
-These verbs share two properties: they are all \obj{e-}initial, and they are all movement verbs.
-However, the same dataset also contains \obj{e-}initial movement verbs with a first person marker \obj{t͡ʃ-} \exref{aku-106}.
-As for the etymology of these four verbs, we can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
-For the other three verbs, we have not encountered potential cognates, but given their \obj{e-}initial nature and their semantics, it seems plausible that they, too, are derived from transitive verbs with meanings like \qu{to put up} etc.
-In any case, \akuriyo is only sparsely documented, and there are many open questions about it; the \akuriyo data in \textcite{gildea1994akuriyo} are not fully reliable.
-
-\ex<aku-106>\akuriyo \parencite[][84]{gildea1994akuriyo}\\
-\begingl
-\gla t͡ʃ-ewai//
-\glb \gl{1}-sit.down//
-\glft \qu{I sat down.}//
-\endgl
-\xe
+%
+%\subsection{The \akuriyo movement verbs}
+%\label{sec:movement}
+%As discussed in \cref{sec:akuriyo}, four \akuriyo \gl{s_a_} verbs are attested as not having the first person marker \obj{t͡ʃ-} in \textcite{gildea1994akuriyo}; instead, they have \obj{w-} or its phonologically reduced form ∅.
+%These verbs share two properties: they are all \obj{e-}initial, and they are all movement verbs.
+%However, the same dataset also contains \obj{e-}initial movement verbs with a first person marker \obj{t͡ʃ-} \exref{aku-106}.
+%As for the etymology of these four verbs, we can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
+%For the other three verbs, we have not encountered potential cognates, but given their \obj{e-}initial nature and their semantics, it seems plausible that they, too, are derived from transitive verbs with meanings like \qu{to put up} etc.
+%In any case, \akuriyo is only sparsely documented, and there are many open questions about it; the \akuriyo data in \textcite{gildea1994akuriyo} are not fully reliable.
+%
+%\ex<aku-106>\akuriyo \parencite[][84]{gildea1994akuriyo}\\
+%\begingl
+%\gla t͡ʃ-ewai//
+%\glb \gl{1}-sit.down//
+%\glft \qu{I sat down.}//
+%\endgl
+%\xe

### 2021-10-18 workflow/process_data.py
added 'bathe' data to CLDF, proper modeling of cognates

@@ -217,6 +217,14 @@ pyd.x = ["Cognateset_ID"]
 pyd.y = ["Orig_Language", "Form", "Language_ID"]
 pyd.content_string = "Affected"
 pyd.x_sort = verb_list
+
+repl_dic = {
+    "DETRZ+come": "come"
+}
+
+overview["Cognateset_ID"] = overview["Cognateset_ID"].replace(repl_dic)
+
+
 result = pyd.compose_paradigm(overview)
 result["Lg"] = pd.Categorical([lvl[2] for lvl in result.index.str.split(".")], lg_list)
 result["Orig"] = pd.Categorical(
@@ -247,12 +255,14 @@ def modify_index(idx):
 
 result.index = result.index.map(modify_index)
 
-# replace cogset ids with reconstructed forms and translations
+# replace cogset ids with reconstructed forms and translationsm
+cs_df["Gloss"] = cs_df["Meaning"].replace(" ", ".")
 t_d = dict(zip(cs_df["ID"], cs_df["Gloss"]))
 f_d = dict(zip(cs_df["ID"], cs_df["Form"]))
+
 result.columns = [
     result.columns.map(lambda x: f"\\rc{{{f_d[x]}}}"),
-    result.columns.map(lambda x: f"\\qu{{{t_d[x]}}}"),
+    result.columns.map(lambda x: f"\\qu{{to {t_d[x]}}}"),
 ]
 
 # add nice-looking checkmarks and stuff

### 2021-10-20 documents/floats/defecate.tex
various revisions
cleaned up partial and normal alignment generation

--- 2021-10-05 documents/floats/defecate.tex
@@
 \centering
-\caption{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
+\caption[\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb]{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
 \label{tab:defecate}
@@
 \ikpeng &   \obj{atke} &              \gl{s_p_} \\
-\bakairi &     \obj{əeke} &              \gl{s_a_} \\
 \trio   &   \obj{weka} &              \gl{s_a_} \\
@@
 \maqui  &   \obj{weka} &              \gl{s_p_} \\
-\panare  &  \obj{(a)iʔka} &                      ? \\
+\panare &   \obj{iʔka} &                      ? \\
 \bottomrule

### 2021-10-22 workflow/lingpy_alignments.py
working on comparative tables with aligned segments

@@ -1,0 +1,146 @@
+import lingpy
+from lingpy.compare.partial import Partial
+import pandas as pd
+import os
+from difflib import SequenceMatcher
+
+tempfile = "lingpy_temp.tsv"
+
+cldf_lingpy = {
+    "Language_ID": "DOCULECT",
+    "Parameter_ID": "CONCEPT",
+    "Form": "IPA",
+    "Cognateset_ID": "COGIDS",
+    "Segments": "TOKENS",
+    "Source": "SOURCE",
+    "Alignment": "ALIGNMENT",
+    "Comment": "COMMENT",
+    "ID": "CLDF_ID"
+}
+
+def rename_dict(df):
+    for col in df.columns:
+        if col not in cldf_lingpy:
+            cldf_lingpy[col] = col.upper()
+    lingpy_cldf = {y: x for x, y in cldf_lingpy.items()}
+    return cldf_lingpy, lingpy_cldf
+
+lingpy_cldf = {y: x for x, y in cldf_lingpy.items()}
+
+def merge_sequences(seq1,seq2):
+    sm=SequenceMatcher(a=seq1,b=seq2)
+    res = []
+    for (op, start1, end1, start2, end2) in sm.get_opcodes():
+        if op == 'equal' or op=='delete':
+            #This range appears in both sequences, or only in the first one.
+            res += seq1[start1:end1]
+        elif op == 'insert':
+            #This range appears in only the second sequence.
+            res += seq2[start2:end2]
+        elif op == 'replace':
+            #There are different ranges in each sequence - add both.
+            res += seq1[start1:end1]
+            res += seq2[start2:end2]
+    return res
+    
+def get_order(lol):
+    lol = sorted(lol, key=len, reverse=True)
+    out = lol[0]
+    for i in range(1, len(lol)):
+        out = merge_sequences(out, lol[i])
+    return out
+
+def is_empty(slot):
+    if list(set(slot.split())) == ["-"]:
+        return True
+    else:
+        return False
+
+def join_slots(slotlist, almsep):
+    out = [slotlist[0]]
+    for slot in slotlist[1::]:
+        if is_empty(out[-1]):
+            out.append("-")
+        else:
+            out.append("+")
+        out.append(slot)
+    return " ".join(out)
+
+def buffer_alignments(df, cogid="COGIDS", cogsep=" ", almid="ALIGNMENT", almsep=" + "):
[…]

### 2021-10-22 workflow/process_data.py

@@ -6,11 +6,37 @@ import cldf_helpers as cldfh
 import pynterlinear as pynt
 import numpy as np
 import re
+from segments import Tokenizer, Profile
+from lingpy_alignments import calculate_alignment
 
 lg_list = list(cah.lg_order().keys())
 cognate_list = ["go", "say", "come", "be_1", "be_2", "go_down", "bathe_intr"]
 cs_df = pd.read_csv("../data/cognate_sets.csv")
 v_df = pd.read_csv("../data/verb_stem_data.csv")
+v_df.rename(columns={"Meaning": "Parameter_ID"}, inplace=True)
+segments = open("../data/segments.txt").read()
+segment_list = [{"Grapheme": x, "mapping": x} for x in segments.split("\n")]
+t = Tokenizer(Profile(*segment_list))
+cognatesets = pd.read_csv("../data/cldf/cognatesets.csv")
+
+#lingpy does not like cogid 0
+cognatesets.index = cognatesets.index+1
+cognum = dict(zip(cognatesets["ID"], cognatesets.index.astype(str)))
+numcog = dict(zip(cognatesets.index.astype(str), cognatesets["ID"]))
+
+print("Assigning cognates")
+
+def str2numcog(cogsets):
+    return " ".join([cognum[x] for x in cogsets.split("+")])
+
+def num2strcog(cogids):
+    return "+".join([numcog[i] for i in cogids.split(" ")])
+
+def segmentify(form):
+    form = re.sub("[()]", "", form)
+    form = form.replace("-", "+")
+    form = form.strip("+")
+    return t(form)
 
 #functions for creating latex tables
 
@@ -90,147 +116,149 @@ def get_sources(df, parens=True):
     pyd.content_string = tmp
     return source_string
 
+def sort_lg(df):
+    df.Language_ID = df.Language_ID.astype("category")
+    df.Language_ID.cat.set_categories(lg_list, ordered=True, inplace=True)
+    df.sort_values(["Language_ID"], inplace=True)
+
 def print_shorthand(abbrev):
     return "\\" + cah.get_shorthand(abbrev)
 
-#compile overview of individual cognate sets
-def print_cognate_table(df, verb, print_class=False):
-    #filter by verb
-    df_v = df[df["Cognateset_ID"] == verb]
-    #get bib references
-    df_v["Source"] = df_v["Source"].fillna("")
-    src = list(df_v["Source"])
[…]

### 2021-10-22 documents/floats/defecate.tex
'to defecate'

--- 2021-10-05 documents/floats/defecate.tex
@@
 \centering
-\caption{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
+\caption[Reflexes of \qu{to defecate} ]{Reflexes of \qu{to defecate} \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}{meira2005bakairi}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
 \label{tab:defecate}
@@
 \maqui   &   \obj{weka} &              \gl{s_p_} \\
-\panare  &  \obj{(a)iʔka} &                      ? \\
+\panare  &   \obj{iʔka} &                      ? \\
+\panare  &  \obj{aiʔka} &                      ? \\
 \bottomrule

### 2021-10-22 workflow/process_data.py

@@ -140,6 +140,14 @@ def sort_lg(df):
 def print_shorthand(abbrev):
     return "\\" + cah.get_shorthand(abbrev)
 
+def extract_sources(df, src_str="Source"):
+    df[src_str] = df[src_str].fillna("")
+    src = list(df[src_str])
+    if "" in src:
+        src.remove("")
+    sources = cldfh.cite_a_bunch(src, parens=True)
+    df.drop(columns=[src_str], inplace=True)
+    return sources
 
 def get_obj_str(lg):
     if lg[0] == "P":
@@ -147,16 +155,19 @@ def get_obj_str(lg):
     else:
         return "obj"
 
+def add_obj_markdown(df, lg_col="Language_ID", form_col="Form"):
+    df[form_col] = df.apply(
+        lambda x: objectify(x[form_col], get_obj_str(x[lg_col])), axis=1
+    )
+
+def repl_lg_id(df):
+    df["Language_ID"] = df["Language_ID"].apply(print_shorthand)
+    df.rename(columns={"Language_ID": "Language"}, inplace=True)
 
 # prepare overviews of class-switching 'go down' and 'defecate'
 print("\nClass membership of 'to go down':")
 gd_df = v_df[v_df["Parameter_ID"] == "go_down"]
-gd_df["Source"] = gd_df["Source"].fillna("")
-src = list(gd_df["Source"])
-if "" in src:
-    src.remove("")
-sources = cldfh.cite_a_bunch(src, parens=True)
-gd_df.drop(columns=["Source"], inplace=True)
+sources = extract_sources(gd_df)
 
 grouped = gd_df.groupby(gd_df.Cog_Cert)
 temp_df1 = grouped.get_group(1.0)
@@ -187,11 +198,8 @@ gd_df["Class"] = gd_df.apply(
     else x["Class"],
     axis=1,
 )
-gd_df["Form"] = gd_df.apply(
-    lambda x: objectify(x["Form"], get_obj_str(x["Language_ID"])), axis=1
-)
-gd_df["Language_ID"] = gd_df["Language_ID"].apply(print_shorthand)
-gd_df.rename(columns={"Language_ID": "Language"}, inplace=True)
+add_obj_markdown(gd_df)
+repl_lg_id(gd_df)
 gd_df.set_index("Language", drop=True, inplace=True)
 tabular = print_latex(gd_df, keep_index=True)
 save_float(
[…]

### 2021-10-22 workflow/process_data.py
verb table function

@@ -19,14 +19,14 @@ segments = open("../data/segments.txt").read()
 segment_list = [{"Grapheme": x, "mapping": x} for x in segments.split("\n")]
 t = Tokenizer(Profile(*segment_list))
 cognatesets = pd.read_csv("../data/cldf/cognatesets.csv")
-cognatesets = cognatesets.append({"ID": "?", "Description": "Unknown material"}, ignore_index=True)
+cognatesets = cognatesets.append(
+    {"ID": "?", "Description": "Unknown material"}, ignore_index=True
+)
 
 # lingpy does not like cogid 0
 cognatesets.index = cognatesets.index + 1
 cognum = dict(zip(cognatesets["ID"], cognatesets.index.astype(str)))
 numcog = dict(zip(cognatesets.index.astype(str), cognatesets["ID"]))
-
-print("Assigning cognates")
 
 
 def str2numcog(cogsets):
@@ -45,8 +45,6 @@ def segmentify(form):
 
 
 # functions for creating latex tables
-
-
 def print_latex(df, ex=False, keep_index=False):
     if keep_index:
         df.columns.name = df.index.name
@@ -140,6 +138,7 @@ def sort_lg(df):
 def print_shorthand(abbrev):
     return "\\" + cah.get_shorthand(abbrev)
 
+
 def extract_sources(df, src_str="Source"):
     df[src_str] = df[src_str].fillna("")
     src = list(df[src_str])
@@ -149,20 +148,24 @@ def extract_sources(df, src_str="Source"):
     df.drop(columns=[src_str], inplace=True)
     return sources
 
+
 def get_obj_str(lg):
     if lg[0] == "P":
         return "rc"
     else:
         return "obj"
 
+
 def add_obj_markdown(df, lg_col="Language_ID", form_col="Form"):
     df[form_col] = df.apply(
         lambda x: objectify(x[form_col], get_obj_str(x[lg_col])), axis=1
     )
 
+
 def repl_lg_id(df):
     df["Language_ID"] = df["Language_ID"].apply(print_shorthand)
     df.rename(columns={"Language_ID": "Language"}, inplace=True)
+
 
 # prepare overviews of class-switching 'go down' and 'defecate'
 print("\nClass membership of 'to go down':")
@@ -178,7 +181,8 @@ temp_df1 = calculate_alignment(temp_df1, fuzzy=True)
[…]

### 2021-10-22 documents/floats/go.tex
floats for 'go' and 'say'

\begin{table}
\centering
\caption[Reflexes of \rc{ɨtə(mə)} \qu{to go}]{Reflexes of \rc{ɨtə(mə)} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6; p.c., Spike Gildea]{franchetto2002kuikuro}}
\label{tab:go}
\begin{tabular}[t]{@{}lllllll@{}}
\toprule
Language &             Form & \multicolumn{5}{l}{Alignment} \\
\midrule
\kaxui    &     \obj{to(mo)} &           &  t &  o &  m &  o \\
\hixka    &      \obj{(ɨ)to} &         ɨ &  t &  o &    &    \\
\waiwai   &   \obj{(e)to(m)} &         e &  t &  o &  m &    \\
\arara    &        \obj{ɨdo} &         ɨ &  d &  o &    &    \\
\ikpeng   &       \obj{aran} &         a &  r &  a &  n &    \\
\ikpeng   &        \obj{ero} &         e &  r &  o &    &    \\
\bakairi  &      \obj{(ɨ)tə} &         ɨ &  t &  ə &    &    \\
\trio     &      \obj{tə(n)} &           &  t &  ə &  n &    \\
\akuriyo  &  \obj{(ə)tə(mɨ)} &         ə &  t &  ə &  m &  ɨ \\
\carijo   &       \obj{təmə} &           &  t &  ə &  m &  ə \\
\wayana   &   \obj{(ɨ)tə(m)} &         ɨ &  t &  ə &  m &    \\
\kalina   &         \obj{to} &           &  t &  o &    &    \\
\kalina   &       \obj{(ɨ)ʔ} &         ɨ &  ʔ &    &    &    \\
\maqui    &    \obj{ɨtə(mə)} &         ɨ &  t &  ə &  m &  ə \\
\ingariko &        \obj{ətə} &         ə &  t &  ə &    &    \\
\pemon    &      \obj{(e)tə} &         e &  t &  ə &    &    \\
\macushi  &      \obj{(a)tɨ} &         a &  t &  ɨ &    &    \\
\panare   &      \obj{tə(n)} &           &  t &  ə &  n &    \\
\yawarana &         \obj{tə} &           &  t &  ə &    &    \\
\mapoyo   &         \obj{tə} &           &  t &  ə &    &    \\
\uxc      &      \obj{(e)te} &         e &  t &  e &    &    \\
\yukpa    &         \obj{to} &           &  t &  o &    &    \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-22 documents/floats/say.tex
floats for 'go' and 'say'

\begin{table}
\centering
\caption[Reflexes of \rc{ka(ti)} \qu{to say}]{Reflexes of \rc{ka(ti)} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102; p.c., Spike Gildea]{mattei1994diccionario}}
\label{tab:say}
\begin{tabular}[t]{@{}llllll@{}}
\toprule
Language &          Form & \multicolumn{4}{l}{Alignment} \\
\midrule
\kaxui   &   \obj{ka(s)} &         k &  a &  s &    \\
\hixka   &   \obj{ka(h)} &         k &  a &  h &    \\
\waiwai  &   \obj{ka(s)} &         k &  a &  s &    \\
\arara   &      \obj{ke} &         k &  e &    &    \\
\ikpeng  &      \obj{ke} &         k &  e &    &    \\
\bakairi &      \obj{ke} &         k &  e &    &    \\
\trio    &      \obj{ka} &         k &  a &    &    \\
\akuriyo &      \obj{ka} &         k &  a &    &    \\
\carijo  &      \obj{ka} &         k &  a &    &    \\
\wayana  &   \obj{ka(i)} &         k &  a &    &  i \\
\apalai  &  \obj{ka(ʃi)} &         k &  a &  ʃ &  i \\
\kalina  &      \obj{ka} &         k &  a &    &    \\
\kapon   &      \obj{ka} &         k &  a &    &    \\
\pemon   &      \obj{ka} &         k &  a &    &    \\
\macushi &      \obj{ka} &         k &  a &    &    \\
\panare  &   \obj{ka(h)} &         k &  a &  h &    \\
\uxc     &      \obj{ki} &         k &  i &    &    \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-22 workflow/process_data.py

@@ -243,6 +243,7 @@ def print_aligned_table(verb, caption, fuzzy=False):
     fields = ["Language_ID", "Form", ""]
     df = v_df[v_df["Parameter_ID"] == verb]
     sources = extract_sources(df)
+    df["Segments"] = df.apply(lambda x: segmentify(x["Form"]), axis=1)
     df["Cognateset_ID"] = df["Cognateset_ID"].map(str2numcog)
     df = calculate_alignment(df, fuzzy=fuzzy)
     df = df[fields]
@@ -265,7 +266,8 @@ def print_aligned_table(verb, caption, fuzzy=False):
     )
 # df.to_csv("../data/comp_tables/come.csv", index=False)
 print_aligned_table("come", r"Reflexes of \qu{to come}", fuzzy=True)
-print_aligned_table("go", r"Reflexes of \rc{ka(ti)} \qu{to go}")
+print_aligned_table("go", r"Reflexes of \rc{ɨtə(mə)} \qu{to go}")
+print_aligned_table("say", r"Reflexes of \rc{ka(ti)} \qu{to say}")
 
 # come_aligned = pd.read_csv("../data/comp_tables/come.csv", keep_default_na=False)
 # # columns = pd.DataFrame(come_aligned.columns.tolist())

### 2021-10-22 workflow/lingpy_alignments.py
'to bathe' tables, extension-verb crosstabulation

@@ -59,7 +59,7 @@ def is_empty(slot):
 def join_slots(slotlist, almsep):
     out = [slotlist[0]]
     for slot in slotlist[1::]:
-        if is_empty(out[-1]):
+        if (is_empty(out[-1]) or is_empty(slot)):
             out.append("-")
         else:
             out.append("+")
@@ -86,7 +86,6 @@ def buffer_alignments(df, cogid="COGIDS", cogsep=" ", almid="ALIGNMENT", almsep=
     c[almid] = c[[x for x in order]].values.tolist()
     # c[almid] = c[almid].apply(lambda x: almsep.join(x))
     c[almid] = c[almid].apply(join_slots, almsep=almsep)
-    # print(c)
     df[almid] = c[almid]
     # print(df)
 

### 2021-10-22 workflow/process_data.py

@@ -139,13 +139,14 @@ def print_shorthand(abbrev):
     return "\\" + cah.get_shorthand(abbrev)
 
 
-def extract_sources(df, src_str="Source"):
+def extract_sources(df, src_str="Source", keep=False):
     df[src_str] = df[src_str].fillna("")
     src = list(df[src_str])
     if "" in src:
         src.remove("")
     sources = cldfh.cite_a_bunch(src, parens=True)
-    df.drop(columns=[src_str], inplace=True)
+    if not keep:
+        df.drop(columns=[src_str], inplace=True)
     return sources
 
 
@@ -239,10 +240,10 @@ save_float(
     short=r"Reflexes of \qu{to defecate} ",
 )
 
-def print_aligned_table(verb, caption, fuzzy=False):
+def print_aligned_table(df, verb="new_verb", caption="", fuzzy=False, only_tab=False, do_sources=True):
     fields = ["Language_ID", "Form", ""]
-    df = v_df[v_df["Parameter_ID"] == verb]
-    sources = extract_sources(df)
+    if do_sources:
+        sources = extract_sources(df)
     df["Segments"] = df.apply(lambda x: segmentify(x["Form"]), axis=1)
     df["Cognateset_ID"] = df["Cognateset_ID"].map(str2numcog)
     df = calculate_alignment(df, fuzzy=fuzzy)
@@ -253,47 +254,84 @@ def print_aligned_table(verb, caption, fuzzy=False):
             df.columns.values[i] = "Alignment"
             break
     df["Form"] = df["Form"].str.replace("+", "", regex=True)
+    print(f"Comparative table for 'to {verb}'")
     print(df)
     add_obj_markdown(df)
     repl_lg_id(df)
     df.set_index("Language", drop=True, inplace=True)
     tabular = print_latex(df, keep_index=True)
-    save_float(
-        tabular,
-        verb,
-        caption + " " + sources,
-        short=caption,
-    )
-# df.to_csv("../data/comp_tables/come.csv", index=False)
-print_aligned_table("come", r"Reflexes of \qu{to come}", fuzzy=True)
-print_aligned_table("go", r"Reflexes of \rc{ɨtə(mə)} \qu{to go}")
-print_aligned_table("say", r"Reflexes of \rc{ka(ti)} \qu{to say}")
+    if only_tab:
+        if do_sources:
+            return tabular, sources
+        else:
[…]

### 2021-10-22 workflow/process_data.py
Finished rewriting latex generation script

@@ -348,7 +348,7 @@ f.close()
 e_df = pd.read_csv("../data/extensions.csv")
 i_df = pd.read_csv("../data/inflection_data.csv")
 i_df = i_df[i_df["Inflection"] == "1"]
-verb_list = ["say", "go", "be-1", "be-2", "come", "go down", "bathe (INTR)"]
+verb_list = ["say", "go", "be_1", "be_2", "come", "go_down", "bathe_intr"]
 
 #determine whether verb was affected by extensions based on cognacy of prefixes
 def identify_affected(cogset, value):
@@ -403,7 +403,7 @@ for i, row in e_df.iterrows():
 pyd.x = ["Cognateset_ID"]
 pyd.y = ["Orig_Language", "Form", "Language_ID"]
 pyd.content_string = "Affected"
-# pyd.x_sort = verb_list
+pyd.x_sort = verb_list
 
 repl_dic = {
     "DETRZ+come": "come",
@@ -414,86 +414,82 @@ repl_dic = {
 overview["Cognateset_ID"] = overview["Verb_Cognateset_ID"].replace(repl_dic)
 #
 result = pyd.compose_paradigm(overview)
+result["Lg"] = pd.Categorical([lvl[2] for lvl in result.index.str.split(".")], lg_list)
+result["Orig"] = pd.Categorical(
+    [lvl[0] for lvl in result.index.str.split(".")], lg_list
+)
+result["Form"] = [lvl[1] for lvl in result.index.str.split(".")]
+temp_list = ["Orig", "Form", "Lg"]
+result.set_index(temp_list, inplace=True)
+result.sort_index(inplace=True, level="Lg")
+result.replace({"n/n": "n", "n/y": "(y)", "": "–"}, inplace=True)
+result.index.names = ["", "", ""]
+print("\nOverview of extensions and (un-)affected verbs:")
 print(result)
-# result["Lg"] = pd.Categorical([lvl[2] for lvl in result.index.str.split(".")], lg_list)
-# result["Orig"] = pd.Categorical(
-#     [lvl[0] for lvl in result.index.str.split(".")], lg_list
-# )
-# result["Form"] = [lvl[1] for lvl in result.index.str.split(".")]
-# temp_list = ["Orig", "Form", "Lg"]
-# result.set_index(temp_list, inplace=True)
[…]

### 2021-10-25 workflow/process_data.py
suppress pandas warnings

@@ -1,6 +1,6 @@
 import pyradigms as pyd
+import warnings; warnings.filterwarnings("ignore")
 import pandas as pd
-
 pd.options.mode.chained_assignment = None
 import cariban_helpers as cah
 import cldf_helpers as cldfh

### 2021-10-25 documents/floats/aku1sa.tex
new autogenerated tables

\begin{table}
\centering
\caption[\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes]{\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes \parencites[77, 79, 82, 84, 85, 86, 87]{gildea1994akuriyo}}
\label{tab:aku1sa}
\begin{tabular}[t]{@{}ll@{}}
\toprule
      first person \obj{k-} &               first person \obj{t͡ʃ-} \\
\midrule
 \obj{əempa-} \qu{to learn} &  \obj{epɨ-} \qu{to bathe (\gl{intr})} \\
\obj{əət͡ʃena-} \qu{to cry} &      \obj{ekɨrɨka-} \qu{to stay back} \\
\obj{əiwa-} \qu{to tremble} &            \obj{entapo-} \qu{to yawn} \\
 \obj{əməmɨ-} \qu{to enter} &       \obj{etonema-} \qu{to lie down} \\
\obj{ətajiŋka-} \qu{to run} &          \obj{ewai-} \qu{to sit down} \\
  \obj{əturu-} \qu{to talk} & \obj{ehpa-} \qu{to bathe (\gl{intr})} \\
   \obj{əʔnɨkɨ-} \qu{sleep} &                                       \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-25 documents/floats/ppekverbs.tex
new autogenerated tables

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[76, 130, 131, 137, 374, 397]{von1892bakairi}[68]{ikpengpacheco1997}[4; p.c., Angela Fabíola Alves Chagas]{meira2003bakairi}}
\label{tab:ppekverbs}
\begin{tabular}[t]{@{}lllll@{}}
\toprule
{} &          \PPek &          \arara &                       \ikpeng &        \bakairi \\
\midrule
\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &     \obj{u-ge-} \\
\qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &      \obj{w-i-} \\
\qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &      \obj{w-a-} \\
\qu{be-2}    &  \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &      \obj{w-i-} \\
\qu{come}    &    \rc{w-epɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &   \obj{k-əewɨ-} \\
\qu{go down} &   \rc{w-ɨptə-} &  \obj{w-iptoŋ-} &  \obj{{\normalfont ?}-iptoŋ-} &  \obj{k-ɨtəgɨ-} \\
\qu{go}      &   \rc{w-ɨtən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &     \obj{u-tə-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-25 documents/floats/ptirverbs.tex
new autogenerated tables

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112, 113, 114, 115, 165]{meira1998proto}}
\label{tab:ptirverbs}
\begin{tabular}[t]{@{}llll@{}}
\toprule
{} &          \PTir &          \trio &                     \akuriyo \\
\midrule
\qu{go}   &  \rc{wɨ-təmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
\qu{say}  &    \rc{wɨ-ka-} &   \obj{wɨ-ka-} &                 \obj{wɨ-ka-} \\
\qu{come} &  \rc{w-əʔepɨ-} &  \obj{w-əepɨ-} &  \obj{{\normalfont ∅}-eepɨ-} \\
\qu{be-1} &      \rc{w-a-} &     \obj{w-a-} &     \obj{{\normalfont ∅}-a-} \\
\qu{be-2} &    \rc{w-eʔi-} &    \obj{w-ei-} &   \obj{{\normalfont ?}-eʔi-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-25 documents/floats/pwaiverbs.tex
new autogenerated tables

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197, 198; p.c., Spike Gildea]{hixkaryanaderby1985}}
\label{tab:pwaiverbs}
\begin{tabular}[t]{@{}llll@{}}
\toprule
{} &        \PWai &        \hixka &         \waiwai \\
\midrule
\qu{say}  &  \rc{wɨ-ka-} &   \obj{ɨ-ka-} &    \obj{wɨ-ka-} \\
\qu{be-1} &   \rc{w-ah-} &   \obj{w-ah-} &      \obj{w-a-} \\
\qu{be-2} &  \rc{w-eʃi-} &  \obj{w-eʃe-} &   \obj{w-eeʃi-} \\
\qu{go}   &  \rc{wɨ-to-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-25 documents/floats/pxinw.tex
new autogenerated tables

\begin{table}
\centering
\caption[Loss of \rc{w} in \ikpeng]{Loss of \rc{w} in \ikpeng \parencites[44, 70]{souza1993arara}[118]{alves2013verbo}[143]{alves2017arara}[21, 164]{ikpengpacheco2001}[9]{desouza2010arara}[40]{campetela1997analise}}
\label{tab:pxinw}
\begin{tabular}[t]{@{}lll@{}}
\toprule
             Meaning &       \arara &     \ikpeng \\
\midrule
    \qu{to defecate} &  \obj{watke} &  \obj{atke} \\
       \qu{\gl{dat}} &   \obj{wɨna} &   \obj{ɨna} \\
            \qu{dog} & \obj{wokori} & \obj{akari} \\
\qu{capuchin monkey} &   \obj{tawe} &   \obj{tae} \\
          \qu{sleep} &  \obj{wɨnkɨ} &  \obj{ɨnkɨ} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-25 documents/sections/extensions.tex
new autogenerated tables

@@
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-While there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} forms are attested inflected for first person \parencite[401]{gildea2018reconstructing}.
+While there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} forms are attested with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
 
@@
 %
-While clearly not a regular sound change, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \exref{ikpwloss}.
+While clearly not a regular sound change, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \cref{tab:pxinw}.
 Thus, we suggest that the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}, and was subject to phonological erosion of \rc{w}, a phenomenon which will also be seen in other languages discussed in following sections.
 
-\ex<ikpwloss> %\arara-\ikpeng \obj{w}:Ø correspondences\\
-\begin{tabular}[t]{@{}lll@{}}
-\PXin & \arara & \ikpeng\\
-\rc{wɨna} \qu{\gl{dat}} & \obj{wɨna} & \obj{ɨna}\\
-\rc{wokori} \qu{dog, jaguar} & \obj{wokori} & \obj{akari}\\
-\rc{tawe} \qu{capuchin monkey} & \obj{tawe} & \obj{tae}\\
-\rc{watke} \qu{to defecate} & \obj{watke} & \obj{atke}\\
-\rc{wɨnkɨ} \qu{to sleep} & \obj{wɨnkɨ} & \obj{ɨnkɨ}\\
-& (\obj{o-wɨnkɨ} \qu{sleep!}) & (\obj{j-ɨnkɨ-lɨ} \qu{s/he slept})\\
-\end{tabular}\\
-\parencites[80]{alves2017arara}[164, 67, 150]{ikpengpacheco2001}[9]{desouza2010arara}[68]{souza1993arara}[91]{alves2013verbo}[40]{campetela1997analise} 
-\xe
+%\ex<ikpwloss> %\arara-\ikpeng \obj{w}:Ø correspondences\\
+%\begin{tabular}[t]{@{}lll@{}}
+%\PXin & \arara & \ikpeng\\
+%\rc{wɨna} \qu{\gl{dat}} & \obj{wɨna} & \obj{ɨna}\\
+%\rc{wokori} \qu{dog, jaguar} & \obj{wokori} & \obj{akari}\\
+%\rc{tawe} \qu{capuchin monkey} & \obj{tawe} & \obj{tae}\\
+%\rc{watke} \qu{to defecate} & \obj{watke} & \obj{atke}\\
+%\rc{wɨnkɨ} \qu{to sleep} & \obj{wɨnkɨ} & \obj{ɨnkɨ}\\
+%& (\obj{o-wɨnkɨ} \qu{sleep!}) & (\obj{j-ɨnkɨ-lɨ} \qu{s/he slept})\\
+%\end{tabular}\\
+%\parencites[80]{alves2017arara}[164, 67, 150]{ikpengpacheco2001}[9]{desouza2010arara}[68]{souza1993arara}[91]{alves2013verbo}[40]{campetela1997analise} 
+%\xe
 
-\begin{table}
-	\centering
-	\caption{Verbs preserving \gl{1}\gl{s_a_} in \PPek}
-	\label{tab:pekverbs}
-	\begin{tabular}{@{}lllll@{}}
-	\mytoprule
-& \PPek & \arara & \ikpeng & \bakairi \\
-\mytoprule
-\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} & \obj{u-ge-} \\
-\qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{Ø-ip-} & \obj{w-i-} \\
-\qu{be-1} & \rc{w-ap-} & \obj{w-ap-} & -- & \obj{w-a-} \\
-\qu{be-2} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{Ø-et͡ʃi-} & \obj{w-i-} \\
-\qu{come} & \rc{w-epɨ}, \rc{k-əd-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
-\qu{go down} & \rc{w-[ɨ/i]ptə-} & \obj{w-iptoŋ-} & ?\obj{-iptoŋ-} & \obj{k-ɨtəgɨ-} \\
-\qu{go} & \rc{w-ɨtən-} & \obj{w-ɨdo-} & \obj{k-aran-} & \obj{u-tə-} \\
-	\mybottomrule
-	\end{tabular}
-\end{table}
+\input{floats/pxinw}
 
-\Cref{tab:pekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with our \PPek reconstruction.
+\input{floats/ppekverbs}
+
+\Cref{tab:ppekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with our \PPek reconstruction.
 The form for \rc{ge} \qu{to say} is straightforward to reconstruct; the \bakairi change of \rc{wɨ-} to \obj{u-} is not unexpected given correspondences like \bakairi \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
@@
 \pex<hixɨ>
-\a<wai-108> \waiwai \parencite[][71]{waiwaihawkins1998}
+\a<wai-108> \waiwai \parencite[][71]{waiwaihawkins1998}\\
 \begingl
@@
 \endgl
-\a<hix-118> \hixka \parencite[][124]{hixkaryanaderby1985}
+\a<hix-118> \hixka \parencite[][124]{hixkaryanaderby1985}\\
 \begingl
@@
 
-There are three more verbs which did not take innovative \rc{k-} in \PWai.
-Two roots are part of the copula \rc{eʃi}/\obj{a(h)}, straightforwardly reconstructed in \exref{parpar.parcop}.
-The other is \qu{to go} \exref{parpar.pargo}, somewhat of a special case.
-
-\pex<parpar>
-\a<parcop> Parukotoan \qu{to be}\\
-\begin{tabular}[t]{@{}llll@{}}
-& \PPar & \hixka & \waiwai \\
-\gl{1}\gl{s_a_} & \rc{\emp{w-}eʃi-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eʃe-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eeʃi-}/\obj{\emp{w-}a-}\\
-\gl{2}\gl{s_a_} & \rc{m-eʃi-}/\obj{m-ah-} & \obj{m-eʃe-}/\obj{m-ah-} & \obj{m-eeʃi-}/\obj{m-a-}\\
-\gl{1+2}\gl{s_a_} & \rc{t-eʃi-}/\obj{t-ah-} & \obj{t-eʃe-}/\obj{t-ah-} & \obj{t͡ʃ-eeʃi-}/\obj{t-a-}\\
-\gl{3}\gl{s_a_} & \rc{n-eʃi-}/\obj{n-ah-} & \obj{n-eʃe-}/\obj{n-ah-} & \obj{n-eeʃi-}/\obj{n-a-}\\
-\end{tabular}\\
-(\cite[197--198]{hixkaryanaderby1985} and Spike Gildea, p.c.)
-\a<pargo> Parukotoan \qu{to go}\\
-\begin{tabular}[t]{@{}llll@{}}
-& \PPar & \hixka & \waiwai\\
-\gl{1}\gl{s_a_} & \rc{\emp{wɨ-}to-} & \obj{\emp{ɨ-}to-} & \obj{kɨ\emp{w-}to-}\\
-\gl{2}\gl{s_a_} & \rc{mɨ-to-} & \obj{mɨ-to-} & \obj{mɨɨ-to-}\\
-\gl{1+2}\gl{s_a_} & \rc{tɨt-to-} & \obj{tɨ-to-} & \obj{tɨ-to-}/\obj{tɨh-t͡ʃe-}\\
-\gl{3}\gl{s_a_} & \rc{nɨ-to-} & \obj{n-to-} & \obj{nɨɨ-to-}\\
-\end{tabular}\\
-(\cites[69--70, 211]{hixkaryanaderby1985}[179]{waiwaihawkins1998} and Spike Gildea, p.c.)
-\xe
-%
+There are three more verbs which did not take innovative \rc{k-} in \PWai; they are shown along with \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
+While the two roots for \qu{to be} are straightforwardly reconstructible, \qu{to go} is somewhat of a special case.
 While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \gl{1}\gl{s_a_} \obj{k-} with the old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
@@
 In any case, \hixka \qu{to go} was clearly not affected by the extension of \rc{k-}, allowing us to reconstruct a \PWai first person form \rc{wɨ-to-}.
+
+\input{floats/pwaiverbs}
+
+%\pex<parpar>
+%\a<parcop> Parukotoan \qu{to be}\\
+%\begin{tabular}[t]{@{}llll@{}}
+%& \PPar & \hixka & \waiwai \\
+%\gl{1}\gl{s_a_} & \rc{\emp{w-}eʃi-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eʃe-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eeʃi-}/\obj{\emp{w-}a-}\\
+%\gl{2}\gl{s_a_} & \rc{m-eʃi-}/\obj{m-ah-} & \obj{m-eʃe-}/\obj{m-ah-} & \obj{m-eeʃi-}/\obj{m-a-}\\
+%\gl{1+2}\gl{s_a_} & \rc{t-eʃi-}/\obj{t-ah-} & \obj{t-eʃe-}/\obj{t-ah-} & \obj{t͡ʃ-eeʃi-}/\obj{t-a-}\\
+%\gl{3}\gl{s_a_} & \rc{n-eʃi-}/\obj{n-ah-} & \obj{n-eʃe-}/\obj{n-ah-} & \obj{n-eeʃi-}/\obj{n-a-}\\
+%\end{tabular}\\
+%(\cite[197--198]{hixkaryanaderby1985} and Spike Gildea, p.c.)
+%\a<pargo> Parukotoan \qu{to go}\\
+%\begin{tabular}[t]{@{}llll@{}}
+%& \PPar & \hixka & \waiwai\\
+%\gl{1}\gl{s_a_} & \rc{\emp{wɨ-}to-} & \obj{\emp{ɨ-}to-} & \obj{kɨ\emp{w-}to-}\\
+%\gl{2}\gl{s_a_} & \rc{mɨ-to-} & \obj{mɨ-to-} & \obj{mɨɨ-to-}\\
+%\gl{1+2}\gl{s_a_} & \rc{tɨt-to-} & \obj{tɨ-to-} & \obj{tɨ-to-}/\obj{tɨh-t͡ʃe-}\\
+%\gl{3}\gl{s_a_} & \rc{nɨ-to-} & \obj{n-to-} & \obj{nɨɨ-to-}\\
+%\end{tabular}\\
+%(\cites[69--70, 211]{hixkaryanaderby1985}[179]{waiwaihawkins1998} and Spike Gildea, p.c.)
+%\xe
+%
 
@@
 %
-As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the first four verbs in \exref{ptirw} as being reconstructible to \PTar with \rc{w-}, for which we offer our reconstructed \PTir forms.
+As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the first four verbs in \cref{tab:ptirverbs} as being reconstructible to \PTar with \rc{w-}, for which we offer our reconstructed \PTir forms.
 To this list, we add the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}}, which has first person \obj{w-} at least in \trio.
 
-\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-} \parencites[112--115]{meira1998proto}[294]{triomeira1999}\\ %[85]{gildea1994akuriyo}
-\begin{tabular}[t]{@{}llll@{}}
-& \PTir & \trio & \akuriyo\\
-\qu{go} & \rc{wɨ-tə(mɨ)-} & \obj{wɨ-tə(mɨ)-} & \obj{ə-təmɨ-} \\% (\obj{wɨ-təemɨ-})\\
-\qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
-\qu{come} & \rc{w-əʔepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
-\qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
-& \rc{w-eʔi-} & \obj{w-ei-} & ?-\obj{eʔi-}\\
-\end{tabular}
+%\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-} \parencites[112--115]{meira1998proto}[294]{triomeira1999}\\ %[85]{gildea1994akuriyo}
+%\begin{tabular}[t]{@{}llll@{}}
+%& \PTir & \trio & \akuriyo\\
+%\qu{go} & \rc{wɨ-tə(mɨ)-} & \obj{wɨ-tə(mɨ)-} & \obj{ə-təmɨ-} \\% (\obj{wɨ-təemɨ-})\\
+%\qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
+%\qu{come} & \rc{w-əʔepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
+%\qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
+%& \rc{w-eʔi-} & \obj{w-ei-} & ?-\obj{eʔi-}\\
+%\end{tabular}
 %\a<aku12> \akuriyo \obj{kəʔ-eepɨ} \qu{we came} \parencite[114]{meira1998proto}
@@
 %\a<tricome> \trio \obj{n-epɨ} \qu{s/he came} \parencite[114]{meira1998proto}
-\xe
+%\xe
+
+\input{floats/ptirverbs}
 
@@
 
-\begin{table}
-	\centering
-	\caption{\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes}
-	\label{tab:aku1sa}
-	\begin{tabular}{@{}lll@{}}
-	\mytoprule
-first person \obj{k-} & first person \obj{t͡ʃ-} \\
-\mymidrule
-\obj{ənɨkɨ} \qu{to sleep} & \obj{eepɨ} \qu{to bathe} \\
-\obj{əməmɨ} \qu{to enter} & \obj{ewai} \qu{to sit down} \\
-\obj{əturu} \qu{to talk} & \obj{etonema} \qu{to lie down} \\
-\obj{əət͡ʃena} \qu{to cry} & \obj{ekɨɨrɨka} \qu{to stay back} \\
-\obj{ətajiŋka} \qu{to run} & \obj{entapo} \qu{to yawn} \\
-\obj{əiwa} \qu{to tremble} &  \\
-\obj{əempa} \qu{to learn} \\
-	\mybottomrule
-	\end{tabular}
-\end{table}
+%\begin{table}
+%	\centering
+%	\caption{\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes}
+%	\label{tab:aku1sa}
+%	\begin{tabular}{@{}lll@{}}
+%	\mytoprule
+%first person \obj{k-} & first person \obj{t͡ʃ-} \\
+%\mymidrule
+%\obj{ənɨkɨ} \qu{to sleep} & \obj{eepɨ} \qu{to bathe} \\
+%\obj{əməmɨ} \qu{to enter} & \obj{ewai} \qu{to sit down} \\
+%\obj{əturu} \qu{to talk} & \obj{etonema} \qu{to lie down} \\
+%\obj{əət͡ʃena} \qu{to cry} & \obj{ekɨɨrɨka} \qu{to stay back} \\
+%\obj{ətajiŋka} \qu{to run} & \obj{entapo} \qu{to yawn} \\
+%\obj{əiwa} \qu{to tremble} &  \\
+%\obj{əempa} \qu{to learn} \\
+%	\mybottomrule
+%	\end{tabular}
+%\end{table}
 
-%Let us first consider a scenario where \obj{k-} was the older \gl{1}\gl{s_a_} marker, occurring on both \obj{e}- and \obj{ə}-initial verbs.
-%Next, \obj{t-}/\obj{t͡ʃ-} would have come in and replaced \obj{k-} before \obj{e}, but (mostly) not before \obj{ə}, resulting in the observable distribution.
-%This scenario is not very satisfactory, since \obj{k-} is replaced in one completely unmotivated phonological environment, but not in the other.
-%
-%The other scenario is that \rc{t-} was the old first person marker occurring on the vast majority of \gl{s_a_} verbs.
-%Subsequently, it developed into \rc{t͡ʃ-} / \_\obj{e}, but remained \rc{t-} / \_\obj{ə}; \obj{t͡ʃ}\goodtilde\obj{ʃ}\goodtilde\obj{s} and \obj{t} are phonemically contrastive in \akuriyo \parencite[16]{meira1998proto}.
-%Thus, \obj{k-} can be seen as replacing \obj{t-}, but not \obj{t͡ʃ-}.
-%The occurrences of \obj{t-} / \_\obj{ə} mentioned by \citeauthor{meira1998proto} are then remnants of the earlier prevalence of \obj{t-}.
-%While this scenario also lacks a clear motivation, it at least explains the largely phonologically complementary distribution of \obj{k-} and \obj{t-} by the fact that the marker which was replaced had a phonemically different form than the one which was preserved.
-%The other scenario which sees \rc{k-} as older would require \rc{t-} replacing \obj{k-} only in a specific environment.\footnote{Note that this process would be plausible even if the instances of \obj{t-} / \_\obj{ə} were due to \trio influence, as suggested by \textcite[108]{meira1998proto} -- these would then be a later reintroduction, replacing \obj{k-}, instead of remnants of the introduction of it.}
-
-It has been speculated that the \trio \gl{s_a_} verb with a first person \obj{k-} marker, \obj{weka}/\obj{oeka} \qu{to defecate} might have something to do with \akuriyo \obj{k-}, which would potentially make the innovation of \rc{k-} a \PTir matter \parencite[116]{meira1998proto}.
-However, this hypothesis faces a problem in that the first person form for \akuriyo \qu{to defecate} is \obj{j-ereina-} \parencite[88]{gildea1994akuriyo}.
-Another hypothesis is that it originated in the corresponding \kaxui form \obj{ku-weka-} (form provided by Spike Gildea, p.c.), although the occurrence of \obj{o} in \trio would still need explanation.
-%Rather, it seems likely that this irregular first person form originated in a class switch from \gl{s_p_} to \gl{s_a_} (\Cref{sec:resistant_verbs}), although it is unclear how exactly that switch resulted in the irregular form.
-
-\subsection{\carijo \obj{j-}}
-\label{sec:carijo}
-As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since almost all \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} is attested.}
-This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \exref{carpar}.
-
-\ex<carpar> \carijo \parencites[106]{meira1998proto}[173]{robayo2000avance}\\
-\begin{tabular}[t]{@{}lll@{}}
-& \obj{tuda} \qu{to arrive} & \obj{eharaga} \qu{to dance}\\
-\gl{1} & \obj{ji-tuda-} & \obj{j-eharaga-}\\
-\gl{2} & \obj{mɨ-tuda-} & \obj{m-eharaga-}\\
-\gl{1+2} & \obj{kɨsi-tuda-} & \obj{kɨs-eharaga-}\\
-\gl{3} & \obj{ni-tuda-} & \obj{n-eharaga-}\\
-\end{tabular}
-\xe
-%
-Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by taking a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance} \exref{car-30}.
-%The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
-
-\ex<car-30>\carijo \parencite[][79]{koch1908hiana}\\
-\begingl
-\glpreamble hẹnẹ(x)tónoko-mā́ɽẹ y-e̱-hẹnẹ́(x)yai̯//
-\gla hɨnəhtono-ko=marə j-e-hɨnəhɨ-jai//
-\glb enemy-\gl{pl}=with \gl{1}-\gl{detrz}-kill-\gl{npst}.\gl{cert}//
-\glft \qu{I fight with the enemies.}//
-\endgl
-\xe
-%
-
-%\pex<car1p>\carijo \parencite[][59, 53]{guerrero2016carijo}
-%\begin{multicols}{2}
-%\a<car-1>
-%\begingl
-%\gla jɨ-hɨnəh-ɨ//
-%\glb \gl{3}>\gl{1}-kill-\gl{pfv}//
-%\glft \qu{S/he killed me.}//
-%\endgl
-%\a<car-31>
-%\begingl
-%\gla j-owo-rɨ//
-%\glb \gl{1}\gl{poss}-uncle-\gl{pert}//
-%\glft \qu{my uncle}//
-%\endgl
-%\end{multicols}
-%\xe
-%
-
-As noted in \cref{sec:taranoan}, this extension also erased any traces of a potential \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
-However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
-
-\pex<car1w>\carijo \parencite[][5, 42]{guerrero2016karihona}
-\a<car-24>
-\begingl
-\gla wɨ-tə-e=rehe//
-\glb \gl{1}-go-\gl{npst}=\gl{frust}//
-\glft \qu{I almost go (but I am not going to go).}//
-\endgl
-\a<car-25>
-\begingl
-\gla əji-marə-ne w-a-e//
-\glb \gl{2}-with-\gl{pl} \gl{1}-be-\gl{npst}//
-\glft \qu{I am with you all.}//
-\endgl
-\xe
-%
-%Interestingly enough, the verb \qu{to be} only shows \obj{w-} in its \obj{a} root allomorph, but not in the \obj{et͡ʃi} root allomorph, where it has \obj{j-} \exref{car1y.mivida-12}.
-%In a somewhat foreshadowing move, I will note that \PTar \rc{(ət)epɨ} \qu{to come} was also affected by the \carijo extension, as seen in \exref{car1y.car-18}.
-%As discussed in \cref{sec:taranoan}, this verb preserves \rc{w-} in \akuriyo and \trio \exref{tri-105}.
-%
-%\pex<car1y>\carijo
-%\a<mivida-12>
-%\begingl
-%\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
-%\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
-%\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
-%\glft \qu{Then I was already grown up.} \parencite[][177]{robayo1989rame}//
-%\endgl
-%\a<car-18>
-%\begingl
-%\gla əji-wa-e j-eh-ɨ//
-%\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
-%\glft \qu{I came looking for you.} \parencite[][102]{guerrero2019carijo}//
-%\endgl
-%\xe
-%
-%\ex<tri-105>\trio \parencite[][337]{triocarlin2004}\\
-%\begingl
-%\gla ene w-əe//
-%\glb see.\gl{sup} \gl{1}\gl{s_a_}-come//
-%\glft \qu{I came to see him/her.}//
-%\endgl
-%\xe
-
-\subsection{\yukpa \obj{j-}}
-\label{sec:yukpa}
-\yukpa has lost most of its \setone constructions, mostly employing innovative ones \parencite{meira2006syntactic}.
-It preserves the \setone prefixes in the immediate past, where the \gl{1}\gl{s_p_} marker appears to have been extended to \gl{s_a_} verbs, as evidenced by the paradigms in \exref{yukpaintr}.
-The loss of split-\gl{s} in favor of the \gl{1}\gl{s_p_} marker is accompanied by the \gl{2}\gl{s_a_} marker \obj{m(ɨ)-} displacing \gl{2}\gl{s_p_} \rc{ə(j)-}, and \gl{1+2} being lost as an inflectional value.
-%It has lost the split-\gl{s} system in general
-
-\ex<yukpaintr> \yukpa \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}\\
-\begin{tabular}[t]{@{}llll@{}}
-& \obj{otum} \qu{to wash self} & \obj{nɨ} \qu{to sleep} & \obj{ata} \qu{to fall}\\
-\gl{1} & \obj{j-otum-} & \obj{jɨ-nɨ-} & \obj{j-ata-}\\
-\gl{2} & \obj{m-otum-} & \obj{mɨ-nɨ-} & \obj{m-ata-}\\
-\gl{3} & \obj{n-otum-} & \obj{nɨ-nɨ-} & \obj{n-ata-}\\
-\end{tabular}
-\xe
-%
-The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of \detrz, like \obj{otum} \qu{to wash self} in \exref{yukpaintr}.
-The analysis that this is a reflex of the \gl{1}\gl{s_p_} marker \rc{u(j)-} is supported by the same form occurring in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.
-
-\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
-\a<yuk-5>
-\begingl
-\gla aw j-esare//
-\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
-\glft \qu{S/he saw me.}//
-\endgl
-\a<yuk-1>
-\begingl
-\gla aw {\normalfont ∅}-esare//
-\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
-\glft \qu{I saw it.}//
-\endgl
-\xe
-On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian,sec:akuriyo}, we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
-The same zero marking is attested for a single intransitive verb: \obj{to} \qu{to go} \exref{yuk-7}.
-Based on regular C-initial verbs taking \obj{jɨ-}, like \qu{to sleep} in \exref{yukpaintr}, it is clear that also C-initial \obj{to} \qu{to go} preserved the old \gl{1}\gl{s_a_} marker \rc{w(ɨ)-}, which was lost due to phonological erosion.
-
-\ex<yuk-7>\yukpa \parencite[][139]{meira2006syntactic}\\
-\begingl
-\gla aw {\normalfont ∅}-to//
-\glb \gl{1}\gl{pro} \gl{1}\gl{s_a_}-go//
-\glft \qu{I went.}//
-\endgl
-\xe
+\input{floats/aku1sa.tex}

### 2021-10-25 workflow/process_data.py

@@ -15,6 +15,22 @@ cognate_list = ["go", "say", "come", "be_1", "be_2", "go_down", "bathe_intr"]
 cs_df = pd.read_csv("../data/cognate_sets.csv")
 v_df = pd.read_csv("../data/verb_stem_data.csv")
 v_df.rename(columns={"Meaning": "Parameter_ID"}, inplace=True)
+e_df = pd.read_csv("../data/extensions.csv")
+i_df = pd.read_csv("../data/inflection_data.csv")
+i_df["Form"] = i_df["Form"].str.replace("+", "", regex=True)
+
+m_df = pd.read_csv("../data/cldf/parameters.csv")
+m_df["Description"] = m_df["Description"].str.strip("'").map(pynt.get_expex_code)
+mean_dic = dict(zip(m_df["ID"], m_df["Description"]))
+
+concept_dic = {
+    "DETRZ1+bathe_1": "bathe_intr",
+    "DETRZ+come": "come",
+    "DETRZ1": "bathe_intr"
+}
+
+i_df["Concept"] = i_df.apply(lambda x: concept_dic[x["Verb_Cognateset_ID"]] if x["Verb_Cognateset_ID"] in concept_dic else x["Verb_Cognateset_ID"], axis=1)
+
 segments = open("../data/segments.txt").read()
 segment_list = [{"Grapheme": x, "mapping": x} for x in segments.split("\n")]
 t = Tokenizer(Profile(*segment_list))
@@ -28,6 +44,15 @@ cognatesets.index = cognatesets.index + 1
 cognum = dict(zip(cognatesets["ID"], cognatesets.index.astype(str)))
 numcog = dict(zip(cognatesets.index.astype(str), cognatesets["ID"]))
 
+# cs_df["Gloss"] = cs_df["Meaning"].replace(" ", ".")
+cog_verb_df = cs_df[~(cs_df["Form"].str.contains("-"))]
+cog_trans_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Meaning"]))
+cog_form_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Form"]))
+cog_form_dic["bathe_intr"] = "e-pɨ"
+cog_form_dic["come"] = "(ət-)jəpɨ"
+cog_trans_dic["bathe_intr"] = r"bathe"
+cog_trans_dic["be_1"] = r"be-1"
+cog_trans_dic["be_2"] = r"be-2"
 
 def str2numcog(cogsets):
[…]

### 2021-10-25 workflow/compile_cldf_irregular.py
various improvements on the CLDF dataset

@@ -15,6 +15,7 @@ cogsets = pd.read_csv("../data/cognate_sets.csv")
 examples = pd.read_csv("../data/examples.csv")
 v_forms = v_forms[v_forms["Inflection"].isin(["1"])]
 v_forms.reset_index(inplace=True, drop=True)
+lex_df = pd.read_csv("../data/other_lexemes.csv")
 
 print("Compiling meanings")
 bathe["Meaning"] = bathe["Transitivity"].apply(
@@ -70,8 +71,14 @@ form_meanings.rename(columns={"Parameter_ID": "ID"}, inplace=True)
 form_meanings.drop(columns=["Inflection", "Meaning"], inplace=True)
 form_meanings.drop_duplicates(subset=["ID"], inplace=True)
 
+lex_meanings = lex_df[["Meaning"]]
+lex_meanings["Name"] = lex_meanings["Meaning"].str.replace(" ", ".")
+lex_meanings["Description"] = "'" + lex_meanings["Meaning"] + "'"
+lex_meanings["ID"] = lex_meanings["Meaning"].str.replace(" ", "_")
+lex_meanings.drop(columns=["Meaning"], inplace=True)
+lex_meanings.drop_duplicates(subset=["ID"], inplace=True)
 
-meanings = verb_meanings.append(form_meanings)
+meanings = verb_meanings.append(form_meanings).append(lex_meanings)
 
 print("Processing examples")
 # prepare examples
@@ -119,8 +126,10 @@ def split_cogset_ids(
 ):
     out = []
     for i, row in df[[id, cogset_col]].iterrows():
+        if pd.isnull(row[cogset_col]): continue
         if sep in row[cogset_col]:
             for j, cog in enumerate(row[cogset_col].split(sep)):
+                if cog == "?": continue
                 out.append(
                     {
                         "ID": row[id] + "_" + bare + "_" + suffixes[j],
@@ -130,6 +139,7 @@ def split_cogset_ids(
                     }
                 )
         else:
+            if row[cogset_col] == "?": continue
             out.append(
                 {
                     "ID": row[id] + "_" + bare,
@@ -158,7 +168,7 @@ cogsets["Description"] = (
     + cogsets["Form"].astype(str)
     + " '"
     + cogsets["Meaning"].astype(str)
[…]

### 2021-10-25 documents/sections/extensions.tex
fixed proto-pekodian 'to go'

@@
 For \arara, \textcite{alves2017arara} reports six\footnote{Seven under her analysis, which treats the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.} \gl{s_a_} verbs which have a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
-We take these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.
+%We take these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.
 
@@
 
-\input{floats/aku1sa.tex}
+\input{floats/aku1sa}
+
+%Let us first consider a scenario where \obj{k-} was the older \gl{1}\gl{s_a_} marker, occurring on both \obj{e}- and \obj{ə}-initial verbs.
+%Next, \obj{t-}/\obj{t͡ʃ-} would have come in and replaced \obj{k-} before \obj{e}, but (mostly) not before \obj{ə}, resulting in the observable distribution.
+%This scenario is not very satisfactory, since \obj{k-} is replaced in one completely unmotivated phonological environment, but not in the other.
+%
+%The other scenario is that \rc{t-} was the old first person marker occurring on the vast majority of \gl{s_a_} verbs.
+%Subsequently, it developed into \rc{t͡ʃ-} / \_\obj{e}, but remained \rc{t-} / \_\obj{ə}; \obj{t͡ʃ}\goodtilde\obj{ʃ}\goodtilde\obj{s} and \obj{t} are phonemically contrastive in \akuriyo \parencite[16]{meira1998proto}.
+%Thus, \obj{k-} can be seen as replacing \obj{t-}, but not \obj{t͡ʃ-}.
+%The occurrences of \obj{t-} / \_\obj{ə} mentioned by \citeauthor{meira1998proto} are then remnants of the earlier prevalence of \obj{t-}.
+%While this scenario also lacks a clear motivation, it at least explains the largely phonologically complementary distribution of \obj{k-} and \obj{t-} by the fact that the marker which was replaced had a phonemically different form than the one which was preserved.
+%The other scenario which sees \rc{k-} as older would require \rc{t-} replacing \obj{k-} only in a specific environment.\footnote{Note that this process would be plausible even if the instances of \obj{t-} / \_\obj{ə} were due to \trio influence, as suggested by \textcite[108]{meira1998proto} -- these would then be a later reintroduction, replacing \obj{k-}, instead of remnants of the introduction of it.}
+
+It has been speculated that the \trio \gl{s_a_} verb with a first person \obj{k-} marker, \obj{weka}/\obj{oeka} \qu{to defecate} might have something to do with \akuriyo \obj{k-}, which would potentially make the innovation of \rc{k-} a \PTir matter \parencite[116]{meira1998proto}.
+However, this hypothesis faces a problem in that the first person form for \akuriyo \qu{to defecate} is \obj{j-ereina-} \parencite[88]{gildea1994akuriyo}.
+Another hypothesis is that it originated in the corresponding \kaxui form \obj{ku-weka-} (form provided by Spike Gildea, p.c.), although the occurrence of \obj{o} in \trio would still need explanation.
+%Rather, it seems likely that this irregular first person form originated in a class switch from \gl{s_p_} to \gl{s_a_} (\Cref{sec:resistant_verbs}), although it is unclear how exactly that switch resulted in the irregular form.
+
+\subsection{\carijo \obj{j-}}
+\label{sec:carijo}
+As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since almost all \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} is attested.}
+This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \exref{carpar}.
+
+\ex<carpar> \carijo \parencites[106]{meira1998proto}[173]{robayo2000avance}\\
+\begin{tabular}[t]{@{}lll@{}}
+& \obj{tuda} \qu{to arrive} & \obj{eharaga} \qu{to dance}\\
+\gl{1} & \obj{ji-tuda-} & \obj{j-eharaga-}\\
+\gl{2} & \obj{mɨ-tuda-} & \obj{m-eharaga-}\\
+\gl{1+2} & \obj{kɨsi-tuda-} & \obj{kɨs-eharaga-}\\
+\gl{3} & \obj{ni-tuda-} & \obj{n-eharaga-}\\
+\end{tabular}
+\xe
+%
+Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by taking a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance} \exref{car-30}.
+%The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
+
+\ex<car-30>\carijo \parencite[][79]{koch1908hiana}\\
+\begingl
+\glpreamble hẹnẹ(x)tónoko-mā́ɽẹ y-e̱-hẹnẹ́(x)yai̯//
+\gla hɨnəhtono-ko=marə j-e-hɨnəhɨ-jai//
+\glb enemy-\gl{pl}=with \gl{1}-\gl{detrz}-kill-\gl{npst}.\gl{cert}//
+\glft \qu{I fight with the enemies.}//
+\endgl
+\xe
+%
+
+%\pex<car1p>\carijo \parencite[][59, 53]{guerrero2016carijo}
+%\begin{multicols}{2}
+%\a<car-1>
+%\begingl
+%\gla jɨ-hɨnəh-ɨ//
+%\glb \gl{3}>\gl{1}-kill-\gl{pfv}//
+%\glft \qu{S/he killed me.}//
+%\endgl
+%\a<car-31>
+%\begingl
+%\gla j-owo-rɨ//
+%\glb \gl{1}\gl{poss}-uncle-\gl{pert}//
+%\glft \qu{my uncle}//
+%\endgl
+%\end{multicols}
+%\xe
+%
+
+As noted in \cref{sec:taranoan}, this extension also erased any traces of a potential \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
+However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
+
+\pex<car1w>\carijo \parencite[][5, 42]{guerrero2016karihona}
+\a<car-24>
+\begingl
+\gla wɨ-tə-e=rehe//
+\glb \gl{1}-go-\gl{npst}=\gl{frust}//
+\glft \qu{I almost go (but I am not going to go).}//
+\endgl
+\a<car-25>
+\begingl
+\gla əji-marə-ne w-a-e//
+\glb \gl{2}-with-\gl{pl} \gl{1}-be-\gl{npst}//
+\glft \qu{I am with you all.}//
+\endgl
+\xe
+%
+%Interestingly enough, the verb \qu{to be} only shows \obj{w-} in its \obj{a} root allomorph, but not in the \obj{et͡ʃi} root allomorph, where it has \obj{j-} \exref{car1y.mivida-12}.
+%In a somewhat foreshadowing move, I will note that \PTar \rc{(ət)epɨ} \qu{to come} was also affected by the \carijo extension, as seen in \exref{car1y.car-18}.
+%As discussed in \cref{sec:taranoan}, this verb preserves \rc{w-} in \akuriyo and \trio \exref{tri-105}.
+%
+%\pex<car1y>\carijo
+%\a<mivida-12>
+%\begingl
+%\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
+%\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
+%\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
+%\glft \qu{Then I was already grown up.} \parencite[][177]{robayo1989rame}//
+%\endgl
+%\a<car-18>
+%\begingl
+%\gla əji-wa-e j-eh-ɨ//
+%\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
+%\glft \qu{I came looking for you.} \parencite[][102]{guerrero2019carijo}//
+%\endgl
+%\xe
+%
+%\ex<tri-105>\trio \parencite[][337]{triocarlin2004}\\
+%\begingl
+%\gla ene w-əe//
+%\glb see.\gl{sup} \gl{1}\gl{s_a_}-come//
+%\glft \qu{I came to see him/her.}//
+%\endgl
+%\xe
+
+\subsection{\yukpa \obj{j-}}
+\label{sec:yukpa}
+\yukpa has lost most of its \setone constructions, mostly employing innovative ones \parencite{meira2006syntactic}.
+It preserves the \setone prefixes in the immediate past, where the \gl{1}\gl{s_p_} marker appears to have been extended to \gl{s_a_} verbs, as evidenced by the paradigms in \exref{yukpaintr}.
+The loss of split-\gl{s} in favor of the \gl{1}\gl{s_p_} marker is accompanied by the \gl{2}\gl{s_a_} marker \obj{m(ɨ)-} displacing \gl{2}\gl{s_p_} \rc{ə(j)-}, and \gl{1+2} being lost as an inflectional value.
+%It has lost the split-\gl{s} system in general
+
+\ex<yukpaintr> \yukpa \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}\\
+\begin{tabular}[t]{@{}llll@{}}
+& \obj{otum} \qu{to wash self} & \obj{nɨ} \qu{to sleep} & \obj{ata} \qu{to fall}\\
+\gl{1} & \obj{j-otum-} & \obj{jɨ-nɨ-} & \obj{j-ata-}\\
+\gl{2} & \obj{m-otum-} & \obj{mɨ-nɨ-} & \obj{m-ata-}\\
+\gl{3} & \obj{n-otum-} & \obj{nɨ-nɨ-} & \obj{n-ata-}\\
+\end{tabular}
+\xe
+%
+The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of \detrz, like \obj{otum} \qu{to wash self} in \exref{yukpaintr}.
+The analysis that this is a reflex of the \gl{1}\gl{s_p_} marker \rc{u(j)-} is supported by the same form occurring in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.
+
+\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
+\a<yuk-5>
+\begingl
+\gla aw j-esare//
+\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
+\glft \qu{S/he saw me.}//
+\endgl
+\a<yuk-1>
+\begingl
+\gla aw {\normalfont ∅}-esare//
+\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
+\glft \qu{I saw it.}//
+\endgl
+\xe
+On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian,sec:akuriyo}, we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
+The same zero marking is attested for a single intransitive verb: \obj{to} \qu{to go} \exref{yuk-7}.
+Based on regular C-initial verbs taking \obj{jɨ-}, like \qu{to sleep} in \exref{yukpaintr}, it is clear that also C-initial \obj{to} \qu{to go} preserved the old \gl{1}\gl{s_a_} marker \rc{w(ɨ)-}, which was lost due to phonological erosion.
+
+\ex<yuk-7>\yukpa \parencite[][139]{meira2006syntactic}\\
+\begingl
+\gla aw {\normalfont ∅}-to//
+\glb \gl{1}\gl{pro} \gl{1}\gl{s_a_}-go//
+\glft \qu{I went.}//
+\endgl
+\xe

### 2021-10-26 documents/sections/extensions.tex
rewriting of extensions section

@@
 \cref{sec:waiwaian} takes a closer look at the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
-\cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir or \PTar, reflected in modern \trio and \akuriyo.
+\cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir (or \PTar), reflected in modern \trio and \akuriyo.
 \crefrange{sec:akuriyo}{sec:yukpa} look at innovative first person markers which are only attested in single modern languages:
@@
 
-For \arara, \textcite{alves2017arara} reports six\footnote{Seven under her analysis, which treats the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.} \gl{s_a_} verbs which have a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
+For \arara, \textcite{alves2017arara} describes six\footnote{Seven under her analysis, which sees the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.} \gl{s_a_} verbs which have a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
+From a comparative view this list is not quite complete, as there is also a reflex of the copula \rc{a[p]}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
+However, its inflectional morphology features verbal \setone prefixes, including first person \obj{w-} \exref{araap}.
 %We take these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.
 
+\begin{multicols}{2}
 \ex<arairr> \arara \parencite[153]{alves2017arara} \\
@@
 %
-From a comparative view, this list is not quite complete, as there is also a reflex of the copula \rc{a[p]}, which is analayzed as a postposition by \textcite[199--201]{alves2017arara}; it is however inflected with \setone prefixes and also shows first person \obj{w-} \exref{araap}.
-
-\ex<araap>
+\ex<araap> \arara \parencite[200]{alves2017arara}\\
 \begin{tabular}[t]{@{}ll@{}}
@@
 \xe
+\end{multicols}
 
@@
 Luckily, while \textcite{von1892bakairi} did not accurately record all phonemic distinctions in \bakairi \parencite{meira2005bakairi}, he does provide inflected forms of cognates to the \arara verbs in \exref{arairr}.
-We present them according to the phonemic analysis of \bakairi by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.
+We present them according to our understanding of the phonemic analysis of \bakairi by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.
 
@@
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-While there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} forms are attested with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
+Further, while there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} forms are attested with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
 
@@
 \xe
-
+%
 %\ex<ikp-182>\ikpeng \parencite[][320]{chagas2019ikpeng}\\
@@
 %
-While clearly not a regular sound change, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \cref{tab:pxinw}.
-Thus, we suggest that the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}, and was subject to phonological erosion of \rc{w}, a phenomenon which will also be seen in other languages discussed in following sections.
 
@@
 
-\input{floats/pxinw}
-
 \input{floats/ppekverbs}
@@
 \Cref{tab:ppekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with our \PPek reconstruction.
-The form for \rc{ge} \qu{to say} is straightforward to reconstruct; the \bakairi change of \rc{wɨ-} to \obj{u-} is not unexpected given correspondences like \bakairi \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
-\qu{to bathe} is straightforwardly reconstructible as \rc{ipɨ}.
+The presence of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests to us that it may be cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
+Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \cref{tab:pxinw}.
+While it is by no means a regular sound change, it allows us to securely connect the two prefixes.
+Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+Thus, we reconstruct a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} to \PPek, identical to \arara.
+
+As for the forms of the verb stems, a few comments are necessary:
 For \qu{to be}, \ikpeng \obj{e} is very likely the original vowel, given the \PC form \rc{eti} \pcref{sec:be}.
 
-The forms for \qu{to come} are not fully cognate; \ikpeng and \bakairi both show a reflex of the detransitivizer in combination with a root reconstructible as \rc{epɨ}.
+For \qu{to go down}, we reconstruct \rc{ɨ} as the initial vowel rather than \rc{i}, see \cref{sec:godown}.
+Further, the forms are not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
+However, the addition of a final \obj{ŋ} in \PXin is attested elsewhere: \begin{inlinelist}
+\item \PC \rc{əne} \qu{to see}, \arara and \ikpeng \obj{eneŋ} %\parencites[8]{gildea2007greenberg}[56]{alves2017arara}[25]{ikpengpacheco2001}
+\item \PC \rc{əta} \qu{to hear}, \arara \obj{taŋ}, \ikpeng \obj{iraŋ} %\parencites[8]{gildea2007greenberg}[144]{alves2017arara}[270]{ikpengpacheco2001}
+\item \PC \rc{ənə} \qu{to eat meat}, \arara \obj{oŋoŋ} \qu{to bite} %\parencites[8]{gildea2007greenberg}[57]{alves2017arara}
+\end{inlinelist} \parencites[8]{gildea2007greenberg}[56, 144, 57]{alves2017arara}[25, 270]{ikpengpacheco2001}.
+Also, based on the fact \bakairi has generally lost much segmental material, we suggest that \bakairi \obj{gɨ} is another addition to a root \rc{ɨptə}, rather than a conservative form.
+
+\input{floats/pxinw}
+
+The forms for \qu{to come} are not fully cognate, either: \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
 In contrast, the \arara first person form is directly based on this root \rc{epɨ}.
-However, other person values show a reflex of \detrz in \arara, too \exref{ara-123}.
+However, other person values show a reflex of \rc{əd-} in \arara, too \exref{ara-123}.
 
@@
 On the other hand, both \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm.
+Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-ebɨ}.
 
-The forms for \qu{to go down} are also not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
-Rather, it seems that some roots gained a final \obj{ŋ} in \PXin, for so far unclear reasons: \begin{inlinelist}
-\item \PC \rc{əne} \qu{to see}, \arara and \ikpeng \obj{eneŋ} \parencites[8]{gildea2007greenberg}[56]{alves2017arara}[25]{ikpengpacheco2001}
-\item \PC \rc{əta} \qu{to hear}, \arara \obj{taŋ}, \ikpeng \obj{iraŋ} \parencites[8]{gildea2007greenberg}[144]{alves2017arara}[270]{ikpengpacheco2001}
-\item \PC \rc{ənə} \qu{to eat meat}, \arara \obj{oŋoŋ} \qu{to bite} \parencites[8]{gildea2007greenberg}[57]{alves2017arara}
+The V-initial nature of \qu{to go} is evidenced in its Xinguan forms, while \bakairi \phon{\rc{wɨ}}{\obj{u}} resulted in a shift of the morpheme boundary, see also \cref{sec:go}.
+The \ikpeng form \obj{aran} is made compatible with our reconstruction \rc{ɨtən} by considering that \ikpeng \obj{a} is an attested outcome of \rc{ə} (\PXin \rc{o}/\obj{e}), for example \begin{inlinelist}
+ \item \obj{akari} \qu{dog} in \cref{tab:pxinw} above
+ \item \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg
+ \item \obj{jaj} \qu{tree} \parencite[98]{ikpengpacheco2001} from \PC \rc{jəje}
  \end{inlinelist}.
-Also, based on the fact \bakairi has generally lost much segmental material, we suggest that \bakairi \obj{gɨ} is yet another addition to a root \rc{ɨptə} or \rc{iptə}, rather than a conservative form.
-
-The forms for \qu{to go} are all cognate, but while \arara and \bakairi point to \rc{ɨtə}, the \ikpeng form seems rather divergent.
-However, it is in fact compatible with a reconstruction \rc{ɨtən}, when one considers that \ikpeng \obj{a} is an attested outcome of \rc{ə} (\PXin \rc{o}/\obj{e}), for example \obj{akari} \qu{dog} in \exref{ikpwloss} above, \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg}, or \obj{jaj} \qu{tree} from \PC \rc{jəje}.
 This attested change of \rc{ə} to \obj{a} need only be preceded by a assimilatory lowering of initial \rc{ɨ} to \rc{ə}, to yield the form \obj{aran} from \rc{ɨtən}.
-Other \ikpeng reflexes of \qu{to go} offer evidence for the suggested intermediate stage \rc{ətən}: \obj{ero-lɨ} \qu{s/he went} \parencite[25]{ikpengpacheco2001}.
+Other \ikpeng reflexes of \qu{to go} offer evidence for the suggested intermediate stage \rc{\underline{ə}t\underline{ə}n}: \obj{\underline{e}r\underline{o}-lɨ} \qu{s/he went} \parencite[25]{ikpengpacheco2001}.
 
@@
 Seven verbs can be reconstructed as having resisted this innovation, and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
-For \qu{to come}, two distinct forms \rc{w-epɨ-} and \rc{k-əd-epɨ-} can be reconstructed, although the latter may be two parallel developments in \bakairi and \ikpeng.
 In later, individual developments, \bakairi introduced \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
+Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ} and introduced \obj{k-}.
 
@@
 Its \hixka counterpart also takes a reflex of \gl{1}\gl{s_a_} \rc{w-}, which has the form \obj{ɨ-} \exref{hixɨ.hix-118}.
-A formally identical prefix is found on transitive verbs in \gl{1}>\gl{3} scenarios \exref{hixɨ.hix-17}, where it also corresponds to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.
+A formally identical prefix is found on transitive verbs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, where it also corresponds to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.
 
@@
 %
-Thus, similar to the the case of \ikpeng, \hixka has lost the \rc{w} of the \envr{}{C} allomorph \rc{wɨ-}.
+This correspondence allows us to establish \hixka \obj{ɨ-} a reflex of \rc{wɨ-}, similar to \ikpeng \pcref{sec:pekodian}.
 Notably, \textcite{hixkaryanaderby1985} does not see this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_}, but as the regular \gl{1}>\gl{3} marker, because he considers \hixka \obj{ka} \qu{to say} to be transitive, an issue which will be discussed in \cref{sec:say}.
 
-There are three more verbs which did not take innovative \rc{k-} in \PWai; they are shown along with \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
+There are three more verbs which did not take innovative \rc{k-} in \PWai, shown alongside \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
 While the two roots for \qu{to be} are straightforwardly reconstructible, \qu{to go} is somewhat of a special case.
@@
 %
-In any case, \hixka \qu{to go} was clearly not affected by the extension of \rc{k-}, allowing us to reconstruct a \PWai first person form \rc{wɨ-to-}.
+In any case, \hixka \qu{to go} was clearly not affected by the extension of \rc{k-}, allowing us to reconstruct a \PWai first person form \rc{wɨ-tom-}.
 
@@
 
-Summing up, we reconstruct the three verbs \rc{et͡ʃi}/\obj{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{to} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
+Summing up, we reconstruct the four verbs \rc{et͡ʃi} and \rc{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{tom} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
 

### 2021-10-26 workflow/compile_cldf_irregular.py
incorporated regular verbs

@@ -77,6 +77,7 @@ lex_meanings["Description"] = "'" + lex_meanings["Meaning"] + "'"
 lex_meanings["ID"] = lex_meanings["Meaning"].str.replace(" ", "_")
 lex_meanings.drop(columns=["Meaning"], inplace=True)
 lex_meanings.drop_duplicates(subset=["ID"], inplace=True)
+lex_meanings = lex_meanings[~(lex_meanings["ID"].isin(verb_meanings["ID"]))]
 
 meanings = verb_meanings.append(form_meanings).append(lex_meanings)
 

### 2021-10-26 workflow/process_data.py

@@ -1,6 +1,9 @@
 import pyradigms as pyd
-import warnings; warnings.filterwarnings("ignore")
+import warnings
+
+warnings.filterwarnings("ignore")
 import pandas as pd
+
 pd.options.mode.chained_assignment = None
 import cariban_helpers as cah
 import cldf_helpers as cldfh
@@ -22,14 +25,20 @@ i_df["Form"] = i_df["Form"].str.replace("+", "", regex=True)
 m_df = pd.read_csv("../data/cldf/parameters.csv")
 m_df["Description"] = m_df["Description"].str.strip("'").map(pynt.get_expex_code)
 mean_dic = dict(zip(m_df["ID"], m_df["Description"]))
+trans_dic = {x: "\\qu{%s}" % y for x, y in mean_dic.items()}
 
 concept_dic = {
     "DETRZ1+bathe_1": "bathe_intr",
     "DETRZ+come": "come",
-    "DETRZ1": "bathe_intr"
+    "DETRZ1": "bathe_intr",
 }
 
-i_df["Concept"] = i_df.apply(lambda x: concept_dic[x["Verb_Cognateset_ID"]] if x["Verb_Cognateset_ID"] in concept_dic else x["Verb_Cognateset_ID"], axis=1)
+i_df["Concept"] = i_df.apply(
+    lambda x: concept_dic[x["Verb_Cognateset_ID"]]
+    if x["Verb_Cognateset_ID"] in concept_dic
+    else x["Verb_Cognateset_ID"],
+    axis=1,
+)
 
 segments = open("../data/segments.txt").read()
 segment_list = [{"Grapheme": x, "mapping": x} for x in segments.split("\n")]
@@ -53,6 +62,7 @@ cog_form_dic["come"] = "(ət-)jəpɨ"
 cog_trans_dic["bathe_intr"] = r"bathe"
 cog_trans_dic["be_1"] = r"be-1"
 cog_trans_dic["be_2"] = r"be-2"
+
 
 def str2numcog(cogsets):
     return " ".join([cognum[x] for x in cogsets.split("+")])
@@ -84,7 +94,7 @@ def print_latex(df, ex=False, keep_index=False):
         del lines[-2]
         return "\n".join(lines)
     else:
-        return "\n".join(lines).replace("\toprule", "\mytoprule")
+        return "\n".join(lines).replace(r"\toprule", "\mytoprule")
 
 
[…]

### 2021-10-26 documents/floats/pekreg.tex
regular verbs in data and docs

\begin{table}
\centering
\caption[Regular Pekodian \gl{s_a_} verbs]{Regular Pekodian \gl{s_a_} verbs \textcites[4]{meira2003bakairi}[150]{alves2017arara}[52]{ikpengpacheco2001}}
\label{tab:pekreg}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &        \bakairi &                       \arara &                        \ikpeng \\
{} &   \qu{to go up} &                \qu{to dance} &                    \qu{to run} \\
\midrule
\gl{1}   &    \obj{k-əku-} &               \obj{k-origu-} &                \obj{k-aranme-} \\
\gl{2}   &    \obj{m-əku-} &               \obj{m-origu-} &                \obj{m-aranme-} \\
\gl{1+2} &  \obj{kɨd-əku-} &             \obj{kud-origu-} &               \obj{kw-aranme-} \\
\gl{3}   &    \obj{n-əku-} &  \obj{{\normalfont ∅}-origu} &  \obj{{\normalfont ∅}-aranme-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-26 documents/floats/ptirreg.tex
regular verbs in data and docs

\begin{table}
\centering
\caption[Regular \PTir \gl{s_a_} verbs]{Regular \PTir \gl{s_a_} verbs \parencites[292, 294]{triomeira1999}[87]{gildea1994akuriyo}}
\label{tab:ptirreg}
\begin{tabular}[t]{@{}lllllll@{}}
\mytoprule
{} & \multicolumn{3}{l}{\qu{to bathe (\gl{intr})}} & \multicolumn{3}{l}{\qu{to sleep}} \\
{} &                     \PTir &          \trio &        \akuriyo &             \PTir &              \trio &           \akuriyo \\
\midrule
\gl{1}   &             \rc{t͡ʃ-epɨ-} &   \obj{s-epɨ-} &  \obj{t͡ʃ-epɨ-} &    \rc{t-əənɨkɨ-} &    \obj{t-əənɨkɨ-} &    \obj{k-əənɨkɨ-} \\
\gl{2}   &               \rc{m-epɨ-} &   \obj{m-epɨ-} &    \obj{m-epɨ-} &    \rc{m-əənɨkɨ-} &    \obj{m-əənɨkɨ-} &    \obj{m-əənɨkɨ-} \\
\gl{1+2} &              \rc{ke-epɨ-} &  \obj{ke-epɨ-} &   \obj{ke-epɨ-} &  \rc{kɨt-əənɨkɨ-} &  \obj{kɨt-əənɨkɨ-} &  \obj{kəʔ-əənɨkɨ-} \\
\gl{3}   &               \rc{n-epɨ-} &   \obj{n-epɨ-} &    \obj{n-epɨ-} &    \rc{n-əənɨkɨ-} &    \obj{n-əənɨkɨ-} &    \obj{n-əənɨkɨ-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-26 documents/floats/pwaireg.tex
regular verbs in data and docs

\begin{table}
\centering
\caption[Regular \PWai verbs]{Regular \PWai verbs \parencites[30]{waiwaihawkins1998}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
\label{tab:pwaireg}
\begin{tabular}[t]{@{}lllllll@{}}
\mytoprule
{} & \multicolumn{3}{l}{\qu{to fall} (\gl{s_a_})} & \multicolumn{3}{l}{\qu{to sleep} (\gl{s_p_})} \\
{} &                    \PWai &           \hixka &            \waiwai &                     \PWai &          \hixka &            \waiwai \\
\midrule
\gl{1}   &           \rc{k-eɸurka-} &  \obj{k-ehurka-} &    \obj{k-eɸɨrka-} &           \rc{kɨ-wɨnɨkɨ-} &  \obj{kɨ-nɨkɨ-} &   \obj{kɨ-wɨnɨkɨ-} \\
\gl{2}   &           \rc{m-eɸurka-} &  \obj{m-ehurka-} &    \obj{m-eɸɨrka-} &            \rc{o-wɨnɨkɨ-} &  \obj{o-wnɨkɨ-} &   \obj{mɨ-wɨnɨkɨ-} \\
\gl{1+2} &           \rc{t-eɸurka-} &  \obj{t-ehurka-} &  \obj{t͡ʃ-eɸɨrka-} &          \rc{tɨt-wɨnɨkɨ-} &  \obj{tɨ-nɨkɨ-} &  \obj{tɨt-wɨnɨkɨ-} \\
\gl{3}   &           \rc{ɲ-eɸurka-} &  \obj{ɲ-ehurka-} &    \obj{ɲ-eɸɨrka-} &           \rc{nɨ-wɨnɨkɨ-} &  \obj{nɨ-nɨkɨ-} &   \obj{nɨ-wɨnɨkɨ-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-26 documents/floats/aku1sa.tex
regular verbs in data and docs
added more regular verbs

@@
 \centering
-\caption[\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes]{\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes \parencites[77, 79, 82, 84, 85, 86, 87]{gildea1994akuriyo}}
+\caption[Regular \akuriyo \gl{1}\gl{s_a_} markers]{Regular \akuriyo \gl{1}\gl{s_a_} markers \parencites[77, 79, 82, 84, 85, 86, 87]{gildea1994akuriyo}}
 \label{tab:aku1sa}
 \begin{tabular}[t]{@{}ll@{}}
-\toprule
+\mytoprule
       first person \obj{k-} &               first person \obj{t͡ʃ-} \\
@@
   \obj{əturu-} \qu{to talk} & \obj{ehpa-} \qu{to bathe (\gl{intr})} \\
-   \obj{əʔnɨkɨ-} \qu{sleep} &                                       \\
+\obj{əənɨkɨ-} \qu{to sleep} &                                       \\
 \bottomrule

### 2021-10-26 documents/floats/pekreg.tex
added more regular verbs

@@
 \centering
-\caption[Regular Pekodian \gl{s_a_} verbs]{Regular Pekodian \gl{s_a_} verbs \textcites[4]{meira2003bakairi}[150]{alves2017arara}[52]{ikpengpacheco2001}}
+\caption[Regular Pekodian \gl{s_a_} verbs]{Regular Pekodian \gl{s_a_} verbs \parencites[4]{meira2003bakairi}[150]{alves2017arara}[52]{ikpengpacheco2001}}
 \label{tab:pekreg}
@@
 \mytoprule
-{} &        \bakairi &                       \arara &                        \ikpeng \\
-{} &   \qu{to go up} &                \qu{to dance} &                    \qu{to run} \\
+{} & \bakairi \qu{to go up} &         \arara \qu{to dance} &            \ikpeng \qu{to run} \\
 \midrule

### 2021-10-26 documents/floats/pwaireg.tex
added more regular verbs

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \PWai verbs \parencites[30]{waiwaihawkins1998}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}
@@
 \mytoprule
-{} & \multicolumn{3}{l}{\qu{to fall} (\gl{s_a_})} & \multicolumn{3}{l}{\qu{to sleep} (\gl{s_p_})} \\
-{} &                    \PWai &           \hixka &            \waiwai &                     \PWai &          \hixka &            \waiwai \\
+{} & \multicolumn{2}{l}{\PWai} & \multicolumn{2}{l}{\hixka} & \multicolumn{2}{l}{\waiwai} \\
+{} &     \qu{to fall} &      \qu{to sleep} &     \qu{to fall} &   \qu{to sleep} &       \qu{to fall} &      \qu{to sleep} \\
 \midrule
-\gl{1}   &           \rc{k-eɸurka-} &  \obj{k-ehurka-} &    \obj{k-eɸɨrka-} &           \rc{kɨ-wɨnɨkɨ-} &  \obj{kɨ-nɨkɨ-} &   \obj{kɨ-wɨnɨkɨ-} \\
-\gl{2}   &           \rc{m-eɸurka-} &  \obj{m-ehurka-} &    \obj{m-eɸɨrka-} &            \rc{o-wɨnɨkɨ-} &  \obj{o-wnɨkɨ-} &   \obj{mɨ-wɨnɨkɨ-} \\
-\gl{1+2} &           \rc{t-eɸurka-} &  \obj{t-ehurka-} &  \obj{t͡ʃ-eɸɨrka-} &          \rc{tɨt-wɨnɨkɨ-} &  \obj{tɨ-nɨkɨ-} &  \obj{tɨt-wɨnɨkɨ-} \\
-\gl{3}   &           \rc{ɲ-eɸurka-} &  \obj{ɲ-ehurka-} &    \obj{ɲ-eɸɨrka-} &           \rc{nɨ-wɨnɨkɨ-} &  \obj{nɨ-nɨkɨ-} &   \obj{nɨ-wɨnɨkɨ-} \\
+\gl{1}   &  \obj{k-eɸurka-} &   \obj{kɨ-wɨnɨkɨ-} &  \obj{k-ehurka-} &  \obj{kɨ-nɨkɨ-} &    \obj{k-eɸɨrka-} &   \obj{kɨ-wɨnɨkɨ-} \\
+\gl{2}   &  \obj{m-eɸurka-} &    \obj{o-wɨnɨkɨ-} &  \obj{m-ehurka-} &  \obj{o-wnɨkɨ-} &    \obj{m-eɸɨrka-} &   \obj{mɨ-wɨnɨkɨ-} \\
+\gl{1+2} &  \obj{t-eɸurka-} &  \obj{tɨt-wɨnɨkɨ-} &  \obj{t-ehurka-} &  \obj{tɨ-nɨkɨ-} &  \obj{t͡ʃ-eɸɨrka-} &  \obj{tɨt-wɨnɨkɨ-} \\
+\gl{3}   &  \obj{ɲ-eɸurka-} &   \obj{nɨ-wɨnɨkɨ-} &  \obj{ɲ-ehurka-} &  \obj{nɨ-nɨkɨ-} &    \obj{ɲ-eɸɨrka-} &   \obj{nɨ-wɨnɨkɨ-} \\
 \bottomrule

### 2021-10-26 documents/floats/yukreg.tex
added more regular verbs

\begin{table}
\centering
\caption[Regular \yukpa verbs]{Regular \yukpa verbs \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}}
\label{tab:yukreg}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &  \qu{to fall} & \qu{to sleep} & \qu{to wash self} \\
\midrule
\gl{1} &  \obj{j-ata-} &  \obj{jɨ-nɨ-} &     \obj{j-otum-} \\
\gl{2} &  \obj{m-ata-} &  \obj{mɨ-nɨ-} &     \obj{m-otum-} \\
\gl{3} &  \obj{n-ata-} &  \obj{nɨ-nɨ-} &     \obj{n-otum-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-10-26 documents/sections/extensions.tex
regular verbs in data and docs
added more regular verbs

@@
 \textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work has been done on \PPek morphosyntax.
-All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \exref{pekreg}.
+All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.
 Its presence can thus be reconstructed to \PPek.
 
-\ex<pekreg>
-\begin{tabular}[t]{@{}llll@{}}
-& \bakairi \qu{to go up}  & \arara \qu{to dance}  & \ikpeng \qu{to run} \\
-& \parencite[4]{meira2003bakairi} & \parencite[150]{alves2017arara} &  \parencite[52]{ikpengpacheco2001}\\
-\gl{1}\gl{s_a_} & \obj{\emp{k-}əku-} & \obj{\emp{k-}origu-} & \obj{\emp{k-}aranme-} \\
-\gl{2}\gl{s_a_} & \obj{m-əku-} & \obj{m-origu-} & \obj{m-aranme-} \\
-\gl{1+2}\gl{s_a_} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
-\gl{3}\gl{s_a_} & \obj{n-əku-} & ∅\obj{-origu-} & ∅\obj{-aranme-} \\
-\end{tabular}
-\xe
+%\ex<pekreg>
+%\begin{tabular}[t]{@{}llll@{}}
+%& \bakairi \qu{to go up}  & \arara \qu{to dance}  & \ikpeng \qu{to run} \\
+%& \parencite[4]{meira2003bakairi} & \parencite[150]{alves2017arara} &  \parencite[52]{ikpengpacheco2001}\\
+%\gl{1}\gl{s_a_} & \obj{\emp{k-}əku-} & \obj{\emp{k-}origu-} & \obj{\emp{k-}aranme-} \\
+%\gl{2}\gl{s_a_} & \obj{m-əku-} & \obj{m-origu-} & \obj{m-aranme-} \\
+%\gl{1+2}\gl{s_a_} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
+%\gl{3}\gl{s_a_} & \obj{n-əku-} & ∅\obj{-origu-} & ∅\obj{-aranme-} \\
+%\end{tabular}
+%\xe
+
+\input{floats/pekreg}
 
@@
  \item \obj{akari} \qu{dog} in \cref{tab:pxinw} above
- \item \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg
+ \item \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg}
  \item \obj{jaj} \qu{tree} \parencite[98]{ikpengpacheco2001} from \PC \rc{jəje}
@@
 This extension was briefly discussed in \cref{sec:extensions_intro}; the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was later extended to \gl{1}\gl{s_a_} in \PWai.
-This created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \exref{wais}.
+This created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \pcref{tab:pwaireg}.
 
-\ex<wais> Waiwaian \rc{eɸurka} \qu{to fall} (\gl{s_a_}) and \rc{wɨnɨkɨ} \qu{to sleep} (\gl{s_p_})\\
-\begin{tabular}[t]{@{}lllllll@{}}
-& \multicolumn{2}{l}{\PWai} & \multicolumn{2}{l}{\hixka}  & \multicolumn{2}{l}{\waiwai} \\
- & \gl{s_a_} & \gl{s_p_} & \gl{s_a_} & \gl{s_p_} & \multicolumn{2}{l}{\gl{s}} \\
-\gl{1} & \rc{\emp{k-}eɸurka-} & \rc{kɨ-wɨnɨkɨ-} & \obj{\emp{k-}ehurka-} & \obj{kɨ-nɨkɨ-} & \obj{\emp{k-}eɸɨrka-} & \obj{kɨ-wɨnk-} \\
-\gl{2} & \rc{m-eɸurka-} & \rc{o-wɨnɨkɨ-} & \obj{m-ehurka-} & \obj{o-wnɨkɨ-} & \obj{m-eɸɨrka-} & \obj{mɨ-wɨnk-} \\
-\gl{1+2} & \rc{t-eɸurka-} & \rc{tɨt-wɨnɨkɨ-} & \obj{t-ehurka-} & \obj{tɨ-nɨkɨ-} & \obj{t͡ʃ-eɸɨrka-} & \obj{tɨt-wɨnk-} \\
-\gl{3} & \rc{ɲ-eɸurka-} & \rc{nɨ-wɨnɨkɨ-} & \obj{ɲ-ehurka-} & \obj{nɨ-nɨkɨ-} & \obj{ɲ-eɸɨrka-} & \obj{nɨ-wɨnk-} \\
-\end{tabular}\\
-\parencites[510]{howard2001wrought}[189--191]{hixkaryanaderby1985}[209--211]{hawkins1953waiwai}[50]{waiwaihawkins1998}
-\xe
+\input{floats/pwaireg}
+
+%\ex<wais> Waiwaian \rc{eɸurka} \qu{to fall} (\gl{s_a_}) and \rc{wɨnɨkɨ} \qu{to sleep} (\gl{s_p_})\\
+%\begin{tabular}[t]{@{}lllllll@{}}
+%& \multicolumn{2}{l}{\PWai} & \multicolumn{2}{l}{\hixka}  & \multicolumn{2}{l}{\waiwai} \\
+% & \gl{s_a_} & \gl{s_p_} & \gl{s_a_} & \gl{s_p_} & \multicolumn{2}{l}{\gl{s}} \\
+%\gl{1} & \rc{\emp{k-}eɸurka-} & \rc{kɨ-wɨnɨkɨ-} & \obj{\emp{k-}ehurka-} & \obj{kɨ-nɨkɨ-} & \obj{\emp{k-}eɸɨrka-} & \obj{kɨ-wɨnk-} \\
+%\gl{2} & \rc{m-eɸurka-} & \rc{o-wɨnɨkɨ-} & \obj{m-ehurka-} & \obj{o-wnɨkɨ-} & \obj{m-eɸɨrka-} & \obj{mɨ-wɨnk-} \\
+%\gl{1+2} & \rc{t-eɸurka-} & \rc{tɨt-wɨnɨkɨ-} & \obj{t-ehurka-} & \obj{tɨ-nɨkɨ-} & \obj{t͡ʃ-eɸɨrka-} & \obj{tɨt-wɨnk-} \\
+%\gl{3} & \rc{ɲ-eɸurka-} & \rc{nɨ-wɨnɨkɨ-} & \obj{ɲ-ehurka-} & \obj{nɨ-nɨkɨ-} & \obj{ɲ-eɸɨrka-} & \obj{nɨ-wɨnk-} \\
+%\end{tabular}\\
+%\parencites[510]{howard2001wrought}[189--191]{hixkaryanaderby1985}[209--211]{hawkins1953waiwai}[50]{waiwaihawkins1998}
+%\xe
 %
@@
 The \setone paradigms of \trio and \akuriyo contain an interesting puzzle: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
+
+\input{floats/ptirreg}
 %This is illustrated by comparing reconstructible \PTir forms with more conservative \wayana \exref{sat} and \kalina \exref{13w}, which can be taken as representative for \PC.
@@
 As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since almost all \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} is attested.}
-This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \exref{carpar}.
-
-\ex<carpar> \carijo \parencites[106]{meira1998proto}[173]{robayo2000avance}\\
-\begin{tabular}[t]{@{}lll@{}}
-& \obj{tuda} \qu{to arrive} & \obj{eharaga} \qu{to dance}\\
-\gl{1} & \obj{ji-tuda-} & \obj{j-eharaga-}\\
-\gl{2} & \obj{mɨ-tuda-} & \obj{m-eharaga-}\\
-\gl{1+2} & \obj{kɨsi-tuda-} & \obj{kɨs-eharaga-}\\
-\gl{3} & \obj{ni-tuda-} & \obj{n-eharaga-}\\
-\end{tabular}
-\xe
+This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \pcref{tab:carreg}.
+%
+\input{floats/carreg}
 %
@@
 \yukpa has lost most of its \setone constructions, mostly employing innovative ones \parencite{meira2006syntactic}.
-It preserves the \setone prefixes in the immediate past, where the \gl{1}\gl{s_p_} marker appears to have been extended to \gl{s_a_} verbs, as evidenced by the paradigms in \exref{yukpaintr}.
+It preserves the \setone prefixes in the immediate past, where the \gl{1}\gl{s_p_} marker appears to have been extended to \gl{s_a_} verbs, as evidenced by the paradigms in \cref{tab:yukreg}.
 The loss of split-\gl{s} in favor of the \gl{1}\gl{s_p_} marker is accompanied by the \gl{2}\gl{s_a_} marker \obj{m(ɨ)-} displacing \gl{2}\gl{s_p_} \rc{ə(j)-}, and \gl{1+2} being lost as an inflectional value.
@@
 
-\ex<yukpaintr> \yukpa \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}\\
-\begin{tabular}[t]{@{}llll@{}}
-& \obj{otum} \qu{to wash self} & \obj{nɨ} \qu{to sleep} & \obj{ata} \qu{to fall}\\
-\gl{1} & \obj{j-otum-} & \obj{jɨ-nɨ-} & \obj{j-ata-}\\
-\gl{2} & \obj{m-otum-} & \obj{mɨ-nɨ-} & \obj{m-ata-}\\
-\gl{3} & \obj{n-otum-} & \obj{nɨ-nɨ-} & \obj{n-ata-}\\
-\end{tabular}
-\xe
-%
-The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of \detrz, like \obj{otum} \qu{to wash self} in \exref{yukpaintr}.
+%\ex<yukpaintr> \yukpa \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}\\
+%\begin{tabular}[t]{@{}llll@{}}
+%& \obj{otum} \qu{to wash self} & \obj{nɨ} \qu{to sleep} & \obj{ata} \qu{to fall}\\
+%\gl{1} & \obj{j-otum-} & \obj{jɨ-nɨ-} & \obj{j-ata-}\\
+%\gl{2} & \obj{m-otum-} & \obj{mɨ-nɨ-} & \obj{m-ata-}\\
+%\gl{3} & \obj{n-otum-} & \obj{nɨ-nɨ-} & \obj{n-ata-}\\
+%\end{tabular}
+%\xe
+
+\input{floats/yukreg}
+
+The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of \detrz, like \obj{otum} \qu{to wash self} in \cref{tab:yukreg}.
 The analysis that this is a reflex of the \gl{1}\gl{s_p_} marker \rc{u(j)-} is supported by the same form occurring in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.

### 2021-10-26 workflow/process_data.py

@@ -286,14 +286,14 @@ out.fillna("", inplace=True)
 save_float(
     print_latex(out),
     "aku1sa",
-    r"\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes " + sources,
-    short=r"\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes",
+    r"Regular \akuriyo \gl{1}\gl{s_a_} markers " + sources,
+    short=r"Regular \akuriyo \gl{1}\gl{s_a_} markers",
 )
 
 # regular proto-waiwaian verbs
 meanings = ["fall", "sleep"]
 lgs = ["PWai", "hix", "wai"]
-pyd.x = ["Meaning", "Language_ID"]
+pyd.x = ["Language_ID", "Meaning"]
 pyd.y = ["Inflection"]
 pyd.filters = {"Language_ID": lgs, "Meaning": meanings}
 pyd.sort_orders = {
@@ -308,7 +308,7 @@ print(tabular)
 
 tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name[1])))
 tabular.rename(
-    columns={"sleep": r"\qu{to sleep} (\gl{s_p_})", "fall": "\qu{to fall} (\gl{s_a_})"},
+    columns={"sleep": r"\qu{to sleep}", "fall": "\qu{to fall}"},
     level="Meaning",
     inplace=True,
 )
@@ -319,7 +319,7 @@ tabular.columns.names = [None, None]
 save_float(
     print_latex(tabular, keep_index=True),
     "pwaireg",
-    "Regular \PWai verbs " + sources,
+    "Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai " + sources,
     short="Regular \PWai verbs",
 )
 
@@ -328,6 +328,7 @@ tir_reg = i_df[~(pd.isnull(i_df["Verb_Cognateset_ID"]))]
 meanings = ["bathe_intr", "sleep"]
 lgs = ["PTir", "tri", "aku"]
 pyd.filters = {"Language_ID": lgs, "Meaning": meanings}
+pyd.x = ["Meaning", "Language_ID"]
 pyd.sort_orders = {
     "Language_ID": lgs,
     "Meaning": meanings,
@@ -365,9 +366,8 @@ for lg, meaning in {"bak": "go_up", "ara": "dance", "ikp": "run"}.items():
[…]

### 2021-10-26 documents/floats/pwaireg.tex
removed 'sleep' cognates

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-10-26 documents/sections/proto_carib.tex
various revisions
removed 'sleep' cognates

@@
 	\centering
-	\caption{\PC \setone (main clause) person markers \parencites[495]{meira2010origin}[497]{gildea2016referential}}
+	\caption[\PC \setone (main clause) person markers]{\PC \setone (main clause) person markers \parencites[495]{meira2010origin}[497]{gildea2016referential}}
 	\label{tab:pcpers}
@@
 	\centering
-	\caption{Underived \PC \gl{s_a_} verbs \parencite[30]{gildea2007greenberg}}
+	\caption[Underived \PC \gl{s_a_} verbs]{Underived \PC \gl{s_a_} verbs \parencite[30]{gildea2007greenberg}}
 	\label{tab:underived2007}
@@
 \mymidrule
-\rc{tə(mə)} & \qu{to go}\\
+\rc{tə[mə]} & \qu{to go}\\
 \rc{ətepɨ} & \qu{to come\subs{1}}\\
-\rc{ka(ti)} & \qu{to say}\\
-\rc{əmə(mɨ)} & \qu{to enter}\\
+\rc{ka[ti]} & \qu{to say}\\
+\rc{əmə[mɨ]} & \qu{to enter}\\
 \rc{eti} & \qu{to dwell, be\subs{2}}\\
-\rc{a(p)} & \qu{to be\subs{1}, say}\\
+\rc{a[p]} & \qu{to be\subs{1}, say}\\
 \rc{əməkɨ} & \qu{to come\subs{2}}\\
@@
 \begin{table}[h]
-\caption{Some examples for completed extensions \parencite[90--92]{gildea1998}}
+\caption[Some examples for completed extensions]{Some examples for completed extensions \parencite[90--92]{gildea1998}}
 \label{tab:completed}
@@
 
-%For instance, while the extension of \PWai \gl{1}\gl{s_p_} \rc{k-} affected most verbs, like \rc{eɸurka} \qu{to fall}, verbs like \rc{ka(ti)} \qu{to say} were not affected \exref{pwaiext}.
+%For instance, while the extension of \PWai \gl{1}\gl{s_p_} \rc{k-} affected most verbs, like \rc{eɸurka} \qu{to fall}, verbs like \rc{ka[ti]} \qu{to say} were not affected \exref{pwaiext}.
 

### 2021-10-26 workflow/process_data.py
fixed object markdown, removed invisible character

@@ -306,7 +306,7 @@ sources = get_sources(i_df)
 
 print(tabular)
 
-tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name[1])))
+tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name[0])))
 tabular.rename(
     columns={"sleep": r"\qu{to sleep}", "fall": "\qu{to fall}"},
     level="Meaning",

### 2021-10-28 documents/sections/extensions.tex
went through 3.1-3.4

@@
 \label{sec:extensions}
-As stated in \cref{sec:extensions_intro}, the person marker extensions which did not spread through the entire lexicon all have in common that they feature innovative \gl{1}\gl{s_a_} markers.
+As stated in \cref{sec:extensions_intro}, the person marker extensions which did not spread through the entire lexicon all have in common that they feature innovative first person markers on verbs that are (at least historically) members of the \gl{s_a_} class.
 There are six attested such innovations, three of which can be reconstructed to intermediate proto-languages.
@@
 It consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
-\textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work has been done on \PPek morphosyntax.
-All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.
-Its presence can thus be reconstructed to \PPek.
+\textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work on \PPek morphosyntax can be found in the literature.
+All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.\footnote{In \cref{tab:pekreg} and the remainder of this paper, we omit any possible TAM suffixes, since our focus lies on the prefixes and stems, and since full paradigms containing the same TAM suffix are rarely found.}
+We can thus reconstruct a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
 
@@
 \begin{tabular}[t]{@{}ll@{}}
-\gl{1} & \obj{w-aptam}\\
+\gl{1} & \obj{w-aptam} \qu{when/if I was}\\
 \gl{2} & \obj{m-od-aptam}\\
@@
 In his brief discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.
-The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}, contrasting with regular \obj{əku} \qu{to go up} in \exref{pekreg} above.
-\qu{to bathe} is also found in the \obj{w-}list for \arara.
+The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}, contrasting with regular \obj{əku} \qu{to go up} in \cref{tab:pekreg} above.
 
@@
 %
-While \textcite[4]{meira2003bakairi} does list \obj{ge} \qu{to say}, \obj{tə} \qu{to go}, and \obj{əe(wɨ)} \qu{to come} as examples of \gl{s_a_} verbs, he does not indicate whether they belong to the class of \gl{s_a_}-1 verbs, with first person \obj{k-}, or the \gl{s_a_}-2 verbs, with \obj{w-}.\footnote{It should be noted at this point that \textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
+Of course, \qu{to bathe} is also found in the \obj{w-}list for \arara.
+While \textcite[4]{meira2003bakairi} does list \obj{ge} \qu{to say}, \obj{tə} \qu{to go}, and \obj{əe(wɨ)} \qu{to come} as examples of \gl{s_a_} verbs, he does not indicate whether they belong to the class of \gl{s_a_}-1 verbs, with first person \obj{k-}, or the \gl{s_a_}-2 verbs, with \obj{w-}.%
+\footnote{
+It should be noted at this point that \textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
 However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
 Given the lack of data for other verbs, we will not further discuss this potential additional pattern.
-If the characterization by \citeauthor{meira2003bakairi} is accurate, then the pattern is fully parallel to the distribution of the first person prefixes.}
+If the characterization by \citeauthor{meira2003bakairi} is accurate, then the pattern is fully parallel to the distribution of the first person prefixes.
+}
 Luckily, while \textcite{von1892bakairi} did not accurately record all phonemic distinctions in \bakairi \parencite{meira2005bakairi}, he does provide inflected forms of cognates to the \arara verbs in \exref{arairr}.
-We present them according to our understanding of the phonemic analysis of \bakairi by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi} in \exref{bakverbs}.
+We present them in \exref{bakverbs} according to our current understanding of \bakairi phonology and verbal morphology, based on \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
 
@@
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-Further, while there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} forms are attested with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
+Further, while there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} reflexes seem to occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
 
@@
 \Cref{tab:ppekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with our \PPek reconstruction.
-The presence of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests to us that it may be cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
+The presence and distribution of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
 Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \cref{tab:pxinw}.
 While it is by no means a regular sound change, it allows us to securely connect the two prefixes.
-Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
-Thus, we reconstruct a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} to \PPek, identical to \arara.
+Similarly, the supposed change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+Thus, we reconstruct a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} to \PPek, identical to the \arara one in form and distribution.
 
@@
 Further, the forms are not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
-However, the addition of a final \obj{ŋ} in \PXin is attested elsewhere: \begin{inlinelist}
+However, the addition of a final \obj{ŋ} in \PXin is attested elsewhere,%
+\footnote{
+\begin{inlinelist}
 \item \PC \rc{əne} \qu{to see}, \arara and \ikpeng \obj{eneŋ} %\parencites[8]{gildea2007greenberg}[56]{alves2017arara}[25]{ikpengpacheco2001}
@@
 \item \PC \rc{ənə} \qu{to eat meat}, \arara \obj{oŋoŋ} \qu{to bite} %\parencites[8]{gildea2007greenberg}[57]{alves2017arara}
-\end{inlinelist} \parencites[8]{gildea2007greenberg}[56, 144, 57]{alves2017arara}[25, 270]{ikpengpacheco2001}.
-Also, based on the fact \bakairi has generally lost much segmental material, we suggest that \bakairi \obj{gɨ} is another addition to a root \rc{ɨptə}, rather than a conservative form.
+\end{inlinelist} \parencites[8]{gildea2007greenberg}[56, 144, 57]{alves2017arara}[25, 270]{ikpengpacheco2001}.}
+and based on the fact \bakairi has generally lost much segmental material, we suggest that both \obj{ŋ} and \obj{gɨ} are later additions to a root \rc{ɨptə}.
 
@@
 In contrast, the \arara first person form is directly based on this root \rc{epɨ}.
-However, other person values show a reflex of \rc{əd-} in \arara, too \exref{ara-123}.
+However, reflexes of \rc{əd-epɨ} can be found elsewhere in the \arara paradigm \exref{ara-123}.
 
@@
 On the other hand, both \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-ebɨ}.
+Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}.
 
-The V-initial nature of \qu{to go} is evidenced in its Xinguan forms, while \bakairi \phon{\rc{wɨ}}{\obj{u}} resulted in a shift of the morpheme boundary, see also \cref{sec:go}.
-The \ikpeng form \obj{aran} is made compatible with our reconstruction \rc{ɨtən} by considering that \ikpeng \obj{a} is an attested outcome of \rc{ə} (\PXin \rc{o}/\obj{e}), for example \begin{inlinelist}
+The V-initial nature of \qu{to go} is evidenced in its Xinguan forms, while \bakairi \phon{\rc{wɨ}}{\obj{u}} resulted in a shift of the morpheme boundary (see also \cref{sec:go}).
+The \ikpeng form \obj{aran} is compatible with our reconstruction \rc{ɨtən}, when one considers that \ikpeng \obj{a} is an attested outcome of \rc{ə}.\footnote{\begin{inlinelist}
  \item \obj{akari} \qu{dog} in \cref{tab:pxinw} above
@@
  \item \obj{jaj} \qu{tree} \parencite[98]{ikpengpacheco2001} from \PC \rc{jəje}
- \end{inlinelist}.
+ \end{inlinelist}.}
 This attested change of \rc{ə} to \obj{a} need only be preceded by a assimilatory lowering of initial \rc{ɨ} to \rc{ə}, to yield the form \obj{aran} from \rc{ɨtən}.
@@
 Seven verbs can be reconstructed as having resisted this innovation, and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
-In later, individual developments, \bakairi introduced \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
-Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ} and introduced \obj{k-}.
+In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
+Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ}, accompanied by the extension of \obj{k-}.
 
@@
 This created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \pcref{tab:pwaireg}.
-
+%
 \input{floats/pwaireg}
-
+%
 %\ex<wais> Waiwaian \rc{eɸurka} \qu{to fall} (\gl{s_a_}) and \rc{wɨnɨkɨ} \qu{to sleep} (\gl{s_p_})\\
@@
 %
-\waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
-Its \hixka counterpart also takes a reflex of \gl{1}\gl{s_a_} \rc{w-}, which has the form \obj{ɨ-} \exref{hixɨ.hix-118}.
-A formally identical prefix is found on transitive verbs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, where it also corresponds to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.
+Not all verbs were affected by this extension; \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
+Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, a potential reflex of \gl{1}\gl{s_a_} \rc{w(ɨ)-}.
+A formally identical prefix occurs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, which regularly corresponds to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.
 
@@
 %
-This correspondence allows us to establish \hixka \obj{ɨ-} a reflex of \rc{wɨ-}, similar to \ikpeng \pcref{sec:pekodian}.
+This correspondence allows us to establish \hixka \obj{ɨ-} a reflex of \rc{w(ɨ)-}, similar to \ikpeng \pcref{sec:pekodian}.
 Notably, \textcite{hixkaryanaderby1985} does not see this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_}, but as the regular \gl{1}>\gl{3} marker, because he considers \hixka \obj{ka} \qu{to say} to be transitive, an issue which will be discussed in \cref{sec:say}.
@@
 There are three more verbs which did not take innovative \rc{k-} in \PWai, shown alongside \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
-While the two roots for \qu{to be} are straightforwardly reconstructible, \qu{to go} is somewhat of a special case.
-While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \gl{1}\gl{s_a_} \obj{k-} with the old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
+The two roots for \qu{to be} are straightforwardly reconstructible, whereas \qu{to go} is somewhat of a special case.
+While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
 Alternatively, this form was influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \exref{waiwto}.
@@
 The \setone paradigms of \trio and \akuriyo contain an interesting puzzle: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
+This resulted in a regular \gl{1}\gl{s_a_} marker \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}; the latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
 
@@
 %
-The question of how this happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
+The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
 In fact, even the issue of \emph{when} this happened is open.
@@
 %
-As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} lists the first four verbs in \cref{tab:ptirverbs} as being reconstructible to \PTar with \rc{w-}, for which we offer our reconstructed \PTir forms.
+As for intransitive verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar; we offer our reconstructed \PTir forms.
 To this list, we add the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}}, which has first person \obj{w-} at least in \trio.
@@
 %
-For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for the first person, and \rc{eepɨ} for the others, based on the idiosyncratic form in \trio and the vowel length in \akuriyo.
-\akuriyo (and \carijo) then levelled the original distribution of \rc{əepɨ} and \rc{eepɨ}, similar to what we have suggested for Pekodian \pcref{sec:pekodian}.
+For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for the first person, and \rc{eepɨ} for the others, based on the idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
+\akuriyo (and \carijo) then levelled this original distribution, similar to what we have suggested for Pekodian \pcref{sec:pekodian}.
 We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly cognate with \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form must have been \rc{əʔepɨ}.
@@
 
-Further, here are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
+Apart from these \PTir verbs with \rc{w(ɨ)-}, there are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
 They have \gl{1}\gl{s_a_} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
@@
 \label{sec:carijo}
-As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since almost all \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} is attested.}
+As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.}
 This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \pcref{tab:carreg}.

### 2021-10-28 documents/floats/go.tex
new autogenerated tables
fixed proto-pekodian 'to go'
regular verbs in data and docs
further revision of verbs section

@@
 \centering
-\caption[Reflexes of \rc{ɨtə(mə)} \qu{to go}]{Reflexes of \rc{ɨtə(mə)} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6; p.c., Spike Gildea]{franchetto2002kuikuro}}
+\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6; p.c., Spike Gildea]{franchetto2002kuikuro}}
 \label{tab:go}
 \begin{tabular}[t]{@{}lllllll@{}}
-\toprule
-Language &             Form & \multicolumn{5}{l}{Alignment} \\
+\mytoprule
+Language & \multicolumn{6}{l}{Form} \\
 \midrule
-\kaxui    &     \obj{to(mo)} &           &  t &  o &  m &  o \\
-\hixka    &      \obj{(ɨ)to} &         ɨ &  t &  o &    &    \\
-\waiwai   &   \obj{(e)to(m)} &         e &  t &  o &  m &    \\
+\kaxui    &     \obj{to[mo]} &    &  t &  o &  m &  o \\
+\hixka    &      \obj{[ɨ]to} &  ɨ &  t &  o &    &    \\
+\waiwai   &   \obj{[e]to[m]} &  e &  t &  o &  m &    \\
 \arara    &        \obj{ɨdo} &  ɨ &  d &  o &    &    \\
@@
 \ikpeng   &        \obj{ero} &  e &  r &  o &    &    \\
-\bakairi  &      \obj{(ɨ)tə} &         ɨ &  t &  ə &    &    \\
-\trio     &      \obj{tə(n)} &           &  t &  ə &  n &    \\
-\akuriyo  &  \obj{(ə)tə(mɨ)} &         ə &  t &  ə &  m &  ɨ \\
+\bakairi  &      \obj{[ɨ]tə} &  ɨ &  t &  ə &    &    \\
+\trio     &      \obj{tə[n]} &    &  t &  ə &  n &    \\
+\akuriyo  &  \obj{[ə]tə[mɨ]} &  ə &  t &  ə &  m &  ɨ \\
 \carijo   &       \obj{təmə} &    &  t &  ə &  m &  ə \\
-\wayana   &   \obj{(ɨ)tə(m)} &         ɨ &  t &  ə &  m &    \\
+\wayana   &   \obj{[ɨ]tə[m]} &  ɨ &  t &  ə &  m &    \\
 \kalina   &         \obj{to} &    &  t &  o &    &    \\
-\kalina   &       \obj{(ɨ)ʔ} &         ɨ &  ʔ &    &    &    \\
-\maqui    &    \obj{ɨtə(mə)} &         ɨ &  t &  ə &  m &  ə \\
+\kalina   &       \obj{[ɨ]ʔ} &  ɨ &  ʔ &    &    &    \\
+\maqui    &    \obj{ɨtə[mə]} &  ɨ &  t &  ə &  m &  ə \\
 \ingariko &        \obj{ətə} &  ə &  t &  ə &    &    \\
-\pemon    &      \obj{(e)tə} &         e &  t &  ə &    &    \\
-\macushi  &      \obj{(a)tɨ} &         a &  t &  ɨ &    &    \\
-\panare   &      \obj{tə(n)} &           &  t &  ə &  n &    \\
+\pemon    &      \obj{[e]tə} &  e &  t &  ə &    &    \\
+\macushi  &      \obj{[a]tɨ} &  a &  t &  ɨ &    &    \\
+\panare   &      \obj{tə[n]} &    &  t &  ə &  n &    \\
 \yawarana &         \obj{tə} &    &  t &  ə &    &    \\
 \mapoyo   &         \obj{tə} &    &  t &  ə &    &    \\
-\uxc      &      \obj{(e)te} &         e &  t &  e &    &    \\
+\uxc      &      \obj{[e]te} &  e &  t &  e &    &    \\
 \yukpa    &         \obj{to} &    &  t &  o &    &    \\

### 2021-10-28 documents/floats/pwaireg.tex
further revision of verbs section

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}
@@
 \midrule
-\gl{1}   &  \obj{k-eɸurka-} &   \obj{kɨ-wɨnɨkɨ-} &  \obj{k-ehurka-} &  \obj{kɨ-nɨkɨ-} &    \obj{k-eɸɨrka-} &   \obj{kɨ-wɨnɨkɨ-} \\
-\gl{2}   &  \obj{m-eɸurka-} &    \obj{o-wɨnɨkɨ-} &  \obj{m-ehurka-} &  \obj{o-wnɨkɨ-} &    \obj{m-eɸɨrka-} &   \obj{mɨ-wɨnɨkɨ-} \\
-\gl{1+2} &  \obj{t-eɸurka-} &  \obj{tɨt-wɨnɨkɨ-} &  \obj{t-ehurka-} &  \obj{tɨ-nɨkɨ-} &  \obj{t͡ʃ-eɸɨrka-} &  \obj{tɨt-wɨnɨkɨ-} \\
-\gl{3}   &  \obj{ɲ-eɸurka-} &   \obj{nɨ-wɨnɨkɨ-} &  \obj{ɲ-ehurka-} &  \obj{nɨ-nɨkɨ-} &    \obj{ɲ-eɸɨrka-} &   \obj{nɨ-wɨnɨkɨ-} \\
+\gl{1}   &  \rc{k-eɸurka-} &   \rc{kɨ-wɨnɨkɨ-} &  \obj{k-ehurka-} &  \obj{kɨ-nɨkɨ-} &    \obj{k-eɸɨrka-} &   \obj{kɨ-wɨnɨkɨ-} \\
+\gl{2}   &  \rc{m-eɸurka-} &    \rc{o-wɨnɨkɨ-} &  \obj{m-ehurka-} &  \obj{o-wnɨkɨ-} &    \obj{m-eɸɨrka-} &   \obj{mɨ-wɨnɨkɨ-} \\
+\gl{1+2} &  \rc{t-eɸurka-} &  \rc{tɨt-wɨnɨkɨ-} &  \obj{t-ehurka-} &  \obj{tɨ-nɨkɨ-} &  \obj{t͡ʃ-eɸɨrka-} &  \obj{tɨt-wɨnɨkɨ-} \\
+\gl{3}   &  \rc{ɲ-eɸurka-} &   \rc{nɨ-wɨnɨkɨ-} &  \obj{ɲ-ehurka-} &  \obj{nɨ-nɨkɨ-} &    \obj{ɲ-eɸɨrka-} &   \obj{nɨ-wɨnɨkɨ-} \\
 \bottomrule

### 2021-10-28 documents/floats/say.tex
new autogenerated tables
regular verbs in data and docs
further revision of verbs section

@@
 \centering
-\caption[Reflexes of \rc{ka(ti)} \qu{to say}]{Reflexes of \rc{ka(ti)} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102; p.c., Spike Gildea]{mattei1994diccionario}}
+\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102; p.c., Spike Gildea]{mattei1994diccionario}}
 \label{tab:say}
 \begin{tabular}[t]{@{}llllll@{}}
-\toprule
-Language &          Form & \multicolumn{4}{l}{Alignment} \\
+\mytoprule
+Language & \multicolumn{5}{l}{Form} \\
 \midrule
-\kaxui   &   \obj{ka(s)} &         k &  a &  s &    \\
-\hixka   &   \obj{ka(h)} &         k &  a &  h &    \\
-\waiwai  &   \obj{ka(s)} &         k &  a &  s &    \\
+\kaxui   &   \obj{ka[s]} &  k &  a &  s &    \\
+\hixka   &   \obj{ka[h]} &  k &  a &  h &    \\
+\waiwai  &   \obj{ka[s]} &  k &  a &  s &    \\
 \arara   &      \obj{ke} &  k &  e &    &    \\
@@
 \carijo  &      \obj{ka} &  k &  a &    &    \\
-\wayana  &   \obj{ka(i)} &         k &  a &    &  i \\
-\apalai  &  \obj{ka(ʃi)} &         k &  a &  ʃ &  i \\
+\wayana  &   \obj{ka[i]} &  k &  a &    &  i \\
+\apalai  &  \obj{ka[ʃi]} &  k &  a &  ʃ &  i \\
 \kalina  &      \obj{ka} &  k &  a &    &    \\
@@
 \macushi &      \obj{ka} &  k &  a &    &    \\
-\panare  &   \obj{ka(h)} &         k &  a &  h &    \\
+\panare  &   \obj{ka[h]} &  k &  a &  h &    \\
 \uxc     &      \obj{ki} &  k &  i &    &    \\

### 2021-10-28 documents/sections/extensions.tex
further revision of verbs section

@@
 
-The V-initial nature of \qu{to go} is evidenced in its Xinguan forms, while \bakairi \phon{\rc{wɨ}}{\obj{u}} resulted in a shift of the morpheme boundary (see also \cref{sec:go}).
+The V-initial nature of \qu{to go} is evidenced in its Xinguan forms, while \bakairi \phon{\rc{wɨ}}{\obj{u}} muddied the morpheme boundary; other forms are V-initial \pcref{sec:go}.
 The \ikpeng form \obj{aran} is compatible with our reconstruction \rc{ɨtən}, when one considers that \ikpeng \obj{a} is an attested outcome of \rc{ə}.\footnote{\begin{inlinelist}
@@
 %
+Further, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car-33}.
+%We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
+Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would either expect \obj{wɨ-} or \obj{ji-}.
+We analyse this ∅ as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
+Further, one may suggest the difference to C-initial \obj{tə} \qu{to go} to result from that verb's former V-initial nature \pcref{sec:go}.
+
+\ex<car-33>\carijo \perscomm{David Felipe Guerrero-Beltran}\\
+\begingl
+\glpreamble dëmëmara kae ëwï iya//
+\gla n-tə-mə=mara ∅-ka-e əwɨ i-ja//
+\glb \gl{3}-go-\gl{pst}=\gl{dub} \gl{1}-say-\gl{npst}.\gl{cert} \gl{1}\gl{pro} \gl{3}-\gl{obl}//
+\glft \qu{“Did s/he leave?”, I say to him.}//
+\endgl
+\xe
+
+%
 %Interestingly enough, the verb \qu{to be} only shows \obj{w-} in its \obj{a} root allomorph, but not in the \obj{et͡ʃi} root allomorph, where it has \obj{j-} \exref{car1y.mivida-12}.
@@
 It preserves the \setone prefixes in the immediate past, where the \gl{1}\gl{s_p_} marker appears to have been extended to \gl{s_a_} verbs, as evidenced by the paradigms in \cref{tab:yukreg}.
-The loss of split-\gl{s} in favor of the \gl{1}\gl{s_p_} marker is accompanied by the \gl{2}\gl{s_a_} marker \obj{m(ɨ)-} displacing \gl{2}\gl{s_p_} \rc{ə(j)-}, and \gl{1+2} being lost as an inflectional value.
+The loss of split-\gl{s} in favor of the \gl{1}\gl{s_p_} marker is accompanied by the \gl{2}\gl{s_a_} marker \obj{m(ɨ)-} extending to \gl{s_p_} verbs, and \gl{1+2} being lost as an inflectional value.
 %It has lost the split-\gl{s} system in general
@@
 
-The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of \detrz, like \obj{otum} \qu{to wash self} in \cref{tab:yukreg}.
+The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of the detransitivizer \detrz, like \obj{otum} \qu{to wash self} in \cref{tab:yukreg}.
 The analysis that this is a reflex of the \gl{1}\gl{s_p_} marker \rc{u(j)-} is supported by the same form occurring in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.
@@
 On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian,sec:akuriyo}, we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
-The same zero marking is attested for a single intransitive verb: \obj{to} \qu{to go} \exref{yuk-7}.
-Based on regular C-initial verbs taking \obj{jɨ-}, like \qu{to sleep} in \exref{yukpaintr}, it is clear that also C-initial \obj{to} \qu{to go} preserved the old \gl{1}\gl{s_a_} marker \rc{w(ɨ)-}, which was lost due to phonological erosion.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian}, we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
+
+In intransitive verbs, this first-person zero marking is attested only for \obj{to} \qu{to go} \exref{yuk-7}.
+It contrasts with regular C-initial verbs taking \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
+Thus, we conclude that the extension of \obj{j-} in \yukpa only stopped short of \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.
 

### 2021-10-28 documents/sections/verbs.tex
floats for 'go' and 'say'
further revision of verbs section

@@
 
-\ex<yukcop> cliticized copula forms in \yukpa \parencite[143--144]{meira2006syntactic}\\
+\ex<yukcop> \yukpa \parencite[143--144]{meira2006syntactic}\\
 \begin{tabular}[t]{@{}lll@{}}
@@
 \label{sec:say}
-Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, illustrated with modern imperative forms in \exref{kati}.
+Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, best visible in the imperative forms of some languages \exref{kati}.
 
-\pex<kati> \PC \rc{kati-kə} \qu{say!}
-\a \apalai \obj{kaʃi-ko} \parencite[35]{koehn1986apalai}
-\a \wayana \obj{kai-kə} \parencite[181]{wayanatavares2005}
-\a \hixka \obj{kas-ko} \parencite[128]{hixkaryanaderby1985}
-\a \panare \obj{kah-kə} \parencite[102]{mattei1994diccionario}
+\ex<kati> \apalai\\
+\obj{kaʃi-ko} \qu{say!} \parencite[35]{koehn1986apalai}
 \xe
-
-This verb was not affected by the extensions found in \PPek, \PWai, and \PTir.
-It was not a potential target of \akuriyo \obj{k-}; we do not know the first person forms of its \carijo and \yukpa reflexes.
+%
+A comparison of the longest forms attested in each language is shown in \cref{tab:say}.
+Segments of cognate forms in this and other tables were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.
+%
+\input{floats/say.tex}
+%
+This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, and \carijo (\crefrange{sec:pekodian}{sec:carijo}).
+We do not know the first person forms of its \yukpa reflex \obj{ka}.
 
 As briefly mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes this verb as transitive in \hixka.
-This analytical choice is not only motivated by avoiding an idiosyncratic intransitive first person prefix \obj{ɨ-}, with the usual prefix being \obj{k-}.
+This analytical choice is not only motivated by avoiding an idiosyncratic intransitive first person prefix \obj{ɨ-}:
 \hixka \obj{ka} also behaves like a transitive verb in other ways, for instance by showing the complementary distribution of the third person marker \obj{n-} and preceding objects -- in this case direct speech or ideophones \exref{hixsay}.
@@
 The exceptionality of \obj{ka} \qu{to say} taking \obj{-po} \qu{\gl{caus}.\gl{tr}} has also been noted for \kalina \parencite[82]{courtz2008carib} and \wayana \parencite[258]{wayanatavares2005}.
+The agent nominalizer \rc{-ne} gave rise to the \panare gnomic verbal suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
+The occurrence of \obj{-ɲe} on \obj{ka} has led to differing categorizations as transitive \parencite[214]{panarepayne2013} or intransitive \parencite[102]{mattei1994diccionario}.
 Reflexes of another transitive causativizer \rc{-metipo} \parencite{gildea2015valency} are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
-\panare has innovated a gnomic or nonspecific verbal suffix \obj{-ɲe} from the agent nominalizer \rc{-ne} \parencite[184]{gildea1998}.
-Its occurrence on \obj{ka} leads \parencite[214]{panarepayne2013} to analyze the verb as transitive, in contrast to \textcite[102]{mattei1994diccionario}, who categorizes it as intransitive.
 
@@
 \hixka has lost the main intransitive criteria, making its reflex look more like a transitive verb.
+It is not attested as being affected by any of the person marker extensions under discussion.
 
@@
 It is true that many reflexes are clearly \obj{t}-initial, for example \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}.
-However, a comparison of selected forms with unambiguously C-initial \rc{ka[ti]} \qu{to say} suggests that an initial vowel \rc{ɨ} should be reconstructed \exref{gosaycomp},\footnote{Many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.} which was subsequently reanalyzed as part of the prefixes in many languages.
+However, once one considers all forms of the various reflexes of this verb \pcref{tab:go}, an initial vowel \rc{ɨ} must clearly be reconstructed -- in contrast to unambiguously C-initial \rc{ka[ti]} \qu{to say}.\footnote{One may add that many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.} which was subsequently reanalyzed as part of the prefixes in many languages.
+This verb was not affected by the any of the extensions discussed in \cref{sec:extensions}.
 
-\ex<gosaycomp>
-\begin{tabular}[t]{@{}lllll@{}}
-& go-\gl{imp} & go-\gl{neg} & say-\gl{imp} & say-\gl{neg} \\
-\wayana & \obj{ɨtə-kə} & \obj{ɨtə-ra} & \obj{kai-kə} & \obj{ka-ra} \\
-\hixka & \obj{ɨto-ko} & \obj{ɨto-hra} & \obj{kas-ko} & \obj{ka-hra} \\
-\apalai & \obj{ɨto-ko} & \obj{ɨto-pɨra} & \obj{kaʃi-ko} & \obj{ka-ra} \\
-\end{tabular}\\
-\parencites[66, 98]{camargo2010wayana}[235, 258]{wayanatavares2005}[47, 54 194]{hixkaryanaderby1985}[65]{derbyshire1965textos}[kuruaz 033, 055]{koehns1994textos}[100]{camargo2002lexico}[107]{koehn1986apalai}
-\xe
+\input{floats/go.tex}
 %
-
-This verb was not affected by the any of the extensions discussed in \cref{sec:extensions}.
+%\ex<gosaycomp>
+%\begin{tabular}[t]{@{}lllll@{}}
+%& go-\gl{imp} & go-\gl{neg} & say-\gl{imp} & say-\gl{neg} \\
+%\wayana & \obj{ɨtə-kə} & \obj{ɨtə-ra} & \obj{kai-kə} & \obj{ka-ra} \\
+%\hixka & \obj{ɨto-ko} & \obj{ɨto-hra} & \obj{kas-ko} & \obj{ka-hra} \\
+%\apalai & \obj{ɨto-ko} & \obj{ɨto-pɨra} & \obj{kaʃi-ko} & \obj{ka-ra} \\
+%\end{tabular}\\
+%\parencites[66, 98]{camargo2010wayana}[235, 258]{wayanatavares2005}[47, 54 194]{hixkaryanaderby1985}[65]{derbyshire1965textos}[kuruaz 033, 055]{koehns1994textos}[100]{camargo2002lexico}[107]{koehn1986apalai}
+%\xe
+%%
 
@@
 This verb is reconstructed as monomorphemic \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
-Different languages present rather different forms, which we argue are all (partially) cognate, with the reconstructed \PC verb having the form \rc{(ət-)jəpɨ}, with an optional detransitivizer.
-Reflexes of a bare \rc{j-}initial root \rc{jəpɨ} are found in the Pemongan languages and \kaxui.
-However, the one form in \kaxui only occurs in the Progressive \exref{kaxprog.kax-82}, and \obj{j(-)} may be a reflex the \settwo third person marker \rc{i-} combined with  the \gl{s_a_} class marker \rc{w-}, as suggested by the occurrence of the latter with other person values \exref{kaxprog.kax-54}.
-On the other hand, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
+Crucially, the \rc{ət} part is not reflected in the majority of forms, and many languages have reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ} for the \rc{epɨ} part.
+We argue that these forms go back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizer and a root \rc{jəpɨ}.
+Evidence for the originally \rc{j}-initial nature of the root is found in the Pemongan languages and \kaxui; the presence of \rc{əpɨ} and \rc{epɨ} reflexes in \kaxui raises the question of whether this \obj{j} is part of the root.
+The \kaxui form \obj{johɨ} only occurs in the Progressive \exref{kaxprog.kax-82}, and indeed \obj{j(-)} may be a reflex the \settwo third person marker \rc{i-} combined with the \gl{s_a_} class marker \rc{w-}, overtly present with other person values \exref{kaxprog.kax-54}.
+However, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
 Thus, it seems more likely that the \obj{j} is indeed part of the root, rather than an outcome of \rc{i-w-}.
@@
 \xe
+%
 
-Other languages -- across the family -- have forms reflecting the same root without the initial consonant, i.e.\ \rc{əpɨ}: \kalina, \panare, \yawarana, and \uxc.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
-Also found across the family are forms reflecting \rc{epɨ}, in \kaxui, \arara, \trio, \carijo, \maqui, and \mapoyo.
-The \akuriyo form \obj{eepɨ} looks like a reflex thereof as well, but the length is unexpected, and is analyzed by \textcite[]{meira1998proto} as reflecting an earlier diphthong \rc{əe}.
-A unifying account of these forms starts from a root \rc{jəpɨ}, which is the subject to two major sound changes: \begin{inlinelist}
+Most forms on which \posscite{gildea2007greenberg} \rc{ətepɨ} seems to be based do not show evidence for \rc{j}.
+However, we interpret the \obj{i} in the \akawaio form \obj{əsipɨ} as additional evidence of the sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of bare \rc{jəpɨ}, \obj{ipɨ}.
+
+There are many forms going seemingly reflecting \rc{əpɨ} and \rc{epɨ}, distributed widely through the family.
+A unifying account of these forms sees the root \rc{jəpɨ} undergoing two major sound changes: \begin{inlinelist}
  \item \rc{j}-loss
@@
 On the other hand, forms like \maqui \obj{ehə} must be the result of \phonl{\rc{ə}}{\rc{e}}{\rc{j}}, with subsequent loss of \rc{j}.
+The \akuriyo form \obj{eepɨ} looks like a reflex thereof as well, but the length is unexpected, and is analyzed by \textcite[]{meira1998proto} as reflecting an earlier diphthong \rc{əe}.
 
-We have so far ignored the forms which form the basis for \posscite{gildea2007greenberg} \rc{ətepɨ}; these are compatible with our reconstructed root \rc{jəpɨ}, with the addition of the detransitivizer \rc{ət(e)-}: \rc{ətjəpɨ}.
-While most of the modern forms with a reflex of the detransitivizer are lacking evidence for \rc{j} (\rc{ətepɨ}), the \obj{i} in the \akawaio form \obj{əsipɨ} is evidence of the sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of bare \rc{jəpɨ}, \obj{ipɨ}.
+%Other languages -- across the family -- have forms reflecting the same root without the initial consonant, i.e.\ \rc{əpɨ}: \kalina, \panare, \yawarana, and \uxc.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
+%Also found across the family are forms reflecting \rc{epɨ}, in \kaxui, \arara, \trio, \carijo, \maqui, and \mapoyo.
+
+Our interpretation of the \rc{ət} portion as a detransitivizer is primarily based on its form, as well as its paradigmatically conditioned occurrence in some languages.
 Although the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
@@
 
-While these sound changes and the addition of \rc{ət(e)-} do account for the majority of the forms in \cref{tab:come},\footnote{
+While these sound changes and the addition of \rc{ət-} do account for the majority of the forms in \cref{tab:come},\footnote{
 Apart from aforementioned \akuriyo \obj{eepɨ}, another exception is the \apalai form \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
@@
 Not only do very closely related languages show different forms, like \yawarana and \mapoyo, but distinct forms can even be found within the same language, usually conditioned by different prefixes.
-This was discussed in \cref{sec:pekodian} for \arara, which has reflexes of \rc{ətjəpɨ} and \rc{jəpɨ} within the same paradigm.
+This was discussed in \cref{sec:pekodian} for \arara, which has reflexes of \rc{ət-jəpɨ} and \rc{jəpɨ} within the same paradigm.
 A similar situation is found in \trio, where the \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}

### 2021-10-28 documents/sections/extensions.tex
incorporated uncertainty about 'to go down'

@@
 
-Apart from these \PTir verbs with \rc{w(ɨ)-}, there are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
-They have \gl{1}\gl{s_a_} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
-At least for \qu{to go down}, the \akuriyo cognate suggests that the irregular first person form \rc{p-ɨhtə-} can be reconstructed to \PTir.
-Whatever their origins, they were not affected by the extension of \rc{t-}.
+%Apart from these \PTir verbs with \rc{w(ɨ)-}, there are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
+%They have \gl{1}\gl{s_a_} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
+%At least for \qu{to go down}, the \akuriyo cognate suggests that the irregular first person form \rc{p-ɨhtə-} can be reconstructed to \PTir.
+%Whatever their origins, they were not affected by the extension of \rc{t-}.
+%
+%\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
+%\begin{tabular}[t]{@{}lll@{}}
+%& \trio & \akuriyo\\
+%\qu{go down} & \obj{p-ɨhtə-} & \obj{p-ɨtə-}\\
+%\qu{defecate} & \obj{k-oeka-} & ?\\
+%\end{tabular}
+%\xe
 
-\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
-\begin{tabular}[t]{@{}lll@{}}
-& \trio & \akuriyo\\
-\qu{go down} & \obj{p-ɨhtə-} & \obj{p-ɨtə-}\\
-\qu{defecate} & \obj{k-oeka-} & ?\\
-\end{tabular}
-\xe
-
-Finally, \textcite{gildea1994akuriyo} also recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-} \exref{akumov.aku}, all \obj{e}-initial movement verbs.
+In addition, \textcite{gildea1994akuriyo} recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-} \exref{akumov.aku}, all \obj{e}-initial movement verbs.
 We have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
@@
 
+In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in \akuriyo, also reconstructible to \PTir \exref{tridown}.
+It was not affected by the extension of \akuriyo \obj{k-}, but its relative chronology to the extension of \PTir \rc{t-} is unclear (see also \cref{sec:godown}).
+
+\ex<tridown> Tiriyoan \qu{to go down} \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
+\begin{tabular}[t]{@{}ll@{}}
+\trio & \obj{p-ɨhtə-} \\
+\akuriyo & \obj{p-ɨtə-}\\
+\end{tabular}
+\xe
+
 %Let us first consider a scenario where \obj{k-} was the older \gl{1}\gl{s_a_} marker, occurring on both \obj{e}- and \obj{ə}-initial verbs.
@@
 
-It has been speculated that the \trio \gl{s_a_} verb with a first person \obj{k-} marker, \obj{weka}/\obj{oeka} \qu{to defecate} might have something to do with \akuriyo \obj{k-}, which would potentially make the innovation of \rc{k-} a \PTir matter \parencite[116]{meira1998proto}.
-However, this hypothesis faces a problem in that the first person form for \akuriyo \qu{to defecate} is \obj{j-ereina-} \parencite[88]{gildea1994akuriyo}.
-Another hypothesis is that it originated in the corresponding \kaxui form \obj{ku-weka-} (form provided by Spike Gildea, p.c.), although the occurrence of \obj{o} in \trio would still need explanation.
+%It has been speculated that the \trio \gl{s_a_} verb with a first person \obj{k-} marker, \obj{weka}/\obj{oeka} \qu{to defecate} might have something to do with \akuriyo \obj{k-}, which would potentially make the innovation of \rc{k-} a \PTir matter \parencite[116]{meira1998proto}.
+%However, this hypothesis faces a problem in that the first person form for \akuriyo \qu{to defecate} is \obj{j-ereina-} \parencite[88]{gildea1994akuriyo}.
+%Another hypothesis is that it originated in the corresponding \kaxui form \obj{ku-weka-} (form provided by Spike Gildea, p.c.), although the occurrence of \obj{o} in \trio would still need explanation.
 %Rather, it seems likely that this irregular first person form originated in a class switch from \gl{s_p_} to \gl{s_a_} (\Cref{sec:resistant_verbs}), although it is unclear how exactly that switch resulted in the irregular form.

### 2021-10-28 documents/sections/verbs.tex
incorporated uncertainty about 'to go down'

@@
 \label{sec:godown}
-Reflexes of this verb were not affected by the extensions of \rc{t-} in \PTir \pcref{sec:taranoan} and \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
-Rather, an irregular first person form \rc{p-ɨptə} can be reconstructed for \PTir.
-While it originally resisted the extension of \rc{k-} in \PPek, \bakairi subsequently introduced it \pcref{sec:pekodian}.
-It was also affected by the extension of \obj{j-} in \carijo \exref{car-32}.
+Reflexes of this verb were not affected by the extensions of \rc{k-} in \PPek \pcref{sec:pekodian} and \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
+The resistance against the former extension was overcome in \bakairi.
+In \akuriyo, the verb already had an irregularly inflected first person form with \obj{p-}, inherited from \PTir.
 
-\ex<car-32>\carijo \perscommpar{David Felipe Guerrero}\\
+At first sight, it may seem that it was affected by the independent extensions of \obj{j-} in \carijo \exref{gowhere.car-18} and \yukpa \exref{gowhere.yuk-14}.
+
+\pex<gowhere>
+\a<car-18> \carijo \parencite[][102]{guerrero2019carijo}\\
 \begingl
-\gla irə wat͡ʃinakano tae j-ehɨtə-e//
-\glb \gl{inan}.\gl{ana} body.of.water along.bounded \gl{1}-go.down-\gl{npst}//
-\glft \qu{…I go down through that guachinacán.}//
+\gla əji-wa-e j-eh-ɨ//
+\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
+\glft \qu{I came looking for you.}//
+\endgl
+\a<yuk-14> \yukpa \parencite[][]{meira2003primeras}\\
+\begingl
+\glpreamble aw yéwtu//
+\gla aw j-ewuhtu//
+\glb \gl{1}-go.down//
+\glft \qu{I went down.}//
 \endgl
@@
 %
-While it seems that \yukpa \obj{ewu(h)tu} \qu{to go down} \parencite{meira2003primeras} is cognate, it was affected by the extension of \obj{j-}.
-%todo maybe insert a yukpa example here? the form is just jewuhtu
+However, a look at the comparative picture suggests a much more complicated situation.
+\cref{tab:godown} shows all attested cognates, some already discussed, including verb class membership and aligned segments.
+While a form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
+The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
 
-\input{floats/godown}
-
-A comparative view \pcref{tab:godown} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
-The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
 In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb in the synchronic analysis of \wayana.
@@
 
-\pex<waygodown>\wayana \parencite[][200]{wayanatavares2005}
-\a<way-73>
+\pex<gowhere>
+\a<car-18> \carijo \parencite[][102]{guerrero2019carijo}\\
 \begingl
-\glpreamble ïwïptëë//
-\gla ɨ-w-ɨptə-rɨ//
-\glb \gl{1}-\gl{s_a_}-go.down-\gl{nmlz}//
-\glft \qu{my going down}//
+\gla əji-wa-e j-eh-ɨ//
+\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
+\glft \qu{I came looking for you.}//
 \endgl
-\a<way-71>
+\a<yuk-14> \yukpa \parencite[][]{meira2003primeras}\\
 \begingl
-\gla əw-ɨptə-k//
-\glb \gl{2}-go.down-\gl{imp}//
-\glft \qu{Go down!}//
+\glpreamble aw yéwtu//
+\gla aw j-ewuhtu//
+\glb \gl{1}\gl{pro} \gl{1}-go.down//
+\glft \qu{I went down.}//
 \endgl
@@
 Its causativized form is \obj{ɨptəka} \parencite[255]{wayanatavares2005}, containing a reflex of the transitivizer \rc{-ka}, which was restricted to \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
-This leads us to posit the hypothesis that the verb may have been a normal member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class -- fully so in the languages where it is \gl{s_a_}.
+This leads us to posit the hypothesis that the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class.
+This in turn implies that all reflexes of this verb with \gl{s_a_} status were members of the \gl{s_p_} class at some point, but switched classes.
+
+\input{floats/godown}
 
 \wayana-external comparative evidence supports this hypothesis:
-The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:descend_cog} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
+The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:godown} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
 Both causativized forms contain a reflex of the (rare) transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
@@
 Thus, it seems that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.
+
+These circumstances make it impossible to answer the question of whether it was affected by the extensions in \PTir, \carijo, and \yukpa.
+For \PTir, we cannot establish a relative chronology of the verb class change, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
+It could have been an \gl{s_p_} verb at the time of the extension, it could have been an \gl{s_a_} verb and remained unaffected, or it could have been affected, but with a subsequent erasure of the results by the introduction of \rc{p-}.
+For \carijo and \yukpa, we cannot know whether the verb switched verb class before the breakdown of the entire split-\gl{s} system -- this is a plausible scenario especially in the former case.
+In all three cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, but it is also possible that it was not even a potential target due to its \gl{s_p_} status.
+On the other hand, the class change must have happened before the split-up of \trio and \akuriyo, and therefore this verb resisted the extension of \akuriyo \obj{k-}.
+Likewise, the class change must have taken place before the extension of \PPek \rc{k-}, since other scenarios would require an \gl{s_p_} verb taking on regular \gl{s_a_} markers, but selecting the archaic first person marker.
 
@@
 
-\subsection{\trio \obj{weka}/\obj{oeka} \qu{to defecate}}
-\label{sec:shit}
-The comparative picture for the resistant \trio \gl{s_a_} verb \qu{to defecate} \pcref{tab:defecate} shows that \rc{weka} was clearly originally an \gl{s_p_} verb.
-As in the case of \qu{to go down} \pcref{sec:godown}, this verb has a mixed \gl{s_a_}/\gl{s_p_} status in \wayana.
-The \bakairi reflex \obj{əeke} seems to contain a reflex of \rc{ət-}, which would explain its status as an \gl{s_a_} verb.
-While the class membership of \panare \obj{aiʔka} and \obj{iʔka} \qu{to defecate} is unknown, the former seems to contain a reflex of the detransitivizer, while the latter only consists of the root, meaning that one would assume \gl{s_a_} and \gl{s_p_} membership, respectively.
-
-\input{floats/defecate}
+%\subsection{\trio \obj{weka}/\obj{oeka} \qu{to defecate}}
+%\label{sec:shit}
+%The comparative picture for the resistant \trio \gl{s_a_} verb \qu{to defecate} \pcref{tab:defecate} shows that \rc{weka} was clearly originally an \gl{s_p_} verb.
+%As in the case of \qu{to go down} \pcref{sec:godown}, this verb has a mixed \gl{s_a_}/\gl{s_p_} status in \wayana.
+%The \bakairi reflex \obj{əeke} seems to contain a reflex of \rc{ət-}, which would explain its status as an \gl{s_a_} verb.
+%While the class membership of \panare \obj{aiʔka} and \obj{iʔka} \qu{to defecate} is unknown, the former seems to contain a reflex of the detransitivizer, while the latter only consists of the root, meaning that one would assume \gl{s_a_} and \gl{s_p_} membership, respectively.
+%
+%\input{floats/defecate}
 %

### 2021-10-28 workflow/process_data.py

@@ -319,7 +319,8 @@ tabular.columns.names = [None, None]
 save_float(
     print_latex(tabular, keep_index=True),
     "pwaireg",
-    "Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai " + sources,
+    "Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai "
+    + sources,
     short="Regular \PWai verbs",
 )
 
@@ -385,11 +386,11 @@ save_float(
     short="Regular Pekodian \gl{s_a_} verbs",
 )
 
-#regular carijo and yukpa verbs
+# regular carijo and yukpa verbs
 pyd.x = ["Meaning"]
 for lg, meanings in {
     "car": ["arrive", "dance"],
-    "yuk": ["wash_self", "sleep", "fall"]
+    "yuk": ["wash_self", "sleep", "fall"],
 }.items():
     pyd.filters = {"Language_ID": [lg], "Meaning": meanings}
     tabular = pyd.compose_paradigm(i_df, multi_index=True)
@@ -400,14 +401,14 @@ for lg, meanings in {
     tabular.rename(columns=trans_dic, inplace=True)
     tabular = tabular.apply(lambda x: x.apply(objectify))
     tabular.columns.names = [None]
-    
+
     save_float(
         print_latex(tabular, keep_index=True),
         lg + "reg",
         f"Regular {print_shorthand(lg)} verbs " + sources,
         short=f"Regular {print_shorthand(lg)} verbs",
     )
-    
+
 i_df = i_df[~(pd.isnull(i_df["Verb_Cognateset_ID"]))]
 
 # pekodian lexical comparison
@@ -530,7 +531,7 @@ def print_aligned_table(
     sort_lg(df)
     for i, col in enumerate(df.columns):
         if col == "":
-            df.columns.values[i] = "Alignment"
+            # df.columns.values[i] = "Alignment"
             break
     df["Form"] = df["Form"].str.replace("+", "", regex=True)
     print(f"Comparative table for 'to {verb}'")
@@ -569,12 +570,12 @@ print_aligned_table(
 print_aligned_table(
     v_df[v_df["Parameter_ID"] == "go"],
     verb="go",
-    caption=r"Reflexes of \rc{ɨtə(mə)} \qu{to go}",
[…]

### 2021-10-29 documents/sections/verbs.tex
more work on 'to go down'

@@
 Reflexes of this verb were not affected by the extensions of \rc{k-} in \PPek \pcref{sec:pekodian} and \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
-The resistance against the former extension was overcome in \bakairi.
-In \akuriyo, the verb already had an irregularly inflected first person form with \obj{p-}, inherited from \PTir.
+The resistance against the former extension was subsequently overcome in \bakairi; its fate in \ikpeng is unknown.
+When \akuriyo extended \obj{k-}, the verb already had an irregularly inflected first person form with \obj{p-}, inherited from \PTir.
 
@@
 \pex<gowhere>
-\a<car-18> \carijo \parencite[][102]{guerrero2019carijo}\\
+\a<car-32> \carijo \perscommpar{David Felipe Guerrero}\\
 \begingl
-\gla əji-wa-e j-eh-ɨ//
-\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
-\glft \qu{I came looking for you.}//
-\endgl
-\a<yuk-14> \yukpa \parencite[][]{meira2003primeras}\\
-\begingl
-\glpreamble aw yéwtu//
-\gla aw j-ewuhtu//
-\glb \gl{1}-go.down//
-\glft \qu{I went down.}//
-\endgl
-\xe
-%
-However, a look at the comparative picture suggests a much more complicated situation.
-\cref{tab:godown} shows all attested cognates, some already discussed, including verb class membership and aligned segments.
-While a form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
-The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
-
-In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb in the synchronic analysis of \wayana.
-It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
-It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.
-
-\pex<gowhere>
-\a<car-18> \carijo \parencite[][102]{guerrero2019carijo}\\
-\begingl
-\gla əji-wa-e j-eh-ɨ//
-\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
-\glft \qu{I came looking for you.}//
+\gla irə wat͡ʃinakano tae j-ehɨtə-e//
+\glb \gl{inan}.\gl{ana} body.of.water along.bounded \gl{1}-go.down-\gl{npst}//
+\glft \qu{…I go down through that guachinacán.}//
 \endgl
@@
 %
-Its causativized form is \obj{ɨptəka} \parencite[255]{wayanatavares2005}, containing a reflex of the transitivizer \rc{-ka}, which was restricted to \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
+However, a look at the comparative picture suggests a much more complicated situation.
+\cref{tab:godown} shows all attested forms, including verb class membership where applicable, as well as aligned segments for forms we judge to be fully cognate.
+For those forms, \rc{ɨpɨtə} can be reconstructed to \PC, but different (proto-)languages do not agree about the class of this verb.
+The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
+
+\input{floats/godown}
+
+In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb in the synchronic analysis of \wayana.
+It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
+It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.
+
+\pex<waygodown>\wayana \parencite[][200]{wayanatavares2005}
+\a<way-73>
+\begingl
+\glpreamble ïwïptëë//
+\gla ɨ-w-ɨptə-rɨ//
+\glb \gl{1}-\gl{s_a_}-go.down-\gl{nmlz}//
+\glft \qu{my going down}//
+\endgl
+\a<way-71>
+\begingl
+\gla əw-ɨptə-k//
+\glb \gl{2}-go.down-\gl{imp}//
+\glft \qu{Go down!}//
+\endgl
+\xe
+%
+Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the transitivizer \rc{-ka} was restricted to \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
 This leads us to posit the hypothesis that the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class.
 This in turn implies that all reflexes of this verb with \gl{s_a_} status were members of the \gl{s_p_} class at some point, but switched classes.
-
-\input{floats/godown}
 
@@
 
-These circumstances make it impossible to answer the question of whether it was affected by the extensions in \PTir, \carijo, and \yukpa.
+These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by the extensions in \PTir, \carijo, and \yukpa.
 For \PTir, we cannot establish a relative chronology of the verb class change, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
-It could have been an \gl{s_p_} verb at the time of the extension, it could have been an \gl{s_a_} verb and remained unaffected, or it could have been affected, but with a subsequent erasure of the results by the introduction of \rc{p-}.
-For \carijo and \yukpa, we cannot know whether the verb switched verb class before the breakdown of the entire split-\gl{s} system -- this is a plausible scenario especially in the former case.
-In all three cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, but it is also possible that it was not even a potential target due to its \gl{s_p_} status.
+For \carijo and \yukpa, we cannot know whether the verb switched class before the breakdown of the entire split-\gl{s} system.
+While there is no language-internal evidence for such a switch, \qu{to go down} does have an inclination for class switches; in the case of \carijo, that could have already happened at the \PTar stage.
+In all three cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, resisting and keeping its old marker, but it is also possible that it was not even a potential target due to its \gl{s_p_} status.
 On the other hand, the class change must have happened before the split-up of \trio and \akuriyo, and therefore this verb resisted the extension of \akuriyo \obj{k-}.
-Likewise, the class change must have taken place before the extension of \PPek \rc{k-}, since other scenarios would require an \gl{s_p_} verb taking on regular \gl{s_a_} markers, but selecting the archaic first person marker.
+Likewise, it seems very likely that the class change took place before the extension of \PPek \rc{k-}.
+Otherwise, the newly-turned-\gl{s_a_} verb would have taken on archaic and lexically heavily restricted \rc{w-}, either in \PTar, \PXin, or \arara.
 

### 2021-10-29 documents/floats/pwaireg.tex
some work on 'to bathe'

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-10-29 documents/sections/verbs.tex
some work on 'to bathe'

@@
 \label{sec:bathe}
-This verb only emerged as resisting an extension in the case of \PPek \pcref{sec:pekodian}, but is widely attested in the family.
-Verbs for intransitive \qu{to bathe} are usually based on transitive verbs in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in Pemongan, \panare, \kalina, and \maqui\footnote{For \maqui, note that while intransitive \obj{eʔhi} points to \rc{e-kupi}, transitive \obj{ɨhɨ} looks like a reflex of \rc{pɨ}.} \pcref{tab:bathe}.
+This verb resisted extensions in \PPek \pcref{sec:pekodian} and \akuriyo \cref{sec:akuriyo}.
+Verbs for intransitive \qu{to bathe} are usually based on transitive verbs in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in some Venezuelan languages.
 As we have shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
 Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch, changing \rc{e-} to \rc{i}.
-The reasons for this are unknown; we are not aware of \rc{i-} as a regular development of the detransitivizer in Pekodian, see also \textcite[506]{meira2010origin}.
+The reasons for this are entirely mysterious; we are not aware of \rc{i-} as a reflex of the detransitivizer in Pekodian, see also \textcite[506]{meira2010origin}.
 However, it should be noted that other languages also show unexpected developments in this verb, consider the apparent addition of \obj{ew-} in \hixka or the chaotic mixture of \rc{pɨ} and \rc{kupi} in languages spoken in Venezuela.

### 2021-10-29 workflow/process_data.py

@@ -631,10 +631,12 @@ for table in bathe_tables:
     )
     if table[2] == "bathe_intr_1":
         bathe_out += r"""\begin{subtable}[t]{.49\linewidth}
+\centering
 """
     if table[2] == "bathe_tr_1":
         bathe_out += r"""\end{subtable}
-\begin{subtable}[t]{.49\linewidth}"""
+\begin{subtable}[t]{.49\linewidth}
+\centering"""
     bathe_out += r"""\caption{%s}
 \label{tab:%s}
 %s""" % (

### 2021-10-29 workflow/process_data.py
fixed overview table

@@ -669,31 +669,34 @@ i_df = i_df[i_df["Inflection"] == "1"]
 verb_list = ["say", "go", "be_1", "be_2", "come", "go_down", "bathe_intr"]
 
 # determine whether verb was affected by extensions based on cognacy of prefixes
-def identify_affected(cogset, value, verb):
-    # we either don't know what form occurs or we think no potential form exists
-    if value in ["?", "–"]:
-        return value
-    # we think the verb does not occur at all in the language
-    elif pd.isnull(value):
-        return "–"
+def identify_affected(cogset, row):
+    value = row["Prefix_Cognateset_ID"]
+    verb = row["Verb_Cognateset_ID"]
+    lg = row["Language_ID"]
     # 'go down' used to be an Sp verb, the relative chronology of the class
     # change and the extensions, and whether the class occurred at all
     # are not known in some cases
-    elif verb == "go_down":
+    if verb == "go_down" and lg not in ["PPek", "ara", "ikp", "bak"]:
         # the extension of PTir *t- could have happened before or after the class change
         # the extension of car and yuk j- could have happened before or after the collapse
         # of the system. the verb might have changed class or not.
-        if cogset in ["1t", "1p"]:
+        if (cogset in ["1t", "1p"] or lg in ["PWai", "wai", "hix"]):
             return "N/A"
         # we have to assume that when aku extended k-
         # it was already an Sa verb
-        elif cogset == "k":
+        elif lg == "aku":
             return "n"
         # the verb didn't exist in PWai, and while it could have switched classes
         # *after* the extension of PPek *k-, why would it have acquired the old Sa
         # marker instead of the new one?
         else:
             raise ValueError("This combination should not occur")
+    # we either don't know what form occurs or we think no potential form exists
+    elif value in ["?", "–"]:
+        return value
+    # we think the verb does not occur at all in the language
+    elif pd.isnull(value):
+        return "–"
     else:
         # the new person marker occurs on the verb
         if cogset == value:
@@ -706,6 +709,14 @@ def identify_affected(cogset, value, verb):
         else:
             return "n"
 
+overview_legend = {
[…]

### 2021-10-31 documents/floats/pwaireg.tex
some LaTeX fixes

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[30]{waiwaihawkins1998}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-10-31 documents/sections/verbs.tex
fixed overview table
some LaTeX fixes

@@
 \cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstructed as \gl{s_a_} verbs that were not derived from transitive verbs.
-\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek; \cref{sec:bathe} investigates \PPek \rc{ɨpɨ} \qu{to bathe}, and \cref{sec:shit} \trio \qu{to defecate}.
-%, and \cref{sec:movement} the \akuriyo movement verbs.
+\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek, and  \cref{sec:bathe} investigates \PPek \rc{ɨpɨ} \qu{to bathe}.
 The \obj{e}-initial verbs not affected by the extension of \rc{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
@@
 
-There are many forms going seemingly reflecting \rc{əpɨ} and \rc{epɨ}, distributed widely through the family.
+There are many forms going seemingly reflecting \rc{əpɨ} and \rc{epɨ}, distributed widely in the family.
 A unifying account of these forms sees the root \rc{jəpɨ} undergoing two major sound changes: \begin{inlinelist}
@@
 
-\subsection{\PPek \rc{ipɨ} \qu{to bathe}}
+\subsection{\qu{to bathe}}
 \label{sec:bathe}

### 2021-10-31 documents/sections/proto_carib.tex
went through 3.1-3.4
fine-tuning 2.1-2.2

@@
 \section{Cariban verbal person marking and marker extensions}
-The archaic person markers discussed in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system.
-This section provides the necessary background for said ancestral system \pcref{sec:pc_person}, the split-\gl{s} system which will be relevant later \pcref{sec:split}, and the concept of person marker extensions \pcref{sec:extensions_intro}.
+The archaic first person markers investigated in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system that did not affect the whole lexicon.
+To provide readers with the necessary background, we first introduce said ancestral system \pcref{sec:pc_person}, then the split-\gl{s} system which is very relevant for our topic \pcref{sec:split}, and finally the concept of person marker extensions \pcref{sec:extensions_intro}.
 
@@
 \label{sec:pc_person}
-\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.\footnote{I use standard IPA symbols in my transcription of Cariban languages, with the exception of coronal rhotics, which I simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
-In languages with strong morphophonological processes and/or subphonemic orthography I show the original transcription in an additional surface line when presented in an interlinearized glossed example.
-I follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
+\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.%
+\footnote{We use standard IPA symbols in ourtranscription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in an interlinearized glossed example.
+We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
 The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
-The locuphoric markers, those referring to \gl{sap} participants, had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
-This was not the case for third person referents, which only had one marker, since third person only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, and third-person prefixes did not distinguish different different third persons.
-Local scenarios were expressed in a non-transparent manner, by using the \gl{1+2} prefix \rc{k-} in both cases.\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
-This is expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented in this table.}
+The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
+There was a single aliophoric marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
+Local scenarios were expressed in a non-transparent manner, by using the \gl{1+2} prefix \rc{k-} in both cases.%
+\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
+This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented in this table.}
 
@@
 On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones.
-The third person marker in \gl{s_p_} verbs was identical to the one in \gl{3}>\gl{3} scenarios (\rc{n(i)-)}, while the \gl{3}\gl{s_a_} marker did not have an \rc{i}, like the other \gl{s_a_} prefixes.
+The third person marker in \gl{s_p_} verbs was identical to the one in \gl{3}>\gl{3} scenarios (\rc{n(i)-)}.
 
-As in many other split-\gl{s} systems found in languages of the world, the intransitive verbal lexicon was divided into the two classes, with verbs inherently being \gl{s_a_} or \gl{s_p_} verbs.
-That is, while in transitive verbs, the choice of the person marker had a crucial semantic contribution, it is predictable in intransitive verbs.
+As in many other split-\gl{s} systems found in languages of the world, the intransitive verbal lexicon was divided into two classes, with verbs inherently being members of the \gl{s_a_} or the \gl{s_p_} group.
+That is, while in transitive verbs, the choice of person marker had a crucial semantic contribution, it was predictable in intransitive verbs.
 This is illustrated with modern \kalina data in \exref[kar1]{kar2}.
@@
 Rather, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.\footnote{The root \obj{awomɨ} \qu{to get up} is subject to syllable reduction and assimilation to the prefix-initial \obj{j}.}
-
-In fact, the split-\gl{s} system is not only fully lexically conditioned, but there are clear semantic mismatches between class membership and semantics:
-\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, in fact  \qu{you're afraid of pulling up the net} clearly has a second person patient (experiencer).
-Similarly, \qu{to get up} with a \dbqu{patientive} marker is a clear semantic mismatch as well.
+What is more, the choice of prefix does not necessarily convey information about the semantics of the verb, since there are clear mismatches:
+\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker must be considered volitional.
 These mismatches are not isolated cases, as will be discussed in \cref{sec:split}.
@@
 With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
-Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles} for participles, and in \cref{tab:nominalizations} for nominalizations.
+Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.
 %\kaxui does not preserve \rc{w-} in its participles (\obj{tehurkat͡ʃe} \qu{fallen}, not \ungr{tɨwehurkat͡ʃe}), while \apalai only preserves it there, but not in nominalizations (\obj{jepɨtopo} \qu{my bathing place}, not \ungr{joepɨtopo}).
@@
 
-The distinction between \gl{s_a_} and \gl{s_p_} is also borne out in imperatives, the suffix for which can be reconstructed as \PC \rc{-kə}.
+The distinction between \gl{s_a_} and \gl{s_p_} is also borne out in imperatives (\PC \rc{-kə}).
 Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed.
@@
 %\subsubsection{Underived \gl{s_a_} verbs}
-As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitives, and do not have a reflex of \detrz, either.
+As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
 \textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
 \textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
-These verbs will be discussed in more detail in \cref{sec:verbs}.
+Many of these verbs will be discussed in \cref{sec:verbs}.
 
@@
 
-\begin{table}[h]
+\begin{table}
 \caption[Some examples for completed extensions]{Some examples for completed extensions \parencite[90--92]{gildea1998}}

### 2021-11-01 documents/sections/proto_carib.tex
2.1-2.3

@@
 The archaic first person markers investigated in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system that did not affect the whole lexicon.
-To provide readers with the necessary background, we first introduce said ancestral system \pcref{sec:pc_person}, then the split-\gl{s} system which is very relevant for our topic \pcref{sec:split}, and finally the concept of person marker extensions \pcref{sec:extensions_intro}.
+To provide readers with the necessary background, we first introduce said ancestral system \pcref{sec:pc_person}, then the peculiarities of the Cariban split-\gl{s} system \pcref{sec:split}, and finally the concept of person marker extensions \pcref{sec:extensions_intro}.
 
@@
 \PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.%
-\footnote{We use standard IPA symbols in ourtranscription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+\footnote{We use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
 In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in an interlinearized glossed example.
 We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
-The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
+The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.\footnote{Glossing abbreviations: \glossingAbbrevsComma.}
 The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
@@
 \label{sec:split}
-As seen in the previous section, the split-\gl{s} defined two inflectional classes for intransitive verbs within the \setone system.
+As seen in the previous section, the split-\gl{s} system defined two inflectional classes for intransitive verbs within the \setone system.
 %\kalina paradigms illustrating the split for each person are given in \cref{tab:kalinaparadigms}.
@@
 
-The distinction between \gl{s_a_} and \gl{s_p_} is also borne out in imperatives (\PC \rc{-kə}).
-Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed.
+The distinction between \gl{s_a_} and \gl{s_p_} is also borne out in imperatives.
+Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed (both suffixed with \rc{-kə}).
 This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
@@
 %\subsubsection{The detransitivizing prefix}
-There is one further property uniting \gl{s_a_} verbs, which is not based on inflectional morphology.
+There is one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
 As mentioned in \cref{sec:pc_person}, mismatches between the semantics of intransitive verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common.
@@
 \textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
-\begin{quotation}
-Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still exisiting transitive sources) or diachronically (with reconstructible but no longer existing transitive sources) \parencite[201]{meira2000split}
-\end{quotation}
+\begin{quotebox}{\parencite[201]{meira2000split}}
+Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still exisiting transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
+\end{quotebox}
 \textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
@@
 
-\begin{quotation}
+\begin{quotebox}{\parencite[226]{meira2000split}}
 Apparently, the morphological behavior of the \gl{s_a_} verb class is an accidental consequence of the fact that detransitivization, as far back as we can reconstruct, entails all the morphology described […] as typical of \gl{s_a_} verbs.
 The alignment of person-marking prefixes appears not to be driven by any semantic forces in the language; it is as though they were being dragged by the evolution of the reflexive marker.
-\parencite[226]{meira2000split}
-\end{quotation}
+\end{quotebox}
 
-As for the form of this reflexive marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although they have since merged into a single morpheme in modern Cariban languages.
+As for the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although they have since merged into a single morpheme, apparently in all languages.
 Modern reflexes of \detrz show a range of meanings, which can all be characterized as \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
@@
 The original \PC split-\gl{s} system has been subject to change in many languages, in some even to the point of total loss.
-Besides such modifications as for example losing the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} or losing certain person values altogether, these changes have largely been due to person prefixes being extended to new verbs.\footnote{
-One could also imagine a scenario whereby verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
+Besides modifications like losing the \gl{s_a_} class marker \rc{w-} or losing certain person values altogether, these changes have largely been due to person prefixes being extended to new verbs.\footnote{
+In principle, one could also imagine a scenario in which verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
 However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
 Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
-For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system features markers being extended to new verbs.}
+For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
 There have been many such person marker extensions in Cariban languages, and some are still ongoing.
-This is shown by \textcite{gildea1998}, using the Parukotoan languages as an example.
+This is shown by \textcite{gildea1998}, using the three Parukotoan languages as an example.
 We have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription -- and in the case of \kaxui, the addition of ∅/\obj{j-} as an alternative \gl{1}\gl{s_p_} marker \perscommpar{Spike Gildea}.
@@
 \waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
-In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the system.
+In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
 That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
@@
 \pex<wailoss>
+\begin{multicols}{2}
 \a<wai-158> \waiwai \parencite[][98]{waiwaihawkins1998}\\
@@
 \endgl
+\end{multicols}
 \xe
@@
 
-While different cases of loss of split-\gl{s} are discussed by \textcite[91--96]{gildea1998}, this paper focuses on a so far neglected aspect of these person marking extensions.
-We argue that they are executed via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
-
-First of all, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
+While different cases of loss of split-\gl{s} are discussed by \textcite[91--96]{gildea1998}, this paper focuses on a so far neglected aspect of the person marking extensions resulting in this loss.
+As a prerequisite for our story, we argue that they are executed via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
 Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
@@
 Not all person marker extensions spread through the entire lexicon, but rather stopped short of some verbs.
-This is illustrated for \trio first person forms of \gl{s_a_} verbs in \exref{trio1}; regular \gl{s_a_} verbs take phonologically determined allomorphs \obj{s-} (\envr{}{\obj{e}}) and \obj{t-} (\envr{}{\obj{ə}}), but \obj{tən} \qu{to go} takes an \dbqu{irregular} or \dbqu{archaic} prefix \obj{wɨ-}; it was not affected by the spread of innovative \obj{t-}/\obj{s-} and therefore preserves the old prefix.
+This is illustrated for \trio first person forms of \gl{s_a_} verbs in \exref{trio1}; regular \gl{s_a_} verbs take phonologically determined allomorphs \obj{s-} (\envr{}{\obj{e}}) and \obj{t-} (\envr{}{\obj{ə}}), but \obj{tən} \qu{to go} takes an \dbqu{archaic} or \dbqu{irregular} prefix \obj{wɨ-}; it was not affected by the spread of innovative \obj{t-}/\obj{s-} and therefore preserves the old prefix.
 
@@
 %
-This paper is primarily about these incomplete extensions, and the verbs that were not affected by them.
+These incomplete extensions and the verbs that were not affected by them are the central topic of this paper.
 
@@
 While there have been extensions for other person values as well, they never affect \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affect the entire lexicon, at least based on the attested data.
-Besides the completed extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}, examples include the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set in \waimiri \pcref{tab:waimiri}.
+Illustrative examples for extensions affecting the entire lexicon are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set in \waimiri \pcref{tab:waimiri}.%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
 
@@
 The remainder of this paper is structured as follows:
-In \cref{sec:extensions}, we discuss each of these six incomplete extensions individually, listing all attested unaffected verbs, and reconstructing proto-paradigms where necessary.
-In \cref{sec:verbs}, we take a comparative look at the verbs that were not affected by these extensions and searches for possible motivations.
-Finally, we discuss the findings and put them in a general context of language change and morphology in \cref{sec:discussion}.
+In \cref{sec:extensions}, we discuss each of these six incomplete extensions individually, discussing all attested unaffected verbs, and reconstructing proto-paradigms where necessary.
+In \cref{sec:verbs}, we take a comparative look at the verbs that were not affected by these extensions.
+Finally, we search for factors motivating the resistance of these verbs, discuss the findings and put them in a general context of language change and morphology in \cref{sec:discussion}.
 

### 2021-11-01 documents/floats/go.tex
added apalai ɨto 'to go'

@@
 \centering
-\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6; p.c., Spike Gildea]{franchetto2002kuikuro}}
+\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6]{franchetto2002kuikuro}[99; p.c., Spike Gildea]{camargo2002lexico}}
 \label{tab:go}
@@
 \wayana   &   \obj{[ɨ]tə[m]} &  ɨ &  t &  ə &  m &    \\
+\apalai   &        \obj{ɨto} &  ɨ &  t &  o &    &    \\
 \kalina   &         \obj{to} &    &  t &  o &    &    \\

### 2021-11-01 documents/floats/pwaireg.tex
added apalai ɨto 'to go'

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[30]{waiwaihawkins1998}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-11-01 documents/sections/extensions.tex
some LaTeX fixes
3.1

@@
-\section{The innovative \gl{1}\gl{s_a_} markers}
+\section{The innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
 \label{sec:extensions}
@@
 \textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work on \PPek morphosyntax can be found in the literature.
-All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.\footnote{In \cref{tab:pekreg} and the remainder of this paper, we omit any possible TAM suffixes, since our focus lies on the prefixes and stems, and since full paradigms containing the same TAM suffix are rarely found.}
+All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.\footnote{In \cref{tab:pekreg} and the remainder of this paper, we omit any possible TAM suffixes from the attested forms in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.}
 We can thus reconstruct a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
@@
 For \qu{to be}, \ikpeng \obj{e} is very likely the original vowel, given the \PC form \rc{eti} \pcref{sec:be}.
-
 For \qu{to go down}, we reconstruct \rc{ɨ} as the initial vowel rather than \rc{i}, see \cref{sec:godown}.
@@
 The V-initial nature of \qu{to go} is evidenced in its Xinguan forms, while \bakairi \phon{\rc{wɨ}}{\obj{u}} muddied the morpheme boundary; other forms are V-initial \pcref{sec:go}.
-The \ikpeng form \obj{aran} is compatible with our reconstruction \rc{ɨtən}, when one considers that \ikpeng \obj{a} is an attested outcome of \rc{ə}.\footnote{\begin{inlinelist}
+The \ikpeng form \obj{aran} is compatible with our reconstruction \rc{ɨtən} when considering that \ikpeng \obj{a} is an attested outcome of \rc{ə}.\footnote{\begin{inlinelist}
  \item \obj{akari} \qu{dog} in \cref{tab:pxinw} above

### 2021-11-01 documents/sections/extensions.tex
3.2

@@
 This extension was briefly discussed in \cref{sec:extensions_intro}; the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was later extended to \gl{1}\gl{s_a_} in \PWai.
-This created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \pcref{tab:pwaireg}.
+For regularly inflected verbs, this created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \pcref{tab:pwaireg}.
 %
@@
 %
-Not all verbs were affected by this extension; \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
+Not all verbs were affected by this extension, though: \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
 Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, a potential reflex of \gl{1}\gl{s_a_} \rc{w(ɨ)-}.
@@
 The two roots for \qu{to be} are straightforwardly reconstructible, whereas \qu{to go} is somewhat of a special case.
-While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
-Alternatively, this form was influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \exref{waiwto}.
+While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
+Alternatively, this form might have been influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \exref{waiwto}.
 
@@
 
-Summing up, we reconstruct the four verbs \rc{et͡ʃi} and \rc{a(h)} \qu{to be}, \rc{ka(s)} \qu{to say}, and \rc{tom} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
+Summing up, we reconstruct the four verbs \rc{eʃi} and \rc{a[h]} \qu{to be}, \rc{ka[s]} \qu{to say}, and \rc{[ɨ]to[m]} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
 

### 2021-11-01 documents/sections/extensions.tex
3.3

@@
 The \setone paradigms of \trio and \akuriyo contain an interesting puzzle: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
-This resulted in a regular \gl{1}\gl{s_a_} marker \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}; the latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
+This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}; the latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
 
@@
 As for intransitive verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar; we offer our reconstructed \PTir forms.
-To this list, we add the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}}, which has first person \obj{w-} at least in \trio.
+To this list, we add the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}}, which has first person \obj{w-}, at least in \trio.
 
@@
 %
-For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for the first person, and \rc{eepɨ} for the others, based on the idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
+For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for the other person values, based on an idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
 \akuriyo (and \carijo) then levelled this original distribution, similar to what we have suggested for Pekodian \pcref{sec:pekodian}.
-We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly cognate with \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form must have been \rc{əʔepɨ}.
+We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ}.
 

### 2021-11-01 documents/sections/extensions.tex
some changes to extensions and verbs sections

@@
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} might be more recent, with which we agree -- since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
+He also suggests that \obj{k-} might be more recent, with which we agree; since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
 The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
+Since there are no examples of or information about \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.
 
@@
 
+All verbs in \cref{tab:ptirverbs} above did of course not introduce \obj{k-}, either; we do not know the first-person form of the copular verb \obj{eʔi}.
 In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in \akuriyo, also reconstructible to \PTir \exref{tridown}.
-It was not affected by the extension of \akuriyo \obj{k-}, but its relative chronology to the extension of \PTir \rc{t-} is unclear (see also \cref{sec:godown}).
+It was not affected by the extension of \akuriyo \obj{k-}, but whether it existed when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).
 
-\ex<tridown> Tiriyoan \qu{to go down} \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
+\ex<tridown> First person forms of \qu{to go down} \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
 \begin{tabular}[t]{@{}ll@{}}
@@
 %
-Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by taking a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance} \exref{car-30}.
+Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by taking a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}, shown in \exref{car-30}.
 %The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
@@
 %
-Further, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car-33}.
+In addition, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car-33}.
 %We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
@@
 We analyse this ∅ as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
-Further, one may suggest the difference to C-initial \obj{tə} \qu{to go} to result from that verb's former V-initial nature \pcref{sec:go}.
+Further, one may suggest the difference between C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} to result from the latter's earlier V-initial nature \pcref{sec:go}.
 
@@
 It contrasts with regular C-initial verbs taking \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
-Thus, we conclude that the extension of \obj{j-} in \yukpa only stopped short of \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.
+Thus, we conclude that the extension of \obj{j-} in \yukpa stopped short of \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.
 

### 2021-11-01 documents/sections/verbs.tex
some changes to extensions and verbs sections

@@
 In \cref{sec:extensions}, we introduced six distinct extensions of personal prefixes into \gl{1}\gl{s_a_} territory, and identified verbs resistant to each innovation.
-The set of unaffected verbs was rather small in most cases, and there is a considerable etymological overlap between the (proto-)language-specific verb groups.
-In this section, we present these verbs from a comparative perspective and discuss their reconstructability.
-\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstructed as \gl{s_a_} verbs that were not derived from transitive verbs.
-\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek, and  \cref{sec:bathe} investigates \PPek \rc{ɨpɨ} \qu{to bathe}.
+The set of unaffected verbs was rather small in most cases, and (proto-)languages show a considerable degree of overlap in what verbs are unaffected.
+In this section, we present these verbs from a comparative perspective.
+\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstruct as \gl{s_a_} verbs that were not derived from transitive verbs.
+\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek, and  \cref{sec:bathe} investigates \rc{e-pɨ} \qu{to bathe}, of which the \PPek reflex \rc{i-pɨ} resisted the extension of \rc{k-}.
 The \obj{e}-initial verbs not affected by the extension of \rc{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
 
-\subsection{\rc{eti}/\obj{a[p]} \qu{to be}}
+\subsection{\rc{eti} and \rc{a[p]} \qu{to be}}
 \label{sec:be}
@@
 \akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while \obj{eʔi} is not attested in a first-person form. 
-\carijo innovated \obj{j-}, but only in the \obj{et͡ʃi} root allomorph \exref{carcop.mivida-12}; the \obj{a} root preserves \obj{w-} \exref{carcop.car-25}.
+\carijo innovated \obj{j-}, but only in the \obj{et͡ʃi} root allomorph \exref{mivida-12}; the \obj{a} root preserves \obj{w-} \pcref{sec:carijo}.
 \yukpa innovated \obj{j-} for reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
 
-\pex<carcop>\carijo
-\a<mivida-12>
+\ex<mivida-12>\carijo \parencite[][177]{robayo1989rame}\\
 \begingl
@@
 \glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
-\glft \qu{Then I was already grown up.} \parencite[][177]{robayo1989rame}//
-\endgl
-\a<car-25>
-\begingl
-\gla əji-marə-ne w-a-e//
-\glb \gl{2}-\gl{com}-\gl{pl} \gl{1}-be-\gl{npst}//
-\glft \qu{I am with you all.} \parencite[][42]{guerrero2016karihona}//
+\glft \qu{Then I was already grown up.}//
 \endgl
@@
 \ex<kati> \apalai\\
-\obj{kaʃi-ko} \qu{say!} \parencite[35]{koehn1986apalai}
+\obj{kaʃi-ko} \qu{say!}\\
+\parencite[35]{koehn1986apalai}
 \xe
@@
 %
-This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, and \carijo (\crefrange{sec:pekodian}{sec:carijo}).
-We do not know the first person forms of its \yukpa reflex \obj{ka}.
+This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}).
+We do not know the first person prefix of its \yukpa reflex \obj{ka}.
 
@@
 \label{sec:go}
-This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə[mə]}, with the second syllable only rarely occurring, as in the case of \rc{ka[ti]} \qu{to say}.
+This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə[mə]}, with a fleeting second syllable, like with \rc{ka[ti]} \qu{to say}.
 It is true that many reflexes are clearly \obj{t}-initial, for example \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}.
@@
 
-\subsection{\qu{to bathe}}
+\subsection{\rc{e-pɨ} \qu{to bathe}}
 \label{sec:bathe}

### 2021-11-02 documents/sections/verbs.tex
4.4 'to come'

@@
 \label{sec:go}
-This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə[mə]}, with a fleeting second syllable, like with \rc{ka[ti]} \qu{to say}.
+This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə[mə]}, like \rc{ka[ti]} \qu{to say} with a fleeting second syllable.
 It is true that many reflexes are clearly \obj{t}-initial, for example \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}.
-However, once one considers all forms of the various reflexes of this verb \pcref{tab:go}, an initial vowel \rc{ɨ} must clearly be reconstructed -- in contrast to unambiguously C-initial \rc{ka[ti]} \qu{to say}.\footnote{One may add that many inflected forms, like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} are ambiguous, since an epenthetic \obj{ɨ} breaks up CC clusters on the prefix-verb boundary.} which was subsequently reanalyzed as part of the prefixes in many languages.
+However, once one considers all forms of the various reflexes of this verb \pcref{tab:go}, an initial vowel \rc{ɨ} must clearly be reconstructed -- in contrast to unambiguously C-initial \rc{ka[ti]} \qu{to say}.%
+\footnote{As indicated by the brackets in \cref{tab:go}, there are many languages where the initial vowel is only present in some forms.
+Also, the prefix-verb boundary in many inflected forms like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} is ambiguous, since an epenthetic \obj{ɨ} breaks up potential CC clusters.}
 This verb was not affected by the any of the extensions discussed in \cref{sec:extensions}.
@@
 
-This verb is reconstructed as monomorphemic \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
-Crucially, the \rc{ət} part is not reflected in the majority of forms, and many languages have reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ} for the \rc{epɨ} part.
-We argue that these forms go back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizer and a root \rc{jəpɨ}.
-Evidence for the originally \rc{j}-initial nature of the root is found in the Pemongan languages and \kaxui; the presence of \rc{əpɨ} and \rc{epɨ} reflexes in \kaxui raises the question of whether this \obj{j} is part of the root.
-The \kaxui form \obj{johɨ} only occurs in the Progressive \exref{kaxprog.kax-82}, and indeed \obj{j(-)} may be a reflex the \settwo third person marker \rc{i-} combined with the \gl{s_a_} class marker \rc{w-}, overtly present with other person values \exref{kaxprog.kax-54}.
-However, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
-Thus, it seems more likely that the \obj{j} is indeed part of the root, rather than an outcome of \rc{i-w-}.
+This verb is reconstructed as \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
+Crucially, the \rc{ət} part of their reconstruction is not reflected in the majority of forms, and many languages ostensibly have reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ} for the \rc{epɨ} part.
+We analyze these forms as going back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizer and a root \rc{jəpɨ}.
+
+Evidence for the originally \rc{j}-initial nature of the root is found in the Pemongan languages and \kaxui, although the simultaneous presence of \obj{oohɨ} in \kaxui raises the question of whether the \obj{j} is part of the root in \obj{johɨ}.
+This form only occurs in the Progressive \exref{kaxprog.kax-82}, and indeed \obj{j(-)} may be a reflex of \rc{i-w-} \qu{\gl{3}-\gl{s_a_}-}, as the \gl{s_a_} class marker \rc{w-} is present with other person values \exref{kaxprog.kax-54}.
+However, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial \gl{s_a_} verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
+Thus, we are confident that this \obj{j} is indeed part of the root, rather than an outcome of \rc{i-w-}, which in turn allows us to reconstruct \rc{j} back to \PC.
 
@@
 %
-
-Most forms on which \posscite{gildea2007greenberg} \rc{ətepɨ} seems to be based do not show evidence for \rc{j}.
+Most forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for this \rc{j}.
 However, we interpret the \obj{i} in the \akawaio form \obj{əsipɨ} as additional evidence of the sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of bare \rc{jəpɨ}, \obj{ipɨ}.
 
-There are many forms going seemingly reflecting \rc{əpɨ} and \rc{epɨ}, distributed widely in the family.
-A unifying account of these forms sees the root \rc{jəpɨ} undergoing two major sound changes: \begin{inlinelist}
+Turning to the many forms apparently reflecting \rc{əpɨ} and \rc{epɨ}, we find that both are distributed widely in the family, sometimes even co-occurring in the same language.
+A unifying account of these forms requires the root \rc{jəpɨ} to undergo two major sound changes: \begin{inlinelist}
  \item \rc{j}-loss
@@
  \end{inlinelist}.
-Both phenomena are widespread in the family \parencite{meira2010origin}.
-However, these sound changes appear to have applied irregularly, and not always in the same order.
-For example, the \kalina form \obj{opɨ} can only be explained if \rc{j} was lost before the umlaut of \rc{ə} to \rc{e}, which happened elsewhere after \rc{j}.
+Both phenomena are found in other contexts throughout the family \parencite{meira2010origin}.
+For this particular verb, these sound changes appear to have applied irregularly, and not always in the same order.
+For example, the \kalina form \obj{opɨ} can only be explained if \rc{j} was lost before the umlaut of \rc{ə} to \rc{e}, which would have been triggered by \rc{j}.
 On the other hand, forms like \maqui \obj{ehə} must be the result of \phonl{\rc{ə}}{\rc{e}}{\rc{j}}, with subsequent loss of \rc{j}.
@@
 Although the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
-For example, \gl{s_p_} \rc{wənɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
-\waiwai \qu{go to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
+For example, the \PC \gl{s_p_} verb \rc{wɨnɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
+Also, \waiwai \qu{go to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
 The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced by its status in \kaxui, \kalina, \panare \exref{pan-128}, \arara, and \trio.
@@
 While these sound changes and the addition of \rc{ət-} do account for the majority of the forms in \cref{tab:come},\footnote{
-Apart from aforementioned \akuriyo \obj{eepɨ}, another exception is the \apalai form \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
+Apart from aforementioned \akuriyo \obj{eepɨ}, another irregular form is \apalai \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
 While \obj{oepɨ} would be a regular outcome of \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
@@
 the distribution within the family is rather chaotic.
-Not only do very closely related languages show different forms, like \yawarana and \mapoyo, but distinct forms can even be found within the same language, usually conditioned by different prefixes.
-This was discussed in \cref{sec:pekodian} for \arara, which has reflexes of \rc{ət-jəpɨ} and \rc{jəpɨ} within the same paradigm.
-A similar situation is found in \trio, where the \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}
+Apart from the distribution of \rc{əpɨ} and \rc{epɨ}, forms with and without \rc{ət-} can be found within the same language, usually conditioned by different prefixes.
+This was briefly discussed in \cref{sec:pekodian} for \arara and in \cref{sec:taranoan} for \trio (and \PTir and \PTar).
+To illustrate, the \trio \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}%
+\footnote{It should be noted that the paradigmatic distributions of forms with and without \rc{ət-} in different languages share no discernible patterns.}
 
@@
 \xe
-%
+
 Summing up, this verb is highly irregular, both from a synchronic and diachronic perspective.
@@
 
-As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \arara reflex of \rc{jəpɨ}, but not on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}.
+As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}, but not on the \arara reflex of \rc{jəpɨ}.
 Both reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ} and/or \rc{epɨ} (\akuriyo) resisted the introduction of \rc{t-} in \PTir.

### 2021-11-02 documents/sections/verbs.tex
'to go down', 'to bathe', discussion, correlation extension -- explanatory factors

@@
 
-At first sight, it may seem that it was affected by the independent extensions of \obj{j-} in \carijo \exref{gowhere.car-18} and \yukpa \exref{gowhere.yuk-14}.
+At first sight, it may seem that it also was affected by the independent extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
 
@@
 However, a look at the comparative picture suggests a much more complicated situation.
-\cref{tab:godown} shows all attested forms, including verb class membership where applicable, as well as aligned segments for forms we judge to be fully cognate.
-For those forms, \rc{ɨpɨtə} can be reconstructed to \PC, but different (proto-)languages do not agree about the class of this verb.
-The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
+\cref{tab:godown} shows all attested forms, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.
+It turns out that while a form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
+Its reflexes in languages that preserve the split-\gl{s} system are split fairly evenly between \gl{s_a_} and \gl{s_p_}.
 
@@
 
-In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb in the synchronic analysis of \wayana.
+In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb in our synchronic analysis of \wayana.
 It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
@@
 %
-Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the transitivizer \rc{-ka} was restricted to \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
-This leads us to posit the hypothesis that the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class.
-This in turn implies that all reflexes of this verb with \gl{s_a_} status were members of the \gl{s_p_} class at some point, but switched classes.
+Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the restriction of \rc{-ka} to \gl{s_p_} verbs in \PC \parencite{gildea2019overview} points to \gl{s_p_} membership.
+These patterns lead us to posit the hypothesis that the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class.
+This in turn implies that all reflexes of this verb with \gl{s_a_} status at some point switched from the \gl{s_p_} class.
 
@@
 The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:godown} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
-Both causativized forms contain a reflex of the (rare) transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
+Both causativized forms contain a reflex of the transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
 Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms \parencite[263]{triomeira1999}.
@@
 
-These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by the extensions in \PTir, \carijo, and \yukpa.
+These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by the extensions in \PWai, \PTir, \carijo, and \yukpa.
 For \PTir, we cannot establish a relative chronology of the verb class change, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
-For \carijo and \yukpa, we cannot know whether the verb switched class before the breakdown of the entire split-\gl{s} system.
+For \PWai, we lack knowledge not only about the first person form, but even about class membership (in \hixka).
+For \carijo and \yukpa, we cannot know whether the verb potentially switched class before the breakdown of the entire split-\gl{s} system.
 While there is no language-internal evidence for such a switch, \qu{to go down} does have an inclination for class switches; in the case of \carijo, that could have already happened at the \PTar stage.
-In all three cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, resisting and keeping its old marker, but it is also possible that it was not even a potential target due to its \gl{s_p_} status.
+In all four cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, resisting and keeping its old marker, but it is also possible that it was not even a potential target due to its \gl{s_p_} status.
 On the other hand, the class change must have happened before the split-up of \trio and \akuriyo, and therefore this verb resisted the extension of \akuriyo \obj{k-}.
@@
 \label{sec:bathe}
-This verb resisted extensions in \PPek \pcref{sec:pekodian} and \akuriyo \cref{sec:akuriyo}.
-Verbs for intransitive \qu{to bathe} are usually based on transitive verbs in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in some Venezuelan languages.
+This verb resisted extensions in \PPek \pcref{sec:pekodian} and \akuriyo \pcref{sec:akuriyo}, but is a regularly inflected \gl{s_a_} verb elsewhere.
+Verbs for intransitive \qu{to bathe} are usually detransitivized derivations of transitive roots in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
 As we have shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
@@
 The reasons for this are entirely mysterious; we are not aware of \rc{i-} as a reflex of the detransitivizer in Pekodian, see also \textcite[506]{meira2010origin}.
-However, it should be noted that other languages also show unexpected developments in this verb, consider the apparent addition of \obj{ew-} in \hixka or the chaotic mixture of \rc{pɨ} and \rc{kupi} in languages spoken in Venezuela.
+However, it should be noted that other languages also show unexpected developments in this verb, considering the apparent glide insertion in Waiwaian or the chaotic distribution of \rc{pɨ} and \rc{kupi} in Venezuelan languages.
 

### 2021-11-02 documents/floats/pwaireg.tex
comparing factors across extensions and verbs

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-11-02 workflow/process_data.py

@@ -190,8 +190,12 @@ def extract_sources(df, src_str="Source", keep=False):
     return sources
 
 
+def proto_lg(lg):
+    return lg[0] == "P"
+
+
 def get_obj_str(lg):
-    if lg[0] == "P":
+    if proto_lg(lg):
         return "rc"
     else:
         return "obj"
@@ -236,418 +240,418 @@ def reconstructed_form_table(lgs, proto, verbs, caption, name):
     pyd.filters = {}
 
 
-lgs = ["PPek", "ara", "ikp", "bak"]
-verbs = ["say", "bathe_intr", "be_1", "be_2", "come", "go_down", "go"]
-reconstructed_form_table(
-    lgs,
-    "PPek",
-    verbs,
-    r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek",
-    "ppekverbs",
-)
-
-lgs = ["PWai", "hix", "wai"]
-verbs = ["say", "be_1", "be_2", "go"]
-reconstructed_form_table(
-    lgs,
-    "PWai",
-    verbs,
-    r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai",
-    "pwaiverbs",
-)
-
-reconstructed_form_table(
-    ["PTir", "tri", "aku"],
-    "PTir",
-    ["go", "say", "come", "be_1", "be_2"],
-    r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir",
-    "ptirverbs",
-)
-
-akuverbs = i_df[
-    (i_df["Language_ID"] == "aku") & (i_df["Prefix_Cognateset_ID"].isin(["k", "1t"]))
-]
-akuverbs["Form"] = akuverbs["Form"].apply(lambda x: x.split("-", 1)[1])
-akuverbs["Meaning"] = akuverbs["Meaning"].map(mean_dic)
-akuverbs["String"] = akuverbs.apply(combine_form_meaning, axis=1)
-
-k_list = akuverbs[akuverbs["Prefix_Cognateset_ID"] == "k"]["String"].reset_index(
-    drop=True
-)
-t_list = akuverbs[akuverbs["Prefix_Cognateset_ID"] == "1t"]["String"].reset_index(
-    drop=True
-)
-sources = extract_sources(akuverbs)
-out = pd.DataFrame.from_dict(
[…]

### 2021-11-03 documents/sections/proto_carib.tex
discussion and factor comparison

@@
 We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
-The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.\footnote{Glossing abbreviations: \glossingAbbrevsComma.}
+The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.\footnote{Glossing abbreviations: }%\glossingAbbrevsComma.}
 The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.

### 2021-11-03 workflow/process_data.py

@@ -94,7 +94,7 @@ def print_latex(df, ex=False, keep_index=False):
         del lines[-2]
         return "\n".join(lines)
     else:
-        return "\n".join(lines).replace(r"\toprule", "\mytoprule")
+        return "\n".join(lines).replace(r"\toprule", "\mytoprule").replace("%", "\\%")
 
 
 def save_float(tabular, label, caption, filename=None, short=None):
@@ -851,7 +851,7 @@ def m_expl(row):
 cond_map = {
     "yuk_j": ["ə", "e", "a"],
     "aku_k": ["ə"],
-    "car_j": ["e"],
+    "car_j": ["e", "ə"],
     "ppek_k": ["ə", "e"],
     "pwai_k": ["o", "e"],
     "ptir_t": ["ə", "e"],
@@ -873,7 +873,7 @@ def p_expl(row):
 freq_map = {
     "be": 1,
     "go": 1,
-    "come": 0.5,
+    "come": 0, # or 0.5?
     "say": 1,
     "bathe_intr": 0,
     "go_down": 0,
@@ -881,12 +881,21 @@ freq_map = {
 
 
 def f_expl(row):
-    return freq_map[row["Meaning"]]
+    aff = freq_map[row["Meaning"]]
+    if aff == 0.5:
+        return 1
+    if row["Affected"] == "y":
+        if aff == 0:
+            return 1
+        elif aff == 1:
+            return 0
+    else:
+        return aff
 
 
-affectedness["Morphological"] = affectedness.apply(m_expl, axis=1)
-affectedness["Phonological"] = affectedness.apply(p_expl, axis=1)
-affectedness["Frequency"] = affectedness.apply(f_expl, axis=1)
+affectedness["morphological"] = affectedness.apply(m_expl, axis=1)
+affectedness["phonological"] = affectedness.apply(p_expl, axis=1)
+affectedness["frequency"] = affectedness.apply(f_expl, axis=1)
 affectedness.drop(columns=["Verb_Cognateset_ID", "Concept", "Meaning"], inplace=True)
 
 
@@ -911,7 +920,7 @@ def get_print(x, factor):
 
 affected_result = pd.DataFrame()
 
-for factor in ["Morphological", "Phonological", "Frequency"]:
+for factor in ["morphological", "phonological", "frequency"]:
     affected_result[factor] = affectedness.groupby("Extension_ID").apply(
         get_print, factor
     )

### 2021-11-04 workflow/process_data.py
discussion, script fixes

@@ -54,7 +54,7 @@ cognum = dict(zip(cognatesets["ID"], cognatesets.index.astype(str)))
 numcog = dict(zip(cognatesets.index.astype(str), cognatesets["ID"]))
 
 cog_verb_df = cs_df[~(cs_df["Form"].str.contains("-"))]
-cog_trans_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Meaning_ID"]))
+cog_trans_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Meaning"]))
 cog_form_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Form"]))
 cog_form_dic["bathe_intr"] = "e-pɨ"
 cog_form_dic["come"] = "(ət-)jəpɨ"
@@ -239,420 +239,421 @@ def reconstructed_form_table(lgs, proto, verbs, caption, name):
     pyd.filters = {}
 
 
-# lgs = ["PPek", "ara", "ikp", "bak"]
-# verbs = ["say", "bathe_intr", "be_1", "be_2", "come", "go_down", "go"]
-# reconstructed_form_table(
-#     lgs,
-#     "PPek",
-#     verbs,
-#     r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek",
-#     "ppekverbs",
-# )
-#
-# lgs = ["PWai", "hix", "wai"]
-# verbs = ["say", "be_1", "be_2", "go"]
-# reconstructed_form_table(
-#     lgs,
-#     "PWai",
-#     verbs,
-#     r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai",
-#     "pwaiverbs",
-# )
-#
-# reconstructed_form_table(
-#     ["PTir", "tri", "aku"],
-#     "PTir",
-#     ["go", "say", "come", "be_1", "be_2"],
-#     r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir",
-#     "ptirverbs",
-# )
-#
-# akuverbs = i_df[
-#     (i_df["Language_ID"] == "aku") & (i_df["Prefix_Cognateset_ID"].isin(["k", "1t"]))
-# ]
-# akuverbs["Form"] = akuverbs["Form"].apply(lambda x: x.split("-", 1)[1])
-# akuverbs["Meaning"] = akuverbs["Meaning_ID"].map(mean_dic)
[…]

### 2021-11-05 documents/floats/pwaireg.tex
data output

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-11-05 workflow/process_data.py

@@ -1,6 +1,7 @@
 import pyradigms as pyd
 import warnings
 import json
+
 warnings.filterwarnings("ignore")
 import pandas as pd
 
@@ -77,13 +78,39 @@ def segmentify(form):
     form = form.strip("+")
     return t(form)
 
+
 exported_tables = {}
-def export_csv(tabular, label, caption=None, keep_index=False):
-    if caption:
-        print(f"{caption}: ")
-    print(tabular)
+
+
+def export_csv(
+    tabular,
+    label,
+    caption=None,
+    keep_index=False,
+    sources=None,
+    print_i_name=False,
+    print_c_name=False,
+):
+    tabular = tabular.copy()
+    # if caption:
+    #     print(f"{caption}: ")
+    if not print_i_name:
+        tabular.index.name = ""
+    if not print_c_name:
+        if type(tabular.columns) == pd.MultiIndex:
+            tabular.columns.names = [None for x in tabular.columns.names]
+        # for c in tabular.columns:
+        #     if type(c) == tuple:
+        #         c = (None for x in c)
+
+    # print(tabular)
     tabular.to_csv(f"data_output/{label}.csv", index=keep_index)
-    exported_tables[label] = caption
+    exported_tables[label] = {"caption": caption}
+    if sources:
+        src = cldfh.combine_refs(sources)
+        # print("(" + ", ".join(src) + ")")
+        exported_tables[label]["sources"] = src
+
 
 # functions for creating latex tables
 def print_latex(df, ex=False, keep_index=False):
@@ -167,7 +194,7 @@ def get_sources(df, parens=True, latexify=True):
     pyd.content_string = tmp
     ref_list = [j for i in table.values.tolist() for j in i]
     if not latexify:
-        return ref_list
+        return cldfh.combine_refs(ref_list)
     source_string = cldfh.cite_a_bunch(ref_list, parens=parens)
     return source_string
 
@@ -184,15 +211,19 @@ def print_shorthand(abbrev):
 
 shorthand_dic = {x: print_shorthand(x) for x in lg_list}
 
+name_dic = {x: cah.get_name(x) for x in lg_list}
[…]

### 2021-11-06 documents/floats/pwaireg.tex
new captions

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-11-06 workflow/process_data.py

@@ -198,6 +198,11 @@ def get_sources(df, parens=True, latexify=True):
     source_string = cldfh.cite_a_bunch(ref_list, parens=parens)
     return source_string
 
+def repl_latex(string):
+    string = re.sub(r"\\rc\{(.*?)\}", r"\**\1*", string)
+    string = re.sub(r"\\obj\{(.*?)\}", r"*\1*", string)
+    string = re.sub(r"\\qu\{(.*?)\}", r"'\1'", string)
+    return string
 
 def sort_lg(df):
     df.Language_ID = df.Language_ID.astype("category")
@@ -631,7 +636,7 @@ def print_aligned_table(
     export_csv(
         df.replace({"Language_ID": name_dic}).rename(columns={"Language_ID": "Language"}),
         verb,
-        f"Reflexes of 'to {verb}'",
+        repl_latex(caption),
         sources = raw_sources
     )
     add_obj_markdown(df)
@@ -662,7 +667,7 @@ def print_aligned_table(
 print_aligned_table(
     v_df[v_df["Parameter_ID"] == "come"],
     verb="come",
-    caption=r"Reflexes of \qu{to come}",
+    caption=r"Reflexes of \rc{(ət-)jəpɨ} \qu{to come}",
     fuzzy=True,
 )
 print_aligned_table(

### 2021-11-06 documents/floats/pwaireg.tex
better captions

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[30]{waiwaihawkins1998}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-11-06 workflow/process_data.py

@@ -198,8 +198,12 @@ def get_sources(df, parens=True, latexify=True):
     source_string = cldfh.cite_a_bunch(ref_list, parens=parens)
     return source_string
 
+def upper_repl(match):
+     return match.group(1).upper()
+     
 def repl_latex(string):
     string = re.sub(r"\\rc\{(.*?)\}", r"\**\1*", string)
+    string = re.sub(r"\\gl\{(.*?)\}", upper_repl, string)
     string = re.sub(r"\\obj\{(.*?)\}", r"*\1*", string)
     string = re.sub(r"\\qu\{(.*?)\}", r"'\1'", string)
     return string
@@ -510,7 +514,7 @@ label = "pxinw"
 export_csv(
     tabular,
     label,
-    "Loss of *w in Ikpeng",
+    "Loss of \**w* in Ikpeng",
     keep_index=True,
     sources=get_sources(forms, latexify=False),
 )
@@ -566,7 +570,7 @@ export_csv(
         columns={"Language_ID": "Language"}
     ),
     label,
-    "Reflexes of *ɨpɨtə 'to go down'",
+    "Reflexes of \**ɨpɨtə* 'to go down'",
     sources=raw_sources,
 )
 

### 2021-11-06 workflow/process_data.py
better captions

@@ -63,6 +63,8 @@ cog_trans_dic["bathe_intr"] = r"bathe"
 cog_trans_dic["be_1"] = r"be-1"
 cog_trans_dic["be_2"] = r"be-2"
 
+def print_shorthand(abbrev):
+    return "\\" + cah.get_shorthand(abbrev)
 
 def str2numcog(cogsets):
     return " ".join([cognum[x] for x in cogsets.split("+")])
@@ -213,16 +215,6 @@ def sort_lg(df):
     df.Language_ID.cat.set_categories(lg_list, ordered=True, inplace=True)
     df.sort_values(["Language_ID"], inplace=True)
 
-
-def print_shorthand(abbrev):
-    return "\\" + cah.get_shorthand(abbrev)
-
-
-shorthand_dic = {x: print_shorthand(x) for x in lg_list}
-
-name_dic = {x: cah.get_name(x) for x in lg_list}
-
-
 def extract_sources(df, src_str="Source", keep=False, latexify=True):
     df[src_str] = df[src_str].fillna("")
     src = list(df[src_str])
@@ -257,6 +249,19 @@ def repl_lg_id(df):
     df["Language_ID"] = df["Language_ID"].apply(print_shorthand)
     df.rename(columns={"Language_ID": "Language"}, inplace=True)
 
+
+shorthand_dic = {x: print_shorthand(x) for x in lg_list}
+name_dic = {x: cah.get_name(x) for x in lg_list}
+
+ext_lg_dic = {x: y for x, y in dict(zip(e_df["ID"], e_df["Language_ID"])).items()}
+ext_form_dic = {x: y for x, y in dict(zip(e_df["ID"], e_df["Form"])).items()}
+
+def extension_string(id, latex=True):
+    form = objectify(ext_form_dic[id], obj_string=get_obj_str(ext_lg_dic[id]))
+    if latex:
+        return print_shorthand(ext_lg_dic[id]) + " " + form
+    else:
+        return name_dic[ext_lg_dic[id]] + " " + repl_latex(form)
 
 # regular pekodian verbs
 tabular = pd.DataFrame()
@@ -953,7 +958,6 @@ def m_expl(row):
             or row["Form"].split("-")[1][0] == "i"
         )
 
-
 cond_map = {
     "yuk_j": ["ə", "e", "a"],
     "aku_k": ["ə"],
@@ -1031,19 +1035,16 @@ for factor in ["morphological", "phonological", "frequency"]:
 
[…]

### 2021-11-06 workflow/process_data.py
better captions

@@ -1045,6 +1045,7 @@ aff_export.index = aff_export.index.map(lambda x: extension_string(x, latex=Fals
 export_csv(aff_export, label, caption, keep_index=True)
 
 affected_result.index = affected_result.index.map(extension_string)
+affected_result.index.name=None
 
 save_float(
     print_latex(affected_result, keep_index=True),

### 2021-11-06 workflow/process_data.py
better overview table

@@ -907,28 +907,36 @@ result.sort_index(inplace=True, level="Lg")
 result.replace({"": "–"}, inplace=True)
 result.index.names = ["", "", ""]
 label = "overview"
-export_csv(
-    result, label, "Overview of extensions and (un-)affected verbs", keep_index=True
-)
-
-# format for latex
-# rename index
-def modify_index(idx):
-    o = get_obj_str(idx[2][0])
-    if idx[0] != idx[2]:
-        return "\\quad " + print_shorthand(idx[2])
-    else:
-        return print_shorthand(idx[0]) + " " + "\\%s{%s}" % (o, idx[1])
-    return idx
-
-
-result.index = result.index.map(modify_index)
 
 # replace cogset ids with reconstructed forms and translationsm
 result.columns = [
     result.columns.map(lambda x: f"\\rc{{{cog_form_dic[x]}}}"),
     result.columns.map(lambda x: f"\\qu{{{cog_trans_dic[x]}}}"),
 ]
+
+def modify_index(idx, latex=True):
+    o = get_obj_str(idx[2][0])
+    if idx[0] != idx[2]:
+        if latex:
+            return "\\quad " + print_shorthand(idx[2])
+        else:
+            return "    " + name_dic[idx[2]]
+    else:
+        if latex:
+            return print_shorthand(idx[0]) + " " + "\\%s{%s}" % (o, idx[1])
+        else:
+            return name_dic[idx[0]] + " " + repl_latex("\\%s{%s}" % (o, idx[1]))
+    return idx
+
+result_exp = result.copy()
+
+result.index = result.index.map(modify_index)
+result_exp.index = result_exp.index.map(lambda x: modify_index(x, latex=False))
+
+export_csv(
+    result_exp, label, "Overview of extensions and (un-)affected verbs", keep_index=True
+)
+
 # add nice-looking checkmarks and stuff
 repl_dic = {x: y["tex"] for x, y in overview_legend.items()}
 result.replace(

### 2021-11-06 workflow/process_data.py
md

@@ -908,12 +908,6 @@ result.replace({"": "–"}, inplace=True)
 result.index.names = ["", "", ""]
 label = "overview"
 
-# replace cogset ids with reconstructed forms and translationsm
-result.columns = [
-    result.columns.map(lambda x: f"\\rc{{{cog_form_dic[x]}}}"),
-    result.columns.map(lambda x: f"\\qu{{{cog_trans_dic[x]}}}"),
-]
-
 def modify_index(idx, latex=True):
     o = get_obj_str(idx[2][0])
     if idx[0] != idx[2]:
@@ -932,6 +926,18 @@ result_exp = result.copy()
 
 result.index = result.index.map(modify_index)
 result_exp.index = result_exp.index.map(lambda x: modify_index(x, latex=False))
+
+# replace cogset ids with reconstructed forms and translationsm
+result.columns = [
+    result.columns.map(lambda x: f"\\rc{{{cog_form_dic[x]}}}"),
+    result.columns.map(lambda x: f"\\qu{{{cog_trans_dic[x]}}}"),
+]
+
+# replace cogset ids with reconstructed forms and translationsm
+result_exp.columns = [
+    result_exp.columns.map(lambda x: f"\**{cog_form_dic[x]}*"),
+    result_exp.columns.map(lambda x: f"'{cog_trans_dic[x]}'"),
+]
 
 export_csv(
     result_exp, label, "Overview of extensions and (un-)affected verbs", keep_index=True

### 2021-11-07 documents/floats/hixintro.tex
tables for introduction

\begin{table}
\centering
\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510, 511, 513]{howard2001wrought}[27, 69, 70]{hixkaryanaderby1985}}
\label{tab:hixintro}
\begin{tabular}[t]{@{}lllll@{}}
\mytoprule
{} &     \qu{to fall} &  \qu{to be afraid} &          \qu{to walk} &    \qu{to go} \\
\midrule
\gl{1}   &  \obj{k-ehurka-} &  \obj{k-oserʲehɨ-} &  \obj{k-atarʲeknohɨ-} &   \obj{ɨ-to-} \\
\gl{2}   &  \obj{m-ehurka-} &  \obj{m-oserʲehɨ-} &  \obj{m-atarʲeknohɨ-} &  \obj{mɨ-to-} \\
\gl{1+2} &  \obj{t-ehurka-} &  \obj{t-oserʲehɨ-} &  \obj{t-atarʲeknohɨ-} &  \obj{tɨ-to-} \\
\gl{3}   &  \obj{ɲ-ehurka-} &  \obj{n-oserʲehɨ-} &  \obj{n-atarʲeknohɨ-} &   \obj{n-to-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-11-07 documents/floats/triintro.tex
tables for introduction

\begin{table}
\centering
\caption[Some \trio verbs]{Some \trio verbs \parencites[292, 294]{triomeira1999}}
\label{tab:triintro}
\begin{tabular}[t]{@{}lllll@{}}
\mytoprule
{} &      \qu{to sleep} & \qu{to bathe (\gl{intr})} &     \qu{to yawn} &    \qu{to say} \\
\midrule
\gl{1}   &    \obj{t-əənɨkɨ-} &              \obj{s-epɨ-} &  \obj{s-entapo-} &   \obj{wɨ-ka-} \\
\gl{2}   &    \obj{m-əənɨkɨ-} &              \obj{m-epɨ-} &  \obj{m-entapo-} &   \obj{mɨ-ka-} \\
\gl{1+2} &  \obj{kɨt-əənɨkɨ-} &             \obj{ke-epɨ-} &  \obj{k-entapo-} &  \obj{kɨh-ka-} \\
\gl{3}   &    \obj{n-əənɨkɨ-} &              \obj{n-epɨ-} &  \obj{n-entapo-} &    \obj{n-ka-} \\
\bottomrule
\end{tabular}
\end{table}

### 2021-11-07 workflow/compile_cldf_irregular.py

@@ -296,7 +296,7 @@ bib_data = bib_parser.parse_file(bibfile)
 filtered_bib_data = BibliographyData()
 for key, entry in bib_data.entries.items():
     if key in found_refs:
-        print(entry)
+        # print(entry)
         filtered_bib_data.add_entry(entry.key, entry)
 
 s = filtered_bib_data.to_string("bibtex")

### 2021-11-07 workflow/process_data.py

@@ -83,7 +83,6 @@ def segmentify(form):
 
 exported_tables = {}
 
-
 def export_csv(
     tabular,
     label,
@@ -263,223 +262,26 @@ def extension_string(id, latex=True):
     else:
         return name_dic[ext_lg_dic[id]] + " " + repl_latex(form)
 
-# regular pekodian verbs
-tabular = pd.DataFrame()
-sources_raw = []
-pyd.x = ["Language_ID", "Meaning_ID"]
-pyd.y = ["Inflection"]
-for lg, meaning in {"bak": "go_up", "ara": "dance", "ikp": "run"}.items():
-    pyd.filters = {"Language_ID": [lg], "Meaning_ID": [meaning]}
-    pyd.sort_orders = {"Inflection": ["1", "2", "1+2", "3"]}
-    temp = pyd.compose_paradigm(i_df, multi_index=True)
-    sources_raw.extend(get_sources(i_df, latexify=False))
-    tabular = pd.concat([tabular, temp], axis=1)
-
-sources = cldfh.cite_a_bunch(sources_raw, parens=True)
-label = "pekreg"
-tabular.index.names = [None]
-export_csv(
-    tabular.rename(columns=name_dic, level="Language_ID"),
-    label,
-    "Regular Pekodian Sa verbs",
-    keep_index=True,
-    sources=sources_raw,
-)
-
-tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name[1])))
-tabular.rename(columns=trans_dic, level="Meaning_ID", inplace=True)
-tabular.rename(columns=shorthand_dic, level="Language_ID", inplace=True)
-
-tabular.columns.names = [None, None]
-tabular.index = tabular.index.map(pynt.get_expex_code)
-tabular.columns = [" ".join([x, y]) for x, y in tabular.columns]
-
-
-save_float(
-    print_latex(tabular, keep_index=True),
-    label,
-    "Regular Pekodian \gl{s_a_} verbs " + sources,
-    short="Regular Pekodian \gl{s_a_} verbs",
-)
-
-
-def reconstructed_form_table(lgs, proto, verbs, caption, name):
-    pyd.x = ["Language_ID"]
-    pyd.y = ["Concept"]
[…]

### 2021-11-12 documents/floats/hixintro.tex
many changes and additions to introduction

@@
 \centering
-\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510, 511, 513]{howard2001wrought}[27, 69, 70]{hixkaryanaderby1985}}
+\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510, 511, 513, 520]{howard2001wrought}[197, 198]{hixkaryanaderby1985}}
 \label{tab:hixintro}
-\begin{tabular}[t]{@{}lllll@{}}
+\begin{tabular}[t]{@{}llllll@{}}
 \mytoprule
-{} &     \qu{to fall} &  \qu{to be afraid} &          \qu{to walk} &    \qu{to go} \\
+{} &     \qu{to fall} &  \qu{to be afraid} &          \qu{to walk} & \qu{to cut self} &    \qu{to be} \\
 \midrule
-\gl{1}   &  \obj{k-ehurka-} &  \obj{k-oserʲehɨ-} &  \obj{k-atarʲeknohɨ-} &   \obj{ɨ-to-} \\
-\gl{2}   &  \obj{m-ehurka-} &  \obj{m-oserʲehɨ-} &  \obj{m-atarʲeknohɨ-} &  \obj{mɨ-to-} \\
-\gl{1+2} &  \obj{t-ehurka-} &  \obj{t-oserʲehɨ-} &  \obj{t-atarʲeknohɨ-} &  \obj{tɨ-to-} \\
-\gl{3}   &  \obj{ɲ-ehurka-} &  \obj{n-oserʲehɨ-} &  \obj{n-atarʲeknohɨ-} &   \obj{n-to-} \\
+\gl{1}   &  \obj{k-ehurka-} &  \obj{k-oserʲehɨ-} &  \obj{k-atarʲeknohɨ-} &   \obj{k-atama-} &  \obj{w-eʃe-} \\
+\gl{2}   &  \obj{m-ehurka-} &  \obj{m-oserʲehɨ-} &  \obj{m-atarʲeknohɨ-} &   \obj{m-atama-} &  \obj{m-eʃe-} \\
+\gl{1+2} &  \obj{t-ehurka-} &  \obj{t-oserʲehɨ-} &  \obj{t-atarʲeknohɨ-} &   \obj{t-atama-} &  \obj{t-eʃe-} \\
+\gl{3}   &  \obj{ɲ-ehurka-} &  \obj{n-oserʲehɨ-} &  \obj{n-atarʲeknohɨ-} &   \obj{n-atama-} &  \obj{n-eʃe-} \\
 \bottomrule

### 2021-11-12 documents/floats/triintro.tex
many changes and additions to introduction

@@
 \centering
-\caption[Some \trio verbs]{Some \trio verbs \parencites[292, 294]{triomeira1999}}
+\caption[Some \trio verbs]{Some \trio verbs \parencites[292, 294]{triomeira1999}[274]{triocarlin2004}}
 \label{tab:triintro}
-\begin{tabular}[t]{@{}lllll@{}}
+\begin{tabular}[t]{@{}llllll@{}}
 \mytoprule
-{} &      \qu{to sleep} & \qu{to bathe (\gl{intr})} &     \qu{to yawn} &    \qu{to say} \\
+{} &      \qu{to sleep} & \qu{to see self} & \qu{to bathe (\gl{intr})} &     \qu{to yawn} &     \qu{to go} \\
 \midrule
-\gl{1}   &    \obj{t-əənɨkɨ-} &              \obj{s-epɨ-} &  \obj{s-entapo-} &   \obj{wɨ-ka-} \\
-\gl{2}   &    \obj{m-əənɨkɨ-} &              \obj{m-epɨ-} &  \obj{m-entapo-} &   \obj{mɨ-ka-} \\
-\gl{1+2} &  \obj{kɨt-əənɨkɨ-} &             \obj{ke-epɨ-} &  \obj{k-entapo-} &  \obj{kɨh-ka-} \\
-\gl{3}   &    \obj{n-əənɨkɨ-} &              \obj{n-epɨ-} &  \obj{n-entapo-} &    \obj{n-ka-} \\
+\gl{1}   &    \obj{t-əənɨkɨ-} &    \obj{t-əene-} &              \obj{s-epɨ-} &  \obj{s-entapo-} &  \obj{wɨ-tən-} \\
+\gl{2}   &    \obj{m-əənɨkɨ-} &    \obj{m-əene-} &              \obj{m-epɨ-} &  \obj{m-entapo-} &  \obj{mɨ-tən-} \\
+\gl{1+2} &  \obj{kɨt-əənɨkɨ-} &    \obj{k-əene-} &             \obj{ke-epɨ-} &  \obj{k-entapo-} &  \obj{kɨ-tən-} \\
+\gl{3}   &    \obj{n-əənɨkɨ-} &    \obj{n-əene-} &              \obj{n-epɨ-} &  \obj{n-entapo-} &  \obj{nɨ-tən-} \\
 \bottomrule

### 2021-11-12 documents/sections/extensions.tex
discussion and factor comparison
many changes and additions to introduction

@@
 \textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work on \PPek morphosyntax can be found in the literature.
-All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.\footnote{In \cref{tab:pekreg} and the remainder of this paper, we omit any possible TAM suffixes from the attested forms in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.}
+All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.
 We can thus reconstruct a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
@@
 \xe
-%
-%\ex<ikp-182>\ikpeng \parencite[][320]{chagas2019ikpeng}\\
-%\begingl
-%\gla ∅-iptoŋ-lan//
-%\glb \gl{3}\gl{s_a_}-sit.down-\gl{pst}//
-%\glft \qu{She sat down.}//
-%\endgl
-%\xe
-%
-
-%\ex<ikpwloss> %\arara-\ikpeng \obj{w}:Ø correspondences\\
-%\begin{tabular}[t]{@{}lll@{}}
-%\PXin & \arara & \ikpeng\\
-%\rc{wɨna} \qu{\gl{dat}} & \obj{wɨna} & \obj{ɨna}\\
-%\rc{wokori} \qu{dog, jaguar} & \obj{wokori} & \obj{akari}\\
-%\rc{tawe} \qu{capuchin monkey} & \obj{tawe} & \obj{tae}\\
-%\rc{watke} \qu{to defecate} & \obj{watke} & \obj{atke}\\
-%\rc{wɨnkɨ} \qu{to sleep} & \obj{wɨnkɨ} & \obj{ɨnkɨ}\\
-%& (\obj{o-wɨnkɨ} \qu{sleep!}) & (\obj{j-ɨnkɨ-lɨ} \qu{s/he slept})\\
-%\end{tabular}\\
-%\parencites[80]{alves2017arara}[164, 67, 150]{ikpengpacheco2001}[9]{desouza2010arara}[68]{souza1993arara}[91]{alves2013verbo}[40]{campetela1997analise} 
-%\xe
 
@@
 
-%\ex<tri-130> \trio \parencite[][114]{triomeira1999}\\
-%\begingl
-%\gla n-epɨ//
-%\glb \gl{3}-come//
-%\glft \qu{S/he came.}//
-%\endgl
-%\xe
-
 %Apart from these \PTir verbs with \rc{w(ɨ)-}, there are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
@@
 
-%\pex<car1p>\carijo \parencite[][59, 53]{guerrero2016carijo}
-%\begin{multicols}{2}
-%\a<car-1>
-%\begingl
-%\gla jɨ-hɨnəh-ɨ//
-%\glb \gl{3}>\gl{1}-kill-\gl{pfv}//
-%\glft \qu{S/he killed me.}//
-%\endgl
-%\a<car-31>
-%\begingl
-%\gla j-owo-rɨ//
-%\glb \gl{1}\gl{poss}-uncle-\gl{pert}//
-%\glft \qu{my uncle}//
-%\endgl
-%\end{multicols}
-%\xe
-%
-
 As noted in \cref{sec:taranoan}, this extension also erased any traces of a potential \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
@@
 \xe
-
-%
-%Interestingly enough, the verb \qu{to be} only shows \obj{w-} in its \obj{a} root allomorph, but not in the \obj{et͡ʃi} root allomorph, where it has \obj{j-} \exref{car1y.mivida-12}.
-%In a somewhat foreshadowing move, I will note that \PTar \rc{(ət)epɨ} \qu{to come} was also affected by the \carijo extension, as seen in \exref{car1y.car-18}.
-%As discussed in \cref{sec:taranoan}, this verb preserves \rc{w-} in \akuriyo and \trio \exref{tri-105}.
-%
-%\pex<car1y>\carijo
-%\a<mivida-12>
-%\begingl
-%\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
-%\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
-%\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
-%\glft \qu{Then I was already grown up.} \parencite[][177]{robayo1989rame}//
-%\endgl
-%\a<car-18>
-%\begingl
-%\gla əji-wa-e j-eh-ɨ//
-%\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
-%\glft \qu{I came looking for you.} \parencite[][102]{guerrero2019carijo}//
-%\endgl
-%\xe
-%
-%\ex<tri-105>\trio \parencite[][337]{triocarlin2004}\\
-%\begingl
-%\gla ene w-əe//
-%\glb see.\gl{sup} \gl{1}\gl{s_a_}-come//
-%\glft \qu{I came to see him/her.}//
-%\endgl
-%\xe
 

### 2021-11-12 documents/sections/introduction.tex
updated readme
tables for introduction
many changes and additions to introduction

@@
 
-\input{stuff/cariban_tree}
+\subsection{The Cariban language family}
+
+%todo write something here about the family in general
+%maybe a nice little map-cum-tree, eh?
+%\input{floats/cariban_tree}
+
+\subsection{Verbs with conservative first person forms}
+
+\input{floats/hixintro}
+\input{floats/triintro}
+Cariban languages feature relatively rich verbal morphology, both pre- and suffixes, inflecting for person, number, tense, aspect, and evidentiality, combined with a range of valency-modifying derivational markers.
+Person paradigms for four \hixka \gl{s_a_} verbs are given in \cref{tab:hixintro}.%
+\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
+This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms we show.}%
+\footnote{In \cref{tab:hixintro} and the remainder of this paper, we omit any TAM suffixes found in the original forms found in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.}%
+\footnote{
+We use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in an interlinearized glossed example.
+We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
+It is evident that the verb \qu{to be} diverges from otherwise identically inflected verbs by having a different first person marker.
+That prefix is \obj{w-}, while other \gl{s_a_} verbs like \qu{to fall}, have \obj{k(ɨ)-}.
+Another similar example comes from \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
+From a synchronic, language-internal perspective, one could call these prefixes and the verbs on which they appear \textsc{irregular}, contrasting with regular prefixes on regular \gl{s_a_} verbs, \obj{k(ɨ)-} and \obj{t-}/\obj{s-}.
+However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many of the more strict ones \parencite[][e.g.]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
+However, both languages have a few other verbs patterning the same way; for example, the first-person form of \trio \qu{to be} is \obj{w-ei-}.
+Whether or not they fit one's understanding of irregular, these verbs can be analyzed as belonging to a small inflection (sub-)class. %
+%\footnote{That analysis is found in synchronic descriptions of other Cariban languages discussed in more detail later, see \textcites[149]{alves2017arara}[4]{meira2003bakairi}.}
+To answer the question of how these inflectional patterns came into being, one needs to turn to diachrony, starting with the person marking system in \PC.
+
+%
+%\begin{inlinelist}
+%	\item special copula prefix \parencite[188]{hixkaryanaderby1985}
+%	\item inflectional subclass \parencites[149]{alves2017arara}[4]{meira2003bakairi}
+%	\item \dbqu{exceptional cases} \parencite[293]{triomeira1999}
+%	\item phonologically conditioned allomorphs \parencites[139]{meira2006syntactic}{meira2003primeras}[189]{hixkaryanaderby1985}
+%	\item not discussed \parencites{waiwaihawkins1998}{ikpengpacheco2001}{alves2013verbo}{pacheco2003intransitivos}
+%\end{inlinelist}.s
+
+%meira: non-detransitivized verbs
+%alves: non-detransitivized verbs (as a consequence, o-initial)
+
+%\subsection{Cariban verbal person marking and marker extensions}
+%\label{sec:background}
+%The archaic first person markers investigated in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system that did not affect the whole lexicon.
+%To provide readers with the necessary background, we first introduce said ancestral system \pcref{sec:pc_person}, then the peculiarities of the Cariban split-\gl{s} system \pcref{sec:split}, and finally the concept of person marker extensions \pcref{sec:extensions_intro}.
+
+\subsubsection{\PC verbal person marking}
+\label{sec:pc_person}
+\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.%
+The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.\footnote{Glossing abbreviations: }%\glossingAbbrevsComma.}
+The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
+There was a single aliophoric marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
+Local scenarios were expressed in a non-transparent manner, by using the \gl{1+2} prefix \rc{k-} in both cases.%
+
+\input{floats/pcpers}
+
+Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
+That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}), as well as the absence of \rc{i} after all \gl{s_a_} prefixes.
+On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones.
+The third person marker in \gl{s_p_} verbs was identical to the one in \gl{3}>\gl{3} scenarios (\rc{n(i)-)}.
+
+From this diachronic perspective, it becomes evident that both \hixka \obj{w-} and \trio \obj{wɨ-} are reflexes of the \PC \gl{1}\gl{s_a_} prefix \rc{w-}.
+Notably, it is reconstructed as the only \gl{1}\gl{s_a_} prefix, a situation still found in other modern languages like \maqui \pcref{tab:makintro}.
+This means that \hixka \qu{to be} and \trio \qu{to go} are divergent because they preserve the old prefix, while other verbs have taken on new prefixes.
+From a language-specific point of view, these verbs are thus \textsc{conservative}, contrasting with innovative verbs.
+The prefixes may be considered \textsc{relics}, old and restricted to specific lexically conditioned contexts, contrasting with the innovative prefixes found elsewhere.
+One can draw a parallel to the few English nouns like \obj{ox}, which preserve the old plural suffix \obj{-en}.
+It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{Ochse-n}, \obj{Mensch-en}, \obj{Hase-n}.
+Having established the conservative nature of these prefixes, the obvious question is where the new prefixes came from.
+
+\input{floats/makintro}
+
+\subsubsection{Person marker extensions in intransitive verbs}
+\label{sec:extensions_intro}
+The original \PC split-\gl{s} system has been subject to change in many languages, in some even to the point of total loss.
+Besides modifications like losing the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} or losing certain person values altogether, these changes have largely been due to person prefixes being extended to new verbs.%
+\footnote{
+In principle, one could also imagine a scenario in which verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
+However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
+Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
+For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
+There have been many such person marker extensions in Cariban languages, and some are still ongoing.
+This is shown by \textcite{gildea1998}, using the three Parukotoan languages as an example.
+We have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription and the added \kaxui alternative \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
+Apart from segmental changes to individual morphemes, the following restructuring innovations happened in the \setone paradigm in Parukotoan:
+
+\begin{enumerate}
+	\item \PPar \begin{enumerate}
+		\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
+		\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+	\end{enumerate}
+	\item \PWai \begin{enumerate}
+		\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
+		\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
+	\end{enumerate}
+	\item \waiwai \begin{enumerate}
+		\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
+\end{enumerate}
+
+\begin{figure}[hbt]
+	\centering
+	\setlength{\tabcolsep}{2pt}
+	\begin{tikzpicture}[
+		every node/.append style={align=center},
+		level distance=80pt,
+		]
+		\Tree[.{\PC{}\pcone}
+		[.{\PPar{}\ppone}
+		{\kaxui{}\kaxone}
+		[.{\PWai{}\pwone}
+		{\hixka{}\hixone}
+		{\waiwai{}\waione}
+		]
+		]
+		]
+	\end{tikzpicture}
+	\caption{Person marking extensions in Parukotoan, after \textcite[94]{gildea1998}}
+	\label{fig:par_ext}
+\end{figure}
+
+\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows variation in the first person and \gl{1+2} prefixes.
+\waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
+In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
+That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
+Similarly, the \gl{2}\gl{s_p_} prefix \obj{a-} was extended to imperatives of (former) \gl{s_a_} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.
+
+\pex<wailoss>
+\begin{multicols}{2}
+	\a<wai-158> \waiwai \parencite[][98]{waiwaihawkins1998}\\
+	\begingl
+	\gla k-eɸɨrka-t͡ʃhe//
+	\glb \gl{1+2}-fall-\gl{advz}.after//
+	\glft \qu{after our fall}//
+	\endgl
+	\a<kax-138> \kaxui \parencite[][49]{schuring2018kaxuyana}\\
+	\begingl
+	\gla ku-w-ehurka-tpɨrɨ//
+	\glb \gl{1+2}-\gl{s_a_}-fall-\gl{nmlz}.\gl{pst}.\gl{pert}//
+	\glft \qu{our fall}//
+	\endgl
+	\a<wai-112> \waiwai \parencite[][177]{waiwaihawkins1998}\\
+	\begingl
+	\gla a-mo-ko//
+	\glb \gl{2}-come-\gl{imp}//
+	\glft \qu{come!}//
+	\endgl
+	\a<hix-43> \hixka \parencite[][191]{hixkaryanaderby1985}\\
+	\begingl
+	\gla m-omokɨ-no//
+	\glb \gl{2}\gl{s_a_}-come-\gl{imm}//
+	\glft \qu{You have come.}//
+	\endgl
+\end{multicols}
+\xe
+
+%Such extensions, often only affecting one person value at a time, and the eventual loss of split-\gl{s} are not entirely unexpected, given that the system lacks any semantic basis.
+
+While different cases of loss of split-\gl{s} by person marking extensions are discussed by \textcite[91--96]{gildea1998}, this paper focuses on a so far neglected aspect of these extensions.
+As a prerequisite for our story, we argue that they are take place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
+Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
+In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which is what one would expect from a change in progress.
+%todo maybe put in GOOD kaxui data
+Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
+Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
+Our final argument in favor of the lexical diffusion scenario brings us back to the \hixka and \trio verbs in \cref{tab:triintro,tab:hixintro}:
+We know from synchronic \hixka that some \gl{s_a_} verbs take \obj{w-} rather than \obj{k-}.
+We interpret this as the extension of \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_} which began in \PWai stopping short of these verbs, and not spreading through the entire verbal lexicon.
+
+We investigated 19 person marker extensions affecting intransitive verbs, and identified 6 which left a group of verbs untouched, like \PWai \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}.
+These have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages, marked with bold print in \cref{fig:caribantree}.
+Interestingly, these extensions always feature innovative first person markers on \gl{s_a_} verbs.
+While extensions occurred with other person values as well, they never affected \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affected the entire lexicon, at least based on the attested data.
+Illustrative examples for extensions affecting the entire lexicon are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set in \waimiri \pcref{tab:waimiri}.%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
+
+The different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to incomplete extensions points to the split-\gl{s} system playing some role.
+Thus, we will introduce that system before discussing the incomplete extensions in more detail.
+This will also give readers an idea how it is possible for the \gl{s_a_}--\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
+
+\input{floats/completed}
+
+\subsubsection{The \PC split-\gl{s} system}
+\label{sec:split}
+As seen in \cref{sec:pc_person}, the \PC distinction between \gl{s_a_} and \gl{s_p_} verbs was made by two inflectional classes for intransitive verbs within the \setone system, but this was not the only morphological criterion:
+%\begin{inlinelist}
+%	\item presence vs absence of the \gl{s_a_} marker \rc{w-}
+%	\item absence vs presence of the second person prefix \rc{ə(j)-} in imperatives
+%	\item presence vs absence of a derivational detransitivizing prefix
+%\end{inlinelist}.
+%
+Many languages show an \gl{s_a_} class marker in deverbalized forms, which can be reconstructed to \PC as \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
+With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
+Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
+%
+\input{floats/participles}%
+%
+\input{floats/nominalizations}%
+%
+The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives.
+Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed (both suffixed with \rc{-kə}).
+This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
+Both the \gl{s_a_} marker \rc{w-} in participles and nominalizations and the prefixes in imperativees have been lost in some languages, but are reconstructible to \PC.
+
+\input{floats/imperatives}
+
+%\subsubsection{The detransitivizing prefix}
+There is one further property uniting most \gl{s_a_} verbs, which is not based on inflectional morphology.
+As mentioned in \cref{sec:pc_person}, mismatches between the semantics of intransitive verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common.
+However, the Cariban split-\gl{s} system goes further than all other known such systems, in that its division of the verbal lexicon does not follow any discernible semantic criteria whatsoever.
+\textcite{meira2000split} takes a sizable corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
+He shows that neither (non\-)activities, % \parencite[210]{meira2000split}, 
+(non-)agency, % \citeyearpar[212]{meira2000split}, 
+(in-)animacy, % \citeyearpar[213]{meira2000split},
+nor Aktionsart % \citeyearpar[215]{meira2000split} 
+satisfactorily predict the class membership of intransitive verbs.
+
+Rather, the reason for a verb to take the \gl{a}- or \gl{p}-oriented prefix is (at least diachronically) a morphological one.
+\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
+\begin{quotebox}{\parencite[201]{meira2000split}}
+	Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still exisiting transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
+\end{quotebox}
+\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
+\begin{inlinelist}
+	\item there are a few underived \gl{s_a_} verbs, with no detransitivizing prefix
+	\item \gl{s_a_} verbs can develop irregular semantics compared to their transitive counterparts
+	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{s_a_} verb
+	\item some originally derived \gl{s_a_} verbs have lost their transitive counterparts
+	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{s_a_})} from \qu{to dance with (\gl{tr})}
+\end{inlinelist}.
+He also notes that this leads to an inflectional split not based in meaning, but rather morphology:
+
+\begin{quotebox}{\parencite[226]{meira2000split}}
+	Apparently, the morphological behavior of the \gl{s_a_} verb class is an accidental consequence of the fact that detransitivization, as far back as we can reconstruct, entails all the morphology described […] as typical of \gl{s_a_} verbs.
+The alignment of person-marking prefixes appears not to be driven by any semantic forces in the language; it is as though they were being dragged by the evolution of the reflexive marker.
+\end{quotebox}
+
+As for the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although they have since merged into a single morpheme, apparently in all languages.
+Modern reflexes of \detrz show a range of meanings, which can all be characterized as \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
+
+\ex<triodetrz> \trio \parencites[218--219]{meira2000split}[128, 256]{triomeira1999}\\
+\begin{tabular}[t]{@{}llll@{}}
+	\\
+	\obj{nonta}  & → & \obj{e-nonta}, & \qu{abandon each other}\\
+	\qu{abandon} & & \obj{əi-nonta} &  (reciprocal) \\
+	\\
+	\obj{suka} & → & \obj{e-suka}, & \qu{wash self}\\
+	\qu{wash} & & \obj{əi-suka} & (reflexive)\\
+	\\
+	\obj{pahka} & → & \obj{e-pahka} & \qu{break (\gl{intr})}\\
+	\qu{break (\gl{tr})} & & & (anticausative)\\
+	\\
+	\obj{puunəpɨ} & → & \obj{əh-puunəpɨ}, & \qu{think, meditate}\\
+	\qu{think about} & & \obj{əi-puunəpɨ} & (antipassive)\\
+\end{tabular}
+\xe
+%
+The morphological variation featured in \qu{to abandon each other} and \qu{to wash self} is due to the mentioned collapse between the two \PC prefixes:
+\obj{e-} is a reflex of the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
+However, both can occur with either meaning -- at least for these two verbs.
+
+As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
+\textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
+\textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
+Many of these verbs will be discussed in \cref{sec:verbs}.
+
+\input{floats/underived2007.tex}
+
+As in many other split-\gl{s} systems found in languages of the world, the intransitive verbal lexicon was divided into two classes, with verbs inherently being members of the \gl{s_a_} or the \gl{s_p_} group.
+That is, while in transitive verbs, the choice of person marker had a crucial semantic contribution, it was predictable in intransitive verbs.
+This is illustrated with modern \kalina data in \exref[kar1]{kar2}.
+
+\pex<kar1> \kalina
+\a<kar-60>
+\begingl
+\gla mi-kupi-ja//
+\glb \gl{2}>\gl{3}-bathe-\gl{prs}//
+\glft \qu{You bathe him/her.} \parencite[][160]{hoff1968carib}//
+\endgl
+\a<kar-62>
+\begingl
+\gla a-kupi-ja//
+\glb \gl{3}>\gl{2}-bathe-\gl{prs}//
+\glft \qu{S/he bathes you.} \parencite[][63]{yamada2011evidentiality}//
+\endgl
+\xe
+%
+In \exref{kar1}, the choice between the second person \gl{a}- and \gl{p}-oriented markers \obj{mi-} and \obj{a-} depends on the scenario:
+The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
+While intransitive verbs show the same (or very similar) person markers, they contribute no semantic difference here \exref{kar2}.
+
+\pex<kar2> \kalina
+\a<kar-59>
+\begingl
+\gla sipi tɨnka-rɨ m-ekema-non hen//
+\glb net pull-\gl{nmlz} \gl{2}-be.afraid-\gl{prs}.\gl{uncert} eh?//
+\glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
+\endgl
+\a<kar-61>
+\begingl
+\gla aj-awoi-ja//
+\glb \gl{2}-get.up-\gl{prs}//
+\glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
+\endgl
+\xe
+%
+Rather, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.\footnote{The root \obj{awomɨ} \qu{to get up} is subject to syllable reduction and assimilation to the prefix-initial \obj{j}.}
+What is more, the choice of prefix does not necessarily convey information about the semantics of the verb, since there are clear mismatches:
+\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker must be considered volitional.
+These mismatches are not isolated cases, as will be discussed in \cref{sec:split}.
+
+This paper is dedicated to these conservatively inflected verbs, their origins, their similarities, and the reasons behind them.
+It is structured as follows: \cref{sec:background} provides readers with the necessary background on Cariban verbal person marking, especially split-\gl{s} system, and introduces the mechanism by which the innovative prefixes spread to almost all \gl{s_a_} verbs.
+The topic of \cref{sec:extensions} are the six individual innovations which introduced new first person prefixes, but left a small group of verbs unaffected.
+Since these verbs show considerable etymological overlap between languages, they are reconstructed and discussed in more detail in \cref{sec:verbs}.
+Finally, we search for factors motivating the resistance of these verbs, discuss the findings and put them in a general context of language change and morphology in \cref{sec:discussion}.

### 2021-11-12 workflow/process_data.py

@@ -267,14 +267,14 @@ pyd.x = ["Meaning_ID"]
 pyd.y = ["Inflection"]
 pyd.y_sort = person
 for lg, meanings in {
-    "hix": ["fall", "be_afraid", "walk", "go"],
-    "tri": ["sleep", "bathe_intr", "yawn", "say"],
-    "mak": ["eat", "arrive", "go"]
+    "hix": ["fall", "be_afraid", "walk", "cut_self", "be"],
+    "tri": ["sleep", "see_self", "bathe_intr", "yawn", "go"],
+    "mak": ["eat", "arrive", "go", "be"]
     # "yuk": ["wash_self", "sleep", "fall"],
 }.items():
     pyd.filters = {"Language_ID": [lg], "Meaning_ID": meanings, "Inflection": person}
     pyd.x_sort = meanings
-    tabular = pyd.compose_paradigm(i_df, multi_index=False)
+    tabular = pyd.compose_paradigm(i_df[i_df["Verb_Cognateset_ID"] != "be_1"], multi_index=False)
     label = lg + "intro"
     tabular_raw = tabular.rename(columns=mean_dic)
     tabular_raw.columns = [repl_latex(f"'{col}'") for col in tabular_raw.columns]

### 2021-11-15 documents/sections/extensions.tex
extensive restructuring of introduction

@@
-\section{The innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
+\section{Incomplete extensions: the innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
 \label{sec:extensions}
-As stated in \cref{sec:extensions_intro}, the person marker extensions which did not spread through the entire lexicon all have in common that they feature innovative first person markers on verbs that are (at least historically) members of the \gl{s_a_} class.
-There are six attested such innovations, three of which can be reconstructed to intermediate proto-languages.
-The new \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
+As stated in \cref{sec:extensions_intro}, the person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically) members of the \gl{s_a_} class.
+Of the six attested incomplete extensions, three can be reconstructed to intermediate proto-languages, while three others happened in earlier stages of single modern languages.
+The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
 \cref{sec:pekodian} investigates the innovation of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.

### 2021-11-15 documents/sections/introduction.tex
extensive restructuring of introduction

--- 2021-09-10 documents/sections/introduction.tex
@@
 
-\input{stuff/cariban_tree}
+\subsection{The Cariban language family}
+
+%todo write something here about the family in general
+%maybe a nice little map-cum-tree, eh?
+%\input{floats/cariban_tree}
+
+\subsection{Verbs with conservative first person forms}
+
+\input{floats/hixintro}
+\input{floats/triintro}
+Cariban languages feature relatively rich verbal morphology, both pre- and suffixes, inflecting for person, number, tense, aspect, and evidentiality, combined with a range of valency-modifying derivational markers.
+A split-\gl{s} system can be reconstructed to \PC, discussed in more detail in \cref{sec:split}.
+\cref{tab:hixintro} shows person paradigms for four \hixka verbs, members of the \gl{s_a_} inflectional class.%
+\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
+This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms we show.}%
+\footnote{In \cref{tab:hixintro} and the remainder of this paper, we omit any TAM suffixes found in the original forms found in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.}%
+\footnote{
+We use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in an interlinearized glossed example.
+We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
+\footnote{Glossing abbreviations: }%\glossingAbbrevsComma.}
+A comparison of these paradigms shows that the verb \qu{to be} diverges from the others in its first person marker.
+That prefix is \obj{w-}, while other \gl{s_a_} verbs like \qu{to fall} have \obj{k(ɨ)-}.
+A similar example can be found in \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
+In both languages, the verbs on the left of the table are representative for the \gl{s_a_} class, since the vast majority is inflected identically.
+In both languages, there are only a few other verbs inflected identically to the divergent ones on the right; for example, the first-person form of \trio \qu{to be} is \obj{w-ei-} \parencites[339]{triomeira1999}.
+
+In our synchronic analyses of these two languages, we consider these verbs and their first person prefixes \textsc{irregular}, contrasting with regular prefixes on regular \gl{s_a_} verbs, \obj{k(ɨ)-} and \obj{t-}/\obj{s-}.
+However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[][e.g.]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
+For such approaches, these verbs simply belong to a small inflectional (sub-)class. %
+%\footnote{That analysis is found in synchronic descriptions of other Cariban languages discussed in more detail later, see \textcites[149]{alves2017arara}[4]{meira2003bakairi}.}
+Regardless of synchronic analysis, the explanation for these inflectional patterns lies, as so often, in diachrony.
+Our story starts with the verbal person marking system of \PC, discussed in \cref{sec:pc_person}.
+That system was subject to different kinds of innovations; the one responsible for the irregular \hixka and \trio verbs is introduced in \cref{sec:extensions_intro}.
+A particular component of that system, the distinction between \gl{s_a_} and \gl{s_p_} verbs, played a major role in the developments under discussion and is the topic of \cref{sec:split}.
+
+%
+%\begin{inlinelist}
+%	\item special copula prefix \parencite[188]{hixkaryanaderby1985}
+%	\item inflectional subclass \parencites[149]{alves2017arara}[4]{meira2003bakairi}
+%	\item \dbqu{exceptional cases} \parencite[293]{triomeira1999}
+%	\item phonologically conditioned allomorphs \parencites[139]{meira2006syntactic}{meira2003primeras}[189]{hixkaryanaderby1985}
+%	\item not discussed \parencites{waiwaihawkins1998}{ikpengpacheco2001}{alves2013verbo}{pacheco2003intransitivos}
+%\end{inlinelist}.s
+
+%meira: non-detransitivized verbs
+%alves: non-detransitivized verbs (as a consequence, o-initial)
+
+%\subsection{Cariban verbal person marking and marker extensions}
+%\label{sec:background}
+%The archaic first person markers investigated in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system that did not affect the whole lexicon.
+%To provide readers with the necessary background, we first introduce said ancestral system \pcref{sec:pc_person}, then the peculiarities of the Cariban split-\gl{s} system \pcref{sec:split}, and finally the concept of person marker extensions \pcref{sec:extensions_intro}.
+
+\subsubsection{\PC verbal person marking}
+\label{sec:pc_person}
+\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.
+The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
+The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
+There was a single aliophoric marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
+Local scenarios were expressed in a non-transparent manner, both using the \gl{1+2} prefix \rc{k-}.
+
+\input{floats/pcpers}
+
+Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
+That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}), as well as the absence of \rc{i} after all \gl{s_a_} prefixes.
+On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones.
+The third person marker in \gl{s_p_} verbs was identical to the one in \gl{3}>\gl{3} scenarios (\rc{n(i)-)}.
+
+Equipped with knowledge about the ancestral system, it becomes clear that the irregularly inflected \hixka and \trio  preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-} and are therefore \textsc{conservative}.
+They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
+The reflexes of \rc{w-} may be considered \textsc{relics}, old and restricted to specific lexically conditioned contexts, contrasting with the innovative prefixes found elsewhere.
+The verbs and their prefixes are comparable with the few English nouns like \obj{ox}, which preserve the old plural suffix \obj{-en}.
+It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{Ochse-n} \qu{ox-en}, \obj{Name-n} \qu{name-s}, \obj{Hase-n} \qu{hare-s}, \obj{Bär-en} \qu{bear-s}.
+Since the irregular \hixka and \trio prefixes are conservative and the regular prefixes are innovative, the next question to be addressed is where these new prefixes came from.
+
+%\input{floats/makintro}
+
+\subsubsection{Person marker extensions in intransitive verbs}
+\label{sec:extensions_intro}
+In his discussion of the \PC split-\gl{s} system and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different modifications in various languages.
+The main mechanism of change leading to these modifications are \textbf{person marker extensions}, i.e.
+the use of verbal person prefixes being extended to contexts previously occupied by other prefixes.
+%\footnote{
+%In principle, one could also imagine a scenario in which verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
+%However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
+%Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
+%For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
+There have been many different person marker extensions in Cariban languages, and some are still ongoing.
+This is illustrated by \textcite{gildea1998}, using the three Parukotoan languages as an example.
+Apart from segmental changes to individual morphemes, the following innovations happened in the \setone paradigm in Parukotoan:
+
+\begin{enumerate}
+	\item \PPar \begin{enumerate}
+		\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
+		\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+	\end{enumerate}
+	\item \PWai \begin{enumerate}
+		\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
+		\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
+	\end{enumerate}
+	\item \waiwai \begin{enumerate}
+		\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
+\end{enumerate}
+%
+All innovations are person marker extensions except 2b, which combined a pronoun with the linker \rc{j-}.
+They are printed in bold in \cref{fig:par_ext}, which reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and the additional alternative \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
+%
+\begin{figure}[hbt]
+	\centering
+	\setlength{\tabcolsep}{2pt}
+	\fbox{\begin{tikzpicture}[
+		every node/.append style={align=center},
+		level distance=80pt,
+		]
+		\Tree[.{\PC{}\pcone}
+		[.{\PPar{}\ppone}
+		{\kaxui{}\kaxone}
+		[.{\PWai{}\pwone}
+		{\hixka{}\hixone}
+		{\waiwai{}\waione}
+		]
+		]
+		]
+	\end{tikzpicture}}
+	\caption{Person marking extensions in Parukotoan, after \textcite[94]{gildea1998}}
+	\label{fig:par_ext}
+\end{figure}
+%
+\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows the variation in the first person and \gl{1+2} prefixes that is reconstructible to \PPar.
+\waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
+%In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
+%That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
+%Similarly, the \gl{2}\gl{s_p_} prefix \obj{a-} was extended to imperatives of (former) \gl{s_a_} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.
+
+%Such extensions, often only affecting one person value at a time, and the eventual loss of split-\gl{s} are not entirely unexpected, given that the system lacks any semantic basis.
+
+\textcite{gildea1998} discusses person marker extensions only in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment; for our story we will zoom in on a so far neglected aspect of these extensions.
+To begin, we argue that they took place via lexical diffusion, a type of extension \parencite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
+Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
+In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which is what one would expect from a change in progress.
+%todo maybe put in GOOD kaxui data
+Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} convincingly argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
+Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
+
+Our third argument in favor of the lexical diffusion scenario brings us back to the \hixka and \trio verbs with conservative first person forms:
+In both cases, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
+We interpret the continued presence of the old \gl{1}\gl{s_a_} marker in a few verbs as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
+In our investigations of person marker extensions, we identified 19 individual cases affecting intransitive verbs, and found 6 of them to be incomplete.
+These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages. %, marked with bold print in \cref{fig:caribantree}.
+Interestingly, all six extensions featured innovative first person markers on \gl{s_a_} verbs.
+While extensions occurred with other person values as well, they never affected \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affected all potential targets.
+Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
+
+The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affeting them points to the split-\gl{s} system playing some role, a suspicion which will be confirmed in \cref{sec:discussion}.
+Thus, we will discuss that system in a little more detail before moving on to the incomplete extensions.
+This will also give readers an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
+
+\input{floats/completed}
+
+\subsubsection{The morphological basis of the split-\gl{s} system}
+\label{sec:split}
+As seen in \cref{sec:pc_person}, the \PC distinction between \gl{s_a_} and \gl{s_p_} verbs was implemented by two inflection classes within the \setone prerson paradigm, but this was not the only morphological criterion:
+%\begin{inlinelist}
+%	\item presence vs absence of the \gl{s_a_} marker \rc{w-}
+%	\item absence vs presence of the second person prefix \rc{ə(j)-} in imperatives
+%	\item presence vs absence of a derivational detransitivizing prefix
+%\end{inlinelist}.
+%
+Many languages show an \gl{s_a_} class marker in deverbalized forms, which can be reconstructed to \PC as \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
+With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
+Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
+%
+\input{floats/participles}%
+%
+\input{floats/nominalizations}%
+%
+The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives.
+Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed (both suffixed with \rc{-kə}).
+This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
+Both the \gl{s_a_} marker \rc{w-} in participles and nominalizations and the prefixes in imperativees have been lost in some languages, but are reconstructible to \PC.
+However, there was one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
+
+\input{floats/imperatives}
+
+%\subsubsection{The detransitivizing prefix}
+Mismatches between the semantics of intransitive verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with modern \kalina data in \exref[kar1]{kar2}.
+
+\pex<kar1> \kalina
+\a<kar-60>
+\begingl
+\gla mi-kupi-ja//
+\glb \gl{2}>\gl{3}-bathe-\gl{prs}//
+\glft \qu{You bathe him/her.} \parencite[][160]{hoff1968carib}//
+\endgl
+\a<kar-62>
+\begingl
+\gla a-kupi-ja//
+\glb \gl{3}>\gl{2}-bathe-\gl{prs}//
+\glft \qu{S/he bathes you.} \parencite[][63]{yamada2011evidentiality}//
+\endgl
+\xe
+%
+In \exref{kar1}, the choice between the second person \gl{a}- and \gl{p}-oriented markers \obj{mi-} and \obj{a-} depends on the scenario:
+The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
+The intransitive verbs in \exref{kar2} show the same person markers, but there is no choice what marker a specific verb uses.
+
+\pex<kar2> \kalina
+\a<kar-59>
+\begingl
+\gla sipi tɨnka-rɨ m-ekema-non hen//
+\glb net pull-\gl{nmlz} \gl{2}-be.afraid-\gl{prs}.\gl{uncert} eh?//
+\glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
+\endgl
+\a<kar-61>
+\begingl
+\glpreamble aya:woiya //
+\gla aj-awomɨ-ja//
+\glb \gl{2}-get.up-\gl{prs}//
+\glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
+\endgl
+\xe
+%
+Rather, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+It seems that in both cases, the prefix does not contribute to the semantics of the predicate, since there are clear mismatches:
+\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker must be considered volitional.
+\textcite{meira2000split} takes a sizable corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
+He shows that neither (non\-)activities, % \parencite[210]{meira2000split}, 
+(non-)agency, % \citeyearpar[212]{meira2000split}, 
+(in-)animacy, % \citeyearpar[213]{meira2000split},
+nor Aktionsart % \citeyearpar[215]{meira2000split} 
+satisfactorily predict the class membership of intransitive verbs.
+
+Rather, the reason for a verb to take the \gl{a}- or \gl{p}-oriented prefix is (at least diachronically) a morphological one.
+\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
+\begin{quotebox}{\parencite[201]{meira2000split}}
+	Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still exisiting transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
+\end{quotebox}
+\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
+\begin{inlinelist}
+	\item there are a few underived \gl{s_a_} verbs, with no detransitivizing prefix
+	\item \gl{s_a_} verbs can develop irregular semantics compared to their transitive counterparts
+	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{s_a_} verb
+	\item some originally derived \gl{s_a_} verbs have lost their transitive counterparts
+	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{s_a_})} from \qu{to dance with (\gl{tr})}
+\end{inlinelist}.
+He also notes that this leads to an inflectional split not based in meaning, but rather morphology:
+
+\begin{quotebox}{\parencite[226]{meira2000split}}
+	Apparently, the morphological behavior of the \gl{s_a_} verb class is an accidental consequence of the fact that detransitivization, as far back as we can reconstruct, entails all the morphology described […] as typical of \gl{s_a_} verbs.
+The alignment of person-marking prefixes appears not to be driven by any semantic forces in the language; it is as though they were being dragged by the evolution of the reflexive marker.
+\end{quotebox}
+
+As for the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although they have since merged into a single morpheme, apparently in all languages.
+Modern reflexes of \detrz show a range of meanings, which can all be characterized as \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
+
+\ex<triodetrz> \trio \parencites[218--219]{meira2000split}[128, 256]{triomeira1999}\\
+\begin{tabular}[t]{@{}llll@{}}
+	\\
+	\obj{nonta}  & → & \obj{e-nonta}, & \qu{abandon each other}\\
+	\qu{abandon} & & \obj{əi-nonta} &  (reciprocal) \\
+	\\
+	\obj{suka} & → & \obj{e-suka}, & \qu{wash self}\\
+	\qu{wash} & & \obj{əi-suka} & (reflexive)\\
+	\\
+	\obj{pahka} & → & \obj{e-pahka} & \qu{break (\gl{intr})}\\
+	\qu{break (\gl{tr})} & & & (anticausative)\\
+	\\
+	\obj{puunəpɨ} & → & \obj{əh-puunəpɨ}, & \qu{think, meditate}\\
+	\qu{think about} & & \obj{əi-puunəpɨ} & (antipassive)\\
+\end{tabular}
+\xe
+%
+The morphological variation featured in \qu{to abandon each other} and \qu{to wash self} is due to the mentioned collapse between the two \PC prefixes:
+\obj{e-} is a reflex of the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
+However, both can occur with either meaning -- at least for these two verbs.
+%As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
+%\textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
+%\textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
+%Many of these verbs will be discussed in \cref{sec:verbs}.
+
+%\input{floats/underived2007.tex}
+
+\subsubsection{Summary and outlook}
+We identified \gl{s_a_} verbs that irregularly inflect for first person in 9 Cariban languages.
+These irregular forms are actually conservative, unaffected by person marker extensions spread via lexical diffusion.
+Such conservative forms are only found among the first person forms of  (etymological) \gl{s_a_} verbs.
+Our primary goal is to establish what verbs remained unaffected by the individual extensions and to search for factors explaining the patterns, proceeding as follows:
+The topic of \cref{sec:extensions} are the six incomplete innovations, the innovative markers they introduced, and the verbs they left untouched.
+Since these verbs show considerable etymological overlap between languages, they are reconstructed and discussed in more detail in \cref{sec:verbs}.
+Finally, we search for factors motivating the resistance of these verbs, discuss the findings and put them in a general context of language change and morphology in \cref{sec:discussion}.

### 2021-11-15 documents/sections/extensions.tex
revised proto-pekodian *k-

@@
 Of the six attested incomplete extensions, three can be reconstructed to intermediate proto-languages, while three others happened in earlier stages of single modern languages.
-The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
+The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
+We discuss each extension separately, contrasting regular and innovative verbs with irregular and conservative verbs, and reconstructing forms where necessary:
 \cref{sec:pekodian} investigates the innovation of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
@@
 \textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work on \PPek morphosyntax can be found in the literature.
-All three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.
-We can thus reconstruct a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
+However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.
+Thus, it is possible to reconstruct a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
 
@@
 
-For \arara, \textcite{alves2017arara} describes six\footnote{Seven under her analysis, which sees the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.} \gl{s_a_} verbs which have a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
-From a comparative view this list is not quite complete, as there is also a reflex of the copula \rc{a[p]}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
+In the most detailed description of a Pekodian language, \textcite{alves2017arara} describes six%
+\footnote{Seven under her analysis, which sees the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.}
+\arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
+From a comparative perspective, this list is not quite complete, as there is also a reflex of the copula \rc{a[p]}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
 However, its inflectional morphology features verbal \setone prefixes, including first person \obj{w-} \exref{araap}.
@@
 \xe
-%
+
 \ex<araap> \arara \parencite[200]{alves2017arara}\\
@@
 %
-Of course, \qu{to bathe} is also found in the \obj{w-}list for \arara.
+Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
 While \textcite[4]{meira2003bakairi} does list \obj{ge} \qu{to say}, \obj{tə} \qu{to go}, and \obj{əe(wɨ)} \qu{to come} as examples of \gl{s_a_} verbs, he does not indicate whether they belong to the class of \gl{s_a_}-1 verbs, with first person \obj{k-}, or the \gl{s_a_}-2 verbs, with \obj{w-}.%
@@
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-Further, while there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, only \rc{eti} reflexes seem to occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
+Further, while there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, it seems that only reflexes of \rc{eti} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
 
@@
 Similarly, the supposed change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
-Thus, we reconstruct a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} to \PPek, identical to the \arara one in form and distribution.
+Thus, we reconstruct a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} to \PPek, identical to the \arara one both in form and distribution.
 
@@
 For \qu{to be}, \ikpeng \obj{e} is very likely the original vowel, given the \PC form \rc{eti} \pcref{sec:be}.
-For \qu{to go down}, we reconstruct \rc{ɨ} as the initial vowel rather than \rc{i}, see \cref{sec:godown}.
+For \qu{to go down}, we reconstruct \rc{ɨ} as the initial vowel rather than \rc{i} \pcref{sec:godown}.
 Further, the forms are not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
@@
 On the other hand, both \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}.
+Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}; similar issues are found outside of Pekodian \pcref{sec:come}.
 
-The V-initial nature of \qu{to go} is evidenced in its Xinguan forms, while \bakairi \phon{\rc{wɨ}}{\obj{u}} muddied the morpheme boundary; other forms are V-initial \pcref{sec:go}.
+Finally, the V-initial nature of \PPek \qu{to go} is evidenced in its Xinguan forms; while the \bakairi change \phon{\rc{wɨ}}{\obj{u}} obscured the morpheme boundary, other forms are V-initial \pcref{sec:go}.
 The \ikpeng form \obj{aran} is compatible with our reconstruction \rc{ɨtən} when considering that \ikpeng \obj{a} is an attested outcome of \rc{ə}.\footnote{\begin{inlinelist}
@@
 Summing up, an innovative \gl{1}\gl{s_a_} marker \rc{k-} is reconstructible to \PPek.
-Seven verbs can be reconstructed as having resisted this innovation, and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
+Seven verbs can be reconstructed as having resisted this innovation and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
 In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
-Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ}, accompanied by the extension of \obj{k-}.
+Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ}, accompanied by the introduction of first person5 \obj{k-}.
 

### 2021-11-15 documents/sections/extensions.tex
waiwaian *k- review

@@
 \label{sec:waiwaian}
-This extension was briefly discussed in \cref{sec:extensions_intro}; the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was later extended to \gl{1}\gl{s_a_} in \PWai.
+This extension led to the \hixka pattern from \cref{tab:hixintro} in \cref{sec:irreg_intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
+In \PWai, the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was extended to \gl{1}\gl{s_a_}.
 For regularly inflected verbs, this created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \pcref{tab:pwaireg}.
-%
 \input{floats/pwaireg}
-%
-%\ex<wais> Waiwaian \rc{eɸurka} \qu{to fall} (\gl{s_a_}) and \rc{wɨnɨkɨ} \qu{to sleep} (\gl{s_p_})\\
-%\begin{tabular}[t]{@{}lllllll@{}}
-%& \multicolumn{2}{l}{\PWai} & \multicolumn{2}{l}{\hixka}  & \multicolumn{2}{l}{\waiwai} \\
-% & \gl{s_a_} & \gl{s_p_} & \gl{s_a_} & \gl{s_p_} & \multicolumn{2}{l}{\gl{s}} \\
-%\gl{1} & \rc{\emp{k-}eɸurka-} & \rc{kɨ-wɨnɨkɨ-} & \obj{\emp{k-}ehurka-} & \obj{kɨ-nɨkɨ-} & \obj{\emp{k-}eɸɨrka-} & \obj{kɨ-wɨnk-} \\
-%\gl{2} & \rc{m-eɸurka-} & \rc{o-wɨnɨkɨ-} & \obj{m-ehurka-} & \obj{o-wnɨkɨ-} & \obj{m-eɸɨrka-} & \obj{mɨ-wɨnk-} \\
-%\gl{1+2} & \rc{t-eɸurka-} & \rc{tɨt-wɨnɨkɨ-} & \obj{t-ehurka-} & \obj{tɨ-nɨkɨ-} & \obj{t͡ʃ-eɸɨrka-} & \obj{tɨt-wɨnk-} \\
-%\gl{3} & \rc{ɲ-eɸurka-} & \rc{nɨ-wɨnɨkɨ-} & \obj{ɲ-ehurka-} & \obj{nɨ-nɨkɨ-} & \obj{ɲ-eɸɨrka-} & \obj{nɨ-wɨnk-} \\
-%\end{tabular}\\
-%\parencites[510]{howard2001wrought}[189--191]{hixkaryanaderby1985}[209--211]{hawkins1953waiwai}[50]{waiwaihawkins1998}
-%\xe
-%
-Not all verbs were affected by this extension, though: \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
+Not all \gl{s_a_} verbs were affected: \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
 Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, a potential reflex of \gl{1}\gl{s_a_} \rc{w(ɨ)-}.
@@
 %
-This correspondence allows us to establish \hixka \obj{ɨ-} a reflex of \rc{w(ɨ)-}, similar to \ikpeng \pcref{sec:pekodian}.
-Notably, \textcite{hixkaryanaderby1985} does not see this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_}, but as the regular \gl{1}>\gl{3} marker, because he considers \hixka \obj{ka} \qu{to say} to be transitive, an issue which will be discussed in \cref{sec:say}.
+This correspondence allows us to establish \hixka \obj{ɨ-} a reflex of \rc{w(ɨ)-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
+Notably, \textcite{hixkaryanaderby1985} does not see this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_} prefix, but as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive  \pcref{sec:say}.
 

### 2021-11-15 documents/sections/extensions.tex
tiriyoan *t- review

@@
 \textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar.
-The \setone paradigms of \trio and \akuriyo contain an interesting puzzle: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
-This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}; the latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
-
+He faced an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
+This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
+\footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
 \input{floats/ptirreg}
-%This is illustrated by comparing reconstructible \PTir forms with more conservative \wayana \exref{sat} and \kalina \exref{13w}, which can be taken as representative for \PC.
-%
-%\ex<sat>
-%\begin{tabular}[t]{@{}llllll@{}}
-%\gl{1}\gl{s_a_} & \PC & \wayana & \PTir & \trio & \akuriyo\\
-%\qu{yawn} & \rc{w-e-mɨta-po-} & \obj{w-emtapɨ-} & \rc{t͡ʃ-entapo-} & \obj{s-entapo-} & \obj{t͡ʃ-entapo-}\\
-%\qu{bathe} & \rc{w-e-pɨ-} & \obj{w-epɨ-} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-}\\
-%\qu{talk} & \rc{w-ət-uru} & \obj{w-əturu-} & \rc{t-əturu-} & \obj{t-əturu-} & \obj{k-əturu-}\footnotemark\\
-%\qu{learn} & \rc{w-ət-eme-pa} & \obj{w-əhepa-} & \rc{t-əempa-}\footnotemark & \obj{t-əenpa-} & \obj{k-əempa-}\\
-%\end{tabular}\\
-%\parencites[28]{camargo2010wayana}[292, 664]{triomeira1999}[77, 86, 82]{gildea1994akuriyo}[50, 208, 213]{wayanatavares2005}
-%\xe
-%\addtocounter{footnote}{-1}
-%\footnotetext{\PTir \rc{t-} was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
-%\addtocounter{footnote}{1}
-%\footnotetext{
-%The \trio form is not due to dissimilation from \rc{mp} to \obj{np}, but rather a phonemic transcription rendering the syllable-final archiphoneme \pnm{N} as \obj{n} \parencite[31]{triomeira1999}.
-%Further, \PTar intervocalic \rc{t͡ʃ} was lost entirely in \trio (except after \rc{i}) and is reflected as \obj{ʔ} in \akuriyo \parencite[31--32]{meira1998proto}.
-%Given the original shape of the detransitivizers \rc{ət(e)-} and \rc{e-}, \trio \obj{ə-enpa} \qu{to learn} contains a reflex of the former with loss of \rc{t}, so \akuriyo \obj{əempa} would be expected to have a glottal stop: ?\obj{əʔempa}.
-%Its absence could either be due to \trio influence or not being transcribed by \textcite{gildea1994akuriyo}.
-%Since there is no evidence for a glottal stop in \PTir (\rc{əʔempa}), I do not reconstruct it.}
-%
-%\ex<13w> 
-%\begin{tabular}[t]{@{}llllll@{}}
-%\gl{1}>\gl{3}& \PC & \kalina & \PTir & \trio & \akuriyo \\
-%\qu{see} & \rc{t-ene-} & \obj{s-ene-} & \rc{w-ene-} & \obj{w-ene-} & \obj{(w-)ene-} \\
-%\qu{bring} & \rc{t-enepɨ-} & \obj{s-enepɨ-} & \rc{w-enepɨ-} & \obj{w-enepɨ-} & Ø\obj{-enepɨ-} \\
-%\end{tabular}\\
-%\parencites[425--426]{hoff1968carib}[28, 304]{triomeira1999}[86--87]{gildea1994akuriyo}
-%\xe
-%%
-%Thus, \wayana preserves the \PC \gl{1}\gl{s_a_} marker \rc{w-}, while it was replaced by \rc{t-} (\phonr{}{\rc{t͡ʃ-}}{\rc{e}}) in \PTir \exref{sat}.
-%%Both \qu{to yawn} and \qu{to bathe} are transparently derived \gl{s_a_} verbs, from \rc{mɨta-po} \qu{to open the mouth of} and \rc{pɨ} \qu{to wash}, respectively.
-%On the other hand, \gl{1}>\gl{3} scenarios show a reflex of \rc{t-} in \kalina, while they took on \rc{w-} in \PTir \exref{13w}.
-%
 The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
@@
 Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
-We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it happening already in \PTar.}
+We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it taking place already in \PTar.}
 %
-As for intransitive verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar; we offer our reconstructed \PTir forms.
-To this list, we add the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}}, which has first person \obj{w-}, at least in \trio.
+Regarding \gl{s_a_} verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar; we offer our reconstructed \PTir forms.
+To this list, the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}} can be added, which has first person \obj{w-}, at least in \trio.
 
@@
 \akuriyo (and \carijo) then levelled this original distribution, similar to what we have suggested for Pekodian \pcref{sec:pekodian}.
-We agree with this scenario, with the exception that \trio \obj{əepɨ} is clearly a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ}.
+We agree with this scenario, with the exception that \trio \obj{əepɨ} looks like a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ}.
 
@@
 Further, these verbs are not mentioned by \textcite{meira1998proto}, who was also working with \posscite{gildea1994akuriyo} data.
-Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by Meira's data, we cannot reconstruct these verbs as not being affected by the extension of \PTir \rc{t-}.
+Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by the available part of \posscite{meira1998proto} data, we cannot reconstruct these verbs as not being affected by the extension of \PTir \rc{t-}.
 

### 2021-11-15 documents/sections/extensions.tex
tiriyoan *t- review

@@
 \label{sec:taranoan}
-The Taranoan languages were already grouped together by \textcite{girard1971proto}, consisting of closely related \trio and \akuriyo (subsumed here under the moniker Tiriyoan), and the more distant member \carijo.
+The moniker Tiriyoan subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages already identified by \textcite{girard1971proto}, the more distant member being \carijo.
 \textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar.
-He faced an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
+He faces an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
 This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
@@
 %
-Regarding \gl{s_a_} verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar; we offer our reconstructed \PTir forms.
+Regarding \gl{s_a_} verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar -- attentive readers my recognize \trio \qu{to go} -- for which we provide our reconstructed \PTir forms.
 To this list, the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}} can be added, which has first person \obj{w-}, at least in \trio.

### 2021-11-15 documents/sections/extensions.tex
akuriyo k- review

@@
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} might be more recent, with which we agree; since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
+He also suggests that \obj{k-} might be more recent, with which we agree: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
 The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
-Since there are no examples of or information about \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.
+Since there are no examples of, or further information about, \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.
 
@@
 
-All verbs in \cref{tab:ptirverbs} above did of course not introduce \obj{k-}, either; we do not know the first-person form of the copular verb \obj{eʔi}.
+The verbs listed for \PTir in \cref{tab:ptirverbs} above did of course also not introduce \obj{k-} in \akuriyo, although we do not know the first-person form of the copular verb \obj{eʔi}.
 In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in \akuriyo, also reconstructible to \PTir \exref{tridown}.
-It was not affected by the extension of \akuriyo \obj{k-}, but whether it existed when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).
+It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{s_a_} verb when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).
 
-\ex<tridown> First person forms of \qu{to go down} \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
-\begin{tabular}[t]{@{}ll@{}}
-\trio & \obj{p-ɨhtə-} \\
-\akuriyo & \obj{p-ɨtə-}\\
+\ex<tridown> First person forms of \qu{to go down}\\
+\begin{tabular}[t]{@{}lll@{}}
+\trio & \obj{p-ɨhtə-} & \parencites[294]{triomeira1999} \\
+\akuriyo & \obj{p-ɨtə-} & \parencite[84]{gildea1994akuriyo}\\
 \end{tabular}

### 2021-11-15 documents/sections/extensions.tex
karijona j- review

@@
 \label{sec:carijo}
-As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.}
-This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \pcref{tab:carreg}.
+\carijo, the cousin of the Tiriyoan languages, has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
+Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs \pcref{tab:carreg}.
 %
@@
 %
-Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by taking a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}, shown in \exref{car-30}.
+Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}, shown in \exref{car-30}.
 %The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
@@
 However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
+In addition, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car1w.car-33}.
 
-\pex<car1w>\carijo \parencite[][5, 42]{guerrero2016karihona}
+\pex<car1w>\carijo \parencite[][5, 42, personal communication]{guerrero2016karihona}
 \a<car-24>
@@
 \endgl
-\xe
-%
-In addition, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car-33}.
-%We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
-Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would either expect \obj{wɨ-} or \obj{ji-}.
-We analyse this ∅ as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
-Further, one may suggest the difference between C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} to result from the latter's earlier V-initial nature \pcref{sec:go}.
-
-\ex<car-33>\carijo \perscomm{David Felipe Guerrero-Beltran}\\
+\a<car-33>
 \begingl
@@
 \xe
+%We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
+Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would either expect conservative \obj{wɨ-} or innovative \obj{ji-}.
+We analyse this ∅ as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
+Further, it is possible that the divergent development of \rc{w-} on C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} is a result of the latter's originally V-initial nature \pcref{sec:go}.
 

### 2021-11-15 documents/sections/extensions.tex
yukpa revision

@@
 \label{sec:yukpa}
-\yukpa has lost most of its \setone constructions, mostly employing innovative ones \parencite{meira2006syntactic}.
-It preserves the \setone prefixes in the immediate past, where the \gl{1}\gl{s_p_} marker appears to have been extended to \gl{s_a_} verbs, as evidenced by the paradigms in \cref{tab:yukreg}.
-The loss of split-\gl{s} in favor of the \gl{1}\gl{s_p_} marker is accompanied by the \gl{2}\gl{s_a_} marker \obj{m(ɨ)-} extending to \gl{s_p_} verbs, and \gl{1+2} being lost as an inflectional value.
-%It has lost the split-\gl{s} system in general
-
-%\ex<yukpaintr> \yukpa \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}\\
-%\begin{tabular}[t]{@{}llll@{}}
-%& \obj{otum} \qu{to wash self} & \obj{nɨ} \qu{to sleep} & \obj{ata} \qu{to fall}\\
-%\gl{1} & \obj{j-otum-} & \obj{jɨ-nɨ-} & \obj{j-ata-}\\
-%\gl{2} & \obj{m-otum-} & \obj{mɨ-nɨ-} & \obj{m-ata-}\\
-%\gl{3} & \obj{n-otum-} & \obj{nɨ-nɨ-} & \obj{n-ata-}\\
-%\end{tabular}
-%\xe
-
-\input{floats/yukreg}
-
-The normal intransitive first person marker is \obj{j(ɨ)-}, including those with clear reflexes of the detransitivizer \detrz, like \obj{otum} \qu{to wash self} in \cref{tab:yukreg}.
-The analysis that this is a reflex of the \gl{1}\gl{s_p_} marker \rc{u(j)-} is supported by the same form occurring in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.
+The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms; it mostly employs innovative ones \parencite{meira2006syntactic}.
+However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm is found \pcref{tab:yukreg}.
+The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
+%
+\input{floats/yukreg}%
+%
+These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their reflex of the detransitivizer \detrz.
+The same prefix occurs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, meaning that \obj{j(ɨ)-} is the reflex of the \PC \gl{1}\gl{s_p_} marker \rc{u(j)-}.
 
@@
 On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion across the family \pcref{sec:pekodian,sec:waiwaian}, we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family (see \cref{sec:pekodian,sec:waiwaian}), we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}, contrasting with the reflex of \rc{u(j)-}, \obj{j(ɨ)-}.
 
-In intransitive verbs, this first-person zero marking is attested only for \obj{to} \qu{to go} \exref{yuk-7}.
-It contrasts with regular C-initial verbs taking \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
+In intransitive verbs, this first-person zero marking is also attested, but only for \obj{to} \qu{to go} \exref{yuk-7}.
+That verb contrasts with regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
 Thus, we conclude that the extension of \obj{j-} in \yukpa stopped short of \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.

### 2021-11-16 documents/sections/extensions.tex
more work on yukpa

@@
 \label{sec:yukpa}
-The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms; it mostly employs innovative ones \parencite{meira2006syntactic}.
+The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
 However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm is found \pcref{tab:yukreg}.
@@
 On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family (see \cref{sec:pekodian,sec:waiwaian}), we argue that the zero in \gl{1}>\gl{3} scenarios is in fact the \yukpa reflex of \rc{w-}, contrasting with the reflex of \rc{u(j)-}, \obj{j(ɨ)-}.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family (see \cref{sec:pekodian,sec:waiwaian}), we identify the zero marking in \gl{1}>\gl{3} scenarios as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
 
@@
 That verb contrasts with regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
-Thus, we conclude that the extension of \obj{j-} in \yukpa stopped short of \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.
+In summary, the extension of \obj{j-} in \yukpa only spared \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.
 

### 2021-11-16 documents/sections/verbs.tex
discussion and factor comparison
'to be'

@@
 In \cref{sec:extensions}, we introduced six distinct extensions of personal prefixes into \gl{1}\gl{s_a_} territory, and identified verbs resistant to each innovation.
-The set of unaffected verbs was rather small in most cases, and (proto-)languages show a considerable degree of overlap in what verbs are unaffected.
-In this section, we present these verbs from a comparative perspective.
-\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}; these are all verbs that \textcite{gildea2007greenberg} reconstruct as \gl{s_a_} verbs that were not derived from transitive verbs.
+The set of unaffected verbs is rather small in most cases, and (proto-)languages show a considerable degree of overlap in what verbs are unaffected.
+Therefore, this section presents these verbs from a comparative perspective.
+\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}.
 \cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek, and  \cref{sec:bathe} investigates \rc{e-pɨ} \qu{to bathe}, of which the \PPek reflex \rc{i-pɨ} resisted the extension of \rc{k-}.
-The \obj{e}-initial verbs not affected by the extension of \rc{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
+The \obj{e}-initial verbs not affected by the extension of \obj{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
 
@@
 \label{sec:be}
-For a comprehensive comparative overview for this verb, we refer the reader to \textcite[375--382]{gildea2018reconstructing}, who reconstructs two distinct roots serving as verbs \qu{to be} in modern Cariban languages.
-One is the older copula \rc{a[p]}, which can be reconstructed as already having various irregularities in \PC.
-The other is a root \rc{eti} reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
+For a comprehensive comparative overview for these two roots, we refer the reader to \textcite[375--382]{gildea2018reconstructing}.
+\rc{a[p]} is the original copula and can be reconstructed as already having various irregularities in \PC.
+\rc{eti}, the other root, is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
 Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam} value.
@@
 \akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while \obj{eʔi} is not attested in a first-person form. 
-\carijo innovated \obj{j-}, but only in the \obj{et͡ʃi} root allomorph \exref{mivida-12}; the \obj{a} root preserves \obj{w-} \pcref{sec:carijo}.
-\yukpa innovated \obj{j-} for reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
+\carijo innovated \obj{j-}, but only in the reflex of \rc{eti} \exref{mivida-12}; the \obj{a} root preserves \obj{w-} \pcref{sec:carijo}.
+\yukpa introduced \obj{j-} to the reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
 
@@
 %In any case, \akuriyo is only sparsely documented, and there are many open questions about it; the \akuriyo data in \textcite{gildea1994akuriyo} are not fully reliable.
-%
-%\ex<aku-106>\akuriyo \parencite[][84]{gildea1994akuriyo}\\
-%\begingl
-%\gla t͡ʃ-ewai//
-%\glb \gl{1}-sit.down//
-%\glft \qu{I sat down.}//
-%\endgl
-%\xe

### 2021-11-16 documents/sections/verbs.tex
'to say'

@@
 %
-A comparison of the longest forms attested in each language is shown in \cref{tab:say}.
-Segments of cognate forms in this and other tables were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.
+\cref{tab:say} shows a comparison of the longest attested forms for each language.
+Cognate segments in \cref{tab:say,tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.
 %
-\input{floats/say.tex}
+\input{floats/say.tex}%
 %
 This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}).
-We do not know the first person prefix of its \yukpa reflex \obj{ka}.
+We do not know the first person form of its \yukpa reflex \obj{ka}.
 
 As briefly mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes this verb as transitive in \hixka.
-This analytical choice is not only motivated by avoiding an idiosyncratic intransitive first person prefix \obj{ɨ-}:
-\hixka \obj{ka} also behaves like a transitive verb in other ways, for instance by showing the complementary distribution of the third person marker \obj{n-} and preceding objects -- in this case direct speech or ideophones \exref{hixsay}.
+This analytical choice is not only motivated by the desire to avoid an idiosyncratic intransitive first person prefix \obj{ɨ-} instead of \obj{kɨ-}.
+\hixka \obj{ka} also shows the complementary distribution of third person  \obj{n-} and preceding objects typical of transitive verbs in Cariban \parencite[60--81]{gildea1998}.
+Due to its semantics, these objects are either direct speech \exref{hixsay} or ideophones.
 
@@
 In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
-This complementary distribution is otherwise only found with transitive verbs \parencite[59--60]{gildea1998}.
-The verb shows the same pattern, albeit inconsistently, in \trio \parencite[267]{triocarlin2004}.
+At least the \trio cognate shows the same pattern, albeit inconsistently \parencite[267]{triocarlin2004}.
 
-Further comparative evidence also points to \rc{ka[ti]} \qu{to say} showing transitive traits:
+Comparative evidence also points to \rc{ka[ti]} \qu{to say} showing transitive traits:
 \trio \obj{ka} is characterized as the only intransitive verb being able to take the causative suffix \obj{-po} and the agentive nominalizer \obj{-ne} \parencite[263, 169]{triomeira1999}.
@@
 The agent nominalizer \rc{-ne} gave rise to the \panare gnomic verbal suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
-The occurrence of \obj{-ɲe} on \obj{ka} has led to differing categorizations as transitive \parencite[214]{panarepayne2013} or intransitive \parencite[102]{mattei1994diccionario}.
-Reflexes of another transitive causativizer \rc{-metipo} \parencite{gildea2015valency} are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
+The occurrence of \obj{-ɲe} on \obj{ka} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, contrastive with the intransitive analysis by \textcite[102]{mattei1994diccionario}.
+Finally, reflexes of the causativizer \rc{-metipo}, usually restricted to transitive verbs \parencite{gildea2015valency}, are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
 
-Our classification of \qu{to say} as an intransitive verb is supported primarily by prefix patterns:
+Our categorization of \qu{to say} as an intransitive verb is supported primarily by its person prefixes.
 \kalina offers a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.\footnote{Interestingly, the \kalina causativized form \obj{kapo} \qu{to make say} does not have the regular \gl{1}>\gl{3} prefix \obj{s(i)-}, but irregular \obj{w(ɨ)-} \parencite[430]{courtz2008carib}.}
 Similarly, \qu{to say} in Pekodian languages has a reflex of \gl{1}\gl{s} \rc{w-} \pcref{sec:pekodian}, and not \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
-Additionally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{kar-84}, and it is probably reflected in vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
+Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show no \obj{i} for this verb, see \exref{sayintr.aku-154} and  \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib}.
+Finally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and it is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
 
-\ex<kar-84>\kalina \parencite[][202]{courtz2008carib}\\
+\pex<sayintr>
+\a<aku-154> \akuriyo \parencite[][113]{meira1998proto}\\
 \begingl
-\glpreamble Òmakon \`wa oti ywykàpo kaiko.//
+\gla mɨ-ka//
+\glb \gl{2}-say//
+\glft \qu{You said.}//
+\endgl
+\a<kar-84> \kalina \parencite[][202]{courtz2008carib}\\
+\begingl
+\glpreamble Òmakon `wa oti ywykàpo kaiko.//
 \gla o-ʔma-kon ʔwa oti ɨ-wɨ-ka-ʔpo kai-ko//
@@
 \xe
-%
+
 Summing up, this verb could be reconstructed as being intransitive based on its prefixes, but transitive based on some suffixes.
 \hixka has lost the main intransitive criteria, making its reflex look more like a transitive verb.
-It is not attested as being affected by any of the person marker extensions under discussion.
+It is not attested as being affected by any of the person marker extensions  in \cref{sec:extensions}.
 

### 2021-11-16 documents/floats/say.tex
added yukpa ka, generating all tables again

@@
 \centering
-\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102; p.c., Spike Gildea]{mattei1994diccionario}}
+\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63; p.c., Spike Gildea]{largo2011yukpa}}
 \label{tab:say}
@@
 \uxc     &      \obj{ki} &  k &  i &    &    \\
+\yukpa   &      \obj{ka} &  k &  a &    &    \\
 \bottomrule

### 2021-11-16 documents/sections/verbs.tex
added yukpa ka, generating all tables again

@@
 %
-\cref{tab:say} shows a comparison of the longest attested forms for each language.
-Cognate segments in \cref{tab:say,tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.
+\cref{tab:say} shows a comparison of the longest attested forms for each language.%
+\footnote{Cognate segments in \crefrange{tab:say}{tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.}
 %
@@
 In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
-At least the \trio cognate shows the same pattern, albeit inconsistently \parencite[267]{triocarlin2004}.
+At least the \trio cognate shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
 
-Comparative evidence also points to \rc{ka[ti]} \qu{to say} showing transitive traits:
+Derivational suffixes also point to \rc{ka[ti]} \qu{to say} being transitive:
 \trio \obj{ka} is characterized as the only intransitive verb being able to take the causative suffix \obj{-po} and the agentive nominalizer \obj{-ne} \parencite[263, 169]{triomeira1999}.
@@
 The agent nominalizer \rc{-ne} gave rise to the \panare gnomic verbal suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
-The occurrence of \obj{-ɲe} on \obj{ka} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, contrastive with the intransitive analysis by \textcite[102]{mattei1994diccionario}.
+The occurrence of \obj{-ɲe} on \obj{ka} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, contrasting with the intransitive analysis by \textcite[102]{mattei1994diccionario}.
 Finally, reflexes of the causativizer \rc{-metipo}, usually restricted to transitive verbs \parencite{gildea2015valency}, are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.

### 2021-11-16 workflow/process_data.py

@@ -291,7 +291,6 @@ for lg, meanings in {
     tabular.rename(columns=trans_dic, inplace=True)
     tabular = tabular.apply(lambda x: x.apply(objectify))
     tabular.columns.names = [None]
-    print(tabular)
     save_float(
         print_latex(tabular, keep_index=True),
         label,
@@ -299,865 +298,850 @@ for lg, meanings in {
         short=f"Some {print_shorthand(lg)} verbs",
     )
 
-#
-# # regular pekodian verbs
-# tabular = pd.DataFrame()
-# sources_raw = []
-# pyd.x = ["Language_ID", "Meaning_ID"]
-# pyd.y = ["Inflection"]
-# for lg, meaning in {"bak": "go_up", "ara": "dance", "ikp": "run"}.items():
-#     pyd.filters = {"Language_ID": [lg], "Meaning_ID": [meaning]}
-#     pyd.sort_orders = {"Inflection": person}
-#     temp = pyd.compose_paradigm(i_df, multi_index=True)
-#     sources_raw.extend(get_sources(i_df, latexify=False))
-#     tabular = pd.concat([tabular, temp], axis=1)
-#
-# sources = cldfh.cite_a_bunch(sources_raw, parens=True)
-# label = "pekreg"
-# tabular.index.names = [None]
-# export_csv(
-#     tabular.rename(columns=name_dic, level="Language_ID"),
-#     label,
-#     "Regular Pekodian Sa verbs",
-#     keep_index=True,
-#     sources=sources_raw,
-# )
-#
-# tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name[1])))
-# tabular.rename(columns=trans_dic, level="Meaning_ID", inplace=True)
-# tabular.rename(columns=shorthand_dic, level="Language_ID", inplace=True)
-#
-# tabular.columns.names = [None, None]
-# tabular.index = tabular.index.map(pynt.get_expex_code)
-# tabular.columns = [" ".join([x, y]) for x, y in tabular.columns]
-#
-#
-# save_float(
-#     print_latex(tabular, keep_index=True),
-#     label,
-#     "Regular Pekodian \gl{s_a_} verbs " + sources,
-#     short="Regular Pekodian \gl{s_a_} verbs",
[…]

### 2021-11-16 documents/sections/verbs.tex
'to say'

@@
 Similarly, \qu{to say} in Pekodian languages has a reflex of \gl{1}\gl{s} \rc{w-} \pcref{sec:pekodian}, and not \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
-Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show no \obj{i} for this verb, see \exref{sayintr.aku-154} and  \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib}.
+Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show no \obj{i} for this verb, see \exref{sayintr.aku-154} as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for the verb's inflection in individual languages.
 Finally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and it is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.

### 2021-11-16 documents/sections/verbs.tex
'to go' and 'to come'

@@
 \footnote{As indicated by the brackets in \cref{tab:go}, there are many languages where the initial vowel is only present in some forms.
-Also, the prefix-verb boundary in many inflected forms like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} is ambiguous, since an epenthetic \obj{ɨ} breaks up potential CC clusters.}
+Also, the prefix-verb boundary in many inflected forms like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} is ambiguous, since an epenthetic \obj{ɨ} breaks up potential CC clusters.
+Still, when one considers unambiguous forms, the contrast with \rc{ka[ti]} becomes very clear.}
 This verb was not affected by the any of the extensions discussed in \cref{sec:extensions}.
@@
 Crucially, the \rc{ət} part of their reconstruction is not reflected in the majority of forms, and many languages ostensibly have reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ} for the \rc{epɨ} part.
-We analyze these forms as going back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizer and a root \rc{jəpɨ}.
+We analyze these forms as going back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
 
-Evidence for the originally \rc{j}-initial nature of the root is found in the Pemongan languages and \kaxui, although the simultaneous presence of \obj{oohɨ} in \kaxui raises the question of whether the \obj{j} is part of the root in \obj{johɨ}.
-This form only occurs in the Progressive \exref{kaxprog.kax-82}, and indeed \obj{j(-)} may be a reflex of \rc{i-w-} \qu{\gl{3}-\gl{s_a_}-}, as the \gl{s_a_} class marker \rc{w-} is present with other person values \exref{kaxprog.kax-54}.
+Evidence for the originally \rc{j}-initial nature of the root is found in the Pemongan languages and \kaxui, although the coexistence of \obj{oohɨ} and \obj{johɨ} in \kaxui raises the question of whether the \obj{j} in the latter form is actually part of the root.
+The form in question only occurs in the Progressive \exref{kaxprog.kax-82}, and it seems possible that \obj{j(-)} is actually a reflex of \rc{i-w-} \qu{\gl{3}-\gl{s_a_}-}, since the \gl{s_a_} class marker \rc{w-} is present with other person values \exref{kaxprog.kax-54}.
 However, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial \gl{s_a_} verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
-Thus, we are confident that this \obj{j} is indeed part of the root, rather than an outcome of \rc{i-w-}, which in turn allows us to reconstruct \rc{j} back to \PC.
+Thus, the best explanation for this \obj{j} is that is indeed part of the root, rather than an outcome of \rc{i-w-}, which in turn allows us to safely reconstruct \rc{j} back to \PC.
 
@@
 %
-Most forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for this \rc{j}.
-However, we interpret the \obj{i} in the \akawaio form \obj{əsipɨ} as additional evidence of the sequence \rc{jə}, which has the same outcome \obj{i} in the \macushi reflex of bare \rc{jəpɨ}, \obj{ipɨ}.
+Most of the longer forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for this \rc{j}, but we identify the \obj{i} in the \akawaio form \obj{əsipɨ} as a reflex of the sequence \rc{jə}, which has the same outcome in the \macushi reflex of bare \rc{jəpɨ}, \obj{ipɨ}.
 
-Turning to the many forms apparently reflecting \rc{əpɨ} and \rc{epɨ}, we find that both are distributed widely in the family, sometimes even co-occurring in the same language.
+Turning to the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ}, we find that both are distributed widely in the family, sometimes even co-occurring in the same language.
 A unifying account of these forms requires the root \rc{jəpɨ} to undergo two major sound changes: \begin{inlinelist}
@@
 
-Our interpretation of the \rc{ət} portion as a detransitivizer is primarily based on its form, as well as its paradigmatically conditioned occurrence in some languages.
+Our interpretation of the \rc{ət} portion as a detransitivizer is primarily based on its form, as well as its paradigmatically conditioned occurrence in some languages, discussed shortly.
 Although the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
@@
 Also, \waiwai \qu{go to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced by its status in \kaxui, \kalina, \panare \exref{pan-128}, \arara, and \trio.
+The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced by its status in \kaxui, \kalina, \arara, \trio, and \panare \exref{pan-128}.
 
@@
 
-While these sound changes and the addition of \rc{ət-} do account for the majority of the forms in \cref{tab:come},\footnote{
-Apart from aforementioned \akuriyo \obj{eepɨ}, another irregular form is \apalai \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
+While a root \rc{jəpɨ}, aforementioned sound changes, and the addition of \rc{ət-} do account for the majority of the forms in \cref{tab:come},\footnote{
+Apart from aforementioned \akuriyo \obj{eepɨ}, another diachronically irregular form is \apalai \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
 While \obj{oepɨ} would be a regular outcome of \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
 The form may be due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}.
-Alternatively, \apalai \obj{oepɨ} could be a fossilized loan from \wayana, which has replaced its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
+Alternatively, \apalai \obj{oepɨ} could be a fossilized loan from \wayana, which has replaced its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of the intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
 the distribution within the family is rather chaotic.
 Apart from the distribution of \rc{əpɨ} and \rc{epɨ}, forms with and without \rc{ət-} can be found within the same language, usually conditioned by different prefixes.
-This was briefly discussed in \cref{sec:pekodian} for \arara and in \cref{sec:taranoan} for \trio (and \PTir and \PTar).
+This was briefly discussed in \cref{sec:pekodian} for \arara (and \PPek) and in \cref{sec:taranoan} for \trio (\PTar).
 To illustrate, the \trio \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}%
@@
 As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}, but not on the \arara reflex of \rc{jəpɨ}.
-Both reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ} and/or \rc{epɨ} (\akuriyo) resisted the introduction of \rc{t-} in \PTir.
+Both reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ} and/or \rc{jəpɨ} (\akuriyo) resisted the introduction of \rc{t-} in \PTir.
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.

### 2021-11-17 documents/sections/verbs.tex
'to come'

@@
 
-This verb is reconstructed as \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all the attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
-Crucially, the \rc{ət} part of their reconstruction is not reflected in the majority of forms, and many languages ostensibly have reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ} for the \rc{epɨ} part.
-We analyze these forms as going back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
+This verb is reconstructed as \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
+Crucially, the majority do not reflect the \rc{ət} part of their reconstruction, and many forms are ostensibly reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ} for the \rc{epɨ} part.
+We analyze all forms as going back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
 
 Evidence for the originally \rc{j}-initial nature of the root is found in the Pemongan languages and \kaxui, although the coexistence of \obj{oohɨ} and \obj{johɨ} in \kaxui raises the question of whether the \obj{j} in the latter form is actually part of the root.
-The form in question only occurs in the Progressive \exref{kaxprog.kax-82}, and it seems possible that \obj{j(-)} is actually a reflex of \rc{i-w-} \qu{\gl{3}-\gl{s_a_}-}, since the \gl{s_a_} class marker \rc{w-} is present with other person values \exref{kaxprog.kax-54}.
+\obj{johɨ} only occurs with third person in the Progressive \exref{kaxprog.kax-82}, and one might analyze \obj{j(-)} as a reflex of \rc{i-w-} \qu{\gl{3}-\gl{s_a_}-}, since the \gl{s_a_} class marker \rc{w-} is present with other person values \exref{kaxprog.kax-54}.
 However, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial \gl{s_a_} verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
@@
 %
-Most of the longer forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for this \rc{j}, but we identify the \obj{i} in the \akawaio form \obj{əsipɨ} as a reflex of the sequence \rc{jə}, which has the same outcome in the \macushi reflex of bare \rc{jəpɨ}, \obj{ipɨ}.
+Most of the longer forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for the segment \rc{j}, but the \obj{i} in the \akawaio form \obj{əsipɨ} is very likely a reflex of the sequence \rc{jə}.
+This analysis is supported by the reflex \obj{ipɨ} from bare \rc{jəpɨ} in very closely related \macushi.
 
-Turning to the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ}, we find that both are distributed widely in the family, sometimes even co-occurring in the same language.
+Turning to the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ}, we find that both are distributed widely in the family, sometimes even co-occurring in the same language.
 A unifying account of these forms requires the root \rc{jəpɨ} to undergo two major sound changes: \begin{inlinelist}
@@
 Both phenomena are found in other contexts throughout the family \parencite{meira2010origin}.
-For this particular verb, these sound changes appear to have applied irregularly, and not always in the same order.
+However, these sound changes appear to have applied irregularly to this verb, and not always in the same order.
 For example, the \kalina form \obj{opɨ} can only be explained if \rc{j} was lost before the umlaut of \rc{ə} to \rc{e}, which would have been triggered by \rc{j}.
@@
 
-Our interpretation of the \rc{ət} portion as a detransitivizer is primarily based on its form, as well as its paradigmatically conditioned occurrence in some languages, discussed shortly.
-Although the semantics of combining a detransitivizer with an intransitive verb do not really make sense, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
+While a root \rc{jəpɨ}, the two sound changes, and the optional addition of \rc{ət-} do account for the majority of the forms in \cref{tab:come},\footnote{
+	Apart from aforementioned \akuriyo \obj{eepɨ}, another diachronically irregular form is \apalai \obj{oepɨ}, where one would expect \rc{ət-epɨ} to yield \obj{os-epɨ} \parencite[506]{meira2010origin}.
+	Similarly, while \obj{oepɨ} would be a regular outcome of hypothetical \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
+	One possibility is that the form is due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}.
+	Alternatively, \apalai \obj{oepɨ} could be a fossilized loan from \wayana, which has replaced its reflex of \rc{ətjəpɨ}, but where intervocalic \rc{t} was also regularly lost \parencite[63]{wayanatavares2005}.}
+the distribution within the family is rather chaotic.
+In addition to the seemingly unordered distribution of \rc{əpɨ} and \rc{epɨ}, forms with and without \rc{ət-} can be found within the same language, usually conditioned by different prefixes.
+This was briefly discussed in \cref{sec:pekodian} for \arara (and \PPek) and in \cref{sec:taranoan} for \trio (and \PTar).
+To illustrate, the \trio \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.%
+\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}
+It should be noted that forms with and without \rc{ət-} in different languages are not triggered by the same person values.
+
+\ex<tricome> \trio \parencite[294]{triomeira1999}\\
+\begin{tabular}[t]{@{}ll@{}}
+	\gl{1} & \obj{w-əepɨ} \\
+	\gl{2} &  \obj{mən-epɨ} \\ 
+	\gl{1+2} &  \obj{ke-epɨ} \\
+	\gl{3} &  \obj{n-epɨ} \\
+\end{tabular}
+\xe
+
+Our interpretation of the \rc{ət} part as a detransitivizer is based on its form and on the paradigmatically conditioned occurrence in some languages.
+Although the combination of a detransitivizer an an intransitive verb seems semantically illogical, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
 For example, the \PC \gl{s_p_} verb \rc{wɨnɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
@@
 
-While a root \rc{jəpɨ}, aforementioned sound changes, and the addition of \rc{ət-} do account for the majority of the forms in \cref{tab:come},\footnote{
-Apart from aforementioned \akuriyo \obj{eepɨ}, another diachronically irregular form is \apalai \obj{oepɨ}, where the detransitivizer would have the reflex \obj{os-} \parencite[506]{meira2010origin}.
-While \obj{oepɨ} would be a regular outcome of \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
-The form may be due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}.
-Alternatively, \apalai \obj{oepɨ} could be a fossilized loan from \wayana, which has replaced its reflex of \rc{ətjəpɨ}, but where regular sound changes would also have resulted in the loss of the intervocalic \rc{t} \parencite[63]{wayanatavares2005}.}
-the distribution within the family is rather chaotic.
-Apart from the distribution of \rc{əpɨ} and \rc{epɨ}, forms with and without \rc{ət-} can be found within the same language, usually conditioned by different prefixes.
-This was briefly discussed in \cref{sec:pekodian} for \arara (and \PPek) and in \cref{sec:taranoan} for \trio (\PTar).
-To illustrate, the \trio \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}%
-\footnote{It should be noted that the paradigmatic distributions of forms with and without \rc{ət-} in different languages share no discernible patterns.}
-
-\ex<tricome> \trio \parencite[294]{triomeira1999}\\
-\begin{tabular}[t]{@{}ll@{}}
-\gl{1} & \obj{w-əepɨ} \\
-\gl{2} &  \obj{mən-epɨ} \\ 
-\gl{1+2} &  \obj{ke-epɨ} \\
-\gl{3} &  \obj{n-epɨ} \\
-\end{tabular}
-\xe
-
 Summing up, this verb is highly irregular, both from a synchronic and diachronic perspective.
-It seems that reflexes of the detransitivizer \rc{ət(e)-} were optionally added to an \gl{s_a_} verb root \rc{jəpɨ}, which further underwent umlaut and loss of \rc{j}, but in no systematic manner, resulting in the chaotic picture in \cref{tab:come}.
-
+The scenario we suggest involves reflexes of the detransitivizer \rc{ət(e)-} being optionally added to an \gl{s_a_} verb root \rc{jəpɨ}, which further underwent umlaut and loss of \rc{j}, but in no systematic manner, resulting in the chaotic picture in \cref{tab:come}.
 As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}, but not on the \arara reflex of \rc{jəpɨ}.
-Both reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ} and/or \rc{jəpɨ} (\akuriyo) resisted the introduction of \rc{t-} in \PTir.
+Reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ} and/or \rc{jəpɨ} (\akuriyo) resisted the introduction of \PTir \rc{t-}.
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.

### 2021-11-17 documents/floats/yukreg.tex
better overview table
md
'to go down'

@@
 \centering
-\caption[Regular \yukpa verbs]{Regular \yukpa verbs \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}}
+\caption[Regular \yukpa verbs]{Regular \yukpa verbs \parencites[139]{meira2006syntactic}[72, 76]{largo2011yukpa}}
 \label{tab:yukreg}
@@
 \mytoprule
-{} &  \qu{to fall} & \qu{to sleep} & \qu{to wash self} \\
+{} &  \qu{to fall} & \qu{to wash self} & \qu{to sleep} \\
 \midrule
-\gl{1} &  \obj{j-ata-} &  \obj{jɨ-nɨ-} &     \obj{j-otum-} \\
-\gl{2} &  \obj{m-ata-} &  \obj{mɨ-nɨ-} &     \obj{m-otum-} \\
-\gl{3} &  \obj{n-ata-} &  \obj{nɨ-nɨ-} &     \obj{n-otum-} \\
+\gl{1} &  \obj{j-ata-} &     \obj{j-otum-} &  \obj{jɨ-nɨ-} \\
+\gl{2} &  \obj{m-ata-} &     \obj{m-otum-} &  \obj{mɨ-nɨ-} \\
+\gl{3} &  \obj{n-ata-} &     \obj{n-otum-} &  \obj{nɨ-nɨ-} \\
 \bottomrule

### 2021-11-17 documents/sections/verbs.tex
'to go down'

@@
 When \akuriyo extended \obj{k-}, the verb already had an irregularly inflected first person form with \obj{p-}, inherited from \PTir.
-
-At first sight, it may seem that it also was affected by the independent extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
+At first sight, it may seem that it also was affected by the two independent extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
 
@@
 %
-However, a look at the comparative picture suggests a much more complicated situation.
-\cref{tab:godown} shows all attested forms, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.
+However, a broader comparative perspective reveals a much more complicated story.
+\cref{tab:godown} shows all attested cognates of \rc{ɨpɨtə}, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.
 It turns out that while a form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
-Its reflexes in languages that preserve the split-\gl{s} system are split fairly evenly between \gl{s_a_} and \gl{s_p_}.
+Its reflexes in languages that preserve the split-\gl{s} system are distributed fairly evenly between \gl{s_a_} and \gl{s_p_}.
 
@@
 Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the restriction of \rc{-ka} to \gl{s_p_} verbs in \PC \parencite{gildea2019overview} points to \gl{s_p_} membership.
-These patterns lead us to posit the hypothesis that the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class.
-This in turn implies that all reflexes of this verb with \gl{s_a_} status at some point switched from the \gl{s_p_} class.
+These patterns lead us to posit the hypothesis that the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class, reflected only in its \gl{1+2} prefix and the \gl{s_a_} class marker.
+This in turn implies that reflexes of this verb with \gl{s_a_} membership in other languages fully switched from \gl{s_p_} at some point.
 
@@
 Both causativized forms contain a reflex of the transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
-Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms \parencite[263]{triomeira1999}.
+Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms that also feature a reflex of \rc{en-} \parencite[263]{triomeira1999}.
 Thus, it seems that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.
@@
 These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by the extensions in \PWai, \PTir, \carijo, and \yukpa.
-For \PTir, we cannot establish a relative chronology of the verb class change, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
-For \PWai, we lack knowledge not only about the first person form, but even about class membership (in \hixka).
-For \carijo and \yukpa, we cannot know whether the verb potentially switched class before the breakdown of the entire split-\gl{s} system.
-While there is no language-internal evidence for such a switch, \qu{to go down} does have an inclination for class switches; in the case of \carijo, that could have already happened at the \PTar stage.
-In all four cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, resisting and keeping its old marker, but it is also possible that it was not even a potential target due to its \gl{s_p_} status.
-On the other hand, the class change must have happened before the split-up of \trio and \akuriyo, and therefore this verb resisted the extension of \akuriyo \obj{k-}.
-Likewise, it seems very likely that the class change took place before the extension of \PPek \rc{k-}.
+For \PTir, we cannot establish a relative chronology of the verb class switch, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
+For \PWai, we lack knowledge not only about the first person form, but even about class membership.
+Similarly, for \carijo and \yukpa we cannot know whether the verb potentially switched class before the breakdown of the entire split-\gl{s} system.
+While there is no language-internal evidence, it is clear that \qu{to go down} does have an inclination to switch classes; in the case of \carijo, that could have already happened at the \PTar stage.
+In all four cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, resisting it and preserving the old prefix, but it is also possible that it was not even a potential target due to its \gl{s_p_} status at the time.
+On the other hand, the class switch happened before the split-up of \trio and \akuriyo, and therefore this verb resisted the extension of \akuriyo \obj{k-} as an \gl{s_a_} verb.
+Likewise, it seems very likely that the class switch took place before the extension of \PPek \rc{k-}.
 Otherwise, the newly-turned-\gl{s_a_} verb would have taken on archaic and lexically heavily restricted \rc{w-}, either in \PTar, \PXin, or \arara.

### 2021-11-17 workflow/process_data.py

@@ -632,31 +632,31 @@ save_float(
     short=r"Reflexes of \rc{ɨpɨtə} \qu{to go down}",
 )
 
-print("\nClass membership of 'to defecate':")
-s_df = v_df[v_df["Parameter_ID"] == "defecate"]
-sources = extract_sources(s_df)
-s_df.drop(
-    columns=["Parameter_ID", "Cog_Cert", "Comment", "Cognateset_ID"], inplace=True
-)
-s_df["Form"] = s_df["Form"].str.replace("+", "", regex=True)
-sort_lg(s_df)
-print(s_df)
-s_df["Class"] = s_df.apply(
-    lambda x: pynt.get_expex_code(x["Class"])
-    if x["Class"] not in ["?", "–"]
-    else x["Class"],
-    axis=1,
-)
-add_obj_markdown(s_df)
-repl_lg_id(s_df)
-s_df.set_index("Language", drop=True, inplace=True)
-tabular = print_latex(s_df, keep_index=True)
-save_float(
-    tabular,
-    "defecate",
-    r"Reflexes of \qu{to defecate} " + sources,
-    short=r"Reflexes of \qu{to defecate} ",
-)
+# print("\nClass membership of 'to defecate':")
+# s_df = v_df[v_df["Parameter_ID"] == "defecate"]
+# sources = extract_sources(s_df)
+# s_df.drop(
+#     columns=["Parameter_ID", "Cog_Cert", "Comment", "Cognateset_ID"], inplace=True
+# )
+# s_df["Form"] = s_df["Form"].str.replace("+", "", regex=True)
+# sort_lg(s_df)
+# print(s_df)
+# s_df["Class"] = s_df.apply(
+#     lambda x: pynt.get_expex_code(x["Class"])
+#     if x["Class"] not in ["?", "–"]
+#     else x["Class"],
+#     axis=1,
+# )
+# add_obj_markdown(s_df)
+# repl_lg_id(s_df)
+# s_df.set_index("Language", drop=True, inplace=True)
+# tabular = print_latex(s_df, keep_index=True)
+# save_float(
+#     tabular,
+#     "defecate",
+#     r"Reflexes of \qu{to defecate} " + sources,
+#     short=r"Reflexes of \qu{to defecate} ",
+# )
 
 
 def print_aligned_table(

### 2021-11-17 documents/sections/verbs.tex
'to bathe'

@@
 \label{sec:bathe}
-This verb resisted extensions in \PPek \pcref{sec:pekodian} and \akuriyo \pcref{sec:akuriyo}, but is a regularly inflected \gl{s_a_} verb elsewhere.
-Verbs for intransitive \qu{to bathe} are usually detransitivized derivations of transitive roots in Cariban languages, which are reflexes of \rc{pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
+This verb resisted the extensions of \PPek \rc{k-} \pcref{sec:pekodian} and, trivially, \akuriyo \obj{k-} \pcref{sec:akuriyo}.
+It took on new \gl{1}\gl{s_a_} prefixes in \PTir \parentext{\trio \obj{s-epɨ-}, \akuriyo \obj{t͡ʃe+pɨ-} \parencites[292]{triomeira1999}[87]{gildea1994akuriyo}} and \PWai \parentext{\hixka \obj{k-ewehɨ-}, \waiwai \obj{k-ejeɸu-} \parencites[195]{hixkaryanaderby1985}[166]{waiwaihawkins1998}}.
+The fate of its \carijo reflex \obj{ehɨ} \parencites[72]{koch1908hiana} is unknown; we are not aware of a \yukpa cognate.
+
+Verbs for intransitive \qu{to bathe} are usually typical \gl{s_a_} verbs in Cariban languages, derived with a detransitivizer from a transitive root.
+These roots are reflexes of \rc{pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
 As we have shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
 Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch, changing \rc{e-} to \rc{i}.
-The reasons for this are entirely mysterious; we are not aware of \rc{i-} as a reflex of the detransitivizer in Pekodian, see also \textcite[506]{meira2010origin}.
-However, it should be noted that other languages also show unexpected developments in this verb, considering the apparent glide insertion in Waiwaian or the chaotic distribution of \rc{pɨ} and \rc{kupi} in Venezuelan languages.
+The reasons for this are entirely mysterious; reflexes of \rc{i-} are not attested as a reflex of the detransitivizer in Pekodian, see  \textcite[506]{meira2010origin}.
+However, it should be noted that other languages also show unexpected developments in this verb, considering the apparent glide insertion in Waiwaian or the distribution of \rc{pɨ} and \rc{kupi} in Venezuelan languages.
 

### 2021-11-17 documents/floats/pwaireg.tex
hacky fix to properly align ye'kwana 'to bathe'

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[30]{waiwaihawkins1998}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-11-17 workflow/process_data.py

@@ -781,11 +781,15 @@ for table in bathe_tables:
     if table[2] == "bathe_intr_1":
         bathe_out += r"""\begin{subtable}[t]{.49\linewidth}
 \centering
-"""
-    if table[2] == "bathe_tr_1":
+"""        
+    elif table[2] == "bathe_tr_1":
         bathe_out += r"""\end{subtable}
 \begin{subtable}[t]{.49\linewidth}
 \centering"""
+    # hacky fix for aligning ye'kwana ʔ with k rather than akawaio ʔ
+    elif table[2] == "bathe_intr_2":
+        tabular = tabular.replace("&    &    &  ʔ", "&  ʔ &    &   ")
+
     bathe_out += r"""\caption{%s}
 \label{tab:%s}
 %s""" % (

### 2021-11-17 documents/sections/verbs.tex
hacky fix to properly align ye'kwana 'to bathe'

@@
 It took on new \gl{1}\gl{s_a_} prefixes in \PTir \parentext{\trio \obj{s-epɨ-}, \akuriyo \obj{t͡ʃe+pɨ-} \parencites[292]{triomeira1999}[87]{gildea1994akuriyo}} and \PWai \parentext{\hixka \obj{k-ewehɨ-}, \waiwai \obj{k-ejeɸu-} \parencites[195]{hixkaryanaderby1985}[166]{waiwaihawkins1998}}.
-The fate of its \carijo reflex \obj{ehɨ} \parencites[72]{koch1908hiana} is unknown; we are not aware of a \yukpa cognate.
+The first person form of its \carijo reflex \obj{ehɨ} \parencites[72]{koch1908hiana} is unknown; we are not aware of a \yukpa cognate.
 
@@
 Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch, changing \rc{e-} to \rc{i}.
-The reasons for this are entirely mysterious; reflexes of \rc{i-} are not attested as a reflex of the detransitivizer in Pekodian, see  \textcite[506]{meira2010origin}.
+This is an irregular development, since reflexes of \rc{i-} are not attested as a reflex of the detransitivizer in Pekodian \parencite[506]{meira2010origin}; its cause is unknown.
 However, it should be noted that other languages also show unexpected developments in this verb, considering the apparent glide insertion in Waiwaian or the distribution of \rc{pɨ} and \rc{kupi} in Venezuelan languages.
@@
 \input{floats/bathe}
-
-%\subsection{\trio \obj{weka}/\obj{oeka} \qu{to defecate}}
-%\label{sec:shit}
-%The comparative picture for the resistant \trio \gl{s_a_} verb \qu{to defecate} \pcref{tab:defecate} shows that \rc{weka} was clearly originally an \gl{s_p_} verb.
-%As in the case of \qu{to go down} \pcref{sec:godown}, this verb has a mixed \gl{s_a_}/\gl{s_p_} status in \wayana.
-%The \bakairi reflex \obj{əeke} seems to contain a reflex of \rc{ət-}, which would explain its status as an \gl{s_a_} verb.
-%While the class membership of \panare \obj{aiʔka} and \obj{iʔka} \qu{to defecate} is unknown, the former seems to contain a reflex of the detransitivizer, while the latter only consists of the root, meaning that one would assume \gl{s_a_} and \gl{s_p_} membership, respectively.
-%
-%\input{floats/defecate}
-%
-%\subsection{The \akuriyo movement verbs}
-%\label{sec:movement}
-%As discussed in \cref{sec:akuriyo}, four \akuriyo \gl{s_a_} verbs are attested as not having the first person marker \obj{t͡ʃ-} in \textcite{gildea1994akuriyo}; instead, they have \obj{w-} or its phonologically reduced form ∅.
-%These verbs share two properties: they are all \obj{e-}initial, and they are all movement verbs.
-%However, the same dataset also contains \obj{e-}initial movement verbs with a first person marker \obj{t͡ʃ-} \exref{aku-106}.
-%As for the etymology of these four verbs, we can only speak to \obj{erama} \qu{to return}, which is a detransitivized form of \obj{rama} \qu{to put back}, a pair which is also found in other Cariban languages.
-%For the other three verbs, we have not encountered potential cognates, but given their \obj{e-}initial nature and their semantics, it seems plausible that they, too, are derived from transitive verbs with meanings like \qu{to put up} etc.
-%In any case, \akuriyo is only sparsely documented, and there are many open questions about it; the \akuriyo data in \textcite{gildea1994akuriyo} are not fully reliable.

### 2021-11-17 workflow/process_data.py
started revision of discussion

@@ -853,7 +853,7 @@ overview_legend = {
     "n": {"tex": "×", "legend": "not affected"},
     "?": {"tex": "?", "legend": "unknown first person prefix"},
     "–": {"tex": "–", "legend": "does not occur"},
-    "(y)": {"tex": "(\\checkmark)", "legend": "affected with surviving old marker"},
+    "(y)": {"tex": "(\\checkmark)", "legend": "old and new marker combined"},
     "N/A": {"tex": "\\textsc{n/a}", "legend": "not meaningfully answerable"},
 }
 

### 2021-11-20 workflow/process_data.py
more discussion and apalai Sa verb count

@@ -1147,5 +1147,20 @@ save_float(
 )
 pyd.content_string = "Form"
 
+
+# preliminary Sa verb frequency counts from apalai
+apalai_data = pd.read_csv("../data/apalai_sa_verb_stats.csv")
+apalai_data["String"] = apalai_data.apply(combine_form_meaning, axis=1)
+
+export_csv(
+    apalai_data,
+    "apalaicounts",
+    "Frequency counts of Sa verbs from two Apalai texts",
+    keep_index=False,
+    sources="",
+)
+
+print(apalai_data)
+
 with open("data_output/metadata.json", "w") as outfile:
     json.dump(exported_tables, outfile, indent=4)

### 2021-11-20 documents/floats/pwaireg.tex
no latex in csv output

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-11-20 workflow/process_data.py

@@ -1150,7 +1150,6 @@ pyd.content_string = "Form"
 
 # preliminary Sa verb frequency counts from apalai
 apalai_data = pd.read_csv("../data/apalai_sa_verb_stats.csv")
-apalai_data["String"] = apalai_data.apply(combine_form_meaning, axis=1)
 
 export_csv(
     apalai_data,
@@ -1159,6 +1158,9 @@ export_csv(
     keep_index=False,
     sources="",
 )
+
+apalai_data["String"] = apalai_data.apply(combine_form_meaning, axis=1)
+
 
 print(apalai_data)
 

### 2021-11-22 workflow/compile_cldf_irregular.py
apalai verb counts

@@ -280,10 +280,11 @@ for filename in os.listdir(data_path):
     if filename in ["extensions.csv", "cognate_sets.csv", "languages.csv"]:
         continue
     df = pd.read_csv(os.path.join(data_path, filename))
-    for i in df["Source"]:
-        if pd.isnull(i):
-            continue
-        found_refs.append(cldf_helpers.split_ref(i)[0])
+    if "Source" in df.columns:
+        for i in df["Source"]:
+            if pd.isnull(i):
+                continue
+            found_refs.append(cldf_helpers.split_ref(i)[0])
 found_refs = list(set(found_refs))
 found_refs.remove("pc")
 

### 2021-11-22 workflow/process_data.py

@@ -63,8 +63,10 @@ cog_trans_dic["bathe_intr"] = r"bathe"
 cog_trans_dic["be_1"] = r"be-1"
 cog_trans_dic["be_2"] = r"be-2"
 
+
 def print_shorthand(abbrev):
     return "\\" + cah.get_shorthand(abbrev)
+
 
 def str2numcog(cogsets):
     return " ".join([cognum[x] for x in cogsets.split("+")])
@@ -82,6 +84,7 @@ def segmentify(form):
 
 
 exported_tables = {}
+
 
 def export_csv(
     tabular,
@@ -114,13 +117,15 @@ def export_csv(
 
 
 # functions for creating latex tables
-def print_latex(df, ex=False, keep_index=False):
+def print_latex(df, ex=False, keep_index=False, formatters=None):
     if keep_index:
         df.columns.name = df.index.name
         df.index.name = None
 
     with pd.option_context("max_colwidth", 1000):
-        lines = df.to_latex(escape=False, index=keep_index).split("\n")
+        lines = df.to_latex(
+            escape=False, index=keep_index, formatters=formatters
+        ).split("\n")
     lines[0] = lines[0].replace("tabular}{l", "tabular}[t]{@{}l").replace("l}", "l@{}}")
     if ex:
         del lines[1:4]
@@ -184,7 +189,7 @@ def combine_form_meaning(row, latex=True):
     else:
         form = f"""*{row["Form"]}*"""
         meaning = f"""'{row["Meaning"]}'"""
-    out = form + " " + meaning
+    out = form + " " + meaning.replace("_", " ")
     return out
 
 
@@ -199,9 +204,11 @@ def get_sources(df, parens=True, latexify=True):
     source_string = cldfh.cite_a_bunch(ref_list, parens=parens)
     return source_string
 
+
 def upper_repl(match):
-     return match.group(1).upper()
-     
+    return match.group(1).upper()
+
+
 def repl_latex(string):
     string = re.sub(r"\\rc\{(.*?)\}", r"\**\1*", string)
     string = re.sub(r"\\gl\{(.*?)\}", upper_repl, string)
@@ -209,10 +216,12 @@ def repl_latex(string):
     string = re.sub(r"\\qu\{(.*?)\}", r"'\1'", string)
     return string
 
+
 def sort_lg(df):
     df.Language_ID = df.Language_ID.astype("category")
[…]

### 2021-11-23 documents/floats/ppekverbs.tex
regular verbs in data and docs
started cleaning up workflow

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[76, 130, 131, 137, 374, 397]{von1892bakairi}[68]{ikpengpacheco1997}[4; p.c., Angela Fabíola Alves Chagas]{meira2003bakairi}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[76, 130, 131, 137, 374, 397]{von1892bakairi}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[4; p.c., Angela Fabíola Alves Chagas]{meira2003bakairi}}
 \label{tab:ppekverbs}
 \begin{tabular}[t]{@{}lllll@{}}
-\toprule
+\mytoprule
 {} &          \PPek &          \arara &                       \ikpeng &        \bakairi \\
 \midrule
-\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &     \obj{u-ge-} \\
-\qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &      \obj{w-i-} \\
 \qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &      \obj{w-a-} \\
 \qu{be-2}    &  \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &      \obj{w-i-} \\
+\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &     \obj{u-ge-} \\
+\qu{go}      &   \rc{w-ɨtən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &     \obj{u-tə-} \\
 \qu{come}    &    \rc{w-epɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &   \obj{k-əewɨ-} \\
 \qu{go down} &   \rc{w-ɨptə-} &  \obj{w-iptoŋ-} &  \obj{{\normalfont ?}-iptoŋ-} &  \obj{k-ɨtəgɨ-} \\
-\qu{go}      &   \rc{w-ɨtən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &     \obj{u-tə-} \\
+\qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &      \obj{w-i-} \\
 \bottomrule

### 2021-11-23 documents/floats/ptirverbs.tex
regular verbs in data and docs
started cleaning up workflow

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112, 113, 114, 115, 165]{meira1998proto}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[112, 113, 114, 115, 165]{meira1998proto}[292, 294, 339]{triomeira1999}}
 \label{tab:ptirverbs}
 \begin{tabular}[t]{@{}llll@{}}
-\toprule
+\mytoprule
 {} &          \PTir &          \trio &                     \akuriyo \\
 \midrule
-\qu{go}   &  \rc{wɨ-təmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
-\qu{say}  &    \rc{wɨ-ka-} &   \obj{wɨ-ka-} &                 \obj{wɨ-ka-} \\
-\qu{come} &  \rc{w-əʔepɨ-} &  \obj{w-əepɨ-} &  \obj{{\normalfont ∅}-eepɨ-} \\
 \qu{be-1} &      \rc{w-a-} &     \obj{w-a-} &     \obj{{\normalfont ∅}-a-} \\
 \qu{be-2} &    \rc{w-eʔi-} &    \obj{w-ei-} &   \obj{{\normalfont ?}-eʔi-} \\
+\qu{say}  &    \rc{wɨ-ka-} &   \obj{wɨ-ka-} &                 \obj{wɨ-ka-} \\
+\qu{go}   &  \rc{wɨ-təmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
+\qu{come} &  \rc{w-əʔepɨ-} &  \obj{w-əepɨ-} &  \obj{{\normalfont ∅}-eepɨ-} \\
 \bottomrule

### 2021-11-23 documents/floats/pwaireg.tex
started cleaning up workflow

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[30]{waiwaihawkins1998}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-11-23 documents/floats/pwaiverbs.tex
fixed proto-pekodian 'to go'
regular verbs in data and docs
started cleaning up workflow

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197, 198; p.c., Spike Gildea]{hixkaryanaderby1985}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[70, 197, 198]{hixkaryanaderby1985}[4]{hixkaryanaderby1979}[71, 85; p.c., Spike Gildea]{waiwaihawkins1998}}
 \label{tab:pwaiverbs}
 \begin{tabular}[t]{@{}llll@{}}
-\toprule
+\mytoprule
 {} &         \PWai &        \hixka &         \waiwai \\
 \midrule
-\qu{say}  &  \rc{wɨ-ka-} &   \obj{ɨ-ka-} &    \obj{wɨ-ka-} \\
 \qu{be-1} &    \rc{w-ah-} &   \obj{w-ah-} &      \obj{w-a-} \\
 \qu{be-2} &   \rc{w-eʃi-} &  \obj{w-eʃe-} &   \obj{w-eeʃi-} \\
-\qu{go}   &  \rc{wɨ-to-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
+\qu{say}  &   \rc{wɨ-ka-} &   \obj{ɨ-ka-} &    \obj{wɨ-ka-} \\
+\qu{go}   &  \rc{wɨ-tom-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
 \bottomrule

### 2021-11-23 workflow/helper_functions.py

@@ -1,0 +1,352 @@
+from read_data import *
+import re
+import pyradigms as pyd
+import cldf_helpers as cldfh
+import warnings
+import json
+import numpy as np
+from lingpy_alignments import calculate_alignment
+
+warnings.filterwarnings("ignore")
+
+def print_shorthand(abbrev):
+    return "\\" + cah.get_shorthand(abbrev)
+
+shorthand_dic = {x: print_shorthand(x) for x in lg_list}
+
+def str2numcog(cogsets):
+    return " ".join([cognum[x] for x in cogsets.split("+")])
+
+
+def num2strcog(cogids):
+    return "+".join([numcog[i] for i in cogids.split(" ")])
+
+
+def segmentify(form):
+    form = re.sub("[()\[\]]", "", form)
+    form = form.replace("-", "+")
+    form = form.strip("+")
+    return t(form)
+
+
+# functions for creating latex tables
+def print_latex(df, ex=False, keep_index=False, formatters=None):
+    if keep_index:
+        df.columns.name = df.index.name
+        df.index.name = None
+
+    with pd.option_context("max_colwidth", 1000):
+        lines = df.to_latex(
+            escape=False, index=keep_index, formatters=formatters
+        ).split("\n")
+    lines[0] = lines[0].replace("tabular}{l", "tabular}[t]{@{}l").replace("l}", "l@{}}")
+    if ex:
+        del lines[1:4]
+        del lines[-1]
+        del lines[-2]
+        return "\n".join(lines)
+    else:
+        return "\n".join(lines).replace(r"\toprule", "\mytoprule").replace("%", "\\%")
+
+
+def save_float(tabular, label, caption, filename=None, short=None):
+    if not filename:
+        filename = label
+    if not short:
+        shorttext = ""
+    else:
+        shorttext = f"[{short}]"
+    output = f"""\\begin{{table}}
+\centering
+\caption{shorttext}{{{caption}}}
+\label{{tab:{label}}}
+{tabular}\end{{table}}"""
+    f = open(f"../documents/floats/{filename}.tex", "w")
+    f.write(output)
+    f.close()
+
+
+normalfonties = ["Ø", "?", "V", "G", "C", "∅"]
+
+
+def objectify(string, obj_string="obj"):
+    if string in ["", "?", " ", "  ", "–"]:
+        return string
+    if string is None:
+        return ""
+    if string[0] == " ":
+        return string
+    output = []
+    entry_list = string.split(" OR ")
+    for entry in entry_list:
+        out_list = []
[…]

### 2021-11-23 workflow/process_data.py

@@ -1,278 +1,7 @@
-import pyradigms as pyd
-import warnings
-import json
-
-warnings.filterwarnings("ignore")
-import pandas as pd
-
-pd.options.mode.chained_assignment = None
-import cariban_helpers as cah
-import cldf_helpers as cldfh
-import pynterlinear as pynt
-import numpy as np
-import re
-from segments import Tokenizer, Profile
-from lingpy_alignments import calculate_alignment
-
-lg_list = list(cah.lg_order().keys())
-cognate_list = ["go", "say", "come", "be_1", "be_2", "go_down", "bathe_intr"]
-cs_df = pd.read_csv("../data/cognate_sets.csv")
-v_df = pd.read_csv("../data/verb_stem_data.csv")
-v_df.rename(columns={"Meaning_ID": "Parameter_ID"}, inplace=True)
-e_df = pd.read_csv("../data/extensions.csv")
-i_df = pd.read_csv("../data/inflection_data.csv")
-i_df["Form"] = i_df["Form"].str.replace("+", "", regex=True)
-
-m_df = pd.read_csv("../data/cldf/parameters.csv")
-m_df["Description"] = m_df["Description"].str.strip("'").map(pynt.get_expex_code)
-mean_dic = dict(zip(m_df["ID"], m_df["Description"]))
-trans_dic = {x: "\\qu{%s}" % y for x, y in mean_dic.items()}
-
-concept_dic = {
-    "DETRZ1+bathe_1": "bathe_intr",
-    "DETRZ+come": "come",
-    "DETRZ1": "bathe_intr",
-}
-
-i_df["Concept"] = i_df.apply(
-    lambda x: concept_dic[x["Verb_Cognateset_ID"]]
-    if x["Verb_Cognateset_ID"] in concept_dic
-    else x["Verb_Cognateset_ID"],
-    axis=1,
-)
-
-segments = open("../data/segments.txt").read()
-segment_list = [{"Grapheme": x, "mapping": x} for x in segments.split("\n")]
-t = Tokenizer(Profile(*segment_list))
-cognatesets = pd.read_csv("../data/cldf/cognatesets.csv")
-cognatesets = cognatesets.append(
-    {"ID": "?", "Description": "Unknown material"}, ignore_index=True
-)
-
-# lingpy does not like cogid 0
-cognatesets.index = cognatesets.index + 1
-cognum = dict(zip(cognatesets["ID"], cognatesets.index.astype(str)))
[…]

### 2021-11-23 workflow/read_data.py

@@ -1,0 +1,61 @@
+import cariban_helpers as cah
+import pandas as pd
+pd.options.mode.chained_assignment = None
+import pynterlinear as pynt
+from segments import Tokenizer, Profile
+
+lg_list = list(cah.lg_order().keys())
+
+cognate_list = ["go", "say", "come", "be_1", "be_2", "go_down", "bathe_intr"]
+cs_df = pd.read_csv("../data/cognate_sets.csv")
+v_df = pd.read_csv("../data/verb_stem_data.csv")
+v_df.rename(columns={"Meaning_ID": "Parameter_ID"}, inplace=True)
+e_df = pd.read_csv("../data/extensions.csv")
+i_df = pd.read_csv("../data/inflection_data.csv")
+i_df["Form"] = i_df["Form"].str.replace("+", "", regex=True)
+
+m_df = pd.read_csv("../data/cldf/parameters.csv")
+m_df["Description"] = m_df["Description"].str.strip("'").map(pynt.get_expex_code)
+mean_dic = dict(zip(m_df["ID"], m_df["Description"]))
+
+name_dic = {x: cah.get_name(x) for x in lg_list}
+
+ext_lg_dic = {x: y for x, y in dict(zip(e_df["ID"], e_df["Language_ID"])).items()}
+ext_form_dic = {x: y for x, y in dict(zip(e_df["ID"], e_df["Form"])).items()}
+
+trans_dic = {x: "\\qu{%s}" % y for x, y in mean_dic.items()}
+
+concept_dic = {
+    "DETRZ1+bathe_1": "bathe_intr",
+    "DETRZ+come": "come",
+    "DETRZ1": "bathe_intr",
+}
+
+i_df["Concept"] = i_df.apply(
+    lambda x: concept_dic[x["Verb_Cognateset_ID"]]
+    if x["Verb_Cognateset_ID"] in concept_dic
+    else x["Verb_Cognateset_ID"],
+    axis=1,
+)
+
+segments = open("../data/segments.txt").read()
+segment_list = [{"Grapheme": x, "mapping": x} for x in segments.split("\n")]
+t = Tokenizer(Profile(*segment_list))
+cognatesets = pd.read_csv("../data/cldf/cognatesets.csv")
+cognatesets = cognatesets.append(
+    {"ID": "?", "Description": "Unknown material"}, ignore_index=True
+)
+
[…]

### 2021-11-24 workflow/helper_functions.py
more tables in readme

@@ -128,6 +128,13 @@ def repl_latex(string):
     string = re.sub(r"\\obj\{(.*?)\}", r"*\1*", string)
     string = re.sub(r"\\qu\{(.*?)\}", r"'\1'", string)
     return string
+    
+def delatexify(string):
+    string = re.sub(r"\\rc\{(.*?)\}", r"\1", string)
+    string = re.sub(r"\\gl\{(.*?)\}", upper_repl, string)
+    string = re.sub(r"\\obj\{(.*?)\}", r"\1", string)
+    string = re.sub(r"\\qu\{(.*?)\}", r"'\1'", string)
+    return string
 
 
 def sort_lg(df):

### 2021-11-24 workflow/process_data.py

@@ -217,8 +217,9 @@ save_float(
     short=r"Regular \akuriyo \gl{1}\gl{s_a_} markers",
 )
 
-aku_verbs = aku_verbs.applymap(repl_latex)
+aku_verbs = aku_verbs.applymap(delatexify)
 aku_verbs.columns = map(repl_latex, aku_verbs.columns)
+
 export_csv(
     aku_verbs,
     label,
@@ -226,40 +227,40 @@ export_csv(
     sources=raw_sources,
 )
 
-#
-# # regular carijo and yukpa verbs
-# pyd.x = ["Meaning_ID"]
-# pyd.y = ["Inflection"]
-# pyd.y_sort = ["1", "2", "1+2", "3"]
-# for lg, meanings in {
-#     "car": ["arrive", "dance"],
-#     "yuk": ["wash_self", "sleep", "fall"],
-# }.items():
-#     pyd.filters = {"Language_ID": [lg], "Meaning_ID": meanings}
-#     tabular = pyd.compose_paradigm(i_df, multi_index=False)
-#     label = lg + "reg"
-#     export_csv(
-#         tabular,
-#         label,
-#         f"Regular {name_dic[lg]} verbs",
-#         keep_index=True,
-#         sources=get_sources(i_df, latexify=False),
-#     )
-#     sources = get_sources(i_df)
-#     tabular.index = tabular.index.map(pynt.get_expex_code)
-#     tabular.index.name = None
-#     tabular.rename(columns=trans_dic, inplace=True)
-#     tabular = tabular.apply(lambda x: x.apply(objectify))
-#     tabular.columns.names = [None]
-#
-#     save_float(
-#         print_latex(tabular, keep_index=True),
-#         label,
-#         f"Regular {print_shorthand(lg)} verbs " + sources,
-#         short=f"Regular {print_shorthand(lg)} verbs",
-#     )
-#
-# i_df = i_df[~(pd.isnull(i_df["Verb_Cognateset_ID"]))]
+
+# regular carijo and yukpa verbs
+pyd.x = ["Meaning_ID"]
+pyd.y = ["Inflection"]
+pyd.y_sort = person
+for lg, meanings in {
+    "car": ["arrive", "dance"],
+    "yuk": ["wash_self", "sleep", "fall"],
+}.items():
+    pyd.filters = {"Language_ID": [lg], "Meaning_ID": meanings}
+    tabular = pyd.compose_paradigm(i_df, multi_index=False)
+    label = lg + "reg"
+    export_csv(
+        tabular,
[…]

### 2021-11-24 documents/floats/pwaireg.tex
more tables

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[30]{waiwaihawkins1998}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-11-24 workflow/process_data.py

@@ -261,355 +261,356 @@ for lg, meanings in {
     )
 
 i_df = i_df[~(pd.isnull(i_df["Verb_Cognateset_ID"]))]
-#
-# # pekodian lexical comparison
-# forms = pd.read_csv("../data/cldf/forms.csv")
-# lgs = ["ara", "ikp"]
-# meanings = [
-#     "defecate",
-#     "DAT",
-#     "dog",
-#     "capuchin_monkey",
-#     "sleep",
-# ]
-# pyd.x = ["Language_ID"]
-# pyd.y = ["Parameter_ID"]
-# pyd.filters = {"Language_ID": lgs, "Parameter_ID": meanings}
-# pyd.x_sort = lgs
-# pyd.y_sort = meanings
-# tabular = pyd.compose_paradigm(forms)
-# label = "pxinw"
-# export_csv(
-#     tabular,
-#     label,
-#     "Loss of \**w* in Ikpeng",
-#     keep_index=True,
-#     sources=get_sources(forms, latexify=False),
+
+# pekodian lexical comparison
+forms = pd.read_csv("../data/cldf/forms.csv")
+lgs = ["ara", "ikp"]
+meanings = [
+    "defecate",
+    "DAT",
+    "dog",
+    "capuchin_monkey",
+    "sleep",
+]
+pyd.x = ["Language_ID"]
+pyd.y = ["Parameter_ID"]
+pyd.filters = {"Language_ID": lgs, "Parameter_ID": meanings}
+pyd.x_sort = lgs
+pyd.y_sort = meanings
+tabular = pyd.compose_paradigm(forms)
+label = "pxinw"
+export_csv(
+    tabular,
+    label,
+    "Loss of \**w* in Ikpeng",
+    keep_index=True,
+    sources=get_sources(forms, latexify=False),
+)
+sources = get_sources(forms)
+tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name)))
+tabular.columns = tabular.columns.map(print_shorthand)
+
+tabular.index.name = "Meaning"
+tabular.reset_index(inplace=True)
+tabular["Meaning"] = tabular["Meaning"].map(mean_dic)
+tabular["Meaning"] = tabular["Meaning"].map(lambda x: "\\qu{" + x + "}")
+
+save_float(
+    print_latex(tabular),
+    label,
+    r"Loss of \rc{w} in \ikpeng " + sources,
[…]

### 2021-11-24 workflow/helper_functions.py
added verb stems previously only in inflection data

@@ -178,7 +178,16 @@ def repl_lg_id(df):
     df.rename(columns={"Language_ID": "Language"}, inplace=True)
 
 
+def get_verb_citation(v_id, l_id, latex=True):
+    res = v_df[(v_df["Cognateset_ID"] == v_id) & (v_df["Language_ID"] == l_id)]
 
+    if len(res) == 0:
+        raise ValueError(f"No verb entry for {l_id} {v_id}")
+    res = res.iloc[0]
+    if latex:
+        return f"""\\{get_obj_str(l_id)}{{{res["Form"]}}} \\qu{{{mean_dic[res["Parameter_ID"]]}}}"""
+    else:
+        return f"""{res["Form"]} '{mean_dic[res["Parameter_ID"]]}'"""
 
 def extension_string(id, latex=True):
     form = objectify(ext_form_dic[id], obj_string=get_obj_str(ext_lg_dic[id]))

### 2021-11-24 workflow/process_data.py

@@ -689,9 +689,12 @@ for lg in ["PWai", "PPek", "PTir", "aku", "car", "yuk"]:
         explanations[row["ID"]] = test_explanations(row.to_dict())
     r_df = pd.DataFrame(explanations)
     r_df = r_df.applymap(int)
+    r_df.columns = r_df.columns.map(lambda x: get_verb_citation(x, lg))
     r_df["Score"] = r_df.apply(sum, axis=1) / len(r_df.columns)
     r_df.sort_values(by="Score", inplace=True, ascending=False)
     label = f"{lg.lower()}_predictions"
+
+    print(r_df)
     export_csv(
         r_df,
         label,
@@ -699,7 +702,6 @@ for lg in ["PWai", "PPek", "PTir", "aku", "car", "yuk"]:
         keep_index=True,
         print_i_name=True,
     )
-    # r_df.to_csv(f"{lg}_results.csv")
     print(r_df)
 
 print(get_frequency_prediction("go", frequencies))
@@ -860,7 +862,6 @@ print(get_frequency_prediction("go", frequencies))
 #     short=r"Frequency counts of \gl{s_a_} verbs in \apalai",
 # )
 #
-for t in exported_tables.keys():
-    print(f"\\input{{floats/{t}.tex}}")
+
 with open("data_output/metadata.json", "w") as outfile:
     json.dump(exported_tables, outfile, indent=4)

### 2021-11-26 workflow/helper_functions.py
new approach to testing predictions from bybee's model

@@ -30,14 +30,14 @@ def segmentify(form):
 
 
 # functions for creating latex tables
-def print_latex(df, ex=False, keep_index=False, formatters=None):
+def print_latex(df, ex=False, keep_index=False, formatters=None, multicolumn=False, float_format=None):
     if keep_index:
         df.columns.name = df.index.name
         df.index.name = None
 
     with pd.option_context("max_colwidth", 1000):
         lines = df.to_latex(
-            escape=False, index=keep_index, formatters=formatters
+            escape=False, index=keep_index, formatters=formatters, multicolumn=multicolumn,float_format=float_format
         ).split("\n")
     lines[0] = lines[0].replace("tabular}{l", "tabular}[t]{@{}l").replace("l}", "l@{}}")
     if ex:
@@ -46,7 +46,7 @@ def print_latex(df, ex=False, keep_index=False, formatters=None):
         del lines[-2]
         return "\n".join(lines)
     else:
-        return "\n".join(lines).replace(r"\toprule", "\mytoprule").replace("%", "\\%")
+        return "\n".join(lines).replace(r"\toprule", "\mytoprule").replace(r"\midrule", "\mymidrule").replace(r"\bottomrule", "\mybottomrule").replace("%", "\\%")
 
 
 def save_float(tabular, label, caption, filename=None, short=None):
@@ -78,12 +78,14 @@ def objectify(string, obj_string="obj"):
         return string
     output = []
     entry_list = string.split(" OR ")
+    unit_count = 0
     for entry in entry_list:
         out_list = []
         morph_list = entry.split("; ")
         for i, morph in enumerate(morph_list):
+            unit_count += 1
             out_list.append(r"\%s{%s}" % (obj_string, morph))
-            if i > 0:
+            if unit_count > 0:
                 obj_string = "obj"
         output.append("/".join(out_list))
     for i, cell in enumerate(output):
@@ -121,12 +123,21 @@ def get_sources(df, parens=True, latexify=True):
 def upper_repl(match):
     return match.group(1).upper()
 
+web_checkmarks = {
+    1: "✓",
+    0: "✖",
+    True: "✓",
+    False: "✖",
+    "\\checkmark": "✓",
+    "×": "✖"
+}
 
[…]

### 2021-11-26 workflow/process_data.py

@@ -603,11 +603,12 @@ def predict_semantics(form):
     pass
 
 
-def predict_morphology(form):
+def predict_detrz(form):
     if "DETRZ" in form["ID"] and form["Form"][0] != "i":
-        return True
+        prediction = True
     else:
-        return False
+        prediction = False
+    return "\\gl{detrz}", None, prediction
 
 
 phon_preds = {
@@ -622,56 +623,97 @@ phon_preds = {
 
 def predict_phonology(form):
     if form["Form"][0] in phon_preds[form["Language_ID"]]:
-        return True
+        prediction = True
     else:
-        return False
-    return False
+        prediction = False
+    if len(phon_preds[form["Language_ID"]]) > 4:
+        detail = "\\envr{}{V}"
+    else:
+        obj = objectify(" OR ".join(phon_preds[form["Language_ID"]]), get_obj_str(form['Language_ID']))
+        detail = f"""\\envr{{}}{{{obj}}}"""
+    return "phono", detail , prediction
+    
+def predict_inflection(form):
+    #akuriyo is the only case where the previous first person inflection was not *w-
+    if form["Language_ID"] == "aku":
+        detail = "\\obj{k-}"
+        # this assumes t-ə as distinct from t͡ʃ-e
+        if form["Form"][0] == "ə":
+            prediction = True
+        else:
+            prediction = False
+    else:
+        detail = "\\rc{w-}"
+        prediction = True
+    return "infl", detail, prediction
+
+factors = [predict_detrz, predict_phonology, predict_inflection]
+
+# preliminary Sa verb frequency counts from apalai
+apalai_data = pd.read_csv("../data/apalai_sa_verb_stats.csv")
+frequencies = dict(zip(apalai_data["ID"], apalai_data["High_Frequency"]))
+
+label = "apalaicounts"
+
+export_csv(
+    apalai_data,
+    label,
+    "Frequency counts of Sa verbs in two Apalai texts",
+    keep_index=False,
+    sources="",
+)
+
+apalai_data["Verb"] = apalai_data.apply(combine_form_meaning, axis=1)
+apalai_data.drop(columns=["Form", "Meaning"], inplace=True)
+
+apalai_data.rename(
+    columns={"% Sa": "% \gl{s_a_} verb tokens", "% Words": "% word tokens"},
[…]

### 2021-11-27 documents/floats/pwaireg.tex
incorporated new results into discussion

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[30]{waiwaihawkins1998}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}
@@
 \mytoprule
-{} & \multicolumn{2}{l}{\PWai} & \multicolumn{2}{l}{\hixka} & \multicolumn{2}{l}{\waiwai} \\
+{} &           \PWai &                   &           \hixka &                 &            \waiwai &                    \\
 {} &    \qu{to fall} &     \qu{to sleep} &     \qu{to fall} &   \qu{to sleep} &       \qu{to fall} &      \qu{to sleep} \\
-\midrule
+\mymidrule
 \gl{1}   &  \rc{k-eɸurka-} &   \rc{kɨ-wɨnɨkɨ-} &  \obj{k-ehurka-} &  \obj{kɨ-nɨkɨ-} &    \obj{k-eɸɨrka-} &   \obj{kɨ-wɨnɨkɨ-} \\
@@
 \gl{3}   &  \rc{ɲ-eɸurka-} &   \rc{nɨ-wɨnɨkɨ-} &  \obj{ɲ-ehurka-} &  \obj{nɨ-nɨkɨ-} &    \obj{ɲ-eɸɨrka-} &   \obj{nɨ-wɨnɨkɨ-} \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2021-11-27 workflow/helper_functions.py

@@ -192,8 +192,10 @@ def repl_lg_id(df):
 def get_verb_citation(v_id, l_id, latex=True, as_tuple=False):
     res = v_df[(v_df["Cognateset_ID"] == v_id) & (v_df["Language_ID"] == l_id)]
     if len(res) == 0:
-        raise ValueError(f"No verb entry for {l_id} {v_id}")
-    res = res.iloc[0]
+        res = {"Form": "əturu", "Parameter_ID": "talk"}
+        # raise ValueError(f"No verb entry for {l_id} {v_id}")
+    else:
+        res = res.iloc[0]
     form = res["Form"].replace("+", "")
     if latex:
         obj, gloss = f"""\\{get_obj_str(l_id)}{{{form}}}""", f"""\\qu{{{mean_dic[res["Parameter_ID"]].replace("to ", "")}}}"""

### 2021-11-27 workflow/process_data.py

@@ -713,13 +713,12 @@ def test_explanations(word_form):
 affectedness = overview[overview["Language_ID"] == overview["Orig_Language"]]
 affectedness = affectedness[affectedness["Affected"].isin(["y", "n"])]
 affectedness = affectedness[affectedness["Concept"].isin(verb_list)]
-affectedness.to_csv("patterns.csv", index=False)
 affectedness["Form"] = affectedness["Form"].apply(lambda x: x.split("-")[1])
 
 # # add fake Sa verbs
 # for lg, conds in phon_preds.items():
-#     for i in range(0, 100):
-#         df = df.append({"Form": conds[0]+"turu", "Language_ID": lg, "Verb_Cognateset_ID": f"DETRZ+talk{i}", "Affected": "y", "Meaning_ID": "talk"}, ignore_index=True)
+#     for i in range(0, 1000):
+#         affectedness = affectedness.append({"Form": conds[0]+"turu", "Language_ID": lg, "Verb_Cognateset_ID": f"DETRZ+talk{i}", "Affected": "y", "Meaning_ID": "talk"}, ignore_index=True)
 
 bool_map = {
     "latex": {
@@ -766,9 +765,10 @@ for i, ext in e_df.iterrows():
         caption=f"Evaluating predictions for {name_dic[lg]}",
         keep_index=True,
     )
-
+    
+    r_df["Score"] *= 100
     save_float(
-        print_latex(r_df, keep_index=True, multicolumn=False, float_format="%.2f%%"),
+        print_latex(r_df, keep_index=True, multicolumn=False, float_format="%.1f%%"),
         label,
         f"Evaluating predictions for {print_shorthand(lg)}",
     )
@@ -794,6 +794,8 @@ caption = "Overview of prediction accuracy"
 label = "resultsoverview"
 
 res_df = pd.DataFrame(all_explanations)
+res_df *= 100
+print(res_df)
 
 export_csv(
     res_df.rename(columns=lambda x: extension_string(x, latex=False), index=repl_latex),
@@ -803,7 +805,7 @@ export_csv(
 )
 
 save_float(
-    print_latex(res_df.rename(columns=extension_string).transpose(), keep_index=True, multicolumn=False, float_format="%.2f%%"),
+    print_latex(res_df.rename(columns=extension_string).transpose(), keep_index=True, multicolumn=False, float_format="%.1f%%"),
     label,
     caption,
 )

### 2021-11-27 documents/floats/ppekverbs.tex
new approach to testing predictions from bybee's model
some minor fixes, better PTar form for 'to go down'

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[76, 130, 131, 137, 374, 397]{von1892bakairi}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[4; p.c., Angela Fabíola Alves Chagas]{meira2003bakairi}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[76, 130, 131, 374, 397]{von1892bakairi}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
 \label{tab:ppekverbs}
@@
 {} &          \PPek &          \arara &                       \ikpeng &       \bakairi \\
-\midrule
+\mymidrule
 \qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &     \obj{w-a-} \\
@@
 \qu{come}    &    \rc{w-epɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &  \obj{k-əewɨ-} \\
-\qu{go down} &   \rc{w-ɨptə-} &  \obj{w-iptoŋ-} &  \obj{{\normalfont ?}-iptoŋ-} &  \obj{k-ɨtəgɨ-} \\
+\qu{go down} &   \rc{w-ɨptə-} &                 &                               &                \\
 \qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &     \obj{w-i-} \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2021-11-27 workflow/process_data.py

@@ -539,6 +539,7 @@ repl_dic = {
     "DETRZ+come": "come",
     "DETRZ1+bathe_1": "bathe_intr",
     "DETRZ1": "bathe_intr",
+    "go_down+?": "go_down"
 }
 
 overview["Cognateset_ID"] = overview["Verb_Cognateset_ID"].replace(repl_dic)

### 2021-11-27 documents/floats/pwaireg.tex
new crosses for readme

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[30]{waiwaihawkins1998}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-11-27 workflow/helper_functions.py

@@ -125,11 +125,11 @@ def upper_repl(match):
 
 web_checkmarks = {
     1: "✓",
-    0: "✖",
+    0: "×",
     True: "✓",
-    False: "✖",
+    False: "×",
     "\\checkmark": "✓",
-    "×": "✖"
+    "×": "×"
 }
 
 def repl_latex(string):
@@ -363,6 +363,7 @@ def export_csv(
     sources=None,
     print_i_name=False,
     print_c_name=False,
+    float_format=None
 ):
     tabular = tabular.copy()
     # if caption:
@@ -377,7 +378,7 @@ def export_csv(
         #         c = (None for x in c)
 
     # print(tabular)
-    tabular.to_csv(f"data_output/{label}.csv", index=keep_index)
+    tabular.to_csv(f"data_output/{label}.csv", index=keep_index, float_format=float_format)
     exported_tables[label] = {"caption": caption}
     if sources:
         src = cldfh.combine_refs(sources)

### 2021-11-27 workflow/process_data.py

@@ -784,6 +784,7 @@ for i, ext in e_df.iterrows():
         caption=f"Predictions for {name_dic[lg]}",
         keep_index=True,
         print_i_name=True,
+        float_format="%.1f%%"
     )
     save_float(
         print_latex(p_df, keep_index=True, multicolumn=False),
@@ -803,6 +804,7 @@ export_csv(
     label,
     caption=caption,
     keep_index=True,
+    float_format="%.1f%%"
 )
 
 save_float(

### 2021-12-02 documents/sections/extensions.tex
work on introduction

@@
 As stated in \cref{sec:extensions_intro}, the person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically) members of the \gl{s_a_} class.
-Of the six attested incomplete extensions, three can be reconstructed to intermediate proto-languages, while three others happened in earlier stages of single modern languages.
+Of the six attested incomplete extensions, three can be reconstructed to intermediate proto-languages, while three others happened in pre-modern stages of single languages.
 The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
-We discuss each extension separately, contrasting regular and innovative verbs with irregular and conservative verbs, and reconstructing forms where necessary:
+We discuss each extension separately, contrasting regular (innovative) verbs with irregular (conservative) ones, and reconstructing forms where necessary:
 \cref{sec:pekodian} investigates the innovation of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
 \cref{sec:waiwaian} takes a closer look at the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
-\cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir (or \PTar), reflected in modern \trio and \akuriyo.
-\crefrange{sec:akuriyo}{sec:yukpa} look at innovative first person markers which are only attested in single modern languages:
+\cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
+\crefrange{sec:akuriyo}{sec:yukpa} look at innovative first person markers only attested in single languages:
 \obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.
@@
 Given the lack of data for other verbs, we will not further discuss this potential additional pattern.
-If the characterization by \citeauthor{meira2003bakairi} is accurate, then the pattern is fully parallel to the distribution of the first person prefixes.
+If the characterization by \citeauthor{meira2003bakairi} is accurate, then verbs with innovative first person prefixes have conservative \gl{1+2} prefixes, and vice versa.
 }
@@
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-Further, while there are reflexes of \rc{a[p]} \qu{to be} in \ikpeng, it seems that only reflexes of \rc{eti} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
+Further, while reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, it seems that only reflexes of \rc{eti} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
 

### 2021-12-02 documents/sections/introduction.tex
waiwaian *k- review
work on introduction

@@
 \label{sec:intro}
+The Cariban language family is one of the larger of South America, with between 60'000 and 100'000 speakers unevenly distributed between 22 to 25 extant languages \parencite[441]{gildea2012classification}.
+The family is concentrated in Venezuela, the Guianas and Northern Brazil, with a two Western and four Southern outliers.
+\cref{fig:phylomap} gives an overview of the geographical distribution and genealogical position of the extant Cariban languages.
+Cariban languages feature relatively rich verbal morphology, both pre- and suffixes, inflecting for person, number, tense, aspect, and evidentiality, combined with a range of valency-modifying affixes.
+Many also have a split-\gl{s} system, which can be reconstructed to \PC \pcref{sec:split}.
+For more detailed linguistic overviews of and comparative work on the family, we refer readers to \textcites{gildea1998}{derbyshire1999carib}{meira2002first}{meira2005southern}{meira2006cariban}{gildea2007greenberg}{meira2010origin}{gildea2010story}{gildea2012classification}{matter2021cariban}{gildea2019overview}.
 
-\subsection{The Cariban language family}
-
-%todo write something here about the family in general
-%maybe a nice little map-cum-tree, eh?
-%\input{floats/cariban_tree}
-
-\subsection{Verbs with conservative first person forms}
+\input{floats/cariban_map.tex}
 
@@
 \input{floats/triintro}
-Cariban languages feature relatively rich verbal morphology, both pre- and suffixes, inflecting for person, number, tense, aspect, and evidentiality, combined with a range of valency-modifying derivational markers.
-A split-\gl{s} system can be reconstructed to \PC, discussed in more detail in \cref{sec:split}.
-\cref{tab:hixintro} shows person paradigms for four \hixka verbs, members of the \gl{s_a_} inflectional class.%
-\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
-This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms we show.}%
-\footnote{In \cref{tab:hixintro} and the remainder of this paper, we omit any TAM suffixes found in the original forms found in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.}%
-\footnote{
-We use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
-In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in an interlinearized glossed example.
-We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
-\footnote{Glossing abbreviations: }%\glossingAbbrevsComma.}
-A comparison of these paradigms shows that the verb \qu{to be} diverges from the others in its first person marker.
-That prefix is \obj{w-}, while other \gl{s_a_} verbs like \qu{to fall} have \obj{k(ɨ)-}.
-A similar example can be found in \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
-In both languages, the verbs on the left of the table are representative for the \gl{s_a_} class, since the vast majority is inflected identically.
-In both languages, there are only a few other verbs inflected identically to the divergent ones on the right; for example, the first-person form of \trio \qu{to be} is \obj{w-ei-} \parencites[339]{triomeira1999}.
 
-In our synchronic analyses of these two languages, we consider these verbs and their first person prefixes \textsc{irregular}, contrasting with regular prefixes on regular \gl{s_a_} verbs, \obj{k(ɨ)-} and \obj{t-}/\obj{s-}.
+Some languages show a small group of verbs with a divergent first person inflection.
+This is illustrated for \hixka in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
+This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms we show.
+In \cref{tab:hixintro} other inflectional tables, we omit any TAM suffixes found in the original forms found in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.
+Further, we use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in interlinearized examples.
+We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.
+Glossing abbreviations: }%\glossingAbbrevsComma.
+which shows person paradigms for four verbs, all members of the \gl{s_a_} inflectional class.
+In this language, the verb \qu{to be} diverges in its first person marker (\obj{w-}), contrasting with other \gl{s_a_} verbs like \qu{to fall}, which have \obj{k(ɨ)-}.
+A similar pattern can be found in \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
+In both languages, the inflectional patterns of the verbs on the left of the table is representative for the vast majority of \gl{s_a_} verbs.
+%In both languages, there are only a few other verbs inflected identically to the divergent ones on the right; for example, the first-person form of \trio \qu{to be} is \obj{w-ei-} \parencites[339]{triomeira1999}.
+
+In the literature, such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}.
+In a language-particular synchronic analysis, these verbs and their first person prefixes may be considered to be \textsc{irregular}, contrasting with regular prefixes, like \hixka \obj{k(ɨ)-} and \trio \obj{t-}/\obj{s-}, on regular verbs.
 However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[][e.g.]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
-For such approaches, these verbs simply belong to a small inflectional (sub-)class. %
-%\footnote{That analysis is found in synchronic descriptions of other Cariban languages discussed in more detail later, see \textcites[149]{alves2017arara}[4]{meira2003bakairi}.}
-Regardless of synchronic analysis, the explanation for these inflectional patterns lies, as so often, in diachrony.
-Our story starts with the verbal person marking system of \PC, discussed in \cref{sec:pc_person}.
-That system was subject to different kinds of innovations; the one responsible for the irregular \hixka and \trio verbs is introduced in \cref{sec:extensions_intro}.
-A particular component of that system, the distinction between \gl{s_a_} and \gl{s_p_} verbs, played a major role in the developments under discussion and is the topic of \cref{sec:split}.
+For such approaches, these verbs simply belong to a small inflectional (sub-)class, an analysis suggested for the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
 
-%
+Regardless of the synchronic analysis, the reason for these inflectional patterns can be found in the diachrony of the languages in question.
+The purpose of this paper is to approach the patterns from a comparative perspective and to provide a unifying diachronic account, for which we first introduce the main clause verbal person marking system of \PC, showing that irregular verbs are conservative \pcref{sec:pc_person}.
+That system was subject to many different kinds of innovations, one of which is responsible for  the irregular first person forms \pcref{sec:extensions_intro}.
+A particular component of the system, the distinction between \gl{s_a_} and \gl{s_p_} verbs, played a major role in the developments under discussion and is the topic of \cref{sec:split}.
+\cref{sec:outlook} summarizes the main points of this section and provides an outlook for the following sections.
+
+% analyses:
 %\begin{inlinelist}
@@
 %	\item not discussed \parencites{waiwaihawkins1998}{ikpengpacheco2001}{alves2013verbo}{pacheco2003intransitivos}
-%\end{inlinelist}.s
-
+%\end{inlinelist}.
 %meira: non-detransitivized verbs
@@
 
-\subsubsection{\PC verbal person marking}
+\subsection{\PC verbal person marking and conservative verbs}
 \label{sec:pc_person}
@@
 
-Equipped with knowledge about the ancestral system, it becomes clear that the irregularly inflected \hixka and \trio  preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-} and are therefore \textsc{conservative}.
+Equipped with this knowledge about the ancestral system, it becomes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \textsc{conservative}.
 They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
 The reflexes of \rc{w-} may be considered \textsc{relics}, old and restricted to specific lexically conditioned contexts, contrasting with the innovative prefixes found elsewhere.
-The verbs and their prefixes are comparable with the few English nouns like \obj{ox}, which preserve the old plural suffix \obj{-en}.
-It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{Ochse-n} \qu{ox-en}, \obj{Name-n} \qu{name-s}, \obj{Hase-n} \qu{hare-s}, \obj{Bär-en} \qu{bear-s}.
+These verbs and their prefixes are comparable with the few English nouns like \obj{ɒks}, which preserve the old plural suffix \obj{-ən}.
+It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{ɔks-n} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
 Since the irregular \hixka and \trio prefixes are conservative and the regular prefixes are innovative, the next question to be addressed is where these new prefixes came from.
@@
 
-\subsubsection{Person marker extensions in intransitive verbs}
+\subsection{Person marker extensions in intransitive verbs}
 \label{sec:extensions_intro}
@@
 All innovations are person marker extensions except 2b, which combined a pronoun with the linker \rc{j-}.
-They are printed in bold in \cref{fig:par_ext}, which reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and the additional alternative \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
+They are printed in bold in \cref{fig:par_ext}, which reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
 %
@@
 
-\textcite{gildea1998} discusses person marker extensions only in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment; for our story we will zoom in on a so far neglected aspect of these extensions.
-To begin, we argue that they took place via lexical diffusion, a type of extension \parencite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment; for our story of conservative verbs we will zoom in on a particular aspect of these extensions.
+To begin, we argue that they took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
 First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
@@
 
-Our third argument in favor of the lexical diffusion scenario brings us back to the \hixka and \trio verbs with conservative first person forms:
+Our third argument in favor of the lexical diffusion scenario brings us back to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
 In both cases, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
-We interpret the continued presence of the old \gl{1}\gl{s_a_} marker in a few verbs as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
-In our investigations of person marker extensions, we identified 19 individual cases affecting intransitive verbs, and found 6 of them to be incomplete.
+We interpret the continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
+In our family-wide investigation of person marker extensions, we identified 19 distinct extensions affecting intransitive verbs, and found 6 of them to be incomplete.
 These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages. %, marked with bold print in \cref{fig:caribantree}.
+
 Interestingly, all six extensions featured innovative first person markers on \gl{s_a_} verbs.
 While extensions occurred with other person values as well, they never affected \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affected all potential targets.
-Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
-
-The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affeting them points to the split-\gl{s} system playing some role, a suspicion which will be confirmed in \cref{sec:discussion}.
-Thus, we will discuss that system in a little more detail before moving on to the incomplete extensions.
+Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
+%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
+The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affecting them points to the split-\gl{s} system playing some role, so we will discuss its main properties in \cref{sec:split}.
 This will also give readers an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
@@
 
-\subsubsection{The morphological basis of the split-\gl{s} system}
+\subsection{The Cariban split-\gl{s} system}
 \label{sec:split}
-As seen in \cref{sec:pc_person}, the \PC distinction between \gl{s_a_} and \gl{s_p_} verbs was implemented by two inflection classes within the \setone prerson paradigm, but this was not the only morphological criterion:
+As seen in \cref{sec:pc_person}, the \PC distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by two inflection classes within the \setone person paradigm, but this was not the only inflectional difference:
 %\begin{inlinelist}
@@
 %
-Many languages show an \gl{s_a_} class marker in deverbalized forms, which can be reconstructed to \PC as \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
+Many languages show an \gl{s_a_} class marker in deverbalized forms, originating in \PC \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
 With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
@@
 %
-The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives.
-Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed (both suffixed with \rc{-kə}).
+\input{floats/imperatives}
+%
+The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives, where the latter took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while the former received no prefix (both suffixed with \rc{-kə}).
 This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
-Both the \gl{s_a_} marker \rc{w-} in participles and nominalizations and the prefixes in imperativees have been lost in some languages, but are reconstructible to \PC.
-However, there was one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
+Both the \gl{s_a_} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
+%However, there was one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
 
-\input{floats/imperatives}
-
-%\subsubsection{The detransitivizing prefix}
-Mismatches between the semantics of intransitive verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with modern \kalina data in \exref[kar1]{kar2}.
+In the modern languages maintaining the split-\gl{s} system, mismatches between the semantics of verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with \kalina data in \exref[kar1]{kar2}.
 
@@
 The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
-The intransitive verbs in \exref{kar2} show the same person markers, but there is no choice what marker a specific verb uses.
+The intransitive verbs in \exref{kar2} show the same person markers, but the choice of marker depends on the verb, rather than semantics.
 
@@
 %
-Rather, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
-It seems that in both cases, the prefix does not contribute to the semantics of the predicate, since there are clear mismatches:
+Thus, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
 \qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker must be considered volitional.
@@
 nor Aktionsart % \citeyearpar[215]{meira2000split} 
-satisfactorily predict the class membership of intransitive verbs.
+satisfactorily predict the class membership of intransitive verbs in any of the languages.
 
-Rather, the reason for a verb to take the \gl{a}- or \gl{p}-oriented prefix is (at least diachronically) a morphological one.
+Rather, the reason for a verb to take \gl{a}- or \gl{p}-oriented prefixes is (at least diachronically) a morphological one.
 \textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
@@
 
-As for the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although they have since merged into a single morpheme, apparently in all languages.
+As for the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although on verbs their reflexes have been merged in a single morpheme in modern languages.
 Modern reflexes of \detrz show a range of meanings, which can all be characterized as \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
@@
 
-\subsubsection{Summary and outlook}
+\subsection{Summary and outlook}
+\label{sec:outlook}
 We identified \gl{s_a_} verbs that irregularly inflect for first person in 9 Cariban languages.
-These irregular forms are actually conservative, unaffected by person marker extensions spread via lexical diffusion.
+These irregular forms are actually conservative, belonging to a few verbs unaffected by person marker extensions spread via lexical diffusion.
 Such conservative forms are only found among the first person forms of  (etymological) \gl{s_a_} verbs.
-Our primary goal is to establish what verbs remained unaffected by the individual extensions and to search for factors explaining the patterns, proceeding as follows:
+The goal is to establish what verbs remained unaffected by the individual extensions and to find reasons for their conservativeness, for which we proceed as follows:
 The topic of \cref{sec:extensions} are the six incomplete innovations, the innovative markers they introduced, and the verbs they left untouched.

### 2021-12-02 documents/sections/extensions.tex
fixed pekodian 'go down' in comparative tables

@@
 In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
-Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ}, accompanied by the introduction of first person5 \obj{k-}.
+Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ}, accompanied by the introduction of first person \obj{k-}.
 
@@
 \label{sec:waiwaian}
-This extension led to the \hixka pattern from \cref{tab:hixintro} in \cref{sec:irreg_intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
+This extension led to the \hixka pattern from \cref{tab:hixintro} in \cref{sec:intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
 In \PWai, the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was extended to \gl{1}\gl{s_a_}.

### 2021-12-02 workflow/helper_functions.py

@@ -219,7 +219,7 @@ def reconstructed_form_table(lgs, proto, verbs, caption, name):
     pyd.filters = {"Inflection": ["1"], "Language_ID": lgs, "Concept": verbs}
     pyd.x_sort = lgs
     pyd.y_sort = ["be_1", "be_2", "say", "go", "come", "go_down", "bathe_intr"]
-    tabular = pyd.compose_paradigm(i_df)
+    tabular = pyd.compose_paradigm(i_df.replace({"go_down+?": "go_down"}))
     sources = get_sources(i_df)
     raw_sources = get_sources(i_df, latexify=False)
     tabular.index.name = None

### 2021-12-03 documents/floats/pwaireg.tex
moved verb discussion to verbs section, fixed lingpy alignments, fixed map

@@
-\begin{table}
+\begin{table}[htbp]
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[30]{waiwaihawkins1998}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-12-03 documents/sections/extensions.tex
moved verb discussion to verbs section, fixed lingpy alignments, fixed map

@@
 While it is by no means a regular sound change, it allows us to securely connect the two prefixes.
-Similarly, the supposed change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+Similarly, the suggested change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
 Thus, we reconstruct a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} to \PPek, identical to the \arara one both in form and distribution.
-
-As for the forms of the verb stems, a few comments are necessary:
-For \qu{to be}, \ikpeng \obj{e} is very likely the original vowel, given the \PC form \rc{eti} \pcref{sec:be}.
-For \qu{to go down}, we reconstruct \rc{ɨ} as the initial vowel rather than \rc{i} \pcref{sec:godown}.
-Further, the forms are not fully cognate; \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}.
-However, the addition of a final \obj{ŋ} in \PXin is attested elsewhere,%
-\footnote{
-\begin{inlinelist}
-\item \PC \rc{əne} \qu{to see}, \arara and \ikpeng \obj{eneŋ} %\parencites[8]{gildea2007greenberg}[56]{alves2017arara}[25]{ikpengpacheco2001}
-\item \PC \rc{əta} \qu{to hear}, \arara \obj{taŋ}, \ikpeng \obj{iraŋ} %\parencites[8]{gildea2007greenberg}[144]{alves2017arara}[270]{ikpengpacheco2001}
-\item \PC \rc{ənə} \qu{to eat meat}, \arara \obj{oŋoŋ} \qu{to bite} %\parencites[8]{gildea2007greenberg}[57]{alves2017arara}
-\end{inlinelist} \parencites[8]{gildea2007greenberg}[56, 144, 57]{alves2017arara}[25, 270]{ikpengpacheco2001}.}
-and based on the fact \bakairi has generally lost much segmental material, we suggest that both \obj{ŋ} and \obj{gɨ} are later additions to a root \rc{ɨptə}.
 
@@
 
-The forms for \qu{to come} are not fully cognate, either: \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
+For the reconstructed forms of the \PPek verb stems we refer to \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
+The verb stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
 In contrast, the \arara first person form is directly based on this root \rc{epɨ}.
@@
 On the other hand, both \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following the line of reasoning used by \textcite[114]{meira1998proto} (see also \cref{sec:taranoan}) for a similar pattern in the three Taranoan languages, we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}; similar issues are found outside of Pekodian \pcref{sec:come}.
-
-Finally, the V-initial nature of \PPek \qu{to go} is evidenced in its Xinguan forms; while the \bakairi change \phon{\rc{wɨ}}{\obj{u}} obscured the morpheme boundary, other forms are V-initial \pcref{sec:go}.
-The \ikpeng form \obj{aran} is compatible with our reconstruction \rc{ɨtən} when considering that \ikpeng \obj{a} is an attested outcome of \rc{ə}.\footnote{\begin{inlinelist}
- \item \obj{akari} \qu{dog} in \cref{tab:pxinw} above
- \item \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg}
- \item \obj{jaj} \qu{tree} \parencite[98]{ikpengpacheco2001} from \PC \rc{jəje}
- \end{inlinelist}.}
-This attested change of \rc{ə} to \obj{a} need only be preceded by a assimilatory lowering of initial \rc{ɨ} to \rc{ə}, to yield the form \obj{aran} from \rc{ɨtən}.
-Other \ikpeng reflexes of \qu{to go} offer evidence for the suggested intermediate stage \rc{\underline{ə}t\underline{ə}n}: \obj{\underline{e}r\underline{o}-lɨ} \qu{s/he went} \parencite[25]{ikpengpacheco2001}.
+Following the line of reasoning used by \textcite[114]{meira1998proto} for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}; this verb shows similar throughout the family \pcref{sec:come}.
 
@@
 \label{sec:waiwaian}
-This extension led to the \hixka pattern from \cref{tab:hixintro} in \cref{sec:intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
+This person marker extension led to the \hixka pattern from \cref{tab:hixintro} in \cref{sec:intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
 In \PWai, the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was extended to \gl{1}\gl{s_a_}.
@@
 %
-This correspondence allows us to establish \hixka \obj{ɨ-} a reflex of \rc{w(ɨ)-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
-Notably, \textcite{hixkaryanaderby1985} does not see this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_} prefix, but as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive  \pcref{sec:say}.
+This correspondence allows us to establish \hixka \obj{ɨ-} as a reflex of \rc{wɨ-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
+Notably, \textcite{hixkaryanaderby1985} does not analyze this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_} prefix, but as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive \pcref{sec:say}.
 
 There are three more verbs which did not take innovative \rc{k-} in \PWai, shown alongside \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
-The two roots for \qu{to be} are straightforwardly reconstructible, whereas \qu{to go} is somewhat of a special case.
+The two forms for \qu{to be} are straightforwardly reconstructible, whereas \qu{to go} is somewhat of a special case.
 While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
@@
 
-%\pex<parpar>
-%\a<parcop> Parukotoan \qu{to be}\\
-%\begin{tabular}[t]{@{}llll@{}}
-%& \PPar & \hixka & \waiwai \\
-%\gl{1}\gl{s_a_} & \rc{\emp{w-}eʃi-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eʃe-}/\obj{\emp{w-}ah-} & \obj{\emp{w-}eeʃi-}/\obj{\emp{w-}a-}\\
-%\gl{2}\gl{s_a_} & \rc{m-eʃi-}/\obj{m-ah-} & \obj{m-eʃe-}/\obj{m-ah-} & \obj{m-eeʃi-}/\obj{m-a-}\\
-%\gl{1+2}\gl{s_a_} & \rc{t-eʃi-}/\obj{t-ah-} & \obj{t-eʃe-}/\obj{t-ah-} & \obj{t͡ʃ-eeʃi-}/\obj{t-a-}\\
-%\gl{3}\gl{s_a_} & \rc{n-eʃi-}/\obj{n-ah-} & \obj{n-eʃe-}/\obj{n-ah-} & \obj{n-eeʃi-}/\obj{n-a-}\\
-%\end{tabular}\\
-%(\cite[197--198]{hixkaryanaderby1985} and Spike Gildea, p.c.)
-%\a<pargo> Parukotoan \qu{to go}\\
-%\begin{tabular}[t]{@{}llll@{}}
-%& \PPar & \hixka & \waiwai\\
-%\gl{1}\gl{s_a_} & \rc{\emp{wɨ-}to-} & \obj{\emp{ɨ-}to-} & \obj{kɨ\emp{w-}to-}\\
-%\gl{2}\gl{s_a_} & \rc{mɨ-to-} & \obj{mɨ-to-} & \obj{mɨɨ-to-}\\
-%\gl{1+2}\gl{s_a_} & \rc{tɨt-to-} & \obj{tɨ-to-} & \obj{tɨ-to-}/\obj{tɨh-t͡ʃe-}\\
-%\gl{3}\gl{s_a_} & \rc{nɨ-to-} & \obj{n-to-} & \obj{nɨɨ-to-}\\
-%\end{tabular}\\
-%(\cites[69--70, 211]{hixkaryanaderby1985}[179]{waiwaihawkins1998} and Spike Gildea, p.c.)
-%\xe
-%
-
-Summing up, we reconstruct the four verbs \rc{eʃi} and \rc{a[h]} \qu{to be}, \rc{ka[s]} \qu{to say}, and \rc{[ɨ]to[m]} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
+%Summing up, we reconstruct the four verbs \rc{eʃi} and \rc{a[h]} \qu{to be}, \rc{ka[s]} \qu{to say}, and \rc{[ɨ]to[m]} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
 
@@
 \label{sec:taranoan}
-The moniker Tiriyoan subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages already identified by \textcite{girard1971proto}, the more distant member being \carijo.
+The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages already identified by \textcite{girard1971proto}, with \carijo as a more distant member.
 \textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar.
@@
 \input{floats/ptirreg}
-The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
+The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario in which both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
 In fact, even the issue of \emph{when} this happened is open.
@@
 Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
-We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it taking place already in \PTar.}
-%
-Regarding \gl{s_a_} verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar -- attentive readers my recognize \trio \qu{to go} -- for which we provide our reconstructed \PTir forms.
-To this list, the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}} can be added, which has first person \obj{w-}, at least in \trio.
-
-%\pex<ptirw> \PTir \gl{1}\gl{s_a_} \rc{w-} \parencites[112--115]{meira1998proto}[294]{triomeira1999}\\ %[85]{gildea1994akuriyo}
-%\begin{tabular}[t]{@{}llll@{}}
-%& \PTir & \trio & \akuriyo\\
-%\qu{go} & \rc{wɨ-tə(mɨ)-} & \obj{wɨ-tə(mɨ)-} & \obj{ə-təmɨ-} \\% (\obj{wɨ-təemɨ-})\\
-%\qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
-%\qu{come} & \rc{w-əʔepɨ-} & \obj{w-əepɨ-} & \obj{Ø-eepɨ-} \\
-%\qu{be} & \rc{w-ae-} & \obj{w-ae-} & Ø\obj{-aʔe-} \\
-%& \rc{w-eʔi-} & \obj{w-ei-} & ?-\obj{eʔi-}\\
-%\end{tabular}
-%\a<aku12> \akuriyo \obj{kəʔ-eepɨ} \qu{we came} \parencite[114]{meira1998proto}
-%\a<aku13> \akuriyo \obj{(w)i-toka} \qu{I hit him/her} \parencite[86]{gildea1994akuriyo}
-%\a<tricome> \trio \obj{n-epɨ} \qu{s/he came} \parencite[114]{meira1998proto}
-%\xe
+We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it taking place already in \PTar; this decision is irrelevant for our analysis.}
 
@@
 
+Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which we provide our reconstructed \PTir forms.
+To this list, the other copular root \rc{eʔi} can be added, whose \trio reflex retains first person \obj{w-}.
 We agree with \posscite[113]{meira1998proto} identification of the idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} as a reflex of \rc{wɨ-}.
-Both components of the irregular change \rc{wɨ-} > \obj{ə-} -- loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə} -- are found in other person prefixes \exref{exp}.
+Both components of the irregular change \rc{wɨ-} > \obj{ə-} -- loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə} -- are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
 
@@
 %
+
 For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for the other person values, based on an idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
-\akuriyo (and \carijo) then levelled this original distribution, similar to what we have suggested for Pekodian \pcref{sec:pekodian}.
-We agree with this scenario, with the exception that \trio \obj{əepɨ} looks like a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ}.
+Both \akuriyo and \carijo then levelled this original distribution, similar to what we have suggested for the Pekodian languages \pcref{sec:pekodian}.
+We agree with this scenario, with the exception that we identify \trio \obj{əepɨ} as a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
 
@@
 Further, these verbs are not mentioned by \textcite{meira1998proto}, who was also working with \posscite{gildea1994akuriyo} data.
-Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by the available part of \posscite{meira1998proto} data, we cannot reconstruct these verbs as not being affected by the extension of \PTir \rc{t-}.
+Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by the available part of \posscite{meira1998proto} data, we cannot confidently reconstruct these verbs as resisting the extension of \PTir \rc{t-}.
 
@@
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} might be more recent, with which we agree: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{e} is reconstructible to \PTir, the most likely scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-}.
-The few \obj{t-} mentioned by \textcite{meira1998proto} were then either reintroduced under \trio influence, or are the last remnants of the replacement of \rc{t-}.
-Since there are no examples of, or further information about, \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.
+He also suggests that \obj{k-} might be more recent, with which we agree: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
+The few \obj{t-} mentioned by \textcite{meira1998proto} were potentially reintroduced under \trio influence.
+However, since there are no examples of, or further information about, \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.
 
@@
 
-The verbs listed for \PTir in \cref{tab:ptirverbs} above did of course also not introduce \obj{k-} in \akuriyo, although we do not know the first-person form of the copular verb \obj{eʔi}.
-In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in \akuriyo, also reconstructible to \PTir \exref{tridown}.
+The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} did of course also not introduce \obj{k-} in \akuriyo, although we do not know the first-person form of the copular verb \obj{eʔi}.
+In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in, likely inherited from \PTir \exref{tridown}.
 It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{s_a_} verb when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).
@@
 \carijo, the cousin of the Tiriyoan languages, has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
-Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs \pcref{tab:carreg}.
+Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs, seen in the paradigms in \cref{tab:carreg}.
 %
@@
 
-As noted in \cref{sec:taranoan}, this extension also erased any traces of a potential \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
+As noted in \cref{sec:taranoan}, this extension also erased any traces of a putative \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
 However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
@@
 %We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
-Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would either expect conservative \obj{wɨ-} or innovative \obj{ji-}.
-We analyse this ∅ as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
+Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would expect \obj{ka} \qu{to say} to either have conservative \obj{wɨ-} or innovative \obj{ji-}, so the zero is unexpected.
+We analyse it as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
 Further, it is possible that the divergent development of \rc{w-} on C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} is a result of the latter's originally V-initial nature \pcref{sec:go}.
@@
 The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
-However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm is found \pcref{tab:yukreg}.
-The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
+However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm \pcref{tab:yukreg} is found.
+The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (now former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
 %
@@
 %
-These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their reflex of the detransitivizer \detrz.
-The same prefix occurs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, meaning that \obj{j(ɨ)-} is the reflex of the \PC \gl{1}\gl{s_p_} marker \rc{u(j)-}.
+These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their semantics and the reflex of \detrz.
+The same prefix occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, meaning that \obj{j(ɨ)-} is the reflex of the \PC \gl{1}\gl{s_p_} marker \rc{u(j)-}.
+On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
 
@@
 \xe
-On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family (see \cref{sec:pekodian,sec:waiwaian}), we identify the zero marking in \gl{1}>\gl{3} scenarios as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
+%
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, we identify the zero marking in \gl{1}>\gl{3} scenarios as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
 
-In intransitive verbs, this first-person zero marking is also attested, but only for \obj{to} \qu{to go} \exref{yuk-7}.
-That verb contrasts with regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
-In summary, the extension of \obj{j-} in \yukpa only spared \obj{to} \qu{to go}, which preserves a phonologically reduced reflex of \rc{wɨ-}.
+In intransitive verbs, this first-person zero marking is attested in a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
+It diverges from regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
+It is thus the only verb attested as having resisted the extension of \obj{j-} in \yukpa.
 

### 2021-12-03 documents/sections/introduction.tex
moved verb discussion to verbs section, fixed lingpy alignments, fixed map

@@
 This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms we show.
-In \cref{tab:hixintro} other inflectional tables, we omit any TAM suffixes found in the original forms found in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.
+In \cref{tab:hixintro} and other paradigm tables, we omit any TAM suffixes found in the original forms found in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.
 Further, we use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.

### 2021-12-03 documents/sections/verbs.tex
work on introduction
moved verb discussion to verbs section, fixed lingpy alignments, fixed map

@@
-\section{Resistant verbs from a comparative perspective}
+\section{Conservative verbs: the usual suspects}
 \label{sec:verbs}
@@
 \label{sec:be}
-For a comprehensive comparative overview for these two roots, we refer the reader to \textcite[375--382]{gildea2018reconstructing}.
+For a comprehensive comparative overview of these two roots, we refer the reader to \textcite[375--382]{gildea2018reconstructing}; we will not discuss them in detail here.
 \rc{a[p]} is the original copula and can be reconstructed as already having various irregularities in \PC.
@@
 This verb was not affected by the any of the extensions discussed in \cref{sec:extensions}.
+
+%The \ikpeng form \obj{aran} is compatible with our reconstruction \rc{ɨtən} when considering that \ikpeng \obj{a} is an attested outcome of \rc{ə}: \begin{inlinelist}
+%	\item \obj{akari} \qu{dog} in \cref{tab:pxinw} above
+%	\item \obj{anma} \qu{path} \parencite[24]{ikpengpacheco2001} from \PC \rc{ətema} \parencite[12]{gildea2007greenberg}
+%	\item \obj{jaj} \qu{tree} \parencite[98]{ikpengpacheco2001} from \PC \rc{jəje}
+%\end{inlinelist}.
+%This attested change of \rc{ə} to \obj{a} need only be preceded by a assimilatory lowering of initial \rc{ɨ} to \rc{ə}, to yield the form \obj{aran} from \rc{ɨtən}.
 
@@
 However, a broader comparative perspective reveals a much more complicated story.
-\cref{tab:godown} shows all attested cognates of \rc{ɨpɨtə}, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.
-It turns out that while a form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
+\cref{tab:godown} shows all attested reflexes of \rc{ɨpɨtə}, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.\footnote{Our reconstruction of \PPek \rc{ɨptə} treats the additional elements found in the daughter languages as non-cognate.
+\textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}, and at least the addition of a final \obj{ŋ} in \PXin is attested elsewhere:
+		\begin{inlinelist}
+			\item \PC \rc{əne} \qu{to see}, \arara and \ikpeng \obj{eneŋ} %\parencites[8]{gildea2007greenberg}[56]{alves2017arara}[25]{ikpengpacheco2001}
+			\item \PC \rc{əta} \qu{to hear}, \arara \obj{taŋ}, \ikpeng \obj{iraŋ} %\parencites[8]{gildea2007greenberg}[144]{alves2017arara}[270]{ikpengpacheco2001}
+			\item \PC \rc{ənə} \qu{to eat meat}, \arara \obj{oŋoŋ} \qu{to bite} %\parencites[8]{gildea2007greenberg}[57]{alves2017arara}
+		\end{inlinelist} \parencites[8]{gildea2007greenberg}[56, 144, 57]{alves2017arara}[25, 270]{ikpengpacheco2001}.}
+It turns out that while the form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
 Its reflexes in languages that preserve the split-\gl{s} system are distributed fairly evenly between \gl{s_a_} and \gl{s_p_}.
@@
 
-\input{floats/bathe}
+%\input{floats/bathe}

### 2021-12-03 workflow/helper_functions.py

@@ -56,7 +56,7 @@ def save_float(tabular, label, caption, filename=None, short=None):
         shorttext = ""
     else:
         shorttext = f"[{short}]"
-    output = f"""\\begin{{table}}
+    output = f"""\\begin{{table}}[htbp]
 \centering
 \caption{shorttext}{{{caption}}}
 \label{{tab:{label}}}

### 2021-12-03 workflow/lingpy_alignments.py

@@ -120,7 +120,7 @@ def calculate_alignment(df, fuzzy=False):
     else:
         lex = lingpy.Wordlist(tempfile)
         alm = lingpy.Alignments(lex, ref="cogids")
-    alm.align()
+    alm.align(model="sca", method="library")
     alm.output("tsv", filename="lingpy_temp", ignore="all", prettify=False)
 
     df = pd.read_csv(tempfile, sep="\t", dtype="str")

### 2021-12-05 documents/sections/extensions.tex
rewriting and restructuring

@@
-\section{Incomplete extensions: the innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
+\section{Inflections and verbs: innovation and resistance}
+\label{sec:data}
+As stated in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
+\cref{sec:extensions} introduces the six identified incomplete extensions, the prefixes they introduced and the verbs they spared.
+Since there is considerable etymological overlap in which verbs these are, the seven verbs showing a considerable degree of resistance are compared and reconstructed in \cref{sec:verbs}.
+The goal of this section is to identify distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
+
+\subsection{Incomplete extensions: the innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
 \label{sec:extensions}
-As stated in \cref{sec:extensions_intro}, the person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically) members of the \gl{s_a_} class.
+As stated in \cref{sec:extensions_intro}, the person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically, in the case of loss of split-\gl{s}) members of the \gl{s_a_} class.
 Of the six attested incomplete extensions, three can be reconstructed to intermediate proto-languages, while three others happened in pre-modern stages of single languages.
 The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
-We discuss each extension separately, contrasting regular (innovative) verbs with irregular (conservative) ones, and reconstructing forms where necessary:
+Each extension is discussed separately, contrasting regular (innovative) verbs with irregular (conservative) ones, and reconstructing forms where necessary:
 \cref{sec:pekodian} investigates the innovation of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
-\cref{sec:waiwaian} takes a closer look at the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
+\cref{sec:waiwaian} is about the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
 \cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
-\crefrange{sec:akuriyo}{sec:yukpa} look at innovative first person markers only attested in single languages:
+The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative first person markers only attested in single languages:
 \obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.
 
-\subsection{\PPek \rc{k-}}
+\subsubsection{\PPek \rc{k-}}
 \label{sec:pekodian}
@@
 
-In the most detailed description of a Pekodian language, \textcite{alves2017arara} describes six%
+In the most detailed description of a Pekodian language, \textcite{alves2017arara} names six%
 \footnote{Seven under her analysis, which sees the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.}
-\arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, listed in \exref{arairr}.
+\arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
 From a comparative perspective, this list is not quite complete, as there is also a reflex of the copula \rc{a[p]}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
@@
 However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
-Given the lack of data for other verbs, we will not further discuss this potential additional pattern.
+Given the lack of data for other verbs, this potential additional pattern will not be further discussed.
 If the characterization by \citeauthor{meira2003bakairi} is accurate, then verbs with innovative first person prefixes have conservative \gl{1+2} prefixes, and vice versa.
@@
 Luckily, while \textcite{von1892bakairi} did not accurately record all phonemic distinctions in \bakairi \parencite{meira2005bakairi}, he does provide inflected forms of cognates to the \arara verbs in \exref{arairr}.
-We present them in \exref{bakverbs} according to our current understanding of \bakairi phonology and verbal morphology, based on \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
+They are presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
 
@@
 
-\Cref{tab:ppekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with our \PPek reconstruction.
+\Cref{tab:ppekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with reconstructed \PPek forms.
 The presence and distribution of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
 Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \cref{tab:pxinw}.
-While it is by no means a regular sound change, it allows us to securely connect the two prefixes.
+While it is by no means a regular sound change, it permits identifying the two prefixes as cognate.
 Similarly, the suggested change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
-Thus, we reconstruct a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} to \PPek, identical to the \arara one both in form and distribution.
+Thus, a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to the \arara one both in form and distribution.
+In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
 
@@
 
-For the reconstructed forms of the \PPek verb stems we refer to \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
+For the reconstructed forms of the \PPek verb stems readers are referred to \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
 The verb stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
@@
 On the other hand, both \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following the line of reasoning used by \textcite[114]{meira1998proto} for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), we suggest that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}; this verb shows similar throughout the family \pcref{sec:come}.
+Following the line of reasoning used by \textcite[114]{meira1998proto} for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), it is likely that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}; this verb shows similar patterns throughout the family \pcref{sec:come}.
 
-Summing up, an innovative \gl{1}\gl{s_a_} marker \rc{k-} is reconstructible to \PPek.
-Seven verbs can be reconstructed as having resisted this innovation and preserving \gl{1}\gl{s_a_} \rc{w(ɨ)-} in \PPek.
-In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
-Further, both languages regularized the paradigm of \qu{to come} to \rc{əd-epɨ}, accompanied by the introduction of first person \obj{k-}.
-
-\subsection{\PWai \rc{k-}}
+\subsubsection{\PWai \rc{k-}}
 \label{sec:waiwaian}
-This person marker extension led to the \hixka pattern from \cref{tab:hixintro} in \cref{sec:intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
+This person marker extension led to the \hixka pattern shown in \cref{sec:intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
 In \PWai, the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was extended to \gl{1}\gl{s_a_}.
@@
 %
-This correspondence allows us to establish \hixka \obj{ɨ-} as a reflex of \rc{wɨ-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
+This regular correspondence in transitive verbs points to \hixka \obj{ɨ-} on intransitive verbs as another reflex of \rc{wɨ-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
 Notably, \textcite{hixkaryanaderby1985} does not analyze this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_} prefix, but as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive \pcref{sec:say}.
@@
 %
-In any case, \hixka \qu{to go} was clearly not affected by the extension of \rc{k-}, allowing us to reconstruct a \PWai first person form \rc{wɨ-tom-}.
+In any case, this was a subsequent development, as the prefix pattern in \hixka \qu{to go} clearly shows that \PWai first person form \rc{wɨ-tom-} can be reconstructed.
 
@@
 
-\subsection{\PTir \rc{t-}}
+\subsubsection{\PTir \rc{t-}}
 \label{sec:taranoan}
@@
 Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
-We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it taking place already in \PTar; this decision is irrelevant for our analysis.}
+Here, a conservative stance is taken and the innovation is assumed to be \PTir, but it could have well already taken place in \PTar.
+This decision has no effect on the results of this study.}
 
@@
 
-Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which we provide our reconstructed \PTir forms.
-To this list, the other copular root \rc{eʔi} can be added, whose \trio reflex retains first person \obj{w-}.
-We agree with \posscite[113]{meira1998proto} identification of the idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} as a reflex of \rc{wɨ-}.
+Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.
+To this list, the second copular root \rc{eʔi} can be added, whose \trio reflex retains first person \obj{w-}.
+The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is identified as a reflex of \rc{wɨ-} by 
+\textcite[113]{meira1998proto}, which is plausible and supported by additional evidence:
 Both components of the irregular change \rc{wɨ-} > \obj{ə-} -- loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə} -- are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
@@
 For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for the other person values, based on an idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
-Both \akuriyo and \carijo then levelled this original distribution, similar to what we have suggested for the Pekodian languages \pcref{sec:pekodian}.
-We agree with this scenario, with the exception that we identify \trio \obj{əepɨ} as a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
+Both \akuriyo and \carijo then levelled this original distribution, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
+This scenario is plausible, with the exception that \trio \obj{əepɨ} is a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
 
@@
 In addition, \textcite{gildea1994akuriyo} recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-} \exref{akumov.aku}, all \obj{e}-initial movement verbs.
-We have only found a \trio cognate for \obj{erama} \qu{to return}, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
+Only \obj{erama} \qu{to return} has an attested \trio cognate, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
 Further, these verbs are not mentioned by \textcite{meira1998proto}, who was also working with \posscite{gildea1994akuriyo} data.
-Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by the available part of \posscite{meira1998proto} data, we cannot confidently reconstruct these verbs as resisting the extension of \PTir \rc{t-}.
+Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by the available part of \posscite{meira1998proto} data, these verbs cannot be confidently reconstructed as resisting the extension of \PTir \rc{t-}.
 
@@
 
-\subsection{\akuriyo \obj{k-}}
+\subsubsection{\akuriyo \obj{k-}}
 \label{sec:akuriyo}
 After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
-It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in the \akuriyo data available to us \parencite{gildea1994akuriyo}, with all relevant verbs shown in \Cref{tab:aku1sa}.
+It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data, with all relevant verbs shown in \Cref{tab:aku1sa}.
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} might be more recent, with which we agree: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
-The few \obj{t-} mentioned by \textcite{meira1998proto} were potentially reintroduced under \trio influence.
-However, since there are no examples of, or further information about, \obj{ə}-initial verbs with \obj{t-}, we cannot discuss these cases.
+He also suggests that \obj{k-} might be more recent, which is very plausible: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
+The few \obj{t-} mentioned by \textcite{meira1998proto} were then potentially reintroduced under \trio influence.
+However, since there are no examples of, or further information about, \obj{ə}-initial verbs with \obj{t-}, these cases cannot be discussed further.
 
@@
 
-The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} did of course also not introduce \obj{k-} in \akuriyo, although we do not know the first-person form of the copular verb \obj{eʔi}.
-In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in, likely inherited from \PTir \exref{tridown}.
+The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} did of course also not introduce \obj{k-} in \akuriyo, although the first-person form of the copular verb \obj{eʔi} is presently unknown.
+In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in, which is apparently reconstructible to \PTir \exref{tridown}.
 It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{s_a_} verb when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).
@@
 
-\subsection{\carijo \obj{j-}}
+\subsubsection{\carijo \obj{j-}}
 \label{sec:carijo}
 \carijo, the cousin of the Tiriyoan languages, has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
-Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs, seen in the paradigms in \cref{tab:carreg}.
+Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs, reflected in the paradigms in \cref{tab:carreg}.
 %
-\input{floats/carreg}
+\input{floats/carreg}%
 %
-Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}, shown in \exref{car-30}.
+Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight} \exref{car-30}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}.
 %The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
@@
 Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would expect \obj{ka} \qu{to say} to either have conservative \obj{wɨ-} or innovative \obj{ji-}, so the zero is unexpected.
-We analyse it as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
-Further, it is possible that the divergent development of \rc{w-} on C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} is a result of the latter's originally V-initial nature \pcref{sec:go}.
+It is analysed here as a reflex of \rc{wɨ-}, primarily due to the loss of \rc{w} in \ikpeng and \hixka.
+While that loss was clearly more regular, an already irregular marker undergoing idiosyncratic phonological erosion is not surprising, see \akuriyo \rc{wɨ-} > \obj{ə-} in the preceding section.
+Alternatively, it is possible that the divergent development of \rc{w-} on C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} is a result of the latter's originally V-initial nature \pcref{sec:go}.
 
-\subsection{\yukpa \obj{j-}}
+\subsubsection{\yukpa \obj{j-}}
 \label{sec:yukpa}
@@
 %
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, we identify the zero marking in \gl{1}>\gl{3} scenarios as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identifiedn  as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
 

### 2021-12-05 documents/sections/introduction.tex
rewriting and restructuring

@@
 \label{sec:intro}
-The Cariban language family is one of the larger of South America, with between 60'000 and 100'000 speakers unevenly distributed between 22 to 25 extant languages \parencite[441]{gildea2012classification}.
-The family is concentrated in Venezuela, the Guianas and Northern Brazil, with a two Western and four Southern outliers.
+The Cariban language family is one of the largest of South America, with between 60'000 and 100'000 speakers unevenly distributed between 22 to 25 extant languages \parencite[441]{gildea2012classification}.
+The family is concentrated in Venezuela, the Guianas and Northern Brazil, with two Western and four Southern outliers.
 \cref{fig:phylomap} gives an overview of the geographical distribution and genealogical position of the extant Cariban languages.
@@
 
-Some languages show a small group of verbs with a divergent first person inflection.
+Some Cariban languages show a small group of verbs which diverge in their first person inflection pattern.
 This is illustrated for \hixka in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
@@
 In the literature, such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}.
-In a language-particular synchronic analysis, these verbs and their first person prefixes may be considered to be \textsc{irregular}, contrasting with regular prefixes, like \hixka \obj{k(ɨ)-} and \trio \obj{t-}/\obj{s-}, on regular verbs.
+In a language-specific synchronic analysis, these verbs and their first person prefixes may be considered to be \textsc{irregular}, contrasting with regular prefixes, like \hixka \obj{k(ɨ)-} and \trio \obj{t-}/\obj{s-}, on regular verbs.
 However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[][e.g.]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
-For such approaches, these verbs simply belong to a small inflectional (sub-)class, an analysis suggested for the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
+For such approaches, these verbs simply belong to a small inflectional (sub-)class, an analysis found in the literature for the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
 
 Regardless of the synchronic analysis, the reason for these inflectional patterns can be found in the diachrony of the languages in question.
-The purpose of this paper is to approach the patterns from a comparative perspective and to provide a unifying diachronic account, for which we first introduce the main clause verbal person marking system of \PC, showing that irregular verbs are conservative \pcref{sec:pc_person}.
-That system was subject to many different kinds of innovations, one of which is responsible for  the irregular first person forms \pcref{sec:extensions_intro}.
-A particular component of the system, the distinction between \gl{s_a_} and \gl{s_p_} verbs, played a major role in the developments under discussion and is the topic of \cref{sec:split}.
-\cref{sec:outlook} summarizes the main points of this section and provides an outlook for the following sections.
+The purpose of this paper is to approach the patterns from a comparative perspective and to provide a unifying diachronic account, proceeding as follows.
+In \cref{sec:background}, readers are provided with necessary background knowledge about \PC verbs, and the mechanism of person marker extensions is introduced.
+In \cref{sec:data}, six incomplete person marker extensions and their extent in the lexicon are described.
+Since there is a surprising amount of agreement about what verbs remain conservative, they are compared and reconstructed.
+\cref{sec:motivations} uses \posscite{bybee1985morphology} network model of morphology to find explanations for the verbs (not) affected by each extension.
+\cref{sec:discussion} summarizes and discusses the results, and puts them in a general context of language change.
 
@@
 %alves: non-detransitivized verbs (as a consequence, o-initial)
-
-%\subsection{Cariban verbal person marking and marker extensions}
-%\label{sec:background}
-%The archaic first person markers investigated in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system that did not affect the whole lexicon.
-%To provide readers with the necessary background, we first introduce said ancestral system \pcref{sec:pc_person}, then the peculiarities of the Cariban split-\gl{s} system \pcref{sec:split}, and finally the concept of person marker extensions \pcref{sec:extensions_intro}.
-
-\subsection{\PC verbal person marking and conservative verbs}
-\label{sec:pc_person}
-\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.
-The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
-The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
-There was a single aliophoric marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
-Local scenarios were expressed in a non-transparent manner, both using the \gl{1+2} prefix \rc{k-}.
-
-\input{floats/pcpers}
-
-Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
-That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}), as well as the absence of \rc{i} after all \gl{s_a_} prefixes.
-On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones.
-The third person marker in \gl{s_p_} verbs was identical to the one in \gl{3}>\gl{3} scenarios (\rc{n(i)-)}.
-
-Equipped with this knowledge about the ancestral system, it becomes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \textsc{conservative}.
-They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
-The reflexes of \rc{w-} may be considered \textsc{relics}, old and restricted to specific lexically conditioned contexts, contrasting with the innovative prefixes found elsewhere.
-These verbs and their prefixes are comparable with the few English nouns like \obj{ɒks}, which preserve the old plural suffix \obj{-ən}.
-It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{ɔks-n} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
-Since the irregular \hixka and \trio prefixes are conservative and the regular prefixes are innovative, the next question to be addressed is where these new prefixes came from.
-
-%\input{floats/makintro}
-
-\subsection{Person marker extensions in intransitive verbs}
-\label{sec:extensions_intro}
-In his discussion of the \PC split-\gl{s} system and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different modifications in various languages.
-The main mechanism of change leading to these modifications are \textbf{person marker extensions}, i.e.
-the use of verbal person prefixes being extended to contexts previously occupied by other prefixes.
-%\footnote{
-%In principle, one could also imagine a scenario in which verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
-%However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
-%Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
-%For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
-There have been many different person marker extensions in Cariban languages, and some are still ongoing.
-This is illustrated by \textcite{gildea1998}, using the three Parukotoan languages as an example.
-Apart from segmental changes to individual morphemes, the following innovations happened in the \setone paradigm in Parukotoan:
-
-\begin{enumerate}
-	\item \PPar \begin{enumerate}
-		\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
-		\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
-		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
-	\end{enumerate}
-	\item \PWai \begin{enumerate}
-		\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
-		\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
-	\end{enumerate}
-	\item \waiwai \begin{enumerate}
-		\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
-\end{enumerate}
-%
-All innovations are person marker extensions except 2b, which combined a pronoun with the linker \rc{j-}.
-They are printed in bold in \cref{fig:par_ext}, which reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
-%
-\begin{figure}[hbt]
-	\centering
-	\setlength{\tabcolsep}{2pt}
-	\fbox{\begin{tikzpicture}[
-		every node/.append style={align=center},
-		level distance=80pt,
-		]
-		\Tree[.{\PC{}\pcone}
-		[.{\PPar{}\ppone}
-		{\kaxui{}\kaxone}
-		[.{\PWai{}\pwone}
-		{\hixka{}\hixone}
-		{\waiwai{}\waione}
-		]
-		]
-		]
-	\end{tikzpicture}}
-	\caption{Person marking extensions in Parukotoan, after \textcite[94]{gildea1998}}
-	\label{fig:par_ext}
-\end{figure}
-%
-\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows the variation in the first person and \gl{1+2} prefixes that is reconstructible to \PPar.
-\waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
-%In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
-%That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
-%Similarly, the \gl{2}\gl{s_p_} prefix \obj{a-} was extended to imperatives of (former) \gl{s_a_} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.
-
-%Such extensions, often only affecting one person value at a time, and the eventual loss of split-\gl{s} are not entirely unexpected, given that the system lacks any semantic basis.
-
-\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment; for our story of conservative verbs we will zoom in on a particular aspect of these extensions.
-To begin, we argue that they took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
-First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
-Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
-In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which is what one would expect from a change in progress.
-%todo maybe put in GOOD kaxui data
-Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} convincingly argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
-Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
-
-Our third argument in favor of the lexical diffusion scenario brings us back to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
-In both cases, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
-We interpret the continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
-In our family-wide investigation of person marker extensions, we identified 19 distinct extensions affecting intransitive verbs, and found 6 of them to be incomplete.
-These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages. %, marked with bold print in \cref{fig:caribantree}.
-
-Interestingly, all six extensions featured innovative first person markers on \gl{s_a_} verbs.
-While extensions occurred with other person values as well, they never affected \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affected all potential targets.
-Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
-%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
-The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affecting them points to the split-\gl{s} system playing some role, so we will discuss its main properties in \cref{sec:split}.
-This will also give readers an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
-
-\input{floats/completed}
-
-\subsection{The Cariban split-\gl{s} system}
-\label{sec:split}
-As seen in \cref{sec:pc_person}, the \PC distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by two inflection classes within the \setone person paradigm, but this was not the only inflectional difference:
-%\begin{inlinelist}
-%	\item presence vs absence of the \gl{s_a_} marker \rc{w-}
-%	\item absence vs presence of the second person prefix \rc{ə(j)-} in imperatives
-%	\item presence vs absence of a derivational detransitivizing prefix
-%\end{inlinelist}.
-%
-Many languages show an \gl{s_a_} class marker in deverbalized forms, originating in \PC \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
-With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
-Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
-%
-\input{floats/participles}%
-%
-\input{floats/nominalizations}%
-%
-\input{floats/imperatives}
-%
-The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives, where the latter took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while the former received no prefix (both suffixed with \rc{-kə}).
-This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
-Both the \gl{s_a_} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
-%However, there was one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
-
-In the modern languages maintaining the split-\gl{s} system, mismatches between the semantics of verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with \kalina data in \exref[kar1]{kar2}.
-
-\pex<kar1> \kalina
-\a<kar-60>
-\begingl
-\gla mi-kupi-ja//
-\glb \gl{2}>\gl{3}-bathe-\gl{prs}//
-\glft \qu{You bathe him/her.} \parencite[][160]{hoff1968carib}//
-\endgl
-\a<kar-62>
-\begingl
-\gla a-kupi-ja//
-\glb \gl{3}>\gl{2}-bathe-\gl{prs}//
-\glft \qu{S/he bathes you.} \parencite[][63]{yamada2011evidentiality}//
-\endgl
-\xe
-%
-In \exref{kar1}, the choice between the second person \gl{a}- and \gl{p}-oriented markers \obj{mi-} and \obj{a-} depends on the scenario:
-The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
-The intransitive verbs in \exref{kar2} show the same person markers, but the choice of marker depends on the verb, rather than semantics.
-
-\pex<kar2> \kalina
-\a<kar-59>
-\begingl
-\gla sipi tɨnka-rɨ m-ekema-non hen//
-\glb net pull-\gl{nmlz} \gl{2}-be.afraid-\gl{prs}.\gl{uncert} eh?//
-\glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
-\endgl
-\a<kar-61>
-\begingl
-\glpreamble aya:woiya //
-\gla aj-awomɨ-ja//
-\glb \gl{2}-get.up-\gl{prs}//
-\glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
-\endgl
-\xe
-%
-Thus, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
-In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
-\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker must be considered volitional.
-\textcite{meira2000split} takes a sizable corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
-He shows that neither (non\-)activities, % \parencite[210]{meira2000split}, 
-(non-)agency, % \citeyearpar[212]{meira2000split}, 
-(in-)animacy, % \citeyearpar[213]{meira2000split},
-nor Aktionsart % \citeyearpar[215]{meira2000split} 
-satisfactorily predict the class membership of intransitive verbs in any of the languages.
-
-Rather, the reason for a verb to take \gl{a}- or \gl{p}-oriented prefixes is (at least diachronically) a morphological one.
-\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
-\begin{quotebox}{\parencite[201]{meira2000split}}
-	Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still exisiting transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
-\end{quotebox}
-\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
-\begin{inlinelist}
-	\item there are a few underived \gl{s_a_} verbs, with no detransitivizing prefix
-	\item \gl{s_a_} verbs can develop irregular semantics compared to their transitive counterparts
-	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{s_a_} verb
-	\item some originally derived \gl{s_a_} verbs have lost their transitive counterparts
-	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{s_a_})} from \qu{to dance with (\gl{tr})}
-\end{inlinelist}.
-He also notes that this leads to an inflectional split not based in meaning, but rather morphology:
-
-\begin{quotebox}{\parencite[226]{meira2000split}}
-	Apparently, the morphological behavior of the \gl{s_a_} verb class is an accidental consequence of the fact that detransitivization, as far back as we can reconstruct, entails all the morphology described […] as typical of \gl{s_a_} verbs.
-The alignment of person-marking prefixes appears not to be driven by any semantic forces in the language; it is as though they were being dragged by the evolution of the reflexive marker.
-\end{quotebox}
-
-As for the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although on verbs their reflexes have been merged in a single morpheme in modern languages.
-Modern reflexes of \detrz show a range of meanings, which can all be characterized as \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
-
-\ex<triodetrz> \trio \parencites[218--219]{meira2000split}[128, 256]{triomeira1999}\\
-\begin{tabular}[t]{@{}llll@{}}
-	\\
-	\obj{nonta}  & → & \obj{e-nonta}, & \qu{abandon each other}\\
-	\qu{abandon} & & \obj{əi-nonta} &  (reciprocal) \\
-	\\
-	\obj{suka} & → & \obj{e-suka}, & \qu{wash self}\\
-	\qu{wash} & & \obj{əi-suka} & (reflexive)\\
-	\\
-	\obj{pahka} & → & \obj{e-pahka} & \qu{break (\gl{intr})}\\
-	\qu{break (\gl{tr})} & & & (anticausative)\\
-	\\
-	\obj{puunəpɨ} & → & \obj{əh-puunəpɨ}, & \qu{think, meditate}\\
-	\qu{think about} & & \obj{əi-puunəpɨ} & (antipassive)\\
-\end{tabular}
-\xe
-%
-The morphological variation featured in \qu{to abandon each other} and \qu{to wash self} is due to the mentioned collapse between the two \PC prefixes:
-\obj{e-} is a reflex of the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
-However, both can occur with either meaning -- at least for these two verbs.
-%As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
-%\textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
-%\textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
-%Many of these verbs will be discussed in \cref{sec:verbs}.
-
-%\input{floats/underived2007.tex}
-
-\subsection{Summary and outlook}
-\label{sec:outlook}
-We identified \gl{s_a_} verbs that irregularly inflect for first person in 9 Cariban languages.
-These irregular forms are actually conservative, belonging to a few verbs unaffected by person marker extensions spread via lexical diffusion.
-Such conservative forms are only found among the first person forms of  (etymological) \gl{s_a_} verbs.
-The goal is to establish what verbs remained unaffected by the individual extensions and to find reasons for their conservativeness, for which we proceed as follows:
-The topic of \cref{sec:extensions} are the six incomplete innovations, the innovative markers they introduced, and the verbs they left untouched.
-Since these verbs show considerable etymological overlap between languages, they are reconstructed and discussed in more detail in \cref{sec:verbs}.
-Finally, we search for factors motivating the resistance of these verbs, discuss the findings and put them in a general context of language change and morphology in \cref{sec:discussion}.

### 2021-12-05 documents/sections/proto_carib.tex
data output
rewriting and restructuring

--- 2021-11-01 documents/sections/proto_carib.tex
@@
-\section{Cariban verbal person marking and marker extensions}
-The archaic first person markers investigated in this paper are the result of incomplete person marker extensions, i.e., modifications to the ancestral prefix system that did not affect the whole lexicon.
-To provide readers with the necessary background, we first introduce said ancestral system \pcref{sec:pc_person}, then the peculiarities of the Cariban split-\gl{s} system \pcref{sec:split}, and finally the concept of person marker extensions \pcref{sec:extensions_intro}.
+\section{The origins of conservative first person inflections}
+\label{sec:background}
+The irregular first person prefixes introduced in \cref{sec:intro} are relics, inherited from the ancestral \PC system \pcref{sec:pc_person}.
+That system was subject to many different kinds of innovations; responsible for the patterns discussed are person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
+A particular component of the system, the distinction between \gl{s_a_} and \gl{s_p_} verbs, plays a role in what extensions are incomplete and is the topic of \cref{sec:split}.
+%\cref{sec:outlook} summarizes the main points of this section and provides an outlook for the rest of the paper.
 
-\subsection{Verbal person marking in \PC}
+\subsection{\PC person marking and inflectional relics}
 \label{sec:pc_person}
-\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.%
-\footnote{We use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
-In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in an interlinearized glossed example.
-We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
-The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.\footnote{Glossing abbreviations: \glossingAbbrevsComma.}
+\PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.
+The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
 The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
 There was a single aliophoric marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
-Local scenarios were expressed in a non-transparent manner, by using the \gl{1+2} prefix \rc{k-} in both cases.%
-\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
-This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented in this table.}
+Local scenarios were expressed in a non-transparent manner, both using the \gl{1+2} prefix \rc{k-}.
 
-\begin{table}
-	\centering
-	\caption[\PC \setone (main clause) person markers]{\PC \setone (main clause) person markers \parencites[495]{meira2010origin}[497]{gildea2016referential}}
-	\label{tab:pcpers}
-\begin{subtable}[b]{.49\linewidth}
-\caption{Transitive}
-\label{tab:pctrans}
-\centering
-	\begin{tabular}{@{}lllll@{}}
-	\mytoprule
-\gl{a}/\gl{p}&		\gl{1}	&	\gl{2}		&	\gl{1+2}	&	\gl{3}	\\
-\mymidrule
-\gl{1}	&		&	\rc{k-}	&				&	\rc{t(i)-}		\\	
-\gl{2}	&	\rc{k-}			&&				&	\rc{m(i)-}		\\
-\gl{1+2}&		&				&				&	\rc{kɨt(i)-}		\\
-\gl{3}	&	\rc{u(j)-}	&	\rc{ə(j)-}	&	\rc{k-}			&	\rc{n(i)-}		\\
-	\mybottomrule
-	\end{tabular}
-\end{subtable}%
-\begin{subtable}[b]{.49\linewidth}
-\caption{Intransitive}
-\label{tab:pcintrans}
-\centering
-\begin{tabular}{@{}lll@{}}
-\mytoprule
-& \gl{s_a_} & \gl{s_p_}  \\
-\mymidrule
-\gl{1} & \rc{w-} & \rc{u(j)-} \\
-\gl{2} & \rc{m-} & \rc{ə(j)-}\\
-\gl{1+2} & \rc{kɨt-} & \rc{k-}\\
-\gl{3} & \rc{n-} & \rc{n(i)-}\\
-\mybottomrule
-\end{tabular}	
-\end{subtable}
-\end{table}
+\input{floats/pcpers}
 
@@
 
-As in many other split-\gl{s} systems found in languages of the world, the intransitive verbal lexicon was divided into two classes, with verbs inherently being members of the \gl{s_a_} or the \gl{s_p_} group.
-That is, while in transitive verbs, the choice of person marker had a crucial semantic contribution, it was predictable in intransitive verbs.
-This is illustrated with modern \kalina data in \exref[kar1]{kar2}.
+Equipped with this knowledge about the ancestral system, it becomes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \textsc{conservative}.
+They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
+The reflexes of \rc{w-} may be considered \textsc{relics}, old and restricted to specific lexically conditioned contexts, contrasting with the innovative prefixes found elsewhere.
+These verbs and their prefixes are comparable with the few English nouns like \obj{ɒks}, which preserve the old plural suffix \obj{-ən}.
+It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{ɔks-n} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
+Since the irregular \hixka and \trio prefixes are conservative and the regular prefixes are innovative, the next question to be addressed is where these new prefixes came from.
+
+%\input{floats/makintro}
+
+\subsection{Person marker extensions in intransitive verbs}
+\label{sec:extensions_intro}
+In his discussion of the \PC split-\gl{s} system and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different modifications in various languages.
+The main mechanism of change leading to these modifications are \textbf{person marker extensions}, i.e.
+the use of verbal person prefixes being extended to contexts previously occupied by other prefixes.
+%\footnote{
+%In principle, one could also imagine a scenario in which verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
+%However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
+%Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
+%For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
+There have been many different person marker extensions in Cariban languages, and some are still ongoing.
+This is illustrated by \textcite{gildea1998}, using the three Parukotoan languages as an example.
+Apart from segmental changes to individual morphemes, the following innovations happened in the \setone paradigm in Parukotoan:
+
+\begin{enumerate}
+	\item \PPar \begin{enumerate}
+		\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
+		\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+	\end{enumerate}
+	\item \PWai \begin{enumerate}
+		\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
+		\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
+	\end{enumerate}
+	\item \waiwai \begin{enumerate}
+		\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
+\end{enumerate}
+%
+All innovations are person marker extensions except 2b, which combined a pronoun with the linker \rc{j-}.
+They are printed in bold in \cref{fig:par_ext}, which reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
+%
+\begin{figure}[hbt]
+	\centering
+	\setlength{\tabcolsep}{2pt}
+	\fbox{\begin{tikzpicture}[
+			every node/.append style={align=center},
+			level distance=80pt,
+			]
+			\Tree[.{\PC{}\pcone}
+			[.{\PPar{}\ppone}
+			{\kaxui{}\kaxone}
+			[.{\PWai{}\pwone}
+			{\hixka{}\hixone}
+			{\waiwai{}\waione}
+			]
+			]
+			]
+	\end{tikzpicture}}
+	\caption{Person marking extensions in Parukotoan, after \textcite[94]{gildea1998}}
+	\label{fig:par_ext}
+\end{figure}
+%
+\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows the variation in the first person and \gl{1+2} prefixes that is reconstructible to \PPar.
+\waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
+%In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
+%That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
+%Similarly, the \gl{2}\gl{s_p_} prefix \obj{a-} was extended to imperatives of (former) \gl{s_a_} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.
+
+%Such extensions, often only affecting one person value at a time, and the eventual loss of split-\gl{s} are not entirely unexpected, given that the system lacks any semantic basis.
+
+\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment; for our story of conservative verbs we will zoom in on a particular aspect of these extensions.
+To begin, we argue that they took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
+Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
+In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which is what one would expect from a change in progress.
+%todo maybe put in GOOD kaxui data
+Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} convincingly argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
+Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
+
+Our third argument in favor of the lexical diffusion scenario brings us back to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
+In both cases, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
+We interpret the continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
+In our family-wide investigation of person marker extensions, we identified 19 distinct extensions affecting intransitive verbs, and found 6 of them to be incomplete.
+These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages. %, marked with bold print in \cref{fig:caribantree}.
+
+Interestingly, all six extensions featured innovative first person markers on \gl{s_a_} verbs.
+While extensions occurred with other person values as well, they never affected \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affected all potential targets.
+Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
+%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
+The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affecting them points to the split-\gl{s} system playing some role, so we will discuss its main properties in \cref{sec:split}.
+This will also give readers an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
+
+\input{floats/completed}
+
+\subsection{The Cariban split-\gl{s} system}
+\label{sec:split}
+As seen in \cref{sec:pc_person}, the \PC distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by two inflection classes within the \setone person paradigm, but this was not the only inflectional difference:
+%\begin{inlinelist}
+%	\item presence vs absence of the \gl{s_a_} marker \rc{w-}
+%	\item absence vs presence of the second person prefix \rc{ə(j)-} in imperatives
+%	\item presence vs absence of a derivational detransitivizing prefix
+%\end{inlinelist}.
+%
+Many languages show an \gl{s_a_} class marker in deverbalized forms, originating in \PC \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
+With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
+Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
+%
+\input{floats/participles}%
+%
+\input{floats/nominalizations}%
+%
+\input{floats/imperatives}
+%
+The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives, where the latter took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while the former received no prefix (both suffixed with \rc{-kə}).
+This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
+Both the \gl{s_a_} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
+%However, there was one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
+
+In the modern languages maintaining the split-\gl{s} system, mismatches between the semantics of verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with \kalina data in \exref[kar1]{kar2}.
 
@@
 The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
-While intransitive verbs show the same (or very similar) person markers, they contribute no semantic difference here \exref{kar2}.
+The intransitive verbs in \exref{kar2} show the same person markers, but the choice of marker depends on the verb, rather than semantics.
 
@@
 \begingl
-\gla aj-awoi-ja//
+\glpreamble aya:woiya //
+\gla aj-awomɨ-ja//
 \glb \gl{2}-get.up-\gl{prs}//
@@
 %
-Rather, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.\footnote{The root \obj{awomɨ} \qu{to get up} is subject to syllable reduction and assimilation to the prefix-initial \obj{j}.}
-What is more, the choice of prefix does not necessarily convey information about the semantics of the verb, since there are clear mismatches:
+Thus, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
 \qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker must be considered volitional.
-These mismatches are not isolated cases, as will be discussed in \cref{sec:split}.
-
-\subsection{Defining features and origins of the split-\gl{s} system}
-\label{sec:split}
-As seen in the previous section, the split-\gl{s} system defined two inflectional classes for intransitive verbs within the \setone system.
-%\kalina paradigms illustrating the split for each person are given in \cref{tab:kalinaparadigms}.
-%\gl{3}\gl{s_a_} differs from \gl{3}\gl{s_p_} by \begin{inlinelist}
-%\item not showing the allomorph \obj{ni-} before consonants
-%\item not triggering the umlaut of verb-initial \phon{\obj{o}}{\obj{e}}
-%\end{inlinelist}, which is shown in \cref{tab:kalinaparadigms}.
-%
-%\begin{table}[h]
-%	\centering
-%	\caption{\kalina \gl{s_a_} and \gl{s_p_} person paradigms \parencites[178]{hoff1968carib}[430, 73]{courtz2008carib}[481]{meira2010origin}}
-%	\label{tab:kalinaparadigms}
-%	\begin{tabular}{@{}lll@{}}
-%	\mytoprule
-%Person & \gl{s_a_} \obj{opɨ} \qu{to come} & \gl{s_p_} \obj{[o/e]kanumɨ} \qu{to run}\\
-%	\mymidrule
-%\gl{1} & \obj{w-opɨ-} & \obj{j-ekanumɨ-}\\
-%\gl{2} & \obj{m-opɨ-} & \obj{aj-ekanumɨ-}\\
-%\gl{1+2} & \obj{kɨt-opɨ-} & \obj{k-okanumɨ-}\\
-%\gl{3} & \obj{n-opɨ-} & \obj{n-ekajumɨ-}\\
-%	\mybottomrule
-%	\end{tabular}
-%\end{table}
-%
-However, there were some other morphological criteria distinguishing \gl{s_a_} from \gl{s_p_} verbs in \PC:
-\begin{inlinelist}
-\item presence vs absence of the \gl{s_a_} marker \rc{w-}
-\item absence vs presence of the second person prefix \rc{ə(j)-} in imperatives
-\item presence vs absence of a derivational detransitivizing prefix
-\end{inlinelist}.
-
-Many languages show an \gl{s_a_} class marker in deverbalized forms, which can be reconstructed to \PC as \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
-With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
-Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.
-%\kaxui does not preserve \rc{w-} in its participles (\obj{tehurkat͡ʃe} \qu{fallen}, not \ungr{tɨwehurkat͡ʃe}), while \apalai only preserves it there, but not in nominalizations (\obj{jepɨtopo} \qu{my bathing place}, not \ungr{joepɨtopo}).
-
-\input{floats/participles}
-
-\input{floats/nominalizations}
-
-The distinction between \gl{s_a_} and \gl{s_p_} is also borne out in imperatives.
-Here, \gl{s_p_} verbs took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while \gl{s_a_} verbs were unprefixed (both suffixed with \rc{-kə}).
-This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
-As in the case of the \gl{s_a_} marker \rc{w-} participles and nominalizations, some languages have lost the distinction between \gl{s_a_} and \gl{s_p_} verbs in imperatives, for example \panare.
-
-\input{floats/imperatives}
-
-%\subsubsection{The detransitivizing prefix}
-There is one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
-As mentioned in \cref{sec:pc_person}, mismatches between the semantics of intransitive verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common.
-However, the Cariban split-\gl{s} system goes further than all other known such systems, in that its division of the verbal lexicon does not follow any discernible semantic criteria whatsoever.
 \textcite{meira2000split} takes a sizable corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
@@
 However, both can occur with either meaning -- at least for these two verbs.
+%As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
+%\textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
+%\textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
+%Many of these verbs will be discussed in \cref{sec:verbs}.
 
-%\subsubsection{Underived \gl{s_a_} verbs}
-As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
-\textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
-\textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
-Many of these verbs will be discussed in \cref{sec:verbs}.
-
-\begin{table}
-	\centering
-	\caption[Underived \PC \gl{s_a_} verbs]{Underived \PC \gl{s_a_} verbs \parencite[30]{gildea2007greenberg}}
-	\label{tab:underived2007}
-	\begin{tabular}{@{}lll@{}}
-	\mytoprule
-Form & Meaning\\
-\mymidrule
-\rc{tə[mə]} & \qu{to go}\\
-\rc{ətepɨ} & \qu{to come\subs{1}}\\
-\rc{ka[ti]} & \qu{to say}\\
-\rc{əmə[mɨ]} & \qu{to enter}\\
-\rc{eti} & \qu{to dwell, be\subs{2}}\\
-\rc{a[p]} & \qu{to be\subs{1}, say}\\
-\rc{əməkɨ} & \qu{to come\subs{2}}\\
-	\mybottomrule
-	\end{tabular}
-\end{table}
-
-\subsection{Person marker extensions in intransitive verbs}
-\label{sec:extensions_intro}
-The original \PC split-\gl{s} system has been subject to change in many languages, in some even to the point of total loss.
-Besides modifications like losing the \gl{s_a_} class marker \rc{w-} or losing certain person values altogether, these changes have largely been due to person prefixes being extended to new verbs.\footnote{
-In principle, one could also imagine a scenario in which verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
-However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
-Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
-For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
-There have been many such person marker extensions in Cariban languages, and some are still ongoing.
-This is shown by \textcite{gildea1998}, using the three Parukotoan languages as an example.
-We have reproduced his tables as a tree diagram in \cref{fig:par_ext}, with adapted transcription -- and in the case of \kaxui, the addition of ∅/\obj{j-} as an alternative \gl{1}\gl{s_p_} marker \perscommpar{Spike Gildea}.
-Apart from segmental changes to individual morphemes, the following restructuring innovations happened in the \setone paradigm in Parukotoan:
-
-\begin{enumerate}
-\item \PPar \begin{enumerate}
-	\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
-	\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
-	\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
-\end{enumerate}
-\item \PWai \begin{enumerate}
-\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
-\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
-\end{enumerate}
-\item \waiwai \begin{enumerate}
-\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
-\end{enumerate}
-
-\begin{figure}[hbt]
-\centering
-\setlength{\tabcolsep}{2pt}
-\begin{tikzpicture}[
-	every node/.append style={align=center},
-	level distance=80pt,
-]
-\Tree[.{\PC{}\pcone}
-[.{\PPar{}\ppone}
-{\kaxui{}\kaxone}
-[.{\PWai{}\pwone}
-{\hixka{}\hixone}
-{\waiwai{}\waione}
-]
-]
-]
-\end{tikzpicture}
-  \caption{Person marking extensions in Parukotoan, after \textcite[94]{gildea1998}}
-  \label{fig:par_ext}
-\end{figure}
-
-\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows variation in the first person and \gl{1+2} prefixes.
-\waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
-In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
-That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
-Similarly, the \gl{2}\gl{s_p_} prefix \obj{a-} was extended to imperatives of (former) \gl{s_a_} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.
-
-\pex<wailoss>
-\begin{multicols}{2}
-\a<wai-158> \waiwai \parencite[][98]{waiwaihawkins1998}\\
-\begingl
-\gla k-eɸɨrka-t͡ʃhe//
-\glb \gl{1+2}-fall-\gl{advz}.after//
-\glft \qu{after our fall}//
-\endgl
-\a<kax-138> \kaxui \parencite[][49]{schuring2018kaxuyana}\\
-\begingl
-\gla ku-w-ehurka-tpɨrɨ//
-\glb \gl{1+2}-\gl{s_a_}-fall-\gl{nmlz}.\gl{pst}.\gl{pert}//
-\glft \qu{our fall}//
-\endgl
-\a<wai-112> \waiwai \parencite[][177]{waiwaihawkins1998}\\
-\begingl
-\gla a-mo-ko//
-\glb \gl{2}-come-\gl{imp}//
-\glft \qu{come!}//
-\endgl
-\a<hix-43> \hixka \parencite[][191]{hixkaryanaderby1985}\\
-\begingl
-\gla m-omokɨ-no//
-\glb \gl{2}\gl{s_a_}-come-\gl{imm}//
-\glft \qu{You have come.}//
-\endgl
-\end{multicols}
-\xe
-
-%Such extensions, often only affecting one person value at a time, and the eventual loss of split-\gl{s} are not entirely unexpected, given that the system lacks any semantic basis.
-
-While different cases of loss of split-\gl{s} are discussed by \textcite[91--96]{gildea1998}, this paper focuses on a so far neglected aspect of the person marking extensions resulting in this loss.
-As a prerequisite for our story, we argue that they are executed via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
-First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
-Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
-In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which is what one would expect from a change in progress.
-%todo put in GOOD kaxui data
-Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
-Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
-Finally, the process of lexical diffusion is seen most clearly where it was incomplete.
-Not all person marker extensions spread through the entire lexicon, but rather stopped short of some verbs.
-This is illustrated for \trio first person forms of \gl{s_a_} verbs in \exref{trio1}; regular \gl{s_a_} verbs take phonologically determined allomorphs \obj{s-} (\envr{}{\obj{e}}) and \obj{t-} (\envr{}{\obj{ə}}), but \obj{tən} \qu{to go} takes an \dbqu{archaic} or \dbqu{irregular} prefix \obj{wɨ-}; it was not affected by the spread of innovative \obj{t-}/\obj{s-} and therefore preserves the old prefix.
-
-\ex<trio1> \trio \gl{1}\gl{s_a_} verbs \parencite[292, 294]{triomeira1999}\\
-\begin{tabular}[t]{@{}ll@{}}
-\obj{s-epɨ} & \qu{I bathed} \\
-\obj{s-entapo} & \qu{I yawned}\\
-\obj{t-əturu} & \qu{I talked}\\
-\obj{t-əənɨkɨ} & \qu{I slept}\\
-\obj{wɨ-tən} & \qu{I went}\\
-\end{tabular}
-\xe
-%
-These incomplete extensions and the verbs that were not affected by them are the central topic of this paper.
-
+%\input{floats/underived2007.tex}
 %
 %
-%For instance, while most \gl{s_a_} verbs in \hixka and \waiwai have a first person marker \obj{k-}, not all do:
-%
-%\ex<TAG> \waiwai \qu{to say} \perscommpar{Spike Gildea}\\
-%\begin{tabular}[t]{@{}ll@{}}
-%\gl{1} & \obj{wɨɨ-ka-}\\
-%\gl{2} & \obj{mɨɨ-ka-}\\
-%\gl{1+2} & \obj{tɨt-ka-}\\
-%\gl{3} & \obj{nɨɨ-ka-}\\
-%\end{tabular}
-%\xe
-%%
-%From a synchronic perspective, \waiwai \qu{to say} has an irregular first person marker \obj{wɨ-} instead of regular \obj{kɨ-}.
-%From a diachronic perspective, this verb is archaic, preserving the old \gl{1}\gl{s_a_} marker \rc{w-}.
-
-That is, apparently ongoing changes like the situation discussed above for \kaxui will not be investigated in detail.
-The same is true for extensions that affected the entire verbal lexicon, although we will briefly comment on them.
-We have investigated the 19 person marker extensions we are aware of, but only 6 of them left a group of irregularly or archaically inflected verbs.
-Interestingly, these verbs are always \gl{s_a_} verbs, and the irregular marker is always a first person one.
-While there have been extensions for other person values as well, they never affect \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affect the entire lexicon, at least based on the attested data.
-Illustrative examples for extensions affecting the entire lexicon are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set in \waimiri \pcref{tab:waimiri}.%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
-
-\begin{table}
-\caption[Some examples for completed extensions]{Some examples for completed extensions \parencite[90--92]{gildea1998}}
-\label{tab:completed}
-\begin{subtable}[h]{0.24\textwidth}
-\centering
-\caption{\apalai}
-\label{tab:apalai}
-\begin{tabular}{@{}lll@{}}
-\mytoprule
-& \gl{s_a_} & \gl{s_p_}\\
-\mymidrule
-\gl{1} & ∅/\obj{ɨ-} & \obj{ɨ-}/\obj{j-}\\
-\gl{2} & \obj{m(ɨ)-} & \obj{o-}\\
-\gl{1+2} & \multicolumn{2}{c}{\emp{\obj{s(ɨ)-}}}\\
-\gl{3} & \multicolumn{2}{c}{\obj{n(ɨ)-}}\\
-\mybottomrule
-\end{tabular}
-\end{subtable}
-\hfill
-\begin{subtable}[h]{0.24\textwidth}
-\centering
-\caption{\panare}
-\label{tab:panare}
-\begin{tabular}{@{}lll@{}}
-\mytoprule
-& \gl{s_a_} & \gl{s_p_}\\
-\mymidrule
-\gl{1} & \obj{w(ɨ)-} & ∅/\obj{j-}\\
-\gl{2} & \multicolumn{2}{c}{\emp{\obj{m(ɨ)-}}}\\
-\gl{1+2} & \multicolumn{2}{c}{\obj{n(ɨ)-}}\\
-\gl{3} & \multicolumn{2}{c}{\obj{n(ɨ)-}}\\
-\mybottomrule
-\end{tabular}
-\end{subtable}
-\hfill
-%\begin{subtable}[h]{0.24\textwidth}
-%\centering
-%\caption{\pemon}
-%\label{tab:pemon}
-%\begin{tabular}{@{}lll@{}}
-%\mytoprule
-%& \gl{s_a_} & \gl{s_p_}\\
-%\mymidrule
-%\gl{1} & \multicolumn{2}{c}{\emp{∅}}\\
-%\gl{2} & \multicolumn{2}{c}{\emp{\obj{m-}}}\\
-%\gl{1+2} & \multicolumn{2}{c}{\obj{n-}}\\
-%\gl{3} & \multicolumn{2}{c}{\obj{n-}}\\
-%\mybottomrule
-%\end{tabular}
-%\end{subtable}
-%\hfill
-\begin{subtable}[h]{0.24\textwidth}
-\centering
-\caption{\waimiri}
-\label{tab:waimiri}
-\begin{tabular}{@{}ll@{}}
-\mytoprule
-& \gl{s}\\
-\mymidrule
-\gl{1} & \emp{\obj{w(ɨ)-}/\obj{i-}}\\
-\gl{2} & \emp{\obj{m(ɨ)-}}\\
-\gl{1+2} & \emp{\obj{h(ɨ)-}}\\
-\gl{3} & \obj{n-}/∅\\
-\mybottomrule
-\end{tabular}
-\end{subtable}
-\end{table}
-
-The remainder of this paper is structured as follows:
-In \cref{sec:extensions}, we discuss each of these six incomplete extensions individually, discussing all attested unaffected verbs, and reconstructing proto-paradigms where necessary.
-In \cref{sec:verbs}, we take a comparative look at the verbs that were not affected by these extensions.
-Finally, we search for factors motivating the resistance of these verbs, discuss the findings and put them in a general context of language change and morphology in \cref{sec:discussion}.
-
-%For instance, while the extension of \PWai \gl{1}\gl{s_p_} \rc{k-} affected most verbs, like \rc{eɸurka} \qu{to fall}, verbs like \rc{ka[ti]} \qu{to say} were not affected \exref{pwaiext}.
-
-%\ex<pwaiext> Incomplete extension of \PWai \rc{k-} to \gl{1}\gl{s_a_}\\
-%Sources: \textcites[510]{howard2001wrought}[227, 39, 4, 198, 249]{hixkaryanaderby1985}[175, 174, 43, 55, 60]{waiwaihawkins1998} \\
-%\begin{tabular}[t]{@{}llll@{}}
-%& \PWai & \hixka & \waiwai \\
-%\qu{fall} & \rc{k-eɸurka-} & \obj{k-ehurka-} & \obj{k-eɸɨrka}\\
-%\qu{choke self} & \rc{k-oseʃewnuk-} & \obj{k-oseʃemnuk-} & \obj{k-eʃeʃewnuk-}\\
-%\qu{treat self medically} & \rc{k-osoht͡ʃema-} & \obj{k-osoht͡ʃema-} & \obj{k-eseht͡ʃem-}\\
-%\qu{come} & \rc{k-omokɨ-} & \obj{k-omokɨ-} & \obj{k-mok-}\\
-%\qu{say} & \rc{wɨ-ka-} & \obj{ɨ-ka-} & \obj{wɨɨ-ka-}\\
-%\qu{be} & \rc{w-eʃi-} & \obj{w-eʃe-} & \obj{w-eeʃi-}\\
-%\end{tabular}
-%\xe
-%
+%\subsection{Summary and outlook}
+%\label{sec:outlook}
+%\gl{s_a_} verbs irregularly inflected for first person were identified in 9 Cariban languages.
+%These irregular forms are actually conservative, belonging to a few verbs unaffected by person marker extensions spread via lexical diffusion.
+%Such conservative forms are only found among the first person forms of  (etymological) \gl{s_a_} verbs.

### 2021-12-05 documents/sections/verbs.tex
rewriting and restructuring

@@
-\section{Conservative verbs: the usual suspects}
+\subsection{Conservative verbs in comparison}
 \label{sec:verbs}
-In \cref{sec:extensions}, we introduced six distinct extensions of personal prefixes into \gl{1}\gl{s_a_} territory, and identified verbs resistant to each innovation.
+In \cref{sec:extensions}, six incomplete extensions of personal prefixes into \gl{1}\gl{s_a_} territory were presented, and the verbs resistant to these innovations were identified.
 The set of unaffected verbs is rather small in most cases, and (proto-)languages show a considerable degree of overlap in what verbs are unaffected.
-Therefore, this section presents these verbs from a comparative perspective.
+Therefore, this section investigates these verbs from a comparative perspective.
 \cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}.
-\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek, and  \cref{sec:bathe} investigates \rc{e-pɨ} \qu{to bathe}, of which the \PPek reflex \rc{i-pɨ} resisted the extension of \rc{k-}.
+\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek, and  \cref{sec:bathe} investigates \rc{e-pɨ} \qu{to bathe}, of which the \PPek reflex \rc{ipɨ} resisted the extension of \rc{k-}.
 The \obj{e}-initial verbs not affected by the extension of \obj{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
 
-\subsection{\rc{eti} and \rc{a[p]} \qu{to be}}
+\subsubsection{\rc{eti} and \rc{a[p]} \qu{to be}}
 \label{sec:be}
-For a comprehensive comparative overview of these two roots, we refer the reader to \textcite[375--382]{gildea2018reconstructing}; we will not discuss them in detail here.
+For a comprehensive comparative overview of these two roots, readers are refferd to \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
 \rc{a[p]} is the original copula and can be reconstructed as already having various irregularities in \PC.
-\rc{eti}, the other root, is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
+\rc{eti}, the other root, is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
 Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam} value.
@@
 
-\subsection{\rc{ka[ti]} \qu{to say}}
+\subsubsection{\rc{ka[ti]} \qu{to say}}
 \label{sec:say}
@@
 This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}).
-We do not know the first person form of its \yukpa reflex \obj{ka}.
+The first person form of its \yukpa reflex \obj{ka} is unattested.
 
-As briefly mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes this verb as transitive in \hixka.
-This analytical choice is not only motivated by the desire to avoid an idiosyncratic intransitive first person prefix \obj{ɨ-} instead of \obj{kɨ-}.
+As mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes this verb as transitive in \hixka.
+This analytical choice is not only motivated by a desire to avoid an idiosyncratic intransitive first person prefix \obj{ɨ-} instead of regular \obj{kɨ-}.
 \hixka \obj{ka} also shows the complementary distribution of third person  \obj{n-} and preceding objects typical of transitive verbs in Cariban \parencite[60--81]{gildea1998}.
-Due to its semantics, these objects are either direct speech \exref{hixsay} or ideophones.
+Due to the semantics of \qu{to say}, these objects are either ideophones or direct speech \exref{hixsay}.
 
@@
 In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
-At least the \trio cognate shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
+Looking beyond \hixka, at least the \trio cognate shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
 
@@
 
-Our categorization of \qu{to say} as an intransitive verb is supported primarily by its person prefixes.
-\kalina offers a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.\footnote{Interestingly, the \kalina causativized form \obj{kapo} \qu{to make say} does not have the regular \gl{1}>\gl{3} prefix \obj{s(i)-}, but irregular \obj{w(ɨ)-} \parencite[430]{courtz2008carib}.}
-Similarly, \qu{to say} in Pekodian languages has a reflex of \gl{1}\gl{s} \rc{w-} \pcref{sec:pekodian}, and not \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
-Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show no \obj{i} for this verb, see \exref{sayintr.aku-154} as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for the verb's inflection in individual languages.
+The categorization of \qu{to say} as an intransitive verb is supported primarily by its person prefixes.
+\kalina offers a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.
+Similarly, \qu{to say} has a reflex of \gl{1}\gl{s_a_} \rc{w-} in Pekodian languages \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
+Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show \obj{ɨ} instead of \obj{i} for this verb, see \exref{sayintr.aku-154} as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for the verb's inflection in such languages.
 Finally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and it is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
@@
 
-Summing up, this verb could be reconstructed as being intransitive based on its prefixes, but transitive based on some suffixes.
+In summary, this verb can be reconstructed as being intransitive based on its prefixes, but transitive based on some suffixes.
 \hixka has lost the main intransitive criteria, making its reflex look more like a transitive verb.
@@
 
-\subsection{\rc{ɨtə[mə]} \qu{to go}}
+\subsubsection{\rc{ɨtə[mə]} \qu{to go}}
 \label{sec:go}
@@
 
-\subsection{\rc{(ət-)jəpɨ} \qu{to come}}
+\subsubsection{\rc{(ət-)jəpɨ} \qu{to come}}
 \label{sec:come}
@@
 This verb is reconstructed as \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
-Crucially, the majority do not reflect the \rc{ət} part of their reconstruction, and many forms are ostensibly reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ} for the \rc{epɨ} part.
-We analyze all forms as going back to a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
+Crucially, the majority do not reflect the \rc{ət} part of their reconstruction, and many forms are ostensibly reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ}, rather than \rc{epɨ}.
+A unifying account of all these forms is made possible by reconstructing a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
 
-Evidence for the originally \rc{j}-initial nature of the root is found in the Pemongan languages and \kaxui, although the coexistence of \obj{oohɨ} and \obj{johɨ} in \kaxui raises the question of whether the \obj{j} in the latter form is actually part of the root.
+Evidence for the originally \rc{j}-initial nature of the root is found only in the Pemongan languages and \kaxui, although the coexistence of the forms \obj{oohɨ} and \obj{johɨ} in \kaxui raises the question of whether the \obj{j} in the latter form is actually part of the root.
 \obj{johɨ} only occurs with third person in the Progressive \exref{kaxprog.kax-82}, and one might analyze \obj{j(-)} as a reflex of \rc{i-w-} \qu{\gl{3}-\gl{s_a_}-}, since the \gl{s_a_} class marker \rc{w-} is present with other person values \exref{kaxprog.kax-54}.
 However, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial \gl{s_a_} verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
-Thus, the best explanation for this \obj{j} is that is indeed part of the root, rather than an outcome of \rc{i-w-}, which in turn allows us to safely reconstruct \rc{j} back to \PC.
+Thus, the best explanation for this \obj{j} is that is indeed part of the root, rather than an outcome of \rc{i-w-}, which in turn permits to confidently reconstruct \rc{j} back to \PC.
 
@@
 %
-Most of the longer forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for the segment \rc{j}, but the \obj{i} in the \akawaio form \obj{əsipɨ} is very likely a reflex of the sequence \rc{jə}.
+Most of the morphologically complex forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for the segment \rc{j}, but the \obj{i} in the \akawaio form \obj{əsipɨ} is very likely a reflex of the sequence \rc{jə}.
 This analysis is supported by the reflex \obj{ipɨ} from bare \rc{jəpɨ} in very closely related \macushi.
 
-Turning to the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ}, we find that both are distributed widely in the family, sometimes even co-occurring in the same language.
+Turning to the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ}, it becomes evident are distributed widely in the family, sometimes even co-occurring in the same language.
 A unifying account of these forms requires the root \rc{jəpɨ} to undergo two major sound changes: \begin{inlinelist}
@@
  \end{inlinelist}.
-Both phenomena are found in other contexts throughout the family \parencite{meira2010origin}.
-However, these sound changes appear to have applied irregularly to this verb, and not always in the same order.
+While both sound changes are found in other contexts throughout the family \parencite{meira2010origin}, they appear to have applied irregularly to this verb, and not always in the same order.
 For example, the \kalina form \obj{opɨ} can only be explained if \rc{j} was lost before the umlaut of \rc{ə} to \rc{e}, which would have been triggered by \rc{j}.
 On the other hand, forms like \maqui \obj{ehə} must be the result of \phonl{\rc{ə}}{\rc{e}}{\rc{j}}, with subsequent loss of \rc{j}.
-The \akuriyo form \obj{eepɨ} looks like a reflex thereof as well, but the length is unexpected, and is analyzed by \textcite[]{meira1998proto} as reflecting an earlier diphthong \rc{əe}.
+The \akuriyo form \obj{eepɨ} looks like a reflex thereof as well, but the length is unexpected, and is analyzed by \textcite[]{meira1998proto} as a coalescence of \rc{e} and \rc{əe}.
 
@@
 	Apart from aforementioned \akuriyo \obj{eepɨ}, another diachronically irregular form is \apalai \obj{oepɨ}, where one would expect \rc{ət-epɨ} to yield \obj{os-epɨ} \parencite[506]{meira2010origin}.
-	Similarly, while \obj{oepɨ} would be a regular outcome of hypothetical \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
+	Similarly, while \obj{oepɨ} would be a regular outcome of a hypothetical \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
 	One possibility is that the form is due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}.
@@
 To illustrate, the \trio \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.%
-\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious.}
+\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious \parencite[115]{meira1998proto}.}
 It should be noted that forms with and without \rc{ət-} in different languages are not triggered by the same person values.
@@
 
-Our interpretation of the \rc{ət} part as a detransitivizer is based on its form and on the paradigmatically conditioned occurrence in some languages.
-Although the combination of a detransitivizer an an intransitive verb seems semantically illogical, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
+The interpretation of the \rc{ət} part as a detransitivizer is justifiable due to its form and its paradigmatically conditioned occurrence in some languages.
+Although the combination of a detransitivizer and an intransitive verb seems semantically illogical, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
 For example, the \PC \gl{s_p_} verb \rc{wɨnɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
 Also, \waiwai \qu{go to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced by its status in \kaxui, \kalina, \arara, \trio, and \panare \exref{pan-128}.
+The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced the class membership of its reflexes in \kaxui, \kalina, \arara, \trio, and \panare \exref{pan-128}.
 
@@
 Summing up, this verb is highly irregular, both from a synchronic and diachronic perspective.
-The scenario we suggest involves reflexes of the detransitivizer \rc{ət(e)-} being optionally added to an \gl{s_a_} verb root \rc{jəpɨ}, which further underwent umlaut and loss of \rc{j}, but in no systematic manner, resulting in the chaotic picture in \cref{tab:come}.
+The scenario suggested here involves reflexes of the detransitivizer \rc{ət(e)-} being optionally added to an \gl{s_a_} verb root \rc{jəpɨ}, which further underwent umlaut and loss of \rc{j}, but in no systematic manner, resulting in the chaotic picture in \cref{tab:come}.
 As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}, but not on the \arara reflex of \rc{jəpɨ}.
-Reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ} and/or \rc{jəpɨ} (\akuriyo) resisted the introduction of \PTir \rc{t-}.
+Reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ}/\rc{jəpɨ} (\akuriyo) resisted the introduction of \PTir \rc{t-}.
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
@@
 
-\subsection{\rc{ɨpɨtə} \qu{to go down}}
+\subsubsection{\rc{ɨpɨtə} \qu{to go down}}
 \label{sec:godown}
@@
 When \akuriyo extended \obj{k-}, the verb already had an irregularly inflected first person form with \obj{p-}, inherited from \PTir.
-At first sight, it may seem that it also was affected by the two independent extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
+At first sight, it may seem that it was also affected by the two independent extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
 
@@
 However, a broader comparative perspective reveals a much more complicated story.
-\cref{tab:godown} shows all attested reflexes of \rc{ɨpɨtə}, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.\footnote{Our reconstruction of \PPek \rc{ɨptə} treats the additional elements found in the daughter languages as non-cognate.
+\cref{tab:godown} shows all attested reflexes of \rc{ɨpɨtə}, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.%
+\footnote{The reconstruction of \PPek \rc{ɨptə} in \cref{tab:godown} treats the suffixed elements found in the daughter languages as non-cognate.
 \textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}, and at least the addition of a final \obj{ŋ} in \PXin is attested elsewhere:
@@
 
-In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb in our synchronic analysis of \wayana.
+In one language, \wayana, the verb shows traits of both classes, necessitating an analysis as a \dbqu{mixed} verb in a synchronic description of \wayana.
 It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
@@
 Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the restriction of \rc{-ka} to \gl{s_p_} verbs in \PC \parencite{gildea2019overview} points to \gl{s_p_} membership.
-These patterns lead us to posit the hypothesis that the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class, reflected only in its \gl{1+2} prefix and the \gl{s_a_} class marker.
-This in turn implies that reflexes of this verb with \gl{s_a_} membership in other languages fully switched from \gl{s_p_} at some point.
+These patterns are suggestive of a scenario in which the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class, reflected only in its \gl{1+2} prefix and the \gl{s_a_} class marker.
+This in turn implies that reflexes of this verb with \gl{s_a_} membership in other languages fully switched from \gl{s_p_} in their inflectional patterns.
 
@@
 Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms that also feature a reflex of \rc{en-} \parencite[263]{triomeira1999}.
-Thus, it seems that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.
+Thus, it appears that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.
 
 These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by the extensions in \PWai, \PTir, \carijo, and \yukpa.
-For \PTir, we cannot establish a relative chronology of the verb class switch, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
-For \PWai, we lack knowledge not only about the first person form, but even about class membership.
-Similarly, for \carijo and \yukpa we cannot know whether the verb potentially switched class before the breakdown of the entire split-\gl{s} system.
-While there is no language-internal evidence, it is clear that \qu{to go down} does have an inclination to switch classes; in the case of \carijo, that could have already happened at the \PTar stage.
+For \PTir, one cannot establish a relative chronology of the verb class switch, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
+For \PWai, not only the first person form of the verb is unknown, but even its class membership.
+Similarly, for \carijo and \yukpa one cannot say whether the verb potentially switched class before the breakdown of the entire split-\gl{s} system.
+While there is no language-internal evidence for such a switch, it is clear that \qu{to go down} does have an inclination to switch classes; in the case of \carijo, that could have already happened at the \PTar stage.
 In all four cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, resisting it and preserving the old prefix, but it is also possible that it was not even a potential target due to its \gl{s_p_} status at the time.
+
 On the other hand, the class switch happened before the split-up of \trio and \akuriyo, and therefore this verb resisted the extension of \akuriyo \obj{k-} as an \gl{s_a_} verb.
-Likewise, it seems very likely that the class switch took place before the extension of \PPek \rc{k-}.
-Otherwise, the newly-turned-\gl{s_a_} verb would have taken on archaic and lexically heavily restricted \rc{w-}, either in \PTar, \PXin, or \arara.
+Likewise, it is very likely that the class switch took place before the extension of \PPek \rc{k-}.
+Otherwise, the newly-turned-\gl{s_a_} verb would have taken on conservative and lexically heavily restricted \rc{w-}, either in \PPek, \PXin, or \arara.
 
-\subsection{\rc{e-pɨ} \qu{to bathe}}
+\subsubsection{\rc{e-pɨ} \qu{to bathe}}
 \label{sec:bathe}
@@
 It took on new \gl{1}\gl{s_a_} prefixes in \PTir \parentext{\trio \obj{s-epɨ-}, \akuriyo \obj{t͡ʃe+pɨ-} \parencites[292]{triomeira1999}[87]{gildea1994akuriyo}} and \PWai \parentext{\hixka \obj{k-ewehɨ-}, \waiwai \obj{k-ejeɸu-} \parencites[195]{hixkaryanaderby1985}[166]{waiwaihawkins1998}}.
-The first person form of its \carijo reflex \obj{ehɨ} \parencites[72]{koch1908hiana} is unknown; we are not aware of a \yukpa cognate.
+The first person form of its \carijo reflex \obj{ehɨ} \parencites[72]{koch1908hiana} is unknown; an unattested \yukpa cognate may exist.
 
@@
 These roots are reflexes of \rc{pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
-As we have shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
+As was shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
 Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch, changing \rc{e-} to \rc{i}.
-This is an irregular development, since reflexes of \rc{i-} are not attested as a reflex of the detransitivizer in Pekodian \parencite[506]{meira2010origin}; its cause is unknown.
-However, it should be noted that other languages also show unexpected developments in this verb, considering the apparent glide insertion in Waiwaian or the distribution of \rc{pɨ} and \rc{kupi} in Venezuelan languages.
+This is an irregular development, since there are no attested reflexes of the detransitivizer in Pekodian that would go back to \PPek \rc{i-} \parencite[506]{meira2010origin}; its cause is unknown.
+However, it should be noted that other languages also show unexpected developments in this verb, considering the glide insertions in Waiwaian or the distribution of \rc{pɨ} and \rc{kupi} in Venezuelan languages.
 
-%\input{floats/bathe}
+\input{floats/bathe}

### 2021-12-05 workflow/helper_functions.py

@@ -56,7 +56,7 @@ def save_float(tabular, label, caption, filename=None, short=None):
         shorttext = ""
     else:
         shorttext = f"[{short}]"
-    output = f"""\\begin{{table}}[htbp]
+    output = f"""\\begin{{table}}[h]
 \centering
 \caption{shorttext}{{{caption}}}
 \label{{tab:{label}}}

### 2021-12-06 documents/sections/reasons.tex
rewriting section 4

\section{Explaining conservativeness: a network morphology approach}
\label{sec:motivations}
As stated in \cref{sec:verbsummary}, the fact that the same verbs retain conservative first person prefixes in six individual innovations is in need of explanation.
Perhaps the most well-known contribution regarding conservativeness, innovativeness, and (ir-){}re\-gu\-la\-ri\-ty in the lexicon is \posscite{bybee1985morphology} network model of morphology, which seems well-suited for the data at hand.
It aims \dbqu{to account for cross-linguistic, diachronic and acquisition patterns in complex morphological systems} \parencite[428]{bybee1995regular}.
It does so by modeling shared morphological properties such as inflectional patterns as emerging from connections of differing strength between lexemes.
A classic example is a network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
This network is attracting new verbs in certain dialects, like \obj{sniːk}--\obj{snʌk} or \obj{brɪŋ}--\obj{brʌŋ} \parencite[129--130]{bybee1985morphology}.
These verbs are recruited based on the lexical connection they form with prototypical members of the group, and accordingly develop irregular or \dbqu{strong} past tense forms.

As possible bases of these connections between lexemes, \textcite[118]{bybee1985morphology} suggests the criteria of semantic, phonological, and morphological similarity; the English strong verbs are an example for a phonologically motivated network.
Another important factor in the model is frequency, since more frequent words have a higher lexical strength \parencite[119]{bybee1985morphology}.
This higher lexical strength diminishes the influence from other lexemes, meaning that high-frequency items are more likely to resist innovations.
For the Cariban inflectional patterns under investigation, the model would predict that a) semantically\slash{}phonologically\slash\hspace{0pt}morphologically similar verbs will be affected by person marker extensions, and b) high-frequency verbs will tend to resist these extensions and thus remain conservative.

When one considers the groups of verbs with innovative first person markers (those not in \cref{tab:overview}), one can easily perceive several factors which could serve as the thread connecting a lexical network.
Perhaps the most obvious similarity is that they all have a reflex of the detransitivizer \detrz, a hallmark of \gl{s_a_} verbs \pcref{sec:split}.
%This was already noted by \textcite{meira1998proto} for the group of \PTar verbs taking irregular first person \rc{w-}:
%\begin{quotebox}{\parencite[112]{meira1998proto}}
%	This category includes a small number of stems, among which ‘to go’, ‘to come’, ‘to say’, [...]\footnote{The original list includes \qu{to go down} and \qu{to defecate}. While these verbs are indeed underived \gl{s_a_} verbs in \trio, no irregular first person \rc{w-} can be reconstructed. As their inclusion in the list was most likely an error, we omit them here.} and the copula. These are basically the verbs that are not synchronically or diachronically detransitivized, yet belong to the A conjugation.
%\end{quotebox} % ‘to go down’, ‘to defecate’,
This also has a phonological consequence: all derived \gl{s_a_} verbs begin with reflexes of \rc{ə} or \rc{e}, meaning that networks with a phonological basis are also plausible.
A more trivial connection between these other verbs is that they are all \gl{s_a_} verbs, and thus share inflectional (morphological) patterns.
In the following analysis, the \gl{1}\gl{s_a_} (pre-innovation) prefix will serve as a proxy for fhese inflectional patterns.
This is made possible by the fact that \gl{s_a_} verbs all had a reflex of \rc{w-} at the time of innovation, making the first person prefix representative for the inflectional class.
The exception to this is \akuriyo, where innovative \gl{1}\gl{s_a_} \obj{k-} and idiosyncratic \obj{p-} on \qu{to go down} created inflectional subclasses of \gl{s_a_} verbs.
There are no semantic patterns that would be obvious, which is not unsurprising given the lack of semantic patterns in the split-\gl{s} system overall \pcref{sec:split}.
Thus, for each extension, there are three hypotheses as to what connected the members of the responsible network: a reflex of \gl{detrz}, their stem-initial phoneme, or a specific \gl{1}\gl{s_a_} prefix.

It is intuitively obvious that many of the conservative verbs in \cref{tab:overview} are high-frequency verbs, which would mean high lexical strength and conservativeness according to the network model.
A major obstacle to confirming this intuition is the lack of frequency counts for individual lexemes for any Cariban language.
The only statement in the literature is \posscite[75]{courtz2008carib} claim of \kalina underived \gl{s_a_} verbs being the most frequent ones: \dbqu{It is difficult […] to imagine an intransitive or transitive origin for some of the most frequent middle verbs}.
This claim is supported neither by frequency counts nor accompanied by a list of verbs, but the unmentioned verbs likely correspond to the five first columns of \cref{tab:overview}, all underived \gl{s_a_} verbs in \kalina.
To ameliorate the dearth of data, a count of \gl{s_a_} verbs in three glossed texts from \textcite{koehns1994textos} will serve as a second source of frequency information, the results of which are shown in \cref{tab:apalaicounts}.
The \apalai data agree with the above interpretation of \posscite{courtz2008carib} claim; defining \dbqu{high frequency} as having an above average count produces the exact same five verbs.
While it is not at all certain that the interpretation of \posscite{courtz2008carib} claim and the small \apalai sample are representative of discourse patterns in the Cariban (proto-)languages under discussion, their use as a tool for identifying high-frequency verbs is necessitated by the absence of alternatives.

\input{floats/apalaicounts.tex}

%\textcite[153]{alves2017arara}: \dbqu{Observe que todos os 12 verbos indexados pelo morfema \obj{k-} [\gl{s_a_}] apresentam a raiz iniciada pela vogal /o/}.

Thus, in addition to the three hypotheses for possible network factors, each can be combined with frequency; high-frequency verbs are predicted to not undergo innovation, even though the factor in question would put them in the same lexical network as regular \gl{s_a_} verbs.
This leaves each extension with six possible explanations for which verbs are affected and which are not.
To test the hypotheses, for each explanation the expected behavior of each verb was predicted, illustrated in \cref{tab:ptir-predictions} for \PTir.
For example, \rc{eʔi} \qu{to be} is expected to participate in innovations spreading in a phonologically defined network, being \rc{e}-initial, as well as in an inflectionally defined one, since it shared \rc{w-} with other \gl{s_a_} verbs.
However, it would not have belonged to a network defined by the presence of a detransitivizer.
If frequency is taken into account, it is expected to remain conservative regardless of the nature of the network.
These predictions were then checked against the data in \cref{tab:overview}, to see how many potentially conservative verb each explanation predicted correctly.
This yielded a score of what proportion of potentially conservative verbs had their behavior accurately predicted, illustrated for \PTir in \cref{tab:ptir-evaluations} and summed up in \cref{tab:resultsoverview}.

\input{floats/ptir-predictions.tex}
\input{floats/ptir-evaluations.tex}

\input{floats/resultsoverview.tex}

It is important to understand that the scores in \cref{tab:resultsoverview} only refer to the group of seven verbs in \cref{tab:overview}, i.e. those that are attested as resisting at least one extension.
For each extension, there were also many run-of-the-mill \gl{s_a_} verbs, all taking on the new person marker, except for the \akuriyo \obj{e}-initial verbs.%
\footnote{While there are a few \gl{s_a_} verbs not transparently derived from transitive verbs \parencites[252]{triomeira1999}[222]{meira2000split}[30]{gildea2007greenberg}, which are not featured in \cref{tab:overview}, these are mostly \rc{ə}-initial and were likely productively derived at some point.
	The verbs to which this does not apply, like \trio \obj{wa} \qu{to dance} \parencites[252]{triomeira1999}, are all instances of \gl{s_p_} verbs switching classes.
	Since none of them is attested as being an \gl{s_a_} verb at the time of a person marker extension, they are not relevant for this study.}
To illustrate, if one adds 1'000 regular \gl{s_a_} verbs per language to the data -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six explanations consistently predict the behavior of 99.99+\% verbs correctly.
However, the available data simply does not allow such large-scale tests for Cariban languages, so the present investigation is restricted to the edge cases.

The extent of the extensions in both \PWai and \PPek is fully predicted by the presence or absence of a detransitivizer.
In both cases, only the underived%
\footnote{Note that for \PPek, the idiosyncratic evolution of \rc{e-pɨ} \qu{to bathe (\gl{intr})} to \rc{ipɨ}  made the verb morphologically opaque.}
\gl{s_a_} verbs were not affected, all other \gl{s_a_} verbs taking \rc{k-}.
Not shown in \cref{tab:resultsoverview} are subsequent evolutions in the Pekodian daughter languages, which can largely be argued to also be due to the detransitivizer:
First, both \ikpeng and \bakairi regularized the paradigm to use forms with detransitivizer for first person \pcref{sec:pekodian}, and both introduced \obj{k-}.%
\footnote{If one instead assumes that first person \rc{w-ebɨ-} and \rc{k-əd-ebɨ-} already co-existed in \PPek, the clear correlation between \rc{k-} and the detransitivizer remains.}
Second, the subsequent introduction of \obj{k-} to \ikpeng \obj{aran} \qu{to go} (< \rc{ɨtən}) suggests a potential reanalysis of \obj{ar} as a detransitivizer.

Three extensions are fully predicted by phonological criteria, those in \akuriyo, \carijo, and \yukpa.
As already discussed, \akuriyo \obj{k-} only appears on \obj{ə}-initial verbs \pcref{sec:akuriyo}.
In \carijo, the extension of \obj{j-} affected all \obj{e}- and \obj{ə}-initial verbs, including \obj{eh} \qu{to come} or \obj{et͡ʃi} \qu{to be}, which do not have a detransitivizing prefix.
Only \obj{ka} \qu{say}, \obj{təmə} \qu{go}, and \obj{a} \qu{be-1} did not take on \obj{j-}.
Similarly, the extension of \yukpa \obj{j-} can succinctly be characterized as affecting all vowel-initial verbs; the only attested conservative verb is C-initial \obj{to} \qu{to go}.

Inflectional morphology as a network basis only played a potential role in the case of \akuriyo.
However, this explanation only works if \obj{t-} and \obj{t͡ʃ-}, which were of course also phonologically conditioned, are analyzed as distinct morphemes.

When additionally considering putative conservatory frequency effects, prediction accuracy was improved in 8 cases, stagnated in 7 cases, and worsened in 3 cases.
The three cases where the tentative model of verb frequency arrives at incorrect predictions are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
Including frequency in the model led to overall improvements, up to 100\% prediction accuracy for all three potential factors in \PTir, as well as for the phonological criterion in \PPek and the inflection criterion in \PWai.

Overall, one must conclude that the patterns of most extensions are correctly predicted not by a single explanation, but rather by 3 to 4 different ones.
The exceptions are \carijo and \yukpa, where a lexical network with a phonolgical basis emerges as an unambiguous winner, while frequency-based explanations fare much worse.
%It may be worth noting that the three extensions which we do not reconstruct to a proto-language (ti.e., which are more recent) are best explained by phonological conditioning factors.
For the other extensions, the network model gives no unambiguous answer to the question of what combination of factors caused innovative first person markers to spread the way they did.
This in turn is due to the fact that three of the factors used in the model -- detransitivizer, phonology, frequency -- largely converge in their predictions:
The most frequent verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.

### 2021-12-06 documents/sections/verbs.tex
rewriting section 4

@@
 \input{floats/bathe}
+
+\subsection{Summary}
+\label{sec:verbsummary}
+In \cref{sec:verbs}, the verbs which were unaffected by the extensions in \cref{sec:extensions} were reconstructed, and affected reflexes in the languages under discussion were identified, where possible.
+\cref{tab:overview} gives an overview of what verbs were affected by which extensions, except for \obj{e}-initial \akuriyo verbs unaffected by the extension of \obj{k-}, as they are a large and predictable group.
+In some cases, the verb does not occur at all or just not in first person forms (–), in others the first person form is unknown (?), and in the case of \qu{to go down} the question of affectedness is not meaningfully answerable (\textsc{n/a}), for any of the reasons discussed in \cref{sec:godown}.
+Every \checkmark stands for a verb affected by an extension, while × represents conservatively inflected verbs.
+This overview makes clear just how pervasive the tendency for these verbs to resist innovative markers is, as they do so in different languages.
+
+\input{floats/overview.tex}
+
+It is astonishing that the same 1-7 verbs retain their old first person marker in 6 distinct developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs take on innovative markers.
+This suggest that there is some strong motivation for these verbs to not be affected by innovative markers.
+The question arises what properties unite these verbs and make them so conservative across different Cariban languages.
+Possible answers to this question will be discussed in \cref{sec:motivations}, using \posscite{bybee1985morphology} network model of morphology.

### 2021-12-06 workflow/process_data.py

@@ -315,6 +315,7 @@ temp_df1 = grouped.get_group(1.0)
 temp_df2 = grouped.get_group(0.5)
 temp_df1["Segments"] = temp_df1.apply(lambda x: segmentify(x["Form"]), axis=1)
 temp_df1["Cognateset_ID"] = temp_df1["Cognateset_ID"].map(str2numcog)
+print(temp_df1)
 temp_df1 = calculate_alignment(temp_df1, fuzzy=True)
 temp_df1["Cognateset_ID"] = temp_df1["Cognateset_ID"].map(num2strcog)
 temp_df1.rename(columns={"Cognateset_ID1": "Cognateset_ID"}, inplace=True)
@@ -476,7 +477,8 @@ for table in bathe_tables:
         table[2],
         tabular,
     )
-bathe_out += r"\end{subtable}\end{table}"
+bathe_out += r"""\end{subtable}
+\end{table}"""
 f = open(f"../documents/floats/bathe.tex", "w")
 f.write(bathe_out)
 f.close()
@@ -593,7 +595,7 @@ for v in overview_legend.values():
 legend += "\\end{legendlist}"
 
 save_float(
-    print_latex(result, keep_index=True) + legend,
+    print_latex(result, keep_index=True) + legend + "\n",
     label,
     "Overview of extensions and (un-)affected verbs",
 )

### 2021-12-06 documents/floats/pwaireg.tex
class membership of proto-taranoan 'to go down' is not reconstructible

--- 2021-12-02 documents/floats/pwaireg.tex
@@
-\begin{table}
+\begin{table}[h]
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-12-06 documents/sections/introduction.tex
improved map

@@
 The family is concentrated in Venezuela, the Guianas and Northern Brazil, with two Western and four Southern outliers.
-\cref{fig:phylomap} gives an overview of the geographical distribution and genealogical position of the extant Cariban languages.
-Cariban languages feature relatively rich verbal morphology, both pre- and suffixes, inflecting for person, number, tense, aspect, and evidentiality, combined with a range of valency-modifying affixes.
-Many also have a split-\gl{s} system, which can be reconstructed to \PC \pcref{sec:split}.
-For more detailed linguistic overviews of and comparative work on the family, we refer readers to \textcites{gildea1998}{derbyshire1999carib}{meira2002first}{meira2005southern}{meira2006cariban}{gildea2007greenberg}{meira2010origin}{gildea2010story}{gildea2012classification}{matter2021cariban}{gildea2019overview}.
+\cref{fig:phylomap} gives an overview of the geographical distribution and genealogical affiliation of the extant Cariban languages.
+%Cariban languages feature relatively rich verbal morphology, both pre- and suffixes, inflecting for person, number, tense, aspect, and evidentiality, combined with a range of valency-modifying affixes.
+%Many also have a split-\gl{s} system, which can be reconstructed to \PC \pcref{sec:split}.
+For linguistic overviews of and large-scale comparative work on the family, readers are referred to \textcites{gildea1998}{derbyshire1999carib}{meira2002first}{meira2005southern}{meira2006cariban}{gildea2007greenberg}{meira2010origin}{gildea2010story}{gildea2012classification}{matter2021cariban}{gildea2019overview}.
 
@@
 This is illustrated for \hixka in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
-This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms we show.
-In \cref{tab:hixintro} and other paradigm tables, we omit any TAM suffixes found in the original forms found in the literature, since a) our focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.
-Further, we use standard IPA symbols in our transcription of Cariban languages, with the exception of coronal rhotics, which we simply represent with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
-In languages with strong morphophonological processes and/or subphonemic orthography we show the original transcription in an additional surface line when presented in interlinearized examples.
-We follow \textcite{gildea2018reconstructing} in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.
+This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms shown here.
+In \cref{tab:hixintro} and other paradigm tables, any TAM suffixes found in the original forms found in the literature are omitted, since a) the focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.
+Further, standard IPA symbols are used in the transcription of Cariban languages, with the exception of coronal rhotics, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+In languages with strong morphophonological processes and/or subphonemic orthography the original transcription is shown in an additional surface line when presented in interlinearized examples.
+\textcite{gildea2018reconstructing} is followed in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.
 Glossing abbreviations: }%\glossingAbbrevsComma.

### 2021-12-06 documents/sections/proto_carib.tex
improved map

@@
 
-\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment; for our story of conservative verbs we will zoom in on a particular aspect of these extensions.
-To begin, we argue that they took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment, but a particular aspect of these extensions is also relevant for conservatively inflected intransitive verbs.
+To begin, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
 First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
@@
 
-Our third argument in favor of the lexical diffusion scenario brings us back to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
+The third argument in favor of the lexical diffusion scenario relates to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
 In both cases, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
-We interpret the continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
-In our family-wide investigation of person marker extensions, we identified 19 distinct extensions affecting intransitive verbs, and found 6 of them to be incomplete.
-These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages. %, marked with bold print in \cref{fig:caribantree}.
+The continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs is best interpreted as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
+In a family-wide investigation of person marker extensions, 19 distinct extensions affecting intransitive verbs were identified, 6 of them incomplete.
+These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages, and will be discussed in \cref{sec:data}.
 
@@
 %, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
-The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affecting them points to the split-\gl{s} system playing some role, so we will discuss its main properties in \cref{sec:split}.
+The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affecting them points to the split-\gl{s} system playing some role, so its main properties will be discussed in \cref{sec:split}.
 This will also give readers an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
@@
 \label{sec:split}
-As seen in \cref{sec:pc_person}, the \PC distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by two inflection classes within the \setone person paradigm, but this was not the only inflectional difference:
+As seen in \cref{sec:pc_person}, the distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by two inflection classes within the \PC \setone person paradigm, but this was not the only inflectional difference:
 %\begin{inlinelist}
@@
 However, both can occur with either meaning -- at least for these two verbs.
+In the next section, it will become evident that many of the verbs not affected by person marker extensions belong to the small group of \gl{s_a_} verbs without a reflex of \detrz.
 %As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.

### 2021-12-07 documents/sections/extensions.tex
rewriting section 4
improved map
first draft of conclusion, finetuning

@@
 \label{sec:data}
-As stated in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
-\cref{sec:extensions} introduces the six identified incomplete extensions, the prefixes they introduced and the verbs they spared.
-Since there is considerable etymological overlap in which verbs these are, the seven verbs showing a considerable degree of resistance are compared and reconstructed in \cref{sec:verbs}.
-The goal of this section is to identify distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
+As was shown in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
+\cref{sec:extensions} presents the six known incomplete extensions, the prefixes they introduced and the verbs they spared.
+Since the latter show considerable etymological overlap across languages, the resistant verbs are compared and reconstructed in \cref{sec:verbs}.
+In addition, their reflexes which did get affected by one of the extensions are identified, where existent.
+%The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
 
@@
 As stated in \cref{sec:extensions_intro}, the person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically, in the case of loss of split-\gl{s}) members of the \gl{s_a_} class.
-Of the six attested incomplete extensions, three can be reconstructed to intermediate proto-languages, while three others happened in pre-modern stages of single languages.
+Of the six attested incomplete extensions, half can be reconstructed to intermediate proto-languages, while the others happened in pre-modern stages of single languages.
 The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
-Each extension is discussed separately, contrasting regular (innovative) verbs with irregular (conservative) ones, and reconstructing forms where necessary:
-\cref{sec:pekodian} investigates the innovation of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
-\cref{sec:waiwaian} is about the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
-\cref{sec:taranoan} concerns the extension of \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
+For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and forms are reconstructed where necessary:
+\cref{sec:pekodian} investigates the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
+\cref{sec:waiwaian} concerns the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
+\cref{sec:taranoan} focuses on the extension of \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
 The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative first person markers only attested in single languages:
@@
 \begin{multicols}{2}
-\ex<arairr> \arara \parencite[153]{alves2017arara} \\
+\ex<arairr> \arara \parencite[153]{alves2017arara}\\
 \begin{tabular}[t]{@{}ll@{}}

### 2021-12-07 documents/sections/introduction.tex
first draft of conclusion, finetuning

@@
 %Many also have a split-\gl{s} system, which can be reconstructed to \PC \pcref{sec:split}.
-For linguistic overviews of and large-scale comparative work on the family, readers are referred to \textcites{gildea1998}{derbyshire1999carib}{meira2002first}{meira2005southern}{meira2006cariban}{gildea2007greenberg}{meira2010origin}{gildea2010story}{gildea2012classification}{matter2021cariban}{gildea2019overview}.
+For linguistic overviews of and  comparative work on the family, readers are referred to \textcites{gildea1998}{derbyshire1999carib}{meira2002first}{meira2005southern}{meira2006cariban}{gildea2007greenberg}{meira2010origin}{gildea2010story}{gildea2012classification}{matter2021cariban}{gildea2019overview}.
 
@@
 
-Some Cariban languages show a small group of verbs which diverge in their first person inflection pattern.
+Some Cariban languages show a small group of verbs which diverge in their first person inflection pattern, a topic which has not received much attention in the literature.
 This is illustrated for \hixka in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
@@
 which shows person paradigms for four verbs, all members of the \gl{s_a_} inflectional class.
-In this language, the verb \qu{to be} diverges in its first person marker (\obj{w-}), contrasting with other \gl{s_a_} verbs like \qu{to fall}, which have \obj{k(ɨ)-}.
+In this language, the verb \qu{to be} diverges in its first person marker (\obj{w-}), contrasting with other \gl{s_a_} verbs like \qu{to fall}, which have \obj{k-}.
 A similar pattern can be found in \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
@@
 
-In the literature, such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}.
-In a language-specific synchronic analysis, these verbs and their first person prefixes may be considered to be \textsc{irregular}, contrasting with regular prefixes, like \hixka \obj{k(ɨ)-} and \trio \obj{t-}/\obj{s-}, on regular verbs.
-However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[][e.g.]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
-For such approaches, these verbs simply belong to a small inflectional (sub-)class, an analysis found in the literature for the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
+Such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}, but have only been subject to comparative analysis for \posscite{meira1998proto} reconstruction of \PTar.
+In a synchronic analysis of a language, these verbs and their first person prefixes may be called \textsc{irregular}, contrasting with regular prefixes, like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}, on regular verbs.
+However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[e.g.,][]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
+For such approaches, these verbs simply belong to a small inflectional (sub-)class, an analysis applied to the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
 
 Regardless of the synchronic analysis, the reason for these inflectional patterns can be found in the diachrony of the languages in question.
-The purpose of this paper is to approach the patterns from a comparative perspective and to provide a unifying diachronic account, proceeding as follows.
-In \cref{sec:background}, readers are provided with necessary background knowledge about \PC verbs, and the mechanism of person marker extensions is introduced.
+The purpose of this paper is to approach the patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows.
+In \cref{sec:background}, readers are provided with necessary background knowledge about \PC verbs, and it is shown that the mechanism of person marker extension is responsible for the \hixka and \trio patterns.
 In \cref{sec:data}, six incomplete person marker extensions and their extent in the lexicon are described.
@@
 \cref{sec:motivations} uses \posscite{bybee1985morphology} network model of morphology to find explanations for the verbs (not) affected by each extension.
-\cref{sec:discussion} summarizes and discusses the results, and puts them in a general context of language change.
+\cref{sec:discussion} summarizes and discusses the results.
 

### 2021-12-07 documents/sections/proto_carib.tex
first draft of conclusion, finetuning

@@
 Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
-That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}), as well as the absence of \rc{i} after all \gl{s_a_} prefixes.
+That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
 On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones.
@@
 It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{ɔks-n} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
-Since the irregular \hixka and \trio prefixes are conservative and the regular prefixes are innovative, the next question to be addressed is where these new prefixes came from.
+Since the regular \hixka and \trio prefixes are innovative, the next question is where they came from.
 
@@
 
-\subsection{Person marker extensions in intransitive verbs}
+\subsection{Person marker extensions and lexical diffusion}
 \label{sec:extensions_intro}
@@
 All innovations are person marker extensions except 2b, which combined a pronoun with the linker \rc{j-}.
-They are printed in bold in \cref{fig:par_ext}, which reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
+\cref{fig:par_ext} shows them in bold and reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
 %
@@
 
-\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment, but a particular aspect of these extensions is also relevant for conservatively inflected intransitive verbs.
+\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment, but this study focuses on a particular aspect of theirs.
 To begin, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
@@
 In both cases, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
-The continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs is best interpreted as the extension stopping short of these verbs, rather than affecting all targets (\gl{s_a_} verbs).
+The continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs is best interpreted as the extension stopping short of these verbs, rather than spreading through the entire \gl{s_a_} lexicon.
 In a family-wide investigation of person marker extensions, 19 distinct extensions affecting intransitive verbs were identified, 6 of them incomplete.
-These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages, and will be discussed in \cref{sec:data}.
+These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages, and are treated individually in \cref{sec:data}.
 
 Interestingly, all six extensions featured innovative first person markers on \gl{s_a_} verbs.
-While extensions occurred with other person values as well, they never affected \gl{s_a_} verbs, only \gl{s_p_} ones, and they always affected all potential targets.
-Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare},\footnote{The presence of the third person marker \obj{n-} for \gl{1+2} is due to the wholesale loss of that inflectional value.} or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
+All other (complete) extensions%
+\footnote{As an honorable mention, when \ikpeng introduced a new mixed person paradigm to \setone constructions, \obj{a} \qu{to be} and \obj{ke} \qu{to say} retained third person \obj{n-} \parencite[12]{matter2019arara}.
+	However, the spread of the innovative third person markers had an entirely different pathway from the extensions discussed here, spreading from subordinate to main clauses, rather than from verbs to verbs.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
+Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
 %, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
-The markedly different behavior of \gl{s_a_} and \gl{s_p_} verbs with regards to the extent of extensions affecting them points to the split-\gl{s} system playing some role, so its main properties will be discussed in \cref{sec:split}.
-This will also give readers an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
+This starkly different behavior of \gl{s_a_} and \gl{s_p_} verbs when subject to extensions points to the split-\gl{s} system playing some role, so its main properties will be discussed in \cref{sec:split}.
+Readers will also get an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
 
@@
 Many languages show an \gl{s_a_} class marker in deverbalized forms, originating in \PC \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
-With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes.
-Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
+With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes \pcref{tab:participles,tab:nominalizations}.
+%Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
 %
@@
 %
-\input{floats/imperatives}
+\input{floats/imperatives}%
 %
-The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives, where the latter took the \gl{p}-oriented second person prefix \rc{ə(j)-}, while the former received no prefix (both suffixed with \rc{-kə}).
-This is illustrated with reflexes in various modern languages in \cref{tab:imperatives}.
+The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives, where the latter took the \gl{p}-oriented second person prefix \rc{ə(j)-} while the former received no prefix \pcref{tab:imperatives}.
+%This is illustrated with reflexes in various modern languages in \pcref{tab:imperatives}.
 Both the \gl{s_a_} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
@@
 The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
-The intransitive verbs in \exref{kar2} show the same person markers, but the choice of marker depends on the verb, rather than semantics.
+The intransitive verbs in \exref{kar2} show the same person markers, but the choice of marker depends on the verb rather than semantics.
 
@@
 %
-Thus, \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+\obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
 In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
-\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker must be considered volitional.
+\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker is clearly volitional.
 \textcite{meira2000split} takes a sizable corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
@@
 
-As for the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct  prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although on verbs their reflexes have been merged in a single morpheme in modern languages.
-Modern reflexes of \detrz show a range of meanings, which can all be characterized as \dbqu{detransitive}; this range is illustrated with \trio examples in \exref{triodetrz}.
+Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although on verbs their reflexes have been merged in a single morpheme in modern languages.
+Modern reflexes of \detrz show a range of meanings summarizable as \dbqu{detransitive}, illustrated with \trio \gl{s_a_} verbs in \exref{triodetrz}.
 

### 2021-12-07 documents/sections/reasons.tex
first draft of conclusion, finetuning

@@
 Overall, one must conclude that the patterns of most extensions are correctly predicted not by a single explanation, but rather by 3 to 4 different ones.
-The exceptions are \carijo and \yukpa, where a lexical network with a phonolgical basis emerges as an unambiguous winner, while frequency-based explanations fare much worse.
+The exceptions are \carijo and \yukpa, where a lexical network with a phonological basis emerges as the unambiguous winner, while frequency-based explanations fare much worse.
 %It may be worth noting that the three extensions which we do not reconstruct to a proto-language (ti.e., which are more recent) are best explained by phonological conditioning factors.
-For the other extensions, the network model gives no unambiguous answer to the question of what combination of factors caused innovative first person markers to spread the way they did.
+For the other extensions, the network model gives no unambiguous answer to the question of what (combination of) factors caused innovative first person markers to spread the way they did.
 This in turn is due to the fact that three of the factors used in the model -- detransitivizer, phonology, frequency -- largely converge in their predictions:
-The most frequent verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.
+The most frequent \gl{s_a_} verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.

### 2021-12-07 workflow/process_data.py

@@ -620,7 +620,7 @@ phon_preds = {
     "PPek": ["ə", "e"],
     "aku": ["ə"],
     "car": ["ə", "e"],
-    "yuk": ["o", "a", "e", "i", "u"],
+    "yuk": ["o", "a", "e", "i", "u", "ɨ"],
 }
 
 

### 2021-12-08 documents/sections/extensions.tex
trimming

@@
 \label{sec:extensions}
-As stated in \cref{sec:extensions_intro}, the person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically, in the case of loss of split-\gl{s}) members of the \gl{s_a_} class.
-Of the six attested incomplete extensions, half can be reconstructed to intermediate proto-languages, while the others happened in pre-modern stages of single languages.
+As stated in \cref{sec:extensions_intro}, the six person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically, in the case of loss of split-\gl{s}) members of the \gl{s_a_} class.
+Of these extensions, half can be reconstructed to intermediate proto-languages, while the others happened in pre-modern stages of single languages.
 The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
-For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and forms are reconstructed where necessary:
-\cref{sec:pekodian} investigates the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
-\cref{sec:waiwaian} concerns the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
-\cref{sec:taranoan} focuses on the extension of \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
-The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative first person markers only attested in single languages:
+For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and verb forms are reconstructed where necessary:
+\cref{sec:pekodian} details the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
+\cref{sec:waiwaian} treats the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
+\cref{sec:taranoan} focuses on innovative \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
+The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative \gl{1}\gl{s_a_} markers only found in single languages:
 \obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.
@@
 \label{sec:pekodian}
-The Pekodian branch was suggested by \textcite{meira2005southern}, as the result of fieldwork on \bakairi by Meira and the availability of more material on \ikpeng.
+%The Pekodian branch was suggested by \textcite{meira2005southern}, as the result of fieldwork on \bakairi by Meira and the availability of more material on \ikpeng.
 % \parencites{pacheco1997relativizacao}{ikpengpacheco1997}{campetela1997analise}{pacheco1998nominalizacao}{ikpengpacheco2001}{campetela2002aspectos}{pacheco2002aspectos}{pacheco2003intransitivos}.
-It consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
-\textcite{meira2005southern} focused on phonological and lexical properties, so no reconstructive work on \PPek morphosyntax can be found in the literature.
-However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}.
-Thus, it is possible to reconstruct a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
+The Pekodian branch consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
+In their comparative work, \textcite{meira2005southern} focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
+However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}, which allow the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
 
@@
 \arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
-From a comparative perspective, this list is not quite complete, as there is also a reflex of the copula \rc{a[p]}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
+There is also a reflex of the copula \rc{a[p]}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
 However, its inflectional morphology features verbal \setone prefixes, including first person \obj{w-} \exref{araap}.
@@
 
-In his brief discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.
+In his brief but exact discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.
 The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}, contrasting with regular \obj{əku} \qu{to go up} in \cref{tab:pekreg} above.
@@
 %
-Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
-While \textcite[4]{meira2003bakairi} does list \obj{ge} \qu{to say}, \obj{tə} \qu{to go}, and \obj{əe(wɨ)} \qu{to come} as examples of \gl{s_a_} verbs, he does not indicate whether they belong to the class of \gl{s_a_}-1 verbs, with first person \obj{k-}, or the \gl{s_a_}-2 verbs, with \obj{w-}.%
+%Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
+While \textcite[4]{meira2003bakairi} states that \bakairi cognates of \arara verbs in \exref{arairr} (\obj{ge} \qu{to say}, \obj{tə} \qu{to go}, and \obj{əe[wɨ]} \qu{to come}) are \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.%
 \footnote{
@@
 }
-Luckily, while \textcite{von1892bakairi} did not accurately record all phonemic distinctions in \bakairi \parencite{meira2005bakairi}, he does provide inflected forms of cognates to the \arara verbs in \exref{arairr}.
-They are presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
+Inflected forms can be found in \textcite{von1892bakairi}, presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
 
@@
 
-All available descriptions of the third Pekodian language, \ikpeng, list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
-However, most \ikpeng cognates of the \arara verbs with \gl{1}\gl{s_a_} \obj{w-} actually do not take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
+All descriptions of \ikpeng, the third Pekodian language, list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
+However, most \ikpeng cognates of the verbs in question do not actually take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
 The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-Further, while reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, it seems that only reflexes of \rc{eti} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
+Further, while reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, apparently only reflexes of \rc{eti} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
 
@@
 
-\Cref{tab:ppekverbs} gives an overview of the first person forms of the seven verbs under discussion, along with reconstructed \PPek forms.
-The presence and distribution of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
-Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in \cref{tab:pxinw}.
-While it is by no means a regular sound change, it permits identifying the two prefixes as cognate.
-Similarly, the suggested change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+Reconstructed \PPek forms of the verbs in question are given in \Cref{tab:ppekverbs}.
+Newly identified \ikpeng \obj{ɨ-}/Ø can be shown to be a reflex of \PXin \rc{w(ɨ)-} based on other (albeit irregular) cases of loss of \rc{w} \pcref{tab:pxinw}.
+%The presence and distribution of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
+%Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in .
+%While it is by no means a regular sound change, it permits identifying the two prefixes as cognate.
+Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
 Thus, a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to the \arara one both in form and distribution.
@@
 
-For the reconstructed forms of the \PPek verb stems readers are referred to \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
+Reconstructions of verb stems, \PPek and other, are discussed in \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
 The verb stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
@@
 %
-On the other hand, both \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following the line of reasoning used by \textcite[114]{meira1998proto} for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), it is likely that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently regularized the paradigm to only use \rc{əd-epɨ}; this verb shows similar patterns throughout the family \pcref{sec:come}.
+In contrast, \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
+Following the line of reasoning used by \textcite[114]{meira1998proto} for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), it is likely that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.
 
@@
 \label{sec:waiwaian}
-This person marker extension led to the \hixka pattern shown in \cref{sec:intro} and was one of the Parukotoan extensions discussed in \cref{sec:extensions_intro}.
-In \PWai, the new \gl{1}\gl{s_p_} prefix \rc{k-}, already innovated at the \PPar stage, was extended to \gl{1}\gl{s_a_}.
-For regularly inflected verbs, this created a unified \gl{1}\gl{s} category, reflected in both \hixka and \waiwai \pcref{tab:pwaireg}.
+This extension, one of the Parukotoan innovations discussed in \cref{sec:extensions_intro}, created the \hixka pattern from \cref{sec:intro}.
+In \PWai, the new \gl{1}\gl{s_p_} prefix \rc{k-}, innovated in \PPar, was further extended to \gl{1}\gl{s_a_}.
+For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
 \input{floats/pwaireg}
+
 Not all \gl{s_a_} verbs were affected: \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
-Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, a potential reflex of \gl{1}\gl{s_a_} \rc{w(ɨ)-}.
-A formally identical prefix occurs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, which regularly corresponds to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.
+Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, which also occurs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, corresponding to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.
 
@@
 %
-This regular correspondence in transitive verbs points to \hixka \obj{ɨ-} on intransitive verbs as another reflex of \rc{wɨ-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
-Notably, \textcite{hixkaryanaderby1985} does not analyze this \obj{ɨ-} as an irregular \gl{1}\gl{s_a_} prefix, but as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive \pcref{sec:say}.
+The regular correspondence in transitive verbs points to \hixka \obj{ɨ-} on intransitive verbs as another reflex of \rc{wɨ-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
+Notably, \textcite{hixkaryanaderby1985} analyzes this \obj{ɨ-} as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive \pcref{sec:say}.
 

### 2021-12-08 tests/test_infl_stems.py
testing consistency between stems and inflected forms

@@ -1,21 +1,61 @@
 import pandas as pd
+import itertools
+import re
+
 i_df = pd.read_csv("../data/inflection_data.csv")
 s_df = pd.read_csv("../data/verb_stem_data.csv")
 b_df = pd.read_csv("../data/bathe_data.csv")
 
+def extract_abbrevs(full_string):
+    def iterate(string):
+        for repl in [r"\1", ""]:
+            variant = re.sub(r"\[(.*?)\]", repl, string, 1)
+            if ("[" in variant):
+                for i in iterate(variant):
+                    yield i
+            else:
+                yield variant.replace("]", "")
+
+    
+    all_abbrevs = list(itertools.chain(*[iterate(s) for s in full_string.split(", ")]))
+    # Deduplicate while keeping the order:
+    seen = set()
+    return [x for x in all_abbrevs if not (x in seen or seen.add(x))]
+
 s_df = s_df.append(b_df)
-s_df["Form"] = s_df["Form"].replace({"\[": "", "\]": "", "\+": ""}, regex=True)
+# s_df["Short"] = s_df["Form"].replace({"\[.*?\]": ""}, regex=True)
+# s_df["Form"] = s_df["Form"].replace({"\[": "", "\]": ""}, regex=True)
+s_df["Forms"] = s_df["Form"].map(extract_abbrevs)
+
 
 stems = {}
 for lg, df in s_df.groupby("Language_ID"):
-    stems[lg] = dict(zip(df["Cognateset_ID"], df["Form"]))
+    stems[lg] = {}
+    for i, row in df.iterrows():
+        cogid = row["Cognateset_ID"]
+        if cogid not in stems[lg]:
+            stems[lg][cogid] = row["Forms"]
+        else:
+            stems[lg][cogid].extend(row["Forms"])
 
 for i, row in i_df.iterrows():
     verb = row["Verb_Cognateset_ID"]
-    lg = row["Language_ID"]
-    form = row["Form"]
     if pd.isnull(verb): continue
+    form = row["Form"]
+    if "-" not in form: continue
+    lg = row["Language_ID"]
     if lg not in stems:
         raise ValueError(f"No data for language {lg}!")
     if verb not in stems[lg]:
         print(f"No data for {lg} {form} '{verb}'!")
+        continue
+    if row["Inflection"] not in ["IMP"]:
+        i = 1
+    else:
+        i = 0
+    stem = form.split("-")[i]
[…]

### 2021-12-08 documents/floats/go.tex
added verb stems previously only in inflection data
new approach to testing predictions from bybee's model
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
added table of Werikyana progressive paradigms

@@
-\begin{table}
+\begin{table}[h]
 \centering
-\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6]{franchetto2002kuikuro}[99; p.c., Spike Gildea]{camargo2002lexico}}
+\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[80, 153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6]{franchetto2002kuikuro}[99; p.c., Spike Gildea]{camargo2002lexico}}
 \label{tab:go}
@@
 \mytoprule
-Language & \multicolumn{6}{l}{Form} \\
-\midrule
+Language &             Form &    &    &    &    &    \\
+\mymidrule
 \kaxui    &     \obj{to[mo]} &    &  t &  o &  m &  o \\
+\PWai     &    \rc{[ɨ]to[m]} &  ɨ &  t &  o &  m &    \\
 \hixka    &      \obj{[ɨ]to} &  ɨ &  t &  o &    &    \\
 \waiwai   &   \obj{[e]to[m]} &  e &  t &  o &  m &    \\
+\PPek     &        \rc{ɨtən} &  ɨ &  t &  ə &  n &    \\
 \arara    &        \obj{ɨdo} &  ɨ &  d &  o &    &    \\
+\arara    &         \obj{to} &    &  t &  o &    &    \\
 \ikpeng   &       \obj{aran} &  a &  r &  a &  n &    \\
@@
 \bakairi  &      \obj{[ɨ]tə} &  ɨ &  t &  ə &    &    \\
+\PTir     &        \rc{təmɨ} &    &  t &  ə &  m &  ɨ \\
 \trio     &      \obj{tə[n]} &    &  t &  ə &  n &    \\
@@
 \yukpa    &         \obj{to} &    &  t &  o &    &    \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2021-12-08 documents/floats/kaxprog.tex
added table of Werikyana progressive paradigms

\begin{table}[h]
\centering
\caption{\kaxui \obj{johɨ} \qu{to come} compared with other verbs (Spike Gildea, p.c.)}
\label{tab:kaxprog}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &                   \qu{to come} &                   \qu{to dream} &                    \qu{to go} \\
\mymidrule
\gl{1}   &  \obj{{\normalfont ∅}-w-oohɨ-} &  \obj{{\normalfont ∅}-w-osone-} &  \obj{{\normalfont ∅}-wɨ-to-} \\
\gl{2}   &                 \obj{o-w-ohɨ-} &                \obj{o-w-osone-} &                 \obj{o-w-to-} \\
\gl{1+2} &                \obj{ku-w-ohɨ-} &               \obj{ku-w-osone-} &                \obj{kɨ-w-to-} \\
\gl{3}   &    \obj{{\normalfont ∅}-johɨ-} &    \obj{{\normalfont ∅}-osone-} &                   \obj{i-to-} \\
\mybottomrule
\end{tabular}
\end{table}

### 2021-12-08 documents/floats/ppekverbs.tex
fixed pekodian 'go down' in comparative tables
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
added table of Werikyana progressive paradigms

--- 2021-11-23 documents/floats/ppekverbs.tex
@@
-\begin{table}
+\begin{table}[h]
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[76, 130, 131, 137, 374, 397]{von1892bakairi}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[4; p.c., Angela Fabíola Alves Chagas]{meira2003bakairi}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[76, 130, 131, 374, 397]{von1892bakairi}{aŋ-ka}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
 \label{tab:ppekverbs}
@@
 {} &          \PPek &          \arara &                       \ikpeng &        \bakairi \\
-\midrule
+\mymidrule
 \qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &      \obj{w-a-} \\
@@
 \qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &      \obj{w-i-} \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2021-12-08 documents/floats/pwaireg.tex
added table of Werikyana progressive paradigms

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-12-08 documents/floats/say.tex
added verb stems previously only in inflection data
new approach to testing predictions from bybee's model
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
added table of Werikyana progressive paradigms

@@
-\begin{table}
+\begin{table}[h]
 \centering
-\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63; p.c., Spike Gildea]{largo2011yukpa}}
+\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[267]{meira2005bakairi}[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209, 209, 279]{ikpengpacheco2001}[80, 153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63; p.c., Spike Gildea]{largo2011yukpa}}
 \label{tab:say}
@@
 \mytoprule
-Language & \multicolumn{5}{l}{Form} \\
-\midrule
+Language &          Form &    &    &    &    \\
+\mymidrule
 \kaxui   &   \obj{ka[s]} &  k &  a &  s &    \\
-\hixka   &   \obj{ka[h]} &  k &  a &  h &    \\
+\PWai    &    \rc{ka[s]} &  k &  a &  s &    \\
+\hixka   &   \obj{ka[s]} &  k &  a &  s &    \\
 \waiwai  &   \obj{ka[s]} &  k &  a &  s &    \\
+\PPek    &       \rc{ge} &  g &  e &    &    \\
+\PPek    &       \rc{ke} &  k &  e &    &    \\
+\arara   &      \obj{ge} &  g &  e &    &    \\
 \arara   &      \obj{ke} &  k &  e &    &    \\
-\ikpeng  &      \obj{ke} &  k &  e &    &    \\
+\ikpeng  &      \obj{ge} &  g &  e &    &    \\
+\ikpeng  &    \obj{k[e]} &  k &  e &    &    \\
 \bakairi &      \obj{ke} &  k &  e &    &    \\
+\bakairi &      \obj{ge} &  g &  e &    &    \\
+\PTir    &       \rc{ka} &  k &  a &    &    \\
 \trio    &      \obj{ka} &  k &  a &    &    \\
@@
 \yukpa   &      \obj{ka} &  k &  a &    &    \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2021-12-08 documents/sections/verbs.tex
added table of Werikyana progressive paradigms

@@
 
-Evidence for the originally \rc{j}-initial nature of the root is found only in the Pemongan languages and \kaxui, although the coexistence of the forms \obj{oohɨ} and \obj{johɨ} in \kaxui raises the question of whether the \obj{j} in the latter form is actually part of the root.
-\obj{johɨ} only occurs with third person in the Progressive \exref{kaxprog.kax-82}, and one might analyze \obj{j(-)} as a reflex of \rc{i-w-} \qu{\gl{3}-\gl{s_a_}-}, since the \gl{s_a_} class marker \rc{w-} is present with other person values \exref{kaxprog.kax-54}.
-However, while C-initial verbs do show a clear reflex of third person \rc{i-} \exref{kaxprog.kax-77}, regular V-initial \gl{s_a_} verbs do not show \obj{j-}, but Ø \exref{kaxprog.kax-83}.
-Thus, the best explanation for this \obj{j} is that is indeed part of the root, rather than an outcome of \rc{i-w-}, which in turn permits to confidently reconstruct \rc{j} back to \PC.
-
-\pex<kaxprog>\kaxui \perscomm{Spike Gildea}
-\a<kax-82>
-\begingl
-\gla johɨ-rɨ//
-\glb \gl{3}.come-\gl{prog}//
-\glft \qu{S/he is coming.}//
-\endgl
-\a<kax-54>
-\begingl
-\gla o-w-ohɨ-rɨ//
-\glb \gl{2}-\gl{s_a_}-come-\gl{prog}//
-\glft \qu{You are coming.}//
-\endgl
-\a<kax-77>
-\begingl
-\gla i-nkɨ-rɨ//
-\glb \gl{3}-sleep-\gl{prog}//
-\glft \qu{S/he is sleeping.}//
-\endgl
-\a<kax-83>
-\begingl
-\gla Ø-osone-rɨ//
-\glb \gl{3}-dream-\gl{prog}//
-\glft \qu{S/he is dreaming.}//
-\endgl
-\xe
+Only the Pemongan languages and \kaxui point to an \rc{j}-initial root, and \kaxui \obj{johɨ} is very rare in contrast to more frequent \obj{o(o)hɨ}.
+It only occurs in the third person form of the Progressive, meaning that the \obj{j} may be a reflex of third person \rc{i-}.
+As \cref{tab:kaxprog} shows, regular \obj{o}-initial verbs have no third person prefix, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}.
+Even if \qu{to come} once shared this \obj{i-}, the circumstances are strongly in favor of a \obj{j}-initial \kaxui root, allowing the reconstruction of \PC \rc{jəpɨ}.
+%
+\input{floats/kaxprog}%
+%\pex<kaxprog>\kaxui \perscomm{Spike Gildea}
+%\a<kax-82>
+%\begingl
+%\gla johɨ-rɨ//
+%\glb \gl{3}.come-\gl{prog}//
+%\glft \qu{S/he is coming.}//
+%\endgl
+%\a<kax-54>
+%\begingl
+%\gla o-w-ohɨ-rɨ//
+%\glb \gl{2}-\gl{s_a_}-come-\gl{prog}//
+%\glft \qu{You are coming.}//
+%\endgl
+%\a<kax-77>
+%\begingl
+%\gla i-nkɨ-rɨ//
+%\glb \gl{3}-sleep-\gl{prog}//
+%\glft \qu{S/he is sleeping.}//
+%\endgl
+%\a<kax-83>
+%\begingl
+%\gla Ø-osone-rɨ//
+%\glb \gl{3}-dream-\gl{prog}//
+%\glft \qu{S/he is dreaming.}//
+%\endgl
+%\xe
 %

### 2021-12-08 workflow/process_data.py

@@ -260,6 +260,28 @@ for lg, meanings in {
         short=f"Regular {print_shorthand(lg)} verbs",
     )
 
+# comparison of werikyana set 2 paradigms of Sa verbs
+prog_df = i_df[i_df["Inflection"].str.contains("PROG")]
+# remove prog suffix
+prog_df["Inflection"] = prog_df["Inflection"].str.replace(".PROG", "")
+pyd.x = ["Meaning_ID"]
+pyd.y = ["Inflection"]
+pyd.z = []
+pyd.filters = {}
+pyd.y_sort = person
+table = pyd.compose_paradigm(prog_df)
+table.index.name = None
+table.index = table.index.map(pynt.get_expex_code)
+table = table[["come", "dream", "go"]]
+table.columns = table.columns.map(lambda x: f"\\qu{{to {x}}}")
+table = table.applymap(objectify)
+save_float(
+    print_latex(table, keep_index=True),
+    "kaxprog",
+    "\\kaxui \\obj{johɨ} \\qu{to come} compared with other verbs (Spike Gildea, p.c.)",
+)
+
+
 i_df = i_df[~(pd.isnull(i_df["Verb_Cognateset_ID"]))]
 
 # pekodian lexical comparison
@@ -315,7 +337,6 @@ temp_df1 = grouped.get_group(1.0)
 temp_df2 = grouped.get_group(0.5)
 temp_df1["Segments"] = temp_df1.apply(lambda x: segmentify(x["Form"]), axis=1)
 temp_df1["Cognateset_ID"] = temp_df1["Cognateset_ID"].map(str2numcog)
-print(temp_df1)
 temp_df1 = calculate_alignment(temp_df1, fuzzy=True)
 temp_df1["Cognateset_ID"] = temp_df1["Cognateset_ID"].map(num2strcog)
 temp_df1.rename(columns={"Cognateset_ID1": "Cognateset_ID"}, inplace=True)
@@ -799,7 +820,6 @@ label = "resultsoverview"
 
 res_df = pd.DataFrame(all_explanations)
 res_df *= 100
-print(res_df)
 
 export_csv(
     res_df.rename(columns=lambda x: extension_string(x, latex=False), index=repl_latex),

### 2021-12-08 documents/sections/extensions.tex
trimming pekodian

@@
 The Pekodian branch consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
-In their comparative work, \textcite{meira2005southern} focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
-However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-}, as evidenced by the paradigms in \cref{tab:pekreg}, which allow the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
+The contribution establishing the branch \parencite{meira2005southern} focuses on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
+However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
 
@@
 
-In the most detailed description of a Pekodian language, \textcite{alves2017arara} names six%
+The most detailed description of a Pekodian language \parencite{alves2017arara} names six%
 \footnote{Seven under her analysis, which sees the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.}
@@
 
-In his brief but exact discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.
-The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}, contrasting with regular \obj{əku} \qu{to go up} in \cref{tab:pekreg} above.
+In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.\footnote{
+	It should be noted at this point that \textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
+	However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
+	Given the lack of data for other verbs, this potential additional pattern will not be further discussed.
+	If the characterization by \citeauthor{meira2003bakairi} is accurate, then verbs with innovative first person prefixes have conservative \gl{1+2} prefixes, and vice versa.
+}
+The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}.
 
@@
 %Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
-While \textcite[4]{meira2003bakairi} states that \bakairi cognates of \arara verbs in \exref{arairr} (\obj{ge} \qu{to say}, \obj{tə} \qu{to go}, and \obj{əe[wɨ]} \qu{to come}) are \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.%
-\footnote{
-It should be noted at this point that \textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
-However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
-Given the lack of data for other verbs, this potential additional pattern will not be further discussed.
-If the characterization by \citeauthor{meira2003bakairi} is accurate, then verbs with innovative first person prefixes have conservative \gl{1+2} prefixes, and vice versa.
-}
+While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of \arara verbs in \exref{arairr} as \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.
 Inflected forms can be found in \textcite{von1892bakairi}, presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
@@
 
-All descriptions of \ikpeng, the third Pekodian language, list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
-However, most \ikpeng cognates of the verbs in question do not actually take \obj{k-}, but rather \obj{ɨ-} or Ø, as shown in \exref{ikpw}.
+All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
+However, most \ikpeng cognates of the verbs in question do not take\obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}.
 The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-Further, while reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, apparently only reflexes of \rc{eti} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
+While reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, apparently only reflexes of \rc{eti} \qu{to be} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
 
@@
 
-Reconstructed \PPek forms of the verbs in question are given in \Cref{tab:ppekverbs}.
-Newly identified \ikpeng \obj{ɨ-}/Ø can be shown to be a reflex of \PXin \rc{w(ɨ)-} based on other (albeit irregular) cases of loss of \rc{w} \pcref{tab:pxinw}.
+Reconstructed \PPek forms of conservatively inflected verbs are given in \Cref{tab:ppekverbs}.
+Newly identified \ikpeng \obj{ɨ-}/Ø is demonstrably a reflex of \PXin \rc{w(ɨ)-}, based on other (albeit irregular) cases of loss of \rc{w} \pcref{tab:pxinw}.
 %The presence and distribution of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
@@
 %While it is by no means a regular sound change, it permits identifying the two prefixes as cognate.
-Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences, like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
-Thus, a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to the \arara one both in form and distribution.
+Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences: \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+Thus, a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to its \arara reflex in form and distribution.
 In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.

### 2021-12-08 documents/sections/extensions.tex
trimming waiwaian

@@
 \label{sec:waiwaian}
-This extension, one of the Parukotoan innovations discussed in \cref{sec:extensions_intro}, created the \hixka pattern from \cref{sec:intro}.
-In \PWai, the new \gl{1}\gl{s_p_} prefix \rc{k-}, innovated in \PPar, was further extended to \gl{1}\gl{s_a_}.
+This extension, one of the Parukotoan innovations discussed in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
+In \PWai, the \gl{1}\gl{s_p_} prefix \rc{k-} innovated in \PPar was further extended to \gl{1}\gl{s_a_}.
 For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
@@
 There are three more verbs which did not take innovative \rc{k-} in \PWai, shown alongside \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
-The two forms for \qu{to be} are straightforwardly reconstructible, whereas \qu{to go} is somewhat of a special case.
+The two forms for \qu{to be} are straightforwardly reconstructible, whereas \qu{to go} is a special case.
 While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
-Alternatively, this form might have been influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \exref{waiwto}.
-
-\pex<waiwto> \waiwai reflexes of the \gl{s_a_} class marker \rc{w-}
-\a \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}
-\a \obj{o-\emp{w}to-t͡ʃhe} \qu{after I went} \parencite[165]{waiwaihawkins1998}
-\a \obj{kɨ-\emp{w}to-me} \qu{for us to go} \parencite[204]{waiwaihawkins1998}
-\xe
+Alternatively, this form may have been influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}
 %
-In any case, this was a subsequent development, as the prefix pattern in \hixka \qu{to go} clearly shows that \PWai first person form \rc{wɨ-tom-} can be reconstructed.
+%\pex<waiwto> \waiwai reflexes of the \gl{s_a_} class marker \rc{w-}
+%\a \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}
+%\a \obj{o-\emp{w}to-t͡ʃhe} \qu{after I went} \parencite[165]{waiwaihawkins1998}
+%\a \obj{kɨ-\emp{w}to-me} \qu{for us to go} \parencite[204]{waiwaihawkins1998}
+%\xe
+%
+In any case, the prefix pattern in \hixka \qu{to go} clearly shows that a \PWai first person form \rc{wɨ-tom-} can be reconstructed.
 

### 2021-12-08 documents/sections/extensions.tex
trimming tiriyoan

@@
 The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages already identified by \textcite{girard1971proto}, with \carijo as a more distant member.
-\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar.
-He faces an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
+\textcite{meira1998proto} contributes an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
 This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
@@
 Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
-Here, a conservative stance is taken and the innovation is assumed to be \PTir, but it could have well already taken place in \PTar.
-This decision has no effect on the results of this study.}
+Here, a conservative stance is taken and the innovation is arbitrarily assumed to be \PTir.
+This decision does not affect the results of this study.}
 
@@
 Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.
-To this list, the second copular root \rc{eʔi} can be added, whose \trio reflex retains first person \obj{w-}.
-The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is identified as a reflex of \rc{wɨ-} by 
-\textcite[113]{meira1998proto}, which is plausible and supported by additional evidence:
-Both components of the irregular change \rc{wɨ-} > \obj{ə-} -- loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə} -- are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
+To this list the second copular root \rc{eʔi} can be added, whose \trio reflex retains first person \obj{w-}.
+The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is plausibly identified as a reflex of \rc{wɨ-} by 
+\textcite[113]{meira1998proto}, which is supported by additional evidence:
+Both components of the irregular change \rc{wɨ-} > \obj{ə-} (loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
 
-\pex<exp>
-\a<aku-164> \akuriyo \parencite[][86]{gildea1994akuriyo}\\
+\pex<exp>\akuriyo
+\a<aku-164>
 \begingl
-\gla (w)i-toka//
+\gla wi-toka//
 \glb \gl{1}>\gl{3}-hit//
-\glft \qu{I hit him/her.}//
+\glft \qu{I hit him/her.} \parencite[][86]{gildea1994akuriyo}//
 \endgl
-\a<aku-159> \akuriyo \parencite[][114]{meira1998proto}\\
+\a<aku-159>
 \begingl
@@
 \glb \gl{1+2}-come//
-\glft \qu{We came.}//
+\glft \qu{We came.} \parencite[][114]{meira1998proto}//
 \endgl
@@
 
-For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for the other person values, based on an idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
+For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for other persons, based on an idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
 Both \akuriyo and \carijo then levelled this original distribution, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
@@
 
-In addition, \textcite{gildea1994akuriyo} recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-} \exref{akumov.aku}, all \obj{e}-initial movement verbs.
-Only \obj{erama} \qu{to return} has an attested \trio cognate, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
-Further, these verbs are not mentioned by \textcite{meira1998proto}, who was also working with \posscite{gildea1994akuriyo} data.
-Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by the available part of \posscite{meira1998proto} data, these verbs cannot be confidently reconstructed as resisting the extension of \PTir \rc{t-}.
-
-\pex<akumov>
-\a<aku> \akuriyo \gl{1}\gl{s_a_} \rc{w-} \parencite[84--86]{gildea1994akuriyo}\\
-\begin{tabular}[t]{@{}ll@{}}
-\qu{return} & Ø\obj{-erama-}\\
-\qu{get up} & Ø\obj{-eokahtə-}\\
-\qu{jump} & \obj{w-ejahka-}\\
-\qu{go out} & \obj{w-ekɨrɨka-}\\
-\end{tabular}
-\a<tri> \trio \obj{s-erama-} \parencite[301]{triomeira1999}
-\xe
+%In addition, \textcite{gildea1994akuriyo} recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-} \exref{akumov.aku}, all \obj{e}-initial movement verbs.
+%Only \obj{erama} \qu{to return} has an attested \trio cognate, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
+%Further, these verbs are not mentioned by \textcite{meira1998proto}, who was also working with \posscite{gildea1994akuriyo} data.
+%Given that this data potentially has strong \trio and/or \wayana influence \parencite[253]{gildea1998} and the lack of support by the available part of \posscite{meira1998proto} data, these verbs cannot be confidently reconstructed as resisting the extension of \PTir \rc{t-}.
+%
+%\pex<akumov>
+%\a<aku> \akuriyo \gl{1}\gl{s_a_} \rc{w-} \parencite[84--86]{gildea1994akuriyo}\\
+%\begin{tabular}[t]{@{}ll@{}}
+%\qu{return} & Ø\obj{-erama-}\\
+%\qu{get up} & Ø\obj{-eokahtə-}\\
+%\qu{jump} & \obj{w-ejahka-}\\
+%\qu{go out} & \obj{w-ekɨrɨka-}\\
+%\end{tabular}
+%\a<tri> \trio \obj{s-erama-} \parencite[301]{triomeira1999}
+%\xe
 
@@
 \label{sec:akuriyo}
-After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated yet another \gl{1}\gl{s_a_} marker: \obj{k-}.
-It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data, with all relevant verbs shown in \Cref{tab:aku1sa}.
+After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{s_a_} marker: \obj{k-}.
+It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
 \textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
 He also suggests that \obj{k-} might be more recent, which is very plausible: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
-The few \obj{t-} mentioned by \textcite{meira1998proto} were then potentially reintroduced under \trio influence.
-However, since there are no examples of, or further information about, \obj{ə}-initial verbs with \obj{t-}, these cases cannot be discussed further.
+The few \obj{t-} mentioned by \textcite{meira1998proto} were then perhaps reintroduced under \trio influence.
+However, since there are no examples of \obj{ə}-initial verbs with \obj{t-}, or further information about them, these cases cannot be discussed further.
 
@@
 
-The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} did of course also not introduce \obj{k-} in \akuriyo, although the first-person form of the copular verb \obj{eʔi} is presently unknown.
+The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} did of course also not introduce \obj{k-} in \akuriyo, although the first-person form of the copular verb \obj{eʔi} is unknown.
 In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in, which is apparently reconstructible to \PTir \exref{tridown}.
@@
 \carijo, the cousin of the Tiriyoan languages, has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
-Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs, reflected in the paradigms in \cref{tab:carreg}.
+Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs \pcref{tab:carreg}.
 %
@@
 It is analysed here as a reflex of \rc{wɨ-}, primarily due to the loss of \rc{w} in \ikpeng and \hixka.
-While that loss was clearly more regular, an already irregular marker undergoing idiosyncratic phonological erosion is not surprising, see \akuriyo \rc{wɨ-} > \obj{ə-} in the preceding section.
+While those cases of loss were more regular, an already irregular marker undergoing idiosyncratic phonological erosion is not that surprising, see \akuriyo \rc{wɨ-} > \obj{ə-} in the preceding section.
 Alternatively, it is possible that the divergent development of \rc{w-} on C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} is a result of the latter's originally V-initial nature \pcref{sec:go}.

### 2021-12-08 documents/sections/extensions.tex
yukpa trimming

@@
 These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their semantics and the reflex of \detrz.
-The same prefix occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, meaning that \obj{j(ɨ)-} is the reflex of the \PC \gl{1}\gl{s_p_} marker \rc{u(j)-}.
-On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
+The same prefix occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, so \obj{j(ɨ)-} is the reflex of the \PC \gl{1}S\textsubscript{(P)} marker \rc{u(j)-}.
+In contrast, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
 
@@
 %
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identifiedn  as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
 
 In intransitive verbs, this first-person zero marking is attested in a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
-It diverges from regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
-It is thus the only verb attested as having resisted the extension of \obj{j-} in \yukpa.
+It diverges from regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.
+It can thus be identified as having resisted the extension of \obj{j-} in \yukpa.
 

### 2021-12-08 documents/sections/verbs.tex
trimming verbs

@@
 \label{sec:be}
-For a comprehensive comparative overview of these two roots, readers are refferd to \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
+For a comprehensive comparative overview of these two roots, readers are referred to \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
 \rc{a[p]} is the original copula and can be reconstructed as already having various irregularities in \PC.
-\rc{eti}, the other root, is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
-Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam} value.
+\rc{eti} is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
+Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam}.
 Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
@@
 \label{sec:say}
-Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, best visible in the imperative forms of some languages \exref{kati}.
-
-\ex<kati> \apalai\\
-\obj{kaʃi-ko} \qu{say!}\\
-\parencite[35]{koehn1986apalai}
-\xe
+Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, best visible in the imperative forms of some languages.% \exref{kati}.
+%
+%\ex<kati> \apalai\\
+%\obj{kaʃi-ko} \qu{say!}\\
+%\parencite[35]{koehn1986apalai}
+%\xe
 %
@@
 In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}).
-In \exref{hixsay.ordpl-35}, it does not occur, because \qu{they said} is preceded by direct speech.
+It does not occur in \exref{hixsay.ordpl-35}, because \qu{they said} is preceded by direct speech.
 Looking beyond \hixka, at least the \trio cognate shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
@@
 Similarly, \qu{to say} has a reflex of \gl{1}\gl{s_a_} \rc{w-} in Pekodian languages \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
-Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show \obj{ɨ} instead of \obj{i} for this verb, see \exref{sayintr.aku-154} as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for the verb's inflection in such languages.
+Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show \obj{ɨ} instead of \obj{i} for this verb, see \akuriyo in \exref{sayintr.aku-154}, as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for cognate forms in other such languages.
 Finally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and it is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
@@
 
-In summary, this verb can be reconstructed as being intransitive based on its prefixes, but transitive based on some suffixes.
+In summary, this verb can be reconstructed as being intransitive based on its (inflectional) prefixes, but transitive based on some (derivational) suffixes.
 \hixka has lost the main intransitive criteria, making its reflex look more like a transitive verb.
-It is not attested as being affected by any of the person marker extensions  in \cref{sec:extensions}.
 
@@
 Still, when one considers unambiguous forms, the contrast with \rc{ka[ti]} becomes very clear.}
-This verb was not affected by the any of the extensions discussed in \cref{sec:extensions}.
+This verb was not affected by the any of the extensions in \cref{sec:extensions}.
 
@@
 
-This verb is reconstructed as \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all attested reflexes \pcref{tab:come} suggests a somewhat more complex story.
+This verb is reconstructed as \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all attested reflexes \pcref{tab:come} points to a more complex story.
 Crucially, the majority do not reflect the \rc{ət} part of their reconstruction, and many forms are ostensibly reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ}, rather than \rc{epɨ}.
-A unifying account of all these forms is made possible by reconstructing a \PC verb of the form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
+A unifying account of all these forms is achieved by reconstructing a \PC form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
 
@@
 %
-Most of the morphologically complex forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} do not show evidence for the segment \rc{j}, but the \obj{i} in the \akawaio form \obj{əsipɨ} is very likely a reflex of the sequence \rc{jə}.
+Most of the (non-Pemongan) morphologically complex forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} show no segmental trace of \rc{j}, but the \obj{i} in \akawaio \obj{əsipɨ} is likely a reflex of the sequence \rc{jə}.
 This analysis is supported by the reflex \obj{ipɨ} from bare \rc{jəpɨ} in very closely related \macushi.
 
-Turning to the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ}, it becomes evident are distributed widely in the family, sometimes even co-occurring in the same language.
-A unifying account of these forms requires the root \rc{jəpɨ} to undergo two major sound changes: \begin{inlinelist}
+As for the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ}, they are distributed widely in the family, sometimes even co-occurring in the same language.
+A unifying account requires the root \rc{jəpɨ} to undergo two major sound changes: \begin{inlinelist}
  \item \rc{j}-loss
@@
 
-The interpretation of the \rc{ət} part as a detransitivizer is justifiable due to its form and its paradigmatically conditioned occurrence in some languages.
+The interpretation of the \rc{ət} part as a detransitivizer is based on its form and its paradigmatically conditioned occurrence in some languages.
 Although the combination of a detransitivizer and an intransitive verb seems semantically illogical, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
@@
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
-It is unknown whether there is a \yukpa reflex of this verb, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come} and was thus not a potential target of innovative \rc{k-}.
+It is unknown whether there is a \yukpa reflex of this verb, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come}.
 

### 2021-12-09 documents/sections/verbs.tex
go down trimming

@@
 The resistance against the former extension was subsequently overcome in \bakairi; its fate in \ikpeng is unknown.
-When \akuriyo extended \obj{k-}, the verb already had an irregularly inflected first person form with \obj{p-}, inherited from \PTir.
-At first sight, it may seem that it was also affected by the two independent extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
+When \akuriyo extended \obj{k-}, the verb already had a first person form irregularly inflected with \obj{p-}, inherited from \PTir.
+One might think that it was also affected by the extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
 
@@
 %
-However, a broader comparative perspective reveals a much more complicated story.
-\cref{tab:godown} shows all attested reflexes of \rc{ɨpɨtə}, including verb class membership where applicable; parenthesized forms indicate uncertainty about cognacy status.%
-\footnote{The reconstruction of \PPek \rc{ɨptə} in \cref{tab:godown} treats the suffixed elements found in the daughter languages as non-cognate.
-\textcite{meira2005southern} make no mention of a regular correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}, and at least the addition of a final \obj{ŋ} in \PXin is attested elsewhere:
+However, a broader comparative perspective reveals a much more complicated story \pcref{tab:godown}.%
+\footnote{In \cref{tab:godown}, parenthesized forms indicate uncertainty about cognacy status.
+The reconstruction of \PPek \rc{ɨptə} treats the suffixed elements found in the daughter languages as non-cognate.
+\textcite{meira2005southern} identify no correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}, and at least the addition of a final \obj{ŋ} in \PXin is attested elsewhere:
 		\begin{inlinelist}
@@
 		\end{inlinelist} \parencites[8]{gildea2007greenberg}[56, 144, 57]{alves2017arara}[25, 270]{ikpengpacheco2001}.}
-It turns out that while the form \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about the class of this verb.
+It turns out that while a verb\rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about its class.
 Its reflexes in languages that preserve the split-\gl{s} system are distributed fairly evenly between \gl{s_a_} and \gl{s_p_}.
@@
 
-In one language, \wayana, the verb shows traits of both classes, necessitating an analysis as a \dbqu{mixed} verb in a synchronic description of \wayana.
+The verb shows traits of both classes in \wayana, necessitating an analysis as a \dbqu{mixed} verb in a synchronic description of that language.
 It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
@@
 %
-Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the restriction of \rc{-ka} to \gl{s_p_} verbs in \PC \parencite{gildea2019overview} points to \gl{s_p_} membership.
-These patterns are suggestive of a scenario in which the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class, reflected only in its \gl{1+2} prefix and the \gl{s_a_} class marker.
-This in turn implies that reflexes of this verb with \gl{s_a_} membership in other languages fully switched from \gl{s_p_} in their inflectional patterns.
+Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the \PC  causativizer \rc{-ka} was restricted to \gl{s_p_} verbs \parencite{gildea2019overview}.
+These patterns point to a scenario where the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class, taking a \gl{1+2}\gl{s_a_} prefix and the \gl{s_a_} class marker.
+This in turn implies that \gl{s_a_} reflexes of this verb in other languages fully switched from \gl{s_p_} in their inflectional patterns.
 
@@
 Both causativized forms contain a reflex of the transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
-Besides the irregular first person \obj{p-}, \trio \obj{ɨhtə} shows other irregularities, in particular in its causativized forms that also feature a reflex of \rc{en-} \parencite[263]{triomeira1999}.
-Thus, it appears that this verb was originally \gl{s_p_}, but then switched its class in four and a half languages of the family, for so far unknown reasons.
+\trio \obj{ɨhtə} has irregular causativized forms that also feature a reflex of \rc{en-} \parencite[263]{triomeira1999}.
+Thus, it appears that this verb was originally \gl{s_p_}, but then switched its class in four \dbqu{and a half} languages of the family, for so far unknown reasons.
 
-These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by the extensions in \PWai, \PTir, \carijo, and \yukpa.
+These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by some extensions.
 For \PTir, one cannot establish a relative chronology of the verb class switch, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
-For \PWai, not only the first person form of the verb is unknown, but even its class membership.
-Similarly, for \carijo and \yukpa one cannot say whether the verb potentially switched class before the breakdown of the entire split-\gl{s} system.
-While there is no language-internal evidence for such a switch, it is clear that \qu{to go down} does have an inclination to switch classes; in the case of \carijo, that could have already happened at the \PTar stage.
-In all four cases, it is possible that the verb had \gl{s_a_} status at the time of the extension, resisting it and preserving the old prefix, but it is also possible that it was not even a potential target due to its \gl{s_p_} status at the time.
+For \PWai, both its first person form and its inflectional class are unknown.
+For \carijo and \yukpa, one cannot rule out the possibility that the verb switched classes before the breakdown of the split-\gl{s} system.
+While no language-internal evidence supports that scenario, \qu{to go down} clearly has an inclination to switch classes; in the case of \carijo, that could have already happened at the \PTar stage.
+In all four cases, the verb could have had \gl{s_a_} status at the time of the extension, resisting it and preserving the old prefix, but it could also have had \gl{s_p_} status and thus not even have been a potential target.
 
-On the other hand, the class switch happened before the split-up of \trio and \akuriyo, and therefore this verb resisted the extension of \akuriyo \obj{k-} as an \gl{s_a_} verb.
+On the other hand, the class switch happened before the split of \PTir, and therefore this verb resisted the extension of \akuriyo \obj{k-} as an \gl{s_a_} verb.
 Likewise, it is very likely that the class switch took place before the extension of \PPek \rc{k-}.

### 2021-12-09 documents/floats/ppekverbs.tex
more trimming, fixed verb citations

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[76, 130, 131, 374, 397]{von1892bakairi}{aŋ-ka}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[76, 130, 131, 374, 397]{von1892bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
 \label{tab:ppekverbs}
@@
 \mymidrule
+\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &     \obj{u-ge-} \\
+\qu{go}      &   \rc{w-ɨtən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &     \obj{u-tə-} \\
 \qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &      \obj{w-a-} \\
 \qu{be-2}    &  \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &      \obj{w-i-} \\
-\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &     \obj{u-ge-} \\
-\qu{go}      &   \rc{w-ɨtən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &     \obj{u-tə-} \\
 \qu{come}    &    \rc{w-epɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &   \obj{k-əewɨ-} \\

### 2021-12-09 documents/floats/pwaireg.tex
more trimming, fixed verb citations

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-12-09 documents/sections/extensions.tex
more trimming, fixed verb citations

@@
 All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
-However, most \ikpeng cognates of the verbs in question do not take\obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}.
+However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}.
 The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
@@
 %While it is by no means a regular sound change, it permits identifying the two prefixes as cognate.
-Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in other correspondences: \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in correspondences like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
 Thus, a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to its \arara reflex in form and distribution.
@@
 
-Reconstructions of verb stems, \PPek and other, are discussed in \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
-The verb stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
+Reconstructions of verb stems are deferred to \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
+The stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
 In contrast, the \arara first person form is directly based on this root \rc{epɨ}.
@@
 In contrast, \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following the line of reasoning used by \textcite[114]{meira1998proto} for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), it is likely that the idiosyncratic pattern in \arara is reconstructible to \PPek, and that \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.
+Following \posscite[114]{meira1998proto} line of reasoning for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), the idiosyncratic pattern in \arara should be reconstructed to \PPek, while \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.
 
@@
 There are three more verbs which did not take innovative \rc{k-} in \PWai, shown alongside \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
-The two forms for \qu{to be} are straightforwardly reconstructible, whereas \qu{to go} is a special case.
+The two forms of \qu{to be} are unproblematic, whereas \qu{to go} is a special case.
 While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
@@
 %
-In any case, the prefix pattern in \hixka \qu{to go} clearly shows that a \PWai first person form \rc{wɨ-tom-} can be reconstructed.
+In any case, the first person form \hixka \qu{to go} clearly points to \PWai \rc{wɨ-tom-}.
 

### 2021-12-09 documents/sections/proto_carib.tex
more trimming, fixed verb citations

@@
 That system was subject to many different kinds of innovations; responsible for the patterns discussed are person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
-A particular component of the system, the distinction between \gl{s_a_} and \gl{s_p_} verbs, plays a role in what extensions are incomplete and is the topic of \cref{sec:split}.
+A particular component of the system, the \gl{s_a_} vs \gl{s_p_} distinction, plays a role in what extensions are incomplete and is discussed in \cref{sec:split}.
 %\cref{sec:outlook} summarizes the main points of this section and provides an outlook for the rest of the paper.
@@
 That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
-On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones.
-The third person marker in \gl{s_p_} verbs was identical to the one in \gl{3}>\gl{3} scenarios (\rc{n(i)-)}.
+On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones, with \gl{3}\gl{s_p_} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
 
@@
 These verbs and their prefixes are comparable with the few English nouns like \obj{ɒks}, which preserve the old plural suffix \obj{-ən}.
-It, too, was once more widespread, being the normal plural suffix of the weak inflection, compare German \obj{ɔks-n} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
+It was once more widespread as the normal plural suffix of the weak inflection, compare German \obj{ɔks-ən} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
 Since the regular \hixka and \trio prefixes are innovative, the next question is where they came from.
@@
 In his discussion of the \PC split-\gl{s} system and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different modifications in various languages.
-The main mechanism of change leading to these modifications are \textbf{person marker extensions}, i.e.
-the use of verbal person prefixes being extended to contexts previously occupied by other prefixes.
+The main mechanism of change leading to these modifications are \textsc{person marker extensions}, i.e. verbal person prefixes being extended to paradigm cells previously occupied by other prefixes.
 %\footnote{
@@
 There have been many different person marker extensions in Cariban languages, and some are still ongoing.
-This is illustrated by \textcite{gildea1998}, using the three Parukotoan languages as an example.
+\textcite{gildea1998} illustrates this using the three Parukotoan languages as an example.
 Apart from segmental changes to individual morphemes, the following innovations happened in the \setone paradigm in Parukotoan:
@@
 \hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows the variation in the first person and \gl{1+2} prefixes that is reconstructible to \PPar.
-\waiwai, on the other hand, has lost the system entirely, which notably happened via distinct innovations at three different diachronic stages.
+\waiwai has lost the system entirely, which notably happened via three diachronically distinct innovations.
 %In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
@@
 
-\textcite{gildea1998} discusses person marker extensions in the context of the loss of the split-\gl{s} system and the accompanying changes to indexing alignment, but this study focuses on a particular aspect of theirs.
-To begin, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+\textcite{gildea1998} discusses person marker extensions relating to the loss of split-\gl{s} and the accompanying changes to indexing alignment, but a different aspect of these extensions is relevant for this study.
+Namely, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
 First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
@@
 %todo maybe put in GOOD kaxui data
-Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} convincingly argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
+Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} convincingly argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
 Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
@@
 	However, the spread of the innovative third person markers had an entirely different pathway from the extensions discussed here, spreading from subordinate to main clauses, rather than from verbs to verbs.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
-Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, or the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
+Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, and the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
 %, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
 This starkly different behavior of \gl{s_a_} and \gl{s_p_} verbs when subject to extensions points to the split-\gl{s} system playing some role, so its main properties will be discussed in \cref{sec:split}.
-Readers will also get an idea how it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
+This will also explain why it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
 
@@
 \label{sec:split}
-As seen in \cref{sec:pc_person}, the distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by two inflection classes within the \PC \setone person paradigm, but this was not the only inflectional difference:
+As seen in \cref{sec:pc_person}, the distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only inflectional difference:
 %\begin{inlinelist}
@@
 %This is illustrated with reflexes in various modern languages in \pcref{tab:imperatives}.
-Both the \gl{s_a_} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
+%Both the \gl{s_a_} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
 %However, there was one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
 
-In the modern languages maintaining the split-\gl{s} system, mismatches between the semantics of verbs and their \gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with \kalina data in \exref[kar1]{kar2}.
-
-\pex<kar1> \kalina
-\a<kar-60>
-\begingl
-\gla mi-kupi-ja//
-\glb \gl{2}>\gl{3}-bathe-\gl{prs}//
-\glft \qu{You bathe him/her.} \parencite[][160]{hoff1968carib}//
-\endgl
-\a<kar-62>
-\begingl
-\gla a-kupi-ja//
-\glb \gl{3}>\gl{2}-bathe-\gl{prs}//
-\glft \qu{S/he bathes you.} \parencite[][63]{yamada2011evidentiality}//
-\endgl
-\xe
+In the modern languages maintaining the split-\gl{s} system, mismatches between the semantics of verbs and their \gl{s_a_} or \gl{s_p_} membership are common, exemplified with \kalina data in \exref{kar2}.
 %
-In \exref{kar1}, the choice between the second person \gl{a}- and \gl{p}-oriented markers \obj{mi-} and \obj{a-} depends on the scenario:
-The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
-The intransitive verbs in \exref{kar2} show the same person markers, but the choice of marker depends on the verb rather than semantics.
-
+%\gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with \kalina data in \exref[kar1]{kar2}.
+%
+%\pex<kar1> \kalina
+%\a<kar-60>
+%\begingl
+%\gla mi-kupi-ja//
+%\glb \gl{2}>\gl{3}-bathe-\gl{prs}//
+%\glft \qu{You bathe him/her.} \parencite[][160]{hoff1968carib}//
+%\endgl
+%\a<kar-62>
+%\begingl
+%\gla a-kupi-ja//
+%\glb \gl{3}>\gl{2}-bathe-\gl{prs}//
+%\glft \qu{S/he bathes you.} \parencite[][63]{yamada2011evidentiality}//
+%\endgl
+%\xe
+%%
+%In \exref{kar1}, the choice between the second person \gl{a}- and \gl{p}-oriented markers \obj{mi-} and \obj{a-} depends on the scenario:
+%The transitive verb \obj{kupi} \qu{to bathe} takes \obj{mi-} in \gl{2}>\gl{3} scenarios \exref{kar1.kar-60}, but \obj{a-} in \gl{3}>\gl{2} scenarios \exref{kar1.kar-62}.
+%The intransitive verbs in \exref{kar2} show the same person markers, but the choice of marker depends on the verb rather than semantics.
+%
 \pex<kar2> \kalina
@@
 %
-\obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, since it is an \gl{s_a_} verb \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+In \exref{kar2.kar-59}, the \gl{s_a_} verb \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
 In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
 \qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker is clearly volitional.
-\textcite{meira2000split} takes a sizable corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
+\textcite{meira2000split} investigates a corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
 He shows that neither (non\-)activities, % \parencite[210]{meira2000split}, 
@@
 \end{quotebox}
-\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
-\begin{inlinelist}
-	\item there are a few underived \gl{s_a_} verbs, with no detransitivizing prefix
-	\item \gl{s_a_} verbs can develop irregular semantics compared to their transitive counterparts
-	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{s_a_} verb
-	\item some originally derived \gl{s_a_} verbs have lost their transitive counterparts
-	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{s_a_})} from \qu{to dance with (\gl{tr})}
-\end{inlinelist}.
-He also notes that this leads to an inflectional split not based in meaning, but rather morphology:
+%\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
+%\begin{inlinelist}
+%	\item there are a few underived \gl{s_a_} verbs, with no detransitivizing prefix
+%	\item \gl{s_a_} verbs can develop irregular semantics compared to their transitive counterparts
+%	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{s_a_} verb
+%	\item some originally derived \gl{s_a_} verbs have lost their transitive counterparts
+%	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{s_a_})} from \qu{to dance with (\gl{tr})}
+%\end{inlinelist}.
+He notes that this leads to an inflectional split not based in meaning, but rather morphology:
 
@@
 
-Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although on verbs their reflexes have been merged in a single morpheme in modern languages.
+Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although on verbs their reflexes have been merged into a single morpheme in modern languages.
 Modern reflexes of \detrz show a range of meanings summarizable as \dbqu{detransitive}, illustrated with \trio \gl{s_a_} verbs in \exref{triodetrz}.
@@
 %
-The morphological variation featured in \qu{to abandon each other} and \qu{to wash self} is due to the mentioned collapse between the two \PC prefixes:
+The morphological variation featured in \qu{to abandon each other} and \qu{to wash self} is due to the collapse between the two \PC prefixes:
 \obj{e-} is a reflex of the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.

### 2021-12-09 documents/sections/reasons.tex
more trimming, fixed verb citations

@@
 \label{sec:motivations}
-As stated in \cref{sec:verbsummary}, the fact that the same verbs retain conservative first person prefixes in six individual innovations is in need of explanation.
 Perhaps the most well-known contribution regarding conservativeness, innovativeness, and (ir-){}re\-gu\-la\-ri\-ty in the lexicon is \posscite{bybee1985morphology} network model of morphology, which seems well-suited for the data at hand.
@@
 Another important factor in the model is frequency, since more frequent words have a higher lexical strength \parencite[119]{bybee1985morphology}.
-This higher lexical strength diminishes the influence from other lexemes, meaning that high-frequency items are more likely to resist innovations.
-For the Cariban inflectional patterns under investigation, the model would predict that a) semantically\slash{}phonologically\slash\hspace{0pt}morphologically similar verbs will be affected by person marker extensions, and b) high-frequency verbs will tend to resist these extensions and thus remain conservative.
+This diminishes the influence from other lexemes, meaning that high-frequency items are more likely to resist innovations.
+For the Cariban first person patterns, the model would predict that a) semantically\slash{}phonologically\slash\hspace{0pt}morphologically similar verbs will be affected by person marker extensions, and b) high-frequency verbs will tend to resist these extensions and thus remain conservative.
 
-When one considers the groups of verbs with innovative first person markers (those not in \cref{tab:overview}), one can easily perceive several factors which could serve as the thread connecting a lexical network.
-Perhaps the most obvious similarity is that they all have a reflex of the detransitivizer \detrz, a hallmark of \gl{s_a_} verbs \pcref{sec:split}.
+When considering the groups of verbs with innovative first person markers (those not in \cref{tab:overview}), there are several factors which could serve as the thread connecting a lexical network.
+The most obvious similarity is that they all have a reflex of the detransitivizer \detrz \parentext{see e.g.
+\textcite[112]{meira1998proto} for Taranoan}, a hallmark of \gl{s_a_} verbs \pcref{sec:split}.
 %This was already noted by \textcite{meira1998proto} for the group of \PTar verbs taking irregular first person \rc{w-}:
@@
 %\end{quotebox} % ‘to go down’, ‘to defecate’,
-This also has a phonological consequence: all derived \gl{s_a_} verbs begin with reflexes of \rc{ə} or \rc{e}, meaning that networks with a phonological basis are also plausible.
-A more trivial connection between these other verbs is that they are all \gl{s_a_} verbs, and thus share inflectional (morphological) patterns.
-In the following analysis, the \gl{1}\gl{s_a_} (pre-innovation) prefix will serve as a proxy for fhese inflectional patterns.
-This is made possible by the fact that \gl{s_a_} verbs all had a reflex of \rc{w-} at the time of innovation, making the first person prefix representative for the inflectional class.
-The exception to this is \akuriyo, where innovative \gl{1}\gl{s_a_} \obj{k-} and idiosyncratic \obj{p-} on \qu{to go down} created inflectional subclasses of \gl{s_a_} verbs.
-There are no semantic patterns that would be obvious, which is not unsurprising given the lack of semantic patterns in the split-\gl{s} system overall \pcref{sec:split}.
+Since all derived \gl{s_a_} verbs begin with reflexes of \rc{ə} or \rc{e} \parentext{see e.g. \textcite[153]{alves2017arara} for \arara}, phonologically based networks are a second possibility.
+%A more trivial connection between these other verbs is that they are all \gl{s_a_} verbs, and thus share inflectional (morphological) patterns.
+%In the following analysis, the \gl{1}\gl{s_a_} (pre-innovation) prefix will serve as a proxy for fhese inflectional patterns.
+%This is made possible by the fact that \gl{s_a_} verbs all had a reflex of \rc{w-} at the time of innovation, making the first person prefix representative for the inflectional class.
+%The exception to this is \akuriyo, where innovative \gl{1}\gl{s_a_} \obj{k-} and idiosyncratic \obj{p-} on \qu{to go down} created inflectional subclasses of \gl{s_a_} verbs.
+A third common trait is the \gl{s_a_} (sub-)class membership, represented by pre-innovation first person \rc{w-} (\obj{t-} in \akuriyo).
+There are no obvious semantic patterns, which is not surprising given the absence of a semantic basis in the split-\gl{s} system \pcref{sec:split}.
 Thus, for each extension, there are three hypotheses as to what connected the members of the responsible network: a reflex of \gl{detrz}, their stem-initial phoneme, or a specific \gl{1}\gl{s_a_} prefix.
 
-It is intuitively obvious that many of the conservative verbs in \cref{tab:overview} are high-frequency verbs, which would mean high lexical strength and conservativeness according to the network model.
-A major obstacle to confirming this intuition is the lack of frequency counts for individual lexemes for any Cariban language.
+It is intuitively obvious that many of the conservative verbs in \cref{tab:overview} are high-frequency verbs, which would cause conservativeness according to the network model.
+Going beyond intuition is difficult due to the lack of frequency counts for individual lexemes for any Cariban language.
 The only statement in the literature is \posscite[75]{courtz2008carib} claim of \kalina underived \gl{s_a_} verbs being the most frequent ones: \dbqu{It is difficult […] to imagine an intransitive or transitive origin for some of the most frequent middle verbs}.
-This claim is supported neither by frequency counts nor accompanied by a list of verbs, but the unmentioned verbs likely correspond to the five first columns of \cref{tab:overview}, all underived \gl{s_a_} verbs in \kalina.
-To ameliorate the dearth of data, a count of \gl{s_a_} verbs in three glossed texts from \textcite{koehns1994textos} will serve as a second source of frequency information, the results of which are shown in \cref{tab:apalaicounts}.
-The \apalai data agree with the above interpretation of \posscite{courtz2008carib} claim; defining \dbqu{high frequency} as having an above average count produces the exact same five verbs.
-While it is not at all certain that the interpretation of \posscite{courtz2008carib} claim and the small \apalai sample are representative of discourse patterns in the Cariban (proto-)languages under discussion, their use as a tool for identifying high-frequency verbs is necessitated by the absence of alternatives.
+This claim is supported neither by frequency counts nor accompanied by a list of verbs, but that list is likely identical with the five first columns of \cref{tab:overview}, all underived \gl{s_a_} verbs in \kalina.
+To improve the situation, a count of \gl{s_a_} verbs in three glossed \apalai texts from \textcite{koehns1994textos} will serve as a second source of frequency information, the results of which are shown in \cref{tab:apalaicounts}.
+The count data agree with the above interpretation of \posscite{courtz2008carib} claim: defining \dbqu{high frequency} as having an above average count yields the exact same five verbs.
+While it is uncertain that the interpretation of \posscite{courtz2008carib} claim and the small \apalai sample are truly representative of discourse patterns in the Cariban (proto-)languages under discussion, the absence of alternatives necessitates their use as a tool for modeling frequency.
 
@@
 
-Thus, in addition to the three hypotheses for possible network factors, each can be combined with frequency; high-frequency verbs are predicted to not undergo innovation, even though the factor in question would put them in the same lexical network as regular \gl{s_a_} verbs.
-This leaves each extension with six possible explanations for which verbs are affected and which are not.
-To test the hypotheses, for each explanation the expected behavior of each verb was predicted, illustrated in \cref{tab:ptir-predictions} for \PTir.
-For example, \rc{eʔi} \qu{to be} is expected to participate in innovations spreading in a phonologically defined network, being \rc{e}-initial, as well as in an inflectionally defined one, since it shared \rc{w-} with other \gl{s_a_} verbs.
-However, it would not have belonged to a network defined by the presence of a detransitivizer.
-If frequency is taken into account, it is expected to remain conservative regardless of the nature of the network.
-These predictions were then checked against the data in \cref{tab:overview}, to see how many potentially conservative verb each explanation predicted correctly.
-This yielded a score of what proportion of potentially conservative verbs had their behavior accurately predicted, illustrated for \PTir in \cref{tab:ptir-evaluations} and summed up in \cref{tab:resultsoverview}.
+Thus, each of the three hypotheses for possible network factors can be combined with frequency: the members of the lexical network formed by the factor are predicted to undergo innovation, but high-frequency verbs are exempt.
+The resulting six hypotheses for possible explanations were tested by predicting the expected behavior of each verb in each extension, illustrated in \cref{tab:ptir-predictions} for \PTir.
+For example, \rc{eʔi} \qu{to be} is expected to participate (\checkmark) in innovations spreading in a phonologically defined network (being \rc{e}-initial), as well as in an inflectionally defined one (sharing \rc{w-} with other \gl{s_a_} verbs).
+On the other hand, a network based on the detransitivizer would predict it to not take on new suffixes (×).
+If frequency is taken into account, it is expected to remain conservative regardless of the basis of the network.
 
 \input{floats/ptir-predictions.tex}
-\input{floats/ptir-evaluations.tex}
 
-\input{floats/resultsoverview.tex}
-
-It is important to understand that the scores in \cref{tab:resultsoverview} only refer to the group of seven verbs in \cref{tab:overview}, i.e.
-those that are attested as resisting at least one extension.
-For each extension, there were also many run-of-the-mill \gl{s_a_} verbs, all taking on the new person marker, except for the \akuriyo \obj{e}-initial verbs.%
+These predictions were checked against the data in \cref{tab:overview}, counting verbs which had their behavior (in-)correctly predicted.
+The resulting scores are illustrated for the extension of \PTir \rc{t-} in \cref{tab:ptir-evaluations} and summed up for all extensions in \cref{tab:resultsoverview}.
+Notably, the scores in \cref{tab:resultsoverview} only refer to the seven verbs which are attested as resisting at least one extension.
+For each extension, there were also many run-of-the-mill \gl{s_a_} verbs which were all affected, except for the \akuriyo \obj{e}-initial verbs.%
 \footnote{While there are a few \gl{s_a_} verbs not transparently derived from transitive verbs \parencites[252]{triomeira1999}[222]{meira2000split}[30]{gildea2007greenberg}, which are not featured in \cref{tab:overview}, these are mostly \rc{ə}-initial and were likely productively derived at some point.
@@
 	Since none of them is attested as being an \gl{s_a_} verb at the time of a person marker extension, they are not relevant for this study.}
-To illustrate, if one adds 1'000 regular \gl{s_a_} verbs per language to the data -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six explanations consistently predict the behavior of 99.99+\% verbs correctly.
-However, the available data simply does not allow such large-scale tests for Cariban languages, so the present investigation is restricted to the edge cases.
+To illustrate, if one adds 1'000 simulated derived \gl{s_a_} verbs per language to the data -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six explanations consistently predict the behavior of 99.99+\% verbs correctly.
+However, the present investigation is restricted to the edge cases, since the available data simply does not allow such large-scale tests for Cariban languages.
+
+\input{floats/ptir-evaluations.tex}
+\input{floats/resultsoverview.tex}
 
@@
 \gl{s_a_} verbs were not affected, all other \gl{s_a_} verbs taking \rc{k-}.
-Not shown in \cref{tab:resultsoverview} are subsequent evolutions in the Pekodian daughter languages, which can largely be argued to also be due to the detransitivizer:
+Not shown in \cref{tab:resultsoverview} are subsequent evolutions in the Pekodian daughter languages, which largely support a detransitivizer-based explanation:
 First, both \ikpeng and \bakairi regularized the paradigm to use forms with detransitivizer for first person \pcref{sec:pekodian}, and both introduced \obj{k-}.%
 \footnote{If one instead assumes that first person \rc{w-ebɨ-} and \rc{k-əd-ebɨ-} already co-existed in \PPek, the clear correlation between \rc{k-} and the detransitivizer remains.}
-Second, the subsequent introduction of \obj{k-} to \ikpeng \obj{aran} \qu{to go} (< \rc{ɨtən}) suggests a potential reanalysis of \obj{ar} as a detransitivizer.
+Second, the development of \PPek \rc{ɨtən} \qu{to go} to \ikpeng \obj{aran} may have led to reanalysis of \obj{ar} as a detransitivizer, accompanied by the introduction of \obj{k-}.
 
-Three extensions are fully predicted by phonological criteria, those in \akuriyo, \carijo, and \yukpa.
-As already discussed, \akuriyo \obj{k-} only appears on \obj{ə}-initial verbs \pcref{sec:akuriyo}.
+The extent of three extensions (in \akuriyo, \carijo, and \yukpa) is correctly predicted by phonological criteria.
+As discussed in \cref{sec:akuriyo}, \akuriyo \obj{k-} only appears on \obj{ə}-initial verbs.
 In \carijo, the extension of \obj{j-} affected all \obj{e}- and \obj{ə}-initial verbs, including \obj{eh} \qu{to come} or \obj{et͡ʃi} \qu{to be}, which do not have a detransitivizing prefix.
@@
 Inflectional morphology as a network basis only played a potential role in the case of \akuriyo.
-However, this explanation only works if \obj{t-} and \obj{t͡ʃ-}, which were of course also phonologically conditioned, are analyzed as distinct morphemes.
+However, this explanation only works if \obj{t-} and \obj{t͡ʃ-} are analyzed as distinct morphemes.
+Since they can also be seen phonologically conditioned allomorphs, the prediction is identical to the phonological one.
 
-When additionally considering putative conservatory frequency effects, prediction accuracy was improved in 8 cases, stagnated in 7 cases, and worsened in 3 cases.
-The three cases where the tentative model of verb frequency arrives at incorrect predictions are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
-Including frequency in the model led to overall improvements, up to 100\% prediction accuracy for all three potential factors in \PTir, as well as for the phonological criterion in \PPek and the inflection criterion in \PWai.
+When additionally considering the assumed conservative effects of frequency, prediction accuracy was improved in 8 cases, stagnated in 7 cases, and worsened in three cases.
+These three cases where the tentative model of verb frequency arrives at incorrect predictions are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
+Overall, including frequency in the model led to improvements, up to 100\% prediction accuracy for all three potential factors in \PTir, as well as for the phonological criterion in \PPek and the inflection criterion in \PWai.
 
-Overall, one must conclude that the patterns of most extensions are correctly predicted not by a single explanation, but rather by 3 to 4 different ones.
-The exceptions are \carijo and \yukpa, where a lexical network with a phonological basis emerges as the unambiguous winner, while frequency-based explanations fare much worse.
+In summary, one can conclude that the patterns of most extensions are correctly predicted not by a single explanation, but rather by 3 to 4 different ones.
+The exceptions are \carijo and \yukpa, where a phonologically defined lexical network emerges as the unambiguous winner, while frequency-based explanations fare much worse.
 %It may be worth noting that the three extensions which we do not reconstruct to a proto-language (ti.e., which are more recent) are best explained by phonological conditioning factors.
 For the other extensions, the network model gives no unambiguous answer to the question of what (combination of) factors caused innovative first person markers to spread the way they did.
-This in turn is due to the fact that three of the factors used in the model -- detransitivizer, phonology, frequency -- largely converge in their predictions:
+This in turn is due to the fact that three of the factors in the model -- detransitivizer, phonology, frequency -- strongly converge in their predictions:
 The most frequent \gl{s_a_} verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.

### 2021-12-09 documents/sections/verbs.tex
more trimming, fixed verb citations

@@
 The set of unaffected verbs is rather small in most cases, and (proto-)languages show a considerable degree of overlap in what verbs are unaffected.
-Therefore, this section investigates these verbs from a comparative perspective.
-\cref{sec:be} treats both roots of the copula \rc{eti}/\obj{a[p]} \qu{to be}, \cref{sec:say} \rc{ka[ti]} \qu{to say}, \cref{sec:go} \rc{ɨtə[mə]} \qu{to go}, and \cref{sec:come} \rc{(ət)jəpɨ} \qu{to come}.
-\cref{sec:godown} takes a look at \rc{ɨpɨtə} \qu{to go down}, which is resistant in \PTir and \PPek, and  \cref{sec:bathe} investigates \rc{e-pɨ} \qu{to bathe}, of which the \PPek reflex \rc{ipɨ} resisted the extension of \rc{k-}.
+Here,  resistant verbs are investigated from a comparative perspective:
+\rc{ka[ti]} \qu{to say} \pcref{sec:say}, \rc{ɨtə[mə]} \qu{to go} \pcref{sec:go}, both roots of the copula \rc{eti}/\obj{a[p]} \pcref{sec:be}, \rc{(ət)jəpɨ} \qu{to come} \pcref{sec:come},  \rc{ɨpɨtə} \qu{to go down} \pcref{sec:godown}, and  \cref{sec:bathe} investigates \rc{e-pɨ} \qu{to bathe}, of which the \PPek reflex \rc{ipɨ} resisted the extension of \rc{k-}.
 The \obj{e}-initial verbs not affected by the extension of \obj{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
-
-\subsubsection{\rc{eti} and \rc{a[p]} \qu{to be}}
-\label{sec:be}
-For a comprehensive comparative overview of these two roots, readers are referred to \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
-\rc{a[p]} is the original copula and can be reconstructed as already having various irregularities in \PC.
-\rc{eti} is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
-Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam}.
-Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
-\akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while \obj{eʔi} is not attested in a first-person form. 
-\carijo innovated \obj{j-}, but only in the reflex of \rc{eti} \exref{mivida-12}; the \obj{a} root preserves \obj{w-} \pcref{sec:carijo}.
-\yukpa introduced \obj{j-} to the reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
-
-\ex<mivida-12>\carijo \parencite[][177]{robayo1989rame}\\
-\begingl
-\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
-\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
-\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
-\glft \qu{Then I was already grown up.}//
-\endgl
-\xe
-
-\ex<yukcop> \yukpa \parencite[143--144]{meira2006syntactic}\\
-\begin{tabular}[t]{@{}lll@{}}
-& \gl{npst} & \gl{pst}\\
-\gl{1} & \obj{=j-a(-s)}&\obj{=j-e}\\
-\gl{2} & \obj{=mak(o)}&\obj{=m-e}\\
-\gl{3} & \obj{=mak(o)}&\obj{=n-e}\\
-\end{tabular}
-\xe
 
@@
 %%
+
+\subsubsection{\rc{eti} and \rc{a[p]} \qu{to be}}
+\label{sec:be}
+For a comprehensive comparative overview of these two roots, readers are referred to \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
+\rc{a[p]} is the original copula and can be reconstructed as already having various irregularities in \PC.
+\rc{eti} is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
+Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam}.
+Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
+\akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while \obj{eʔi} is not attested in a first-person form. 
+\carijo innovated \obj{j-}, but only in the reflex of \rc{eti} \exref{mivida-12}; the \obj{a} root preserves \obj{w-} \pcref{sec:carijo}.
+\yukpa introduced \obj{j-} to the reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
+
+\ex<mivida-12>\carijo \parencite[][177]{robayo1989rame}\\
+\begingl
+\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
+\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
+\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
+\glft \qu{Then I was already grown up.}//
+\endgl
+\xe
+
+\ex<yukcop> \yukpa \parencite[143--144]{meira2006syntactic}\\
+\begin{tabular}[t]{@{}lll@{}}
+	& \gl{npst} & \gl{pst}\\
+	\gl{1} & \obj{=j-a(-s)}&\obj{=j-e}\\
+	\gl{2} & \obj{=mak(o)}&\obj{=m-e}\\
+	\gl{3} & \obj{=mak(o)}&\obj{=n-e}\\
+\end{tabular}
+\xe
 
@@
 \pex<waygodown>\wayana \parencite[][200]{wayanatavares2005}
+\begin{multicols}{2}
 \a<way-73>
@@
 \label{sec:bathe}
-This verb resisted the extensions of \PPek \rc{k-} \pcref{sec:pekodian} and, trivially, \akuriyo \obj{k-} \pcref{sec:akuriyo}.
-It took on new \gl{1}\gl{s_a_} prefixes in \PTir \parentext{\trio \obj{s-epɨ-}, \akuriyo \obj{t͡ʃe+pɨ-} \parencites[292]{triomeira1999}[87]{gildea1994akuriyo}} and \PWai \parentext{\hixka \obj{k-ewehɨ-}, \waiwai \obj{k-ejeɸu-} \parencites[195]{hixkaryanaderby1985}[166]{waiwaihawkins1998}}.
+This verb resisted the extensions of \PPek \rc{k-} \pcref{sec:pekodian} and \akuriyo \obj{k-} \pcref{sec:akuriyo}.
+It took on new \gl{1}\gl{s_a_} prefixes in \PTir \parentext{\trio \obj{s-epɨ-}, \akuriyo \obj{t͡ʃ-epɨ-} \parencites[292]{triomeira1999}[87]{gildea1994akuriyo}} and \PWai \parentext{\hixka \obj{k-ewehɨ-}, \waiwai \obj{k-ejeɸu-} \parencites[195]{hixkaryanaderby1985}[166]{waiwaihawkins1998}}.
 The first person form of its \carijo reflex \obj{ehɨ} \parencites[72]{koch1908hiana} is unknown; an unattested \yukpa cognate may exist.
@@
 Verbs for intransitive \qu{to bathe} are usually typical \gl{s_a_} verbs in Cariban languages, derived with a detransitivizer from a transitive root.
-These roots are reflexes of \rc{pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
-As was shown in \cref{sec:pekodian}, \PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp(ɨ)} (\gl{tr}).
-Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, intransitive \qu{to bathe} is divergent in this branch, changing \rc{e-} to \rc{i}.
-This is an irregular development, since there are no attested reflexes of the detransitivizer in Pekodian that would go back to \PPek \rc{i-} \parencite[506]{meira2010origin}; its cause is unknown.
-However, it should be noted that other languages also show unexpected developments in this verb, considering the glide insertions in Waiwaian or the distribution of \rc{pɨ} and \rc{kupi} in Venezuelan languages.
+These roots are reflexes of \rc{[ɨ]pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
+\PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp} (\gl{tr}).
+Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, \qu{to bathe (\gl{intr})} diverges by changing \rc{e-} to \rc{i}.
+This is an irregular development, since there are no attested reflexes of a Pekodian detransitivizer \rc{i-} \parencite[506]{meira2010origin}; its cause is unknown.
+However, other languages also show unexpected developments in this verb, like the glide insertions in Waiwaian or the distribution of \rc{[ɨ]pɨ} and \rc{kupi} in Venezuelan languages.
 
@@
 \label{sec:verbsummary}
-In \cref{sec:verbs}, the verbs which were unaffected by the extensions in \cref{sec:extensions} were reconstructed, and affected reflexes in the languages under discussion were identified, where possible.
-\cref{tab:overview} gives an overview of what verbs were affected by which extensions, except for \obj{e}-initial \akuriyo verbs unaffected by the extension of \obj{k-}, as they are a large and predictable group.
+In \cref{sec:verbs}, the verbs which were unaffected by the extensions in \cref{sec:extensions} were reconstructed, and affected reflexes in the languages under discussion were identified.
+\cref{tab:overview} gives an overview of what verbs were affected by which extensions (except for \obj{e}-initial \akuriyo verbs unaffected by the extension of \obj{k-}, as they are a large and predictable group).
 In some cases, the verb does not occur at all or just not in first person forms (–), in others the first person form is unknown (?), and in the case of \qu{to go down} the question of affectedness is not meaningfully answerable (\textsc{n/a}), for any of the reasons discussed in \cref{sec:godown}.
 Every \checkmark stands for a verb affected by an extension, while × represents conservatively inflected verbs.
-This overview makes clear just how pervasive the tendency for these verbs to resist innovative markers is, as they do so in different languages.
+This overview makes clear just how strongly these verbs tend to resist person marker extensions, as they do so in different languages.
 
@@
 It is astonishing that the same 1-7 verbs retain their old first person marker in 6 distinct developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs take on innovative markers.
-This suggest that there is some strong motivation for these verbs to not be affected by innovative markers.
-The question arises what properties unite these verbs and make them so conservative across different Cariban languages.
-Possible answers to this question will be discussed in \cref{sec:motivations}, using \posscite{bybee1985morphology} network model of morphology.
+This suggests that there is something about these verbs which causes them to behave in such a way.
+Possible answers to the question of what makes them different from regular \gl{s_a_} verbs will be discussed in \cref{sec:motivations}, using \posscite{bybee1985morphology} network model of morphology.

### 2021-12-09 workflow/helper_functions.py

@@ -198,7 +198,7 @@ def get_verb_citation(v_id, l_id, latex=True, as_tuple=False):
         res = res.iloc[0]
     form = res["Form"].replace("+", "")
     if latex:
-        obj, gloss = f"""\\{get_obj_str(l_id)}{{{form}}}""", f"""\\qu{{{mean_dic[res["Parameter_ID"]].replace("to ", "")}}}"""
+        obj, gloss = f"""\\{get_obj_str(l_id)}{{{form}}}""", f"""\\qu{{{cog_trans_dic[cog_meaning_dic[res["Cognateset_ID"]]].replace("to ", "")}}}"""
     else:
         obj, gloss = f"""{form}""",  f"""'{mean_dic[res["Parameter_ID"]]}'"""
     if as_tuple:
@@ -213,12 +213,12 @@ def extension_string(id, latex=True):
     else:
         return name_dic[ext_lg_dic[id]] + " " + repl_latex(form)
 
-def reconstructed_form_table(lgs, proto, verbs, caption, name):
+def reconstructed_form_table(lgs, proto, verbs, caption, name, sort):
     pyd.x = ["Language_ID"]
     pyd.y = ["Concept"]
     pyd.filters = {"Inflection": ["1"], "Language_ID": lgs, "Concept": verbs}
     pyd.x_sort = lgs
-    pyd.y_sort = ["be_1", "be_2", "say", "go", "come", "go_down", "bathe_intr"]
+    pyd.y_sort = sort
     tabular = pyd.compose_paradigm(i_df.replace({"go_down+?": "go_down"}))
     sources = get_sources(i_df)
     raw_sources = get_sources(i_df, latexify=False)

### 2021-12-09 workflow/process_data.py

@@ -1,5 +1,6 @@
 from helper_functions import *
 
+verb_list = ["say", "go", "be_1", "be_2", "come", "go_down", "bathe_intr"]
 
 # tables for the introduction, featuring "regular" and "irregular" verbs from Trio and Hixkaryana
 person = ["1", "2", "1+2", "3"]
@@ -86,6 +87,7 @@ reconstructed_form_table(
     verbs,
     r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek",
     "ppekverbs",
+    verb_list
 )
 
 
@@ -136,6 +138,7 @@ reconstructed_form_table(
     verbs,
     r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai",
     "pwaiverbs",
+    verb_list
 )
 
 
@@ -184,6 +187,7 @@ reconstructed_form_table(
     ["go", "say", "come", "be_1", "be_2"],
     r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir",
     "ptirverbs",
+    verb_list
 )
 
 # regular akuriyo Sa verbs
@@ -456,7 +460,7 @@ bathe_tables = [
         "bathe_intr_3",
         True,
     ),
-    (tr, r"Reflexes of \rc{(ɨ)pɨ} \qu{to bathe (\gl{tr})}", "bathe_tr_1", False),
+    (tr, r"Reflexes of \rc{[ɨ]pɨ} \qu{to bathe (\gl{tr})}", "bathe_tr_1", False),
     (tr_1, r"Reflexes of \rc{kupi} \qu{to bathe (\gl{tr})}", "bathe_tr_2", False),
 ]
 
@@ -506,7 +510,6 @@ f.close()
 
 # overview of what extensions affected what verbs
 i_df = i_df[i_df["Inflection"] == "1"]
-verb_list = ["say", "go", "be_1", "be_2", "come", "go_down", "bathe_intr"]
 
 
 overview_legend = {
@@ -603,9 +606,9 @@ export_csv(
 )
 
 # add nice-looking checkmarks and stuff
-repl_dic = {x: y["tex"] for x, y in overview_legend.items()}
+bool_repl_dic = {x: y["tex"] for x, y in overview_legend.items()}
 result.replace(
-    repl_dic,
+    bool_repl_dic,
     inplace=True,
 )
 
@@ -769,13 +772,17 @@ for i, ext in e_df.iterrows():
[…]

### 2021-12-09 workflow/read_data.py

@@ -52,6 +52,7 @@ cognum = dict(zip(cognatesets["ID"], cognatesets.index.astype(str)))
 numcog = dict(zip(cognatesets.index.astype(str), cognatesets["ID"]))
 
 cog_verb_df = cs_df[~(cs_df["Form"].str.contains("-"))]
+
 cog_trans_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Meaning"]))
 cog_form_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Form"]))
 cog_form_dic["bathe_intr"] = "e-pɨ"
@@ -59,3 +60,7 @@ cog_form_dic["come"] = "(ət-)jəpɨ"
 cog_trans_dic["bathe_intr"] = r"bathe"
 cog_trans_dic["be_1"] = r"be-1"
 cog_trans_dic["be_2"] = r"be-2"
+
+cog_meaning_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["ID"]))
+cog_meaning_dic["DETRZ1+bathe_1"] = "bathe_intr"
+cog_meaning_dic["DETRZ+come"] = "come"

### 2021-12-09 documents/floats/pwaireg.tex
added brackets and initial vowels to some proto-forms

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2021-12-10 documents/sections/introduction.tex
finetuning readthrough

@@
 The family is concentrated in Venezuela, the Guianas and Northern Brazil, with two Western and four Southern outliers.
-\cref{fig:phylomap} gives an overview of the geographical distribution and genealogical affiliation of the extant Cariban languages.
+\cref{fig:phylomap} shows the geographical distribution and genealogical affiliation of the extant Cariban languages.
 %Cariban languages feature relatively rich verbal morphology, both pre- and suffixes, inflecting for person, number, tense, aspect, and evidentiality, combined with a range of valency-modifying affixes.
@@
 
-Some Cariban languages show a small group of verbs which diverge in their first person inflection pattern, a topic which has not received much attention in the literature.
-This is illustrated for \hixka in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
+In some Cariban languages, a small group of verbs show a divergent first person inflection pattern, a topic which has not received much attention in the literature.
+This is illustrated with person paradigms of four \hixka verbs in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
 This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms shown here.
@@
 \textcite{gildea2018reconstructing} is followed in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.
-Glossing abbreviations: }%\glossingAbbrevsComma.
-which shows person paradigms for four verbs, all members of the \gl{s_a_} inflectional class.
-In this language, the verb \qu{to be} diverges in its first person marker (\obj{w-}), contrasting with other \gl{s_a_} verbs like \qu{to fall}, which have \obj{k-}.
-A similar pattern can be found in \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
-In both languages, the inflectional patterns of the verbs on the left of the table is representative for the vast majority of \gl{s_a_} verbs.
+Glossing abbreviations: } %\glossingAbbrevsComma.
+all members of the \gl{s_a_} inflectional class.
+In this language, the verb \qu{to be} diverges from other \gl{s_a_} verbs like \qu{to fall} by having a first person marker \obj{w-}, rather than \obj{k-}.
+A similar pattern exists in \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
+In both languages, the first person prefix of the verbs on the left is representative for the vast majority of \gl{s_a_} verbs.
 %In both languages, there are only a few other verbs inflected identically to the divergent ones on the right; for example, the first-person form of \trio \qu{to be} is \obj{w-ei-} \parencites[339]{triomeira1999}.
 
-Such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}, but have only been subject to comparative analysis for \posscite{meira1998proto} reconstruction of \PTar.
-In a synchronic analysis of a language, these verbs and their first person prefixes may be called \textsc{irregular}, contrasting with regular prefixes, like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}, on regular verbs.
+Such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}, but have only been subject to comparative scrutiny in \posscite{meira1998proto} reconstruction of \PTar.
+In a synchronic analysis of a language, these verbs and their first person prefixes may be called \textsc{irregular}, contrasting with regular prefixes (like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}) on regular verbs.
 However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[e.g.,][]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
@@
 
-Regardless of the synchronic analysis, the reason for these inflectional patterns can be found in the diachrony of the languages in question.
-The purpose of this paper is to approach the patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows.
-In \cref{sec:background}, readers are provided with necessary background knowledge about \PC verbs, and it is shown that the mechanism of person marker extension is responsible for the \hixka and \trio patterns.
-In \cref{sec:data}, six incomplete person marker extensions and their extent in the lexicon are described.
-Since there is a surprising amount of agreement about what verbs remain conservative, they are compared and reconstructed.
-\cref{sec:motivations} uses \posscite{bybee1985morphology} network model of morphology to find explanations for the verbs (not) affected by each extension.
-\cref{sec:discussion} summarizes and discusses the results.
+Ignoring the specifics of synchronic analysis, the cause for the divergent inflectional patterns lies in the diachrony of the languages in question.
+The goal of this study is to approach the patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows:
+In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns like the \hixka and \trio ones.
+In \cref{sec:data}, six incomplete person marker extensions and the verbs unaffected by them are described.
+Since conservative verbs show a considerable etymological overlap between languages, they are further discussed and reconstructed.
+\cref{sec:motivations} uses \posscite{bybee1985morphology} network model of morphology to search explanations for the verbs (un-)affected by each extension.
+\cref{sec:discussion} summarizes and discusses the results of the study.
 

### 2021-12-10 documents/sections/proto_carib.tex
finetuning readthrough

@@
 \label{sec:background}
-The irregular first person prefixes introduced in \cref{sec:intro} are relics, inherited from the ancestral \PC system \pcref{sec:pc_person}.
-That system was subject to many different kinds of innovations; responsible for the patterns discussed are person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
-A particular component of the system, the \gl{s_a_} vs \gl{s_p_} distinction, plays a role in what extensions are incomplete and is discussed in \cref{sec:split}.
+The irregular first person prefixes from \cref{sec:intro} are relics, inherited from the ancestral \PC system \pcref{sec:pc_person}.
+That system underwent much innovation; the mechanism responsible for the irregular forms are person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
+A specific aspect of the system, the \gl{s_a_} vs \gl{s_p_} distinction, plays a role in incomeplete extensions and is discussed in \cref{sec:split}.
 %\cref{sec:outlook} summarizes the main points of this section and provides an outlook for the rest of the paper.
@@
 \PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.
-The choice of person marker in transitive verbs can be characterized as being conditioned by a basic person hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
+Person indexation in transitive verbs was conditioned by a basic hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
 The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
@@
 Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
-That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}>\gl{3} \rc{t(i)-} vs \gl{1}\gl{s_a_} \rc{w-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
-On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones, with \gl{3}\gl{s_p_} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
+That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}\gl{s_a_} \rc{w-} vs \gl{1}>\gl{3} \rc{t(i)-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
+On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones, and \gl{3}\gl{s_p_} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
 
-Equipped with this knowledge about the ancestral system, it becomes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \textsc{conservative}.
+Knowledge about the ancestral system makes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \textsc{conservative}.
 They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
-The reflexes of \rc{w-} may be considered \textsc{relics}, old and restricted to specific lexically conditioned contexts, contrasting with the innovative prefixes found elsewhere.
+The reflexes of \rc{w-} are \textsc{relics}, old and restricted to a few lexemes, contrasting with the innovative prefixes found elsewhere.
 These verbs and their prefixes are comparable with the few English nouns like \obj{ɒks}, which preserve the old plural suffix \obj{-ən}.
 It was once more widespread as the normal plural suffix of the weak inflection, compare German \obj{ɔks-ən} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
-Since the regular \hixka and \trio prefixes are innovative, the next question is where they came from.
+
+Since the regular \hixka and \trio prefixes are innovative, one may ask where they came from.
 
@@
 \label{sec:extensions_intro}
-In his discussion of the \PC split-\gl{s} system and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different modifications in various languages.
-The main mechanism of change leading to these modifications are \textsc{person marker extensions}, i.e. verbal person prefixes being extended to paradigm cells previously occupied by other prefixes.
+In his discussion of the \PC split-\gl{s} system \pcref{sec:split} and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different changes in daughter languages.
+The main mechanism of these changes are \textsc{person marker extensions}, i.e. person prefixes being extended to verbal paradigm cells previously occupied by other prefixes.
 %\footnote{
@@
 %For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
-There have been many different person marker extensions in Cariban languages, and some are still ongoing.
-\textcite{gildea1998} illustrates this using the three Parukotoan languages as an example.
+There have been quite a few person marker extensions in Cariban languages, some still ongoing.
+\textcite{gildea1998} illustrates this with the three Parukotoan languages \kaxui, \hixka, and \waiwai.
 Apart from segmental changes to individual morphemes, the following innovations happened in the \setone paradigm in Parukotoan:
 
-\begin{enumerate}
-	\item \PPar \begin{enumerate}
-		\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
-		\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
-		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
-	\end{enumerate}
-	\item \PWai \begin{enumerate}
-		\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
-		\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
-	\end{enumerate}
-	\item \waiwai \begin{enumerate}
-		\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
-\end{enumerate}
+\ex
+\begin{tabular}[t]{ll}
+\PPar & \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}\\
+& \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)\\
+& \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)\\
+\PWai & \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}\\
+& \rc{owɨ(ro) j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}\\
+\waiwai & \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}
+\end{tabular}
+\xe
+%
+%\begin{enumerate}
+%	\item \PPar \begin{enumerate}
+%		\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
+%		\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+%		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+%	\end{enumerate}
+%	\item \PWai \begin{enumerate}
+%		\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
+%		\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
+%	\end{enumerate}
+%	\item \waiwai \begin{enumerate}
+%		\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
+%\end{enumerate}
 %
@@
 %
-\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows the variation in the first person and \gl{1+2} prefixes that is reconstructible to \PPar.
+\hixka has preserved split-\gl{s} only in the second person prefixes, while \kaxui still shows the variation reconstructible to \PPar in its first person and \gl{1+2} prefixes.
 \waiwai has lost the system entirely, which notably happened via three diachronically distinct innovations.
@@
 
-\textcite{gildea1998} discusses person marker extensions relating to the loss of split-\gl{s} and the accompanying changes to indexing alignment, but a different aspect of these extensions is relevant for this study.
-Namely, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}; this hypothesis is supported by three facts.
+For \textcite{gildea1998}, person marker extensions are relevant for loss of split-\gl{s} and the accompanying changes to indexing alignment, whereas this study focuses on a different aspect of these extensions.
+Namely, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}, a hypothesis supported by three facts.
 First, the variation in first person and \gl{1+2} prefixes described above for \kaxui is not completely free.
-Rather, some verbs only allow for example first person \obj{k-}, but not \obj{j-}, while others can occur with both, which is the expected pattern in a lexical diffusion scenario.
-In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which is what one would expect from a change in progress.
+Some verbs only allow e.g.\ first person \obj{k-}, but not \obj{j-}, while others can occur with both, a pattern expected in a lexical diffusion scenario.
+In addition, this is speaker-dependent \perscommpar{Spike Gildea}, which points to an ongoing change.
 %todo maybe put in GOOD kaxui data
-Second, while there is no detailed diachronic scenario for the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} convincingly argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading at the same time.
-Whether this gradual switch was along ordered lines or not, lexical diffusion must have played a role.
+Second, while there is no detailed diachronic account of the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading simultaneously.
+Whether or not this gradual switch followed ordered lines, lexical diffusion must have played a role.
 
-The third argument in favor of the lexical diffusion scenario relates to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
-In both cases, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
-The continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs is best interpreted as the extension stopping short of these verbs, rather than spreading through the entire \gl{s_a_} lexicon.
-In a family-wide investigation of person marker extensions, 19 distinct extensions affecting intransitive verbs were identified, 6 of them incomplete.
-These incomplete extensions have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages, and are treated individually in \cref{sec:data}.
+The third argument in favor of the lexical diffusion scenario goes back to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
+Both innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
+The continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs is the result of the extension stopping short of these verbs, rather than spreading through the entire \gl{s_a_} lexicon.
+In a family-wide search, 18 distinct extensions affecting intransitive verbs were identified, 6 of them incomplete.
+The latter have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages \pcref{sec:data}.
 
-Interestingly, all six extensions featured innovative first person markers on \gl{s_a_} verbs.
+Interestingly, all six featured innovative first person markers on \gl{s_a_} verbs.
 All other (complete) extensions%
-\footnote{As an honorable mention, when \ikpeng introduced a new mixed person paradigm to \setone constructions, \obj{a} \qu{to be} and \obj{ke} \qu{to say} retained third person \obj{n-} \parencite[12]{matter2019arara}.
-	However, the spread of the innovative third person markers had an entirely different pathway from the extensions discussed here, spreading from subordinate to main clauses, rather than from verbs to verbs.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
+\footnote{As an honorable mention, when \ikpeng replaced third person \setone with \settwo prefixes, \obj{a} \qu{to be} and \obj{ke} \qu{to say} retained \obj{n-} \parencite[12]{matter2019arara}.
+	However, the spread of the innovative markers had an entirely different pathway from the extensions discussed here, spreading from subordinate to main clauses, rather than from verbs to verbs.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
 Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, and the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
 %, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
-This starkly different behavior of \gl{s_a_} and \gl{s_p_} verbs when subject to extensions points to the split-\gl{s} system playing some role, so its main properties will be discussed in \cref{sec:split}.
-This will also explain why it is possible for the \gl{s_a_}/\gl{s_p_} distinction to be lost only for a single person, or for \gl{s_p_} verbs to take on \gl{s_a_} markers with apparent semantic impunity.
+The starkly different behavior of \gl{s_a_} and \gl{s_p_} verbs regarding extensions points to the split-\gl{s} system playing a role, so its main properties will be discussed in \cref{sec:split}.
+It will also be made clear how the \gl{s_a_}/\gl{s_p_} distinction can be lost for a single person, or how \gl{s_p_} verbs can take on \gl{s_a_} markers with apparent semantic impunity.
 
@@
 %
-Many languages show an \gl{s_a_} class marker in deverbalized forms, originating in \PC \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
+%todo CONTINUE HERE
+\textcite[208]{meira2000split} Many languages show an \gl{s_a_} class marker in deverbalized forms, originating in \PC \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
 With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes \pcref{tab:participles,tab:nominalizations}.
@@
 %
-\input{floats/participles}%
+%\input{floats/participles}%
 %
-\input{floats/nominalizations}%
+%\input{floats/nominalizations}%
 %
-\input{floats/imperatives}%
+%\input{floats/imperatives}%
 %
-The distinction between \gl{s_a_} and \gl{s_p_} was also borne out in imperatives, where the latter took the \gl{p}-oriented second person prefix \rc{ə(j)-} while the former received no prefix \pcref{tab:imperatives}.
+The distinction between \gl{s_a_} and \gl{s_p_} was also reflected in imperatives, where the latter took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} while the former were unprefixed .
 %This is illustrated with reflexes in various modern languages in \pcref{tab:imperatives}.
@@
 
-In the modern languages maintaining the split-\gl{s} system, mismatches between the semantics of verbs and their \gl{s_a_} or \gl{s_p_} membership are common, exemplified with \kalina data in \exref{kar2}.
+In the modern instantiations of split-\gl{s}, mismatches between the semantics of verbs and their \gl{s_a_} or \gl{s_p_} membership are common, exemplified with \kalina data in \exref{kar2}.
 %

### 2021-12-11 documents/sections/proto_carib.tex
more finetuning

@@
 \footnote{As an honorable mention, when \ikpeng replaced third person \setone with \settwo prefixes, \obj{a} \qu{to be} and \obj{ke} \qu{to say} retained \obj{n-} \parencite[12]{matter2019arara}.
-	However, the spread of the innovative markers had an entirely different pathway from the extensions discussed here, spreading from subordinate to main clauses, rather than from verbs to verbs.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
+	However, the spread of the innovative markers had an entierly different dynamic, spreading from subordinate to main clauses.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
 Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, and the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
@@
 \label{sec:split}
-As seen in \cref{sec:pc_person}, the distinction between \gl{s_a_} and \gl{s_p_} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only inflectional difference:
-%\begin{inlinelist}
-%	\item presence vs absence of the \gl{s_a_} marker \rc{w-}
-%	\item absence vs presence of the second person prefix \rc{ə(j)-} in imperatives
-%	\item presence vs absence of a derivational detransitivizing prefix
-%\end{inlinelist}.
-%
-%todo CONTINUE HERE
-\textcite[208]{meira2000split} Many languages show an \gl{s_a_} class marker in deverbalized forms, originating in \PC \rc{w-}.\footnote{See \textcite[227]{meira2000split}, who identifies reflexes of this morpheme as having \dbqu{no purpose other than being \qu{class markers}, without any obvious semantic or functional load}.}
-With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes \pcref{tab:participles,tab:nominalizations}.
+As seen in \cref{sec:pc_person}, the split between \gl{s_a_} and \gl{s_p_} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only inflectional criterion:
+In deverbalized forms, \gl{s_a_} verbs took a class marker \rc{w-}, while \gl{s_p_} verbs lacked that prefix \parencites[89, 141--142]{gildea1998}[208]{meira2000split}.
+The distinction between \gl{s_a_} and \gl{s_p_} was also reflected in imperatives, where the latter took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} while the former were unprefixed \parencites[89]{gildea1998}[208]{meira2000split}.
+
+%\parencite[208]{meira2000split} Many languages show reflexes \gl{s_a_}  in deverbalized forms, originating in .
+%With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes \pcref{tab:participles,tab:nominalizations}.
 %Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
@@
 %
-The distinction between \gl{s_a_} and \gl{s_p_} was also reflected in imperatives, where the latter took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} while the former were unprefixed .
+%The distinction between \gl{s_a_} and \gl{s_p_} was also reflected in imperatives, where the latter took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} while the former were unprefixed .
 %This is illustrated with reflexes in various modern languages in \pcref{tab:imperatives}.
@@
 
-In the modern instantiations of split-\gl{s}, mismatches between the semantics of verbs and their \gl{s_a_} or \gl{s_p_} membership are common, exemplified with \kalina data in \exref{kar2}.
+In modern instantiations of Cariban split-\gl{s}, mismatches between the semantics of verbs and their \gl{s_a_} or \gl{s_p_} status are common, exemplified with \kalina data in \exref{kar2}.
 %
@@
 \gla sipi tɨnka-rɨ m-ekema-non hen//
-\glb net pull-\gl{nmlz} \gl{2}-be.afraid-\gl{prs}.\gl{uncert} eh?//
+\glb net pull-\gl{nmlz} \gl{2}\gl{s_a_}-be.afraid-\gl{prs}.\gl{uncert} eh?//
 \glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
@@
 \gla aj-awomɨ-ja//
-\glb \gl{2}-get.up-\gl{prs}//
+\glb \gl{2}\gl{s_p_}-get.up-\gl{prs}//
 \glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
@@
 %
-In \exref{kar2.kar-59}, the \gl{s_a_} verb \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+The \gl{s_a_} verb \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
 In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
@@
 Rather, the reason for a verb to take \gl{a}- or \gl{p}-oriented prefixes is (at least diachronically) a morphological one.
-\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
+\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a derivational detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
 \begin{quotebox}{\parencite[201]{meira2000split}}
@@
 
-Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although on verbs their reflexes have been merged into a single morpheme in modern languages.
-Modern reflexes of \detrz show a range of meanings summarizable as \dbqu{detransitive}, illustrated with \trio \gl{s_a_} verbs in \exref{triodetrz}.
+Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although their reflexes on verbs have been merged into a single morpheme in modern languages.
+Reflexes of \detrz show a range of meanings summarizable as \dbqu{detransitive}, illustrated with \trio \gl{s_a_} verbs in \exref{triodetrz}.
 
@@
 %
-The morphological variation featured in \qu{to abandon each other} and \qu{to wash self} is due to the collapse between the two \PC prefixes:
-\obj{e-} is a reflex of the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
+The morphological variation in \qu{to abandon each other} and \qu{to wash self} is due to the collapse of the two \PC prefixes:
+\obj{e-} comes from the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
 However, both can occur with either meaning -- at least for these two verbs.
-In the next section, it will become evident that many of the verbs not affected by person marker extensions belong to the small group of \gl{s_a_} verbs without a reflex of \detrz.
+In the next section, it will become apparent that many of the verbs not affected by person marker extensions belong to the small group of \gl{s_a_} verbs without a reflex of \detrz.
 %As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.

### 2021-12-11 workflow/process_data.py

@@ -283,6 +283,7 @@ save_float(
     print_latex(table, keep_index=True),
     "kaxprog",
     "\\kaxui \\obj{johɨ} \\qu{to come} compared with other verbs (Spike Gildea, p.c.)",
+    "\\kaxui \\obj{johɨ} \\qu{to come} compared with other verbs"
 )
 
 

### 2021-12-11 documents/sections/extensions.tex
extension finetuning

@@
 \label{sec:data}
-As was shown in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
-\cref{sec:extensions} presents the six known incomplete extensions, the prefixes they introduced and the verbs they spared.
-Since the latter show considerable etymological overlap across languages, the resistant verbs are compared and reconstructed in \cref{sec:verbs}.
-In addition, their reflexes which did get affected by one of the extensions are identified, where existent.
+As shown in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
+\cref{sec:extensions} presents the six identified incomplete extensions, the prefixes they introduced and the verbs they spared.
+Since the latter show considerable etymological overlap across languages, these resistant verbs are compared and reconstructed in \cref{sec:verbs}.
+Where existent, their reflexes which did get affected by one of the extensions are identified.
 %The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
@@
 \label{sec:extensions}
-As stated in \cref{sec:extensions_intro}, the six person marker extensions which did not affect all potential targets all have in common that they feature innovative first person markers on verbs that are (at least historically, in the case of loss of split-\gl{s}) members of the \gl{s_a_} class.
-Of these extensions, half can be reconstructed to intermediate proto-languages, while the others happened in pre-modern stages of single languages.
+As stated in \cref{sec:extensions_intro}, the six person marker extensions which did not affect all potential targets have in common that they introduced innovative first person markers on \gl{s_a_} verbs.
+Of these extensions, half can be reconstructed to intermediate proto-languages, and half happened in pre-modern stages of single languages.
 The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
-For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and verb forms are reconstructed where necessary:
+For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and verb forms are reconstructed where necessary.
 \cref{sec:pekodian} details the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
@@
 The Pekodian branch consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
-The contribution establishing the branch \parencite{meira2005southern} focuses on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
+The contribution establishing the branch \parencite{meira2005southern} focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
 However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
@@
 
-In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.\footnote{
-	It should be noted at this point that \textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
+In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.\footnote{\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
 	However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
@@
 }
-The verb used to illustrate the first group is \obj{i} \qu{to bathe} \exref{bak-33}.
+The first group is illustrated with \obj{i} \qu{to bathe} \exref{bak-33}.
 
@@
 %Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
-While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of \arara verbs in \exref{arairr} as \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.
-Inflected forms can be found in \textcite{von1892bakairi}, presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
+While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of the \arara verbs in \exref{arairr} as \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.
+However, inflected forms can be found in \textcite{von1892bakairi}, presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
 
@@
 All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
-However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}.
-The exception is \qu{to go}, which has \obj{k-} \exref{ikp-170}.
+However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}, with the exception of \obj{k-}prefixed \qu{to go} \exref{ikp-170}.
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
@@
 In contrast, \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following \posscite[114]{meira1998proto} line of reasoning for a similar pattern in the three Taranoan languages (see also \cref{sec:taranoan}), the idiosyncratic pattern in \arara should be reconstructed to \PPek, while \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.
+Following \posscite[114]{meira1998proto} line of reasoning for a similar pattern in Taranoan (see also \cref{sec:taranoan}), the idiosyncratic pattern in \arara can be reconstructed to \PPek, meaning that \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.
 
@@
 \label{sec:waiwaian}
-This extension, one of the Parukotoan innovations discussed in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
-In \PWai, the \gl{1}\gl{s_p_} prefix \rc{k-} innovated in \PPar was further extended to \gl{1}\gl{s_a_}.
+This extension, one of the Parukotoan innovations shown in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
+\PWai further extended the \gl{1}\gl{s_p_} prefix \rc{k-} innovated in \PPar to \gl{1}\gl{s_a_}.
 For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
@@
 
-There are three more verbs which did not take innovative \rc{k-} in \PWai, shown alongside \rc{ka} \qu{to say} in \cref{tab:pwaiverbs}.
-The two forms of \qu{to be} are unproblematic, whereas \qu{to go} is a special case.
+There are three more verbs which did not take innovative \rc{k-} in \PWai \pcref{tab:pwaiverbs}.
+The two forms for \qu{to be} are unproblematic, whereas \qu{to go} is a special case.
 While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
@@
 \label{sec:taranoan}
-The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages already identified by \textcite{girard1971proto}, with \carijo as a more distant member.
+The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages identified by \textcite{girard1971proto}, with \carijo as a more distant member.
 \textcite{meira1998proto} contributes an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
@@
 Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.
-To this list the second copular root \rc{eʔi} can be added, whose \trio reflex retains first person \obj{w-}.
-The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is plausibly identified as a reflex of \rc{wɨ-} by 
-\textcite[113]{meira1998proto}, which is supported by additional evidence:
-Both components of the irregular change \rc{wɨ-} > \obj{ə-} (loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
+The other copular root \rc{eʔi} can be added as a fifth verb, whose \trio reflex retains first person \obj{w-}.
+The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is identified as a reflex of \rc{wɨ-} by 
+\textcite[113]{meira1998proto}, which is supported the fact that both components of the irregular change \rc{wɨ-} > \obj{ə-} (loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
 
 \pex<exp>\akuriyo
+\begin{multicols}{2}
 \a<aku-164>
@@
 
-For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for other persons, based on an idiosyncratic paradigmatic pattern in \trio and the vowel length in \akuriyo.
+For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for other persons, based on such a paradigmatic pattern in \trio and the vowel length in \akuriyo.
 Both \akuriyo and \carijo then levelled this original distribution, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
-This scenario is plausible, with the exception that \trio \obj{əepɨ} is a reflex of \rc{ət-epɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
+This scenario is plausible, with the exception that \trio \obj{əepɨ} is a reflex of \rc{ət-jəpɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
 

### 2021-12-11 documents/sections/extensions.tex
more extension finetuning

@@
 \label{sec:akuriyo}
-After the split-up of \PTir, when \rc{t-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{s_a_} marker: \obj{k-}.
+After the split of \PTir, when \rc{t-} and \rc{t͡ʃ-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{s_a_} marker \obj{k-}.
 It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
-\textcite[107]{meira1998proto} largely confirms the distribution shown here, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} might be more recent, which is very plausible: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
+\textcite[107]{meira1998proto} largely confirms that distribution, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
+He also suggests that \obj{k-} might be more recent, which is plausible: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
 The few \obj{t-} mentioned by \textcite{meira1998proto} were then perhaps reintroduced under \trio influence.
-However, since there are no examples of \obj{ə}-initial verbs with \obj{t-}, or further information about them, these cases cannot be discussed further.
+However, since there are no examples of, or more information about, \obj{ə}-initial verbs with \obj{t-}, these cases cannot be discussed further.
 
@@
 
-The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} did of course also not introduce \obj{k-} in \akuriyo, although the first-person form of the copular verb \obj{eʔi} is unknown.
-In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-} in, which is apparently reconstructible to \PTir \exref{tridown}.
-It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{s_a_} verb when \PTir \rc{t-} was extended is unclear (see \cref{sec:godown}).
+The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} of course also resisted the extension of \obj{k-} in \akuriyo, although the first-person form of the copular verb \obj{eʔi} is unknown.
+In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-}, apparently reconstructible to \PTir \exref{tridown}.
+It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{s_a_} verb when \PTir \rc{t-} was introduced is unclear (see \cref{sec:godown}).
 
@@
 \carijo, the cousin of the Tiriyoan languages, has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
-Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs \pcref{tab:carreg}.
+Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regular verbs \pcref{tab:carreg}.
 %
@@
 %We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
-Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would expect \obj{ka} \qu{to say} to either have conservative \obj{wɨ-} or innovative \obj{ji-}, so the zero is unexpected.
+Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would expect \obj{ka} \qu{to say} to either have conservative \obj{wɨ-} or innovative \obj{ji-}, so the zero needs explanation.
 It is analysed here as a reflex of \rc{wɨ-}, primarily due to the loss of \rc{w} in \ikpeng and \hixka.
-While those cases of loss were more regular, an already irregular marker undergoing idiosyncratic phonological erosion is not that surprising, see \akuriyo \rc{wɨ-} > \obj{ə-} in the preceding section.
-Alternatively, it is possible that the divergent development of \rc{w-} on C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} is a result of the latter's originally V-initial nature \pcref{sec:go}.
+While those developments were more regular, an already irregular marker undergoing idiosyncratic phonological erosion is not that surprising, see \akuriyo \rc{wɨ-} > \obj{ə-} in the preceding section.
+Alternatively, the divergent development of \rc{w-} on \obj{ka} \qu{to say} and \obj{tə} \qu{to go} may be due to the latter's originally V-initial nature \pcref{sec:go}.
 
@@
 The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
-However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm \pcref{tab:yukreg} is found.
+However, it does preserve the \setone prefixes in the immediate past, which shows a unified intransitive paradigm \pcref{tab:yukreg}.
 The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (now former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
@@
 These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their semantics and the reflex of \detrz.
-The same prefix occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, so \obj{j(ɨ)-} is the reflex of the \PC \gl{1}S\textsubscript{(P)} marker \rc{u(j)-}.
+Since \obj{j(ɨ)-} is the reflex of the \PC \gl{1}S\textsubscript{(P)} marker \rc{u(j)-} \parencite[92]{gildea1998}, it also occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.
 In contrast, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.
@@
 %
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.
+Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \gl{1}\gl{s_a_} \rc{w-}.
 
 In intransitive verbs, this first-person zero marking is attested in a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
-It diverges from regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.
+It diverges from regular C-initial verbs with \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.
 It can thus be identified as having resisted the extension of \obj{j-} in \yukpa.

### 2021-12-12 documents/sections/verbs.tex
more finetuning

@@
 \label{sec:verbs}
-In \cref{sec:extensions}, six incomplete extensions of personal prefixes into \gl{1}\gl{s_a_} territory were presented, and the verbs resistant to these innovations were identified.
-The set of unaffected verbs is rather small in most cases, and (proto-)languages show a considerable degree of overlap in what verbs are unaffected.
-Here,  resistant verbs are investigated from a comparative perspective:
-\rc{ka[ti]} \qu{to say} \pcref{sec:say}, \rc{ɨtə[mə]} \qu{to go} \pcref{sec:go}, both roots of the copula \rc{eti}/\obj{a[p]} \pcref{sec:be}, \rc{(ət)jəpɨ} \qu{to come} \pcref{sec:come},  \rc{ɨpɨtə} \qu{to go down} \pcref{sec:godown}, and  \cref{sec:bathe} investigates \rc{e-pɨ} \qu{to bathe}, of which the \PPek reflex \rc{ipɨ} resisted the extension of \rc{k-}.
-The \obj{e}-initial verbs not affected by the extension of \obj{k-} in \akuriyo \pcref{sec:akuriyo} will not be discussed here, as they are a large and phonologically coherent group.
+In \cref{sec:extensions}, six incomplete extensions of personal prefixes into \gl{1}\gl{s_a_} territory and the verbs unaffected by them were identified.
+This set is rather small in most cases, and many of the verbs remain conservative in different (proto-)languages.
+Here, these verbs are investigated from a comparative perspective:
+\rc{ka[ti]} \qu{to say} \pcref{sec:say}, \rc{ɨtə[mə]} \qu{to go} \pcref{sec:go}, both roots of the copula \rc{eti}/\obj{a[p]} \pcref{sec:be}, \rc{(ət)jəpɨ} \qu{to come} \pcref{sec:come},  \rc{ɨpɨtə} \qu{to go down} \pcref{sec:godown}, and \rc{e-pɨ} \qu{to bathe} \pcref{sec:bathe}.
+The large and phonologically coherent group of \obj{e}-initial verbs not affected by the extension of \akuriyo \obj{k-} \pcref{sec:akuriyo} will not be discussed.
 
@@
 \label{sec:say}
-Most reflexes of this verb are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, best visible in the imperative forms of some languages.% \exref{kati}.
+This verb was not affected by any of the extensions in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}), while the first person form of its \yukpa reflex \obj{ka} is unattested.
+Most reflexes are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, best visible in the imperative forms of some languages. % \exref{kati}.
 %
@@
 \cref{tab:say} shows a comparison of the longest attested forms for each language.%
-\footnote{Cognate segments in \crefrange{tab:say}{tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for easier recognition of correspondences.}
+\footnote{Cognate segments in \crefrange{tab:say}{tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for improved exposition of correspondences.}
 %
@@
 %
-This verb was not affected by any of the extensions found in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}).
-The first person form of its \yukpa reflex \obj{ka} is unattested.
 
-As mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes this verb as transitive in \hixka.
-This analytical choice is not only motivated by a desire to avoid an idiosyncratic intransitive first person prefix \obj{ɨ-} instead of regular \obj{kɨ-}.
+As mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes the \hixka reflex as transitive, an analytical choice not only motivated by a desire to avoid an idiosyncratic intransitive first person prefix \obj{ɨ-} instead of regular \obj{kɨ-}.
 \hixka \obj{ka} also shows the complementary distribution of third person  \obj{n-} and preceding objects typical of transitive verbs in Cariban \parencite[60--81]{gildea1998}.
-Due to the semantics of \qu{to say}, these objects are either ideophones or direct speech \exref{hixsay}.
+Due to its semantics, these objects are either ideophones or direct speech \exref{hixsay}.
 
@@
 \xe
-In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}).
-It does not occur in \exref{hixsay.ordpl-35}, because \qu{they said} is preceded by direct speech.
-Looking beyond \hixka, at least the \trio cognate shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
+In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}), while does not occur in \exref{hixsay.ordpl-35} where the verb is preceded by direct speech.
+Looking beyond \hixka, at least the \trio reflex shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
 
-Derivational suffixes also point to \rc{ka[ti]} \qu{to say} being transitive:
-\trio \obj{ka} is characterized as the only intransitive verb being able to take the causative suffix \obj{-po} and the agentive nominalizer \obj{-ne} \parencite[263, 169]{triomeira1999}.
-The exceptionality of \obj{ka} \qu{to say} taking \obj{-po} \qu{\gl{caus}.\gl{tr}} has also been noted for \kalina \parencite[82]{courtz2008carib} and \wayana \parencite[258]{wayanatavares2005}.
-The agent nominalizer \rc{-ne} gave rise to the \panare gnomic verbal suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
-The occurrence of \obj{-ɲe} on \obj{ka} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, contrasting with the intransitive analysis by \textcite[102]{mattei1994diccionario}.
-Finally, reflexes of the causativizer \rc{-metipo}, usually restricted to transitive verbs \parencite{gildea2015valency}, are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
+\rc{ka[ti]} \qu{to say} also shows transitive patterns in its derivational suffixes:
+In \trio, it is the only intransitive verb taking transitive \obj{-po} (\gl{caus}) and the \obj{-ne} (\gl{agt}.\gl{nmlz}) \parencite[263, 169]{triomeira1999}.
+It also exceptionally takes the former suffix in \kalina \parencite[82]{courtz2008carib} and \wayana \parencite[258]{wayanatavares2005}.
+The agent nominalizer \rc{-ne} became the \panare inflectional suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
+The combination of \obj{ka} and \obj{-ɲe} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, although \textcite[102]{mattei1994diccionario} arrived at the opposite conclusion.
+Finally, reflexes of the \parentext{\parencite{gildea2015valency}} causativizer \rc{-metipo} are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
 
-The categorization of \qu{to say} as an intransitive verb is supported primarily by its person prefixes.
-\kalina offers a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.
-Similarly, \qu{to say} has a reflex of \gl{1}\gl{s_a_} \rc{w-} in Pekodian languages \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
-Additionally, languages which differentiate direct prefixes from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} show \obj{ɨ} instead of \obj{i} for this verb, see \akuriyo in \exref{sayintr.aku-154}, as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for cognate forms in other such languages.
-Finally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and it is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
+Arugments in favor of intransitive \qu{to say} primarily come from its inflectional prefixes.
+\kalina has a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.
+Similarly, \PPek \rc{ke} \qu{to say} took \gl{1}\gl{s_a_} \rc{w-} \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
+Additionally, languages differentiating transitive from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} have \obj{ɨ}-final prefixes, see \akuriyo in \exref{sayintr.aku-154}, as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for cognate forms in other such languages.
+Finally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
 
@@
 \label{sec:go}
-This verb is reconstructed by \textcite{gildea2007greenberg} as \rc{tə[mə]}, like \rc{ka[ti]} \qu{to say} with a fleeting second syllable.
-It is true that many reflexes are clearly \obj{t}-initial, for example \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}.
-However, once one considers all forms of the various reflexes of this verb \pcref{tab:go}, an initial vowel \rc{ɨ} must clearly be reconstructed -- in contrast to unambiguously C-initial \rc{ka[ti]} \qu{to say}.%
-\footnote{As indicated by the brackets in \cref{tab:go}, there are many languages where the initial vowel is only present in some forms.
-Also, the prefix-verb boundary in many inflected forms like e.g.\ \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} is ambiguous, since an epenthetic \obj{ɨ} breaks up potential CC clusters.
-Still, when one considers unambiguous forms, the contrast with \rc{ka[ti]} becomes very clear.}
 This verb was not affected by the any of the extensions in \cref{sec:extensions}.
+\textcite{gildea2007greenberg} reconstruct it as \rc{tə[mə]}, like \rc{ka[ti]} \qu{to say} with a fleeting second syllable.
+While many reflexes are clearly \obj{t}-initial \parentext{e.g.
+\hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}}, an initial vowel \rc{ɨ} must clearly be reconstructed \pcref{tab:go}%
+\footnote{The brackets in \cref{tab:go} show that languages have the initial vowel only in some forms.
+Also, the prefix-stem boundary in many inflected forms like \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} is ambiguous, since epenthetic \obj{ɨ} breaks up CC clusters on these boundaries.
+Still, when considering unambiguous forms, the contrast with C-initial \rc{ka[ti]} becomes very clear.}
 
@@
 \label{sec:be}
-For a comprehensive comparative overview of these two roots, readers are referred to \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
-\rc{a[p]} is the original copula and can be reconstructed as already having various irregularities in \PC.
+\rc{a[p]} is the older copula and already had various irregularities in \PC \parencite{gildea2018reconstructing}.
 \rc{eti} is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
-Various modern languages use reflexes of these two roots in a suppletive manner, conditioned by person and\slash{}or \gl{tam}.
+Modern reflexes of these roots are used suppletively, conditioned by person and\slash{}or \gl{tam}.
 Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
-\akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while \obj{eʔi} is not attested in a first-person form. 
+\akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while the first person form of \obj{eʔi} is not attested. 
 \carijo innovated \obj{j-}, but only in the reflex of \rc{eti} \exref{mivida-12}; the \obj{a} root preserves \obj{w-} \pcref{sec:carijo}.
@@
 \xe
+%
+A comprehensive comparative overview of these two roots is given by \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
 
@@
 \label{sec:come}
+This verb can be reconstructed as \PC \rc{(ət-)jəpɨ}, morphologically segmentable into an optional detransitivizing prefix and a root \rc{jəpɨ}.
+As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}, but not on the \arara reflex of \rc{jəpɨ}.
+Reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ}/\rc{jəpɨ} (\akuriyo) resisted the introduction of \PTir \rc{t-}.
+\carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
+No \yukpa reflex of this verb is attested, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come}.
+
+\ex<car-18>\carijo \parencite[][102]{guerrero2019carijo}\\
+\begingl
+\gla əji-wa-e j-eh-ɨ//
+\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
+\glft \qu{I came looking for you.}//
+\endgl
+\xe
 
@@
 
-This verb is reconstructed as \rc{ətepɨ} by \textcite[30]{gildea2007greenberg}, but an inspection of all attested reflexes \pcref{tab:come} points to a more complex story.
-Crucially, the majority do not reflect the \rc{ət} part of their reconstruction, and many forms are ostensibly reflexes of \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ}, rather than \rc{epɨ}.
-A unifying account of all these forms is achieved by reconstructing a \PC form \rc{(ət-)jəpɨ}, morphologically segmentable into a detransitivizing prefix and a root \rc{jəpɨ}.
+This reconstruction disagrees with \posscite[30]{gildea2007greenberg} \rc{ətepɨ}, due to the comparative evidence in \cref{tab:come}.
+Crucially, most reflexes do not reflect the \rc{ət} part of \posscite{gildea2007greenberg} reconstruction, and many ostensibly go back to \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ}, rather than \rc{epɨ}.
 
@@
 It only occurs in the third person form of the Progressive, meaning that the \obj{j} may be a reflex of third person \rc{i-}.
-As \cref{tab:kaxprog} shows, regular \obj{o}-initial verbs have no third person prefix, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}.
-Even if \qu{to come} once shared this \obj{i-}, the circumstances are strongly in favor of a \obj{j}-initial \kaxui root, allowing the reconstruction of \PC \rc{jəpɨ}.
+However, regular \obj{o}-initial \kaxui verbs have no third person prefix \pcref{tab:kaxprog}, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}.
+Even if \qu{to come} once shared this \obj{i-}, the \obj{j} is best analyzed as part of the root in \kaxui, allowing the reconstruction of \PC \rc{jəpɨ}.
 %
@@
 %
-Most of the (non-Pemongan) morphologically complex forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} show no segmental trace of \rc{j}, but the \obj{i} in \akawaio \obj{əsipɨ} is likely a reflex of the sequence \rc{jə}.
-This analysis is supported by the reflex \obj{ipɨ} from bare \rc{jəpɨ} in very closely related \macushi.
+The morphologically complex (non-Pemongan) forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} show no segmental trace of \rc{j}, but the \obj{i} in \akawaio \obj{əsipɨ} likely reflects the sequence \rc{jə}.
+This analysis is supported by bare \rc{jəpɨ} becoming \obj{ipɨ} in closely related \macushi.
 
-As for the many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ}, they are distributed widely in the family, sometimes even co-occurring in the same language.
-A unifying account requires the root \rc{jəpɨ} to undergo two major sound changes: \begin{inlinelist}
+The many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ} are distributed unsystematically in the family, sometimes even co-occurring in the same language.
+A unifying account requires the root \rc{jəpɨ} to undergo two main sound changes: \begin{inlinelist}
  \item \rc{j}-loss
@@
  \end{inlinelist}.
-While both sound changes are found in other contexts throughout the family \parencite{meira2010origin}, they appear to have applied irregularly to this verb, and not always in the same order.
-For example, the \kalina form \obj{opɨ} can only be explained if \rc{j} was lost before the umlaut of \rc{ə} to \rc{e}, which would have been triggered by \rc{j}.
+Both sound changes are found in other contexts throughout the family \parencite{meira2010origin}, but they appear to have applied inconsistently to this verb, and not always in the same order.
+Forms like \kalina \obj{opɨ} can only be explained if \rc{j} was lost before the umlauting of \rc{ə} to \rc{e}, which would have been triggered by \rc{j}.
 On the other hand, forms like \maqui \obj{ehə} must be the result of \phonl{\rc{ə}}{\rc{e}}{\rc{j}}, with subsequent loss of \rc{j}.
@@
 The scenario suggested here involves reflexes of the detransitivizer \rc{ət(e)-} being optionally added to an \gl{s_a_} verb root \rc{jəpɨ}, which further underwent umlaut and loss of \rc{j}, but in no systematic manner, resulting in the chaotic picture in \cref{tab:come}.
-As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}, but not on the \arara reflex of \rc{jəpɨ}.
-Reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ}/\rc{jəpɨ} (\akuriyo) resisted the introduction of \PTir \rc{t-}.
-\carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
-It is unknown whether there is a \yukpa reflex of this verb, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come}.
-
-\ex<car-18>\carijo \parencite[][102]{guerrero2019carijo}\\
-\begingl
-\gla əji-wa-e j-eh-ɨ//
-\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
-\glft \qu{I came looking for you.}//
-\endgl
-\xe
 

### 2021-12-13 documents/floats/comepara.tex
changed reconstruction of 'to come'

\begin{table}[h]
\centering
\caption[\rc{ət-epɨ} \qu{to come} in paradigms]{\rc{ət-epɨ} \qu{to come} in paradigms \parencites[113, 150, 153, 156]{alves2017arara}[294; p.c., Spike Gildea]{triomeira1999}}
\label{tab:comepara}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &          \kaxui &          \arara &          \trio \\
\mymidrule
\gl{1}   &   \obj{w-oohɨ-} &    \obj{w-ebɨ-} &  \obj{w-əepɨ-} \\
\gl{2}   &   \obj{m-oohɨ-} &  \obj{m-odebɨ-} &  \obj{mən-epɨ} \\
\gl{1+2} &  \obj{kɨs-ohɨ-} &  \obj{kud-ebɨ-} &   \obj{ke-epɨ} \\
\gl{3}   &    \obj{n-ehɨ-} &    \obj{t-ebɨ-} &    \obj{n-epɨ} \\
\mybottomrule
\end{tabular}
\end{table}

### 2021-12-13 documents/sections/verbs.tex
changed reconstruction of 'to come'

@@
 
-\subsubsection{\rc{(ət-)jəpɨ} \qu{to come}}
+\subsubsection{\rc{(ət-)epɨ} \qu{to come}}
 \label{sec:come}
-This verb can be reconstructed as \PC \rc{(ət-)jəpɨ}, morphologically segmentable into an optional detransitivizing prefix and a root \rc{jəpɨ}.
-As discussed in \cref{sec:pekodian}, innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ətjəpɨ}, but not on the \arara reflex of \rc{jəpɨ}.
-Reflexes of \rc{ətjəpɨ} (\trio) and of \rc{ətjəpɨ}/\rc{jəpɨ} (\akuriyo) resisted the introduction of \PTir \rc{t-}.
+Innovative \rc{k-} was introduced on the \ikpeng and \bakairi reflexes of \rc{ət-epɨ}, but not on the \arara reflex of \rc{epɨ} \pcref{sec:pekodian}.
+The reflex of \rc{ət-epɨ} resisted the introduction of \PTir \rc{t-} \pcref{sec:taranoan}.
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
@@
 \xe
+%
+\posscite{gildea2007greenberg} reconstruction \rc{ətepɨ} can be segmented into an optional prefix \rc{ət-} and a root \rc{epɨ}, since reflexes can be grouped into those with a reflex of \rc{ət-} and those without \pcref{tab:come}.
+As seen in \cref{sec:pekodian,sec:taranoan,tab:comepara}, this division can exist within a single paradigm.
 
@@
 
-This reconstruction disagrees with \posscite[30]{gildea2007greenberg} \rc{ətepɨ}, due to the comparative evidence in \cref{tab:come}.
-Crucially, most reflexes do not reflect the \rc{ət} part of \posscite{gildea2007greenberg} reconstruction, and many ostensibly go back to \rc{əpɨ}, \rc{jepɨ}, or \rc{jəpɨ}, rather than \rc{epɨ}.
+\input{floats/comepara}
 
-Only the Pemongan languages and \kaxui point to an \rc{j}-initial root, and \kaxui \obj{johɨ} is very rare in contrast to more frequent \obj{o(o)hɨ}.
+Long \rc{ət-epɨ} lost \rc{t} in \trio, \apalai, and \bakairi, but only the first development is due to regular sound changes \parencite[31--32]{meira1998proto}.
+\akuriyo reflects \rc{əe} as \obj{ee}, with \obj{eepɨ} being betrayed as not a reflex of \rc{epɨ} by the otherwise unexpected vowel length \parencite[114--115]{meira1998proto}.
+Other languages coalesced \rc{əe} to \rc{əə}, with \rc{əəpɨ} being reflected in \uxc \parencite[452]{gildea2012classification}, \kalina, \kaxui, \panare, and \yawarana.
+In \kalina \parencite[509--510]{meira2010origin}, \rc{e}-initial transitive verbs detransitivized with \rc{ət-} underwent the same development (also yielding vowel length), while \rc{əte} > \rc{əə} is irregular at least in \kaxui and \panare.
+
+The Pemongan languages and \kaxui point to \rc{jəpɨ} rather than \rc{epɨ}, although \kaxui \obj{johɨ} is very rare in contrast to more frequent \obj{o(o)hɨ}.
 It only occurs in the third person form of the Progressive, meaning that the \obj{j} may be a reflex of third person \rc{i-}.
-However, regular \obj{o}-initial \kaxui verbs have no third person prefix \pcref{tab:kaxprog}, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}.
-Even if \qu{to come} once shared this \obj{i-}, the \obj{j} is best analyzed as part of the root in \kaxui, allowing the reconstruction of \PC \rc{jəpɨ}.
+However, regular \obj{o}-initial \kaxui verbs have no third person prefix \pcref{tab:kaxprog}, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}, suggesting that \obj{j} is indeed part of the root.
+Putative \PC \rc{jəpɨ} would yield \rc{epɨ} by the two widespread sound changes of \rc{ə}-umlaut after \rc{j} and \rc{j}-loss \parencite{meira2010origin}, but one would expect evidence of a sequence \rc{jə} in \yawarana, \mapoyo, or \panare \parencite{gildea2010story}.
+The \kaxui form \obj{ehɨ} would also not be accounted for, as \rc{ə} was not umlauted after \rc{j} in Parukotoan \parencite[501--502]{meira2010origin}.
 %
 \input{floats/kaxprog}%
-%\pex<kaxprog>\kaxui \perscomm{Spike Gildea}
-%\a<kax-82>
-%\begingl
-%\gla johɨ-rɨ//
-%\glb \gl{3}.come-\gl{prog}//
-%\glft \qu{S/he is coming.}//
-%\endgl
-%\a<kax-54>
-%\begingl
-%\gla o-w-ohɨ-rɨ//
-%\glb \gl{2}-\gl{s_a_}-come-\gl{prog}//
-%\glft \qu{You are coming.}//
-%\endgl
-%\a<kax-77>
-%\begingl
-%\gla i-nkɨ-rɨ//
-%\glb \gl{3}-sleep-\gl{prog}//
-%\glft \qu{S/he is sleeping.}//
-%\endgl
-%\a<kax-83>
-%\begingl
-%\gla Ø-osone-rɨ//
-%\glb \gl{3}-dream-\gl{prog}//
-%\glft \qu{S/he is dreaming.}//
-%\endgl
-%\xe
-%
-The morphologically complex (non-Pemongan) forms corresponding to \posscite{gildea2007greenberg} \rc{ətepɨ} show no segmental trace of \rc{j}, but the \obj{i} in \akawaio \obj{əsipɨ} likely reflects the sequence \rc{jə}.
-This analysis is supported by bare \rc{jəpɨ} becoming \obj{ipɨ} in closely related \macushi.
 
-The many forms seemingly reflecting \rc{əpɨ} and \rc{epɨ} rather than \obj{jəpɨ} are distributed unsystematically in the family, sometimes even co-occurring in the same language.
-A unifying account requires the root \rc{jəpɨ} to undergo two main sound changes: \begin{inlinelist}
- \item \rc{j}-loss
- \item \rc{ə}-umlaut after \rc{j}
- \end{inlinelist}.
-Both sound changes are found in other contexts throughout the family \parencite{meira2010origin}, but they appear to have applied inconsistently to this verb, and not always in the same order.
-Forms like \kalina \obj{opɨ} can only be explained if \rc{j} was lost before the umlauting of \rc{ə} to \rc{e}, which would have been triggered by \rc{j}.
-On the other hand, forms like \maqui \obj{ehə} must be the result of \phonl{\rc{ə}}{\rc{e}}{\rc{j}}, with subsequent loss of \rc{j}.
-The \akuriyo form \obj{eepɨ} looks like a reflex thereof as well, but the length is unexpected, and is analyzed by \textcite[]{meira1998proto} as a coalescence of \rc{e} and \rc{əe}.
-
-%Other languages -- across the family -- have forms reflecting the same root without the initial consonant, i.e.\ \rc{əpɨ}: \kalina, \panare, \yawarana, and \uxc.\footnote{\obj{e} is the regular outcome of \rc{ə} in \uxc; \rc{e} became \obj{i} \parencite[176]{meira2005southern}.}
-%Also found across the family are forms reflecting \rc{epɨ}, in \kaxui, \arara, \trio, \carijo, \maqui, and \mapoyo.
-
-While a root \rc{jəpɨ}, the two sound changes, and the optional addition of \rc{ət-} do account for the majority of the forms in \cref{tab:come},\footnote{
-	Apart from aforementioned \akuriyo \obj{eepɨ}, another diachronically irregular form is \apalai \obj{oepɨ}, where one would expect \rc{ət-epɨ} to yield \obj{os-epɨ} \parencite[506]{meira2010origin}.
-	Similarly, while \obj{oepɨ} would be a regular outcome of a hypothetical \rc{ə-jəpɨ}, the \envr{}{C} allomorph of the detransitivizer is \obj{e-} in \apalai.
-	One possibility is that the form is due to borrowing from \trio, which has lost intervocalic \rc{t} to create \obj{əepɨ}.
-	Alternatively, \apalai \obj{oepɨ} could be a fossilized loan from \wayana, which has replaced its reflex of \rc{ətjəpɨ}, but where intervocalic \rc{t} was also regularly lost \parencite[63]{wayanatavares2005}.}
-the distribution within the family is rather chaotic.
-In addition to the seemingly unordered distribution of \rc{əpɨ} and \rc{epɨ}, forms with and without \rc{ət-} can be found within the same language, usually conditioned by different prefixes.
-This was briefly discussed in \cref{sec:pekodian} for \arara (and \PPek) and in \cref{sec:taranoan} for \trio (and \PTar).
-To illustrate, the \trio \setone paradigm shows a reflex of \rc{ətepɨ} (< \rc{ətjəpɨ}) for first, but of \rc{epɨ} (< \rc{jəpɨ}) for the other persons \exref{tricome}.%
-\footnote{While the \gl{1+2} form is a regular outcome of \rc{kɨt-epɨ}, the second person form is mysterious \parencite[115]{meira1998proto}.}
-It should be noted that forms with and without \rc{ət-} in different languages are not triggered by the same person values.
-
-\ex<tricome> \trio \parencite[294]{triomeira1999}\\
-\begin{tabular}[t]{@{}ll@{}}
-	\gl{1} & \obj{w-əepɨ} \\
-	\gl{2} &  \obj{mən-epɨ} \\ 
-	\gl{1+2} &  \obj{ke-epɨ} \\
-	\gl{3} &  \obj{n-epɨ} \\
-\end{tabular}
-\xe
-
-The interpretation of the \rc{ət} part as a detransitivizer is based on its form and its paradigmatically conditioned occurrence in some languages.
-Although the combination of a detransitivizer and an intransitive verb seems semantically illogical, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
+The clear segmentability of \rc{ət-} in combination with its form suggest that it is a detransitivizing prefix.
+Although the combination of a detransitivizer and an intransitive verb makes no sense semantically, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
 For example, the \PC \gl{s_p_} verb \rc{wɨnɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
-Also, \waiwai \qu{go to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-The parallels to \qu{to sleep} end here, since bare \rc{jəpɨ} \qu{to come} apparently already was an \gl{s_a_} verb, as evidenced the class membership of its reflexes in \kaxui, \kalina, \arara, \trio, and \panare \exref{pan-128}.
-
-\ex<pan-128>\panare \parencite[][65]{panarepayne2013}\\
-\begingl
-\gla ju-w-əəpɨ-n ka=m kanoʔ//
-\glb \gl{3}-\gl{s_a_}-come-\gl{nspec} \gl{q}=\gl{2}.\gl{aux} rain//
-\glft \qu{Do you think it is gonna rain?}//
-\endgl
-\xe
-
-Summing up, this verb is highly irregular, both from a synchronic and diachronic perspective.
-The scenario suggested here involves reflexes of the detransitivizer \rc{ət(e)-} being optionally added to an \gl{s_a_} verb root \rc{jəpɨ}, which further underwent umlaut and loss of \rc{j}, but in no systematic manner, resulting in the chaotic picture in \cref{tab:come}.
+\waiwai \qu{to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
+However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently already was an \gl{s_a_} verb (although its reflexes in languages with split-\gl{s} always co-occur with reflexes of \rc{ət-epɨ}).
 

### 2021-12-13 workflow/helper_functions.py

@@ -151,7 +151,7 @@ def delatexify(string):
 def sort_lg(df):
     df.Language_ID = df.Language_ID.astype("category")
     df.Language_ID.cat.set_categories(lg_list, ordered=True, inplace=True)
-    df.sort_values(["Language_ID"], inplace=True)
+    df.sort_values(["Language_ID", "Form"], inplace=True)
 
 
 def extract_sources(df, src_str="Source", keep=False, latexify=True):
@@ -256,6 +256,7 @@ def print_aligned_table(
         raw_sources = extract_sources(df, latexify=False, keep=True)
     df["Segments"] = df.apply(lambda x: segmentify(x["Form"]), axis=1)
     df["Cognateset_ID"] = df["Cognateset_ID"].map(str2numcog)
+    df.index += 1
     df = calculate_alignment(df, fuzzy=fuzzy)
     df = df[fields]
     sort_lg(df)

### 2021-12-13 workflow/process_data.py

@@ -1,9 +1,9 @@
 from helper_functions import *
 
 verb_list = ["say", "go", "be_1", "be_2", "come", "go_down", "bathe_intr"]
+person = ["1", "2", "1+2", "3"]
 
 # tables for the introduction, featuring "regular" and "irregular" verbs from Trio and Hixkaryana
-person = ["1", "2", "1+2", "3"]
 pyd.x = ["Meaning_ID"]
 pyd.y = ["Inflection"]
 pyd.y_sort = person
@@ -281,9 +281,9 @@ table.columns = table.columns.map(lambda x: f"\\qu{{to {x}}}")
 table = table.applymap(objectify)
 save_float(
     print_latex(table, keep_index=True),
-    "kaxprog",
-    "\\kaxui \\obj{johɨ} \\qu{to come} compared with other verbs (Spike Gildea, p.c.)",
-    "\\kaxui \\obj{johɨ} \\qu{to come} compared with other verbs"
+    label="kaxprog",
+    caption="\\kaxui \\qu{to come} and other \gl{s_a_} verbs (Spike Gildea, p.c.)",
+    short="\\kaxui \\qu{to come} and other \gl{s_a_} verbs"
 )
 
 
@@ -418,7 +418,7 @@ save_float(
 print_aligned_table(
     v_df[v_df["Parameter_ID"] == "come"],
     verb="come",
-    caption=r"Reflexes of \rc{(ət-)jəpɨ} \qu{to come}",
+    caption=r"Reflexes of \rc{(ət-)epɨ} \qu{to come}",
     fuzzy=True,
 )
 print_aligned_table(
@@ -430,6 +430,28 @@ print_aligned_table(
     v_df[v_df["Parameter_ID"] == "say"],
     verb="say",
     caption=r"Reflexes of \rc{ka[ti]} \qu{to say}",
+)
+
+#some paradigms for 'to come' 
+pyd.x = ["Language_ID"]
+pyd.y = ["Inflection"]
+pyd.y_sort = person
+pyd.x_sort = lg_list
+pyd.filters = {"Meaning_ID": ["come"], "Inflection": person, "Language_ID": ["ara", "tri", "kax"]}
+table = pyd.compose_paradigm(i_df)
+sources = get_sources(i_df)
+
+table.rename(columns=shorthand_dic, inplace=True)
+table.index = table.index.map(pynt.get_expex_code)
+table.index.name = ""
+table = table.applymap(objectify)
[…]

### 2021-12-13 workflow/read_data.py

@@ -29,6 +29,7 @@ concept_dic = {
     "DETRZ1+bathe_1": "bathe_intr",
     "DETRZ+come": "come",
     "DETRZ1": "bathe_intr",
+    "DETRZ": "come"
 }
 
 i_df["Concept"] = i_df.apply(
@@ -56,7 +57,7 @@ cog_verb_df = cs_df[~(cs_df["Form"].str.contains("-"))]
 cog_trans_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Meaning"]))
 cog_form_dic = dict(zip(cog_verb_df["ID"], cog_verb_df["Form"]))
 cog_form_dic["bathe_intr"] = "e-pɨ"
-cog_form_dic["come"] = "(ət-)jəpɨ"
+cog_form_dic["come"] = "(ət-)epɨ"
 cog_trans_dic["bathe_intr"] = r"bathe"
 cog_trans_dic["be_1"] = r"be-1"
 cog_trans_dic["be_2"] = r"be-2"

### 2021-12-13 documents/sections/extensions.tex
'come' footnote for ptir

@@
 
-Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.
+Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.\footnote{The present reconstruction of \qu{to come} diverges from \posscite[114--115]{meira1998proto}, who reconstructs \PTar \rc{əepɨ} for first, but \rc{eepɨ} for other persons, based on the paradigmatic pattern in \trio and the vowel length in \akuriyo.
+\akuriyo and \carijo would then have levelled that pattern, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
+Here, the length in \akuriyo \obj{eepɨ} is identified as resulting from coalescence of \rc{əe}, and \trio \obj{əepɨ} as reflecting \rc{ətepɨ} \pcref{sec:come}.
+This yields \PTir \rc{(əʔ)epɨ} and \PTar \rc{(ət͡ʃ-)epɨ}.}
 The other copular root \rc{eʔi} can be added as a fifth verb, whose \trio reflex retains first person \obj{w-}.
@@
 
-For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for other persons, based on such a paradigmatic pattern in \trio and the vowel length in \akuriyo.
-Both \akuriyo and \carijo then levelled this original distribution, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
-This scenario is plausible, with the exception that \trio \obj{əepɨ} is a reflex of \rc{ət-jəpɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
+%For \qu{to come}, \textcite[114--115]{meira1998proto} reconstructs \PTar \rc{əepɨ} for first person, and \rc{eepɨ} for other persons, based on such a paradigmatic pattern in \trio and the vowel length in \akuriyo.
+%Both \akuriyo and \carijo then levelled this original distribution, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
+%This scenario is plausible, with the exception that \trio \obj{əepɨ} is a reflex of \rc{ət-jəpɨ} \pcref{sec:come}, meaning that the \PTir form would have been \rc{əʔepɨ} (\PTar \rc{ət͡ʃepɨ}).
 

### 2021-12-13 documents/sections/verbs.tex
'come' footnote for ptir

@@
 \waiwai \qu{to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently already was an \gl{s_a_} verb (although its reflexes in languages with split-\gl{s} always co-occur with reflexes of \rc{ət-epɨ}).
+However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently already was an \gl{s_a_} verb (although its reflexes in languages with split-\gl{s} mostly co-occur with reflexes of \rc{ət-epɨ}).
 

### 2021-12-13 documents/sections/verbs.tex
clarified jEpI reconstruction

@@
 However, regular \obj{o}-initial \kaxui verbs have no third person prefix \pcref{tab:kaxprog}, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}, suggesting that \obj{j} is indeed part of the root.
-Putative \PC \rc{jəpɨ} would yield \rc{epɨ} by the two widespread sound changes of \rc{ə}-umlaut after \rc{j} and \rc{j}-loss \parencite{meira2010origin}, but one would expect evidence of a sequence \rc{jə} in \yawarana, \mapoyo, or \panare \parencite{gildea2010story}.
-The \kaxui form \obj{ehɨ} would also not be accounted for, as \rc{ə} was not umlauted after \rc{j} in Parukotoan \parencite[501--502]{meira2010origin}.
+Putative \PC \rc{jəpɨ} would yield \rc{epɨ} via the two widespread sound changes \phonl{\rc{ə}}{\rc{e}}{\rc{j}} and \phonl{\rc{j}}{∅}{\#} \parencite{meira2010origin}.
+%Still, one would expect evidence of the sequence \rc{jə} in \yawarana, \mapoyo, or \panare, and the \kaxui form \obj{ehɨ} would also not be accounted for
+The argument against this reconstruction is that \phonl{\rc{ə}}{\rc{e}}{\rc{j}} did not happen in \PPar, and only inconsistently in \PPP and \mapoyo-\yawarana \parencites[501--502]{meira2010origin}{gildea2010story}.
+Thus, it does not explain \kaxui \obj{ehɨ}, nor the absence of the sequence \rc{jə} in \panare and \mapoyo-\yawarana.
 %

### 2021-12-13 documents/sections/verbs.tex
finetuning 'go down' and 'bathe'

@@
 Reflexes of this verb were not affected by the extensions of \rc{k-} in \PPek \pcref{sec:pekodian} and \obj{k-} in \akuriyo \pcref{sec:akuriyo}.
-The resistance against the former extension was subsequently overcome in \bakairi; its fate in \ikpeng is unknown.
+Its resistance against the former extension was later broken in \bakairi, while its fate in \ikpeng is unknown.
 When \akuriyo extended \obj{k-}, the verb already had a first person form irregularly inflected with \obj{p-}, inherited from \PTir.
-One might think that it was also affected by the extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
+At first sight, it was also affected by the extensions of \obj{j-} in \carijo \exref{gowhere.car-32} and \yukpa \exref{gowhere.yuk-14}.
 
@@
 %
-However, a broader comparative perspective reveals a much more complicated story \pcref{tab:godown}.%
-\footnote{In \cref{tab:godown}, parenthesized forms indicate uncertainty about cognacy status.
-The reconstruction of \PPek \rc{ɨptə} treats the suffixed elements found in the daughter languages as non-cognate.
+However, a family-wide perspective reveals a more complicated story \pcref{tab:godown}.%
+\footnote{
+The cognacy status of parenthesized forms in \cref{tab:godown} is uncertain.
+The reconstruction of \PPek \rc{ɨptə} treats the additional elements in daughter languages as non-cognate.
 \textcite{meira2005southern} identify no correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}, and at least the addition of a final \obj{ŋ} in \PXin is attested elsewhere:
@@
 		\end{inlinelist} \parencites[8]{gildea2007greenberg}[56, 144, 57]{alves2017arara}[25, 270]{ikpengpacheco2001}.}
-It turns out that while a verb\rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about its class.
+While a verb \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about its class.
 Its reflexes in languages that preserve the split-\gl{s} system are distributed fairly evenly between \gl{s_a_} and \gl{s_p_}.
@@
 
-The verb shows traits of both classes in \wayana, necessitating an analysis as a \dbqu{mixed} verb in a synchronic description of that language.
+The verb shows traits of both classes in \wayana, making it a \dbqu{mixed} verb in a synchronic analysis.
 It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
@@
 Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the \PC  causativizer \rc{-ka} was restricted to \gl{s_p_} verbs \parencite{gildea2019overview}.
-These patterns point to a scenario where the verb was a regular member of the \gl{s_p_} class in pre-\wayana, but partially switched to the \gl{s_a_} class, taking a \gl{1+2}\gl{s_a_} prefix and the \gl{s_a_} class marker.
-This in turn implies that \gl{s_a_} reflexes of this verb in other languages fully switched from \gl{s_p_} in their inflectional patterns.
+These patterns point to \qu{to go down} as a regular \gl{s_p_} verb in pre-\wayana, but partially switching to the \gl{s_a_} class by taking a \gl{1+2}\gl{s_a_} prefix and the \gl{s_a_} class marker.
+This in turn implies that (inflectionally defined) \gl{s_a_} reflexes in other languages fully switched from \gl{s_p_} to \gl{s_a_}.
 
-\wayana-external comparative evidence supports this hypothesis:
+\wayana-external evidence supports this hypothesis:
 The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:godown} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
-Both causativized forms contain a reflex of the transitivizer \rc{en-}, which was usually found with \gl{s_p_} verbs \parencite{gildea2019overview}.
-\trio \obj{ɨhtə} has irregular causativized forms that also feature a reflex of \rc{en-} \parencite[263]{triomeira1999}.
-Thus, it appears that this verb was originally \gl{s_p_}, but then switched its class in four \dbqu{and a half} languages of the family, for so far unknown reasons.
+These forms have a reflex of the transitivizer \rc{en-}, occurring with \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
+\trio \obj{ɨhtə} has irregular causativized forms, also with a reflex of \rc{en-} \parencite[263]{triomeira1999}.
+In conclusion, \rc{ɨpɨtə} \qu{to go down} was an \gl{s_p_} verb in \PC, but for unknown reasons switched classes in four \dbqu{and a half} languages of the family.
 
-These circumstances make it impossible to answer the question of whether \qu{to go down} was affected by some extensions.
-For \PTir, one cannot establish a relative chronology of the verb class switch, the introduction of the idiosyncratic marker \rc{p-}, and the extension of \rc{t-}.
-For \PWai, both its first person form and its inflectional class are unknown.
-For \carijo and \yukpa, one cannot rule out the possibility that the verb switched classes before the breakdown of the split-\gl{s} system.
-While no language-internal evidence supports that scenario, \qu{to go down} clearly has an inclination to switch classes; in the case of \carijo, that could have already happened at the \PTar stage.
+This makes it impossible to tell whether it was affected by most extensions under discussion:
+For \PTir, one cannot establish a relative chronology of the class switch, the introduction of idiosyncratic \gl{1}\gl{s_a_} \rc{p-}, and the extension of \rc{t-}.
+Its first person form and its inflectional class in \PWai are unknown.
+For \carijo and \yukpa, one cannot rule out a verb class switch before the breakdown of the split-\gl{s} system.
+While no language-internal evidence supports such a switch, \qu{to go down} is clearly inclined to do so; \carijo may even have inherited it as \gl{s_a_} from \PTar.
 In all four cases, the verb could have had \gl{s_a_} status at the time of the extension, resisting it and preserving the old prefix, but it could also have had \gl{s_p_} status and thus not even have been a potential target.
 
-On the other hand, the class switch happened before the split of \PTir, and therefore this verb resisted the extension of \akuriyo \obj{k-} as an \gl{s_a_} verb.
-Likewise, it is very likely that the class switch took place before the extension of \PPek \rc{k-}.
+On the other hand, a class switch is reconstructible to \PTir, so it was an \gl{s_a_} verb when \akuriyo introduced \obj{k-}.
+Likewise, the class switch most likely took place before the extension of \PPek \rc{k-}.
 Otherwise, the newly-turned-\gl{s_a_} verb would have taken on conservative and lexically heavily restricted \rc{w-}, either in \PPek, \PXin, or \arara.
@@
 
-Verbs for intransitive \qu{to bathe} are usually typical \gl{s_a_} verbs in Cariban languages, derived with a detransitivizer from a transitive root.
+Verbs for intransitive \qu{to bathe} are typical \gl{s_a_} verbs in most Cariban languages, derived with a detransitivizer from a transitive root.
 These roots are reflexes of \rc{[ɨ]pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
-\PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp} (\gl{tr}).
-Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, \qu{to bathe (\gl{intr})} diverges by changing \rc{e-} to \rc{i}.
+\PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp} (\gl{tr}) \pcref{sec:pekodian}.
+Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, \qu{to bathe (\gl{intr})} changed \rc{e-} to \rc{i}.
 This is an irregular development, since there are no attested reflexes of a Pekodian detransitivizer \rc{i-} \parencite[506]{meira2010origin}; its cause is unknown.
-However, other languages also show unexpected developments in this verb, like the glide insertions in Waiwaian or the distribution of \rc{[ɨ]pɨ} and \rc{kupi} in Venezuelan languages.
+%However, other languages also show unexpected developments in this verb, like the glide insertions in Waiwaian or the distribution of \rc{[ɨ]pɨ} and \rc{kupi} in Venezuelan languages.
 

### 2021-12-13 documents/sections/verbs.tex
section 3 summary

@@
 \cref{tab:overview} gives an overview of what verbs were affected by which extensions (except for \obj{e}-initial \akuriyo verbs unaffected by the extension of \obj{k-}, as they are a large and predictable group).
-In some cases, the verb does not occur at all or just not in first person forms (–), in others the first person form is unknown (?), and in the case of \qu{to go down} the question of affectedness is not meaningfully answerable (\textsc{n/a}), for any of the reasons discussed in \cref{sec:godown}.
-Every \checkmark stands for a verb affected by an extension, while × represents conservatively inflected verbs.
-This overview makes clear just how strongly these verbs tend to resist person marker extensions, as they do so in different languages.
+In some cases, the verb does not occur, or at least not in a first person form (–), in others that form is unknown (?), and the question of affectedness is often not meaningfully answerable (\textsc{n/a}) for \qu{to go down} \pcref{sec:godown}.
+Every \checkmark stands for a verb affected by an extension, while × represents conservatively inflected verbs, making clear how strongly these verbs tend to resist person marker extensions in different languages.
+\cref{sec:motivations} will explore possible explanations for the same 1-7 verbs retaining their old \gl{1}\gl{s_a_} marker in 6 independent developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs take on innovative markers.
 
 \input{floats/overview.tex}
-
-It is astonishing that the same 1-7 verbs retain their old first person marker in 6 distinct developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs take on innovative markers.
-This suggests that there is something about these verbs which causes them to behave in such a way.
-Possible answers to the question of what makes them different from regular \gl{s_a_} verbs will be discussed in \cref{sec:motivations}, using \posscite{bybee1985morphology} network model of morphology.

### 2021-12-13 documents/sections/reasons.tex
started trimming section 4

@@
 \label{sec:motivations}
-Perhaps the most well-known contribution regarding conservativeness, innovativeness, and (ir-){}re\-gu\-la\-ri\-ty in the lexicon is \posscite{bybee1985morphology} network model of morphology, which seems well-suited for the data at hand.
-It aims \dbqu{to account for cross-linguistic, diachronic and acquisition patterns in complex morphological systems} \parencite[428]{bybee1995regular}.
+A well-known contribution treating conservativeness, innovativeness, and (ir-){}re\-gu\-la\-ri\-ty in the lexicon is \posscite{bybee1985morphology} network model of morphology, which seems well-suited for the data at hand.
+It aims to \dbqu{account for cross-linguistic, diachronic and acquisition patterns in complex morphological systems} \parencite[428]{bybee1995regular}.
 It does so by modeling shared morphological properties such as inflectional patterns as emerging from connections of differing strength between lexemes.
-A classic example is a network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
-This network is attracting new verbs in certain dialects, like \obj{sniːk}--\obj{snʌk} or \obj{brɪŋ}--\obj{brʌŋ} \parencite[129--130]{bybee1985morphology}.
-These verbs are recruited based on the lexical connection they form with prototypical members of the group, and accordingly develop irregular or \dbqu{strong} past tense forms.
+A prominent example is the network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
+This network created new strong verbs in some dialects, like \obj{sniːk}--\obj{snʌk} or \obj{brɪŋ}--\obj{brʌŋ} \parencite[129--130]{bybee1985morphology}.
+These verbs formed a strong lexical connection with prototypical members of the group, developing new past tense forms as a result.
 
-As possible bases of these connections between lexemes, \textcite[118]{bybee1985morphology} suggests the criteria of semantic, phonological, and morphological similarity; the English strong verbs are an example for a phonologically motivated network.
-Another important factor in the model is frequency, since more frequent words have a higher lexical strength \parencite[119]{bybee1985morphology}.
+English strong verbs are an example for a phonologically motivated network; other possible bases of connections between lexemes are semantic or morphological similarity \parencite[118]{bybee1985morphology}.
+Another important factor in the model is frequency, since more frequent words have higher lexical strength \parencite[119]{bybee1985morphology}.
 This diminishes the influence from other lexemes, meaning that high-frequency items are more likely to resist innovations.
-For the Cariban first person patterns, the model would predict that a) semantically\slash{}phonologically\slash\hspace{0pt}morphologically similar verbs will be affected by person marker extensions, and b) high-frequency verbs will tend to resist these extensions and thus remain conservative.
+For the Cariban first person patterns, the model predicts that semantically\slash{}phonologically\slash\hspace{0pt}morphologically similar verbs are affected by person marker extensions, while high-frequency verbs resist extensions and thus remain conservative.
 
-When considering the groups of verbs with innovative first person markers (those not in \cref{tab:overview}), there are several factors which could serve as the thread connecting a lexical network.
-The most obvious similarity is that they all have a reflex of the detransitivizer \detrz \parentext{see e.g.
+When considering the \gl{s_a_} verbs with innovative first person markers (those not in \cref{tab:overview}), several salient factors emerge as potential bases of lexical networks.
+The most obvious similarity is that they all have a reflex of the detransitivizer \detrz{} \parentext{see e.g.
 \textcite[112]{meira1998proto} for Taranoan}, a hallmark of \gl{s_a_} verbs \pcref{sec:split}.
@@
 %\end{quotebox} % ‘to go down’, ‘to defecate’,
-Since all derived \gl{s_a_} verbs begin with reflexes of \rc{ə} or \rc{e} \parentext{see e.g. \textcite[153]{alves2017arara} for \arara}, phonologically based networks are a second possibility.
+Due to \detrz{} being prefixes, all derived \gl{s_a_} verbs begin with reflexes of \rc{ə} or \rc{e} \parentext{see e.g.
+\textcite[153]{alves2017arara} for \arara{}}, making phonologically based networks a second possibility.
 %A more trivial connection between these other verbs is that they are all \gl{s_a_} verbs, and thus share inflectional (morphological) patterns.
@@
 %The exception to this is \akuriyo, where innovative \gl{1}\gl{s_a_} \obj{k-} and idiosyncratic \obj{p-} on \qu{to go down} created inflectional subclasses of \gl{s_a_} verbs.
-A third common trait is the \gl{s_a_} (sub-)class membership, represented by pre-innovation first person \rc{w-} (\obj{t-} in \akuriyo).
-There are no obvious semantic patterns, which is not surprising given the absence of a semantic basis in the split-\gl{s} system \pcref{sec:split}.
-Thus, for each extension, there are three hypotheses as to what connected the members of the responsible network: a reflex of \gl{detrz}, their stem-initial phoneme, or a specific \gl{1}\gl{s_a_} prefix.
+An inflectional commonality of innovative verbs is their \gl{s_a_} (sub-)class membership, represented by pre-extension \gl{1}\gl{s_a_} \rc{w-} (\obj{t-} in \akuriyo).
+There are no obvious semantic patterns, which is unsurprising considering the absence of a semantic basis in the split-\gl{s} system \pcref{sec:split}.
+Thus, there are three hypotheses for factors connecting members of the networks spreading the extensions: a reflex of \gl{detrz}, their stem-initial phoneme, or a their \gl{1}\gl{s_a_} prefix.
 
-It is intuitively obvious that many of the conservative verbs in \cref{tab:overview} are high-frequency verbs, which would cause conservativeness according to the network model.
+It is intuitively obvious that many of the conservative verbs in \cref{tab:overview} are highly frequent, which would cause conservativeness according to the network model.
 Going beyond intuition is difficult due to the lack of frequency counts for individual lexemes for any Cariban language.
 The only statement in the literature is \posscite[75]{courtz2008carib} claim of \kalina underived \gl{s_a_} verbs being the most frequent ones: \dbqu{It is difficult […] to imagine an intransitive or transitive origin for some of the most frequent middle verbs}.
-This claim is supported neither by frequency counts nor accompanied by a list of verbs, but that list is likely identical with the five first columns of \cref{tab:overview}, all underived \gl{s_a_} verbs in \kalina.
-To improve the situation, a count of \gl{s_a_} verbs in three glossed \apalai texts from \textcite{koehns1994textos} will serve as a second source of frequency information, the results of which are shown in \cref{tab:apalaicounts}.
-The count data agree with the above interpretation of \posscite{courtz2008carib} claim: defining \dbqu{high frequency} as having an above average count yields the exact same five verbs.
-While it is uncertain that the interpretation of \posscite{courtz2008carib} claim and the small \apalai sample are truly representative of discourse patterns in the Cariban (proto-)languages under discussion, the absence of alternatives necessitates their use as a tool for modeling frequency.
+This claim is neither supported by frequency counts nor accompanied by a list of verbs, although that list is likely synonymous with the five first columns of \cref{tab:overview}.
+To improve the situation, a count of \gl{s_a_} verbs in three glossed \apalai texts from \textcite{koehns1994textos} will serve as a second source of frequency information \pcref{tab:apalaicounts}.
+The count data support the above interpretation of \posscite{courtz2008carib} claim, since defining \dbqu{high frequency} as above average yields the exact same five verbs.
+One cannot be sure that the interpretation of \posscite{courtz2008carib} claim and this small \apalai sample are truly representative of discourse patterns in the Cariban (proto-)languages under discussion, but the absence of alternatives necessitates their use for frequency information.
 

### 2021-12-16 documents/floats/go.tex
added brackets and initial vowels to some proto-forms
changed reconstruction of 'to come'
a thousand small fixes

@@
-\begin{table}[h]
+\begin{table}
 \centering
-\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[80, 153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6]{franchetto2002kuikuro}[99; p.c., Spike Gildea]{camargo2002lexico}}
+\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[80, 153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6]{franchetto2002kuikuro}[99]{camargo2002lexico}[14; p.c., Spike Gildea]{lacerda1991waimiri}}
 \label{tab:go}
@@
 \waiwai   &   \obj{[e]to[m]} &  e &  t &  o &  m &    \\
-\PPek     &        \rc{ɨtən} &  ɨ &  t &  ə &  n &    \\
+\PPek     &      \rc{ɨtə[n]} &  ɨ &  t &  ə &  n &    \\
+\arara    &         \obj{to} &    &  t &  o &    &    \\
 \arara    &        \obj{ɨdo} &  ɨ &  d &  o &    &    \\
-\arara    &         \obj{to} &    &  t &  o &    &    \\
 \ikpeng   &       \obj{aran} &  a &  r &  a &  n &    \\
@@
 \bakairi  &      \obj{[ɨ]tə} &  ɨ &  t &  ə &    &    \\
-\PTir     &        \rc{təmɨ} &    &  t &  ə &  m &  ɨ \\
+\PTir     &   \rc{[ə]tə[mɨ]} &  ə &  t &  ə &  m &  ɨ \\
 \trio     &      \obj{tə[n]} &    &  t &  ə &  n &    \\
 \akuriyo  &  \obj{[ə]tə[mɨ]} &  ə &  t &  ə &  m &  ɨ \\
-\carijo   &       \obj{təmə} &    &  t &  ə &  m &  ə \\
+\carijo   &     \obj{tə[mə]} &    &  t &  ə &  m &  ə \\
 \wayana   &   \obj{[ɨ]tə[m]} &  ɨ &  t &  ə &  m &    \\
 \apalai   &        \obj{ɨto} &  ɨ &  t &  o &    &    \\
+\kalina   &       \obj{[ɨ]ʔ} &  ɨ &  ʔ &    &    &    \\
 \kalina   &         \obj{to} &    &  t &  o &    &    \\
-\kalina   &       \obj{[ɨ]ʔ} &  ɨ &  ʔ &    &    &    \\
 \maqui    &    \obj{ɨtə[mə]} &  ɨ &  t &  ə &  m &  ə \\
@@
 \yukpa    &         \obj{to} &    &  t &  o &    &    \\
+\waimiri  &         \obj{ɨʔ} &  ɨ &  ʔ &    &    &    \\
 \mybottomrule

### 2021-12-16 documents/floats/pwaireg.tex
a thousand small fixes

@@
-\begin{table}[h]
+\begin{table}
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[30]{waiwaihawkins1998}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}
@@
 \mytoprule
-{} &           \PWai &                   &           \hixka &                 &            \waiwai &                    \\
-{} &    \qu{to fall} &     \qu{to sleep} &     \qu{to fall} &   \qu{to sleep} &       \qu{to fall} &      \qu{to sleep} \\
+{} &     \qu{to fall} &                  &                    &      \qu{to sleep} &                 &                    \\
+{} &            \PWai &           \hixka &            \waiwai &              \PWai &          \hixka &            \waiwai \\
 \mymidrule
-\gl{1}   &  \rc{k-eɸurka-} &   \rc{kɨ-wɨnɨkɨ-} &  \obj{k-ehurka-} &  \obj{kɨ-nɨkɨ-} &    \obj{k-eɸɨrka-} &   \obj{kɨ-wɨnɨkɨ-} \\
-\gl{2}   &  \rc{m-eɸurka-} &    \rc{o-wɨnɨkɨ-} &  \obj{m-ehurka-} &  \obj{o-wnɨkɨ-} &    \obj{m-eɸɨrka-} &   \obj{mɨ-wɨnɨkɨ-} \\
-\gl{1+2} &  \rc{t-eɸurka-} &  \rc{tɨt-wɨnɨkɨ-} &  \obj{t-ehurka-} &  \obj{tɨ-nɨkɨ-} &  \obj{t͡ʃ-eɸɨrka-} &  \obj{tɨt-wɨnɨkɨ-} \\
-\gl{3}   &  \rc{ɲ-eɸurka-} &   \rc{nɨ-wɨnɨkɨ-} &  \obj{ɲ-ehurka-} &  \obj{nɨ-nɨkɨ-} &    \obj{ɲ-eɸɨrka-} &   \obj{nɨ-wɨnɨkɨ-} \\
+\gl{1}   &  \obj{k-eɸurka-} &  \obj{k-ehurka-} &    \obj{k-eɸɨrka-} &   \obj{kɨ-wɨnɨkɨ-} &  \obj{kɨ-nɨkɨ-} &   \obj{kɨ-wɨnɨkɨ-} \\
+\gl{2}   &  \obj{m-eɸurka-} &  \obj{m-ehurka-} &    \obj{m-eɸɨrka-} &    \obj{o-wɨnɨkɨ-} &  \obj{o-wnɨkɨ-} &   \obj{mɨ-wɨnɨkɨ-} \\
+\gl{1+2} &  \obj{t-eɸurka-} &  \obj{t-ehurka-} &  \obj{t͡ʃ-eɸɨrka-} &  \obj{tɨt-wɨnɨkɨ-} &  \obj{tɨ-nɨkɨ-} &  \obj{tɨt-wɨnɨkɨ-} \\
+\gl{3}   &  \obj{ɲ-eɸurka-} &  \obj{ɲ-ehurka-} &    \obj{ɲ-eɸɨrka-} &   \obj{nɨ-wɨnɨkɨ-} &  \obj{nɨ-nɨkɨ-} &   \obj{nɨ-wɨnɨkɨ-} \\
 \mybottomrule

### 2021-12-16 documents/sections/extensions.tex
a thousand small fixes

@@
-\section{Inflections and verbs: innovation and resistance}
+\section{Prefixes and verbs: innovation and resistance}
 \label{sec:data}
-As shown in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers from incomplete person marker extensions.
+As shown in \cref{sec:extensions_intro}, irregularly inflected first person forms are leftovers of incomplete person marker extensions.
 \cref{sec:extensions} presents the six identified incomplete extensions, the prefixes they introduced and the verbs they spared.
-Since the latter show considerable etymological overlap across languages, these resistant verbs are compared and reconstructed in \cref{sec:verbs}.
-Where existent, their reflexes which did get affected by one of the extensions are identified.
+Since languages show considerable etymological overlap in their conservative verbs, they are compared and reconstructed in \cref{sec:verbs}.
+Further, their reflexes which did get affected by an incomplete extension are identified.
 %The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
@@
 \label{sec:extensions}
-As stated in \cref{sec:extensions_intro}, the six person marker extensions which did not affect all potential targets have in common that they introduced innovative first person markers on \gl{s_a_} verbs.
+As stated in \cref{sec:extensions_intro}, the six person marker extensions which left a group of verbs untouched all introduced innovative first person markers on \gl{s_a_} verbs.
 Of these extensions, half can be reconstructed to intermediate proto-languages, and half happened in pre-modern stages of single languages.
@@
 For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and verb forms are reconstructed where necessary.
-\cref{sec:pekodian} details the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
-\cref{sec:waiwaian} treats the extension of \rc{k-} in \PWai, which was briefly shown in \cref{sec:extensions_intro}.
+\cref{sec:pekodian} demonstrates the extension of \rc{k-} in \PPek, reflected in the three daughter languages \arara, \ikpeng, and \bakairi.
+\cref{sec:waiwaian} details the extension of \rc{k-} in \PWai, briefly shown in \cref{sec:extensions_intro}.
 \cref{sec:taranoan} focuses on innovative \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
-The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative \gl{1}\gl{s_a_} markers only found in single languages:
+The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative \gl{1}\gl{s_a_} markers found in single languages:
 \obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.
@@
 The most detailed description of a Pekodian language \parencite{alves2017arara} names six%
-\footnote{Seven under her analysis, which sees the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.}
+\footnote{Seven in her analysis, treating the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.}
 \arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
-There is also a reflex of the copula \rc{a[p]}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
-However, its inflectional morphology features verbal \setone prefixes, including first person \obj{w-} \exref{araap}.
+There is also a reflex of the copula \rc{a[p]} \cref{sec:be}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
+However, it inflects with verbal \setone prefixes, including first person \obj{w-} \exref{araap}.
 %We take these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.
@@
 
-In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two subclasses of \gl{s_a_} verbs, one taking first person \obj{w-}, and one \obj{k-}.\footnote{\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
+In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two \gl{s_a_} subclasses, defined by \gl{1}\gl{s_a_}\obj{w-} and \obj{k-}, respectively.\footnote{\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
 	However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
@@
 While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of the \arara verbs in \exref{arairr} as \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.
-However, inflected forms can be found in \textcite{von1892bakairi}, presented in \exref{bakverbs} according to the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
+However, inflected forms can be found in \textcite{von1892bakairi}, represented in \exref{bakverbs} based on the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
 
@@
 There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
-While reflexes of \rc{a[p]} \qu{to be} do exist in \ikpeng, apparently only reflexes of \rc{eti} \qu{to be} occur with first person inflectional prefixes \parencite[401]{gildea2018reconstructing}.
+While reflexes of \rc{a[p]} \qu{be-1} do exist in \ikpeng, apparently only reflexes of \rc{eti} \qu{be-2} occur with first person prefixes \parencite[401]{gildea2018reconstructing}.
 
@@
 
-Reconstructions of verb stems are deferred to \cref{sec:verbs}, but a brief comment on \qu{to come} is in order:
-The stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ}.
-In contrast, the \arara first person form is directly based on this root \rc{epɨ}.
-However, reflexes of \rc{əd-epɨ} can be found elsewhere in the \arara paradigm \exref{ara-123}.
+Reconstructions of verb stems are deferred to \cref{sec:verbs}, but a comment on \qu{to come} is in order:
+The stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ} \pcref{sec:come}.
+In contrast, the \arara first person form has a bare reflex of \rc{epɨ}.
+While reflexes of \rc{əd-epɨ} can be found elsewhere in the \arara paradigm \exref{ara-123}, \ikpeng and \bakairi uniformly reflect \rc{əd-ebɨ}.
 
@@
 %
-In contrast, \ikpeng and \bakairi show reflexes of \rc{əd-ebɨ} throughout the whole paradigm.
-Following \posscite[114]{meira1998proto} line of reasoning for a similar pattern in Taranoan (see also \cref{sec:taranoan}), the idiosyncratic pattern in \arara can be reconstructed to \PPek, meaning that \bakairi and \ikpeng independently levelled the paradigm in favor of \rc{əd-epɨ}.
+Following \posscite[114]{meira1998proto} line of reasoning for a similar pattern in Taranoan (see also \cref{sec:taranoan}), the idiosyncratic pattern in \arara can be reconstructed to \PPek, with \bakairi and \ikpeng independently leveling the paradigm in favor of \rc{əd-epɨ}.
 
@@
 This extension, one of the Parukotoan innovations shown in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
-\PWai further extended the \gl{1}\gl{s_p_} prefix \rc{k-} innovated in \PPar to \gl{1}\gl{s_a_}.
+\PWai further extended the \gl{1}\gl{s_p_} prefix \rc{k-} (innovated in \PPar) to \gl{1}\gl{s_a_}.
 For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
@@
 The two forms for \qu{to be} are unproblematic, whereas \qu{to go} is a special case.
-While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with the old \rc{w-}, an etymological analysis also considered by \textcite[90]{gildea1998}.
-Alternatively, this form may have been influenced by deverbalized forms of \qu{to go}, where a reflex of the \gl{s_a_} class marker \rc{w-} has become fossilized \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}
+While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with  old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
+Alternatively, this form may have been influenced by deverbalized forms of \qu{to go}, which show a fossilized of the \gl{s_a_} class marker \rc{w-}  \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}.
 %
@@
 %
-In any case, the first person form \hixka \qu{to go} clearly points to \PWai \rc{wɨ-tom-}.
+Either way, the first person form \hixka \qu{to go} clearly points to \PWai \rc{wɨ-tom-}.
 
@@
 \label{sec:taranoan}
-The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages identified by \textcite{girard1971proto}, with \carijo as a more distant member.
-\textcite{meira1998proto} contributes an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
-This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
+The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages identified by \textcite{girard1971proto}, with \carijo as a sister.
+\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places, creating a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
 \footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
 \input{floats/ptirreg}
-The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario in which both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
+The question of how this switch happened in detail \parencite[107--112]{meira1998proto} is still open, although a scenario seems necessary in which both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
 In fact, even the issue of \emph{when} this happened is open.
@@
 
-Turning to verbs not affected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.\footnote{The present reconstruction of \qu{to come} diverges from \posscite[114--115]{meira1998proto}, who reconstructs \PTar \rc{əepɨ} for first, but \rc{eepɨ} for other persons, based on the paradigmatic pattern in \trio and the vowel length in \akuriyo.
+As for verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs four of the items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar, for which reconstructed \PTir forms are substituted here.\footnote{The present reconstruction of \qu{to come} diverges from \posscite[114--115]{meira1998proto}, who reconstructs \PTar \rc{əepɨ} for first, but \rc{eepɨ} for other persons, based on the paradigmatic pattern in \trio and the vowel length in \akuriyo.
 \akuriyo and \carijo would then have levelled that pattern, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
@@
 This yields \PTir \rc{(əʔ)epɨ} and \PTar \rc{(ət͡ʃ-)epɨ}.}
-The other copular root \rc{eʔi} can be added as a fifth verb, whose \trio reflex retains first person \obj{w-}.
+As a fifth verb \rc{eʔi} \qu{be-1} can be added, whose \trio reflex retains \obj{w-}.
 The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is identified as a reflex of \rc{wɨ-} by 
-\textcite[113]{meira1998proto}, which is supported the fact that both components of the irregular change \rc{wɨ-} > \obj{ə-} (loss of \rc{w} and lowering of \rc{ɨ} to \obj{ə}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
+\textcite[113]{meira1998proto}, which is supported the fact that both components of the idiosyncratic change \rc{wɨ-} > \obj{ə-} (\phon{\rc{w}}{∅} and \phon{\rc{ɨ}}{\obj{ə}}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
 
@@
 After the split of \PTir, when \rc{t-} and \rc{t͡ʃ-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{s_a_} marker \obj{k-}.
-It seems to have replaced \rc{t-} only in specific environments, with the two markers showing a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
+It seems to have replaced \rc{t-} only in specific environments, with \obj{k-} and \obj{t͡ʃ-} showing a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
 \textcite[107]{meira1998proto} largely confirms that distribution, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
-He also suggests that \obj{k-} might be more recent, which is plausible: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
+He also suggests that \obj{k-} may be more recent, which is plausible: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
 The few \obj{t-} mentioned by \textcite{meira1998proto} were then perhaps reintroduced under \trio influence.
-However, since there are no examples of, or more information about, \obj{ə}-initial verbs with \obj{t-}, these cases cannot be discussed further.
+However, since there are no examples of -- or more information about -- \obj{ə}-initial verbs with \obj{t-}, these cases cannot be discussed further.
 
@@
 \label{sec:carijo}
-\carijo, the cousin of the Tiriyoan languages, has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
+\carijo, the third Taranoan language, extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
 Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regular verbs \pcref{tab:carreg}.
@@
 %
-Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight} \exref{car-30}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}.
+Although the split-\gl{s} system was lost entirely, former \gl{s_a_} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight} \exref{car-30}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}.
 %The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
@@
 
-As noted in \cref{sec:taranoan}, this extension also erased any traces of a putative \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
+As noted in \cref{sec:taranoan}, this extension also erased any traces of putative \PTar \gl{1}\gl{s_a_} \rc{t-}.
 However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
@@
 %We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
-Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would expect \obj{ka} \qu{to say} to either have conservative \obj{wɨ-} or innovative \obj{ji-}, so the zero needs explanation.
+Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, \obj{ka} \qu{to say} should either have conservative \obj{wɨ-} or innovative \obj{ji-}, so the zero is unexpected.
 It is analysed here as a reflex of \rc{wɨ-}, primarily due to the loss of \rc{w} in \ikpeng and \hixka.
@@
 \label{sec:yukpa}
-The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
+The divergent nature of \yukpa\footnote{Very little is known about its sister \japreria, whose status as a dialect of \yukpa is contested by \textcite{oquendo2004japreria}.} is \textit{inter alia}t visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
 However, it does preserve the \setone prefixes in the immediate past, which shows a unified intransitive paradigm \pcref{tab:yukreg}.
@@
 
-In intransitive verbs, this first-person zero marking is attested in a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
+In intransitive verbs, this first-person zero marking is attested with a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
 It diverges from regular C-initial verbs with \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.

### 2021-12-16 documents/sections/introduction.tex
a thousand small fixes

@@
 The Cariban language family is one of the largest of South America, with between 60'000 and 100'000 speakers unevenly distributed between 22 to 25 extant languages \parencite[441]{gildea2012classification}.
-The family is concentrated in Venezuela, the Guianas and Northern Brazil, with two Western and four Southern outliers.
+The family is concentrated in Venezuela, the Guianas and Northern Brazil, with three Western and four Southern outliers.
 \cref{fig:phylomap} shows the geographical distribution and genealogical affiliation of the extant Cariban languages.
@@
 %Many also have a split-\gl{s} system, which can be reconstructed to \PC \pcref{sec:split}.
-For linguistic overviews of and  comparative work on the family, readers are referred to \textcites{gildea1998}{derbyshire1999carib}{meira2002first}{meira2005southern}{meira2006cariban}{gildea2007greenberg}{meira2010origin}{gildea2010story}{gildea2012classification}{matter2021cariban}{gildea2019overview}.
+For overviews of and  comparative work on the family, readers are referred to \textcites{gildea1998}{derbyshire1999carib}{meira2002first}{meira2005southern}{meira2006cariban}{gildea2007greenberg}{meira2010origin}{gildea2010story}{gildea2012classification}{matter2021cariban}{gildea2019overview}.
 
@@
 This is illustrated with person paradigms of four \hixka verbs in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
-This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in the paradigms shown here.
-In \cref{tab:hixintro} and other paradigm tables, any TAM suffixes found in the original forms found in the literature are omitted, since a) the focus lies on the prefixes and stems, and b) full paradigms containing the same TAM suffix are rarely found.
-Further, standard IPA symbols are used in the transcription of Cariban languages, with the exception of coronal rhotics, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
-In languages with strong morphophonological processes and/or subphonemic orthography the original transcription is shown in an additional surface line when presented in interlinearized examples.
-\textcite{gildea2018reconstructing} is followed in using \ort{ə} for the proto-vowel reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.
+This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in \cref{tab:hixintro} and other paradigm tables.
+Further, TAM suffixes in the attested forms are omitted, since a) the focus lies on the prefixes and stems, and b) full paradigms with the same TAM suffix are rare in the available sources.
+Standard IPA symbols are used in the transcription of Cariban languages, except for coronal rhotics, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+In languages with strong morphophonological processes and/or non-phonemic orthography the original form is shown in an additional line when presented in interlinearized examples.
+\textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.
 Glossing abbreviations: } %\glossingAbbrevsComma.
@@
 Such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}, but have only been subject to comparative scrutiny in \posscite{meira1998proto} reconstruction of \PTar.
-In a synchronic analysis of a language, these verbs and their first person prefixes may be called \textsc{irregular}, contrasting with regular prefixes (like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}) on regular verbs.
+In synchronic analyses, these verbs and their first person prefixes may be called \textsc{irregular}, contrasting with regular prefixes (like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}) on regular verbs.
 However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[e.g.,][]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
-For such approaches, these verbs simply belong to a small inflectional (sub-)class, an analysis applied to the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
+In such approaches, the verbs simply belong to a small inflectional (sub-)class, an analysis applied to the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
 
-Ignoring the specifics of synchronic analysis, the cause for the divergent inflectional patterns lies in the diachrony of the languages in question.
-The goal of this study is to approach the patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows:
-In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns like the \hixka and \trio ones.
+Ignoring the details of synchronic analysis, the cause for the divergent inflectional patterns is found in the diachrony of the languages in question.
+The goal of this study is to approach these patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows:
+In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns as in \cref{tab:hixintro,tab:triintro} .
 In \cref{sec:data}, six incomplete person marker extensions and the verbs unaffected by them are described.
-Since conservative verbs show a considerable etymological overlap between languages, they are further discussed and reconstructed.
+Since languages show considerable etymological overlap in their conservative verbs, they are further discussed and reconstructed.
 \cref{sec:motivations} uses \posscite{bybee1985morphology} network model of morphology to search explanations for the verbs (un-)affected by each extension.

### 2021-12-16 documents/sections/proto_carib.tex
a thousand small fixes

@@
 That system underwent much innovation; the mechanism responsible for the irregular forms are person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
-A specific aspect of the system, the \gl{s_a_} vs \gl{s_p_} distinction, plays a role in incomeplete extensions and is discussed in \cref{sec:split}.
+A specific aspect of the system, the \gl{s_a_} vs \gl{s_p_} distinction, plays a role in incomplete extensions and is discussed in \cref{sec:split}.
 %\cref{sec:outlook} summarizes the main points of this section and provides an outlook for the rest of the paper.
@@
 There was a single aliophoric marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
-Local scenarios were expressed in a non-transparent manner, both using the \gl{1+2} prefix \rc{k-}.
+Local scenarios were marked identically and non-transparently with the \gl{1+2} prefix \rc{k-}.
 
@@
 Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
-That is, \gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}\gl{s_a_} \rc{w-} vs \gl{1}>\gl{3} \rc{t(i)-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
-On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones, and \gl{3}\gl{s_p_} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
+\gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}\gl{s_a_} \rc{w-} vs \gl{1}>\gl{3} \rc{t(i)-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
+On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones, with \gl{3}\gl{s_p_} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
 
@@
 It was once more widespread as the normal plural suffix of the weak inflection, compare German \obj{ɔks-ən} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
-
-Since the regular \hixka and \trio prefixes are innovative, one may ask where they came from.
+Since the regular \hixka and \trio prefixes are innovative, the question arises how they developed.
 
@@
 In his discussion of the \PC split-\gl{s} system \pcref{sec:split} and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different changes in daughter languages.
-The main mechanism of these changes are \textsc{person marker extensions}, i.e. person prefixes being extended to verbal paradigm cells previously occupied by other prefixes.
+The main mechanism of these changes are \textsc{person marker extensions}, person prefixes being extended to verbal paradigm cells previously occupied by other prefixes.
 %\footnote{
@@
 %For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
-There have been quite a few person marker extensions in Cariban languages, some still ongoing.
+There have been many person marker extensions in Cariban languages, some still ongoing.
 \textcite{gildea1998} illustrates this with the three Parukotoan languages \kaxui, \hixka, and \waiwai.
@@
 \ex
+    \begin{threeparttable}
     	\begin{tabular}[t]{ll}
     		\PPar & \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}\\
-& \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)\\
-& \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)\\
+    		& \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_}\tnote{a}\\
+    		& \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_}\tnote{a}\\
     		\PWai & \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}\\
@@
     	\end{tabular}
+    \begin{tablenotes}
+    	\footnotesize
+    	\item[a] Completed in \PWai, ongoing in \kaxui.
+    \end{tablenotes}
+    	\end{threeparttable}
 \xe
@@
 
-For \textcite{gildea1998}, person marker extensions are relevant for loss of split-\gl{s} and the accompanying changes to indexing alignment, whereas this study focuses on a different aspect of these extensions.
+For \textcite{gildea1998}, person marker extensions are relevant for the loss of split-\gl{s} and the accompanying changes to indexing alignment, whereas this study focuses on a different aspect.
 Namely, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}, a hypothesis supported by three facts.
@@
 
-The third argument in favor of the lexical diffusion scenario goes back to the conservative \hixka and \trio forms in \cref{tab:hixintro,tab:triintro}.
+The third argument in favor of the lexical diffusion scenario goes back to conservative forms like in \hixka and \trio.
 Both innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
@@
 \footnote{As an honorable mention, when \ikpeng replaced third person \setone with \settwo prefixes, \obj{a} \qu{to be} and \obj{ke} \qu{to say} retained \obj{n-} \parencite[12]{matter2019arara}.
-	However, the spread of the innovative markers had an entierly different dynamic, spreading from subordinate to main clauses.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
+	However, the spread of the innovative markers had an entirely different dynamic, spreading from subordinate to main clauses.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
 Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, and the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
@@
 The starkly different behavior of \gl{s_a_} and \gl{s_p_} verbs regarding extensions points to the split-\gl{s} system playing a role, so its main properties will be discussed in \cref{sec:split}.
-It will also be made clear how the \gl{s_a_}/\gl{s_p_} distinction can be lost for a single person, or how \gl{s_p_} verbs can take on \gl{s_a_} markers with apparent semantic impunity.
+This will also clarify how the \gl{s_a_}/\gl{s_p_} distinction can be lost for a single person, or how \gl{s_p_} verbs can take on \gl{s_a_} markers with apparent semantic impunity.
 
@@
 \label{sec:split}
-As seen in \cref{sec:pc_person}, the split between \gl{s_a_} and \gl{s_p_} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only inflectional criterion:
+As seen in \cref{sec:pc_person}, the split between \gl{s_a_} and \gl{s_p_} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only difference:
 In deverbalized forms, \gl{s_a_} verbs took a class marker \rc{w-}, while \gl{s_p_} verbs lacked that prefix \parencites[89, 141--142]{gildea1998}[208]{meira2000split}.
-The distinction between \gl{s_a_} and \gl{s_p_} was also reflected in imperatives, where the latter took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} while the former were unprefixed \parencites[89]{gildea1998}[208]{meira2000split}.
+Also, \gl{s_p_} verbs took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} in imperatives, while \gl{s_a_} verbs were unprefixed \parencites[89]{gildea1998}[208]{meira2000split}.
 
@@
 
-Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although their reflexes on verbs have been merged into a single morpheme in modern languages.
+Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although their reflexes on verbs have been merged into a single morpheme in modern languages.%
+\footnote{In the \PC transitive verb template, derivational prefixes were in a paradigmatic relationship with the third person marker \rc{i-}:
+\rc{m-i-}V \qu{you V it}, \rc{m-e-}V \qu{you V yourself}, \rc{m-əte-}V \qu{you V each other}.
+This analysis is applied to \trio by \textcite[268--269]{triocarlin2004}, who interprets \obj{i-} as marking transitive diathesis.}
 Reflexes of \detrz show a range of meanings summarizable as \dbqu{detransitive}, illustrated with \trio \gl{s_a_} verbs in \exref{triodetrz}.
@@
 %
-The morphological variation in \qu{to abandon each other} and \qu{to wash self} is due to the collapse of the two \PC prefixes:
+The morphological variation in \qu{abandon each other} and \qu{wash self} is due to the collapse of the two \PC prefixes:
 \obj{e-} comes from the reflexive prefix \rc{e-}, while the form \obj{əi-} originates in reciprocal \rc{əte-}.
 However, both can occur with either meaning -- at least for these two verbs.
-In the next section, it will become apparent that many of the verbs not affected by person marker extensions belong to the small group of \gl{s_a_} verbs without a reflex of \detrz.
+In the next section, it will be seen that many of the verbs not affected by person marker extensions belong to the small group of \gl{s_a_} verbs without a reflex of \detrz.
 %As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.

### 2021-12-16 documents/sections/reasons.tex
a thousand small fixes

@@
 \label{sec:motivations}
-A well-known contribution treating conservativeness, innovativeness, and (ir-){}re\-gu\-la\-ri\-ty in the lexicon is \posscite{bybee1985morphology} network model of morphology, which seems well-suited for the data at hand.
+A well-known approach to conservativeness, innovativeness, and (ir-){}re\-gu\-la\-ri\-ty in the lexicon is \posscite{bybee1985morphology} network model of morphology, which seems well-suited for the data at hand.
 It aims to \dbqu{account for cross-linguistic, diachronic and acquisition patterns in complex morphological systems} \parencite[428]{bybee1995regular}.
 It does so by modeling shared morphological properties such as inflectional patterns as emerging from connections of differing strength between lexemes.
-A prominent example is the network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
+A prominent example is a network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
 This network created new strong verbs in some dialects, like \obj{sniːk}--\obj{snʌk} or \obj{brɪŋ}--\obj{brʌŋ} \parencite[129--130]{bybee1985morphology}.
-These verbs formed a strong lexical connection with prototypical members of the group, developing new past tense forms as a result.
+These verbs formed a lexical connection with prototypical members of the group, developing new past tense forms as a result.
 
-English strong verbs are an example for a phonologically motivated network; other possible bases of connections between lexemes are semantic or morphological similarity \parencite[118]{bybee1985morphology}.
+These verbs are an example for a phonologically motivated network; other possible bases of connections between lexemes are semantic or morphological similarity \parencite[118]{bybee1985morphology}.
 Another important factor in the model is frequency, since more frequent words have higher lexical strength \parencite[119]{bybee1985morphology}.
@@
 There are no obvious semantic patterns, which is unsurprising considering the absence of a semantic basis in the split-\gl{s} system \pcref{sec:split}.
-Thus, there are three hypotheses for factors connecting members of the networks spreading the extensions: a reflex of \gl{detrz}, their stem-initial phoneme, or a their \gl{1}\gl{s_a_} prefix.
+Thus, there are three hypotheses for factors connecting members of the networks in which extensions spread: a reflex of \detrz, the stem-initial phoneme, or a specific \gl{1}\gl{s_a_} prefix.
 
@@
 The only statement in the literature is \posscite[75]{courtz2008carib} claim of \kalina underived \gl{s_a_} verbs being the most frequent ones: \dbqu{It is difficult […] to imagine an intransitive or transitive origin for some of the most frequent middle verbs}.
-This claim is neither supported by frequency counts nor accompanied by a list of verbs, although that list is likely synonymous with the five first columns of \cref{tab:overview}.
-To improve the situation, a count of \gl{s_a_} verbs in three glossed \apalai texts from \textcite{koehns1994textos} will serve as a second source of frequency information \pcref{tab:apalaicounts}.
+This claim is neither supported by frequency counts nor accompanied by a list of verbs, although those verbs likely correspond to the five first columns of \cref{tab:overview}.
+To improve the data situation, a count of \gl{s_a_} verbs in three \apalai texts from \textcite{koehns1994textos} will serve as a second source of frequency information \pcref{tab:apalaicounts}.
 The count data support the above interpretation of \posscite{courtz2008carib} claim, since defining \dbqu{high frequency} as above average yields the exact same five verbs.
-One cannot be sure that the interpretation of \posscite{courtz2008carib} claim and this small \apalai sample are truly representative of discourse patterns in the Cariban (proto-)languages under discussion, but the absence of alternatives necessitates their use for frequency information.
+It is questionable if the interpretation of \posscite{courtz2008carib} claim and this small \apalai sample are representative of discourse patterns in the Cariban (proto-)languages discussed here, but their use is made necessary by the absence of alternatives.
 
@@
 
-Thus, each of the three hypotheses for possible network factors can be combined with frequency: the members of the lexical network formed by the factor are predicted to undergo innovation, but high-frequency verbs are exempt.
-The resulting six hypotheses for possible explanations were tested by predicting the expected behavior of each verb in each extension, illustrated in \cref{tab:ptir-predictions} for \PTir.
-For example, \rc{eʔi} \qu{to be} is expected to participate (\checkmark) in innovations spreading in a phonologically defined network (being \rc{e}-initial), as well as in an inflectionally defined one (sharing \rc{w-} with other \gl{s_a_} verbs).
-On the other hand, a network based on the detransitivizer would predict it to not take on new suffixes (×).
+Each of the three hypotheses for possible network factors can be combined with frequency: members of the lexical network connected by a factor are prone to innovation, but high-frequency verbs are exempt.
+The resulting six hypotheses were tested by predicting the expected behavior of each verb in each extension, illustrated in \cref{tab:ptir-predictions} for \PTir.
+For example, \rc{eʔi} \qu{to be} is expected to participate (\checkmark) in innovations spreading in a phonologically defined (\rc{e}-initial) or inflectionally defined (\gl{1}\gl{s_a_} \rc{w-}) network.
+On the other hand, a network based on the detransitivizer predicts it to not undergo innovation (×).
 If frequency is taken into account, it is expected to remain conservative regardless of the basis of the network.
@@
 
-These predictions were checked against the data in \cref{tab:overview}, counting verbs which had their behavior (in-)correctly predicted.
+These predictions were checked against the data in \cref{tab:overview} by tallying verbs with (in-)correctly predicted behavior.
 The resulting scores are illustrated for the extension of \PTir \rc{t-} in \cref{tab:ptir-evaluations} and summed up for all extensions in \cref{tab:resultsoverview}.
-Notably, the scores in \cref{tab:resultsoverview} only refer to the seven verbs which are attested as resisting at least one extension.
-For each extension, there were also many run-of-the-mill \gl{s_a_} verbs which were all affected, except for the \akuriyo \obj{e}-initial verbs.%
-\footnote{While there are a few \gl{s_a_} verbs not transparently derived from transitive verbs \parencites[252]{triomeira1999}[222]{meira2000split}[30]{gildea2007greenberg}, which are not featured in \cref{tab:overview}, these are mostly \rc{ə}-initial and were likely productively derived at some point.
+These scores crucially only consider the seven verbs in \cref{tab:overview}.
+Each extension affected many run-of-the-mill \gl{s_a_} verbs,%
+\footnote{
+While some non-detransitivized \gl{s_a_} verbs \parencites[252]{triomeira1999}[222]{meira2000split}[30]{gildea2007greenberg} are not in \cref{tab:overview}, these are mostly \rc{ə}-initial and were likely productively derived at some point, see \textcite[252]{triomeira1999} for \trio.
 The verbs to which this does not apply, like \trio \obj{wa} \qu{to dance} \parencites[252]{triomeira1999}, are all instances of \gl{s_p_} verbs switching classes.
-	Since none of them is attested as being an \gl{s_a_} verb at the time of a person marker extension, they are not relevant for this study.}
-To illustrate, if one adds 1'000 simulated derived \gl{s_a_} verbs per language to the data -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six explanations consistently predict the behavior of 99.99+\% verbs correctly.
-However, the present investigation is restricted to the edge cases, since the available data simply does not allow such large-scale tests for Cariban languages.
+Since none of them is attested as having \gl{s_a_} status at the time of a person marker extension, they are not relevant for this study.}
+except for the \akuriyo \obj{e}-initial verbs.
+If one simulates the addition of 1'000 regularly inflected detransitivized \gl{s_a_} verbs per language -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six explanations consistently predict the behavior of 99.99+\% verbs correctly.
+Since the descriptive coverage of Cariban languages makes using real such large-scale data impossible, the present investigation is restricted to the edge cases.
 
@@
 In both cases, only the underived%
-\footnote{Note that for \PPek, the idiosyncratic evolution of \rc{e-pɨ} \qu{to bathe (\gl{intr})} to \rc{ipɨ}  made the verb morphologically opaque.}
+\footnote{The assumption is made that the idiosyncratic evolution of \rc{e-pɨ} \qu{to bathe (\gl{intr})} to \rc{ipɨ} \PPek made the verb morphologically opaque.}
 \gl{s_a_} verbs were not affected, all other \gl{s_a_} verbs taking \rc{k-}.
@@
 In \carijo, the extension of \obj{j-} affected all \obj{e}- and \obj{ə}-initial verbs, including \obj{eh} \qu{to come} or \obj{et͡ʃi} \qu{to be}, which do not have a detransitivizing prefix.
-Only \obj{ka} \qu{say}, \obj{təmə} \qu{go}, and \obj{a} \qu{be-1} did not take on \obj{j-}.
-Similarly, the extension of \yukpa \obj{j-} can succinctly be characterized as affecting all vowel-initial verbs; the only attested conservative verb is C-initial \obj{to} \qu{to go}.
+Only \obj{ka} \qu{say}, \obj{tə[mə]} \qu{go}, and \obj{a} \qu{be-1} did not take on \obj{j-}.
+The extension of \yukpa \obj{j-} can succinctly be characterized as affecting all vowel-initial verbs; the only attested conservative verb is C-initial \obj{to} \qu{to go}.
 
 Inflectional morphology as a network basis only played a potential role in the case of \akuriyo.
-However, this explanation only works if \obj{t-} and \obj{t͡ʃ-} are analyzed as distinct morphemes.
-Since they can also be seen phonologically conditioned allomorphs, the prediction is identical to the phonological one.
+However, this explanation rests on the analysis of \obj{t-} and \obj{t͡ʃ-} as distinct morphemes.
+Due to their phonologically conditioned distribution in \PTir, the putative inflectionally defined network is identical to the phonological one.
 
-When additionally considering the assumed conservative effects of frequency, prediction accuracy was improved in 8 cases, stagnated in 7 cases, and worsened in three cases.
-These three cases where the tentative model of verb frequency arrives at incorrect predictions are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
-Overall, including frequency in the model led to improvements, up to 100\% prediction accuracy for all three potential factors in \PTir, as well as for the phonological criterion in \PPek and the inflection criterion in \PWai.
+When adding the expected conservative effects of frequency, prediction accuracy was improved in 8 cases, stagnated in 7 cases, and worsened in three cases.
+These three cases where the crude model of verb frequency arrives at incorrect predictions are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
+Overall, including frequency in the model led to improvements, up to 100\% prediction accuracy for all three factors in \PTir, phonology and inflection in \PWai, and phonology in \PPek.
 
@@
 For the other extensions, the network model gives no unambiguous answer to the question of what (combination of) factors caused innovative first person markers to spread the way they did.
-This in turn is due to the fact that three of the factors in the model -- detransitivizer, phonology, frequency -- strongly converge in their predictions:
+This is due three of the factors in the model -- frequency, detransitivizer, phonology -- strongly converging in their predictions.
 The most frequent \gl{s_a_} verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.

### 2021-12-16 documents/sections/verbs.tex
a thousand small fixes

@@
 Here, these verbs are investigated from a comparative perspective:
-\rc{ka[ti]} \qu{to say} \pcref{sec:say}, \rc{ɨtə[mə]} \qu{to go} \pcref{sec:go}, both roots of the copula \rc{eti}/\obj{a[p]} \pcref{sec:be}, \rc{(ət)jəpɨ} \qu{to come} \pcref{sec:come},  \rc{ɨpɨtə} \qu{to go down} \pcref{sec:godown}, and \rc{e-pɨ} \qu{to bathe} \pcref{sec:bathe}.
+\rc{ka[ti]} \qu{to say} \pcref{sec:say}, \rc{ɨtə[mə]} \qu{to go} \pcref{sec:go}, both roots of the copula \rc{eti} and \rc{a[p]} \pcref{sec:be}, \rc{(ət)jəpɨ} \qu{to come} \pcref{sec:come},  \rc{ɨpɨtə} \qu{to go down} \pcref{sec:godown}, and \rc{e-pɨ} \qu{to bathe} \pcref{sec:bathe}.
 The large and phonologically coherent group of \obj{e}-initial verbs not affected by the extension of \akuriyo \obj{k-} \pcref{sec:akuriyo} will not be discussed.
@@
 \label{sec:say}
-This verb was not affected by any of the extensions in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}), while the first person form of its \yukpa reflex \obj{ka} is unattested.
-Most reflexes are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, best visible in the imperative forms of some languages. % \exref{kati}.
+This verb was not affected by any of the extensions in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}), while no first person form of its \yukpa reflex \obj{ka} is attested.
+Most reflexes are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, visible in the imperative forms of some languages. % \exref{kati}.
 %
@@
 \cref{tab:say} shows a comparison of the longest attested forms for each language.%
-\footnote{Cognate segments in \crefrange{tab:say}{tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for improved exposition of correspondences.}
+\footnote{Cognate segments in \crefrange{tab:say}{tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for improved exposition of correspondences.
+Brackets indicate segments not present in all forms.}
 %
@@
 
-As mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes the \hixka reflex as transitive, an analytical choice not only motivated by a desire to avoid an idiosyncratic intransitive first person prefix \obj{ɨ-} instead of regular \obj{kɨ-}.
-\hixka \obj{ka} also shows the complementary distribution of third person  \obj{n-} and preceding objects typical of transitive verbs in Cariban \parencite[60--81]{gildea1998}.
+As mentioned in \cref{sec:waiwaian}, \textcite{hixkaryanaderby1985} analyzes \hixka \obj{ka[s]} as transitive, a choice not only motivated by a desire to avoid an idiosyncratic intransitive first person prefix \obj{ɨ-} instead of regular \obj{kɨ-}.
+The verb also shows the complementary distribution of third person \obj{n-} and preceding objects typical of transitive verbs in Cariban \parencite[60--81]{gildea1998}.
 Due to its semantics, these objects are either ideophones or direct speech \exref{hixsay}.
@@
 In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}), while does not occur in \exref{hixsay.ordpl-35} where the verb is preceded by direct speech.
-Looking beyond \hixka, at least the \trio reflex shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
+Outside of \hixka, at least the \trio reflex shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
 
-\rc{ka[ti]} \qu{to say} also shows transitive patterns in its derivational suffixes:
-In \trio, it is the only intransitive verb taking transitive \obj{-po} (\gl{caus}) and the \obj{-ne} (\gl{agt}.\gl{nmlz}) \parencite[263, 169]{triomeira1999}.
+Reflexes of \rc{ka[ti]} \qu{to say} also show transitive patterns in their derivational suffixes:
+In \trio, it is the only intransitive verb (inconsistently) taking transitive \obj{-po} (\gl{caus}) and \obj{-ne} (\gl{agt}.\gl{nmlz}) \parencite[263, 169]{triomeira1999}.
 It also exceptionally takes the former suffix in \kalina \parencite[82]{courtz2008carib} and \wayana \parencite[258]{wayanatavares2005}.
 The agent nominalizer \rc{-ne} became the \panare inflectional suffix \obj{-ɲe} on transitive verbs \parencite[184--185]{gildea1998}.
-The combination of \obj{ka} and \obj{-ɲe} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, although \textcite[102]{mattei1994diccionario} arrived at the opposite conclusion.
-Finally, reflexes of the \parentext{\parencite{gildea2015valency}} causativizer \rc{-metipo} are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
+The combination of \obj{ka} and \obj{-ɲe} likely led \textcite[214]{panarepayne2013} to categorize it as transitive, disagreeing with  \textcite[102]{mattei1994diccionario}.
+Finally, reflexes of the transitive causativizer \rc{-metipo} \parencite{gildea2015valency} are found with \obj{ka} in \apalai \parencite[51]{koehn1986apalai} and \waiwai \parencite[52]{waiwaihawkins1998}.
 
-Arugments in favor of intransitive \qu{to say} primarily come from its inflectional prefixes.
+Arguments in favor of intransitive \qu{to say} primarily come from its inflectional prefixes.
 \kalina has a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.
-Similarly, \PPek \rc{ke} \qu{to say} took \gl{1}\gl{s_a_} \rc{w-} \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
+Similarly, \PPek \rc{ke} \qu{to say} had \gl{1}\gl{s_a_} \rc{w-} \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
 Additionally, languages differentiating transitive from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} have \obj{ɨ}-final prefixes, see \akuriyo in \exref{sayintr.aku-154}, as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for cognate forms in other such languages.
@@
 This verb was not affected by the any of the extensions in \cref{sec:extensions}.
-\textcite{gildea2007greenberg} reconstruct it as \rc{tə[mə]}, like \rc{ka[ti]} \qu{to say} with a fleeting second syllable.
-While many reflexes are clearly \obj{t}-initial \parentext{e.g.
-\hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}}, an initial vowel \rc{ɨ} must clearly be reconstructed \pcref{tab:go}%
-\footnote{The brackets in \cref{tab:go} show that languages have the initial vowel only in some forms.
-Also, the prefix-stem boundary in many inflected forms like \trio \obj{wɨtənne} or \arara \obj{wɨdolɨ} \qu{I went} \parencites[43]{triomeira1999}[153]{alves2017arara} is ambiguous, since epenthetic \obj{ɨ} breaks up CC clusters on these boundaries.
-Still, when considering unambiguous forms, the contrast with C-initial \rc{ka[ti]} becomes very clear.}
+\textcite{gildea2007greenberg} reconstruct it as C-initial \rc{tə[mə]}, like \rc{ka[ti]} \qu{to say} with a fleeting second syllable.
+While many reflexes are unambiguously \obj{t}-initial \parentext{e.g.
+\hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}}, the contrast with C-initial \rc{ka[ti]} becomes clear once all forms are considered \pcref{tab:go}.
+\obj{ɨ} is predominant and can tentatively be reconstructed, although languages throughout the family reflect \rc{ə}.
+In many languages \obj{ɨ} only occurs in certain contexts; but even languages with only C-initial roots show ambiguous inflected forms like \trio \obj{wɨtənne} \qu{I went} \parencite[43]{triomeira1999}, since epenthetic \obj{ɨ} breaks up CC clusters on prefix-verb boundaries.
 
@@
 \rc{eti} is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
-Modern reflexes of these roots are used suppletively, conditioned by person and\slash{}or \gl{tam}.
+Reflexes of these roots are used suppletively, conditioned by person and\slash{}or \gl{tam}.
 Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
-\akuriyo \obj{a} was not affected by the extension of \obj{k-} \pcref{sec:akuriyo}, while the first person form of \obj{eʔi} is not attested. 
+\akuriyo \obj{a} was not targeted by the extension of \obj{k-} \pcref{sec:akuriyo}, while no first person form of \obj{eʔi} is attested. 
 \carijo innovated \obj{j-}, but only in the reflex of \rc{eti} \exref{mivida-12}; the \obj{a} root preserves \obj{w-} \pcref{sec:carijo}.
-\yukpa introduced \obj{j-} to the reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries in certain constructions \exref{yukcop}.
+\yukpa introduced \obj{j-} to the reflexes of both \rc{a[p]} and \rc{eti}, which are preserved as encliticized auxiliaries  \exref{yukcop}.
 
@@
 %
-A comprehensive comparative overview of these two roots is given by \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
+A comprehensive comparative overview of both roots is given by \textcite[375--382]{gildea2018reconstructing}; they will not be discussed in detail here.
 
@@
 \posscite{gildea2007greenberg} reconstruction \rc{ətepɨ} can be segmented into an optional prefix \rc{ət-} and a root \rc{epɨ}, since reflexes can be grouped into those with a reflex of \rc{ət-} and those without \pcref{tab:come}.
-As seen in \cref{sec:pekodian,sec:taranoan,tab:comepara}, this division can exist within a single paradigm.
+As seen in \cref{tab:comepara}, this division can exist within a single paradigm.
 
@@
 Long \rc{ət-epɨ} lost \rc{t} in \trio, \apalai, and \bakairi, but only the first development is due to regular sound changes \parencite[31--32]{meira1998proto}.
-\akuriyo reflects \rc{əe} as \obj{ee}, with \obj{eepɨ} being betrayed as not a reflex of \rc{epɨ} by the otherwise unexpected vowel length \parencite[114--115]{meira1998proto}.
+\akuriyo reflects \rc{əte} as \obj{ee}, where \obj{eepɨ} cannot be a reflex of \rc{epɨ} due the otherwise unexpected vowel length \parencite[114--115]{meira1998proto}.
 Other languages coalesced \rc{əe} to \rc{əə}, with \rc{əəpɨ} being reflected in \uxc \parencite[452]{gildea2012classification}, \kalina, \kaxui, \panare, and \yawarana.
-In \kalina \parencite[509--510]{meira2010origin}, \rc{e}-initial transitive verbs detransitivized with \rc{ət-} underwent the same development (also yielding vowel length), while \rc{əte} > \rc{əə} is irregular at least in \kaxui and \panare.
+\rc{e}-initial transitive verbs detransitivized with \rc{ət-} underwent the same development in \kalina, creating minimal pairs for length \parencite[509--510]{meira2010origin}, while \rc{ət-e} > \rc{əə} is irregular at least in \kaxui and \panare.
 
@@
 However, regular \obj{o}-initial \kaxui verbs have no third person prefix \pcref{tab:kaxprog}, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}, suggesting that \obj{j} is indeed part of the root.
-Putative \PC \rc{jəpɨ} would yield \rc{epɨ} via the two widespread sound changes \phonl{\rc{ə}}{\rc{e}}{\rc{j}} and \phonl{\rc{j}}{∅}{\#} \parencite{meira2010origin}.
+Putative \PC \rc{jəpɨ} could yield \rc{epɨ} via the two widespread sound changes \phonl{\rc{ə}}{\rc{e}}{\rc{j}}{} and \phonl{\rc{j}}{∅}{\#} \parencite{meira2010origin}.
 %Still, one would expect evidence of the sequence \rc{jə} in \yawarana, \mapoyo, or \panare, and the \kaxui form \obj{ehɨ} would also not be accounted for
 The argument against this reconstruction is that \phonl{\rc{ə}}{\rc{e}}{\rc{j}} did not happen in \PPar, and only inconsistently in \PPP and \mapoyo-\yawarana \parencites[501--502]{meira2010origin}{gildea2010story}.
-Thus, it does not explain \kaxui \obj{ehɨ}, nor the absence of the sequence \rc{jə} in \panare and \mapoyo-\yawarana.
+Thus, it does not explain \kaxui \obj{ehɨ}, nor the total absence of the sequence \rc{jə} in \panare and \mapoyo-\yawarana.
 %
@@
 \waiwai \qu{to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently already was an \gl{s_a_} verb (although its reflexes in languages with split-\gl{s} mostly co-occur with reflexes of \rc{ət-epɨ}).
+However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently was an \gl{s_a_} verb already (although its reflexes in languages with split-\gl{s} mostly co-occur with reflexes of \rc{ət-epɨ}).
 
@@
 
-The verb shows traits of both classes in \wayana, making it a \dbqu{mixed} verb in a synchronic analysis.
+The verb shows traits of both classes in \wayana, making it a \dbqu{mixed} verb synchronically.
 It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
@@
 Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the \PC  causativizer \rc{-ka} was restricted to \gl{s_p_} verbs \parencite{gildea2019overview}.
-These patterns point to \qu{to go down} as a regular \gl{s_p_} verb in pre-\wayana, but partially switching to the \gl{s_a_} class by taking a \gl{1+2}\gl{s_a_} prefix and the \gl{s_a_} class marker.
-This in turn implies that (inflectionally defined) \gl{s_a_} reflexes in other languages fully switched from \gl{s_p_} to \gl{s_a_}.
+These patterns point to \qu{to go down} being a regular \gl{s_p_} verb in pre-\wayana, but partially switching to the \gl{s_a_} class by taking a \gl{1+2}\gl{s_a_} prefix and the \gl{s_a_} class marker.
+This in turn implies that (inflectionally defined) \gl{s_a_} reflexes in other languages fully switched from \gl{s_p_}.
 
@@
 The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:godown} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
-These forms have a reflex of the transitivizer \rc{en-}, occurring with \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
+These forms reflect the transitivizer \rc{en-}, occurring with \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
 \trio \obj{ɨhtə} has irregular causativized forms, also with a reflex of \rc{en-} \parencite[263]{triomeira1999}.
@@
 While no language-internal evidence supports such a switch, \qu{to go down} is clearly inclined to do so; \carijo may even have inherited it as \gl{s_a_} from \PTar.
-In all four cases, the verb could have had \gl{s_a_} status at the time of the extension, resisting it and preserving the old prefix, but it could also have had \gl{s_p_} status and thus not even have been a potential target.
+In all four cases, the verb could have had \gl{s_a_} or \gl{s_p_} status at the time of the extension, so it is unknown whether it even was a potential target.
 
@@
 \cref{tab:overview} gives an overview of what verbs were affected by which extensions (except for \obj{e}-initial \akuriyo verbs unaffected by the extension of \obj{k-}, as they are a large and predictable group).
-In some cases, the verb does not occur, or at least not in a first person form (–), in others that form is unknown (?), and the question of affectedness is often not meaningfully answerable (\textsc{n/a}) for \qu{to go down} \pcref{sec:godown}.
+In some cases, the verb does not occur, or at least not in a first person form (–), in others that form is unknown (?), and the question of affectedness is often not meaningfully answerable (\textsc{n/a}) for \qu{to go down}.
 Every \checkmark stands for a verb affected by an extension, while × represents conservatively inflected verbs, making clear how strongly these verbs tend to resist person marker extensions in different languages.
-\cref{sec:motivations} will explore possible explanations for the same 1-7 verbs retaining their old \gl{1}\gl{s_a_} marker in 6 independent developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs take on innovative markers.
+\cref{sec:motivations} will explore explanations for the fact that the same 1-7 verbs retained their old \gl{1}\gl{s_a_} marker in 6 independent developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs took on new markers.
 

### 2021-12-16 workflow/helper_functions.py

@@ -56,7 +56,7 @@ def save_float(tabular, label, caption, filename=None, short=None):
         shorttext = ""
     else:
         shorttext = f"[{short}]"
-    output = f"""\\begin{{table}}[h]
+    output = f"""\\begin{{table}}
 \centering
 \caption{shorttext}{{{caption}}}
 \label{{tab:{label}}}

### 2021-12-16 workflow/process_data.py

@@ -94,7 +94,7 @@ reconstructed_form_table(
 # regular proto-waiwaian verbs
 meanings = ["fall", "sleep"]
 lgs = ["PWai", "hix", "wai"]
-pyd.x = ["Language_ID", "Meaning_ID"]
+pyd.x = ["Meaning_ID", "Language_ID"]
 pyd.y = ["Inflection"]
 pyd.filters = {"Language_ID": lgs, "Meaning_ID": meanings}
 pyd.sort_orders = {
@@ -282,8 +282,8 @@ table = table.applymap(objectify)
 save_float(
     print_latex(table, keep_index=True),
     label="kaxprog",
-    caption="\\kaxui \\qu{to come} and other \gl{s_a_} verbs (Spike Gildea, p.c.)",
-    short="\\kaxui \\qu{to come} and other \gl{s_a_} verbs"
+    caption="\\kaxui \gl{s_a_} verbs in the Progressive (Spike Gildea, p.c.)",
+    short="\\kaxui \gl{s_a_} verbs in the Progressive"
 )
 
 
@@ -737,6 +737,27 @@ save_float(
 )
 
 
+affectedness = overview[overview["Language_ID"] == overview["Orig_Language"]]
+affectedness = affectedness[affectedness["Affected"].isin(["y", "n"])]
+affectedness = affectedness[affectedness["Concept"].isin(verb_list)]
+affectedness["Form"] = affectedness["Form"].apply(lambda x: x.split("-")[1])
+
+# # add fake Sa verbs
+# for lg, conds in phon_preds.items():
+#     for i in range(0, 1000):
+#         affectedness = affectedness.append({"Form": conds[0]+"turu", "Language_ID": lg, "Verb_Cognateset_ID": f"DETRZ+talk{i}", "Affected": "y", "Meaning_ID": "talk"}, ignore_index=True)
+
+bool_map = {
+    "latex": {
+        True: "\checkmark",
+        False: "×"
+    },
+    "csv": {
+        True: "correct",
+        False: "incorrect"
+    }
+}
+
 def predict_frequency(lexeme, frequencies):
     if lexeme in frequencies:
         return not frequencies[lexeme]
@@ -761,28 +782,6 @@ def test_explanations(word_form):
         explanations[combo] = predictions[combo] == actual
     return predictions, explanations
 
-
-affectedness = overview[overview["Language_ID"] == overview["Orig_Language"]]
[…]

### 2021-12-16 documents/floats/pwaireg.tex
fixed csv output

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-12-16 workflow/helper_functions.py

@@ -124,8 +124,6 @@ def upper_repl(match):
     return match.group(1).upper()
 
 web_checkmarks = {
-    1: "✓",
-    0: "×",
     True: "✓",
     False: "×",
     "\\checkmark": "✓",
@@ -224,7 +222,7 @@ def reconstructed_form_table(lgs, proto, verbs, caption, name, sort):
     raw_sources = get_sources(i_df, latexify=False)
     tabular.index.name = None
     export_csv(
-        tabular.rename(columns=name_dic),
+        tabular.rename(columns=name_dic).rename(index=plain_trans_dic),
         name,
         f"Irregular {cah.get_name(proto)} verbs",
         True,

### 2021-12-16 workflow/process_data.py

@@ -57,7 +57,7 @@ sources = cldfh.cite_a_bunch(sources_list, parens=True)
 label = "pekreg"
 tabular.index.names = [None]
 export_csv(
-    tabular.rename(columns=name_dic, level="Language_ID"),
+    tabular.rename(columns=name_dic, level="Language_ID").rename(columns=plain_trans_dic, level="Meaning_ID"),
     label,
     "Regular Pekodian Sa verbs",
     keep_index=True,
@@ -159,7 +159,7 @@ tabular = pyd.compose_paradigm(tir_reg, multi_index=True)
 sources = get_sources(tir_reg)
 
 export_csv(
-    tabular.rename(columns=name_dic, level="Language_ID"),
+    tabular.rename(columns=name_dic, level="Language_ID").rename(columns=plain_trans_dic, level="Language_ID"),
     label,
     "Regular Proto-Tiriyoan Sa verbs",
     sources=get_sources(tir_reg, latexify=False),
@@ -244,7 +244,7 @@ for lg, meanings in {
     tabular = pyd.compose_paradigm(i_df, multi_index=False)
     label = lg + "reg"
     export_csv(
-        tabular,
+        tabular.rename(columns=plain_trans_dic),
         label,
         f"Regular {name_dic[lg]} verbs",
         keep_index=True,
@@ -307,7 +307,7 @@ pyd.y_sort = meanings
 tabular = pyd.compose_paradigm(forms)
 label = "pxinw"
 export_csv(
-    tabular,
+    tabular.rename(columns=name_dic),
     label,
     "Loss of \**w* in Ikpeng",
     keep_index=True,
@@ -710,7 +710,7 @@ label = "apalaicounts"
 export_csv(
     apalai_data,
     label,
-    "Frequency counts of Sa verbs in two Apalai texts",
+    "Frequency counts of Sa verbs in three Apalai texts",
     keep_index=False,
     sources="",
 )

### 2021-12-16 workflow/read_data.py

@@ -24,6 +24,7 @@ ext_lg_dic = {x: y for x, y in dict(zip(e_df["ID"], e_df["Language_ID"])).items(
 ext_form_dic = {x: y for x, y in dict(zip(e_df["ID"], e_df["Form"])).items()}
 
 trans_dic = {x: "\\qu{%s}" % y for x, y in mean_dic.items()}
+plain_trans_dic = {x: "'%s'" % y for x, y in mean_dic.items()}
 
 concept_dic = {
     "DETRZ1+bathe_1": "bathe_intr",

### 2021-12-16 workflow/process_data.py
fixed csv output

@@ -852,7 +852,7 @@ res_df = pd.DataFrame(all_explanations)
 res_df *= 100
 
 export_csv(
-    res_df.rename(columns=lambda x: extension_string(x, latex=False), index=repl_latex),
+    res_df.replace(web_checkmarks).rename(columns=lambda x: extension_string(x, latex=False), index=repl_latex),
     label,
     caption=caption,
     keep_index=True,

### 2021-12-16 workflow/process_data.py
overview prettification

@@ -627,7 +627,7 @@ result_exp.columns = [
 ]
 
 export_csv(
-    result_exp, label, "Overview of extensions and (un-)affected verbs", keep_index=True
+    result_exp.replace(web_checkmarks), label, "Overview of extensions and (un-)affected verbs", keep_index=True
 )
 
 # add nice-looking checkmarks and stuff
@@ -852,7 +852,7 @@ res_df = pd.DataFrame(all_explanations)
 res_df *= 100
 
 export_csv(
-    res_df.replace(web_checkmarks).rename(columns=lambda x: extension_string(x, latex=False), index=repl_latex),
+    res_df.rename(columns=lambda x: extension_string(x, latex=False), index=repl_latex),
     label,
     caption=caption,
     keep_index=True,

### 2021-12-16 documents/floats/pwaireg.tex
overview prettification

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189, 190, 196]{hixkaryanaderby1985}[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-12-16 workflow/helper_functions.py

@@ -124,6 +124,8 @@ def upper_repl(match):
     return match.group(1).upper()
 
 web_checkmarks = {
+    "y": "✓",
+    "n": "×",
     True: "✓",
     False: "×",
     "\\checkmark": "✓",

### 2021-12-17 documents/floats/comepara.tex
a thousand small fixes
small fixes

@@
-\begin{table}[h]
+\begin{table}
 \centering
-\caption[\rc{ət-epɨ} \qu{to come} in paradigms]{\rc{ət-epɨ} \qu{to come} in paradigms \parencites[113, 150, 153, 156]{alves2017arara}[294; p.c., Spike Gildea]{triomeira1999}}
+\caption[\rc{(ət-)epɨ} \qu{to come} in paradigms]{\rc{(ət-)epɨ} \qu{to come} in paradigms \parencites[113, 150, 153, 156]{alves2017arara}[294; p.c., Spike Gildea]{triomeira1999}}
 \label{tab:comepara}

### 2021-12-17 documents/floats/participles.tex
regular verbs in data and docs
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
a thousand small fixes
small fixes

--- 2021-09-10 documents/floats/participles.tex
@@
+\begin{table}
+\centering
+\caption[Participles of \gl{s_a_} and \gl{s_p_} verbs]{Participles of \gl{s_a_} and \gl{s_p_} verbs \parencites[39]{schuring2018kaxuyana}[118, 207]{alves2017arara}[333-334]{triomeira1999}[400]{wayanatavares2005}[35]{koehn1986apalai}[kuruaz-154]{koehns1994textos}[430, 433]{hoff1968carib}[232, 244]{panarepayne2013}}
+\label{tab:participles}
 \begin{tabular}[t]{@{}lll@{}}
-\toprule
+\mytoprule
 Language &                               \gl{s_a_} &                        \gl{s_p_} \\
-\midrule
+\mymidrule
 \kaxui  &         \obj{t-ehurka-t͡ʃe} \qu{fallen} &       \obj{tɨ-jaʔ-so} \qu{burnt} \\
@@
 \wayana &     \obj{tə-\emp{w-}epɨ-he} \qu{bathed} &    \obj{t-onopɨ-he} \qu{painted} \\
-\apalai &  \obj{t-\emp{o-}erama-se} \qu{returned} &      \obj{t-anorɨ-se} \qu{dried} \\
+\apalai &        \obj{t-\emp{o-}ɨto-se} \qu{gone} &   \obj{t-ɨhto-se} \qu{gone down} \\
 \kalina &  \obj{tu-\emp{w-}oʔka-se} \qu{come out} &       \obj{t-okari-se} \qu{told} \\
 \panare &  \obj{t-\emp{o-}tatɨhpə-se} \qu{wailed} &  \obj{tɨ-sɨrɨke-t͡ʃe} \qu{tired} \\
-\bottomrule
+\mybottomrule
 \end{tabular}
+\end{table}

### 2021-12-17 documents/floats/ppekverbs.tex
a thousand small fixes
small fixes

@@
-\begin{table}[h]
+\begin{table}
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[76, 130, 131, 374, 397]{von1892bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[76, 130-131, 374, 397]{von1892bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
 \label{tab:ppekverbs}

### 2021-12-17 documents/floats/ptirverbs.tex
new approach to testing predictions from bybee's model
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
more trimming, fixed verb citations
changed reconstruction of 'to come'
a thousand small fixes
small fixes

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[112, 113, 114, 115, 165]{meira1998proto}[292, 294, 339]{triomeira1999}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-115, 165]{meira1998proto}}
 \label{tab:ptirverbs}
@@
 {} &          \PTir &          \trio &                     \akuriyo \\
-\midrule
+\mymidrule
+\qu{say}  &    \rc{wɨ-ka-} &   \obj{wɨ-ka-} &                 \obj{wɨ-ka-} \\
+\qu{go}   &  \rc{w-ɨtəmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
 \qu{be-1} &      \rc{w-a-} &     \obj{w-a-} &     \obj{{\normalfont ∅}-a-} \\
 \qu{be-2} &    \rc{w-eʔi-} &    \obj{w-ei-} &   \obj{{\normalfont ?}-eʔi-} \\
-\qu{say}  &    \rc{wɨ-ka-} &   \obj{wɨ-ka-} &                 \obj{wɨ-ka-} \\
-\qu{go}   &  \rc{wɨ-təmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
 \qu{come} &  \rc{w-əʔepɨ-} &  \obj{w-əepɨ-} &  \obj{{\normalfont ∅}-eepɨ-} \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2021-12-17 documents/floats/pwaireg.tex
small fixes

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[209, 211]{hawkins1953waiwai}[189, 190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189-190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
 \label{tab:pwaireg}

### 2021-12-17 documents/floats/pwaiverbs.tex
new approach to testing predictions from bybee's model
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
more trimming, fixed verb citations
a thousand small fixes
small fixes

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[70, 197, 198]{hixkaryanaderby1985}[4]{hixkaryanaderby1979}[71, 85; p.c., Spike Gildea]{waiwaihawkins1998}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197-198; p.c., Spike Gildea]{hixkaryanaderby1985}}
 \label{tab:pwaiverbs}
@@
 {} &         \PWai &        \hixka &         \waiwai \\
-\midrule
+\mymidrule
+\qu{say}  &   \rc{wɨ-ka-} &   \obj{ɨ-ka-} &    \obj{wɨ-ka-} \\
+\qu{go}   &  \rc{wɨ-tom-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
 \qu{be-1} &    \rc{w-ah-} &   \obj{w-ah-} &      \obj{w-a-} \\
 \qu{be-2} &   \rc{w-eʃi-} &  \obj{w-eʃe-} &   \obj{w-eeʃi-} \\
-\qu{say}  &   \rc{wɨ-ka-} &   \obj{ɨ-ka-} &    \obj{wɨ-ka-} \\
-\qu{go}   &  \rc{wɨ-tom-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2021-12-17 documents/floats/say.tex
changed reconstruction of 'to come'
a thousand small fixes
small fixes

@@
-\begin{table}[h]
+\begin{table}
 \centering
-\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[267]{meira2005bakairi}[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209, 209, 279]{ikpengpacheco2001}[80, 153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63; p.c., Spike Gildea]{largo2011yukpa}}
+\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[267]{meira2005bakairi}[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209, 279]{ikpengpacheco2001}[80, 153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63; p.c., Spike Gildea]{largo2011yukpa}}
 \label{tab:say}
@@
 \ikpeng  &    \obj{k[e]} &  k &  e &    &    \\
+\bakairi &      \obj{ge} &  g &  e &    &    \\
 \bakairi &      \obj{ke} &  k &  e &    &    \\
-\bakairi &      \obj{ge} &  g &  e &    &    \\
 \PTir    &       \rc{ka} &  k &  a &    &    \\

### 2021-12-17 documents/sections/extensions.tex
small fixes

@@
 \cref{sec:extensions} presents the six identified incomplete extensions, the prefixes they introduced and the verbs they spared.
-Since languages show considerable etymological overlap in their conservative verbs, they are compared and reconstructed in \cref{sec:verbs}.
+Since languages show considerable etymological overlap in their conservative verbs, those are compared and reconstructed in \cref{sec:verbs}.
 Further, their reflexes which did get affected by an incomplete extension are identified.
+\cref{sec:verbsummary} summarizes the results.
 %The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
@@
 \arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
-There is also a reflex of the copula \rc{a[p]} \cref{sec:be}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
+There is also a reflex of the copula \rc{a[p]} \pcref{sec:be}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
 However, it inflects with verbal \setone prefixes, including first person \obj{w-} \exref{araap}.
@@
 
-In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two \gl{s_a_} subclasses, defined by \gl{1}\gl{s_a_}\obj{w-} and \obj{k-}, respectively.\footnote{\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
+In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two \gl{s_a_} subclasses, defined by \gl{1}\gl{s_a_} \obj{w-} and \obj{k-}, respectively.\footnote{\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
 	However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
@@
 Not all \gl{s_a_} verbs were affected: \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
-Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, which also occurs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, corresponding to \waiwai \obj{w(ɨ)-} \exref{hixɨ.wai-159}.
+Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, which also occurs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, corresponding to \waiwai \obj{wɨ-} \exref{hixɨ.wai-159}.
 
@@
 While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with  old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
-Alternatively, this form may have been influenced by deverbalized forms of \qu{to go}, which show a fossilized of the \gl{s_a_} class marker \rc{w-}  \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}.
+Alternatively, this form may have been influenced by deverbalized forms of \qu{to go} showing a fossilized of the \gl{s_a_} class marker \rc{w-}  \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}.
 %
@@
 Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
-Here, a conservative stance is taken and the innovation is arbitrarily assumed to be \PTir.
-This decision does not affect the results of this study.}
+The decision taken here to assign the innovation to \PTir does not affect the results of this study.}
 
@@
 \akuriyo and \carijo would then have levelled that pattern, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
-Here, the length in \akuriyo \obj{eepɨ} is identified as resulting from coalescence of \rc{əe}, and \trio \obj{əepɨ} as reflecting \rc{ətepɨ} \pcref{sec:come}.
-This yields \PTir \rc{(əʔ)epɨ} and \PTar \rc{(ət͡ʃ-)epɨ}.}
+Here, both \trio \obj{əepɨ} and \akuriyo \obj{eepɨ} are identified as reflecting \rc{ətepɨ} \pcref{sec:come}, via \PTar \rc{(ət͡ʃ-)epɨ} and \PTir \rc{(əʔ)epɨ}.}
 As a fifth verb \rc{eʔi} \qu{be-1} can be added, whose \trio reflex retains \obj{w-}.
@@
 \label{sec:akuriyo}
-After the split of \PTir, when \rc{t-} and \rc{t͡ʃ-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{s_a_} marker \obj{k-}.
-It seems to have replaced \rc{t-} only in specific environments, with \obj{k-} and \obj{t͡ʃ-} showing a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
+After the split of \PTir, when \rc{t-}\slash{}\rc{t͡ʃ-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{s_a_} marker \obj{k-}.
+\obj{k-} and \obj{t͡ʃ-} show a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
 \textcite[107]{meira1998proto} largely confirms that distribution, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
@@
 \gla əji-marə-ne w-a-e//
-\glb \gl{2}-with-\gl{pl} \gl{1}-be-\gl{npst}//
+\glb \gl{2}-\gl{com}-\gl{pl} \gl{1}-be-\gl{npst}//
 \glft \qu{I am with you all.}//
@@
 \glpreamble dëmëmara kae ëwï iya//
-\gla n-tə-mə=mara ∅-ka-e əwɨ i-ja//
-\glb \gl{3}-go-\gl{pst}=\gl{dub} \gl{1}-say-\gl{npst}.\gl{cert} \gl{1}\gl{pro} \gl{3}-\gl{obl}//
+\gla n-təmə=mara ∅-ka-e əwɨ i-ja//
+\glb \gl{3}-go=\gl{dub} \gl{1}-say-\gl{npst}.\gl{cert} \gl{1}\gl{pro} \gl{3}-\gl{obl}//
 \glft \qu{“Did s/he leave?”, I say to him.}//
@@
 \xe
+
 %We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
@@
 \label{sec:yukpa}
-The divergent nature of \yukpa\footnote{Very little is known about its sister \japreria, whose status as a dialect of \yukpa is contested by \textcite{oquendo2004japreria}.} is \textit{inter alia}t visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
+The divergent nature of \yukpa\footnote{Very little is known about its sister \japreria, whose status as a dialect of \yukpa is contested by \textcite{oquendo2004japreria}.} is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
 However, it does preserve the \setone prefixes in the immediate past, which shows a unified intransitive paradigm \pcref{tab:yukreg}.

### 2021-12-17 documents/sections/introduction.tex
small fixes

@@
 In some Cariban languages, a small group of verbs show a divergent first person inflection pattern, a topic which has not received much attention in the literature.
-This is illustrated with person paradigms of four \hixka verbs in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
+This is illustrated for \hixka in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
 This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in \cref{tab:hixintro} and other paradigm tables.
@@
 Glossing abbreviations: } %\glossingAbbrevsComma.
-all members of the \gl{s_a_} inflectional class.
+with paradigms of four verbs, all members of the \gl{s_a_} inflectional class.
 In this language, the verb \qu{to be} diverges from other \gl{s_a_} verbs like \qu{to fall} by having a first person marker \obj{w-}, rather than \obj{k-}.
-A similar pattern exists in \trio, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}} \pcref{tab:triintro}.
+A similar pattern exists in \trio \pcref{tab:triintro}, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}}.
 In both languages, the first person prefix of the verbs on the left is representative for the vast majority of \gl{s_a_} verbs.
@@
 In synchronic analyses, these verbs and their first person prefixes may be called \textsc{irregular}, contrasting with regular prefixes (like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}) on regular verbs.
-However, there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[e.g.,][]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
+Note that there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[e.g.,][]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
 In such approaches, the verbs simply belong to a small inflectional (sub-)class, an analysis applied to the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
 
-Ignoring the details of synchronic analysis, the cause for the divergent inflectional patterns is found in the diachrony of the languages in question.
+Regardless of the details of synchronic analysis, the cause for the divergent inflectional patterns is found in the diachrony of the languages in question.
 The goal of this study is to approach these patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows:
-In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns as in \cref{tab:hixintro,tab:triintro} .
+In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns like in \cref{tab:hixintro,tab:triintro} .
 In \cref{sec:data}, six incomplete person marker extensions and the verbs unaffected by them are described.
-Since languages show considerable etymological overlap in their conservative verbs, they are further discussed and reconstructed.
+Since languages show considerable etymological overlap in their conservative verbs, these are further discussed and reconstructed.
 \cref{sec:motivations} uses \posscite{bybee1985morphology} network model of morphology to search explanations for the verbs (un-)affected by each extension.

### 2021-12-17 documents/sections/proto_carib.tex
small fixes

@@
-\section{The origins of conservative first person inflections}
+\section{The origins of irregular first person inflections}
 \label{sec:background}
@@
 	\end{tikzpicture}}
-	\caption{Person marking extensions in Parukotoan, after \textcite[94]{gildea1998}}
+	\caption[Person marker extensions in Parukotoan]{Person marker extensions in Parukotoan, after \textcite[94]{gildea1998}}
 	\label{fig:par_ext}
@@
 \footnote{As an honorable mention, when \ikpeng replaced third person \setone with \settwo prefixes, \obj{a} \qu{to be} and \obj{ke} \qu{to say} retained \obj{n-} \parencite[12]{matter2019arara}.
-	However, the spread of the innovative markers had an entirely different dynamic, spreading from subordinate to main clauses.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
+	However, innovative markers were not spreading within a paradigm, but rather from former subordinate to main clauses.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
 Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, and the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
@@
 Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although their reflexes on verbs have been merged into a single morpheme in modern languages.%
-\footnote{In the \PC transitive verb template, derivational prefixes were in a paradigmatic relationship with the third person marker \rc{i-}:
+\footnote{In the \PC transitive verb template, derivational prefixes were in a paradigmatic relationship with the earlier third person marker \rc{i-}:
 \rc{m-i-}V \qu{you V it}, \rc{m-e-}V \qu{you V yourself}, \rc{m-əte-}V \qu{you V each other}.

### 2021-12-17 documents/sections/verbs.tex
small fixes

@@
 
-In summary, this verb can be reconstructed as being intransitive based on its (inflectional) prefixes, but transitive based on some (derivational) suffixes.
+In summary, this verb can be reconstructed as intransitive based on its (inflectional) prefixes, but transitive based on some (derivational) suffixes.
 \hixka has lost the main intransitive criteria, making its reflex look more like a transitive verb.
@@
 \obj{ɨ} is predominant and can tentatively be reconstructed, although languages throughout the family reflect \rc{ə}.
-In many languages \obj{ɨ} only occurs in certain contexts; but even languages with only C-initial roots show ambiguous inflected forms like \trio \obj{wɨtənne} \qu{I went} \parencite[43]{triomeira1999}, since epenthetic \obj{ɨ} breaks up CC clusters on prefix-verb boundaries.
+Since many languages only show \obj{ɨ} in certain contexts, even forms like \trio \obj{wɨtənne} \qu{I went} \parencite[43]{triomeira1999} are ambiguous, since epenthetic \obj{ɨ} breaks up CC clusters on prefix-verb boundaries.
 
@@
 \carijo \obj{ehɨ} shows innovative \obj{j-}, rather than conservative \obj{w-} \exref{car-18}.
-No \yukpa reflex of this verb is attested, and it was fully replaced in \PWai by \rc{omokɨ} \qu{to come}.
+No \yukpa reflex of this verb is attested, and it was replaced in \PWai by \rc{omokɨ} \qu{to come}.
 
@@
 \akuriyo reflects \rc{əte} as \obj{ee}, where \obj{eepɨ} cannot be a reflex of \rc{epɨ} due the otherwise unexpected vowel length \parencite[114--115]{meira1998proto}.
-Other languages coalesced \rc{əe} to \rc{əə}, with \rc{əəpɨ} being reflected in \uxc \parencite[452]{gildea2012classification}, \kalina, \kaxui, \panare, and \yawarana.
+Other languages coalesced \rc{əte} to \rc{əə}, with \rc{əəpɨ} being reflected in \uxc \parencite[452]{gildea2012classification}, \kalina, \kaxui, \panare, and \yawarana.
 \rc{e}-initial transitive verbs detransitivized with \rc{ət-} underwent the same development in \kalina, creating minimal pairs for length \parencite[509--510]{meira2010origin}, while \rc{ət-e} > \rc{əə} is irregular at least in \kaxui and \panare.
@@
 These roots are reflexes of \rc{[ɨ]pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
-\PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨp} (\gl{tr}) \pcref{sec:pekodian}.
+\PPek can be reconstructed as having the pair \rc{ipɨ} (\gl{intr}) / \rc{ɨpɨ} (\gl{tr}) \pcref{sec:pekodian}.
 Thus, while \PPek \qu{to bathe (\gl{tr})} has perfectly regular cognates in other languages of the family, \qu{to bathe (\gl{intr})} changed \rc{e-} to \rc{i}.

### 2021-12-17 workflow/process_data.py

@@ -450,8 +450,8 @@ table = print_latex(table, keep_index=True)
 save_float(
     table,
     "comepara",
-    r"\rc{ət-epɨ} \qu{to come} in paradigms " + sources,
-    short=r"\rc{ət-epɨ} \qu{to come} in paradigms",
+    r"\rc{(ət-)epɨ} \qu{to come} in paradigms " + sources,
+    short=r"\rc{(ət-)epɨ} \qu{to come} in paradigms",
 )
 
 # comparison of intransitive and transitive 'to bathe'

### 2021-12-25 documents/config/cadernos.tex
journal-specific formatting

\documentclass[a4paper, 11pt, twoside]{article}

%LAYOUT
\usepackage[left=2.4cm,right=2.4cm, top=2.5cm, bottom=2.4cm]{geometry}

%FONTS
\usepackage{fontspec}
\setmainfont{Linux Libertine}

%GLOSSES
\usepackage{glossingtool}
\lingset{aboveglftskip=-0.3ex, everygla=\itshape, belowglpreambleskip=-0.5ex, everyglpreamble=\itshape}
\input{config/own_abbreviations.tex}

%FIRST MENTIONS
\newcommand{\firstmention}[1]{\textsc{#1}}

### 2021-12-25 documents/config/linguistic_discovery.tex
journal-specific formatting

\documentclass[letterpaper, 12pt]{article}

%LAYOUT
\usepackage[left=2.54cm, right=2.54cm, top=2.54cm, bottom=2.54cm]{geometry}

%SPACING
%\usepackage{setspace}
%\doublespacing

%GLOSSES
\usepackage{glossingtool}
\definelingstyle{IJAL}{labelwidth=2em,labelanchor=numleft, labeloffset=0pt,labelformat=(A),everylabel=\actualexno, textanchor=normal,textoffset=1em,preambleanchor=text, preambleoffset=0pt,avoidnumlabelclash, appendtopexarg={samplelabel=(\actualexno a)}}
\lingset{aboveglftskip=-0.3ex, everygla=\itshape, belowglpreambleskip=-0.5ex, everyglpreamble=\itshape, lingstyle=IJAL}
\input{config/own_abbreviations.tex}

%FONTS
\usepackage{fontspec}
\setmainfont{Linux Libertine}
\newfontfamily\objfont{Linux Libertine} %Doulos SIL does not have italics...
\renewcommand{\obj}[1]{\textit{\objfont#1}}
\renewcommand{\rc}[1]{*\textit{\objfont#1}}
\renewcommand{\ort}[1]{$\langle$\objfont#1$\rangle$}

\lingset{everygla=\itshape\objfont, everyglpreamble=\itshape\objfont, everyglb=\objfont}
\usepackage{newunicodechar}
\newunicodechar{∅}{{\objfont∅}}

%HEADERS AND FOOTERS
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[L,C,R]{}
\renewcommand{\headrulewidth}{0pt}
%\fancyhead[R]{}
%\fancyfoot[L]{}
\fancyfoot[C]{\thepage}
%\setlength{\headheight}{14.5pt}

%SECTION HEADERS
\usepackage{titlesec}
\titleformat{\section}
{\normalfont\fontsize{14}{14}\bfseries}{\thesection.}{1ex}{}

\titleformat{\subsection}
{\normalfont\fontsize{12}{12}\bfseries}{\thesubsection.}{1ex}{}

%FIRST MENTIONS
\newcommand{\firstmention}[1]{\textit{#1}}

%TITLE AND ABSTRACT
\date{}
\usepackage{etoolbox}
\AtBeginEnvironment{abstract}{\itshape}

\usepackage{placeins}

%FLOATS AND FOOTNOTES
\usepackage[bottom]{footmisc}

% LINENUMBERS
%\usepackage{lineno}
\newcommand{\linenumbers}{}

### 2021-12-25 documents/sections/introduction.tex
journal-specific formatting

@@
 Such divergent verbs have been identified for \hixka \parencite[188]{hixkaryanaderby1985}, \waiwai \parencite[90]{gildea1998}, the three Taranoan languages \parencite[112--115]{meira1998proto}, \bakairi \parencite{meira2003bakairi}, and \arara \parencite[153]{alves2017arara}, but have only been subject to comparative scrutiny in \posscite{meira1998proto} reconstruction of \PTar.
-In synchronic analyses, these verbs and their first person prefixes may be called \textsc{irregular}, contrasting with regular prefixes (like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}) on regular verbs.
+In synchronic analyses, these verbs and their first person prefixes may be called \firstmention{irregular}, contrasting with regular prefixes (like \hixka \obj{kɨ-} and \trio \obj{t-}/\obj{s-}) on regular verbs.
 Note that there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[e.g.,][]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.

### 2021-12-25 documents/sections/proto_carib.tex
journal-specific formatting

@@
 
-Knowledge about the ancestral system makes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \textsc{conservative}.
+Knowledge about the ancestral system makes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \firstmention{conservative}.
 They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
-The reflexes of \rc{w-} are \textsc{relics}, old and restricted to a few lexemes, contrasting with the innovative prefixes found elsewhere.
+The reflexes of \rc{w-} are \firstmention{relics}, old and restricted to a few lexemes, contrasting with the innovative prefixes found elsewhere.
 These verbs and their prefixes are comparable with the few English nouns like \obj{ɒks}, which preserve the old plural suffix \obj{-ən}.
@@
 In his discussion of the \PC split-\gl{s} system \pcref{sec:split} and reconstruction of the intransitive person prefixes, \textcite[88--96]{gildea1998} shows that the system has undergone many different changes in daughter languages.
-The main mechanism of these changes are \textsc{person marker extensions}, person prefixes being extended to verbal paradigm cells previously occupied by other prefixes.
+The main mechanism of these changes are \firstmention{person marker extensions}, person prefixes being extended to verbal paradigm cells previously occupied by other prefixes.
 %\footnote{

### 2021-12-25 documents/sections/reasons.tex
journal-specific formatting

@@
 It does so by modeling shared morphological properties such as inflectional patterns as emerging from connections of differing strength between lexemes.
-A prominent example is a network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
+A well-known example is a network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
 This network created new strong verbs in some dialects, like \obj{sniːk}--\obj{snʌk} or \obj{brɪŋ}--\obj{brʌŋ} \parencite[129--130]{bybee1985morphology}.
@@
 
-It is intuitively obvious that many of the conservative verbs in \cref{tab:overview} are highly frequent, which would cause conservativeness according to the network model.
+It is intuitively obvious that many of the verbs in \cref{tab:overview} are highly frequent, which would cause conservativeness according to the network model.
 Going beyond intuition is difficult due to the lack of frequency counts for individual lexemes for any Cariban language.
@@
 
-These predictions were checked against the data in \cref{tab:overview} by tallying verbs with (in-)correctly predicted behavior.
+These predictions were checked against the data in \cref{tab:overview} by tallying verbs with (in\-/)correctly predicted behavior.
 The resulting scores are illustrated for the extension of \PTir \rc{t-} in \cref{tab:ptir-evaluations} and summed up for all extensions in \cref{tab:resultsoverview}.
-These scores crucially only consider the seven verbs in \cref{tab:overview}.
-Each extension affected many run-of-the-mill \gl{s_a_} verbs,%
+These scores crucially only consider the seven verbs in \cref{tab:overview}, but each extension affected many run-of-the-mill \gl{s_a_} verbs.%
 \footnote{
@@
 Since none of them is attested as having \gl{s_a_} status at the time of a person marker extension, they are not relevant for this study.}
-except for the \akuriyo \obj{e}-initial verbs.
-If one simulates the addition of 1'000 regularly inflected detransitivized \gl{s_a_} verbs per language -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six explanations consistently predict the behavior of 99.99+\% verbs correctly.
+If one simulates the addition of 1'000 regularly inflected detransitivized \gl{s_a_} verbs per language -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six hypotheses consistently predict the behavior of 99.99+\% verbs correctly.
 Since the descriptive coverage of Cariban languages makes using real such large-scale data impossible, the present investigation is restricted to the edge cases.
@@
 
-The extent of the extensions in both \PWai and \PPek is fully predicted by the presence or absence of a detransitivizer.
+The extent of the \PWai and \PPek and extensions is fully predicted by the presence or absence of a detransitivizer.
 In both cases, only the underived%
@@
 As discussed in \cref{sec:akuriyo}, \akuriyo \obj{k-} only appears on \obj{ə}-initial verbs.
-In \carijo, the extension of \obj{j-} affected all \obj{e}- and \obj{ə}-initial verbs, including \obj{eh} \qu{to come} or \obj{et͡ʃi} \qu{to be}, which do not have a detransitivizing prefix.
+In \carijo, the extension of \obj{j-} affected all \obj{e}- and \obj{ə}-initial verbs, including \obj{eh} \qu{to come} and \obj{et͡ʃi} \qu{to be}, which do not have a detransitivizing prefix.
 Only \obj{ka} \qu{say}, \obj{tə[mə]} \qu{go}, and \obj{a} \qu{be-1} did not take on \obj{j-}.
-The extension of \yukpa \obj{j-} can succinctly be characterized as affecting all vowel-initial verbs; the only attested conservative verb is C-initial \obj{to} \qu{to go}.
+The extension of \yukpa \obj{j-} can succinctly be characterized as affecting all vowel-initial verbs, as the only attested conservative verb is C-initial \obj{to} \qu{to go}.
 
@@
 However, this explanation rests on the analysis of \obj{t-} and \obj{t͡ʃ-} as distinct morphemes.
-Due to their phonologically conditioned distribution in \PTir, the putative inflectionally defined network is identical to the phonological one.
+Due to their phonologically conditioned distribution in \PTir, the hypothetical inflectionally defined network is identical to the phonological one.
 
 When adding the expected conservative effects of frequency, prediction accuracy was improved in 8 cases, stagnated in 7 cases, and worsened in three cases.
-These three cases where the crude model of verb frequency arrives at incorrect predictions are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
-Overall, including frequency in the model led to improvements, up to 100\% prediction accuracy for all three factors in \PTir, phonology and inflection in \PWai, and phonology in \PPek.
+These three cases are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
+Overall, including the crude estimate of frequency in the model led to improvements, up to 100\% prediction accuracy for all three factors in \PTir, phonology and inflection in \PWai, and phonology in \PPek.
 
-In summary, one can conclude that the patterns of most extensions are correctly predicted not by a single explanation, but rather by 3 to 4 different ones.
-The exceptions are \carijo and \yukpa, where a phonologically defined lexical network emerges as the unambiguous winner, while frequency-based explanations fare much worse.
+One can conclude that the patterns of most extensions are correctly predicted not by a single explanation, but rather by 3 to 4 different ones.
+The exceptions are \carijo and \yukpa, where a phonologically defined lexical network emerges as the unambiguous winner, while frequency-based explanations fare worse.
 %It may be worth noting that the three extensions which we do not reconstruct to a proto-language (ti.e., which are more recent) are best explained by phonological conditioning factors.
 For the other extensions, the network model gives no unambiguous answer to the question of what (combination of) factors caused innovative first person markers to spread the way they did.
-This is due three of the factors in the model -- frequency, detransitivizer, phonology -- strongly converging in their predictions.
+This is due three of the factors in the model -- frequency, detransitivizer, phonology -- strongly converging in their predictions:
 The most frequent \gl{s_a_} verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.

### 2023-01-13 documents/sections/introduction.tex
final submission

@@
 In languages with strong morphophonological processes and/or non-phonemic orthography the original form is shown in an additional line when presented in interlinearized examples.
-\textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.
-Glossing abbreviations: } %\glossingAbbrevsComma.
+\textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
 with paradigms of four verbs, all members of the \gl{s_a_} inflectional class.

### 2023-01-25 documents/config/cariban_language_names.tex
implemented all recommendations from review 1

\usepackage{xspace}
\providecommand{\apalai}{Apalaí\xspace}
\DeclareAcronym{apa_lg}{short=apa,long=Apalaí,short-format=\scshape}
\providecommand{\trio}{Tiriyó\xspace}
\DeclareAcronym{tri_lg}{short=tri,long=Tiriyó,short-format=\scshape}
\providecommand{\hixka}{Hixkaryána\xspace}
\DeclareAcronym{hix_lg}{short=hix,long=Hixkaryána,short-format=\scshape}
\providecommand{\panare}{Panare\xspace}
\DeclareAcronym{pan_lg}{short=pan,long=Panare,short-format=\scshape}
\providecommand{\macushi}{Macushi\xspace}
\DeclareAcronym{mac_lg}{short=mac,long=Macushi,short-format=\scshape}
\providecommand{\waimiri}{Waimiri-Atroari\xspace}
\DeclareAcronym{wmr_lg}{short=wmr,long=Waimiri-Atroari,short-format=\scshape}
\providecommand{\maqui}{Ye'kwana\xspace}
\DeclareAcronym{mak_lg}{short=mak,long=Ye'kwana,short-format=\scshape}
\providecommand{\waiwai}{Waiwai\xspace}
\DeclareAcronym{wai_lg}{short=wai,long=Waiwai,short-format=\scshape}
\providecommand{\wayana}{Wayana\xspace}
\DeclareAcronym{way_lg}{short=way,long=Wayana,short-format=\scshape}
\providecommand{\kalina}{Kari'ña\xspace}
\DeclareAcronym{kar_lg}{short=kar,long=Kari'ña,short-format=\scshape}
\providecommand{\ikpeng}{Ikpeng\xspace}
\DeclareAcronym{ikp_lg}{short=ikp,long=Ikpeng,short-format=\scshape}
\providecommand{\uxc}{Upper Xingu Carib\xspace}
\DeclareAcronym{uxc_lg}{short=uxc,long=Upper Xingu Carib,short-format=\scshape}
\providecommand{\arara}{Arara\xspace}
\DeclareAcronym{ara_lg}{short=ara,long=Arara,short-format=\scshape}
\providecommand{\ingariko}{Ingarikó\xspace}
\DeclareAcronym{ing_lg}{short=ing,long=Ingarikó,short-format=\scshape}
\providecommand{\pemon}{Pemón\xspace}
\DeclareAcronym{pem_lg}{short=pem,long=Pemón,short-format=\scshape}
\providecommand{\kapon}{Kapón\xspace}
\DeclareAcronym{kap_lg}{short=kap,long=Kapón,short-format=\scshape}
\providecommand{\kalapalo}{Kalapalo\xspace}
\DeclareAcronym{kal_lg}{short=kal,long=Kalapalo,short-format=\scshape}
\providecommand{\kuikuro}{Kuikuro\xspace}
\DeclareAcronym{kui_lg}{short=kui,long=Kuikuro,short-format=\scshape}
\providecommand{\yukpa}{Yukpa\xspace}
\DeclareAcronym{yuk_lg}{short=yuk,long=Yukpa,short-format=\scshape}
\providecommand{\carijo}{Carijona\xspace}
\DeclareAcronym{car_lg}{short=car,long=Carijona,short-format=\scshape}
\providecommand{\bakairi}{Bakairi\xspace}
\DeclareAcronym{bak_lg}{short=bak,long=Bakairi,short-format=\scshape}
\providecommand{\tamanaku}{†Tamanaku\xspace}
\DeclareAcronym{tam_lg}{short=tam,long=†Tamanaku,short-format=\scshape}
\providecommand{\kaxui}{Werikyana\xspace}
\DeclareAcronym{kax_lg}{short=kax,long=Werikyana,short-format=\scshape}
\providecommand{\akuriyo}{Akuriyó\xspace}
\DeclareAcronym{aku_lg}{short=aku,long=Akuriyó,short-format=\scshape}
\providecommand{\japreria}{Japreria\xspace}
\DeclareAcronym{jap_lg}{short=jap,long=Japreria,short-format=\scshape}
\providecommand{\kumana}{†Kumaná\xspace}
\DeclareAcronym{cum_lg}{short=cum,long=†Kumaná,short-format=\scshape}
\providecommand{\palmela}{†Palmela\xspace}
\DeclareAcronym{palm_lg}{short=palm,long=†Palmela,short-format=\scshape}
\providecommand{\piment}{†Pimenteira\xspace}
\DeclareAcronym{pime_lg}{short=pime,long=†Pimenteira,short-format=\scshape}
\providecommand{\arakaju}{†Arakaju\xspace}
\DeclareAcronym{arak_lg}{short=arak,long=†Arakaju,short-format=\scshape}
\providecommand{\pauxiana}{†Pauxiana\xspace}
\DeclareAcronym{pxan_lg}{short=pxan,long=†Pauxiana,short-format=\scshape}
\providecommand{\paravi}{†Paravilhana\xspace}
\DeclareAcronym{para_lg}{short=para,long=†Paravilhana,short-format=\scshape}
\providecommand{\bonari}{†Bonari\xspace}
\DeclareAcronym{bona_lg}{short=bona,long=†Bonari,short-format=\scshape}
\providecommand{\wayu}{†Wayumara\xspace}
\DeclareAcronym{wayu_lg}{short=wayu,long=†Wayumara,short-format=\scshape}
\providecommand{\apiaka}{†Apiaka-Tocantins\xspace}
\DeclareAcronym{apia_lg}{short=apia,long=†Apiaka-Tocantins,short-format=\scshape}
\providecommand{\opon}{†Opón-Carare\xspace}
\DeclareAcronym{opon_lg}{short=opon,long=†Opón-Carare,short-format=\scshape}
\providecommand{\pauxi}{†Pauxi\xspace}
\DeclareAcronym{paux_lg}{short=paux,long=†Pauxi,short-format=\scshape}
\providecommand{\sapara}{†Sapara\xspace}
\DeclareAcronym{sapa_lg}{short=sapa,long=†Sapara,short-format=\scshape}
\providecommand{\tiveri}{†Tiverighotto\xspace}
\DeclareAcronym{tive_lg}{short=tive,long=†Tiverighotto,short-format=\scshape}
\providecommand{\yao}{†Yao\xspace}
\DeclareAcronym{yao_lg}{short=yao,long=†Yao,short-format=\scshape}
\providecommand{\yaruma}{†Yarumá\xspace}
\DeclareAcronym{yaru_lg}{short=yaru,long=†Yarumá,short-format=\scshape}
\providecommand{\koyaima}{†Koyaima\xspace}
\DeclareAcronym{koya_lg}{short=koya,long=†Koyaima,short-format=\scshape}
\providecommand{\hiana}{†Hianakoto-Umawa\xspace}
\DeclareAcronym{hian_lg}{short=hian,long=†Hianakoto-Umawa,short-format=\scshape}
\providecommand{\saluma}{†Salumá\xspace}
\DeclareAcronym{salu_lg}{short=salu,long=†Salumá,short-format=\scshape}
\providecommand{\sikiana}{†Sikiana\xspace}
\DeclareAcronym{siku_lg}{short=siku,long=†Sikiana,short-format=\scshape}
\providecommand{\purukoto}{†Purukoto\xspace}
\DeclareAcronym{puru_lg}{short=puru,long=†Purukoto,short-format=\scshape}
\providecommand{\juma}{†Juma\xspace}
\DeclareAcronym{juma_lg}{short=juma,long=†Juma,short-format=\scshape}
\providecommand{\palenque}{†Palenque\xspace}
\DeclareAcronym{pale_lg}{short=pale,long=†Palenque,short-format=\scshape}
\providecommand{\parukoto}{†Parukoto\xspace}
\DeclareAcronym{paru_lg}{short=paru,long=†Parukoto,short-format=\scshape}
\providecommand{\urukue}{†Urukuena\xspace}
\DeclareAcronym{uruk_lg}{short=uruk,long=†Urukuena,short-format=\scshape}
\providecommand{\wayakule}{†Wayakulé\xspace}
\DeclareAcronym{waya_lg}{short=waya,long=†Wayakulé,short-format=\scshape}
\providecommand{\piritu}{†Píritu\xspace}
\DeclareAcronym{piri_lg}{short=piri,long=†Píritu,short-format=\scshape}
\providecommand{\guake}{†Guake\xspace}
\DeclareAcronym{guak_lg}{short=guak,long=†Guake,short-format=\scshape}
\providecommand{\macoita}{†Macoita\xspace}
\DeclareAcronym{mcta_lg}{short=mcta,long=†Macoita,short-format=\scshape}
\providecommand{\macoa}{†Macoa\xspace}
\DeclareAcronym{maco_lg}{short=maco,long=†Macoa,short-format=\scshape}
\providecommand{\chake}{†Chake\xspace}
\DeclareAcronym{chak_lg}{short=chak,long=†Chake,short-format=\scshape}
\providecommand{\chayma}{†Chayma\xspace}
\DeclareAcronym{chai_lg}{short=chai,long=†Chayma,short-format=\scshape}
\providecommand{\avarikoto}{†Avarikoto\xspace}
\DeclareAcronym{avar_lg}{short=avar,long=†Avarikoto,short-format=\scshape}
\providecommand{\tauli}{Taulipang\xspace}
\DeclareAcronym{tau_lg}{short=tau,long=Taulipang,short-format=\scshape}
\providecommand{\patamona}{Patamona\xspace}
\DeclareAcronym{pat_lg}{short=pat,long=Patamona,short-format=\scshape}
\providecommand{\arekuna}{Arekuna\xspace}
\DeclareAcronym{are_lg}{short=are,long=Arekuna,short-format=\scshape}
\providecommand{\kamara}{Kamarakoto\xspace}
\DeclareAcronym{kam_lg}{short=kam,long=Kamarakoto,short-format=\scshape}
\providecommand{\akawaio}{Akawaio\xspace}
\DeclareAcronym{aka_lg}{short=aka,long=Akawaio,short-format=\scshape}
\providecommand{\kumanagoto}{†Kumanagoto\xspace}
\DeclareAcronym{cgt_lg}{short=cgt,long=†Kumanagoto,short-format=\scshape}
\providecommand{\crichana}{†Crichana\xspace}
\DeclareAcronym{cric_lg}{short=cric,long=†Crichana,short-format=\scshape}
\providecommand{\mapoyo}{Mapoyo\xspace}
\DeclareAcronym{map_lg}{short=map,long=Mapoyo,short-format=\scshape}
\providecommand{\yawarana}{Yawarana\xspace}
\DeclareAcronym{yab_lg}{short=yab,long=Yawarana,short-format=\scshape}
\providecommand{\yawaperi}{†Jawaperi\xspace}
\DeclareAcronym{jawa_lg}{short=jawa,long=†Jawaperi,short-format=\scshape}
\providecommand{\pemono}{†Pémono\xspace}
\DeclareAcronym{pemo_lg}{short=pemo,long=†Pémono,short-format=\scshape}
\providecommand{\pianakoto}{†Pianakoto\xspace}
\DeclareAcronym{pian_lg}{short=pian,long=†Pianakoto,short-format=\scshape}
\providecommand{\PC}{Proto-Cariban\xspace}
\DeclareAcronym{pc_lg}{short=pc,long=Proto-Cariban,short-format=\scshape}
\providecommand{\arina}{†Arinakoto\xspace}
\DeclareAcronym{arin_lg}{short=arin,long=†Arinakoto,short-format=\scshape}
\providecommand{\kumayena}{†Kumayena\xspace}
\DeclareAcronym{kuma_lg}{short=kuma,long=†Kumayena,short-format=\scshape}
\providecommand{\tyrewuju}{Tyrewuju\xspace}
\DeclareAcronym{tyr_lg}{short=tyr,long=Tyrewuju,short-format=\scshape}
\providecommand{\dekwana}{De’kwana\xspace}
\DeclareAcronym{dek_lg}{short=dek,long=De’kwana,short-format=\scshape}
\providecommand{\dhekwana}{Dhe’kwana\xspace}
\DeclareAcronym{dhe_lg}{short=dhe,long=Dhe’kwana,short-format=\scshape}
\newcommand{\PPar}{Proto-Parukotoan\xspace}
\newcommand{\PPars}{\textsc{PPar}\xspace}
\newcommand{\PPem}{Proto-Pemongan\xspace}
\newcommand{\PPems}{\textsc{PPem}\xspace}
\newcommand{\PPek}{Proto-Pekodian\xspace}
\newcommand{\PPeks}{\textsc{PPek}\xspace}
\newcommand{\PPP}{Proto-Panare-Pemongan\xspace}
\newcommand{\PPPs}{\textsc{PPP}\xspace}
\newcommand{\PTir}{Proto-Tiriyoan\xspace}
\newcommand{\PTirs}{\textsc{PTir}\xspace}
\newcommand{\PWai}{Proto-Waiwaian\xspace}
\newcommand{\PWais}{\textsc{PWai}\xspace}
\newcommand{\PXin}{Proto-Xinguan\xspace}
\newcommand{\PXins}{\textsc{PXin}\xspace}
\newcommand{\PTar}{Proto-Taranoan\xspace}
\newcommand{\PTars}{\textsc{PTar}\xspace}
\newcommand{\yabmap}{\mapoyo-\yawarana}
\newcommand{\yabmaps}{\textsc{yab-map}\xspace}
\newcommand{\vzk}{Venezuelan \kalina}
\newcommand{\vzks}{\textsc{vzk}\xspace}
\newcommand{\esk}{Eastern Surinamese \kalina}
\newcommand{\esks}{\textsc{esk}\xspace}
\newcommand{\wsk}{Western Surinamese \kalina}
\newcommand{\wsks}{\textsc{wsk}\xspace}
\renewcommand{\lg}[1]{\acs*{#1_lg}}

### 2023-01-26 review_answers.md
review answers

# Review 1

* **169**: *Example 5 is not aligned*

	* There is nothing that could be “aligned” in this example. I interpreted it as "make the distance between object language and translation identical to example 4", and added translations for all wordforms in the paradigm.

* **188**: *The author Alves Chagas (2013) is known as Chagas (2013)*

	* I've changed the name.

# Review 2

* **Overall**: *My major concern has to do with the reconstructed Proto-Pekodian forms of verbs. Although
some sound correspondences between Bakairi and Ikpeng have been identified in Meira & Franchetto (2005), I am not aware of any existing study that would address the phonological
reconstruction of Proto-Pekodian. This is important because not all assertions regarding
cognacy are equally convincing. For instance, it is not obvious to this reviewer that
Proto-Pekodian *-ɨtən- is a valid reconstruction, and that the purported reflexes of this form
are exact cognates: Ikpeng -aran- seems to point to *-ətən-, Bakairi t, as opposed to *d,
points to an erstwhile consonant cluster (*wɨ-wtə-?); the Arara form could be cognate with
the Bakairi counterpart, but not with the Ikpeng one. If the Ikpeng form is actually
noncognate, the author’s calculations in Part 2 would have to be redone. The irregular
correspondences between -iptoŋ- and -ɨtəgɨ- are likewise suspect. I’d like the author to
include a section, or maybe an appendix, explaining the basic premises of their
reconstruction of Proto-Pekodian.*

	* TBD

* **Overall**: *Another suggestion, which is easier to implement but is perhaps somewhat more radical, is
to split the manuscript in two. Since the author’s calculations in Part 2 are entirely dependent
on the correctness of their findings in Part 1, and since the methodologies employed in these
two parts of the manuscripts are entirely different (comparative method + Diachronic
Construction Grammar vs. quantitative study), I’d recommend the author to consider the
possibility to publish their Part 2 as a separate paper, possibly in a journal specializing on
quantitative linguistics (though of course it still lies within Cadernos’ scope), since its
methodology could be of interest to readers that outside the Americanist community. I leave
this to the author’s discretion.*

	* TBD

* **Data**: *The author reports that no first-person forms of the Ikpeng verb for ‘to go down’ are attested,
with a reference to Angela Chagas. I am not a specialist on Ikpeng, but I have just asked aIkpeng acquaintance (whose name is Fael Ikpeng in the case the author would like to
acknowledge his help), and he gave me the form iptongtxi (∅-iptoŋ-tʃi) for ‘I’ll go down’,
which fits nicely the author’s predictions.*

	* TBD

* **Thoughout**: *The use of ‘to’ in the English glosses of verbs is non-consistent. Please either use it
in all glosses or leave it out.*

	* TBD

* **Throughout**: *For languages with significant dialectal variation, such as Kari’ña or Bakairi, please
indicate the dialect from which a given datum was extracted. For example, it would
be important to know whether Bakairi k-əewɨ- belongs to a dialect that lost
intervocalic rhotics (this would make its derivation from *-ət-epɨ- more credible).*

	* TBD

* **Figure 1**: *Trinidad and Tobago is missing*

	* TBD

* **Figure 1**: *Extinct languages are not shown. If this is on purpose, please modify the label, but I
believe it is important for the reader to know that there used to be Cariban languages
outside the current Cariban-speaking area, such as Pimenteira in Piauí, Cumanagoto
and other groups on the Caribbean coast, and Kari’ña varieties in the Antilles and in
Trinidad.*

	* TBD

* **Figure 1**: *Languages such as Kari’ña really should be represented by meshes/polygons rather
than by points. Bakairi, Upper Xingu Carib, and other languages could be best
represented by several points.*

	* TBD

* **Figure 1**: *The spelling Kapón is not particularly common as opposed to the more common
Kapong (or Akawaio, or Ingarikó).*

	* TBD

* **footnote 1**: *“The presence of a 1+2 person value implies that of a 1+3 value” — I’d leave this out.
In fact, there are many languages where 1+2 exists as a category, but 1+3 isn’t
encoded differently from 1 (which has led some linguists to refer to 1+2 as “the fourth
person”).*

	* TBD

* **footnote 1**: *“Further, TAM suffixes in the attested forms are omitted, since a) the focus lies on the
prefixes and stems, and b) full paradigms with the same TAM suffix are rare in the
available sources” — the justification is legitimate, but the reader might want to know
whether different TAM configurations imply different person indexing patterns, as they
often do. It is often the case that different participants (say, A vs. P) are indexed in
the same affix slot under different TAM values in some languages, and even for the
same notional participant different prefixes (belonging to different sets) may occur. A
clarification is needed on which TAM values are seen as representative in your paper
–– I assume you only consider the ones that select for Gildea’s Set I.*

	* TBD

* **footnote 1**: *“Standard IPA symbols are used in the transcription of Cariban languages” →
“Standard IPA symbols are used for linguistic data” — you also use IPA for Germanic
languages*

	* TBD

* **footnote 1**: *“In languages with strong morphophonological processes and/or non-phonemic
orthography...” — I would make this uniform across the paper, when possible. It would be more reader-friendly if you either systematically included the contemporary
orthographic representation or always left it out. The inconsistency looks particularly
out of place in the Hixkaryana data in 11b–c, and in the Kari’ña examples in 2a–b. I
understand that different sources use different transcription systems, but I have no
doubt that the author has the necessary knowledge to easily convert between
orthographies — could you maybe always follow the conventions used by the
speakers?*

	* TBD

* **footnote 1**: *“Gildea (2018) is followed in using ⟨ə⟩ for the Proto-Cariban reconstructed by Meira &
Franchetto (2005)” → “Gildea (2018) is followed in using ⟨ə⟩ for the Proto-Cariban *ô
reconstructed by Meira & Franchetto (2005)”*

	* TBD

* **section 2.1**: *Consider defining “locuphoric” and “aliophoric”. This reviewer is familiar with
Haspelmath’s preferred terminology, but the reader cannot be expected to — Gildea (1998)
does not use these terms for obvious reasons, and a search on Google returns 14 results for
“aliophoric”. Also note that you use “aliophoric” and “nonlocal” within the same period,
though Haspelmath’s intention was to altogether replace the term “nonlocal” with
“aliophoric”.*

	* TBD

* **lines 60–61**: *Don’t use hyphens for morpheme segmentation within glosses*

	* TBD

* **Figure 2**: *Does Werikyana really lack the (j) in 2SP.P before vowel-initial stems? Does the j-
element actually always show up after owɨ(ro)-, or should it be parenthesized? (also in line
5 of example 1)*

	* TBD

* **line 107**: *Consider defining the term “deverbalized” (does this encompass all nonfinite
forms?), which will be relevant for the reader’s comprehension of the paradigm of the
Proto-Cariban verb for ‘to go’.*

	* TBD

* **between lines 110 and 111 (also 114–121)**: *“In modern instantiations of Cariban split-S,
mismatches between the semantics of verbs and their SA or SP status are common,
exemplified with Kari’ña data in (2)” — This piece of discussion strikes me as a strawman
fallacy, at least the way it is phrased in the current version of the manuscript. It is by now a
well-established fact that split-S does not typically show a 100% Guarani-like correlation with
semantics in many languages, so I can’t see why one would expect there to be such a
correlation in Cariban. If you are contesting claims found in earlier literature, consider citing
them.*

	* TBD

* **lines 127–8**: *“although their reflexes on verbs have been merged into a single morpheme in
modern languages” → “although their reflexes on verbs have been merged into a single
morpheme in some modern languages”*

	* TBD

* **example 4**: *Consider segmenting the TAM suffixes off the roots so as to facilitate the
recognition of cognate roots across Pekodian*

	* TBD

* **example 7d**: *k-ɨtəgɨ-se cannot seem to be a correct restitution of ⟨kχ-itaké-he⟩. Is this an
instance of Proto-Bakairi *ʔs, as reconstructed by Meira (2005), which is reflected as ʔ in
Western Bakairi and as s in Eastern Bakairi?*

	* TBD

* **Section 3.1.1**: *It should be made clear at this point that the discussion of the reconstructed
forms is postponed to Section 3.2, otherwise the reader is left wondering whether the
author’s cognacy judgments can be trusted.*

	* TBD

* **Table 6**: *Consider replacing *w-epɨ- with *w-(əd-)epɨ-. The form *w-əd-epɨ- certainly existed
in Proto-Pekodian, since it has exact cognates elsewhere in Cariban.*

	* TBD

* **Table 8**: *Asterisks are missing in PW forms.
example 11: consider swapping (c) and (d)*

	* TBD

* **line 239**: *“a fossilized of the SA class marker” → “a fossilized SA class marker” (incidentally,
why not consider it a nonfiniteness prefix/nominalizer? “Class marker” is not a transparent
choice, especially if the morpheme in question only shows up in nonfinite forms in PC.)*

	* TBD

* **line 240**: *“the first person form Hixkaryána ‘to go’” — consider rephrasing this
footnote 8: “Here, both Tiriyó əepɨ and Akuriyó eepɨ are identified as reflecting *ətepɨ
(Section 3.2.4), via Proto-Taranoan *(ət͡ʃ-)epɨ and Proto-Tiriyoan *(əʔ)epɨ” — Why the
element *ət- is segmented (and parenthesized) in the Proto-Taranoan and Proto-Tiriyoan
forms, but not in the Proto-Cariban one?*

	* TBD

* **example 12**: *I fail to see what does the example 12a illustrate. If anything, *w is not lost in it.*

	* TBD

* **example 13**: *The form p-ɨ(h)tə- should be discussed in 3.1.3, on Tiriyoan, and not in 3.1.4,
on Akuriyó, since this form is also attested in Tiriyó.*

	* TBD

* **example 15**: *Apart from non-uniform transcription systems (IPA in (a–b), orthography in (c));
-e is glossed as NPST in (a–b), but as NPST.CERT in (c) — is this correct?*

	* TBD

* **lines 332–4**: *“a choice not only motivated by a desire to avoid an idiosyncratic intransitive
first person prefix ɨ- instead of regular kɨ-” — the period is unfinished*

	* TBD

* **Table 15 and subsequent tables**: *This reviewer doesn’t think the aligned presentation of
data contributes much to the discussion — on the contrary, it confuses the reader. Even
though I am familiar with List’s ideology on this matter, the presentation scheme is
suboptimal for presenting allomorphy (ke/ge in Pekodian), and obfuscates the identity
correspondence between PC *ati (and its reflexes) and Pekodian e, Kuhikugu/Kalapalo/Matipu i (I assume *aCi > *ai > *e > Pekodian e, UXC i).*

	* TBD

* **lines 373–375**: *“Since many languages only show ɨ in certain contexts, even forms like Tiriyó
wɨtənne ‘I went’ (Meira 1999: 43) are ambiguous, since epenthetic ɨ breaks up CC clusters
on prefix-verb boundaries” — consider rephrasing this so as to avoid two since-clauses.*

	* TBD

* **example 20**: *I can’t see why would you want to you a two-lines presentation of the data,
since there are no “strong morphophonological processes and/or non-phonemic
orthography” involved.*

	* TBD

* **footnote 13**: *“final ŋ in Proto-Xinguan”: an asterisk is missing. In addition, the explanation
given is a non-explanation. Couldn’t PX *-ŋ be a previously unnoticed retention from
Proto-Cariban?*

	* TBD

* **lines 498–499**: *“These verbs formed a lexical connection with prototypical members of the
group, developing new past tense forms as a result” — did you mean ‘patient participle’?
section 4: Please justify the choice of Apalaí texts for the frequency count — why not the
exceptionally-well described Kari’ña? It is true that there are no ready-to-use corpora for
Cariban languages, but there are many glossed texts available. Another good start would be
the Kuhikugu and Kalapalo texts in Stenzel & Franchetto (2017).
section 4: In frequency counts, it should be made clear if one counts the frequency of
verbal lexemes or specifically of verbs inflected for 1st person (some verb may only be
frequent in the 3rd person).*

	* TBD

* **Abbreviations**: *“S marked like A” → “S marked more like A than like P” (?)*

	* TBD

* **References**: *consider citing the recently published version of Cáceres and Gildea’s
Amazonicas presentation on Tense-Aspect constructions in Yawarana (<http://www.etnolinguistica.org/article:vol10n1-7>).*

	* TBD

* **lines 29–30**: *“like in Tables 1 and 2 .” — remove the space*

	* TBD

* **Table 22**: *something is wrong with the table’s width and the layout of the label*

	* TBD

* **Abbreviations**: *uncertainty.9 → uncertainty.*

	* TBD

* **References**: *(Franchetto 1986): Estudo Etnolinguistico → Estudo etnolinguístico*

	* TBD

* **References**: *The use of capital letters in Spanish and Portuguese titles is inconsistent.*

	* TBD

### 2023-08-04 documents/sections/extensions.tex
font, 'alignment'
implemented all recommendations from review 1
readthrough

@@
 The Pekodian branch consists of closely related \arara and \ikpeng, with \bakairi as a more distant member.
-The contribution establishing the branch \parencite{meira2005southern} focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
+\textcite{meira2005southern}, who established the branch, focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
 However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
@@
 The most detailed description of a Pekodian language \parencite{alves2017arara} names six%
-\footnote{Seven in her analysis, treating the two meanings of \obj{it͡ʃi} \qu{to be, to lie down} as different verbs.}
+\footnote{Seven in her analysis, treating the two meanings of \obj{it͡ʃi} \qu{to be / to lie down} as different verbs.}
 \arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
@@
 \ex<araap> \arara \parencite[200]{alves2017arara}\\
-\begin{tabular}[t]{@{}ll@{}}
-\gl{1} & \obj{w-aptam} \qu{when/if I was}\\
-\gl{2} & \obj{m-od-aptam}\\
-\gl{1+2} & \obj{kud-aptam}\\
-\gl{3} & ∅\obj{-aptam}\\
+\begin{tabular}[t]{@{}lll@{}}
+\gl{1} & \obj{w-aptam} & \qu{when/if I was}\\
+\gl{2} & \obj{m-od-aptam} & \qu{when/if you were}\\
+\gl{1+2} & \obj{kud-aptam} & \qu{when/if we were}\\
+\gl{3} & ∅\obj{-aptam} & \qu{when/if s/he was}\\
 \end{tabular}
@@
 \label{sec:yukpa}
-The divergent nature of \yukpa\footnote{Very little is known about its sister \japreria, whose status as a dialect of \yukpa is contested by \textcite{oquendo2004japreria}.} is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
+The divergent nature of \yukpa\footnote{Very little is known about its sister \japreria, whose status as a dialect of \yukpa is contested by \textcite{oquendo2004japreria}.} is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative main clause verbal constructions \parencite{meira2006syntactic}.
 However, it does preserve the \setone prefixes in the immediate past, which shows a unified intransitive paradigm \pcref{tab:yukreg}.
@@
 
-In intransitive verbs, this first-person zero marking is attested with a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
+Zero-marked first person is attested for a single intransitive verb, \obj{to} \qu{to go} \exref{yuk-7}.
 It diverges from regular C-initial verbs with \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.

### 2023-08-04 documents/sections/introduction.tex
readthrough

@@
 In this language, the verb \qu{to be} diverges from other \gl{s_a_} verbs like \qu{to fall} by having a first person marker \obj{w-}, rather than \obj{k-}.
-A similar pattern exists in \trio \pcref{tab:triintro}, where the verb \qu{to go} has a first-person prefix \obj{wɨ-} while other \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}}.
-In both languages, the first person prefix of the verbs on the left is representative for the vast majority of \gl{s_a_} verbs.
+A similar pattern exists in \trio \pcref{tab:triintro}, where \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}}, while the verb \qu{to go} has a first-person prefix \obj{wɨ-}.
+For both languages, the first person prefix on the first four verbs in the table is representative for the vast majority of \gl{s_a_} verbs.
 %In both languages, there are only a few other verbs inflected identically to the divergent ones on the right; for example, the first-person form of \trio \qu{to be} is \obj{w-ei-} \parencites[339]{triomeira1999}.
@@
 Note that there is no widely accepted definition of irregularity \parencite{stolz2012introduction}, and many stricter definitions \parencite[e.g.,][]{haspelmath2010understanding} require the pattern to occur at a single place in the grammar.
-In such approaches, the verbs simply belong to a small inflectional (sub-)class, an analysis applied to the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
+In such approaches, these verbs simply belong to a small inflectional (sub-)class, an analysis applied to the Pekodian languages \bakairi and \arara \parencites[4]{meira2003bakairi}[149]{alves2017arara}.
 

### 2023-08-04 documents/sections/proto_carib.tex
readthrough

@@
 The irregular first person prefixes from \cref{sec:intro} are relics, inherited from the ancestral \PC system \pcref{sec:pc_person}.
-That system underwent much innovation; the mechanism responsible for the irregular forms are person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
+That system underwent much innovation; the mechanism responsible for the irregular forms involves person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
 A specific aspect of the system, the \gl{s_a_} vs \gl{s_p_} distinction, plays a role in incomplete extensions and is discussed in \cref{sec:split}.
@@
 Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
-\gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}\gl{s_a_} \rc{w-} vs \gl{1}>\gl{3} \rc{t(i)-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
+\gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}\gl{s_a_} \rc{w-} vs \gl{1}>\gl{3} \rc{t[i]-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
 On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones, with \gl{3}\gl{s_p_} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
 
-Knowledge about the ancestral system makes clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \firstmention{conservative}.
+Knowledge about the ancestral system makes evident that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \firstmention{conservative}.
 They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
@@
 %
-All innovations are person marker extensions except 2b, which combined a pronoun with the linker \rc{j-}.
+All innovations are person marker extensions except the innovative first person P form, which combined a pronoun with the linker \rc{j-}.
 \cref{fig:par_ext} shows them in bold and reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
@@
 
-For \textcite{gildea1998}, person marker extensions are relevant for the loss of split-\gl{s} and the accompanying changes to indexing alignment, whereas this study focuses on a different aspect.
+For \textcite{gildea1998}, person marker extensions are relevant for the loss of split-\gl{s} and the resulting changes to indexing alignment, whereas this study focuses on a different aspect.
 Namely, they most likely took place via lexical diffusion, characterized as a type of extension by \textcite[106--115]{harris1995historical}, a hypothesis supported by three facts.
@@
 
-The third argument in favor of the lexical diffusion scenario goes back to conservative forms like in \hixka and \trio.
-Both innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
+The third argument in favor of the lexical diffusion scenario goes back to conservative forms like those in \hixka and \trio.
+In both languages, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
 The continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs is the result of the extension stopping short of these verbs, rather than spreading through the entire \gl{s_a_} lexicon.
@@
 \begin{quotebox}{\parencite[201]{meira2000split}}
-	Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still exisiting transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
+	Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still existing transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
 \end{quotebox}

### 2023-08-04 documents/sections/reasons.tex
readthrough

@@
 For the other extensions, the network model gives no unambiguous answer to the question of what (combination of) factors caused innovative first person markers to spread the way they did.
-This is due three of the factors in the model -- frequency, detransitivizer, phonology -- strongly converging in their predictions:
+This is due three of the candidate factors in the model -- frequency, detransitivizer, phonology -- strongly converging in their predictions:
 The most frequent \gl{s_a_} verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.

### 2023-08-04 documents/sections/verbs.tex
readthrough

@@
 In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}), while does not occur in \exref{hixsay.ordpl-35} where the verb is preceded by direct speech.
-Outside of \hixka, at least the \trio reflex shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
+In the rest of the family, at least the \trio reflex shows the same pattern, albeit inconsistently so \parencite[267]{triocarlin2004}.
 
@@
 \hixka \obj{ntoje} \qu{he went} \parencite[27]{hixkaryanaderby1985}, \trio \obj{təkə} \qu{go!} \parencite[246]{triomeira1999}, or \wayana \obj{kuptəm} \qu{we went} \parencite[195]{wayanatavares2005}}, the contrast with C-initial \rc{ka[ti]} becomes clear once all forms are considered \pcref{tab:go}.
-\obj{ɨ} is predominant and can tentatively be reconstructed, although languages throughout the family reflect \rc{ə}.
+\obj{ɨ}-initial forms are predominant, so the vowel can tentatively be reconstructed, although languages throughout the family reflect \rc{ə}.
 Since many languages only show \obj{ɨ} in certain contexts, even forms like \trio \obj{wɨtənne} \qu{I went} \parencite[43]{triomeira1999} are ambiguous, since epenthetic \obj{ɨ} breaks up CC clusters on prefix-verb boundaries.
@@
 \rc{a[p]} is the older copula and already had various irregularities in \PC \parencite{gildea2018reconstructing}.
-\rc{eti} is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the existence of \obj{it͡ʃi} \qu{to lie down} in \arara \parencite[196]{alves2017arara}.}
+\rc{eti} is reconstructed by \textcites{meira2009property}{gildea2018reconstructing} as originally meaning \qu{to dwell, live}, but serving as a copula already in \PC.\footnote{Such a stative, locative source is also suggested by the meaning \qu{to lie down} of \arara \obj{it͡ʃi} \parencite[196]{alves2017arara}.}
 Reflexes of these roots are used suppletively, conditioned by person and\slash{}or \gl{tam}.
@@
 \waiwai \qu{to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently was an \gl{s_a_} verb already (although its reflexes in languages with split-\gl{s} mostly co-occur with reflexes of \rc{ət-epɨ}).
+However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently was an \gl{s_a_} verb already (although its reflexes in languages preserving the split-\gl{s} system mostly co-occur with reflexes of \rc{ət-epɨ}).
 

### 2023-08-04 review_answers.md
some small reviewer 2 fixes

@@
 	* There is nothing that could be “aligned” in this example.
-I interpreted it as "make the distance between object language and translation identical to example 4", and added translations for all wordforms in the paradigm.
+I interpreted the comment as "make the distance between object language and translation identical to example 4", and added translations for all wordforms in the paradigm.
 
@@
 
-* **Table 8**: *Asterisks are missing in PW forms.
-example 11: consider swapping (c) and (d)*
+* **Table 8**: *Asterisks are missing in PW forms.*
+
+	* TBD
+
+* **example 11**: *consider swapping (c) and (d)*
 
@@
 
-	* TBD
+	* I do not understand what is meant by this comment.
 
@@
 * **lines 498–499**: *“These verbs formed a lexical connection with prototypical members of the
-group, developing new past tense forms as a result” — did you mean ‘patient participle’?
-section 4: Please justify the choice of Apalaí texts for the frequency count — why not the
+group, developing new past tense forms as a result” — did you mean ‘patient participle’?*
+
+	* I did, thank you.
+
+* **section 4**: *Please justify the choice of Apalaí texts for the frequency count — why not the
 exceptionally-well described Kari’ña? It is true that there are no ready-to-use corpora for
 Cariban languages, but there are many glossed texts available. Another good start would be
-the Kuhikugu and Kalapalo texts in Stenzel & Franchetto (2017).
-section 4: In frequency counts, it should be made clear if one counts the frequency of
+the Kuhikugu and Kalapalo texts in Stenzel & Franchetto (2017).*
+
+	* TBD
+
+* **section 4**: *In frequency counts, it should be made clear if one counts the frequency of
 verbal lexemes or specifically of verbs inflected for 1st person (some verb may only be
@@
 
-	* TBD
+	* I only cite works that I reference in the text.
 
@@
 
-	* TBD
+	* Space removed.
 
@@
 	* TBD
-
-* **Abbreviations**: *uncertainty.9 → uncertainty.*
-
-	* TBD
-
-* **References**: *(Franchetto 1986): Estudo Etnolinguistico → Estudo etnolinguístico*
-
-	* TBD
-
-* **References**: *The use of capital letters in Spanish and Portuguese titles is inconsistent.*
-
-	* TBD

### 2023-08-04 documents/sections/introduction.tex
some small reviewer 2 fixes

@@
 The goal of this study is to approach these patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows:
-In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns like in \cref{tab:hixintro,tab:triintro} .
+In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns like in \cref{tab:hixintro,tab:triintro}.
 In \cref{sec:data}, six incomplete person marker extensions and the verbs unaffected by them are described.

### 2023-08-04 documents/sections/verbs.tex
some small reviewer 2 fixes

@@
 The reconstruction of \PPek \rc{ɨptə} treats the additional elements in daughter languages as non-cognate.
-\textcite{meira2005southern} identify no correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}, and at least the addition of a final \obj{ŋ} in \PXin is attested elsewhere:
+\textcite{meira2005southern} identify no correspondence between \bakairi \obj{gɨ} and \ikpeng \obj{ŋ}, and at least the addition of a final \rc{ŋ} in \PXin is attested elsewhere:
 		\begin{inlinelist}

### 2023-08-07 documents/sections/extensions.tex
ikpeng 0-iptong-, fixes #3
fixes #8

@@
 However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}, with the exception of \obj{k-}prefixed \qu{to go} \exref{ikp-170}.
-There is a formally identical \ikpeng cognate of \arara \obj{iptoŋ} \qu{to go down}, but no first person forms are attested \perscommpar{Angela Chagas}.
 While reflexes of \rc{a[p]} \qu{be-1} do exist in \ikpeng, apparently only reflexes of \rc{eti} \qu{be-2} occur with first person prefixes \parencite[401]{gildea2018reconstructing}.
@@
 
-The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} of course also resisted the extension of \obj{k-} in \akuriyo, although the first-person form of the copular verb \obj{eʔi} is unknown.
+The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} of course also resisted the extension of \obj{k-} in \akuriyo.
 In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-}, apparently reconstructible to \PTir \exref{tridown}.

### 2023-08-11 documents/sections/proto_carib.tex
fixes #6

@@
 %
-\pex<kar2> \kalina
-\a<kar-59>
-\begingl
+\ex Kari'ña    \a \label{kar2}
+\begingl \glpreamble sipi tynkary mekemanon, hen? //
 \gla sipi tɨnka-rɨ m-ekema-non hen//
-\glb net pull-\gl{nmlz} \gl{2}\gl{s_a_}-be.afraid-\gl{prs}.\gl{uncert} eh?//
-\glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
+\glb net pull-\gl{nmlz} \gl{2}\gl{s\_a\_}-be.afraid-\gl{prs}.\gl{uncert} eh?//
+\glft ‘You're afraid to pull up the net, aren't you?’ (\cite[253]{courtz2008carib}
+)//
 \endgl
-\a<kar-61>
-\begingl
-\glpreamble aya:woiya //
+
+\a \label{kar2}
+\begingl \glpreamble aya:woiya //
 \gla aj-awomɨ-ja//
-\glb \gl{2}\gl{s_p_}-get.up-\gl{prs}//
-\glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
+\glb \gl{2}\gl{s\_p\_}-get.up-\gl{prs}//
+\glft ‘You are getting up.’ (\cite[167]{hoff1968carib}
+)//
 \endgl
+
 \xe

### 2023-08-16 documents/sections/extensions.tex
fixes #6
fixes #1

@@
 
-In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two \gl{s_a_} subclasses, defined by \gl{1}\gl{s_a_} \obj{w-} and \obj{k-}, respectively.\footnote{\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
-	However, this marker is only illustrated for \qu{to bathe}, both by \textcite{meira2003bakairi} and \textcite{von1892bakairi}.
-	Given the lack of data for other verbs, this potential additional pattern will not be further discussed.
-	If the characterization by \citeauthor{meira2003bakairi} is accurate, then verbs with innovative first person prefixes have conservative \gl{1+2} prefixes, and vice versa.
+In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two \gl{s_a_} subclasses, defined by \gl{1}\gl{s_a_} \obj{w-} and \obj{k-}, respectively.\footnote{
+	\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
+	However, this marker only occurs on \obj{i} \qu{to bathe} and \obj{i} \qu{be-2}; the latter is also attested with \obj{kɨd-} \parencite[112]{pinto2022bakairi}; its cognacy is uncertain.
 }
@@
 \begingl
-\gla wi-toka//
+\gla i-toka//
 \glb \gl{1}>\gl{3}-hit//

### 2023-08-16 workflow/compile_cldf_irregular.py
bringing table generation up to speed

@@ -1,16 +1,16 @@
-import pycldf
-import pandas as pd
-import re
-from segments import Profile, Tokenizer
-from cldfbench.cldf import CLDFWriter
-from cldfbench import CLDFSpec
-import pybtex
 import os
-from pycldf.sources import Source
-import lingpy
-from writio import load
-
+import re
 from pathlib import Path
+
+import lingpy
+import pandas as pd
+import pybtex
+import pycldf
+from cldfbench import CLDFSpec
+from cldfbench.cldf import CLDFWriter
+from pycldf.sources import Source
+from segments import Profile, Tokenizer
+from writio import load
 
 data_dir = Path("../data")
 import logging
@@ -288,7 +288,12 @@ with CLDFWriter(spec) as writer:
         "CognateTable": cognates,
         "ExampleTable": examples,
     }.items():
-        for sepcol, sep in {"Source": "; ", "Alignment": " ", "Analyzed_Word": "\t", "Gloss": "\t"}.items():
+        for sepcol, sep in {
+            "Source": "; ",
+            "Alignment": " ",
+            "Analyzed_Word": "\t",
+            "Gloss": "\t",
+        }.items():
             if sepcol in df.columns:
                 df[sepcol] = df[sepcol].apply(lambda x: x.split(sep))
         for rec in df.to_dict("records"):
@@ -298,7 +303,7 @@ with CLDFWriter(spec) as writer:
                     found_refs.append(s.split("[")[0])
 
     found_refs = list(set(found_refs))
-    bib = pybtex.database.parse_file(data_dir/"cariban.bib")
+    bib = pybtex.database.parse_file(data_dir / "cariban.bib")
     sources = [
         Source.from_entry(k, e) for k, e in bib.entries.items() if k in found_refs
     ]
@@ -310,7 +315,6 @@ with CLDFWriter(spec) as writer:
     )
     sources.append(Source.from_entry("pc", pc))
     writer.cldf.add_sources(*sources)
-
 
     writer.write()
     ds = writer.cldf

### 2023-08-16 workflow/helper_functions.py

@@ -1,420 +1,0 @@
-from read_data import *
-import re
-import pyradigms as pyd
-import cldf_helpers as cldfh
-import warnings
-import json
-import numpy as np
-from lingpy_alignments import calculate_alignment
-
-warnings.filterwarnings("ignore")
-
-
-def print_shorthand(abbrev):
-    return "\\" + cah.get_shorthand(abbrev)
-
-
-shorthand_dic = {x: print_shorthand(x) for x in lg_list}
-
-
-def str2numcog(cogsets):
-    return " ".join([cognum[x] for x in cogsets.split("+")])
-
-
-def num2strcog(cogids):
-    return "+".join([numcog[i] for i in cogids.split(" ")])
-
-
-def segmentify(form):
-    form = re.sub("[()\[\]]", "", form)
-    form = form.replace("-", "+")
-    form = form.strip("+")
-    return t(form)
-
-
-# functions for creating latex tables
-def print_latex(
-    df,
-    ex=False,
-    keep_index=False,
-    formatters=None,
-    multicolumn=False,
-    float_format=None,
-):
-    if keep_index:
-        df.columns.name = df.index.name
-        df.index.name = None
-
-    with pd.option_context("max_colwidth", 1000):
-        lines = df.to_latex(
-            escape=False,
-            index=keep_index,
-            formatters=formatters,
-            multicolumn=multicolumn,
-            float_format=float_format,
-        ).split("\n")
-    lines[0] = lines[0].replace("tabular}{l", "tabular}[t]{@{}l").replace("l}", "l@{}}")
-    if ex:
-        del lines[1:4]
-        del lines[-1]
-        del lines[-2]
-        return "\n".join(lines)
-    else:
-        return (
-            "\n".join(lines)
-            .replace(r"\toprule", "\mytoprule")
-            .replace(r"\midrule", "\mymidrule")
-            .replace(r"\bottomrule", "\mybottomrule")
-            .replace("%", "\\%")
-        )
-
-
-def save_float(tabular, label, caption, filename=None, short=None):
-    if not filename:
-        filename = label
-    if not short:
-        shorttext = ""
-    else:
-        shorttext = f"[{short}]"
-    output = f"""\\begin{{table}}
-\centering
-\caption{shorttext}{{{caption}}}
-\label{{tab:{label}}}
-{tabular}\end{{table}}"""
-    f = open(f"../documents/floats/{filename}.tex", "w")
-    f.write(output)
-    f.close()
[…]

### 2023-08-16 workflow/lingpy_alignments.py

@@ -1,8 +1,9 @@
-import lingpy
-from lingpy.compare.partial import Partial
-import pandas as pd
 import os
 from difflib import SequenceMatcher
+
+import lingpy
+import pandas as pd
+from lingpy.compare.partial import Partial
 
 tempfile = "lingpy_temp.tsv"
 

### 2023-08-16 workflow/process_data.py

@@ -1,22 +1,48 @@
-from helper_functions import *
+import pyradigms
+from pylingdocs.helpers import decorate_gloss_string
 
-verb_list = ["say", "go", "be_1", "be_2", "come", "go_down", "bathe_intr"]
-person = ["1", "2", "1+2", "3"]
+from helpers import (
+    export_csv,
+    get_sources,
+    infl_data,
+    mean_dic,
+    name_dic,
+    objectify,
+    print_latex,
+    print_shorthand,
+    repl_latex,
+    save_float,
+    trans_dic,
+)
+
+verb_list = [
+    "say",
+    "go",
+    "be_1",
+    "be_2",
+    "come",
+    "go_down",
+    "bathe_intr",
+]  # relevant verb cognates
+person = ["1", "2", "1+2", "3"]  # relevant person values
+
 
 # tables for the introduction, featuring "regular" and "irregular" verbs from Trio and Hixkaryana
-pyd.x = ["Meaning_ID"]
-pyd.y = ["Inflection"]
-pyd.y_sort = person
 for lg, meanings in {
     "hix": ["fall", "be_afraid", "walk", "cut_self", "be"],
     "tri": ["sleep", "see_self", "bathe_intr", "yawn", "go"],
     # "mak": ["eat", "arrive", "go", "be"]
     # "yuk": ["wash_self", "sleep", "fall"],
 }.items():
-    pyd.filters = {"Language_ID": [lg], "Meaning_ID": meanings, "Inflection": person}
-    pyd.x_sort = meanings
+    pyd = pyradigms.Pyradigm(
+        infl_data[infl_data["Verb_Cognateset_ID"] != "be_1"],
+        x="Meaning_ID",
+        y="Inflection",
+        filters={"Language_ID": [lg], "Meaning_ID": meanings, "Inflection": person},
+        sort_orders={"Inflection": person, "Meaning_ID": meanings},
+    )
     tabular = pyd.compose_paradigm(
-        i_df[i_df["Verb_Cognateset_ID"] != "be_1"], multi_index=False
+        multi_index=False,
     )
     label = lg + "intro"
     tabular_raw = tabular.rename(columns=mean_dic)
@@ -26,10 +52,10 @@ for lg, meanings in {
         label,
         f"Some {name_dic[lg]} verbs",
         keep_index=True,
[…]

### 2023-08-16 workflow/read_data.py

@@ -1,9 +1,9 @@
-import cariban_helpers as cah
+import karipona as cah
 import pandas as pd
 
 pd.options.mode.chained_assignment = None
 import pynterlinear as pynt
-from segments import Tokenizer, Profile
+from segments import Profile, Tokenizer
 
 lg_list = list(cah.lg_order().keys())
 
@@ -12,8 +12,8 @@ cs_df = pd.read_csv("../data/cognate_sets.csv")
 v_df = pd.read_csv("../data/verb_stem_data.csv")
 v_df.rename(columns={"Meaning_ID": "Parameter_ID"}, inplace=True)
 e_df = pd.read_csv("../data/extensions.csv")
-i_df = pd.read_csv("../data/inflection_data.csv")
-i_df["Form"] = i_df["Form"].str.replace("+", "", regex=True)
+infl_data = pd.read_csv("../data/inflection_data.csv")
+infl_data["Form"] = infl_data["Form"].str.replace("+", "", regex=True)
 
 m_df = pd.read_csv("../data/cldf/parameters.csv")
 m_df["Description"] = m_df["Description"].str.strip("'").map(pynt.get_expex_code)
@@ -34,7 +34,7 @@ concept_dic = {
     "DETRZ": "come",
 }
 
-i_df["Concept"] = i_df.apply(
+infl_data["Concept"] = infl_data.apply(
     lambda x: concept_dic[x["Verb_Cognateset_ID"]]
     if x["Verb_Cognateset_ID"] in concept_dic
     else x["Verb_Cognateset_ID"],

### 2023-08-21 documents/floats/ppekverbs.tex
more data processing

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[76, 130-131, 374, 397]{von1892bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[129, 170]{meiraDBebakairi / pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[111-112]{pinto2022bakairi}{meiraDBebakairi}}
 \label{tab:ppekverbs}
@@
 \mytoprule
-{} &          \PPek &          \arara &                       \ikpeng &        \bakairi \\
+{} &          \PPek &          \arara &                       \ikpeng &             \ebakairi \\
 \mymidrule
-\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &     \obj{u-ge-} \\
-\qu{go}      &   \rc{w-ɨtən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &     \obj{u-tə-} \\
-\qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &      \obj{w-a-} \\
+\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &  \obj{u-gue- / u-ge-} \\
+\qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &            \obj{w-i-} \\
+\qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &          \obj{w-awɨ-} \\
 \qu{be-2}    &  \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &            \obj{w-i-} \\
 \qu{come}    &    \rc{w-epɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &         \obj{k-əewɨ-} \\
-\qu{go down} &   \rc{w-ɨptə-} &  \obj{w-iptoŋ-} &  \obj{{\normalfont ?}-iptoŋ-} &  \obj{k-ɨtəgɨ-} \\
-\qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &      \obj{w-i-} \\
+\qu{go down} &   \rc{w-ɨptə-} &                 &                               &                       \\
+\qu{go}      &   \rc{w-ɨdən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &           \obj{u-də-} \\
 \mybottomrule

### 2023-08-21 documents/floats/pwaireg.tex
more data processing

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[189-190, 196]{hixkaryanaderby1985}[209, 211]{hawkins1953waiwai}[150]{howard2001wrought}[30]{waiwaihawkins1998}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189-190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2023-08-21 documents/floats/pwaiverbs.tex
more data processing

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197-198; p.c., Spike Gildea]{hixkaryanaderby1985}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197-198]{hixkaryanaderby1985}[Spike Gildea]{pc}}
 \label{tab:pwaiverbs}
@@
 \qu{say}  &   \rc{wɨ-ka-} &   \obj{ɨ-ka-} &    \obj{wɨ-ka-} \\
-\qu{go}   &  \rc{wɨ-tom-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
 \qu{be-1} &    \rc{w-ah-} &   \obj{w-ah-} &      \obj{w-a-} \\
 \qu{be-2} &   \rc{w-eʃi-} &  \obj{w-eʃe-} &   \obj{w-eeʃi-} \\
+\qu{go}   &  \rc{wɨ-tom-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
 \mybottomrule

### 2023-08-21 workflow/process_data.py

@@ -1,17 +1,24 @@
+import pandas as pd
 import pyradigms
 from pylingdocs.helpers import decorate_gloss_string
 
 from helpers import (
+    combine_sources,
     export_csv,
+    get_obj_str,
     get_sources,
     infl_data,
     mean_dic,
     name_dic,
     objectify,
+    plain_trans_dic,
     print_latex,
     print_shorthand,
+    reconstructed_form_table,
     repl_latex,
     save_float,
+    shorthand_dic,
+    src,
     trans_dic,
 )
 
@@ -67,19 +74,21 @@ for lg, meanings in {
         short=f"Some {print_shorthand(lg)} verbs",
     )
 
-# regular pekodian verbs
-tabular = pd.DataFrame()
+
+# regular pekodian sa verbs
+entries = []
 sources_list = []
-pyd.x = ["Language_ID", "Meaning_ID"]
+pyd = pyradigms.Pyradigm(infl_data, x=["Language_ID", "Meaning_ID"], y="Inflection")
 # no reconstructable paradigms, so we use regular Sa verbs
-for lg, meaning in {"bak": "go_up", "ara": "dance", "ikp": "run"}.items():
+for lg, meaning in {"ebak": "go_up", "ara": "dance", "ikp": "run"}.items():
     pyd.filters = {"Language_ID": [lg], "Meaning_ID": [meaning]}
     pyd.sort_orders = {"Inflection": person}
-    temp = pyd.compose_paradigm(infl_data, multi_index=True)
-    sources_list.extend(get_sources(infl_data, latexify=False))
-    tabular = pd.concat([tabular, temp], axis=1)
+    temp = pyd.compose_paradigm(with_multi_index=True)
+    entries.append(temp)
+    sources_list.extend(get_sources(pyd, combine=False))
 
-sources = cldfh.cite_a_bunch(sources_list, parens=True)
+tabular = pd.concat(entries, axis=1)
+
 label = "pekreg"
 tabular.index.names = [None]
 export_csv(
@@ -92,22 +101,25 @@ export_csv(
     sources=sources_list,
 )
 
+
 tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name)))
 tabular.rename(columns=trans_dic, level="Meaning_ID", inplace=True)
[…]

### 2023-08-23 documents/floats/aku1sa.tex
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
a thousand small fixes
small fixes
godown

@@
 \centering
-\caption[Regular \akuriyo \gl{1}\gl{s_a_} markers]{Regular \akuriyo \gl{1}\gl{s_a_} markers \parencites[77, 79, 82, 84, 85, 86, 87]{gildea1994akuriyo}}
+\caption[Regular \akuriyo \gl{1}\gl{s_a_} markers]{Regular \akuriyo \gl{1}\gl{s_a_} markers \parencites[77, 79, 82, 84-87]{gildea1994akuriyo}}
 \label{tab:aku1sa}
@@
       first person \obj{k-} &               first person \obj{t͡ʃ-} \\
-\midrule
- \obj{əempa-} \qu{to learn} &  \obj{epɨ-} \qu{to bathe (\gl{intr})} \\
-\obj{əət͡ʃena-} \qu{to cry} &      \obj{ekɨrɨka-} \qu{to stay back} \\
-\obj{əiwa-} \qu{to tremble} &            \obj{entapo-} \qu{to yawn} \\
- \obj{əməmɨ-} \qu{to enter} &       \obj{etonema-} \qu{to lie down} \\
-\obj{ətajiŋka-} \qu{to run} &          \obj{ewai-} \qu{to sit down} \\
-  \obj{əturu-} \qu{to talk} & \obj{ehpa-} \qu{to bathe (\gl{intr})} \\
+\mymidrule
+ \obj{əempa-} \qu{to learn} &            \obj{entapo-} \qu{to yawn} \\
+\obj{əiwa-} \qu{to tremble} &      \obj{ekɨrɨka-} \qu{to stay back} \\
+\obj{ətajiŋka-} \qu{to run} &       \obj{etonema-} \qu{to lie down} \\
+\obj{əət͡ʃena-} \qu{to cry} &          \obj{ewai-} \qu{to sit down} \\
+  \obj{əturu-} \qu{to talk} &  \obj{epɨ-} \qu{to bathe (\gl{intr})} \\
+ \obj{əməmɨ-} \qu{to enter} & \obj{ehpa-} \qu{to bathe (\gl{intr})} \\
 \obj{əənɨkɨ-} \qu{to sleep} &                                       \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2023-08-23 documents/floats/ppekverbs.tex
godown

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[129, 170]{meiraDBebakairi / pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[111-112]{pinto2022bakairi}{meiraDBebakairi}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[129, 170]{meiraDBebakairi / pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[111]{pinto2022bakairi}{meiraDBebakairi}}
 \label{tab:ppekverbs}

### 2023-08-23 documents/floats/ptirverbs.tex
godown

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-115, 165]{meira1998proto}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-114]{meira1998proto}{meiraDBakuriyo}}
 \label{tab:ptirverbs}
@@
 \mymidrule
+\qu{go}   &  \rc{w-ɨtəmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
 \qu{say}  &    \rc{wɨ-ka-} &   \obj{wɨ-ka-} &                 \obj{wɨ-ka-} \\
-\qu{go}   &  \rc{w-ɨtəmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
+\qu{come} &  \rc{w-əʔepɨ-} &  \obj{w-əepɨ-} &  \obj{{\normalfont ∅}-eepɨ-} \\
 \qu{be-1} &      \rc{w-a-} &     \obj{w-a-} &     \obj{{\normalfont ∅}-a-} \\
-\qu{be-2} &    \rc{w-eʔi-} &    \obj{w-ei-} &   \obj{{\normalfont ?}-eʔi-} \\
-\qu{come} &  \rc{w-əʔepɨ-} &  \obj{w-əepɨ-} &  \obj{{\normalfont ∅}-eepɨ-} \\
+\qu{be-2} &    \rc{w-eʔi-} &    \obj{w-ei-} &   \obj{{\normalfont ∅}-eʔi-} \\
 \mybottomrule

### 2023-08-23 documents/floats/pwaireg.tex
godown

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189-190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2023-08-23 documents/floats/pwaiverbs.tex
godown

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197-198]{hixkaryanaderby1985}[Spike Gildea]{pc}}
+\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197]{hixkaryanaderby1985}[Spike Gildea]{pc}}
 \label{tab:pwaiverbs}

### 2023-08-23 documents/floats/pxinw.tex
regular verbs in data and docs
new approach to testing predictions from bybee's model
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
godown

@@
 \centering
-\caption[Loss of \rc{w} in \ikpeng]{Loss of \rc{w} in \ikpeng \parencites[44, 70]{souza1993arara}[118]{alves2013verbo}[143]{alves2017arara}[21, 164]{ikpengpacheco2001}[9]{desouza2010arara}[40]{campetela1997analise}}
+\caption[Loss of \rc{w} in \ikpeng]{Loss of \rc{w} in \ikpeng \parencites[44, 70]{souza1993arara}[118]{chagas2013verbo}[143]{alves2017arara}[21, 164]{ikpengpacheco2001}[9]{desouza2010arara}[40]{campetela1997analise}}
 \label{tab:pxinw}
 \begin{tabular}[t]{@{}lll@{}}
-\toprule
+\mytoprule
              Meaning &       \arara &     \ikpeng \\
-\midrule
+\mymidrule
     \qu{to defecate} &  \obj{watke} &  \obj{atke} \\
@@
 \qu{capuchin monkey} &   \obj{tawe} &   \obj{tae} \\
-          \qu{sleep} &  \obj{wɨnkɨ} &  \obj{ɨnkɨ} \\
-\bottomrule
+       \qu{to sleep} &  \obj{wɨnkɨ} &  \obj{ɨnkɨ} \\
+\mybottomrule
 \end{tabular}

### 2023-08-23 workflow/process_data.py

@@ -3,8 +3,10 @@ import pyradigms
 from pylingdocs.helpers import decorate_gloss_string
 
 from helpers import (
+    combine_form_meaning,
     combine_sources,
-    export_csv,
+    delatexify,
+    export_table,
     get_obj_str,
     get_sources,
     infl_data,
@@ -12,6 +14,7 @@ from helpers import (
     name_dic,
     objectify,
     plain_trans_dic,
+    sort_lg,
     print_latex,
     print_shorthand,
     reconstructed_form_table,
@@ -20,6 +23,9 @@ from helpers import (
     shorthand_dic,
     src,
     trans_dic,
+    v_df,
+    add_obj_markdown,
+    repl_lg_id
 )
 
 verb_list = [
@@ -38,8 +44,6 @@ person = ["1", "2", "1+2", "3"]  # relevant person values
 for lg, meanings in {
     "hix": ["fall", "be_afraid", "walk", "cut_self", "be"],
     "tri": ["sleep", "see_self", "bathe_intr", "yawn", "go"],
-    # "mak": ["eat", "arrive", "go", "be"]
-    # "yuk": ["wash_self", "sleep", "fall"],
 }.items():
     pyd = pyradigms.Pyradigm(
         infl_data[infl_data["Verb_Cognateset_ID"] != "be_1"],
@@ -48,32 +52,18 @@ for lg, meanings in {
         filters={"Language_ID": [lg], "Meaning_ID": meanings, "Inflection": person},
         sort_orders={"Inflection": person, "Meaning_ID": meanings},
     )
-    tabular = pyd.compose_paradigm(
-        multi_index=False,
-    )
-    label = lg + "intro"
-    tabular_raw = tabular.rename(columns=mean_dic)
-    tabular_raw.columns = [repl_latex(f"'{col}'") for col in tabular_raw.columns]
-    export_csv(
-        tabular_raw,
-        label,
-        f"Some {name_dic[lg]} verbs",
-        keep_index=True,
-        sources=get_sources(pyd, latexify=False),
-    )
-    sources = get_sources(pyd)
+    tabular = pyd.compose_paradigm()
     tabular.index = tabular.index.map(decorate_gloss_string)
     tabular.index.name = None
     tabular.rename(columns=trans_dic, inplace=True)
[…]

### 2023-08-23 workflow/read_data.py

@@ -5,6 +5,8 @@ pd.options.mode.chained_assignment = None
 import pynterlinear as pynt
 from segments import Profile, Tokenizer
 
+
+lg_dict = {x["Shorthand"]: x["Name"] for x in cah.languages.values()}
 lg_list = list(cah.lg_order().keys())
 
 cognate_list = ["go", "say", "come", "be_1", "be_2", "go_down", "bathe_intr"]

### 2023-08-23 documents/config/cariban_language_names.tex
more table fixes

@@
 \usepackage{xspace}
+\providecommand{\wbakairi}{Western Bakairi\xspace}
+\DeclareAcronym{wbak_lg}{short=wbak,long=Western Bakairi,short-format=\scshape}
+\providecommand{\ebakairi}{Eastern Bakairi\xspace}
+\DeclareAcronym{ebak_lg}{short=ebak,long=Eastern Bakairi,short-format=\scshape}
+\providecommand{\ingariko}{Ingarikó\xspace}
+\DeclareAcronym{ing_lg}{short=ing,long=Ingarikó,short-format=\scshape}
+\providecommand{\akawaio}{Akawaio\xspace}
+\DeclareAcronym{aka_lg}{short=aka,long=Akawaio,short-format=\scshape}
+\providecommand{\patamona}{Patamona\xspace}
+\DeclareAcronym{pat_lg}{short=pat,long=Patamona,short-format=\scshape}
+\providecommand{\tabajari}{Tabajari\xspace}
+\DeclareAcronym{tab_lg}{short=tab,long=Tabajari,short-format=\scshape}
+\providecommand{\gkar}{Guianese Carib\xspace}
+\DeclareAcronym{gkar_lg}{short=gkar,long=Guianese Carib,short-format=\scshape}
+\providecommand{\aretyry}{Aretyry\xspace}
+\DeclareAcronym{aty_lg}{short=aty,long=Aretyry,short-format=\scshape}
+\providecommand{\tyrewuju}{Tyrewuju\xspace}
+\DeclareAcronym{tyr_lg}{short=tyr,long=Tyrewuju,short-format=\scshape}
+\providecommand{\sikiana}{Sikiana\xspace}
+\DeclareAcronym{siku_lg}{short=siku,long=Sikiana,short-format=\scshape}
+\providecommand{\dekwana}{De’kwana\xspace}
+\DeclareAcronym{dek_lg}{short=dek,long=De’kwana,short-format=\scshape}
+\providecommand{\dhekwana}{Dhe’kwana\xspace}
+\DeclareAcronym{dhe_lg}{short=dhe,long=Dhe’kwana,short-format=\scshape}
+\providecommand{\kamara}{Kamarakoto\xspace}
+\DeclareAcronym{kam_lg}{short=kam,long=Kamarakoto,short-format=\scshape}
+\providecommand{\tauli}{Taulipang\xspace}
+\DeclareAcronym{tau_lg}{short=tau,long=Taulipang,short-format=\scshape}
+\providecommand{\arekuna}{Arekuna\xspace}
+\DeclareAcronym{are_lg}{short=are,long=Arekuna,short-format=\scshape}
+\providecommand{\nahukwa}{Nahukwa\xspace}
+\DeclareAcronym{nah_lg}{short=nah,long=Nahukwa,short-format=\scshape}
+\providecommand{\kuikuro}{Kuikuro\xspace}
+\DeclareAcronym{kui_lg}{short=kui,long=Kuikuro,short-format=\scshape}
+\providecommand{\matipu}{Matipu\xspace}
+\DeclareAcronym{mat_lg}{short=mat,long=Matipu,short-format=\scshape}
+\providecommand{\kalapalo}{Kalapalo\xspace}
+\DeclareAcronym{kal_lg}{short=kal,long=Kalapalo,short-format=\scshape}
+\providecommand{\yukpa}{Yukpa\xspace}
+\DeclareAcronym{yuk_lg}{short=yuk,long=Yukpa,short-format=\scshape}
+\providecommand{\japreria}{Japreria\xspace}
+\DeclareAcronym{jap_lg}{short=jap,long=Japreria,short-format=\scshape}
+\providecommand{\carijo}{Karijona\xspace}
+\DeclareAcronym{car_lg}{short=car,long=Karijona,short-format=\scshape}
+\providecommand{\mapoyo}{Mapoyo\xspace}
+\DeclareAcronym{map_lg}{short=map,long=Mapoyo,short-format=\scshape}
+\providecommand{\yawarana}{Yawarana\xspace}
+\DeclareAcronym{yab_lg}{short=yab,long=Yawarana,short-format=\scshape}
+\providecommand{\panare}{Panare\xspace}
+\DeclareAcronym{pan_lg}{short=pan,long=Panare,short-format=\scshape}
+\providecommand{\maqui}{Ye'kwana\xspace}
+\DeclareAcronym{mak_lg}{short=mak,long=Ye'kwana,short-format=\scshape}
+\providecommand{\pemon}{Pemón\xspace}
+\DeclareAcronym{pem_lg}{short=pem,long=Pemón,short-format=\scshape}
+\providecommand{\kapon}{Kapong\xspace}
+\DeclareAcronym{kap_lg}{short=kap,long=Kapong,short-format=\scshape}
+\providecommand{\waimiri}{Waimiri-Atroari\xspace}
+\DeclareAcronym{wmr_lg}{short=wmr,long=Waimiri-Atroari,short-format=\scshape}
+\providecommand{\macushi}{Macushi\xspace}
+\DeclareAcronym{mac_lg}{short=mac,long=Macushi,short-format=\scshape}
+\providecommand{\waiwai}{Waiwai\xspace}
+\DeclareAcronym{wai_lg}{short=wai,long=Waiwai,short-format=\scshape}
+\providecommand{\hixka}{Hixkaryána\xspace}
+\DeclareAcronym{hix_lg}{short=hix,long=Hixkaryána,short-format=\scshape}
+\providecommand{\kaxui}{Werikyana\xspace}
+\DeclareAcronym{kax_lg}{short=kax,long=Werikyana,short-format=\scshape}
+\providecommand{\trio}{Tiriyó\xspace}
+\DeclareAcronym{tri_lg}{short=tri,long=Tiriyó,short-format=\scshape}
+\providecommand{\akuriyo}{Akuriyó\xspace}
+\DeclareAcronym{aku_lg}{short=aku,long=Akuriyó,short-format=\scshape}
 \providecommand{\apalai}{Apalaí\xspace}
 \DeclareAcronym{apa_lg}{short=apa,long=Apalaí,short-format=\scshape}
-\providecommand{\trio}{Tiriyó\xspace}
-\DeclareAcronym{tri_lg}{short=tri,long=Tiriyó,short-format=\scshape}
-\providecommand{\hixka}{Hixkaryána\xspace}
-\DeclareAcronym{hix_lg}{short=hix,long=Hixkaryána,short-format=\scshape}
-\providecommand{\panare}{Panare\xspace}
-\DeclareAcronym{pan_lg}{short=pan,long=Panare,short-format=\scshape}
-\providecommand{\macushi}{Macushi\xspace}
-\DeclareAcronym{mac_lg}{short=mac,long=Macushi,short-format=\scshape}
-\providecommand{\waimiri}{Waimiri-Atroari\xspace}
-\DeclareAcronym{wmr_lg}{short=wmr,long=Waimiri-Atroari,short-format=\scshape}
-\providecommand{\maqui}{Ye'kwana\xspace}
-\DeclareAcronym{mak_lg}{short=mak,long=Ye'kwana,short-format=\scshape}
-\providecommand{\waiwai}{Waiwai\xspace}
-\DeclareAcronym{wai_lg}{short=wai,long=Waiwai,short-format=\scshape}
 \providecommand{\wayana}{Wayana\xspace}
 \DeclareAcronym{way_lg}{short=way,long=Wayana,short-format=\scshape}
+\providecommand{\ikpeng}{Ikpeng\xspace}
+\DeclareAcronym{ikp_lg}{short=ikp,long=Ikpeng,short-format=\scshape}
+\providecommand{\arara}{Arara\xspace}
+\DeclareAcronym{ara_lg}{short=ara,long=Arara,short-format=\scshape}
 \providecommand{\kalina}{Kari'ña\xspace}
 \DeclareAcronym{kar_lg}{short=kar,long=Kari'ña,short-format=\scshape}
-\providecommand{\ikpeng}{Ikpeng\xspace}
-\DeclareAcronym{ikp_lg}{short=ikp,long=Ikpeng,short-format=\scshape}
 \providecommand{\uxc}{Upper Xingu Carib\xspace}
 \DeclareAcronym{uxc_lg}{short=uxc,long=Upper Xingu Carib,short-format=\scshape}
-\providecommand{\arara}{Arara\xspace}
-\DeclareAcronym{ara_lg}{short=ara,long=Arara,short-format=\scshape}
-\providecommand{\ingariko}{Ingarikó\xspace}
-\DeclareAcronym{ing_lg}{short=ing,long=Ingarikó,short-format=\scshape}
-\providecommand{\pemon}{Pemón\xspace}
-\DeclareAcronym{pem_lg}{short=pem,long=Pemón,short-format=\scshape}
-\providecommand{\kapon}{Kapón\xspace}
-\DeclareAcronym{kap_lg}{short=kap,long=Kapón,short-format=\scshape}
-\providecommand{\kalapalo}{Kalapalo\xspace}
-\DeclareAcronym{kal_lg}{short=kal,long=Kalapalo,short-format=\scshape}
-\providecommand{\kuikuro}{Kuikuro\xspace}
-\DeclareAcronym{kui_lg}{short=kui,long=Kuikuro,short-format=\scshape}
-\providecommand{\yukpa}{Yukpa\xspace}
-\DeclareAcronym{yuk_lg}{short=yuk,long=Yukpa,short-format=\scshape}
-\providecommand{\carijo}{Carijona\xspace}
-\DeclareAcronym{car_lg}{short=car,long=Carijona,short-format=\scshape}
 \providecommand{\bakairi}{Bakairi\xspace}
 \DeclareAcronym{bak_lg}{short=bak,long=Bakairi,short-format=\scshape}
-\providecommand{\tamanaku}{†Tamanaku\xspace}
-\DeclareAcronym{tam_lg}{short=tam,long=†Tamanaku,short-format=\scshape}
-\providecommand{\kaxui}{Werikyana\xspace}
-\DeclareAcronym{kax_lg}{short=kax,long=Werikyana,short-format=\scshape}
-\providecommand{\akuriyo}{Akuriyó\xspace}
-\DeclareAcronym{aku_lg}{short=aku,long=Akuriyó,short-format=\scshape}
-\providecommand{\japreria}{Japreria\xspace}
-\DeclareAcronym{jap_lg}{short=jap,long=Japreria,short-format=\scshape}
-\providecommand{\kumana}{†Kumaná\xspace}
-\DeclareAcronym{cum_lg}{short=cum,long=†Kumaná,short-format=\scshape}
-\providecommand{\palmela}{†Palmela\xspace}
-\DeclareAcronym{palm_lg}{short=palm,long=†Palmela,short-format=\scshape}
-\providecommand{\piment}{†Pimenteira\xspace}
-\DeclareAcronym{pime_lg}{short=pime,long=†Pimenteira,short-format=\scshape}
-\providecommand{\arakaju}{†Arakaju\xspace}
-\DeclareAcronym{arak_lg}{short=arak,long=†Arakaju,short-format=\scshape}
-\providecommand{\pauxiana}{†Pauxiana\xspace}
-\DeclareAcronym{pxan_lg}{short=pxan,long=†Pauxiana,short-format=\scshape}
-\providecommand{\paravi}{†Paravilhana\xspace}
-\DeclareAcronym{para_lg}{short=para,long=†Paravilhana,short-format=\scshape}
-\providecommand{\bonari}{†Bonari\xspace}
-\DeclareAcronym{bona_lg}{short=bona,long=†Bonari,short-format=\scshape}
-\providecommand{\wayu}{†Wayumara\xspace}
-\DeclareAcronym{wayu_lg}{short=wayu,long=†Wayumara,short-format=\scshape}
-\providecommand{\apiaka}{†Apiaka-Tocantins\xspace}
-\DeclareAcronym{apia_lg}{short=apia,long=†Apiaka-Tocantins,short-format=\scshape}
-\providecommand{\opon}{†Opón-Carare\xspace}
-\DeclareAcronym{opon_lg}{short=opon,long=†Opón-Carare,short-format=\scshape}
-\providecommand{\pauxi}{†Pauxi\xspace}
-\DeclareAcronym{paux_lg}{short=paux,long=†Pauxi,short-format=\scshape}
-\providecommand{\sapara}{†Sapara\xspace}
-\DeclareAcronym{sapa_lg}{short=sapa,long=†Sapara,short-format=\scshape}
-\providecommand{\tiveri}{†Tiverighotto\xspace}
-\DeclareAcronym{tive_lg}{short=tive,long=†Tiverighotto,short-format=\scshape}
-\providecommand{\yao}{†Yao\xspace}
-\DeclareAcronym{yao_lg}{short=yao,long=†Yao,short-format=\scshape}
-\providecommand{\yaruma}{†Yarumá\xspace}
-\DeclareAcronym{yaru_lg}{short=yaru,long=†Yarumá,short-format=\scshape}
-\providecommand{\koyaima}{†Koyaima\xspace}
-\DeclareAcronym{koya_lg}{short=koya,long=†Koyaima,short-format=\scshape}
 \providecommand{\hiana}{†Hianakoto-Umawa\xspace}
 \DeclareAcronym{hian_lg}{short=hian,long=†Hianakoto-Umawa,short-format=\scshape}
+\providecommand{\chayma}{†Chayma\xspace}
+\DeclareAcronym{cha_lg}{short=cha,long=†Chayma,short-format=\scshape}
+\providecommand{\kumanagoto}{†Kumanagoto\xspace}
+\DeclareAcronym{cgt_lg}{short=cgt,long=†Kumanagoto,short-format=\scshape}
+\providecommand{\arina}{†Arinakoto\xspace}
+\DeclareAcronym{arin_lg}{short=arin,long=†Arinakoto,short-format=\scshape}
 \providecommand{\saluma}{†Salumá\xspace}
 \DeclareAcronym{salu_lg}{short=salu,long=†Salumá,short-format=\scshape}
-\providecommand{\sikiana}{†Sikiana\xspace}
-\DeclareAcronym{siku_lg}{short=siku,long=†Sikiana,short-format=\scshape}
-\providecommand{\purukoto}{†Purukoto\xspace}
-\DeclareAcronym{puru_lg}{short=puru,long=†Purukoto,short-format=\scshape}
-\providecommand{\juma}{†Juma\xspace}
-\DeclareAcronym{juma_lg}{short=juma,long=†Juma,short-format=\scshape}
-\providecommand{\palenque}{†Palenque\xspace}
-\DeclareAcronym{pale_lg}{short=pale,long=†Palenque,short-format=\scshape}
-\providecommand{\parukoto}{†Parukoto\xspace}
-\DeclareAcronym{paru_lg}{short=paru,long=†Parukoto,short-format=\scshape}
 \providecommand{\urukue}{†Urukuena\xspace}
@@
 \DeclareAcronym{waya_lg}{short=waya,long=†Wayakulé,short-format=\scshape}
-\providecommand{\piritu}{†Píritu\xspace}
-\DeclareAcronym{piri_lg}{short=piri,long=†Píritu,short-format=\scshape}
-\providecommand{\guake}{†Guake\xspace}
-\DeclareAcronym{guak_lg}{short=guak,long=†Guake,short-format=\scshape}
+\providecommand{\pianakoto}{†Pianakoto\xspace}
+\DeclareAcronym{pian_lg}{short=pian,long=†Pianakoto,short-format=\scshape}
+\providecommand{\kumayena}{†Kumayena\xspace}
+\DeclareAcronym{kuma_lg}{short=kuma,long=†Kumayena,short-format=\scshape}
+\providecommand{\parukoto}{†Parukoto\xspace}
+\DeclareAcronym{paru_lg}{short=paru,long=†Parukoto,short-format=\scshape}
+\providecommand{\yawaperi}{†Jawaperi\xspace}
+\DeclareAcronym{jawa_lg}{short=jawa,long=†Jawaperi,short-format=\scshape}
 \providecommand{\macoita}{†Macoita\xspace}
@@
 \DeclareAcronym{chak_lg}{short=chak,long=†Chake,short-format=\scshape}
-\providecommand{\chayma}{†Chayma\xspace}
-\DeclareAcronym{chai_lg}{short=chai,long=†Chayma,short-format=\scshape}
+\providecommand{\koyaima}{†Koyaima\xspace}
+\DeclareAcronym{koya_lg}{short=koya,long=†Koyaima,short-format=\scshape}
+\providecommand{\opon}{†Opón-Carare\xspace}
+\DeclareAcronym{opon_lg}{short=opon,long=†Opón-Carare,short-format=\scshape}
+\providecommand{\guake}{†Guake\xspace}
+\DeclareAcronym{guak_lg}{short=guak,long=†Guake,short-format=\scshape}
+\providecommand{\pemono}{†Pémono\xspace}
+\DeclareAcronym{pno_lg}{short=pno,long=†Pémono,short-format=\scshape}
+\providecommand{\piritu}{†Píritu\xspace}
+\DeclareAcronym{piri_lg}{short=piri,long=†Píritu,short-format=\scshape}
+\providecommand{\tamanaku}{†Tamanaku\xspace}
+\DeclareAcronym{tam_lg}{short=tam,long=†Tamanaku,short-format=\scshape}
+\providecommand{\palenque}{†Palenque\xspace}
+\DeclareAcronym{pale_lg}{short=pale,long=†Palenque,short-format=\scshape}
 \providecommand{\avarikoto}{†Avarikoto\xspace}
 \DeclareAcronym{avar_lg}{short=avar,long=†Avarikoto,short-format=\scshape}
-\providecommand{\tauli}{Taulipang\xspace}
-\DeclareAcronym{tau_lg}{short=tau,long=Taulipang,short-format=\scshape}
-\providecommand{\patamona}{Patamona\xspace}
-\DeclareAcronym{pat_lg}{short=pat,long=Patamona,short-format=\scshape}
-\providecommand{\arekuna}{Arekuna\xspace}
-\DeclareAcronym{are_lg}{short=are,long=Arekuna,short-format=\scshape}
-\providecommand{\kamara}{Kamarakoto\xspace}
-\DeclareAcronym{kam_lg}{short=kam,long=Kamarakoto,short-format=\scshape}
-\providecommand{\akawaio}{Akawaio\xspace}
-\DeclareAcronym{aka_lg}{short=aka,long=Akawaio,short-format=\scshape}
-\providecommand{\kumanagoto}{†Kumanagoto\xspace}
-\DeclareAcronym{cgt_lg}{short=cgt,long=†Kumanagoto,short-format=\scshape}
+\providecommand{\kumana}{†Kumaná\xspace}
+\DeclareAcronym{cum_lg}{short=cum,long=†Kumaná,short-format=\scshape}
+\providecommand{\palmela}{†Palmela\xspace}
+\DeclareAcronym{palm_lg}{short=palm,long=†Palmela,short-format=\scshape}
+\providecommand{\pauxiana}{†Pauxiana\xspace}
+\DeclareAcronym{pxan_lg}{short=pxan,long=†Pauxiana,short-format=\scshape}
+\providecommand{\tiveri}{†Tiverighotto\xspace}
+\DeclareAcronym{tive_lg}{short=tive,long=†Tiverighotto,short-format=\scshape}
+\providecommand{\purukoto}{†Purukoto\xspace}
+\DeclareAcronym{puru_lg}{short=puru,long=†Purukoto,short-format=\scshape}
+\providecommand{\sapara}{†Sapara\xspace}
+\DeclareAcronym{sapa_lg}{short=sapa,long=†Sapara,short-format=\scshape}
+\providecommand{\yao}{†Yao\xspace}
+\DeclareAcronym{yao_lg}{short=yao,long=†Yao,short-format=\scshape}
+\providecommand{\wayu}{†Wayumara\xspace}
+\DeclareAcronym{wayu_lg}{short=wayu,long=†Wayumara,short-format=\scshape}
+\providecommand{\juma}{†Juma\xspace}
+\DeclareAcronym{juma_lg}{short=juma,long=†Juma,short-format=\scshape}
+\providecommand{\paravi}{†Paravilhana\xspace}
+\DeclareAcronym{para_lg}{short=para,long=†Paravilhana,short-format=\scshape}
 \providecommand{\crichana}{†Crichana\xspace}
 \DeclareAcronym{cric_lg}{short=cric,long=†Crichana,short-format=\scshape}
-\providecommand{\mapoyo}{Mapoyo\xspace}
-\DeclareAcronym{map_lg}{short=map,long=Mapoyo,short-format=\scshape}
-\providecommand{\yawarana}{Yawarana\xspace}
-\DeclareAcronym{yab_lg}{short=yab,long=Yawarana,short-format=\scshape}
-\providecommand{\yawaperi}{†Jawaperi\xspace}
-\DeclareAcronym{jawa_lg}{short=jawa,long=†Jawaperi,short-format=\scshape}
-\providecommand{\pemono}{†Pémono\xspace}
-\DeclareAcronym{pemo_lg}{short=pemo,long=†Pémono,short-format=\scshape}
-\providecommand{\pianakoto}{†Pianakoto\xspace}
-\DeclareAcronym{pian_lg}{short=pian,long=†Pianakoto,short-format=\scshape}
+\providecommand{\bonari}{†Bonari\xspace}
+\DeclareAcronym{bona_lg}{short=bona,long=†Bonari,short-format=\scshape}
+\providecommand{\pauxi}{†Pauxi\xspace}
+\DeclareAcronym{paux_lg}{short=paux,long=†Pauxi,short-format=\scshape}
+\providecommand{\yaruma}{†Yarumá\xspace}
+\DeclareAcronym{yaru_lg}{short=yaru,long=†Yarumá,short-format=\scshape}
+\providecommand{\arakaju}{†Arakaju\xspace}
+\DeclareAcronym{arak_lg}{short=arak,long=†Arakaju,short-format=\scshape}
+\providecommand{\apiaka}{†Apiaka-Tocantins\xspace}
+\DeclareAcronym{apia_lg}{short=apia,long=†Apiaka-Tocantins,short-format=\scshape}
+\providecommand{\piment}{†Pimenteira\xspace}
+\DeclareAcronym{pime_lg}{short=pime,long=†Pimenteira,short-format=\scshape}
 \providecommand{\PC}{Proto-Cariban\xspace}
-\DeclareAcronym{pc_lg}{short=pc,long=Proto-Cariban,short-format=\scshape}
-\providecommand{\arina}{†Arinakoto\xspace}
-\DeclareAcronym{arin_lg}{short=arin,long=†Arinakoto,short-format=\scshape}
-\providecommand{\kumayena}{†Kumayena\xspace}
-\DeclareAcronym{kuma_lg}{short=kuma,long=†Kumayena,short-format=\scshape}
-\providecommand{\tyrewuju}{Tyrewuju\xspace}
-\DeclareAcronym{tyr_lg}{short=tyr,long=Tyrewuju,short-format=\scshape}
-\providecommand{\dekwana}{De’kwana\xspace}
-\DeclareAcronym{dek_lg}{short=dek,long=De’kwana,short-format=\scshape}
-\providecommand{\dhekwana}{Dhe’kwana\xspace}
-\DeclareAcronym{dhe_lg}{short=dhe,long=Dhe’kwana,short-format=\scshape}
-\newcommand{\PPar}{Proto-Parukotoan\xspace}
-\newcommand{\PPars}{\textsc{PPar}\xspace}
-\newcommand{\PPem}{Proto-Pemongan\xspace}
-\newcommand{\PPems}{\textsc{PPem}\xspace}
-\newcommand{\PPek}{Proto-Pekodian\xspace}
-\newcommand{\PPeks}{\textsc{PPek}\xspace}
-\newcommand{\PPP}{Proto-Panare-Pemongan\xspace}
-\newcommand{\PPPs}{\textsc{PPP}\xspace}
-\newcommand{\PTir}{Proto-Tiriyoan\xspace}
-\newcommand{\PTirs}{\textsc{PTir}\xspace}
-\newcommand{\PWai}{Proto-Waiwaian\xspace}
-\newcommand{\PWais}{\textsc{PWai}\xspace}
-\newcommand{\PXin}{Proto-Xinguan\xspace}
-\newcommand{\PXins}{\textsc{PXin}\xspace}
-\newcommand{\PTar}{Proto-Taranoan\xspace}
-\newcommand{\PTars}{\textsc{PTar}\xspace}
-\newcommand{\yabmap}{\mapoyo-\yawarana}
-\newcommand{\yabmaps}{\textsc{yab-map}\xspace}
-\newcommand{\vzk}{Venezuelan \kalina}
-\newcommand{\vzks}{\textsc{vzk}\xspace}
-\newcommand{\esk}{Eastern Surinamese \kalina}
-\newcommand{\esks}{\textsc{esk}\xspace}
-\newcommand{\wsk}{Western Surinamese \kalina}
-\newcommand{\wsks}{\textsc{wsk}\xspace}
-\renewcommand{\lg}[1]{\acs*{#1_lg}}
+\DeclareAcronym{PC_lg}{short=PC,long=Proto-Cariban,short-format=\scshape}
+\providecommand{\PPar}{Proto-Parukotoan\xspace}
+\DeclareAcronym{PPar_lg}{short=PPar,long=Proto-Parukotoan,short-format=\scshape}
+\providecommand{\PWai}{Proto-Waiwaian\xspace}
+\DeclareAcronym{PWai_lg}{short=PWai,long=Proto-Waiwaian,short-format=\scshape}
+\providecommand{\PPek}{Proto-Pekodian\xspace}
+\DeclareAcronym{PPek_lg}{short=PPek,long=Proto-Pekodian,short-format=\scshape}
+\providecommand{\PXin}{Proto-Xinguan\xspace}
+\DeclareAcronym{PXin_lg}{short=PXin,long=Proto-Xinguan,short-format=\scshape}
+\providecommand{\PPP}{Proto-Pemong-Panare\xspace}
+\DeclareAcronym{PPP_lg}{short=PPP,long=Proto-Pemong-Panare,short-format=\scshape}
+\providecommand{\PPem}{Proto-Pemongan\xspace}
+\DeclareAcronym{PPem_lg}{short=PPem,long=Proto-Pemongan,short-format=\scshape}
+\providecommand{\PTar}{Proto-Taranoan\xspace}
+\DeclareAcronym{PTar_lg}{short=PTar,long=Proto-Taranoan,short-format=\scshape}
+\providecommand{\PTir}{Proto-Tiriyoan\xspace}
+\DeclareAcronym{PTir_lg}{short=PTir,long=Proto-Tiriyoan,short-format=\scshape}
+\providecommand{\PMan}{Proto-Manapiarean\xspace}
+\DeclareAcronym{PMan_lg}{short=PMan,long=Proto-Manapiarean,short-format=\scshape}
+\providecommand{\PVen}{Proto-Venezuelan\xspace}
+\DeclareAcronym{PVen_lg}{short=PVen,long=Proto-Venezuelan,short-format=\scshape}
+\providecommand{\PCum}{Proto-Cumana\xspace}
+\DeclareAcronym{PCum_lg}{short=PCum,long=Proto-Cumana,short-format=\scshape}

### 2023-08-23 documents/floats/comepara.tex
more table fixes

@@
 \centering
-\caption[\rc{(ət-)epɨ} \qu{to come} in paradigms]{\rc{(ət-)epɨ} \qu{to come} in paradigms \parencites[113, 150, 153, 156]{alves2017arara}[294; p.c., Spike Gildea]{triomeira1999}}
+\caption[\rc{(ət-)epɨ} \qu{to come} in paradigms]{\rc{(ət-)epɨ} \qu{to come} in paradigms \parencites[113, 150, 153, 156]{alves2017arara}[294]{triomeira1999}[Spike Gildea]{pc}}
 \label{tab:comepara}
@@
 \mytoprule
-{} &          \kaxui &          \arara &          \trio \\
+{} &          \arara &          \trio &          \kaxui \\
 \mymidrule
-\gl{1}   &   \obj{w-oohɨ-} &    \obj{w-ebɨ-} &  \obj{w-əepɨ-} \\
-\gl{2}   &   \obj{m-oohɨ-} &  \obj{m-odebɨ-} &  \obj{mən-epɨ} \\
-\gl{1+2} &  \obj{kɨs-ohɨ-} &  \obj{kud-ebɨ-} &   \obj{ke-epɨ} \\
-\gl{3}   &    \obj{n-ehɨ-} &    \obj{t-ebɨ-} &    \obj{n-epɨ} \\
+\gl{1}   &    \obj{w-ebɨ-} &  \obj{w-əepɨ-} &   \obj{w-oohɨ-} \\
+\gl{2}   &  \obj{m-odebɨ-} &  \obj{mən-epɨ} &   \obj{m-oohɨ-} \\
+\gl{1+2} &  \obj{kud-ebɨ-} &   \obj{ke-epɨ} &  \obj{kɨs-ohɨ-} \\
+\gl{3}   &    \obj{t-ebɨ-} &    \obj{n-epɨ} &    \obj{n-ehɨ-} \\
 \mybottomrule

### 2023-08-23 documents/floats/go.tex
more table fixes

--- 2021-11-01 documents/floats/go.tex
@@
 \centering
-\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6]{franchetto2002kuikuro}[99; p.c., Spike Gildea]{camargo2002lexico}}
+\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[80, 153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[5]{guerrero2016karihona}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[Spike Gildea]{pc}[6]{franchetto2002kuikuro}[99]{camargo2002lexico}[14]{lacerda1991waimiri}}
 \label{tab:go}
-\begin{tabular}[t]{@{}lllllll@{}}
+\begin{tabular}[t]{@{}ll@{}}
 \mytoprule
-Language & \multicolumn{6}{l}{Form} \\
-\midrule
-\kaxui    &     \obj{to[mo]} &    &  t &  o &  m &  o \\
-\hixka    &      \obj{[ɨ]to} &  ɨ &  t &  o &    &    \\
-\waiwai   &   \obj{[e]to[m]} &  e &  t &  o &  m &    \\
-\arara    &        \obj{ɨdo} &  ɨ &  d &  o &    &    \\
-\ikpeng   &       \obj{aran} &  a &  r &  a &  n &    \\
-\ikpeng   &        \obj{ero} &  e &  r &  o &    &    \\
-\bakairi  &      \obj{[ɨ]tə} &  ɨ &  t &  ə &    &    \\
-\trio     &      \obj{tə[n]} &    &  t &  ə &  n &    \\
-\akuriyo  &  \obj{[ə]tə[mɨ]} &  ə &  t &  ə &  m &  ɨ \\
-\carijo   &       \obj{təmə} &    &  t &  ə &  m &  ə \\
-\wayana   &   \obj{[ɨ]tə[m]} &  ɨ &  t &  ə &  m &    \\
-\apalai   &        \obj{ɨto} &  ɨ &  t &  o &    &    \\
-\kalina   &         \obj{to} &    &  t &  o &    &    \\
-\kalina   &       \obj{[ɨ]ʔ} &  ɨ &  ʔ &    &    &    \\
-\maqui    &    \obj{ɨtə[mə]} &  ɨ &  t &  ə &  m &  ə \\
-\ingariko &        \obj{ətə} &  ə &  t &  ə &    &    \\
-\pemon    &      \obj{[e]tə} &  e &  t &  ə &    &    \\
-\macushi  &      \obj{[a]tɨ} &  a &  t &  ɨ &    &    \\
-\panare   &      \obj{tə[n]} &    &  t &  ə &  n &    \\
-\yawarana &         \obj{tə} &    &  t &  ə &    &    \\
-\mapoyo   &         \obj{tə} &    &  t &  ə &    &    \\
-\uxc      &      \obj{[e]te} &  e &  t &  e &    &    \\
-\yukpa    &         \obj{to} &    &  t &  o &    &    \\
-\bottomrule
+Language &             Form \\
+\mymidrule
+\kaxui    &     \obj{to[mo]} \\
+\PWai     &    \rc{[ɨ]to[m]} \\
+\hixka    &      \obj{[ɨ]to} \\
+\waiwai   &   \obj{[e]to[m]} \\
+\PPek     &      \rc{ɨtə[n]} \\
+\bakairi  &      \obj{[ɨ]tə} \\
+\arara    &         \obj{to} \\
+\arara    &        \obj{ɨdo} \\
+\ikpeng   &       \obj{aran} \\
+\ikpeng   &        \obj{ero} \\
+\carijo   &     \obj{tə[mə]} \\
+\PTir     &   \rc{[ə]tə[mɨ]} \\
+\trio     &      \obj{tə[n]} \\
+\akuriyo  &  \obj{[ə]tə[mɨ]} \\
+\wayana   &   \obj{[ɨ]tə[m]} \\
+\apalai   &        \obj{ɨto} \\
+\kalina   &       \obj{[ɨ]ʔ} \\
+\kalina   &         \obj{to} \\
+\maqui    &    \obj{ɨtə[mə]} \\
+\ingariko &        \obj{ətə} \\
+\pemon    &      \obj{[e]tə} \\
+\macushi  &      \obj{[a]tɨ} \\
+\panare   &      \obj{tə[n]} \\
+\yawarana &         \obj{tə} \\
+\mapoyo   &         \obj{tə} \\
+\uxc      &      \obj{[e]te} \\
+\yukpa    &         \obj{to} \\
+\waimiri  &         \obj{ɨʔ} \\
+\mybottomrule
 \end{tabular}

### 2023-08-23 documents/floats/pwaireg.tex
more table fixes

--- 2023-08-21 documents/floats/pwaireg.tex
@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189-190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}
@@
 \mymidrule
-\gl{1}   &  \obj{k-eɸurka-} &  \obj{k-ehurka-} &    \obj{k-eɸɨrka-} &   \obj{kɨ-wɨnɨkɨ-} &  \obj{kɨ-nɨkɨ-} &   \obj{kɨ-wɨnɨkɨ-} \\
-\gl{2}   &  \obj{m-eɸurka-} &  \obj{m-ehurka-} &    \obj{m-eɸɨrka-} &    \obj{o-wɨnɨkɨ-} &  \obj{o-wnɨkɨ-} &   \obj{mɨ-wɨnɨkɨ-} \\
-\gl{1+2} &  \obj{t-eɸurka-} &  \obj{t-ehurka-} &  \obj{t͡ʃ-eɸɨrka-} &  \obj{tɨt-wɨnɨkɨ-} &  \obj{tɨ-nɨkɨ-} &  \obj{tɨt-wɨnɨkɨ-} \\
-\gl{3}   &  \obj{ɲ-eɸurka-} &  \obj{ɲ-ehurka-} &    \obj{ɲ-eɸɨrka-} &   \obj{nɨ-wɨnɨkɨ-} &  \obj{nɨ-nɨkɨ-} &   \obj{nɨ-wɨnɨkɨ-} \\
+\gl{1}   &  \rc{k-eɸurka-} &  \obj{k-ehurka-} &    \obj{k-eɸɨrka-} &   \rc{kɨ-wɨnɨkɨ-} &  \obj{kɨ-nɨkɨ-} &   \obj{kɨ-wɨnɨkɨ-} \\
+\gl{2}   &  \rc{m-eɸurka-} &  \obj{m-ehurka-} &    \obj{m-eɸɨrka-} &    \rc{o-wɨnɨkɨ-} &  \obj{o-wnɨkɨ-} &   \obj{mɨ-wɨnɨkɨ-} \\
+\gl{1+2} &  \rc{t-eɸurka-} &  \obj{t-ehurka-} &  \obj{t͡ʃ-eɸɨrka-} &  \rc{tɨt-wɨnɨkɨ-} &  \obj{tɨ-nɨkɨ-} &  \obj{tɨt-wɨnɨkɨ-} \\
+\gl{3}   &  \rc{ɲ-eɸurka-} &  \obj{ɲ-ehurka-} &    \obj{ɲ-eɸɨrka-} &   \rc{nɨ-wɨnɨkɨ-} &  \obj{nɨ-nɨkɨ-} &   \obj{nɨ-wɨnɨkɨ-} \\
 \mybottomrule

### 2023-08-23 documents/floats/say.tex
more table fixes

--- 2021-11-16 documents/floats/say.tex
@@
 \centering
-\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63; p.c., Spike Gildea]{largo2011yukpa}}
+\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[267]{meira2005bakairi}[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209, 279]{ikpengpacheco2001}[80, 153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[Spike Gildea]{pc}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63]{largo2011yukpa}}
 \label{tab:say}
-\begin{tabular}[t]{@{}llllll@{}}
+\begin{tabular}[t]{@{}ll@{}}
 \mytoprule
-Language & \multicolumn{5}{l}{Form} \\
-\midrule
-\kaxui   &   \obj{ka[s]} &  k &  a &  s &    \\
-\hixka   &   \obj{ka[h]} &  k &  a &  h &    \\
-\waiwai  &   \obj{ka[s]} &  k &  a &  s &    \\
-\arara   &      \obj{ke} &  k &  e &    &    \\
-\ikpeng  &      \obj{ke} &  k &  e &    &    \\
-\bakairi &      \obj{ke} &  k &  e &    &    \\
-\trio    &      \obj{ka} &  k &  a &    &    \\
-\akuriyo &      \obj{ka} &  k &  a &    &    \\
-\carijo  &      \obj{ka} &  k &  a &    &    \\
-\wayana  &   \obj{ka[i]} &  k &  a &    &  i \\
-\apalai  &  \obj{ka[ʃi]} &  k &  a &  ʃ &  i \\
-\kalina  &      \obj{ka} &  k &  a &    &    \\
-\kapon   &      \obj{ka} &  k &  a &    &    \\
-\pemon   &      \obj{ka} &  k &  a &    &    \\
-\macushi &      \obj{ka} &  k &  a &    &    \\
-\panare  &   \obj{ka[h]} &  k &  a &  h &    \\
-\uxc     &      \obj{ki} &  k &  i &    &    \\
-\yukpa   &      \obj{ka} &  k &  a &    &    \\
-\bottomrule
+Language &          Form \\
+\mymidrule
+\kaxui   &   \obj{ka[s]} \\
+\PWai    &    \rc{ka[s]} \\
+\hixka   &   \obj{ka[s]} \\
+\waiwai  &   \obj{ka[s]} \\
+\PPek    &       \rc{ge} \\
+\PPek    &       \rc{ke} \\
+\bakairi &      \obj{ge} \\
+\bakairi &      \obj{ke} \\
+\arara   &      \obj{ge} \\
+\arara   &      \obj{ke} \\
+\ikpeng  &      \obj{ge} \\
+\ikpeng  &    \obj{k[e]} \\
+\carijo  &      \obj{ka} \\
+\PTir    &       \rc{ka} \\
+\trio    &      \obj{ka} \\
+\akuriyo &      \obj{ka} \\
+\wayana  &   \obj{ka[i]} \\
+\apalai  &  \obj{ka[ʃi]} \\
+\kalina  &      \obj{ka} \\
+\kapon   &      \obj{ka} \\
+\pemon   &      \obj{ka} \\
+\macushi &      \obj{ka} \\
+\panare  &   \obj{ka[h]} \\
+\uxc     &      \obj{ki} \\
+\yukpa   &      \obj{ka} \\
+\mybottomrule
 \end{tabular}

### 2023-08-23 documents/sections/verbs.tex
more table fixes

@@
 Due to its semantics, these objects are either ideophones or direct speech \exref{hixsay}.
-
-\pex<hixsay>\hixka
-\a<hix-119>
-\begingl
-\glpreamble onɨ wyaro nkekonɨ bɨryekomo, tɨyonɨ wya//
-\gla onɨ wjaro n-ka-jakonɨ bɨrʲekomo tɨ-jonɨ wja//
-\glb this like \gl{3}-say-\gl{rem}.\gl{cont} boy \gl{cor}-mother \gl{obl}//
-\glft \qu{This is what the boy said to his mother.} \parencite[][36]{hixkaryanaderby1985}//
-\endgl
-\a<ordpl-35>
-\begingl
-\glpreamble moro ha, ketxkoná hatá.//
-\gla moro ha ka-jat͡ʃkonɨ hatɨ//
-\glb \gl{med}.\gl{dem}.\gl{inan} \gl{ints} say-\gl{rem}.\gl{cont}.\gl{pl} \gl{hsy}//
-\glft \qu{“That one there” they said.} \parencite[][14]{derbyshire1965textos}//
-\endgl
-\xe
+\input{examples/hixsay}
 In \exref{hixsay.hix-119}, the prefix \obj{n-} occurs because there is no preceding object (\qu{he said it like this}), while does not occur in \exref{hixsay.ordpl-35} where the verb is preceded by direct speech.
@@
 
-\pex<sayintr>
-\a<aku-154> \akuriyo \parencite[][113]{meira1998proto}\\
-\begingl
-\gla mɨ-ka//
-\glb \gl{2}-say//
-\glft \qu{You said.}//
-\endgl
-\a<kar-84> \kalina \parencite[][202]{courtz2008carib}\\
-\begingl
-\glpreamble Òmakon `wa oti ywykàpo kaiko.//
-\gla o-ʔma-kon ʔwa oti ɨ-wɨ-ka-ʔpo kai-ko//
-\glb \gl{2}-child-\gl{pl} \gl{obl} greeting \gl{1}-\gl{s_a_}-say-\gl{pst}.\gl{nmlz} say-\gl{imp}//
-\glft \qu{Pass my greetings to your children.}//
-\endgl
-\xe
+\input{examples/sayintr}
 
@@
 
-\ex<mivida-12>\carijo \parencite[][177]{robayo1989rame}\\
-\begingl
-\glpreamble iretibə et͡ʃinəme gərə jet͡ʃiɨ//
-\gla ireti-bə et͡ʃi-nə=me gərə j-et͡ʃi-ɨ//
-\glb then-from be-\gl{inf}=\gl{attrz} still \gl{1}-be-\gl{pfv}//
-\glft \qu{Then I was already grown up.}//
-\endgl
-\xe
+\input{examples/mivida-12}
 
@@
 
-\ex<car-18>\carijo \parencite[][102]{guerrero2019carijo}\\
-\begingl
-\gla əji-wa-e j-eh-ɨ//
-\glb \gl{2}-search-\gl{sup} \gl{1}-come-\gl{pfv}//
-\glft \qu{I came looking for you.}//
-\endgl
-\xe
+\input{examples/car-18}
 %
@@
 
-\pex<gowhere>
-\a<car-32> \carijo \perscommpar{David Felipe Guerrero}\\
-\begingl
-\gla irə wat͡ʃinakano tae j-ehɨtə-e//
-\glb \gl{inan}.\gl{ana} body.of.water along.bounded \gl{1}-go.down-\gl{npst}//
-\glft \qu{…I go down through that guachinacán.}//
-\endgl
-\a<yuk-14> \yukpa \parencite[][]{meira2003primeras}\\
-\begingl
-\glpreamble aw yéwtu//
-\gla aw j-ewuhtu//
-\glb \gl{1}\gl{pro} \gl{1}-go.down//
-\glft \qu{I went down.}//
-\endgl
-\xe
+\input{examples/gowhere}
 %
@@
 It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.
-
-\pex<waygodown>\wayana \parencite[][200]{wayanatavares2005}
-\begin{multicols}{2}
-\a<way-73>
-\begingl
-\glpreamble ïwïptëë//
-\gla ɨ-w-ɨptə-rɨ//
-\glb \gl{1}-\gl{s_a_}-go.down-\gl{nmlz}//
-\glft \qu{my going down}//
-\endgl
-\a<way-71>
-\begingl
-\gla əw-ɨptə-k//
-\glb \gl{2}-go.down-\gl{imp}//
-\glft \qu{Go down!}//
-\endgl
-\end{multicols}
-\xe
+\input{examples/waygodown}
 %

### 2023-08-23 workflow/process_data.py

@@ -1,7 +1,7 @@
 import pandas as pd
 import pyradigms
 from pylingdocs.helpers import decorate_gloss_string
-
+from writio import dump
 from helpers import (
     combine_form_meaning,
     combine_sources,
@@ -14,6 +14,8 @@ from helpers import (
     name_dic,
     objectify,
     plain_trans_dic,
+    print_aligned_table,
+    glossify_index,
     sort_lg,
     print_latex,
     print_shorthand,
@@ -25,7 +27,9 @@ from helpers import (
     trans_dic,
     v_df,
     add_obj_markdown,
-    repl_lg_id
+    repl_lg_id,
+    translate_cols,
+    x_infl_y_trans,
 )
 
 verb_list = [
@@ -53,11 +57,7 @@ for lg, meanings in {
         sort_orders={"Inflection": person, "Meaning_ID": meanings},
     )
     tabular = pyd.compose_paradigm()
-    tabular.index = tabular.index.map(decorate_gloss_string)
-    tabular.index.name = None
-    tabular.rename(columns=trans_dic, inplace=True)
-    tabular = tabular.apply(lambda x: x.apply(objectify))
-    tabular.columns.names = [None]
+    tabular = x_infl_y_trans(tabular)
     export_table(
         df=tabular,
         label=lg + "intro",
@@ -78,13 +78,11 @@ for lg, meaning in {"ebak": "go_up", "ara": "dance", "ikp": "run"}.items():
     sources_list.extend(get_sources(pyd, combine=False))
 
 tabular = pd.concat(entries, axis=1)
-tabular.index.names = [None]
-tabular = tabular.apply(lambda x: x.apply(objectify, obj_string=get_obj_str(x.name)))
 tabular.rename(columns=trans_dic, level="Meaning_ID", inplace=True)
 tabular.rename(columns=shorthand_dic, level="Language_ID", inplace=True)
 tabular.columns.names = [None, None]
-tabular.index = tabular.index.map(decorate_gloss_string)
 tabular.columns = [" ".join([x, y]) for x, y in tabular.columns]
+tabular = glossify_index(tabular)
 export_table(
     tabular,
     label="pekreg",
@@ -118,22 +116,16 @@ pyd.sort_orders = {
     "Inflection": ["1", "2", "1+2", "3"],
 }
[…]

### 2023-08-23 workflow/read_data.py

@@ -4,7 +4,7 @@ import pandas as pd
 pd.options.mode.chained_assignment = None
 import pynterlinear as pynt
 from segments import Profile, Tokenizer
-
+from writio import load
 
 lg_dict = {x["Shorthand"]: x["Name"] for x in cah.languages.values()}
 lg_list = list(cah.lg_order().keys())

### 2023-08-23 documents/sections/extensions.tex
abbrevs

@@
 \cref{sec:verbsummary} summarizes the results.
-%The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{s_a_} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
+%The main goal of this section is to identify the distributional patterns of innovative and conservative prefixes in the \gl{sa} lexicon, in order to search for potential explanations \pcref{sec:motivations}.
 
-\subsection{Incomplete extensions: the innovative \gl{1}\texorpdfstring{\gl{s_a_}}{Sa} markers}
+\subsection{Incomplete extensions: the innovative \gl{1}\texorpdfstring{\gl{sa}}{Sa} markers}
 \label{sec:extensions}
-As stated in \cref{sec:extensions_intro}, the six person marker extensions which left a group of verbs untouched all introduced innovative first person markers on \gl{s_a_} verbs.
+As stated in \cref{sec:extensions_intro}, the six person marker extensions which left a group of verbs untouched all introduced innovative first person markers on \gl{sa} verbs.
 Of these extensions, half can be reconstructed to intermediate proto-languages, and half happened in pre-modern stages of single languages.
-The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{s_a_} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{s_p_} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{s_p_} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
+The sources of innovative markers vary, but not much: the innovative \gl{1}\gl{sa} prefix is formally identical to the \gl{1+2}\gl{p}/\gl{sp} marker (\PC \rc{k-}) in three cases, to the \gl{1}\gl{p}/\gl{sp} marker (\PC \rc{u(j)-}) in two cases, and to the \gl{1}>\gl{3} marker (\PC \rc{t-}) in one case.
 For each extension, regular (innovative) verbs are contrasted with irregular (conservative) ones, and verb forms are reconstructed where necessary.
@@
 \cref{sec:taranoan} focuses on innovative \rc{t-} in \PTir, reflected in modern \trio and \akuriyo.
-The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative \gl{1}\gl{s_a_} markers found in single languages:
+The topic of \crefrange{sec:akuriyo}{sec:yukpa} are innovative \gl{1}\gl{sa} markers found in single languages:
 \obj{k-} in \akuriyo, and \obj{j-} in \carijo and \yukpa.
@@
 \textcite{meira2005southern}, who established the branch, focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
-However, all three Pekodian languages have a regular \gl{1}\gl{s_a_} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{s_a_} marker \rc{k-}.
+However, all three Pekodian languages have a regular \gl{1}\gl{sa} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{sa} marker \rc{k-}.
 
@@
 %& \parencite[4]{meira2003bakairi} & \parencite[150]{alves2017arara} &  \parencite[52]{ikpengpacheco2001}\\
-%\gl{1}\gl{s_a_} & \obj{\emp{k-}əku-} & \obj{\emp{k-}origu-} & \obj{\emp{k-}aranme-} \\
-%\gl{2}\gl{s_a_} & \obj{m-əku-} & \obj{m-origu-} & \obj{m-aranme-} \\
-%\gl{1+2}\gl{s_a_} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
-%\gl{3}\gl{s_a_} & \obj{n-əku-} & ∅\obj{-origu-} & ∅\obj{-aranme-} \\
+%\gl{1}\gl{sa} & \obj{\emp{k-}əku-} & \obj{\emp{k-}origu-} & \obj{\emp{k-}aranme-} \\
+%\gl{2}\gl{sa} & \obj{m-əku-} & \obj{m-origu-} & \obj{m-aranme-} \\
+%\gl{1+2}\gl{sa} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
+%\gl{3}\gl{sa} & \obj{n-əku-} & ∅\obj{-origu-} & ∅\obj{-aranme-} \\
 %\end{tabular}
@@
 \footnote{Seven in her analysis, treating the two meanings of \obj{it͡ʃi} \qu{to be / to lie down} as different verbs.}
-\arara \gl{s_a_} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
+\arara \gl{sa} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
 There is also a reflex of the copula \rc{a[p]} \pcref{sec:be}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
@@
 
-In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two \gl{s_a_} subclasses, defined by \gl{1}\gl{s_a_} \obj{w-} and \obj{k-}, respectively.\footnote{
+In his brief but precise discussion of \bakairi verbal person marking, \textcite{meira2003bakairi} reports the existence of two \gl{sa} subclasses, defined by \gl{1}\gl{sa} \obj{w-} and \obj{k-}, respectively.\footnote{
 	\textcite{meira2003bakairi} indicates that the same verbs which take first person \obj{w-} in \bakairi also take a \gl{1+2} marker \obj{k-}.
@@
 \gla w-i-də//
-\glb \gl{1}\gl{s_a_}-bathe-\gl{imm}//
+\glb \gl{1}\gl{sa}-bathe-\gl{imm}//
 \glft \qu{I bathed}//
@@
 %Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
-While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of the \arara verbs in \exref{arairr} as \gl{s_a_} verbs, he does not indicate whether they belong to the \gl{s_a_}-1 class with \obj{k-}, or the \gl{s_a_}-2 class with \obj{w-}.
+While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of the \arara verbs in \exref{arairr} as \gl{sa} verbs, he does not indicate whether they belong to the \gl{sa}-1 class with \obj{k-}, or the \gl{sa}-2 class with \obj{w-}.
 However, inflected forms can be found in \textcite{von1892bakairi}, represented in \exref{bakverbs} based on the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
@@
 \gla u-ge-pa//
-\glb \gl{1}\gl{s_a_}-say-\gl{neg}//
+\glb \gl{1}\gl{sa}-say-\gl{neg}//
 \glft \qu{I don't say.}//
@@
 \gla w-i-taki//
-\glb \gl{1}\gl{s_a_}-be-\gl{int}//
+\glb \gl{1}\gl{sa}-be-\gl{int}//
 \glft \qu{I was.}//
@@
 \gla k-əewɨ-lɨ//
-\glb \gl{1}\gl{s_a_}-come-\gl{imm}//
+\glb \gl{1}\gl{sa}-come-\gl{imm}//
 \glft \qu{I came.}//
@@
 \gla k-ɨtəgɨ-se//
-\glb \gl{1}\gl{s_a_}-go.down-\gl{npst}?//
+\glb \gl{1}\gl{sa}-go.down-\gl{npst}?//
 \glft \qu{I go down.}//
@@
 \gla u-tə//
-\glb \gl{1}\gl{s_a_}-go//
+\glb \gl{1}\gl{sa}-go//
 \glft \qu{I go.}//
@@
 \gla tərə w-a-kɨne//
-\glb there \gl{1}\gl{s_a_}-be-\gl{pst}.\gl{cont}//
+\glb there \gl{1}\gl{sa}-be-\gl{pst}.\gl{cont}//
 \glft \qu{I was there.}//
@@
 
-All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{s_a_} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
+All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{sa} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
 However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}, with the exception of \obj{k-}prefixed \qu{to go} \exref{ikp-170}.
@@
 Newly identified \ikpeng \obj{ɨ-}/Ø is demonstrably a reflex of \PXin \rc{w(ɨ)-}, based on other (albeit irregular) cases of loss of \rc{w} \pcref{tab:pxinw}.
-%The presence and distribution of the \ikpeng \gl{1}\gl{s_a_} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{s_a_} \obj{w(ɨ)-}.
+%The presence and distribution of the \ikpeng \gl{1}\gl{sa} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{sa} \obj{w(ɨ)-}.
 %Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in .
@@
 Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in correspondences like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
-Thus, a \gl{1}\gl{s_a_} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to its \arara reflex in form and distribution.
+Thus, a \gl{1}\gl{sa} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to its \arara reflex in form and distribution.
 In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
@@
 \gla m-odebɨ-nɨ//
-\glb \gl{2}\gl{s_a_}-come-\gl{rec}//
+\glb \gl{2}\gl{sa}-come-\gl{rec}//
 \glft \qu{You came.}//
@@
 This extension, one of the Parukotoan innovations shown in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
-\PWai further extended the \gl{1}\gl{s_p_} prefix \rc{k-} (innovated in \PPar) to \gl{1}\gl{s_a_}.
+\PWai further extended the \gl{1}\gl{sp} prefix \rc{k-} (innovated in \PPar) to \gl{1}\gl{sa}.
 For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
@@
 
-Not all \gl{s_a_} verbs were affected: \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
+Not all \gl{sa} verbs were affected: \waiwai \obj{ka} \qu{to say} does not take \obj{kɨ-}, but rather conservative \obj{wɨ-} \exref{hixɨ.wai-108}.
 Its \hixka counterpart has a prefix \obj{ɨ-} \exref{hixɨ.hix-118}, which also occurs in \gl{1}>\gl{3} scenarios in \hixka \exref{hixɨ.hix-17}, corresponding to \waiwai \obj{wɨ-} \exref{hixɨ.wai-159}.
@@
 \gla ro-ʃe-hɨra n-a-je hamɨ kajwerʲe ɨ-ka-jakonɨ//
-\glb \gl{1}-\gl{des}-\gl{neg} \gl{3}-be-\gl{npst}.\gl{uncert} \gl{evid} K. \gl{1}\gl{s_a_}-say-\gl{rem}.\gl{cont}//
+\glb \gl{1}-\gl{des}-\gl{neg} \gl{3}-be-\gl{npst}.\gl{uncert} \gl{evid} K. \gl{1}\gl{sa}-say-\gl{rem}.\gl{cont}//
 \glft \qu{I said (to myself), “Kaywerye evidently doesn't like me”.}//
@@
 While \hixka has the expected \obj{ɨ-}, \waiwai seems to have combined innovative \obj{k-} with  old \rc{w-}, an analysis also considered by \textcite[90]{gildea1998}.
-Alternatively, this form may have been influenced by deverbalized forms of \qu{to go} showing a fossilized of the \gl{s_a_} class marker \rc{w-}  \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}.
+Alternatively, this form may have been influenced by deverbalized forms of \qu{to go} showing a fossilized of the \gl{sa} class marker \rc{w-}  \parentext{e.g., \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}}.
 %
-%\pex<waiwto> \waiwai reflexes of the \gl{s_a_} class marker \rc{w-}
+%\pex<waiwto> \waiwai reflexes of the \gl{sa} class marker \rc{w-}
 %\a \obj{o-\emp{w}to-topo-nho} \qu{my trip} \parencite[92]{waiwaihawkins1998}
@@
 
-%Summing up, we reconstruct the four verbs \rc{eʃi} and \rc{a[h]} \qu{to be}, \rc{ka[s]} \qu{to say}, and \rc{[ɨ]to[m]} \qu{to go} as preserving the old \gl{1}\gl{s_a_} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
+%Summing up, we reconstruct the four verbs \rc{eʃi} and \rc{a[h]} \qu{to be}, \rc{ka[s]} \qu{to say}, and \rc{[ɨ]to[m]} \qu{to go} as preserving the old \gl{1}\gl{sa} marker \rc{w-} in \PWai, while the rest took on innovative \rc{k-}.
 
@@
 The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages identified by \textcite{girard1971proto}, with \carijo as a sister.
-\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places, creating a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
+\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{sa} \rc{w-} seem to have switched places, creating a regular \gl{1}\gl{sa} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
 \footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
@@
 
-%Apart from these \PTir verbs with \rc{w(ɨ)-}, there are two irregularly inflected \gl{s_a_} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
-%They have \gl{1}\gl{s_a_} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
+%Apart from these \PTir verbs with \rc{w(ɨ)-}, there are two irregularly inflected \gl{sa} verbs in \trio, \obj{ɨhtə} \qu{to go down} and \obj{weka}/\obj{oeka} \qu{to defecate} \exref{downshit}.
+%They have \gl{1}\gl{sa} markers \obj{p-} and \obj{k-}, which are otherwise entirely unattested in \trio.\footnote{Although both elements also occur in other irregular forms of these verbs \parencite[180, 325, 331]{triomeira1999}.}
 %At least for \qu{to go down}, the \akuriyo cognate suggests that the irregular first person form \rc{p-ɨhtə-} can be reconstructed to \PTir.
@@
 %
-%\ex<downshit> Idiosyncratic \gl{1}\gl{s_a_} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
+%\ex<downshit> Idiosyncratic \gl{1}\gl{sa} prefixes \parencites[294]{triomeira1999}[84]{gildea1994akuriyo}\\
 %\begin{tabular}[t]{@{}lll@{}}
@@
 %In addition, \textcite{gildea1994akuriyo} recorded four more \akuriyo verbs seemingly not affected by innovative \rc{t-} \exref{akumov.aku}, all \obj{e}-initial movement verbs.
-%Only \obj{erama} \qu{to return} has an attested \trio cognate, which behaves like a regular \gl{s_a_} verb in taking \obj{s-} \exref{akumov.tri}.
+%Only \obj{erama} \qu{to return} has an attested \trio cognate, which behaves like a regular \gl{sa} verb in taking \obj{s-} \exref{akumov.tri}.
 %Further, these verbs are not mentioned by \textcite{meira1998proto}, who was also working with \posscite{gildea1994akuriyo} data.
@@
 %\pex<akumov>
-%\a<aku> \akuriyo \gl{1}\gl{s_a_} \rc{w-} \parencite[84--86]{gildea1994akuriyo}\\
+%\a<aku> \akuriyo \gl{1}\gl{sa} \rc{w-} \parencite[84--86]{gildea1994akuriyo}\\
 %\begin{tabular}[t]{@{}ll@{}}
@@
 
-%In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} largely replaced \gl{1}\gl{s_a_} \rc{w-}.
+%In summary, in \PTir{} -- at the latest -- the \PC \gl{1}>\gl{3} marker \rc{t-} largely replaced \gl{1}\gl{sa} \rc{w-}.
 %Five verbs are solidly reconstructible as preserving \rc{w-} in \PTir, and \rc{ɨhtə} \qu{to go down} with irregular \rc{p-} was also not affected.
@@
 \label{sec:akuriyo}
-After the split of \PTir, when \rc{t-}\slash{}\rc{t͡ʃ-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{s_a_} marker \obj{k-}.
+After the split of \PTir, when \rc{t-}\slash{}\rc{t͡ʃ-} had largely replaced \rc{w-}, \akuriyo innovated a third \gl{1}\gl{sa} marker \obj{k-}.
 \obj{k-} and \obj{t͡ʃ-} show a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
@@
 %	\centering
-%	\caption{\akuriyo \gl{1}\gl{s_a_} markers in Gildea's fieldnotes}
+%	\caption{\akuriyo \gl{1}\gl{sa} markers in Gildea's fieldnotes}
 %	\label{tab:aku1sa}
@@
 The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} of course also resisted the extension of \obj{k-} in \akuriyo.
-In addition, there is an \gl{s_a_} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-}, apparently reconstructible to \PTir \exref{tridown}.
-It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{s_a_} verb when \PTir \rc{t-} was introduced is unclear (see \cref{sec:godown}).
+In addition, there is an \gl{sa} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-}, apparently reconstructible to \PTir \exref{tridown}.
+It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{sa} verb when \PTir \rc{t-} was introduced is unclear (see \cref{sec:godown}).
 
@@
 
-%Let us first consider a scenario where \obj{k-} was the older \gl{1}\gl{s_a_} marker, occurring on both \obj{e}- and \obj{ə}-initial verbs.
+%Let us first consider a scenario where \obj{k-} was the older \gl{1}\gl{sa} marker, occurring on both \obj{e}- and \obj{ə}-initial verbs.
 %Next, \obj{t-}/\obj{t͡ʃ-} would have come in and replaced \obj{k-} before \obj{e}, but (mostly) not before \obj{ə}, resulting in the observable distribution.
@@
 %
-%The other scenario is that \rc{t-} was the old first person marker occurring on the vast majority of \gl{s_a_} verbs.
+%The other scenario is that \rc{t-} was the old first person marker occurring on the vast majority of \gl{sa} verbs.
 %Subsequently, it developed into \rc{t͡ʃ-} / \_\obj{e}, but remained \rc{t-} / \_\obj{ə}; \obj{t͡ʃ}\goodtilde\obj{ʃ}\goodtilde\obj{s} and \obj{t} are phonemically contrastive in \akuriyo \parencite[16]{meira1998proto}.
@@
 
-%It has been speculated that the \trio \gl{s_a_} verb with a first person \obj{k-} marker, \obj{weka}/\obj{oeka} \qu{to defecate} might have something to do with \akuriyo \obj{k-}, which would potentially make the innovation of \rc{k-} a \PTir matter \parencite[116]{meira1998proto}.
+%It has been speculated that the \trio \gl{sa} verb with a first person \obj{k-} marker, \obj{weka}/\obj{oeka} \qu{to defecate} might have something to do with \akuriyo \obj{k-}, which would potentially make the innovation of \rc{k-} a \PTir matter \parencite[116]{meira1998proto}.
 %However, this hypothesis faces a problem in that the first person form for \akuriyo \qu{to defecate} is \obj{j-ereina-} \parencite[88]{gildea1994akuriyo}.
 %Another hypothesis is that it originated in the corresponding \kaxui form \obj{ku-weka-} (form provided by Spike Gildea, p.c.), although the occurrence of \obj{o} in \trio would still need explanation.
-%Rather, it seems likely that this irregular first person form originated in a class switch from \gl{s_p_} to \gl{s_a_} (\Cref{sec:resistant_verbs}), although it is unclear how exactly that switch resulted in the irregular form.
+%Rather, it seems likely that this irregular first person form originated in a class switch from \gl{sp} to \gl{sa} (\Cref{sec:resistant_verbs}), although it is unclear how exactly that switch resulted in the irregular form.
 
@@
 \label{sec:carijo}
-\carijo, the third Taranoan language, extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
-Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regular verbs \pcref{tab:carreg}.
+\carijo, the third Taranoan language, extended the \gl{1}\gl{sp} marker \obj{j(i)-}\footnote{Since all affected \gl{sa} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{sa} verbs \parencite[105--107]{meira1998proto}.
+Combined with the extension of \gl{2}\gl{sa} \obj{m-} and \gl{1+2}\gl{sa} \obj{kɨt-}/\obj{kɨs-} to \gl{sp} verbs, this created a single unified \gl{s} category for regular verbs \pcref{tab:carreg}.
 %
@@
 %
-Although the split-\gl{s} system was lost entirely, former \gl{s_a_} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight} \exref{car-30}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}.
+Although the split-\gl{s} system was lost entirely, former \gl{sa} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight} \exref{car-30}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}.
 %The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.
@@
 
-As noted in \cref{sec:taranoan}, this extension also erased any traces of putative \PTar \gl{1}\gl{s_a_} \rc{t-}.
-However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
+As noted in \cref{sec:taranoan}, this extension also erased any traces of putative \PTar \gl{1}\gl{sa} \rc{t-}.
+However, it did not fully eclipse the old \gl{1}\gl{sa} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
 In addition, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car1w.car-33}.
@@
 However, it does preserve the \setone prefixes in the immediate past, which shows a unified intransitive paradigm \pcref{tab:yukreg}.
-The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (now former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
+The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{sa} \obj{m(ɨ)-} to (now former) \gl{sp} verbs like \obj{nɨ} \qu{to sleep}.
 %
@@
 %
-These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their semantics and the reflex of \detrz.
+These verbs share their first person marker \obj{j(ɨ)-} with former \gl{sa} verbs like \obj{otum} \qu{to wash self}, identifiable by their semantics and the reflex of \detrz.
 Since \obj{j(ɨ)-} is the reflex of the \PC \gl{1}S\textsubscript{(P)} marker \rc{u(j)-} \parencite[92]{gildea1998}, it also occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}.
@@
 %
-Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \gl{1}\gl{s_a_} \rc{w-}.
+Since \PC \gl{1}\gl{sa} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \gl{1}\gl{sa} \rc{w-}.
 
@@
 \gla aw {\normalfont ∅}-to//
-\glb \gl{1}\gl{pro} \gl{1}\gl{s_a_}-go//
+\glb \gl{1}\gl{pro} \gl{1}\gl{sa}-go//
 \glft \qu{I went.}//

### 2023-08-23 documents/sections/introduction.tex
abbrevs

@@
 \textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
-with paradigms of four verbs, all members of the \gl{s_a_} inflectional class.
-In this language, the verb \qu{to be} diverges from other \gl{s_a_} verbs like \qu{to fall} by having a first person marker \obj{w-}, rather than \obj{k-}.
-A similar pattern exists in \trio \pcref{tab:triintro}, where \gl{s_a_} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}}, while the verb \qu{to go} has a first-person prefix \obj{wɨ-}.
-For both languages, the first person prefix on the first four verbs in the table is representative for the vast majority of \gl{s_a_} verbs.
+with paradigms of four verbs, all members of the \gl{sa} inflectional class.
+In this language, the verb \qu{to be} diverges from other \gl{sa} verbs like \qu{to fall} by having a first person marker \obj{w-}, rather than \obj{k-}.
+A similar pattern exists in \trio \pcref{tab:triintro}, where \gl{sa} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}}, while the verb \qu{to go} has a first-person prefix \obj{wɨ-}.
+For both languages, the first person prefix on the first four verbs in the table is representative for the vast majority of \gl{sa} verbs.
 %In both languages, there are only a few other verbs inflected identically to the divergent ones on the right; for example, the first-person form of \trio \qu{to be} is \obj{w-ei-} \parencites[339]{triomeira1999}.

### 2023-08-23 documents/sections/proto_carib.tex
fixes #1
more table fixes
abbrevs

--- 2023-08-04 documents/sections/proto_carib.tex
@@
 The irregular first person prefixes from \cref{sec:intro} are relics, inherited from the ancestral \PC system \pcref{sec:pc_person}.
-That system underwent much innovation; the mechanism responsible for the irregular forms involves person marker extensions not spreading through the entire \gl{s_a_} lexicon \pcref{sec:extensions_intro}.
-A specific aspect of the system, the \gl{s_a_} vs \gl{s_p_} distinction, plays a role in incomplete extensions and is discussed in \cref{sec:split}.
+That system underwent much innovation; the mechanism responsible for the irregular forms involves person marker extensions not spreading through the entire \gl{sa} lexicon \pcref{sec:extensions_intro}.
+A specific aspect of the system, the \gl{sa} vs \gl{sp} distinction, plays a role in incomplete extensions and is discussed in \cref{sec:split}.
 %\cref{sec:outlook} summarizes the main points of this section and provides an outlook for the rest of the paper.
@@
 Formally identical or etymologically related markers occured in intransitive verbs, which showed a split-\gl{s} system \pcref{tab:pcintrans}.
-\gl{s_a_} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}\gl{s_a_} \rc{w-} vs \gl{1}>\gl{3} \rc{t[i]-}) and the absence of \rc{i} after all \gl{s_a_} prefixes.
-On the other hand, \gl{s_p_} verbs took markers fully identical to the \gl{p}-oriented ones, with \gl{3}\gl{s_p_} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
+\gl{sa} verbs took similar markers as the \gl{a}-oriented ones in transitive verbs, with the exception of first person (\gl{1}\gl{sa} \rc{w-} vs \gl{1}>\gl{3} \rc{t[i]-}) and the absence of \rc{i} after all \gl{sa} prefixes.
+On the other hand, \gl{sp} verbs took markers fully identical to the \gl{p}-oriented ones, with \gl{3}\gl{sp} \rc{n(i)-} aligning with \gl{3}>\gl{3} scenarios.
 
-Knowledge about the ancestral system makes evident that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{s_a_} prefix \rc{w-}; they are therefore \firstmention{conservative}.
-They contrast with regular \gl{s_a_} verbs, which are innovative in both languages.
+Knowledge about the ancestral system makes evident that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{sa} prefix \rc{w-}; they are therefore \firstmention{conservative}.
+They contrast with regular \gl{sa} verbs, which are innovative in both languages.
 The reflexes of \rc{w-} are \firstmention{relics}, old and restricted to a few lexemes, contrasting with the innovative prefixes found elsewhere.
@@
 %In principle, one could also imagine a scenario in which verbs of one of the classes are gradually replaced with innovative verbs from the other class, as suggested for \panare by \textcite[225]{meira2000split}.
-%However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{s_a_} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{s_a_}:\gl{s_p_} ratio in Meira's sample.
-%Thus, the hypothesis that \panare has largely replaced \gl{s_p_} with \gl{s_a_} verbs remains to be thoroughly tested.
+%However, the verbs Meira sampled came from the \emp{A}-section of \posscite{mattei1994diccionario} dictionary \perscommpar{Spike Gildea}, and \obj{at-}\slash{}\obj{at͡ʃ-}\slash{}\obj{as-} is a frequent reflex of \detrz in \panare, meaning that many \gl{sa} verbs are \obj{a}-initial, thus explaining the 80:40 \gl{sa}:\gl{sp} ratio in Meira's sample.
+%Thus, the hypothesis that \panare has largely replaced \gl{sp} with \gl{sa} verbs remains to be thoroughly tested.
 %For the moment, the most likely scenario leading to the eventual loss of the split-\gl{s} system involves markers being extended to new verbs.}
@@
     	\begin{tabular}[t]{ll}
-    		\PPar & \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}\\
-    		& \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_}\tnote{a}\\
-    		& \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_}\tnote{a}\\
-    		\PWai & \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}\\
+    		\PPar & \gl{1}\gl{sa} \rc{w-} to \gl{1}>\gl{3}\\
+    		& \gl{1+2} \rc{k-} to \gl{1}\gl{sp}\tnote{a}\\
+    		& \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{sp}\tnote{a}\\
+    		\PWai & \gl{1}\gl{sp} \rc{k-} to \gl{1}\gl{sa}\\
     		& \rc{owɨ(ro) j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}\\
-    		\waiwai & \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}
+    		\waiwai & \gl{2}\gl{sa} \obj{m-} to \gl{2}\gl{sp}
     	\end{tabular}
@@
 %	\item \PPar \begin{enumerate}
-%		\item \gl{1}\gl{s_a_} \rc{w-} to \gl{1}>\gl{3}
-%		\item \gl{1+2} \rc{k-} to \gl{1}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
-%		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{s_p_} (completed in \PWai, ongoing in \kaxui)
+%		\item \gl{1}\gl{sa} \rc{w-} to \gl{1}>\gl{3}
+%		\item \gl{1+2} \rc{k-} to \gl{1}\gl{sp} (completed in \PWai, ongoing in \kaxui)
+%		\item \gl{1+2} \rc{kɨt-} to \gl{1+2}\gl{sp} (completed in \PWai, ongoing in \kaxui)
 %	\end{enumerate}
 %	\item \PWai \begin{enumerate}
-%		\item \gl{1}\gl{s_p_} \rc{k-} to \gl{1}\gl{s_a_}
+%		\item \gl{1}\gl{sp} \rc{k-} to \gl{1}\gl{sa}
 %		\item innovative \rc{owɨro j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}
@@
 %	\item \waiwai \begin{enumerate}
-%		\item \gl{2}\gl{s_a_} \obj{m-} to \gl{2}\gl{s_p_}\end{enumerate}
+%		\item \gl{2}\gl{sa} \obj{m-} to \gl{2}\gl{sp}\end{enumerate}
 %\end{enumerate}
@@
 All innovations are person marker extensions except the innovative first person P form, which combined a pronoun with the linker \rc{j-}.
-\cref{fig:par_ext} shows them in bold and reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{s_p_} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
+\cref{fig:par_ext} shows them in bold and reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{sp} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
 %
@@
 %In this case, the loss of the inflectional classes also entailed the loss of the other morphological traces of the split-\gl{s} system.
-%That is, the \gl{s_a_} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
-%Similarly, the \gl{2}\gl{s_p_} prefix \obj{a-} was extended to imperatives of (former) \gl{s_a_} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.
+%That is, the \gl{sa} class marker \rc{w-} \pcref{sec:split} was lost in \waiwai, as evidenced by the contrast between the \waiwai and \kaxui deverbal forms in \exref{wailoss.wai-158} and \exref{wailoss.kax-138}.
+%Similarly, the \gl{2}\gl{sp} prefix \obj{a-} was extended to imperatives of (former) \gl{sa} verbs \exref[wailoss.wai-112]{wailoss.hix-43}, but only C-initial ones \parencite[62]{waiwaihawkins1998}.
 
@@
 %todo maybe put in GOOD kaxui data
-Second, while there is no detailed diachronic account of the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading simultaneously.
+Second, while there is no detailed diachronic account of the switch of \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{sa} \rc{w-} in the Tiriyoan languages \pcref{sec:taranoan}, \textcite[111--112]{meira1998proto} argues that it must have happened gradually rather than instantaneously, and entailed both markers spreading simultaneously.
 Whether or not this gradual switch followed ordered lines, lexical diffusion must have played a role.
@@
 The third argument in favor of the lexical diffusion scenario goes back to conservative forms like those in \hixka and \trio.
-In both languages, the innovative \gl{1}\gl{s_a_} prefixes were introduced by a person marker extension spreading via lexical diffusion.
-The continued presence of the old \gl{1}\gl{s_a_} prefix in a few verbs is the result of the extension stopping short of these verbs, rather than spreading through the entire \gl{s_a_} lexicon.
+In both languages, the innovative \gl{1}\gl{sa} prefixes were introduced by a person marker extension spreading via lexical diffusion.
+The continued presence of the old \gl{1}\gl{sa} prefix in a few verbs is the result of the extension stopping short of these verbs, rather than spreading through the entire \gl{sa} lexicon.
 In a family-wide search, 18 distinct extensions affecting intransitive verbs were identified, 6 of them incomplete.
@@
 
-Interestingly, all six featured innovative first person markers on \gl{s_a_} verbs.
+Interestingly, all six featured innovative first person markers on \gl{sa} verbs.
 All other (complete) extensions%
 \footnote{As an honorable mention, when \ikpeng replaced third person \setone with \settwo prefixes, \obj{a} \qu{to be} and \obj{ke} \qu{to say} retained \obj{n-} \parencite[12]{matter2019arara}.
-	However, innovative markers were not spreading within a paradigm, but rather from former subordinate to main clauses.} either occurred with other person values and/or targeted \gl{s_p_} verbs.
-Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{s_a_} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{s_p_} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{s_a_} \obj{m(ɨ)-} in to \gl{s_p_} verbs in \panare \pcref{tab:panare}, and the extension of the entire \gl{s_a_} set to \gl{s_p_} verbs in \waimiri \pcref{tab:waimiri}.
-%, as well as the extension of \rc{k-} to \gl{1}\gl{s_p_} and of \rc{tɨt-} to \gl{1+2}\gl{s_p_} in \PWai \pcref{fig:par_ext}.
-The starkly different behavior of \gl{s_a_} and \gl{s_p_} verbs regarding extensions points to the split-\gl{s} system playing a role, so its main properties will be discussed in \cref{sec:split}.
-This will also clarify how the \gl{s_a_}/\gl{s_p_} distinction can be lost for a single person, or how \gl{s_p_} verbs can take on \gl{s_a_} markers with apparent semantic impunity.
+	However, innovative markers were not spreading within a paradigm, but rather from former subordinate to main clauses.} either occurred with other person values and/or targeted \gl{sp} verbs.
+Illustrative examples for complete extensions are shown in \cref{tab:completed}: the extension of \gl{1+2}\gl{sa} \obj{s(ɨ)-} (< \rc{kɨt-}) to \gl{sp} verbs in \apalai \pcref{tab:apalai}, of \gl{2}\gl{sa} \obj{m(ɨ)-} in to \gl{sp} verbs in \panare \pcref{tab:panare}, and the extension of the entire \gl{sa} set to \gl{sp} verbs in \waimiri \pcref{tab:waimiri}.
+%, as well as the extension of \rc{k-} to \gl{1}\gl{sp} and of \rc{tɨt-} to \gl{1+2}\gl{sp} in \PWai \pcref{fig:par_ext}.
+The starkly different behavior of \gl{sa} and \gl{sp} verbs regarding extensions points to the split-\gl{s} system playing a role, so its main properties will be discussed in \cref{sec:split}.
+This will also clarify how the \gl{sa}/\gl{sp} distinction can be lost for a single person, or how \gl{sp} verbs can take on \gl{sa} markers with apparent semantic impunity.
 
@@
 \label{sec:split}
-As seen in \cref{sec:pc_person}, the split between \gl{s_a_} and \gl{s_p_} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only difference:
-In deverbalized forms, \gl{s_a_} verbs took a class marker \rc{w-}, while \gl{s_p_} verbs lacked that prefix \parencites[89, 141--142]{gildea1998}[208]{meira2000split}.
-Also, \gl{s_p_} verbs took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} in imperatives, while \gl{s_a_} verbs were unprefixed \parencites[89]{gildea1998}[208]{meira2000split}.
+As seen in \cref{sec:pc_person}, the split between \gl{sa} and \gl{sp} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only difference:
+In deverbalized forms, \gl{sa} verbs took a class marker \rc{w-}, while \gl{sp} verbs lacked that prefix \parencites[89, 141--142]{gildea1998}[208]{meira2000split}.
+Also, \gl{sp} verbs took the \gl{2}\gl{sp} prefix \rc{ə(j)-} in imperatives, while \gl{sa} verbs were unprefixed \parencites[89]{gildea1998}[208]{meira2000split}.
 
-%\parencite[208]{meira2000split} Many languages show reflexes \gl{s_a_}  in deverbalized forms, originating in .
-%With \gl{s_a_} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{s_p_} verbs took the bare prefixes \pcref{tab:participles,tab:nominalizations}.
+%\parencite[208]{meira2000split} Many languages show reflexes \gl{sa}  in deverbalized forms, originating in .
+%With \gl{sa} verbs, \rc{w-} occurred immediately between the possessive prefixes and the verb stem, while \gl{sp} verbs took the bare prefixes \pcref{tab:participles,tab:nominalizations}.
 %Reflexes of \rc{w-} in languages from different branches are illustrated in \cref{tab:participles,tab:nominalizations} for participles and nominalizations.%
@@
 %
-%The distinction between \gl{s_a_} and \gl{s_p_} was also reflected in imperatives, where the latter took the \gl{2}\gl{s_p_} prefix \rc{ə(j)-} while the former were unprefixed .
+%The distinction between \gl{sa} and \gl{sp} was also reflected in imperatives, where the latter took the \gl{2}\gl{sp} prefix \rc{ə(j)-} while the former were unprefixed .
 %This is illustrated with reflexes in various modern languages in \pcref{tab:imperatives}.
-%Both the \gl{s_a_} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
-%However, there was one further property uniting most \gl{s_a_} verbs, not based on inflectional morphology.
+%Both the \gl{sa} marker \rc{w-} and the prefix pattern in imperatives have been lost in some languages, but are reconstructible to \PC.
+%However, there was one further property uniting most \gl{sa} verbs, not based on inflectional morphology.
 
-In modern instantiations of Cariban split-\gl{s}, mismatches between the semantics of verbs and their \gl{s_a_} or \gl{s_p_} status are common, exemplified with \kalina data in \exref{kar2}.
+In modern instantiations of Cariban split-\gl{s}, mismatches between the semantics of verbs and their \gl{sa} or \gl{sp} status are common, exemplified with \kalina data in \exref{kar2}.
 %
@@
 %
-\pex<kar2> \kalina
-\a<kar-59>
-\begingl
-\gla sipi tɨnka-rɨ m-ekema-non hen//
-\glb net pull-\gl{nmlz} \gl{2}\gl{s_a_}-be.afraid-\gl{prs}.\gl{uncert} eh?//
-\glft \qu{You're afraid to pull up the net, aren't you? } \parencite[][253]{courtz2008carib}//
-\endgl
-\a<kar-61>
-\begingl
-\glpreamble aya:woiya //
-\gla aj-awomɨ-ja//
-\glb \gl{2}\gl{s_p_}-get.up-\gl{prs}//
-\glft \qu{You are getting up.} \parencite[][167]{hoff1968carib}//
-\endgl
-\xe
+\input{examples/kar2.tex}
 %
-The \gl{s_a_} verb \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker \exref{kar2.kar-59}, while the \gl{s_p_} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+The \gl{sa} verb \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker \exref{kar2.kar-59}, while the \gl{sp} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
 In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
@@
 Rather, the reason for a verb to take \gl{a}- or \gl{p}-oriented prefixes is (at least diachronically) a morphological one.
-\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a derivational detransitivizing prefix are treated as \gl{s_a_} verbs, while essentially all others are \gl{s_p_} verbs:
+\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a derivational detransitivizing prefix are treated as \gl{sa} verbs, while essentially all others are \gl{sp} verbs:
 \begin{quotebox}{\parencite[201]{meira2000split}}
-	Almost all verbs in the \gl{s_a_} class are detransitivized forms of transitive verbs, either synchronically (with still existing transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
+	Almost all verbs in the \gl{sa} class are detransitivized forms of transitive verbs, either synchronically (with still existing transitive sources) or diachronically (with reconstructible but no longer existing transitive sources)
 \end{quotebox}
-%\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{s_a_} verbs, rather than being inflectional in nature:
+%\textcite[221--223]{meira2000split} also argues that the detransitivizing prefixes are indeed deriving \gl{sa} verbs, rather than being inflectional in nature:
 %\begin{inlinelist}
-%	\item there are a few underived \gl{s_a_} verbs, with no detransitivizing prefix
-%	\item \gl{s_a_} verbs can develop irregular semantics compared to their transitive counterparts
-%	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{s_a_} verb
-%	\item some originally derived \gl{s_a_} verbs have lost their transitive counterparts
-%	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{s_a_})} from \qu{to dance with (\gl{tr})}
+%	\item there are a few underived \gl{sa} verbs, with no detransitivizing prefix
+%	\item \gl{sa} verbs can develop irregular semantics compared to their transitive counterparts
+%	\item it is unpredictable whether the \gl{a} or \gl{p} argument of the underlying transitive verb becomes the \gl{s} of the derived \gl{sa} verb
+%	\item some originally derived \gl{sa} verbs have lost their transitive counterparts
+%	\item \dbqu{basic} concepts are expressed as derivations of more complex concepts, like \qu{to dance (\gl{sa})} from \qu{to dance with (\gl{tr})}
 %\end{inlinelist}.
@@
 \begin{quotebox}{\parencite[226]{meira2000split}}
-	Apparently, the morphological behavior of the \gl{s_a_} verb class is an accidental consequence of the fact that detransitivization, as far back as we can reconstruct, entails all the morphology described […] as typical of \gl{s_a_} verbs.
+	Apparently, the morphological behavior of the \gl{sa} verb class is an accidental consequence of the fact that detransitivization, as far back as we can reconstruct, entails all the morphology described […] as typical of \gl{sa} verbs.
 The alignment of person-marking prefixes appears not to be driven by any semantic forces in the language; it is as though they were being dragged by the evolution of the reflexive marker.
@@
 This analysis is applied to \trio by \textcite[268--269]{triocarlin2004}, who interprets \obj{i-} as marking transitive diathesis.}
-Reflexes of \detrz show a range of meanings summarizable as \dbqu{detransitive}, illustrated with \trio \gl{s_a_} verbs in \exref{triodetrz}.
+Reflexes of \detrz show a range of meanings summarizable as \dbqu{detransitive}, illustrated with \trio \gl{sa} verbs in \exref{triodetrz}.
 
@@
 However, both can occur with either meaning -- at least for these two verbs.
-In the next section, it will be seen that many of the verbs not affected by person marker extensions belong to the small group of \gl{s_a_} verbs without a reflex of \detrz.
-%As mentioned, there are a few \gl{s_a_} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
+In the next section, it will be seen that many of the verbs not affected by person marker extensions belong to the small group of \gl{sa} verbs without a reflex of \detrz.
+%As mentioned, there are a few \gl{sa} verbs that are not derived from transitive verbs, and do not have a reflex of \detrz, either.
 %\textcite[221]{meira2000split} characterizes these as very old and exceptional, and counts between 5 and 7 verbs, depending on the language.
-%\textcite{gildea2007greenberg} identify 7 such underived \gl{s_a_} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
+%\textcite{gildea2007greenberg} identify 7 such underived \gl{sa} verbs as reconstructible to \PC, shown in \cref{tab:underived2007}.
 %Many of these verbs will be discussed in \cref{sec:verbs}.
@@
 %\label{sec:outlook}
-%\gl{s_a_} verbs irregularly inflected for first person were identified in 9 Cariban languages.
+%\gl{sa} verbs irregularly inflected for first person were identified in 9 Cariban languages.
 %These irregular forms are actually conservative, belonging to a few verbs unaffected by person marker extensions spread via lexical diffusion.
-%Such conservative forms are only found among the first person forms of  (etymological) \gl{s_a_} verbs.
+%Such conservative forms are only found among the first person forms of  (etymological) \gl{sa} verbs.

### 2023-08-23 documents/sections/reasons.tex
some small reviewer 2 fixes
abbrevs

--- 2021-12-25 documents/sections/reasons.tex
@@
 This network created new strong verbs in some dialects, like \obj{sniːk}--\obj{snʌk} or \obj{brɪŋ}--\obj{brʌŋ} \parencite[129--130]{bybee1985morphology}.
-These verbs formed a lexical connection with prototypical members of the group, developing new past tense forms as a result.
+These verbs formed a lexical connection with prototypical members of the group, developing new participle forms as a result.
 
@@
 
-When considering the \gl{s_a_} verbs with innovative first person markers (those not in \cref{tab:overview}), several salient factors emerge as potential bases of lexical networks.
+When considering the \gl{sa} verbs with innovative first person markers (those not in \cref{tab:overview}), several salient factors emerge as potential bases of lexical networks.
 The most obvious similarity is that they all have a reflex of the detransitivizer \detrz{} \parentext{see e.g.
-\textcite[112]{meira1998proto} for Taranoan}, a hallmark of \gl{s_a_} verbs \pcref{sec:split}.
+\textcite[112]{meira1998proto} for Taranoan}, a hallmark of \gl{sa} verbs \pcref{sec:split}.
 %This was already noted by \textcite{meira1998proto} for the group of \PTar verbs taking irregular first person \rc{w-}:
@@
 %	This category includes a small number of stems, among which ‘to go’, ‘to come’, ‘to say’, [...]\footnote{The original list includes \qu{to go down} and \qu{to defecate}.
-While these verbs are indeed underived \gl{s_a_} verbs in \trio, no irregular first person \rc{w-} can be reconstructed.
+While these verbs are indeed underived \gl{sa} verbs in \trio, no irregular first person \rc{w-} can be reconstructed.
 As their inclusion in the list was most likely an error, we omit them here.} and the copula.
@@
 %\end{quotebox} % ‘to go down’, ‘to defecate’,
-Due to \detrz{} being prefixes, all derived \gl{s_a_} verbs begin with reflexes of \rc{ə} or \rc{e} \parentext{see e.g.
+Due to \detrz{} being prefixes, all derived \gl{sa} verbs begin with reflexes of \rc{ə} or \rc{e} \parentext{see e.g.
 \textcite[153]{alves2017arara} for \arara{}}, making phonologically based networks a second possibility.
-%A more trivial connection between these other verbs is that they are all \gl{s_a_} verbs, and thus share inflectional (morphological) patterns.
-%In the following analysis, the \gl{1}\gl{s_a_} (pre-innovation) prefix will serve as a proxy for fhese inflectional patterns.
-%This is made possible by the fact that \gl{s_a_} verbs all had a reflex of \rc{w-} at the time of innovation, making the first person prefix representative for the inflectional class.
-%The exception to this is \akuriyo, where innovative \gl{1}\gl{s_a_} \obj{k-} and idiosyncratic \obj{p-} on \qu{to go down} created inflectional subclasses of \gl{s_a_} verbs.
-An inflectional commonality of innovative verbs is their \gl{s_a_} (sub-)class membership, represented by pre-extension \gl{1}\gl{s_a_} \rc{w-} (\obj{t-} in \akuriyo).
+%A more trivial connection between these other verbs is that they are all \gl{sa} verbs, and thus share inflectional (morphological) patterns.
+%In the following analysis, the \gl{1}\gl{sa} (pre-innovation) prefix will serve as a proxy for fhese inflectional patterns.
+%This is made possible by the fact that \gl{sa} verbs all had a reflex of \rc{w-} at the time of innovation, making the first person prefix representative for the inflectional class.
+%The exception to this is \akuriyo, where innovative \gl{1}\gl{sa} \obj{k-} and idiosyncratic \obj{p-} on \qu{to go down} created inflectional subclasses of \gl{sa} verbs.
+An inflectional commonality of innovative verbs is their \gl{sa} (sub-)class membership, represented by pre-extension \gl{1}\gl{sa} \rc{w-} (\obj{t-} in \akuriyo).
 There are no obvious semantic patterns, which is unsurprising considering the absence of a semantic basis in the split-\gl{s} system \pcref{sec:split}.
-Thus, there are three hypotheses for factors connecting members of the networks in which extensions spread: a reflex of \detrz, the stem-initial phoneme, or a specific \gl{1}\gl{s_a_} prefix.
+Thus, there are three hypotheses for factors connecting members of the networks in which extensions spread: a reflex of \detrz, the stem-initial phoneme, or a specific \gl{1}\gl{sa} prefix.
 
@@
 Going beyond intuition is difficult due to the lack of frequency counts for individual lexemes for any Cariban language.
-The only statement in the literature is \posscite[75]{courtz2008carib} claim of \kalina underived \gl{s_a_} verbs being the most frequent ones: \dbqu{It is difficult […] to imagine an intransitive or transitive origin for some of the most frequent middle verbs}.
+The only statement in the literature is \posscite[75]{courtz2008carib} claim of \kalina underived \gl{sa} verbs being the most frequent ones: \dbqu{It is difficult […] to imagine an intransitive or transitive origin for some of the most frequent middle verbs}.
 This claim is neither supported by frequency counts nor accompanied by a list of verbs, although those verbs likely correspond to the five first columns of \cref{tab:overview}.
-To improve the data situation, a count of \gl{s_a_} verbs in three \apalai texts from \textcite{koehns1994textos} will serve as a second source of frequency information \pcref{tab:apalaicounts}.
+To improve the data situation, a count of \gl{sa} verbs in three \apalai texts from \textcite{koehns1994textos} will serve as a second source of frequency information \pcref{tab:apalaicounts}.
 The count data support the above interpretation of \posscite{courtz2008carib} claim, since defining \dbqu{high frequency} as above average yields the exact same five verbs.
@@
 
-%\textcite[153]{alves2017arara}: \dbqu{Observe que todos os 12 verbos indexados pelo morfema \obj{k-} [\gl{s_a_}] apresentam a raiz iniciada pela vogal /o/}.
+%\textcite[153]{alves2017arara}: \dbqu{Observe que todos os 12 verbos indexados pelo morfema \obj{k-} [\gl{sa}] apresentam a raiz iniciada pela vogal /o/}.
 
@@
 The resulting six hypotheses were tested by predicting the expected behavior of each verb in each extension, illustrated in \cref{tab:ptir-predictions} for \PTir.
-For example, \rc{eʔi} \qu{to be} is expected to participate (\checkmark) in innovations spreading in a phonologically defined (\rc{e}-initial) or inflectionally defined (\gl{1}\gl{s_a_} \rc{w-}) network.
+For example, \rc{eʔi} \qu{to be} is expected to participate (\checkmark) in innovations spreading in a phonologically defined (\rc{e}-initial) or inflectionally defined (\gl{1}\gl{sa} \rc{w-}) network.
 On the other hand, a network based on the detransitivizer predicts it to not undergo innovation (×).
@@
 The resulting scores are illustrated for the extension of \PTir \rc{t-} in \cref{tab:ptir-evaluations} and summed up for all extensions in \cref{tab:resultsoverview}.
-These scores crucially only consider the seven verbs in \cref{tab:overview}, but each extension affected many run-of-the-mill \gl{s_a_} verbs.%
+These scores crucially only consider the seven verbs in \cref{tab:overview}, but each extension affected many run-of-the-mill \gl{sa} verbs.%
 \footnote{
-While some non-detransitivized \gl{s_a_} verbs \parencites[252]{triomeira1999}[222]{meira2000split}[30]{gildea2007greenberg} are not in \cref{tab:overview}, these are mostly \rc{ə}-initial and were likely productively derived at some point, see \textcite[252]{triomeira1999} for \trio.
-The verbs to which this does not apply, like \trio \obj{wa} \qu{to dance} \parencites[252]{triomeira1999}, are all instances of \gl{s_p_} verbs switching classes.
-Since none of them is attested as having \gl{s_a_} status at the time of a person marker extension, they are not relevant for this study.}
-If one simulates the addition of 1'000 regularly inflected detransitivized \gl{s_a_} verbs per language -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six hypotheses consistently predict the behavior of 99.99+\% verbs correctly.
+While some non-detransitivized \gl{sa} verbs \parencites[252]{triomeira1999}[222]{meira2000split}[30]{gildea2007greenberg} are not in \cref{tab:overview}, these are mostly \rc{ə}-initial and were likely productively derived at some point, see \textcite[252]{triomeira1999} for \trio.
+The verbs to which this does not apply, like \trio \obj{wa} \qu{to dance} \parencites[252]{triomeira1999}, are all instances of \gl{sp} verbs switching classes.
+Since none of them is attested as having \gl{sa} status at the time of a person marker extension, they are not relevant for this study.}
+If one simulates the addition of 1'000 regularly inflected detransitivized \gl{sa} verbs per language -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six hypotheses consistently predict the behavior of 99.99+\% verbs correctly.
 Since the descriptive coverage of Cariban languages makes using real such large-scale data impossible, the present investigation is restricted to the edge cases.
@@
 \footnote{The assumption is made that the idiosyncratic evolution of \rc{e-pɨ} \qu{to bathe (\gl{intr})} to \rc{ipɨ} \PPek made the verb morphologically opaque.}
-\gl{s_a_} verbs were not affected, all other \gl{s_a_} verbs taking \rc{k-}.
+\gl{sa} verbs were not affected, all other \gl{sa} verbs taking \rc{k-}.
 Not shown in \cref{tab:resultsoverview} are subsequent evolutions in the Pekodian daughter languages, which largely support a detransitivizer-based explanation:
@@
 For the other extensions, the network model gives no unambiguous answer to the question of what (combination of) factors caused innovative first person markers to spread the way they did.
-This is due three of the factors in the model -- frequency, detransitivizer, phonology -- strongly converging in their predictions:
-The most frequent \gl{s_a_} verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.
+This is due three of the candidate factors in the model -- frequency, detransitivizer, phonology -- strongly converging in their predictions:
+The most frequent \gl{sa} verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{sa} verbs.

### 2023-08-23 documents/sections/verbs.tex
abbrevs

@@
 \label{sec:verbs}
-In \cref{sec:extensions}, six incomplete extensions of personal prefixes into \gl{1}\gl{s_a_} territory and the verbs unaffected by them were identified.
+In \cref{sec:extensions}, six incomplete extensions of personal prefixes into \gl{1}\gl{sa} territory and the verbs unaffected by them were identified.
 This set is rather small in most cases, and many of the verbs remain conservative in different (proto-)languages.
@@
 \kalina has a minimal pair between transitive \obj{ka} \qu{to remove} and intransitive \obj{ka} \qu{to say}, \obj{sikai} \qu{I took it away} vs \obj{wɨkai} \qu{I said} \parencite[288, 45]{courtz2008carib}.
-Similarly, \PPek \rc{ke} \qu{to say} had \gl{1}\gl{s_a_} \rc{w-} \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
-Additionally, languages differentiating transitive from \gl{s_a_} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} have \obj{ɨ}-final prefixes, see \akuriyo in \exref{sayintr.aku-154}, as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for cognate forms in other such languages.
-Finally, the \gl{s_a_} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
+Similarly, \PPek \rc{ke} \qu{to say} had \gl{1}\gl{sa} \rc{w-} \pcref{sec:pekodian}, rather than \gl{1}>\gl{3} \obj{s-} (\bakairi) or \rc{ini-} (\PXin).
+Additionally, languages differentiating transitive from \gl{sa} prefixes by the presence of \obj{i} \parencite[495]{meira2010origin} have \obj{ɨ}-final prefixes, see \akuriyo in \exref{sayintr.aku-154}, as well as \textcites[294]{triomeira1999}[195]{wayanatavares2005}[288]{ikpengpacheco2001}[150]{alves2017arara}[168]{hoff1968carib} for cognate forms in other such languages.
+Finally, the \gl{sa} class marker \obj{w-} occurs on nominalizations in \kalina \exref{sayintr.kar-84}, and is probably reflected as vowel length in the \trio \parencite[333]{triomeira1999} and \wayana \parencite[196]{wayanatavares2005} participles.
 
@@
 Reflexes of these roots are used suppletively, conditioned by person and\slash{}or \gl{tam}.
-Both roots preserved \gl{1}\gl{s_a_} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
+Both roots preserved \gl{1}\gl{sa} \rc{w-} in \PPek, \PWai, and \PTir (\crefrange{sec:pekodian}{sec:taranoan}).
 \akuriyo \obj{a} was not targeted by the extension of \obj{k-} \pcref{sec:akuriyo}, while no first person form of \obj{eʔi} is attested. 
@@
 The clear segmentability of \rc{ət-} in combination with its form suggest that it is a detransitivizing prefix.
-Although the combination of a detransitivizer and an intransitive verb makes no sense semantically, some historical \gl{s_p_} verbs are attested as adding the detransitivizer to become \gl{s_a_} verbs.
-For example, the \PC \gl{s_p_} verb \rc{wɨnɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{s_a_}.
+Although the combination of a detransitivizer and an intransitive verb makes no sense semantically, some historical \gl{sp} verbs are attested as adding the detransitivizer to become \gl{sa} verbs.
+For example, the \PC \gl{sp} verb \rc{wɨnɨkɨ} \qu{to sleep} becomes \trio \obj{əənɨkɨ} \parencite[252]{triomeira1999} and \kalina \obj{əʔnɨkɨ} \parencite[429]{courtz2008carib}, both \gl{sa}.
 \waiwai \qu{to sleep} can be \obj{wɨnɨk} \parencite[30]{waiwaihawkins1998} or \obj{et-wɨnɨk} \parencite[204]{hawkins1953waiwai}.
-However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently was an \gl{s_a_} verb already (although its reflexes in languages preserving the split-\gl{s} system mostly co-occur with reflexes of \rc{ət-epɨ}).
+However, unlike \qu{to sleep}, bare \rc{epɨ} \qu{to come} apparently was an \gl{sa} verb already (although its reflexes in languages preserving the split-\gl{s} system mostly co-occur with reflexes of \rc{ət-epɨ}).
 
@@
 While a verb \rc{ɨpɨtə} can be reconstructed to \PC, different (proto-)languages do not agree about its class.
-Its reflexes in languages that preserve the split-\gl{s} system are distributed fairly evenly between \gl{s_a_} and \gl{s_p_}.
+Its reflexes in languages that preserve the split-\gl{s} system are distributed fairly evenly between \gl{sa} and \gl{sp}.
 
@@
 The verb shows traits of both classes in \wayana, making it a \dbqu{mixed} verb synchronically.
-It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
-It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.
+It takes the first and second person \gl{sp} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{sa} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
+It also shows the \gl{sa} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{sp} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.
 \input{examples/waygodown}
 %
-Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the \PC  causativizer \rc{-ka} was restricted to \gl{s_p_} verbs \parencite{gildea2019overview}.
-These patterns point to \qu{to go down} being a regular \gl{s_p_} verb in pre-\wayana, but partially switching to the \gl{s_a_} class by taking a \gl{1+2}\gl{s_a_} prefix and the \gl{s_a_} class marker.
-This in turn implies that (inflectionally defined) \gl{s_a_} reflexes in other languages fully switched from \gl{s_p_}.
+Its causativized form is \obj{ɨptə-ka} \parencite[255]{wayanatavares2005}; the \PC  causativizer \rc{-ka} was restricted to \gl{sp} verbs \parencite{gildea2019overview}.
+These patterns point to \qu{to go down} being a regular \gl{sp} verb in pre-\wayana, but partially switching to the \gl{sa} class by taking a \gl{1+2}\gl{sa} prefix and the \gl{sa} class marker.
+This in turn implies that (inflectionally defined) \gl{sa} reflexes in other languages fully switched from \gl{sp}.
 
@@
 The \arara causativized form is \obj{eniptoŋ} \parencite[66]{alves2017arara}, and \kalina has a cognate form \obj{enɨʔto} \parencite[263]{courtz2008carib}; \obj{onɨʔto} \qu{to go down} in \cref{tab:godown} is a detransitivized form thereof, lit.\ \qu{to get oneself down}.
-These forms reflect the transitivizer \rc{en-}, occurring with \gl{s_p_} verbs in \PC \parencite{gildea2019overview}.
+These forms reflect the transitivizer \rc{en-}, occurring with \gl{sp} verbs in \PC \parencite{gildea2019overview}.
 \trio \obj{ɨhtə} has irregular causativized forms, also with a reflex of \rc{en-} \parencite[263]{triomeira1999}.
-In conclusion, \rc{ɨpɨtə} \qu{to go down} was an \gl{s_p_} verb in \PC, but for unknown reasons switched classes in four \dbqu{and a half} languages of the family.
+In conclusion, \rc{ɨpɨtə} \qu{to go down} was an \gl{sp} verb in \PC, but for unknown reasons switched classes in four \dbqu{and a half} languages of the family.
 
 This makes it impossible to tell whether it was affected by most extensions under discussion:
-For \PTir, one cannot establish a relative chronology of the class switch, the introduction of idiosyncratic \gl{1}\gl{s_a_} \rc{p-}, and the extension of \rc{t-}.
+For \PTir, one cannot establish a relative chronology of the class switch, the introduction of idiosyncratic \gl{1}\gl{sa} \rc{p-}, and the extension of \rc{t-}.
 Its first person form and its inflectional class in \PWai are unknown.
 For \carijo and \yukpa, one cannot rule out a verb class switch before the breakdown of the split-\gl{s} system.
-While no language-internal evidence supports such a switch, \qu{to go down} is clearly inclined to do so; \carijo may even have inherited it as \gl{s_a_} from \PTar.
-In all four cases, the verb could have had \gl{s_a_} or \gl{s_p_} status at the time of the extension, so it is unknown whether it even was a potential target.
+While no language-internal evidence supports such a switch, \qu{to go down} is clearly inclined to do so; \carijo may even have inherited it as \gl{sa} from \PTar.
+In all four cases, the verb could have had \gl{sa} or \gl{sp} status at the time of the extension, so it is unknown whether it even was a potential target.
 
-On the other hand, a class switch is reconstructible to \PTir, so it was an \gl{s_a_} verb when \akuriyo introduced \obj{k-}.
+On the other hand, a class switch is reconstructible to \PTir, so it was an \gl{sa} verb when \akuriyo introduced \obj{k-}.
 Likewise, the class switch most likely took place before the extension of \PPek \rc{k-}.
-Otherwise, the newly-turned-\gl{s_a_} verb would have taken on conservative and lexically heavily restricted \rc{w-}, either in \PPek, \PXin, or \arara.
+Otherwise, the newly-turned-\gl{sa} verb would have taken on conservative and lexically heavily restricted \rc{w-}, either in \PPek, \PXin, or \arara.
 
@@
 This verb resisted the extensions of \PPek \rc{k-} \pcref{sec:pekodian} and \akuriyo \obj{k-} \pcref{sec:akuriyo}.
-It took on new \gl{1}\gl{s_a_} prefixes in \PTir \parentext{\trio \obj{s-epɨ-}, \akuriyo \obj{t͡ʃ-epɨ-} \parencites[292]{triomeira1999}[87]{gildea1994akuriyo}} and \PWai \parentext{\hixka \obj{k-ewehɨ-}, \waiwai \obj{k-ejeɸu-} \parencites[195]{hixkaryanaderby1985}[166]{waiwaihawkins1998}}.
+It took on new \gl{1}\gl{sa} prefixes in \PTir \parentext{\trio \obj{s-epɨ-}, \akuriyo \obj{t͡ʃ-epɨ-} \parencites[292]{triomeira1999}[87]{gildea1994akuriyo}} and \PWai \parentext{\hixka \obj{k-ewehɨ-}, \waiwai \obj{k-ejeɸu-} \parencites[195]{hixkaryanaderby1985}[166]{waiwaihawkins1998}}.
 The first person form of its \carijo reflex \obj{ehɨ} \parencites[72]{koch1908hiana} is unknown; an unattested \yukpa cognate may exist.
 
-Verbs for intransitive \qu{to bathe} are typical \gl{s_a_} verbs in most Cariban languages, derived with a detransitivizer from a transitive root.
+Verbs for intransitive \qu{to bathe} are typical \gl{sa} verbs in most Cariban languages, derived with a detransitivizer from a transitive root.
 These roots are reflexes of \rc{[ɨ]pɨ}, or \rc{kupi} in some Venezuelan languages \pcref{tab:bathe}.
@@
 Every \checkmark stands for a verb affected by an extension, while × represents conservatively inflected verbs, making clear how strongly these verbs tend to resist person marker extensions in different languages.
-\cref{sec:motivations} will explore explanations for the fact that the same 1-7 verbs retained their old \gl{1}\gl{s_a_} marker in 6 independent developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs took on new markers.
+\cref{sec:motivations} will explore explanations for the fact that the same 1-7 verbs retained their old \gl{1}\gl{sa} marker in 6 independent developments, while a plethora of regular \gl{s}\textsubscript{(\gl{a})} verbs took on new markers.
 

### 2023-08-23 documents/config/cariban_language_names.tex
fix abbrevs & lg names

--- 2023-01-25 documents/config/cariban_language_names.tex
@@
 \usepackage{xspace}
+\providecommand{\wbakairi}{Western Bakairi\xspace}
+\providecommand{\ebakairi}{Eastern Bakairi\xspace}
+\providecommand{\ingariko}{Ingarikó\xspace}
+\providecommand{\akawaio}{Akawaio\xspace}
+\providecommand{\patamona}{Patamona\xspace}
+\providecommand{\yukpa}{Yukpa\xspace}
+\providecommand{\japreria}{Japreria\xspace}
+\providecommand{\carijo}{Karijona\xspace}
+\providecommand{\mapoyo}{Mapoyo\xspace}
+\providecommand{\yawarana}{Yawarana\xspace}
+\providecommand{\panare}{Panare\xspace}
+\providecommand{\maqui}{Ye'kwana\xspace}
+\providecommand{\pemon}{Pemón\xspace}
+\providecommand{\kapon}{Kapong\xspace}
+\providecommand{\waimiri}{Waimiri-Atroari\xspace}
+\providecommand{\macushi}{Macushi\xspace}
+\providecommand{\waiwai}{Waiwai\xspace}
+\providecommand{\hixka}{Hixkaryána\xspace}
+\providecommand{\kaxui}{Werikyana\xspace}
+\providecommand{\trio}{Tiriyó\xspace}
+\providecommand{\akuriyo}{Akuriyó\xspace}
 \providecommand{\apalai}{Apalaí\xspace}
-\DeclareAcronym{apa_lg}{short=apa,long=Apalaí,short-format=\scshape}
-\providecommand{\trio}{Tiriyó\xspace}
-\DeclareAcronym{tri_lg}{short=tri,long=Tiriyó,short-format=\scshape}
-\providecommand{\hixka}{Hixkaryána\xspace}
-\DeclareAcronym{hix_lg}{short=hix,long=Hixkaryána,short-format=\scshape}
-\providecommand{\panare}{Panare\xspace}
-\DeclareAcronym{pan_lg}{short=pan,long=Panare,short-format=\scshape}
-\providecommand{\macushi}{Macushi\xspace}
-\DeclareAcronym{mac_lg}{short=mac,long=Macushi,short-format=\scshape}
-\providecommand{\waimiri}{Waimiri-Atroari\xspace}
-\DeclareAcronym{wmr_lg}{short=wmr,long=Waimiri-Atroari,short-format=\scshape}
-\providecommand{\maqui}{Ye'kwana\xspace}
-\DeclareAcronym{mak_lg}{short=mak,long=Ye'kwana,short-format=\scshape}
-\providecommand{\waiwai}{Waiwai\xspace}
-\DeclareAcronym{wai_lg}{short=wai,long=Waiwai,short-format=\scshape}
 \providecommand{\wayana}{Wayana\xspace}
-\DeclareAcronym{way_lg}{short=way,long=Wayana,short-format=\scshape}
+\providecommand{\ikpeng}{Ikpeng\xspace}
+\providecommand{\arara}{Arara\xspace}
 \providecommand{\kalina}{Kari'ña\xspace}
-\DeclareAcronym{kar_lg}{short=kar,long=Kari'ña,short-format=\scshape}
-\providecommand{\ikpeng}{Ikpeng\xspace}
-\DeclareAcronym{ikp_lg}{short=ikp,long=Ikpeng,short-format=\scshape}
 \providecommand{\uxc}{Upper Xingu Carib\xspace}
-\DeclareAcronym{uxc_lg}{short=uxc,long=Upper Xingu Carib,short-format=\scshape}
-\providecommand{\arara}{Arara\xspace}
-\DeclareAcronym{ara_lg}{short=ara,long=Arara,short-format=\scshape}
-\providecommand{\ingariko}{Ingarikó\xspace}
-\DeclareAcronym{ing_lg}{short=ing,long=Ingarikó,short-format=\scshape}
-\providecommand{\pemon}{Pemón\xspace}
-\DeclareAcronym{pem_lg}{short=pem,long=Pemón,short-format=\scshape}
-\providecommand{\kapon}{Kapón\xspace}
-\DeclareAcronym{kap_lg}{short=kap,long=Kapón,short-format=\scshape}
-\providecommand{\kalapalo}{Kalapalo\xspace}
-\DeclareAcronym{kal_lg}{short=kal,long=Kalapalo,short-format=\scshape}
-\providecommand{\kuikuro}{Kuikuro\xspace}
-\DeclareAcronym{kui_lg}{short=kui,long=Kuikuro,short-format=\scshape}
-\providecommand{\yukpa}{Yukpa\xspace}
-\DeclareAcronym{yuk_lg}{short=yuk,long=Yukpa,short-format=\scshape}
-\providecommand{\carijo}{Carijona\xspace}
-\DeclareAcronym{car_lg}{short=car,long=Carijona,short-format=\scshape}
 \providecommand{\bakairi}{Bakairi\xspace}
-\DeclareAcronym{bak_lg}{short=bak,long=Bakairi,short-format=\scshape}
-\providecommand{\tamanaku}{†Tamanaku\xspace}
-\DeclareAcronym{tam_lg}{short=tam,long=†Tamanaku,short-format=\scshape}
-\providecommand{\kaxui}{Werikyana\xspace}
-\DeclareAcronym{kax_lg}{short=kax,long=Werikyana,short-format=\scshape}
-\providecommand{\akuriyo}{Akuriyó\xspace}
-\DeclareAcronym{aku_lg}{short=aku,long=Akuriyó,short-format=\scshape}
-\providecommand{\japreria}{Japreria\xspace}
-\DeclareAcronym{jap_lg}{short=jap,long=Japreria,short-format=\scshape}
-\providecommand{\kumana}{†Kumaná\xspace}
-\DeclareAcronym{cum_lg}{short=cum,long=†Kumaná,short-format=\scshape}
-\providecommand{\palmela}{†Palmela\xspace}
-\DeclareAcronym{palm_lg}{short=palm,long=†Palmela,short-format=\scshape}
-\providecommand{\piment}{†Pimenteira\xspace}
-\DeclareAcronym{pime_lg}{short=pime,long=†Pimenteira,short-format=\scshape}
-\providecommand{\arakaju}{†Arakaju\xspace}
-\DeclareAcronym{arak_lg}{short=arak,long=†Arakaju,short-format=\scshape}
-\providecommand{\pauxiana}{†Pauxiana\xspace}
-\DeclareAcronym{pxan_lg}{short=pxan,long=†Pauxiana,short-format=\scshape}
-\providecommand{\paravi}{†Paravilhana\xspace}
-\DeclareAcronym{para_lg}{short=para,long=†Paravilhana,short-format=\scshape}
-\providecommand{\bonari}{†Bonari\xspace}
-\DeclareAcronym{bona_lg}{short=bona,long=†Bonari,short-format=\scshape}
-\providecommand{\wayu}{†Wayumara\xspace}
-\DeclareAcronym{wayu_lg}{short=wayu,long=†Wayumara,short-format=\scshape}
-\providecommand{\apiaka}{†Apiaka-Tocantins\xspace}
-\DeclareAcronym{apia_lg}{short=apia,long=†Apiaka-Tocantins,short-format=\scshape}
-\providecommand{\opon}{†Opón-Carare\xspace}
-\DeclareAcronym{opon_lg}{short=opon,long=†Opón-Carare,short-format=\scshape}
-\providecommand{\pauxi}{†Pauxi\xspace}
-\DeclareAcronym{paux_lg}{short=paux,long=†Pauxi,short-format=\scshape}
-\providecommand{\sapara}{†Sapara\xspace}
-\DeclareAcronym{sapa_lg}{short=sapa,long=†Sapara,short-format=\scshape}
-\providecommand{\tiveri}{†Tiverighotto\xspace}
-\DeclareAcronym{tive_lg}{short=tive,long=†Tiverighotto,short-format=\scshape}
-\providecommand{\yao}{†Yao\xspace}
-\DeclareAcronym{yao_lg}{short=yao,long=†Yao,short-format=\scshape}
-\providecommand{\yaruma}{†Yarumá\xspace}
-\DeclareAcronym{yaru_lg}{short=yaru,long=†Yarumá,short-format=\scshape}
-\providecommand{\koyaima}{†Koyaima\xspace}
-\DeclareAcronym{koya_lg}{short=koya,long=†Koyaima,short-format=\scshape}
-\providecommand{\hiana}{†Hianakoto-Umawa\xspace}
-\DeclareAcronym{hian_lg}{short=hian,long=†Hianakoto-Umawa,short-format=\scshape}
-\providecommand{\saluma}{†Salumá\xspace}
-\DeclareAcronym{salu_lg}{short=salu,long=†Salumá,short-format=\scshape}
-\providecommand{\sikiana}{†Sikiana\xspace}
-\DeclareAcronym{siku_lg}{short=siku,long=†Sikiana,short-format=\scshape}
-\providecommand{\purukoto}{†Purukoto\xspace}
-\DeclareAcronym{puru_lg}{short=puru,long=†Purukoto,short-format=\scshape}
-\providecommand{\juma}{†Juma\xspace}
-\DeclareAcronym{juma_lg}{short=juma,long=†Juma,short-format=\scshape}
-\providecommand{\palenque}{†Palenque\xspace}
-\DeclareAcronym{pale_lg}{short=pale,long=†Palenque,short-format=\scshape}
-\providecommand{\parukoto}{†Parukoto\xspace}
-\DeclareAcronym{paru_lg}{short=paru,long=†Parukoto,short-format=\scshape}
-\providecommand{\urukue}{†Urukuena\xspace}
-\DeclareAcronym{uruk_lg}{short=uruk,long=†Urukuena,short-format=\scshape}
-\providecommand{\wayakule}{†Wayakulé\xspace}
-\DeclareAcronym{waya_lg}{short=waya,long=†Wayakulé,short-format=\scshape}
-\providecommand{\piritu}{†Píritu\xspace}
-\DeclareAcronym{piri_lg}{short=piri,long=†Píritu,short-format=\scshape}
-\providecommand{\guake}{†Guake\xspace}
-\DeclareAcronym{guak_lg}{short=guak,long=†Guake,short-format=\scshape}
-\providecommand{\macoita}{†Macoita\xspace}
-\DeclareAcronym{mcta_lg}{short=mcta,long=†Macoita,short-format=\scshape}
-\providecommand{\macoa}{†Macoa\xspace}
-\DeclareAcronym{maco_lg}{short=maco,long=†Macoa,short-format=\scshape}
-\providecommand{\chake}{†Chake\xspace}
-\DeclareAcronym{chak_lg}{short=chak,long=†Chake,short-format=\scshape}
-\providecommand{\chayma}{†Chayma\xspace}
-\DeclareAcronym{chai_lg}{short=chai,long=†Chayma,short-format=\scshape}
-\providecommand{\avarikoto}{†Avarikoto\xspace}
-\DeclareAcronym{avar_lg}{short=avar,long=†Avarikoto,short-format=\scshape}
-\providecommand{\tauli}{Taulipang\xspace}
-\DeclareAcronym{tau_lg}{short=tau,long=Taulipang,short-format=\scshape}
-\providecommand{\patamona}{Patamona\xspace}
-\DeclareAcronym{pat_lg}{short=pat,long=Patamona,short-format=\scshape}
-\providecommand{\arekuna}{Arekuna\xspace}
-\DeclareAcronym{are_lg}{short=are,long=Arekuna,short-format=\scshape}
-\providecommand{\kamara}{Kamarakoto\xspace}
-\DeclareAcronym{kam_lg}{short=kam,long=Kamarakoto,short-format=\scshape}
-\providecommand{\akawaio}{Akawaio\xspace}
-\DeclareAcronym{aka_lg}{short=aka,long=Akawaio,short-format=\scshape}
-\providecommand{\kumanagoto}{†Kumanagoto\xspace}
-\DeclareAcronym{cgt_lg}{short=cgt,long=†Kumanagoto,short-format=\scshape}
-\providecommand{\crichana}{†Crichana\xspace}
-\DeclareAcronym{cric_lg}{short=cric,long=†Crichana,short-format=\scshape}
-\providecommand{\mapoyo}{Mapoyo\xspace}
-\DeclareAcronym{map_lg}{short=map,long=Mapoyo,short-format=\scshape}
-\providecommand{\yawarana}{Yawarana\xspace}
-\DeclareAcronym{yab_lg}{short=yab,long=Yawarana,short-format=\scshape}
-\providecommand{\yawaperi}{†Jawaperi\xspace}
-\DeclareAcronym{jawa_lg}{short=jawa,long=†Jawaperi,short-format=\scshape}
-\providecommand{\pemono}{†Pémono\xspace}
-\DeclareAcronym{pemo_lg}{short=pemo,long=†Pémono,short-format=\scshape}
-\providecommand{\pianakoto}{†Pianakoto\xspace}
-\DeclareAcronym{pian_lg}{short=pian,long=†Pianakoto,short-format=\scshape}
 \providecommand{\PC}{Proto-Cariban\xspace}
-\DeclareAcronym{pc_lg}{short=pc,long=Proto-Cariban,short-format=\scshape}
-\providecommand{\arina}{†Arinakoto\xspace}
-\DeclareAcronym{arin_lg}{short=arin,long=†Arinakoto,short-format=\scshape}
-\providecommand{\kumayena}{†Kumayena\xspace}
-\DeclareAcronym{kuma_lg}{short=kuma,long=†Kumayena,short-format=\scshape}
-\providecommand{\tyrewuju}{Tyrewuju\xspace}
-\DeclareAcronym{tyr_lg}{short=tyr,long=Tyrewuju,short-format=\scshape}
-\providecommand{\dekwana}{De’kwana\xspace}
-\DeclareAcronym{dek_lg}{short=dek,long=De’kwana,short-format=\scshape}
-\providecommand{\dhekwana}{Dhe’kwana\xspace}
-\DeclareAcronym{dhe_lg}{short=dhe,long=Dhe’kwana,short-format=\scshape}
-\newcommand{\PPar}{Proto-Parukotoan\xspace}
-\newcommand{\PPars}{\textsc{PPar}\xspace}
-\newcommand{\PPem}{Proto-Pemongan\xspace}
-\newcommand{\PPems}{\textsc{PPem}\xspace}
-\newcommand{\PPek}{Proto-Pekodian\xspace}
-\newcommand{\PPeks}{\textsc{PPek}\xspace}
-\newcommand{\PPP}{Proto-Panare-Pemongan\xspace}
-\newcommand{\PPPs}{\textsc{PPP}\xspace}
-\newcommand{\PTir}{Proto-Tiriyoan\xspace}
-\newcommand{\PTirs}{\textsc{PTir}\xspace}
-\newcommand{\PWai}{Proto-Waiwaian\xspace}
-\newcommand{\PWais}{\textsc{PWai}\xspace}
-\newcommand{\PXin}{Proto-Xinguan\xspace}
-\newcommand{\PXins}{\textsc{PXin}\xspace}
-\newcommand{\PTar}{Proto-Taranoan\xspace}
-\newcommand{\PTars}{\textsc{PTar}\xspace}
-\newcommand{\yabmap}{\mapoyo-\yawarana}
-\newcommand{\yabmaps}{\textsc{yab-map}\xspace}
-\newcommand{\vzk}{Venezuelan \kalina}
-\newcommand{\vzks}{\textsc{vzk}\xspace}
-\newcommand{\esk}{Eastern Surinamese \kalina}
-\newcommand{\esks}{\textsc{esk}\xspace}
-\newcommand{\wsk}{Western Surinamese \kalina}
-\newcommand{\wsks}{\textsc{wsk}\xspace}
-\renewcommand{\lg}[1]{\acs*{#1_lg}}
+\providecommand{\PPar}{Proto-Parukotoan\xspace}
+\providecommand{\PWai}{Proto-Waiwaian\xspace}
+\providecommand{\PPek}{Proto-Pekodian\xspace}
+\providecommand{\PXin}{Proto-Xinguan\xspace}
+\providecommand{\PPP}{Proto-Pemong-Panare\xspace}
+\providecommand{\PTar}{Proto-Taranoan\xspace}
+\providecommand{\PTir}{Proto-Tiriyoan\xspace}

### 2023-08-23 documents/config/own_abbreviations.tex
latex compilation working again
fix abbrevs & lg names

@@
-\newGlossingAbbrev{inter}{interrogative}
-\DeclareAcronym{tam}{short=TAM,long=Tense-Aspect-Mood,class=other}
+\DeclareAcronym{1}{short=1,long=first~person,short-format=\scshape}
+\DeclareAcronym{1+2}{short=1+2,long=first~person~inclusive,short-format=\scshape}
+\DeclareAcronym{1+3}{short=1+3,long=first~person~exclusive,short-format=\scshape}
+\DeclareAcronym{2}{short=2,long=second~person,short-format=\scshape}
+\DeclareAcronym{3}{short=3,long=third~person,short-format=\scshape}
+\DeclareAcronym{a}{short=A,long=agent-like~argument~of~canonical~transitive~verb,short-format=\scshape}
+\DeclareAcronym{agt}{short=agt,long=agent,short-format=\scshape}
+\DeclareAcronym{ana}{short=ana,long=anaphoric,short-format=\scshape}
+\DeclareAcronym{attrz}{short=attrz,long=attributivizer,short-format=\scshape}
+\DeclareAcronym{caus}{short=caus,long=causative,short-format=\scshape}
+\DeclareAcronym{cert}{short=cert,long=certainty,short-format=\scshape}
+\DeclareAcronym{com}{short=com,long=comitative,short-format=\scshape}
+\DeclareAcronym{cont}{short=cont,long=continuative,short-format=\scshape}
+\DeclareAcronym{cor}{short=cor,long=coreference,short-format=\scshape}
+\DeclareAcronym{dem}{short=dem,long=demonstrative,short-format=\scshape}
+\DeclareAcronym{des}{short=des,long=desiderative,short-format=\scshape}
+\DeclareAcronym{detrz}{short=detrz,long=detransitivizer,short-format=\scshape}
+\DeclareAcronym{dub}{short=dub,long=dubitative,short-format=\scshape}
+\DeclareAcronym{evid}{short=evid,long=evidentiality,short-format=\scshape}
+\DeclareAcronym{frust}{short=frust,long=frustrative,short-format=\scshape}
+\DeclareAcronym{hsy}{short=hsy,long=hearsay/indirect~evidentiality,short-format=\scshape}
+\DeclareAcronym{imm}{short=imm,long=immentiate~past,short-format=\scshape}
+\DeclareAcronym{imp}{short=imp,long=imperative,short-format=\scshape}
+\DeclareAcronym{inan}{short=inan,long=inanimate,short-format=\scshape}
+\DeclareAcronym{inf}{short=inf,long=infinitive,short-format=\scshape}
+\DeclareAcronym{intr}{short=intr,long=intermediate~past,short-format=\scshape}
+\DeclareAcronym{ints}{short=ints,long=intensifier,short-format=\scshape}
+\DeclareAcronym{lk}{short=lk,long=linker,short-format=\scshape}
+\DeclareAcronym{loc}{short=loc,long=locative,short-format=\scshape}
+\DeclareAcronym{med}{short=med,long=medial,short-format=\scshape}
+\DeclareAcronym{neg}{short=neg,long=negation,short-format=\scshape}
+\DeclareAcronym{nmlz}{short=nmlz,long=nominalizer,short-format=\scshape}
+\DeclareAcronym{npst}{short=npst,long=non-past,short-format=\scshape}
+\DeclareAcronym{obl}{short=obl,long=oblique,short-format=\scshape}
+\DeclareAcronym{p}{short=p,long=patient-like~argument~of~canonical~transitive~verb,short-format=\scshape}
+\DeclareAcronym{pfv}{short=pfv,long=perfective,short-format=\scshape}
+\DeclareAcronym{pl}{short=pl,long=plural,short-format=\scshape}
+\DeclareAcronym{pro}{short=pro,long=pronoun,short-format=\scshape}
+\DeclareAcronym{prs}{short=prs,long=present,short-format=\scshape}
+\DeclareAcronym{pst}{short=pst,long=past,short-format=\scshape}
+\DeclareAcronym{rec}{short=rec,long=recent~past,short-format=\scshape}
+\DeclareAcronym{rem}{short=rem,long=remote~past,short-format=\scshape}
+\DeclareAcronym{s}{short=s,long=single~argument~of~canonical~intransitive~verb,short-format=\scshape}
+\DeclareAcronym{sa}{short=S\textsubscript{A},long=S~marked~like~A,short-format=\scshape}
+\DeclareAcronym{sp}{short=S\textsubscript{P},long=S~marked~like~P,short-format=\scshape}
+\DeclareAcronym{sap}{short=sap,long=speech~act~participant,short-format=\scshape}
+\DeclareAcronym{sup}{short=sup,long=supine,short-format=\scshape}
+\DeclareAcronym{tam}{short=TAM,long=tense-aspect-mood,short-format=\scshape}
+\DeclareAcronym{tr}{short=tr,long=transitive,short-format=\scshape}
+\DeclareAcronym{uncert}{short=uncert,long=uncertainty,short-format=\scshape}
+\DeclareAcronym{inter}{short=inter,long=interrogative,short-format=\scshape}
+\DeclareAcronym{int}{short=int,long=intermediate~past,short-format=\scshape}
+\DeclareAcronym{all}{short=all,long=allative,short-format=\scshape}
+\DeclareAcronym{dat}{short=dat,long=dative,short-format=\scshape}

### 2023-08-23 documents/floats/nominalizations.tex
regular verbs in data and docs
new approach to testing predictions from bybee's model
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
fix abbrevs & lg names

--- 2021-09-10 documents/floats/nominalizations.tex
@@
+\begin{table}
+\centering
+\caption[Nominalizations of \gl{sa} and \gl{sp} verbs]{Nominalizations of \gl{sa} and \gl{sp} verbs \parencites[49, 74]{schuring2018kaxuyana}[97]{alves2017arara}[246]{triomeira1999}[130, 409]{wayanatavares2005}[90]{koehn1986apalai}[ner2-003]{koehns1994textos}[135, 392]{hoff1968carib}[390]{panarepayne2013}[23]{mattei1994diccionario}}
+\label{tab:nominalizations}
 \begin{tabular}[t]{@{}lll@{}}
-\toprule
-Language &                                         \gl{s_a_} &                                             \gl{s_p_} \\
-\midrule
+\mytoprule
+Language &                                         \gl{sa} &                                             \gl{sp} \\
+\mymidrule
 \kaxui  &       \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} &            \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} \\
@@
 \apalai &            \obj{j-epɨ-topo} \qu{my bathing place} &   \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} \\
-\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &                  \obj{aj-ereʔna-∅} \qu{your fainting} \\
+\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &    \obj{aj-ereʔna-{\normalfont ∅}} \qu{your fainting} \\
 \panare &     \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} &  \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} \\
-\bottomrule
+\mybottomrule
 \end{tabular}
+\end{table}

### 2023-08-23 documents/floats/participles.tex
fix abbrevs & lg names

--- 2021-09-10 documents/floats/participles.tex
@@
+\begin{table}
+\centering
+\caption[Participles of \gl{sa} and \gl{sp} verbs]{Participles of \gl{sa} and \gl{sp} verbs \parencites[39]{schuring2018kaxuyana}[118, 207]{alves2017arara}[333-334]{triomeira1999}[400]{wayanatavares2005}[35]{koehn1986apalai}[kuruaz-154]{koehns1994textos}[430, 433]{hoff1968carib}[232, 244]{panarepayne2013}}
+\label{tab:participles}
 \begin{tabular}[t]{@{}lll@{}}
-\toprule
-Language &                               \gl{s_a_} &                        \gl{s_p_} \\
-\midrule
+\mytoprule
+Language &                               \gl{sa} &                        \gl{sp} \\
+\mymidrule
 \kaxui  &         \obj{t-ehurka-t͡ʃe} \qu{fallen} &       \obj{tɨ-jaʔ-so} \qu{burnt} \\

### 2023-08-23 documents/floats/pekreg.tex
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
rewriting and restructuring
a thousand small fixes
more data processing
fix abbrevs & lg names

@@
 \centering
-\caption[Regular Pekodian \gl{s_a_} verbs]{Regular Pekodian \gl{s_a_} verbs \parencites[4]{meira2003bakairi}[150]{alves2017arara}[52]{ikpengpacheco2001}}
+\caption[Regular Pekodian \gl{sa} verbs]{Regular Pekodian \gl{sa} verbs \parencites[4]{meira2003bakairi}[150]{alves2017arara}[52]{ikpengpacheco2001}}
 \label{tab:pekreg}
@@
 \mytoprule
-{} & \bakairi \qu{to go up} &         \arara \qu{to dance} &            \ikpeng \qu{to run} \\
-\midrule
+{} & \ebakairi \qu{to go up} &         \arara \qu{to dance} &            \ikpeng \qu{to run} \\
+\mymidrule
 \gl{1}   &            \obj{k-əku-} &               \obj{k-origu-} &                \obj{k-aranme-} \\
@@
 \gl{3}   &            \obj{n-əku-} &  \obj{{\normalfont ∅}-origu} &  \obj{{\normalfont ∅}-aranme-} \\
-\bottomrule
+\mybottomrule
 \end{tabular}

### 2023-08-23 documents/floats/ppekverbs.tex
more table fixes
fix abbrevs & lg names

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[129, 170]{meiraDBebakairi / pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[111]{pinto2022bakairi}{meiraDBebakairi}}
+\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[129, 170]{meiraDBebakairi / pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[111]{pinto2022bakairi}{meiraDBebakairi}}
 \label{tab:ppekverbs}
@@
 \mymidrule
-\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &  \obj{u-gue- / u-ge-} \\
+\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &  \obj{u-gue-} / \obj{u-ge-} \\
 \qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &                  \obj{w-i-} \\

### 2023-08-23 documents/floats/ptirverbs.tex
fix abbrevs & lg names

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-114]{meira1998proto}{meiraDBakuriyo}}
+\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-114]{meira1998proto}{meiraDBakuriyo}}
 \label{tab:ptirverbs}

### 2023-08-23 documents/floats/pwaireg.tex
fix abbrevs & lg names

@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai \parencites[150]{howard2001wrought}[189, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{sa}) and \qu{to sleep} (\gl{sp}) in \PWai \parencites[150]{howard2001wrought}[189, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
 \label{tab:pwaireg}

### 2023-08-23 documents/floats/pwaiverbs.tex
fix abbrevs & lg names

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197]{hixkaryanaderby1985}[Spike Gildea]{pc}}
+\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PWai \parencites[4]{hixkaryanaderby1979}[71, 85]{waiwaihawkins1998}[70, 197]{hixkaryanaderby1985}[Spike Gildea]{pc}}
 \label{tab:pwaiverbs}

### 2023-08-23 workflow/process_data.py

@@ -340,16 +340,27 @@ print_aligned_table(
 
 # some paradigms for 'to come'
 lg_list = ["ara", "tri", "kax"]
-pyd = pyradigms.Pyradigm(infl_data, x="Language_ID", y="Inflection", sort_orders={"Language_ID": lg_list, "Inflection": person}, filters= {
-    "Meaning_ID": ["come"],
-    "Inflection": person,
-    "Language_ID": lg_list,
-})
+pyd = pyradigms.Pyradigm(
+    infl_data,
+    x="Language_ID",
+    y="Inflection",
+    sort_orders={"Language_ID": lg_list, "Inflection": person},
+    filters={
+        "Meaning_ID": ["come"],
+        "Inflection": person,
+        "Language_ID": lg_list,
+    },
+)
 table = pyd.compose_paradigm()
 sources = get_sources(pyd)
 table.rename(columns=shorthand_dic, inplace=True)
 table = glossify_index(table)
-export_table(table, label="comepara", caption=r"\rc{(ət-)epɨ} \qu{to come} in paradigms", sources=sources)
+export_table(
+    table,
+    label="comepara",
+    caption=r"\rc{(ət-)epɨ} \qu{to come} in paradigms",
+    sources=sources,
+)
 
 # comparison of intransitive and transitive 'to bathe'
 df_b = pd.read_csv("../data/bathe_data.csv")

### 2023-08-24 documents/sections/introduction.tex
fixes #18 and #17

@@
 In some Cariban languages, a small group of verbs show a divergent first person inflection pattern, a topic which has not received much attention in the literature.
-This is illustrated for \hixka in \cref{tab:hixintro},\footnote{The presence of a \gl{1+2} person value implies that of a \gl{1+3} value.
-This is usually expressed with a free pronoun combined with third person morphology in Cariban languages, so it is not represented as a distinct value in \cref{tab:hixintro} and other paradigm tables.
-Further, TAM suffixes in the attested forms are omitted, since a) the focus lies on the prefixes and stems, and b) full paradigms with the same TAM suffix are rare in the available sources.
-Standard IPA symbols are used in the transcription of Cariban languages, except for coronal rhotics, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
-In languages with strong morphophonological processes and/or non-phonemic orthography the original form is shown in an additional line when presented in interlinearized examples.
+This is illustrated for \hixka in \cref{tab:hixintro},\footnote{All forms under discussion have prefixes from \posscite[16-18]{gildea1998} "Set I" paradigm, which is conditioned by \gl{tam} suffixes.
+	The latter are omitted in tables, since a) the focus lies on the prefixes and stems, and b) full paradigms with the same \gl{tam} suffix are rare in the available sources.
+Standard IPA symbols are used in the transcription languages, except for coronal rhotics of Cariban, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
 \textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}

### 2023-08-24 documents/sections/introduction.tex
fixes #20; fixes #21

@@
 Standard IPA symbols are used in the transcription languages, except for coronal rhotics of Cariban, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
-\textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
+\textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC \rc{ô} reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
 with paradigms of four verbs, all members of the \gl{sa} inflectional class.

### 2023-08-24 documents/sections/proto_carib.tex
fixes #20; fixes #21

@@
 Person indexation in transitive verbs was conditioned by a basic hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
-The locuphoric markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
-There was a single aliophoric marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
+\gl{sap} markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
+There was a single third person marker \rc{n(i)-}, which only surfaced in nonlocal (\gl{3}>\gl{3}) scenarios, without morphologically expressed distinctions between different third person referents.
 Local scenarios were marked identically and non-transparently with the \gl{1+2} prefix \rc{k-}.

### 2023-08-24 documents/floats/ppekverbs.tex
missing references, last traces of _

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[129, 170]{meiraDBebakairi / pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[111]{pinto2022bakairi}{meiraDBebakairi}}
+\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}{meiraDBebakairi}[111, 129, 170]{pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
 \label{tab:ppekverbs}

### 2023-08-24 documents/sections/extensions.tex
missing references, last traces of _

@@
 
-All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{sa} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{alves2013verbo}.
+All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{sa} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{chagas2013verbo}.
 However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}, with the exception of \obj{k-}prefixed \qu{to go} \exref{ikp-170}.

### 2023-08-24 workflow/process_data.py

@@ -86,7 +86,7 @@ tabular = glossify_index(tabular)
 export_table(
     tabular,
     label="pekreg",
-    caption="Regular Pekodian \gl{s_a_} verbs",
+    caption="Regular Pekodian \gl{sa} verbs",
     sources=sources_list,
 )
 
@@ -98,7 +98,7 @@ reconstructed_form_table(
     lgs,
     "PPek",
     verbs,
-    r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek",
+    r"Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek",
     "ppekverbs",
     verb_list,
 )
@@ -123,7 +123,7 @@ tabular = glossify_index(tabular)
 export_table(
     tabular,
     label="pwaireg",
-    caption="Regular \qu{to fall} (\gl{s_a_}) and \qu{to sleep} (\gl{s_p_}) in \PWai",
+    caption="Regular \qu{to fall} (\gl{sa}) and \qu{to sleep} (\gl{sp}) in \PWai",
     sources=get_sources(pyd),
     short_caption="Regular \\PWai verbs",
 )
@@ -134,7 +134,7 @@ reconstructed_form_table(
     lgs,
     "PWai",
     verbs,
-    r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai",
+    r"Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PWai",
     "pwaiverbs",
     verb_list,
 )
@@ -167,7 +167,7 @@ tabular.columns.names = [None, None]
 export_table(
     df=tabular,
     label="ptirreg",
-    caption="Regular \PTir \gl{s_a_} verbs",
+    caption="Regular \PTir \gl{sa} verbs",
     sources=get_sources(pyd),
 )
 
@@ -176,7 +176,7 @@ reconstructed_form_table(
     ["PTir", "tri", "aku"],
     "PTir",
     ["go", "say", "come", "be_1", "be_2"],
-    r"Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir",
+    r"Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir",
     "ptirverbs",
     verb_list,
 )
@@ -205,7 +205,7 @@ aku_verbs.fillna("", inplace=True)
 export_table(
     aku_verbs,
     label="aku1sa",
[…]

### 2023-08-24 documents/sections/proto_carib.tex
fixes #22

@@
 
-Knowledge about the ancestral system makes evident that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{sa} prefix \rc{w-}; they are therefore \firstmention{conservative}.
+Knowledge about the ancestral system makes it clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{sa} prefix \rc{w-}; they are therefore \firstmention{conservative}.
 They contrast with regular \gl{sa} verbs, which are innovative in both languages.
@@
 These verbs and their prefixes are comparable with the few English nouns like \obj{ɒks}, which preserve the old plural suffix \obj{-ən}.
-It was once more widespread as the normal plural suffix of the weak inflection, compare German \obj{ɔks-ən} \qu{ox-en}, \obj{naːmə-n} \qu{name-s}, \obj{haːzə-n} \qu{hare-s}, \obj{bɛːʁ-ən} \qu{bear-s}.
+It was once more widespread as the normal plural suffix of the weak inflection, compare German \obj{ɔks-ən} \qu{oxen}, \obj{naːmə-n} \qu{names}, \obj{haːzə-n} \qu{hares}, \obj{bɛːʁ-ən} \qu{bears}.
 Since the regular \hixka and \trio prefixes are innovative, the question arises how they developed.

### 2023-08-24 documents/config/own_abbreviations.tex
fixes #26

@@
 \DeclareAcronym{obl}{short=obl,long=oblique,short-format=\scshape}
-\DeclareAcronym{p}{short=p,long=patient-like~argument~of~canonical~transitive~verb,short-format=\scshape}
+\DeclareAcronym{p}{short=P,long=patient-like~argument~of~canonical~transitive~verb,short-format=\scshape}
 \DeclareAcronym{pfv}{short=pfv,long=perfective,short-format=\scshape}
@@
 \DeclareAcronym{rem}{short=rem,long=remote~past,short-format=\scshape}
-\DeclareAcronym{s}{short=s,long=single~argument~of~canonical~intransitive~verb,short-format=\scshape}
+\DeclareAcronym{s}{short=S,long=single~argument~of~canonical~intransitive~verb,short-format=\scshape}
 \DeclareAcronym{sa}{short=S\textsubscript{A},long=S~marked~like~A,short-format=\scshape}

### 2023-08-24 documents/sections/proto_carib.tex
fixes #26

@@
     		\PWai & \gl{1}\gl{sp} \rc{k-} to \gl{1}\gl{sa}\\
-    		& \rc{owɨ(ro) j-} \qu{\gl{1}\gl{pro} \gl{lk}} for \gl{1}\gl{p}\\
+    		& \rc{owɨ(ro) (j-)} \qu{\gl{1}\gl{pro} (\gl{lk})} for \gl{1}\gl{p}\\
     		\waiwai & \gl{2}\gl{sa} \obj{m-} to \gl{2}\gl{sp}
@@
 %
-All innovations are person marker extensions except the innovative first person P form, which combined a pronoun with the linker \rc{j-}.
+All innovations are person marker extensions except the innovative first person P form, which consists of a pronoun, combined with the linker \rc{j-} for V-initial stems.
 \cref{fig:par_ext} shows them in bold and reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{sp} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
@@
 As seen in \cref{sec:pc_person}, the split between \gl{sa} and \gl{sp} verbs was instantiated by inflection classes within the \PC \setone person paradigm, but this was not the only difference:
-In deverbalized forms, \gl{sa} verbs took a class marker \rc{w-}, while \gl{sp} verbs lacked that prefix \parencites[89, 141--142]{gildea1998}[208]{meira2000split}.
+In nonfinite forms (nominalizations and adverbializations), \gl{sa} verbs took a prefix \rc{w-}, while \gl{sp} verbs lacked that prefix \parencites[89, 141--142]{gildea1998}[208]{meira2000split}.
 Also, \gl{sp} verbs took the \gl{2}\gl{sp} prefix \rc{ə(j)-} in imperatives, while \gl{sa} verbs were unprefixed \parencites[89]{gildea1998}[208]{meira2000split}.
@@
 
-In modern instantiations of Cariban split-\gl{s}, mismatches between the semantics of verbs and their \gl{sa} or \gl{sp} status are common, exemplified with \kalina data in \exref{kar2}.
+%In modern instantiations of Cariban split-\gl{s}, mismatches between the semantics of verbs and their \gl{sa} or \gl{sp} status are common, exemplified with \kalina data in \exref{kar2}.
 %
+%\input{examples/kar2.tex}
+%
+%The \gl{sa} verb \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker \exref{kar2.kar-59}, while the \gl{sp} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
+%In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
+%\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker is clearly volitional.
 %\gl{a}- or \gl{p}-oriented inflectional morphology are common, exemplified with \kalina data in \exref[kar1]{kar2}.
@@
 %
-\input{examples/kar2.tex}
-%
-The \gl{sa} verb \obj{ekema} \qu{to be afraid} takes an \gl{a}-oriented marker \exref{kar2.kar-59}, while the \gl{sp} verb \obj{awomɨ} \qu{to get up} takes a \gl{p}-oriented marker \exref{kar2.kar-61}.
-In both cases, the prefix does not appear to contribute to the semantics of the predicate, since there are clear mismatches:
-\qu{to be afraid} with an \dbqu{agentive} marker can hardly be considered a volitional act, while  \qu{to get up} with a \dbqu{patientive} marker is clearly volitional.
+
 \textcite{meira2000split} investigates a corpus of intransitive verbs from \trio, \kalina, \apalai, and \wayana, and categorizes them by applying different criteria commonly encountered in split-\gl{s} systems.
@@
 satisfactorily predict the class membership of intransitive verbs in any of the languages.
-
 Rather, the reason for a verb to take \gl{a}- or \gl{p}-oriented prefixes is (at least diachronically) a morphological one.
-\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a derivational detransitivizing prefix are treated as \gl{sa} verbs, while essentially all others are \gl{sp} verbs:
+\textcite[217--221]{meira2000split} demonstrates that those intransitive verbs which (etymologically) have a derivational detransitivizing prefix are treated as \gl{sa} verbs, while almost all others are \gl{sp} verbs:
 \begin{quotebox}{\parencite[201]{meira2000split}}

### 2023-11-24 documents/floats/ppekverbs.tex
xinguan forms

@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}{meiraDBebakairi}[111, 129, 170]{pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
+\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}{meiraDBebakairi}[83, 111, 129, 170]{pinto2022bakairi}{meiraDBwbakairi}[76, 130, 137, 374, 397, 400]{von1892bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[Fael Ikpeng]{pc}}
 \label{tab:ppekverbs}
-\begin{tabular}[t]{@{}lllll@{}}
+\begin{tabular}[t]{@{}llllllll@{}}
 \mytoprule
-{} &          \PPek &          \arara &                       \ikpeng &                   \ebakairi \\
+{} &                                         \PPek &             \PXin &          \arara &                       \ikpeng &                     \ebakairi &                     \wbakairi &                          \bakairi \\
 \mymidrule
-\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &  \obj{u-gue-} / \obj{u-ge-} \\
-\qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &                  \obj{w-i-} \\
-\qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &                \obj{w-awɨ-} \\
-\qu{be-2}    &  \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &                  \obj{w-i-} \\
-\qu{come}    &    \rc{w-epɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &               \obj{k-əewɨ-} \\
-\qu{go down} &   \rc{w-ɨptə-} &                 &                               &                             \\
-\qu{go}      &   \rc{w-ɨdən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &                 \obj{u-də-} \\
+\qu{say}     &                                   \rc{wɨ-ge-} &       \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &                   \obj{u-ge-} &                   \obj{u-ge-} &                     \ort{u-ɣépa} \\
+\qu{bathe}   &                                   \rc{w-ipɨ-} &       \rc{w-ibɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &                    \obj{w-i-} &                    \obj{w-i-} &                                   \\
+\qu{be-1}    &                                    \rc{w-ap-} &        \rc{w-ap-} &     \obj{w-ap-} &                             – &                \obj{w-a(wɨ)-} &                \obj{w-a(wə)-} &  \ort{töre-wakine}, \ort{wawí-le} \\
+\qu{be-2}    &                                 \rc{w-et͡ʃi-} &     \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &                    \obj{w-i-} &                    \obj{w-i-} &                      \ort{wí-ta} \\
+\qu{come}    &                              \rc{w-(əd-)epɨ-} &  \rc{w-(od-)ebɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &                 \obj{k-əewɨ-} &                 \obj{k-əewə-} &                 \ort{kχ-aewí-le} \\
+\qu{go down} &  \rc{w-ɨptə({\normalfont C}{\normalfont V})-} &     \rc{w-iptoŋ-} &  \obj{w-iptoŋ-} &  \obj{{\normalfont ∅}-iptoŋ-} &  \obj{{\normalfont ∅}-ɨtəgɨ-} &  \obj{{\normalfont ∅}-itəge-} &                \ort{kχ-itaké-he} \\
+\qu{go}      &                                  \rc{w-ɨdən-} &      \rc{w-ɨdon-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &                   \obj{u-də-} &                   \obj{u-də-} &                       \ort{ū́ta} \\
 \mybottomrule

### 2023-11-24 workflow/process_data.py

@@ -92,7 +92,7 @@ export_table(
 
 
 # conservative pekodian verbs
-lgs = ["PPek", "ara", "ikp", "ebak"]
+lgs = ["PPek", "PXin", "ara", "ikp", "ebak", "wbak", "bak"]
 verbs = ["say", "bathe_intr", "be_1", "be_2", "come", "go_down", "go"]
 reconstructed_form_table(
     lgs,

### 2023-11-24 workflow/read_data.py

@@ -34,6 +34,7 @@ concept_dic = {
     "DETRZ+come": "come",
     "DETRZ1": "bathe_intr",
     "DETRZ": "come",
+    "go_down+?": "go_down",
 }
 
 infl_data["Concept"] = infl_data.apply(

### 2023-11-26 documents/floats/aku-1sa.tex
reworked extensions section

\begin{table}
\centering
\caption[Regular \akuriyo \gl{1}\gl{sa} markers]{Regular \akuriyo \gl{1}\gl{sa} markers \parencites[77, 79, 82, 84-87]{gildea1994akuriyo}{meiraDBakuriyo}}
\label{tab:aku-1sa}
\begin{tabular}[t]{@{}ll@{}}
\toprule
first person \obj{k-} & first person \obj{t͡ʃ-} \\
\midrule
\obj{əət͡ʃena-} ‘to cry’ & \obj{epɨ-} ‘to bathe (\gl{intr})’ \\
\obj{əməmɨ-} ‘to enter’ & \obj{ehpa-} ‘to bathe (\gl{intr})’ \\
\obj{əempa-} ‘to learn’ & \obj{etonema-} ‘to lie down’ \\
\obj{ətajiŋka-} ‘to run’ & \obj{ewai-} ‘to sit down’ \\
\obj{əənɨkɨ-} ‘to sleep’ & \obj{ekɨrɨka-} ‘to stay back’ \\
\obj{əturu-} ‘to talk’ & \obj{entapo-} ‘to yawn’ \\
\obj{əiwa-} ‘to tremble’ & \obj{etaka-} ‘to be satiated’ \\
\bottomrule
\end{tabular}
\end{table}

### 2023-11-26 documents/floats/bak-verbs.tex
reworked extensions section

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \bakairi]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \bakairi \parencites{meiraDBebakairi}[83, 111-112, 170]{pinto2022bakairi}{meiraDBwbakairi}[130-131, 137, 374, 397, 400]{von1892bakairi}[4]{meira2003bakairi}}
\label{tab:bak-verbs}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \ebakairi & \wbakairi & \bakairi(19th century) \\
\midrule
\qu{say} & \obj{u-ge-} & \obj{u-ge-} & \ort{u-ɣépa} \\
\qu{bathe} & \obj{w-i-} & \obj{w-i-} &  \\
\qu{be-1} & \obj{w-a(wɨ)-} & \obj{w-a(wə)-} & \ort{wawí-le} \\
\qu{be-2} & \obj{w-i-} & \obj{w-i-} & \ort{wí-ta} \\
\qu{go down} & \obj{{\normalfont ∅}-ɨtəgɨ-} & \obj{{\normalfont ∅}-itəge-} & \ort{kχ-itaké-he} \\
\qu{go} & \obj{u-də-} & \obj{u-də-} & \ort{ū́ta} \\
\bottomrule
\end{tabular}
\end{table}

### 2023-11-26 documents/floats/hix-intro.tex
reworked extensions section

--- 2021-11-12 documents/floats/hixintro.tex
@@
 \centering
-\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510, 511, 513, 520]{howard2001wrought}[197, 198]{hixkaryanaderby1985}}
-\label{tab:hixintro}
+\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510-511, 513, 520]{howard2001wrought}[198]{hixkaryanaderby1985}}
+\label{tab:hix-intro}
 \begin{tabular}[t]{@{}llllll@{}}
-\mytoprule
-{} &     \qu{to fall} &  \qu{to be afraid} &          \qu{to walk} & \qu{to cut self} &    \qu{to be} \\
+\toprule
+ & ‘to fall’ & ‘to be afraid’ & ‘to walk’ & ‘to cut self’ & ‘to be’ \\
 \midrule

### 2023-11-26 documents/floats/ikp-a.tex
reworked extensions section

\begin{table}
\centering
\caption[Lowering of mid vowels in \ikpeng]{Lowering of mid vowels in \ikpeng \parencites[9, 96]{desouza2010arara}[41]{ikpengpacheco1997}[21, 82, 98, 267, 284]{ikpengpacheco2001}[38]{desouza1988arara}[28, 102]{pinto2022bakairi}[131, 172, 187]{alves2017arara}[180, 184]{meira2005southern}}
\label{tab:ikp-a}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \arara & \ikpeng & \ebakairi \\
\midrule
‘dog’ & \obj{wokori} & \obj{okari} / \obj{akari} &  \\
‘path’ & \obj{enma} / \obj{onma} / \obj{anma} & \obj{anma} & \obj{ə̃wã} \\
‘to bring’ & \obj{enebɨ} & \obj{anep} & \obj{ene} \\
‘tree’ & \obj{jej} & \obj{jaj} & \obj{se} \\
‘what’ & \obj{odɨ} & \obj{arɨ} & \obj{ədɨ} \\
\bottomrule
\end{tabular}
\end{table}

### 2023-11-26 documents/floats/ikp-w.tex
reworked extensions section

\begin{table}
\centering
\caption[Loss of \rc{w} in \ikpeng]{Loss of \rc{w} in \ikpeng \parencites[143]{alves2017arara}[21, 164]{ikpengpacheco2001}[9]{desouza2010arara}[41]{ikpengpacheco1997}[44, 70]{souza1993arara}[40]{campetela1997analise}[118]{chagas2013verbo}}
\label{tab:ikp-w}
\begin{tabular}[t]{@{}lll@{}}
\toprule
 & \arara & \ikpeng \\
\midrule
‘DAT’ & \obj{wɨna} & \obj{ɨna} \\
‘dog’ & \obj{wokori} & \obj{okari} / \obj{akari} \\
‘capuchin monkey’ & \obj{tawe} & \obj{tae} \\
‘to sleep’ & \obj{wɨnkɨ} & \obj{ɨnkɨ} \\
‘to defecate’ & \obj{watke} & \obj{atke} \\
\bottomrule
\end{tabular}
\end{table}

### 2023-11-26 documents/floats/pek-verbs.tex
reworked extensions section

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}{meiraDBebakairi}[111-112, 129, 170]{pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[Fael Ikpeng]{pc}}
\label{tab:pek-verbs}
\begin{tabular}[t]{@{}lllll@{}}
\toprule
 & \PPek & \arara & \ikpeng & \ebakairi \\
\midrule
\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} & \obj{u-ge-} \\
\qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{{\normalfont ∅}-ip-} & \obj{w-i-} \\
\qu{be-1} & \rc{w-apɨ-} & \obj{w-ap-} & – & \obj{w-a(wɨ)-} \\
\qu{be-2} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{{\normalfont ∅}-et͡ʃi-} & \obj{w-i-} \\
\qu{come} & \rc{w-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
\qu{go down} & \rc{w-ɨptə({\normalfont C}{\normalfont V})-} & \obj{w-iptoŋ-} & \obj{{\normalfont ∅}-iptoŋ-} & \obj{{\normalfont ∅}-ɨtəgɨ-} \\
\qu{go} & \rc{w-ɨdən-} & \obj{w-ɨdo-} & \obj{k-aran-} & \obj{u-də-} \\
\bottomrule
\end{tabular}
\end{table}

### 2023-11-26 documents/floats/ptir-reg.tex
reworked extensions section

--- 2021-10-26 documents/floats/ptirreg.tex
@@
 \centering
-\caption[Regular \PTir \gl{s_a_} verbs]{Regular \PTir \gl{s_a_} verbs \parencites[292, 294]{triomeira1999}[87]{gildea1994akuriyo}}
-\label{tab:ptirreg}
+\caption[Regular \PTir \gl{sa} verbs]{Regular \PTir \gl{sa} verbs \parencites[292, 294]{triomeira1999}[87]{gildea1994akuriyo}}
+\label{tab:ptir-reg}
 \begin{tabular}[t]{@{}lllllll@{}}
-\mytoprule
-{} & \multicolumn{3}{l}{\qu{to bathe (\gl{intr})}} & \multicolumn{3}{l}{\qu{to sleep}} \\
-{} &                     \PTir &          \trio &        \akuriyo &             \PTir &              \trio &           \akuriyo \\
+\toprule
+ & \multicolumn{3}{l}{‘to bathe (\gl{intr})’} & \multicolumn{3}{l}{‘to sleep’} \\
+ & \PTir & \trio & \akuriyo & \PTir & \trio & \akuriyo \\
 \midrule
-\gl{1}   &             \rc{t͡ʃ-epɨ-} &   \obj{s-epɨ-} &  \obj{t͡ʃ-epɨ-} &    \rc{t-əənɨkɨ-} &    \obj{t-əənɨkɨ-} &    \obj{k-əənɨkɨ-} \\
+\gl{1} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-} / \obj{t͡ʃ-ehpa-} & \rc{t-əənɨkɨ-} & \obj{t-əənɨkɨ-} & \obj{k-əənɨkɨ-} \\
 \gl{2} & \rc{m-epɨ-} & \obj{m-epɨ-} & \obj{m-epɨ-} & \rc{m-əənɨkɨ-} & \obj{m-əənɨkɨ-} & \obj{m-əənɨkɨ-} \\

### 2023-11-26 documents/floats/ptir-verbs.tex
reworked extensions section

--- 2023-08-23 documents/floats/ptirverbs.tex
@@
 \centering
-\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-114]{meira1998proto}{meiraDBakuriyo}}
-\label{tab:ptirverbs}
+\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-115]{meira1998proto}{meiraDBakuriyo}}
+\label{tab:ptir-verbs}
 \begin{tabular}[t]{@{}llll@{}}
-\mytoprule
-{} &          \PTir &          \trio &                     \akuriyo \\
-\mymidrule
+\toprule
+ & \PTir & \trio & \akuriyo \\
+\midrule
 \qu{go} & \rc{w-ɨtəmɨ-} & \obj{wɨ-tən-} & \obj{ə-təmɨ-} \\
@@
 \qu{be-2} & \rc{w-eʔi-} & \obj{w-ei-} & \obj{{\normalfont ∅}-eʔi-} \\
-\mybottomrule
+\bottomrule
 \end{tabular}

### 2023-11-26 documents/floats/ptirreg.tex
more tables
new approach to testing predictions from bybee's model
moved verb discussion to verbs section, fixed lingpy alignments, fixed map
class membership of proto-taranoan 'to go down' is not reconstructible
more trimming, fixed verb citations
fixed csv output
fix abbrevs & lg names
reworked extensions section

@@
 \centering
-\caption[Regular \PTir \gl{s_a_} verbs]{Regular \PTir \gl{s_a_} verbs \parencites[292, 294]{triomeira1999}[87]{gildea1994akuriyo}}
+\caption[Regular \PTir \gl{sa} verbs]{Regular \PTir \gl{sa} verbs \parencites[292, 294]{triomeira1999}[87]{gildea1994akuriyo}}
 \label{tab:ptirreg}
 \begin{tabular}[t]{@{}lllllll@{}}
-\mytoprule
-{} & \multicolumn{3}{l}{\qu{to bathe (\gl{intr})}} & \multicolumn{3}{l}{\qu{to sleep}} \\
-{} &                     \PTir &          \trio &        \akuriyo &             \PTir &              \trio &           \akuriyo \\
+\toprule
+ & \multicolumn{3}{l}{‘to bathe (\gl{intr})’} & \multicolumn{3}{l}{‘to sleep’} \\
+ & \PTir & \trio & \akuriyo & \PTir & \trio & \akuriyo \\
 \midrule
-\gl{1}   &             \rc{t͡ʃ-epɨ-} &   \obj{s-epɨ-} &  \obj{t͡ʃ-epɨ-} &    \rc{t-əənɨkɨ-} &    \obj{t-əənɨkɨ-} &    \obj{k-əənɨkɨ-} \\
+\gl{1} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-} / \obj{t͡ʃ-ehpa-} & \rc{t-əənɨkɨ-} & \obj{t-əənɨkɨ-} & \obj{k-əənɨkɨ-} \\
 \gl{2} & \rc{m-epɨ-} & \obj{m-epɨ-} & \obj{m-epɨ-} & \rc{m-əənɨkɨ-} & \obj{m-əənɨkɨ-} & \obj{m-əənɨkɨ-} \\

### 2023-11-26 documents/floats/pwai-reg.tex
reworked extensions section

--- 2023-08-23 documents/floats/pwaireg.tex
@@
 \centering
-\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{sa}) and \qu{to sleep} (\gl{sp}) in \PWai \parencites[150]{howard2001wrought}[189, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
-\label{tab:pwaireg}
+\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{sa}) and \qu{to sleep} (\gl{sp}) in \PWai \parencites[150]{howard2001wrought}[189-190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
+\label{tab:pwai-reg}
 \begin{tabular}[t]{@{}lllllll@{}}
-\mytoprule
-{} &    \qu{to fall} &                  &                    &     \qu{to sleep} &                 &                    \\
-{} &           \PWai &           \hixka &            \waiwai &             \PWai &          \hixka &            \waiwai \\
-\mymidrule
+\toprule
+ & \multicolumn{3}{l}{‘to fall’} & \multicolumn{3}{l}{‘to sleep’} \\
+ & \PWai & \hixka & \waiwai & \PWai & \hixka & \waiwai \\
+\midrule
 \gl{1} & \rc{k-eɸurka-} & \obj{k-ehurka-} & \obj{k-eɸɨrka-} & \rc{kɨ-wɨnɨkɨ-} & \obj{kɨ-nɨkɨ-} & \obj{kɨ-wɨnɨkɨ-} \\
@@
 \gl{3} & \rc{ɲ-eɸurka-} & \obj{ɲ-ehurka-} & \obj{ɲ-eɸɨrka-} & \rc{nɨ-wɨnɨkɨ-} & \obj{nɨ-nɨkɨ-} & \obj{nɨ-wɨnɨkɨ-} \\
-\mybottomrule
+\bottomrule
 \end{tabular}

### 2023-11-26 documents/floats/xin-verbs.tex
reworked extensions section

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PXin]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PXin \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[Fael Ikpeng]{pc}}
\label{tab:xin-verbs}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \PXin & \arara & \ikpeng \\
\midrule
\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} \\
\qu{bathe} & \rc{w-ibɨ-} & \obj{w-ibɨ-} & \obj{{\normalfont ∅}-ip-} \\
\qu{be-1} & \rc{w-ap-} & \obj{w-ap-} & – \\
\qu{be-2} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{{\normalfont ∅}-et͡ʃi-} \\
\qu{come} & \rc{w-(od)ebɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} \\
\qu{go down} & \rc{w-iptoŋ-} & \obj{w-iptoŋ-} & \obj{{\normalfont ∅}-iptoŋ-} \\
\qu{go} & \rc{w-ɨdon-} & \obj{w-ɨdo-} & \obj{k-aran-} \\
\bottomrule
\end{tabular}
\end{table}

### 2023-11-26 documents/sections/extensions.tex
reworked extensions section

@@
 Since languages show considerable etymological overlap in their conservative verbs, those are compared and reconstructed in \cref{sec:verbs}.
-Further, their reflexes which did get affected by an incomplete extension are identified.
+Further, their reflexes which \textit{did} get affected by an incomplete extension are identified.
 \cref{sec:verbsummary} summarizes the results.
@@
 \textcite{meira2005southern}, who established the branch, focused on phonology and lexicon, so no reconstructions of \PPek morphosyntax are found in the literature.
-However, all three Pekodian languages have a regular \gl{1}\gl{sa} marker \obj{k-} \pcref{tab:pekreg}, allowing the reconstruction of a \PPek \gl{1}\gl{sa} marker \rc{k-}.
+However, all three Pekodian languages have a regular \gl{1}\gl{sa} marker \obj{k-} \pcref{tab:pek-reg}, allowing the reconstruction of a \PPek \gl{1}\gl{sa} marker \rc{k-}.
 
@@
 
-\input{floats/pekreg}
+%\input{floats/pek-reg}
 
-The most detailed description of a Pekodian language \parencite{alves2017arara} names six%
+\textcite{alves2017arara} names six%
 \footnote{Seven in her analysis, treating the two meanings of \obj{it͡ʃi} \qu{to be / to lie down} as different verbs.}
-\arara \gl{sa} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{arairr}.
+\arara \gl{sa} verbs forming a subclass defined by a first person marker \obj{w(ɨ)-} rather than \obj{k-}, shown in \exref{ara.irr}.
 There is also a reflex of the copula \rc{a[p]} \pcref{sec:be}, serving syntactically as a postposition introducing adverbial clauses meaning \qu{if} or \qu{when} \parencite[199--201]{alves2017arara}.
-However, it inflects with verbal \setone prefixes, including first person \obj{w-} \exref{araap}.
+However, it inflects with verbal \setone prefixes, including first person \obj{w-} \exref{ara.ap}.
 %We take these \arara verbs as a starting point for the other Pekodian languages, as \textcite{alves2017arara} gives the most detailed description of person markers for any language of the branch.
 
+\pex<ara> \arara \parencite[153, 200]{alves2017arara}
 \begin{multicols}{2}
-\ex<arairr> \arara \parencite[153]{alves2017arara}\\
+\a<irr> 
 \begin{tabular}[t]{@{}ll@{}}
@@
 \end{tabular}
-\xe
-
-\ex<araap> \arara \parencite[200]{alves2017arara}\\
+\a<ap>
 \begin{tabular}[t]{@{}lll@{}}
@@
 \end{tabular}
+\end{multicols}
 \xe
-\end{multicols}
 
@@
 The first group is illustrated with \obj{i} \qu{to bathe} \exref{bak-33}.
+Additional members were identified in Sérgio Meira's shoebox databases \parencites{meiraDBebakairi}{meiraDBwbakairi}, agreeing with forms attested in \textcite{pinto2022bakairi}.
+19th century \bakairi data recorded by \textcite{von1892bakairi} largely match the modern inflectional pattern, with the exception of \obj{k-}marked \qu{to go down} \pcref{tab:bak-verbs}.
 
@@
 \xe
+
+\input{floats/bak-verbs}
 %
 %Since \qu{to bathe} is also found in the \obj{w-}list for \arara, other \bakairi cognates of these verbs are of interest.
-While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of the \arara verbs in \exref{arairr} as \gl{sa} verbs, he does not indicate whether they belong to the \gl{sa}-1 class with \obj{k-}, or the \gl{sa}-2 class with \obj{w-}.
-However, inflected forms can be found in \textcite{von1892bakairi}, represented in \exref{bakverbs} based on the analyses of \bakairi phonology and verbal morphology by \textcites{wheatley1969bakairi}{meira2003bakairi}{meira2005bakairi}{franchetto2016classes}.
 
-\pex[everyglpreamble=]<bakverbs> \bakairi \parencite[][131, 397, 76, 137, 374, 130]{von1892bakairi}
-\begin{multicols}{2}
-\a<bak-3>
-\begingl
-\glpreamble \ort{u-ɣépa} //
-\gla u-ge-pa//
-\glb \gl{1}\gl{sa}-say-\gl{neg}//
-\glft \qu{I don't say.}//
-\endgl
-\a<bak-5>
-\begingl
-\glpreamble \ort{wi-táki} / \ort{wi-tági} //
-\gla w-i-taki//
-\glb \gl{1}\gl{sa}-be-\gl{int}//
-\glft \qu{I was.}//
-\endgl
-\a<bak-4>
-\begingl
-\glpreamble \ort{kχaewí-le} //
-\gla k-əewɨ-lɨ//
-\glb \gl{1}\gl{sa}-come-\gl{imm}//
-\glft \qu{I came.}//
-\endgl
-\a<bak-6>
-\begingl
-\glpreamble \ort{kχ-itaké-he} //
-\gla k-ɨtəgɨ-se//
-\glb \gl{1}\gl{sa}-go.down-\gl{npst}?//
-\glft \qu{I go down.}//
-\endgl
-\a<bak-2>
-\begingl
-\glpreamble \ort{ū́ta} / \ort{uúta}//
-\gla u-tə//
-\glb \gl{1}\gl{sa}-go//
-\glft \qu{I go.}//
-\endgl
-\a<bak-35>
-\begingl
-\glpreamble \ort{töre-w-akine}//
-\gla tərə w-a-kɨne//
-\glb there \gl{1}\gl{sa}-be-\gl{pst}.\gl{cont}//
-\glft \qu{I was there.}//
-\endgl
-\end{multicols}
-\xe
+%While \textcite[4]{meira2003bakairi} lists some \bakairi cognates of the \arara verbs in \exref{arairr} as \gl{sa} verbs, he does not indicate whether they belong to the \gl{sa}-1 class with \obj{k-}, or the \gl{sa}-2 class with \obj{w-}.
+
+%\pex[everyglpreamble=]<bakverbs> \bakairi \parencite[][131, 397, 76, 137, 374, 130]{von1892bakairi}
+%\begin{multicols}{2}
+%\a<bak-3>
+%\begingl
+%\glpreamble \ort{u-ɣépa} //
+%\gla u-ge-pa//
+%\glb \gl{1}\gl{sa}-say-\gl{neg}//
+%\glft \qu{I don't say.}//
+%\endgl
+%\a<bak-5>
+%\begingl
+%\glpreamble \ort{wi-táki} / \ort{wi-tági} //
+%\gla w-i-taki//
+%\glb \gl{1}\gl{sa}-be-\gl{int}//
+%\glft \qu{I was.}//
+%\endgl
+%\a<bak-4>
+%\begingl
+%\glpreamble \ort{kχaewí-le} //
+%\gla k-əewɨ-lɨ//
+%\glb \gl{1}\gl{sa}-come-\gl{imm}//
+%\glft \qu{I came.}//
+%\endgl
+%\a<bak-6>
+%\begingl
+%\glpreamble \ort{kχ-itaké-he} //
+%\gla k-ɨtəgɨ-se//
+%\glb \gl{1}\gl{sa}-go.down-\gl{npst}?//
+%\glft \qu{I go down.}//
+%\endgl
+%\a<bak-2>
+%\begingl
+%\glpreamble \ort{ū́ta} / \ort{uúta}//
+%\gla u-tə//
+%\glb \gl{1}\gl{sa}-go//
+%\glft \qu{I go.}//
+%\endgl
+%\a<bak-35>
+%\begingl
+%\glpreamble \ort{töre-w-akine}//
+%\gla tərə w-a-kɨne//
+%\glb there \gl{1}\gl{sa}-be-\gl{pst}.\gl{cont}//
+%\glft \qu{I was there.}//
+%\endgl
+%\end{multicols}
+%\xe
 
 All descriptions of \ikpeng list \obj{k-} as the only \gl{1}\gl{sa} marker \parencites[55]{ikpengpacheco1997}[105]{campetela1997analise}[64]{ikpengpacheco2001}[205]{chagas2013verbo}.
-However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or Ø \exref{ikpw}, with the exception of \obj{k-}prefixed \qu{to go} \exref{ikp-170}.
+However, most \ikpeng cognates of the verbs in question do not take \obj{k-}, but rather \obj{ɨ-} or ∅ \exref{ikpw}, with the exception of \obj{k-}prefixed \qu{to go} and \qu{to come} \exref{ikp-comego}.
 While reflexes of \rc{a[p]} \qu{be-1} do exist in \ikpeng, apparently only reflexes of \rc{eti} \qu{be-2} occur with first person prefixes \parencite[401]{gildea2018reconstructing}.
@@
 \begingl
-\gla Ø-et͡ʃi-lɨ//
+\gla ∅-et͡ʃi-lɨ//
 \glb \gl{1}-be-\gl{rec}//
@@
 \begingl
-\gla at͡ʃagotpop Ø-ip-t͡ʃi ik-gwa-kt͡ʃi//
+\gla at͡ʃagotpop ∅-ip-t͡ʃi ik-gwa-kt͡ʃi//
 \glb always \gl{1}-bathe-\gl{npst} river-\gl{loc}.aquatic-\gl{all}//
@@
 \endgl
+\a<ikp-godown>
+\begingl
+\gla ∅-iptoŋ-tʃi//
+\glb \gl{1}-go.down-\gl{npst}//
+\glft \qu{I go down.} (Fael Ikpeng, p.c.) //
+\endgl
 \xe
 
-\ex<ikp-170>\ikpeng \parencite[][80]{ikpengpacheco2001}\\
+\pex<ikp-comego> \ikpeng \parencite[][80]{ikpengpacheco2001}
+\a<ikp-170>
 \begingl
@@
 \endgl
+\a<ikp-come>
+\begingl
+\gla k-arep-lɨ//
+\glb \gl{1}-come-\gl{rec}//
+\glft \qu{I came.}//
+\endgl
 \xe
 
-\input{floats/ppekverbs}
+\input{floats/pek-verbs}
 
-Reconstructed \PPek forms of conservatively inflected verbs are given in \Cref{tab:ppekverbs}.
-Newly identified \ikpeng \obj{ɨ-}/Ø is demonstrably a reflex of \PXin \rc{w(ɨ)-}, based on other (albeit irregular) cases of loss of \rc{w} \pcref{tab:pxinw}.
-%The presence and distribution of the \ikpeng \gl{1}\gl{sa} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{sa} \obj{w(ɨ)-}.
-%Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in .
-%While it is by no means a regular sound change, it permits identifying the two prefixes as cognate.
-Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in correspondences like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
-Thus, a \gl{1}\gl{sa} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to its \arara reflex in form and distribution.
-In later, individual developments, \bakairi extended \obj{k-} to \qu{to go down}, and \ikpeng to \qu{to go}.
+\PPek forms can be reconstructed by comparing the three languages \pcref{tab:pek-verbs}; the \ebakairi dialect is used for its relative conservativeness.
+\arara most faithfully conserves the prefixes I reconstruct for \PPek.
+In \bakairi, \rc{wɨ-} became \obj{u-} before consonants, and apparently eroded to zero before \obj{ɨ}; this correspondence is found elsewhere, e.g.\ \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+Zero is the general pre-vowel reflex in \ikpeng, the only verb with an overt marker being C-initial \qu{to say}.
+The identification of this zero as cognate with \arara \obj{w-} is supported by the existence of identical correspondences in lexical material \pcref{tab:pxinw}; similarly, \ebakairi \qu{to sleep} is \obj{w}-less \obj{ɨkɨ} \parencite[39]{pinto2022bakairi}.
 
-\input{floats/pxinw}
+\input{floats/ikp-w}
 
-Reconstructions of verb stems are deferred to \cref{sec:verbs}, but a comment on \qu{to come} is in order:
+After the separation from \arara, \ikpeng extended first person \obj{k-} to \qu{to come} and \qu{to go}, the former development happening separately in \bakairi.
+The situation for \qu{to go down} is somewhat unexpected, since the 19th centry \bakairi recorded by \textcite{von1892bakairi} shows (innovative) \obj{k-}, but both modern dialects consistently have (conservative) ∅-.
+A likely scenario is that while the innovative prefix started spreading to this verb, it did not take hold.
+There would have been sociolinguistic and/or dialectal variation where the verb occurred with both prefixes for a while.
+Neither of the two Xinguan languages shows any evidence for an innovative prefix on \qu{to go down}.
+
+The identification of \ikpeng \obj{aran} \qu{to go} as cognate with \arara \obj{ɨdo} and \bakairi \obj{də} may require some justification.
+The consonants (and meanings) are a perfect match for the cognate set; while final \obj{n} is not attested elsewhere in Pekodian, it is outside of the branch \pcref{sec:go}.
+The two mismatching vowels are the outcome of a lowering of mid vowels, the source still attested in the root allomorph used in third person forms, \obj{ero} \parencite[54]{ikpengpacheco2001}.
+Lowering to \obj{a} is attested elsewhere in \ikpeng \pcref{tab:ikp-a}; forms like \obj{akari}/\obj{okari} suggest that it can gradually spread via assimilation.
+The suggested trajectory thus starts with the lowering of initial \obj{ɨ} to \obj{e} \pcref{sec:go}, followed by a (simultaneous or subsequent) lowering of both vowels to \obj{a}.
+
+\input{floats/ikp-a.tex}
+
+Reconstructions of verb stems are deferred to \cref{sec:verbs}, but another comment on \qu{to come} is in order:
 The stems are not fully cognate, as \ikpeng and \bakairi both show a reflex of the \PPek detransitivizer \rc{əd-} in combination with a root reconstructible as \rc{epɨ} \pcref{sec:come}.
@@
 
+%Reconstructed \PPek forms of conservatively inflected verbs are given in \Cref{tab:pek-verbs}.
+%Newly identified \ikpeng \obj{ɨ-}/∅ is demonstrably a reflex of \PXin \rc{w(ɨ)-}, based on other (albeit irregular) cases of loss of \rc{w} \pcref{tab:pxinw}.
+%The presence and distribution of the \ikpeng \gl{1}\gl{sa} marker \obj{ɨ-}/Ø suggests that it is cognate with \arara \gl{1}\gl{sa} \obj{w(ɨ)-}.
+%Indeed, \PXin \rc{w} is attested as sometimes being lost in \ikpeng, as evidenced by the correspondences in .
+%While it is by no means a regular sound change, it permits identifying the two prefixes as cognate.
+%Similarly, the change of \rc{wɨ} to \bakairi \obj{u} is found in correspondences like \obj{udo} \parencite{meira2005southern} from \PC \rc{wɨtoto} \qu{person} \parencite[4]{gildea2007greenberg}.
+%Thus, a \gl{1}\gl{sa} prefix \rc{w(ɨ)-} can securely be reconstructed to \PPek, identical to its \arara reflex in form and distribution.
+
 \subsubsection{\PWai \rc{k-}}
@@
 This extension, one of the Parukotoan innovations shown in \cref{sec:extensions_intro}, resulted in the \hixka patterns from \cref{sec:intro}.
-\PWai further extended the \gl{1}\gl{sp} prefix \rc{k-} (innovated in \PPar) to \gl{1}\gl{sa}.
-For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwaireg}.
-\input{floats/pwaireg}
+\PWai further extended the \gl{1}\gl{sp} prefix \rc{k-} that was introduced in \PPar to \gl{1}\gl{sa}.
+For regularly inflected verbs, this created a unified \gl{1}\gl{s} category \pcref{tab:pwai-reg}.
 
@@
 The regular correspondence in transitive verbs points to \hixka \obj{ɨ-} on intransitive verbs as another reflex of \rc{wɨ-}, with a similar phonological reduction as in \ikpeng \pcref{sec:pekodian}.
-Notably, \textcite{hixkaryanaderby1985} analyzes this \obj{ɨ-} as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive \pcref{sec:say}.
+Notably, \textcite{hixkaryanaderby1985} analyzes this \obj{ɨ-} as the regular \gl{1}>\gl{3} prefix, because he considers \hixka \obj{ka} \qu{to say} to be transitive (see \cref{sec:say}).
 
@@
 
+\input{floats/pwai-reg}
 \input{floats/pwaiverbs}
@@
 The moniker Tiriyoan \parencite{glottolog} subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages identified by \textcite{girard1971proto}, with \carijo as a sister.
-\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{sa} \rc{w-} seem to have switched places, creating a regular \gl{1}\gl{sa} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
-\footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
-\input{floats/ptirreg}
+\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar, facing an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{sa} \rc{w-} seem to have switched places, creating a regular \gl{1}\gl{sa} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptir-reg}.%
+\footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo, see \cref{sec:akuriyo}.}
+\input{floats/ptir-reg}
 The question of how this switch happened in detail \parencite[107--112]{meira1998proto} is still open, although a scenario seems necessary in which both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
@@
 
-\input{floats/ptirverbs}
+\input{floats/ptir-verbs}
 
@@
 \akuriyo and \carijo would then have levelled that pattern, similar to what was suggested for the Pekodian languages \pcref{sec:pekodian}.
-Here, both \trio \obj{əepɨ} and \akuriyo \obj{eepɨ} are identified as reflecting \rc{ətepɨ} \pcref{sec:come}, via \PTar \rc{(ət͡ʃ-)epɨ} and \PTir \rc{(əʔ)epɨ}.}
-As a fifth verb \rc{eʔi} \qu{be-1} can be added, whose \trio reflex retains \obj{w-}.
-The idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} is identified as a reflex of \rc{wɨ-} by 
-\textcite[113]{meira1998proto}, which is supported the fact that both components of the idiosyncratic change \rc{wɨ-} > \obj{ə-} (\phon{\rc{w}}{∅} and \phon{\rc{ɨ}}{\obj{ə}}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
+In the reconstruction presented here, both \trio \obj{əepɨ} and \akuriyo \obj{eepɨ} are identified as reflecting \rc{ətepɨ} \pcref{sec:come}, via \PTar \rc{(ət͡ʃ-)epɨ} and \PTir \rc{(əʔ)epɨ}.}
+To this list, a fifth verb \rc{eʔi} \qu{be-1} can be added.
+
+\textcite[113]{meira1998proto} identifies the idiosyncratic \akuriyo first person prefix \obj{ə-} on \qu{to go} as a reflex of \rc{wɨ-}, which is supported the fact that both components of the idiosyncratic change \rc{wɨ-} > \obj{ə-} (\phon{\rc{w}}{∅} and \phon{\rc{ɨ}}{\obj{ə}}) are found in other person prefixes \exref[exp.aku-164]{exp.aku-159}.
 
@@
 \obj{k-} and \obj{t͡ʃ-} show a clear phonologically conditioned distribution in \posscite{gildea1994akuriyo} \akuriyo data \pcref{tab:aku1sa}.
-\textcite[107]{meira1998proto} largely confirms that distribution, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs), albeit without any examples.
+\textcite[107]{meira1998proto} largely confirms that distribution, but mentions \dbqu{several cases of first person \obj{t-} in \akuriyo{}} (on \obj{ə}-initial verbs).
 He also suggests that \obj{k-} may be more recent, which is plausible: since the distribution \envr{\rc{t-}}{\obj{ə}} / \envr{\rc{t͡ʃ-}}{\obj{e}} is reconstructible to \PTir, the most straightforward scenario is \obj{k-} replacing \rc{t-} but not \rc{t͡ʃ-} in \akuriyo.
-The few \obj{t-} mentioned by \textcite{meira1998proto} were then perhaps reintroduced under \trio influence.
-However, since there are no examples of -- or more information about -- \obj{ə}-initial verbs with \obj{t-}, these cases cannot be discussed further.
+His shoebox database \parencite{meiraDBakuriyo} contains two examples of first person \obj{t-}, both with run-of-the-mill \gl{sa} verbs with a detransitivizing prefix.
+Considering that the speaker in question did not speak \akuriyo very well, \posscite[108]{meira1998proto} suggestion that these were perhaps reintroduced under \trio influence is not unrealistic.
 
@@
 
-\input{floats/aku1sa}
+\input{floats/aku-1sa}
 
-The verbs listed for \PTir in \cref{tab:ptirverbs} in \cref{sec:taranoan} of course also resisted the extension of \obj{k-} in \akuriyo.
+The verbs listed for \PTir in \cref{tab:ptir-verbs} in \cref{sec:taranoan} of course also resisted the extension of \obj{k-} in \akuriyo.
 In addition, there is an \gl{sa} verb \obj{ɨ(h)tə} \qu{to go down}, which has an irregular first person marker \obj{p-}, apparently reconstructible to \PTir \exref{tridown}.
-It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{sa} verb when \PTir \rc{t-} was introduced is unclear (see \cref{sec:godown}).
+It was not affected by the extension of \akuriyo \obj{k-}, but whether it was an \gl{sa} verb when \PTir \rc{t-} was introduced in the first place is unclear (see \cref{sec:godown}).
 
@@
 \label{sec:carijo}
-\carijo, the third Taranoan language, extended the \gl{1}\gl{sp} marker \obj{j(i)-}\footnote{Since all affected \gl{sa} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{sa} verbs \parencite[105--107]{meira1998proto}.
-Combined with the extension of \gl{2}\gl{sa} \obj{m-} and \gl{1+2}\gl{sa} \obj{kɨt-}/\obj{kɨs-} to \gl{sp} verbs, this created a single unified \gl{s} category for regular verbs \pcref{tab:carreg}.
+\carijo, the third Taranoan language, extended the \gl{1}\gl{sp} marker \obj{j-} to \gl{sa} verbs \parencite[105--107]{meira1998proto}.
+Combined with the extension of \gl{2}\gl{sa} \obj{m-} and \gl{1+2}\gl{sa} \obj{kɨt-}/\obj{kɨs-} to \gl{sp} verbs, this created a single unified \gl{s} category for regular verbs \pcref{tab:car-reg}.
 %
-\input{floats/carreg}%
+\input{floats/car-reg}%
 %
@@
 
-As noted in \cref{sec:taranoan}, this extension also erased any traces of putative \PTar \gl{1}\gl{sa} \rc{t-}.
+As noted in \cref{sec:taranoan}, this extension would have also erased any traces of a putative \PTar \gl{1}\gl{sa} \rc{t-}.
 However, it did not fully eclipse the old \gl{1}\gl{sa} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
@@
 Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, \obj{ka} \qu{to say} should either have conservative \obj{wɨ-} or innovative \obj{ji-}, so the zero is unexpected.
-It is analysed here as a reflex of \rc{wɨ-}, primarily due to the loss of \rc{w} in \ikpeng and \hixka.
+It is identified here as an eroded reflex of \rc{wɨ-}, similar to the loss of \rc{w} in \ikpeng and \hixka.
 While those developments were more regular, an already irregular marker undergoing idiosyncratic phonological erosion is not that surprising, see \akuriyo \rc{wɨ-} > \obj{ə-} in the preceding section.
@@
 %
-\input{floats/yukreg}%
+\input{floats/yuk-reg}%
 %

### 2023-11-26 documents/sections/introduction.tex
reworked extensions section

@@
 
-\input{floats/hixintro}
-\input{floats/triintro}
+\input{floats/hix-intro}
+\input{floats/tri-intro}
 
 In some Cariban languages, a small group of verbs show a divergent first person inflection pattern, a topic which has not received much attention in the literature.
-This is illustrated for \hixka in \cref{tab:hixintro},\footnote{All forms under discussion have prefixes from \posscite[16-18]{gildea1998} "Set I" paradigm, which is conditioned by \gl{tam} suffixes.
+This is illustrated for \hixka in \cref{tab:hix-intro},\footnote{All forms under discussion have prefixes from \posscite[16-18]{gildea1998} "Set I" paradigm, which is conditioned by \gl{tam} suffixes.
 	The latter are omitted in tables, since a) the focus lies on the prefixes and stems, and b) full paradigms with the same \gl{tam} suffix are rare in the available sources.
-Standard IPA symbols are used in the transcription languages, except for coronal rhotics of Cariban, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
+Standard IPA symbols are used for transcribing language data, except for coronal rhotics, which are simply represented with \ort{r}, rather than \ort{ɽ} for \wayana or \ort{ɾ̠} for \maqui etc.
 \textcite{gildea2018reconstructing} is followed in using \ort{ə} for the \PC \rc{ô} reconstructed by \textcite{meira2005southern}, although it was likely more back \parencite{gildea2010story}.}
@@
 In this language, the verb \qu{to be} diverges from other \gl{sa} verbs like \qu{to fall} by having a first person marker \obj{w-}, rather than \obj{k-}.
-A similar pattern exists in \trio \pcref{tab:triintro}, where \gl{sa} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}}, while the verb \qu{to go} has a first-person prefix \obj{wɨ-}.
+A similar pattern exists in \trio \pcref{tab:tri-intro}, where \gl{sa} verbs have a prefix with phonologically conditioned allomorphs \envr{\obj{t-}}{\obj{ə}} and \envr{\obj{s-}}{\obj{e}}, while the verb \qu{to go} has a first-person prefix \obj{wɨ-}.
 For both languages, the first person prefix on the first four verbs in the table is representative for the vast majority of \gl{sa} verbs.
@@
 The goal of this study is to approach these patterns from a comparative perspective and to provide a diachronic and functional account, proceeding as follows:
-In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns like in \cref{tab:hixintro,tab:triintro}.
+In \cref{sec:background}, relevant aspects of the \PC verbal system are introduced, and it is shown that the mechanism of person marker extensions is responsible for patterns like in \cref{tab:hix-intro,tab:tri-intro}.
 In \cref{sec:data}, six incomplete person marker extensions and the verbs unaffected by them are described.

### 2023-11-26 documents/sections/proto_carib.tex
reworked extensions section

@@
 \PC is reconstructed by \textcite{gildea1998} as using a person paradigm called \setone in its independent verb forms, shown in \cref{tab:pcpers}.
-Person indexation in transitive verbs was conditioned by a basic hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
+Person indexation in transitive verbs can be described as being conditioned by a basic hierarchy \fbox{\gl{1}/\gl{2} > \gl{3}}.
 \gl{sap} markers had two forms, an \gl{a}-oriented one for direct (\gl{sap}>\gl{3}) scenarios and a \gl{p}-oriented one for inverse (\gl{3}>\gl{sap}) scenarios.
@@
 Knowledge about the ancestral system makes it clear that the divergent \hixka and \trio forms in \cref{tab:hixintro,tab:triintro} behave irregularly because they preserve the original \PC \gl{1}\gl{sa} prefix \rc{w-}; they are therefore \firstmention{conservative}.
-They contrast with regular \gl{sa} verbs, which are innovative in both languages.
+They contrast with regular \gl{sa} verbs, which have innovative first person markers in both languages.
 The reflexes of \rc{w-} are \firstmention{relics}, old and restricted to a few lexemes, contrasting with the innovative prefixes found elsewhere.
@@
 %
-All innovations are person marker extensions except the innovative first person P form, which consists of a pronoun, combined with the linker \rc{j-} for V-initial stems.
-\cref{fig:par_ext} shows them in bold and reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{sp} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
+All innovations are person marker extensions except the innovative first person P form, which consists of a free pronoun combined with the linker \rc{j-} for V-initial stems.
+\cref{fig:par_ext} shows the extensions in bold and reproduces \posscite{gildea1998} tables as a tree diagram, with adapted transcription and an additional \kaxui \gl{1}\gl{sp} marker ∅/\obj{j-} \perscommpar{Spike Gildea}.
 %
@@
 Whether or not this gradual switch followed ordered lines, lexical diffusion must have played a role.
+The third argument in favor of the lexical diffusion scenario are the very forms under discussion.
+A situation where some verbs preserve the old person marker fits very well with a scenario where lexical diffusion stops short of these verbs, rather than spreading affecting all potential targets.
 
-The third argument in favor of the lexical diffusion scenario goes back to conservative forms like those in \hixka and \trio.
-In both languages, the innovative \gl{1}\gl{sa} prefixes were introduced by a person marker extension spreading via lexical diffusion.
-The continued presence of the old \gl{1}\gl{sa} prefix in a few verbs is the result of the extension stopping short of these verbs, rather than spreading through the entire \gl{sa} lexicon.
 In a family-wide search, 18 distinct extensions affecting intransitive verbs were identified, 6 of them incomplete.
 The latter have left between 1 and 7 conservatively inflected verbs in 9 Cariban languages \pcref{sec:data}.
-
 Interestingly, all six featured innovative first person markers on \gl{sa} verbs.
@@
 The starkly different behavior of \gl{sa} and \gl{sp} verbs regarding extensions points to the split-\gl{s} system playing a role, so its main properties will be discussed in \cref{sec:split}.
-This will also clarify how the \gl{sa}/\gl{sp} distinction can be lost for a single person, or how \gl{sp} verbs can take on \gl{sa} markers with apparent semantic impunity.
 
@@
 Regarding the form of this marker, \textcite[505--512]{meira2010origin} reconstruct two distinct prefixes for \PC: reciprocal \rc{əte-} and reflexive \rc{e-}, although their reflexes on verbs have been merged into a single morpheme in modern languages.%
-\footnote{In the \PC transitive verb template, derivational prefixes were in a paradigmatic relationship with the earlier third person marker \rc{i-}:
+\footnote{In the \PC verb template, derivational prefixes were in a paradigmatic relationship with the earlier third person marker \rc{i-}:
 \rc{m-i-}V \qu{you V it}, \rc{m-e-}V \qu{you V yourself}, \rc{m-əte-}V \qu{you V each other}.

### 2023-11-26 documents/sections/extensions.tex
yukpa 1-say

@@
 Since \PC \gl{1}\gl{sa} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \gl{1}\gl{sa} \rc{w-}.
+Zero-marked first person is attested for a two intransitive verbs, \obj{to} \qu{to go} and \obj{ka} \qu{to say} \exref{yuk-zero}.
+They diverge from regular C-initial verbs with \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yuk-reg} and can thus be identified as having resisted the extension of \obj{j-} in \yukpa.
 
-Zero-marked first person is attested for a single intransitive verb, \obj{to} \qu{to go} \exref{yuk-7}.
-It diverges from regular C-initial verbs with \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.
-It can thus be identified as having resisted the extension of \obj{j-} in \yukpa.
-
-\ex<yuk-7>\yukpa \parencite[][139]{meira2006syntactic}\\
+\pex<yuk-zero> \yukpa \parencites[139]{meira2006syntactic}{meiraDByukpa}
+\a<yuk-7>
 \begingl
@@
 \endgl
+\a<yuk-say>
+\begingl
+\gla awɨ sa {\normalfont ∅}-ka jɨ-na//
+\glb \gl{1}\gl{pro} thus \gl{1}\gl{sa}-say \gl{3}-\gl{obl}//
+\glft \qu{I said to him,}//
+\endgl
 \xe

### 2023-11-27 documents/sections/verbs.tex
start reworking stems

@@
 
+\input{floats/simple-verbs}
+
 \subsubsection{\rc{ka[ti]} \qu{to say}}
 \label{sec:say}
-This verb was not affected by any of the extensions in \PPek, \PWai, \PTir, \akuriyo, or \carijo (\crefrange{sec:pekodian}{sec:carijo}), while no first person form of its \yukpa reflex \obj{ka} is attested.
-Most reflexes are simply \obj{ka}, but a fleeting syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, visible in the imperative forms of some languages. % \exref{kati}.
+This verb was not affected by any of the extensions in \PPek, \PWai, \PTir, \akuriyo, \carijo, or \yukpa (\crefrange{sec:pekodian}{sec:yukpa}).
+Most reflexes are simply \obj{ka}, but a fleeting second syllable \rc{ti} is reconstructed by \textcite{gildea2007greenberg}, visible in the imperative forms of some languages. % \exref{kati}.
 %
@@
 %
-\cref{tab:say} shows a comparison of the longest attested forms for each language.%
-\footnote{Cognate segments in \crefrange{tab:say}{tab:bathe} were aligned automatically with LingPy \parencite{lingpy268}, for improved exposition of correspondences.
-Brackets indicate segments not present in all forms.}
+\cref{tab:say} shows a comparison of the longest attested forms for each language; brackets indicate segments not present in all forms.
 %
